### 2019-01-15 paper/abstract.tex
Sync

\begin{abstract}

  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
  purus, et dictum nunc justo sit amet elit.

\end{abstract}

### 2019-01-15 paper/main.tex
Sync

\documentclass[conference]{IEEEtran}

%% ODER: format ==         = "\mathrel{==}"
%% ODER: format /=         = "\neq "
%
%
\makeatletter
\@ifundefined{lhs2tex.lhs2tex.sty.read}%
  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
   \newcommand\SkipToFmtEnd{}%
   \newcommand\EndFmtInput{}%
   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
  }\SkipToFmtEnd

\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
\usepackage{amstext}
\usepackage{amssymb}
\usepackage{stmaryrd}
\DeclareFontFamily{OT1}{cmtex}{}
\DeclareFontShape{OT1}{cmtex}{m}{n}
  {<5><6><7><8>cmtex8
   <9>cmtex9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
\DeclareFontShape{OT1}{cmtex}{m}{it}
  {<-> ssub * cmtt/m/it}{}
\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
\DeclareFontShape{OT1}{cmtt}{bx}{n}
  {<5><6><7><8>cmtt8
   <9>cmbtt9
   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
\DeclareFontShape{OT1}{cmtex}{bx}{n}
  {<-> ssub * cmtt/bx/n}{}
\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU

\newcommand{\Sp}{\hskip.33334em\relax}

\newcommand{\Conid}[1]{\mathit{#1}}
\newcommand{\Varid}[1]{\mathit{#1}}
\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
\newcommand{\plus}{\mathbin{+\!\!\!+}}
\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
\newcommand{\sequ}{\mathbin{>\!\!\!>}}
\renewcommand{\leq}{\leqslant}
\renewcommand{\geq}{\geqslant}
\usepackage{polytable}

%mathindent has to be defined
\@ifundefined{mathindent}%
  {\newdimen\mathindent\mathindent\leftmargini}%
  {}%

\def\resethooks{%
  \global\let\SaveRestoreHook\empty
  \global\let\ColumnHook\empty}
\newcommand*{\savecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
\newcommand*{\restorecolumns}[1][default]%
  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
\newcommand*{\aligncolumn}[2]%
  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}

\resethooks

\newcommand{\onelinecommentchars}{\quad-{}- }
\newcommand{\commentbeginchars}{\enskip\{-}
\newcommand{\commentendchars}{-\}\enskip}

\newcommand{\visiblecomments}{%
  \let\onelinecomment=\onelinecommentchars
  \let\commentbegin=\commentbeginchars
  \let\commentend=\commentendchars}

\newcommand{\invisiblecomments}{%
  \let\onelinecomment=\empty
  \let\commentbegin=\empty
  \let\commentend=\empty}

\visiblecomments

\newlength{\blanklineskip}
\setlength{\blanklineskip}{0.66084ex}

\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
\let\hspre\empty
\let\hspost\empty
\newcommand{\NB}{\textbf{NB}}
\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}

\EndFmtInput
\makeatother
%
%
%
%
%
%
% This package provides two environments suitable to take the place
% of hscode, called "plainhscode" and "arrayhscode". 
%
% The plain environment surrounds each code block by vertical space,
% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
% similar to formulas. Note that if these dimensions are changed,
% the spacing around displayed math formulas changes as well.
% All code is indented using \leftskip.
%
% Changed 19.08.2004 to reflect changes in colorcode. Should work with
% CodeGroup.sty.
%
\ReadOnlyOnce{polycode.fmt}%
\makeatletter

\newcommand{\hsnewpar}[1]%
  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}

% can be used, for instance, to redefine the code size, by setting the
% command to \small or something alike
\newcommand{\hscodestyle}{}

% The command \sethscode can be used to switch the code formatting
% behaviour by mapping the hscode environment in the subst directive
% to a new LaTeX environment.

\newcommand{\sethscode}[1]%
  {\expandafter\let\expandafter\hscode\csname #1\endcsname
   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}

% "compatibility" mode restores the non-polycode.fmt layout.

\newenvironment{compathscode}%
  {\par\noindent
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed\)%
   \par\noindent
   \ignorespacesafterend}

\newcommand{\compaths}{\sethscode{compathscode}}

% "plain" mode is the proposed default.
% It should now work with \centering.
% This required some changes. The old version
% is still available for reference as oldplainhscode.

\newenvironment{plainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\hspre\(\let\hspost\)%
   \pboxed}%
  {\endpboxed%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newenvironment{oldplainhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

% Here, we make plainhscode the default environment.

\newcommand{\plainhs}{\sethscode{plainhscode}}
\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
\plainhs

% The arrayhscode is like plain, but makes use of polytable's
% parray environment which disallows page breaks in code blocks.

\newenvironment{arrayhscode}%
  {\hsnewpar\abovedisplayskip
   \advance\leftskip\mathindent
   \hscodestyle
   \let\\=\@normalcr
   \(\parray}%
  {\endparray\)%
   \hsnewpar\belowdisplayskip
   \ignorespacesafterend}

\newcommand{\arrayhs}{\sethscode{arrayhscode}}

% The mathhscode environment also makes use of polytable's parray 
% environment. It is supposed to be used only inside math mode 
% (I used it to typeset the type rules in my thesis).

\newenvironment{mathhscode}%
  {\parray}{\endparray}

\newcommand{\mathhs}{\sethscode{mathhscode}}

% texths is similar to mathhs, but works in text mode.

\newenvironment{texthscode}%
  {\(\parray}{\endparray\)}

\newcommand{\texths}{\sethscode{texthscode}}

% The framed environment places code in a framed box.

\def\codeframewidth{\arrayrulewidth}
\RequirePackage{calc}

\newenvironment{framedhscode}%
  {\parskip=\abovedisplayskip\par\noindent
   \hscodestyle
   \arrayrulewidth=\codeframewidth
   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
   \hline\framedhslinecorrect\\{-1.5ex}%
   \let\endoflinesave=\\
   \let\\=\@normalcr
   \(\pboxed}%
  {\endpboxed\)%
   \framedhslinecorrect\endoflinesave{.5ex}\hline
   \endtabular
   \parskip=\belowdisplayskip\par\noindent
   \ignorespacesafterend}

\newcommand{\framedhslinecorrect}[2]%
  {#1[#2]}

\newcommand{\framedhs}{\sethscode{framedhscode}}

% The inlinehscode environment is an experimental environment
% that can be used to typeset displayed code inline.

\newenvironment{inlinehscode}%
  {\(\def\column##1##2{}%
   \let\>\undefined\let\<\undefined\let\\\undefined
   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
   \def\fromto##1##2##3{##3}%
   \def\nextline{}}{\) }%

\newcommand{\inlinehs}{\sethscode{inlinehscode}}

% The joincode environment is a separate environment that
% can be used to surround and thereby connect multiple code
% blocks.

\newenvironment{joincode}%
  {\let\orighscode=\hscode
   \let\origendhscode=\endhscode
   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
   %\let\SaveRestoreHook=\empty
   %\let\ColumnHook=\empty
   %\let\resethooks=\empty
   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
  {\origendhscode
   \global\let\hscode=\orighscode
   \global\let\endhscode=\origendhscode}%

\makeatother
\EndFmtInput
%

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End:

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{todonotes}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{mathdots}
\usepackage{yhmath}
\usepackage{cancel}
\usepackage{color}
\usepackage{siunitx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{xspace}
\usetikzlibrary{fadings}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros
\newcommand{\tocite}{\textbf{CITATION NEEDED}}
\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
\newcommand{\dragen}{\textbf{DRAGEN}\xspace}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title
\title{Synthesizing Random Generators via Higher-Order Representations}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Authors
\author{
  \IEEEauthorblockN{
    % 1\textsuperscript{st}
    Agust\'in Mista
  }
  \IEEEauthorblockA{
    % \textit{Department of Computer Science and Engineering} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    mista@chalmers.se
  }
\and
\IEEEauthorblockN{
  % 2\textsuperscript{nd}
  Alejandro Russo
  }
  \IEEEauthorblockA{
    % \textit{Department of Computer Science and Engineering} \\
    \textit{Chalmers University of Technology}\\
    Gothenburg, Sweden \\
    russo@chalmers.se
  }
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstract

\begin{abstract}

  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
  purus, et dictum nunc justo sit amet elit.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Keywords
\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections

\section{Introduction}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat\cite{grieco2017}.

The main contribuitions of this paper are:
%
\begin{itemize}
  %
\item We identify two patological scenarios for which standard type-driven
  automatic derivation tools fail to synthesize practical random generators, due
  to a lack of either type structure or domain knowleadge (Section
  \ref{sec:randomtesting}).
  %
\item We present a generation technique able to encode stronger properties of
  the target data by reifying the static information present on the program
  codebase (Section \ref{sec:hrep}).
  %
\item We apply and extend the theory of branching processes to analitically
  predict the average distribution of generated values.
  %
  Furthermore, we use the predictions to perform simulation-based optimization
  of the random generation parameters (Section \ref{sec:synthesis}).
  %
\item We provide an implementation of our ideas in the form of a Haskell library
  to perform automatic derivation of random generators capable to extract
  useful structure information from the user source code.
  %
\end{itemize}
\section{Type-Driven Random Generation} \label{sec:randomtesting}

In this section we briefly describe the common technique used for automatically
deriving random data generators in Haskell using a type-driven approach.
%
Then, we introduce its main drawbacks by considering two scenarios where this
technique gives poor results in practice.
%
We remark that, altough this work makes focus on Haskell data types, this
technique is general enough to be applied to most programming languages with
some level of support for composite types.

Haskell is a strongly typed programming language with a powerful type system.
%
It lets programmers encode a considerable amount of information about the
properties of their systems using data types that can be checked at compilation
time.
%
One of its key aspects is the support for Algebraic Data Types (ADTs).
%
Essentially, an ADTs is a composite type defined by combining other types.
%
In the most basic conception, types can be combined by \textbf{sums} (also known
as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
(or tuples) of other data types.
%
To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
encoding integer expressions:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
\>[11]{}\mathrel{=}{}\<[15]%
\>[15]{}\Conid{Val}\;{}\<[20]%
\>[20]{}\Conid{Int}{}\<[E]%
\\
\>[11]{}\;\;\vert\;{}\<[15]%
\>[15]{}\Conid{Add}\;{}\<[20]%
\>[20]{}\Conid{Exp}\;{}\<[25]%
\>[25]{}\Conid{Exp}{}\<[E]%
\\
\>[11]{}\;\;\vert\;{}\<[15]%
\>[15]{}\Conid{Mul}\;{}\<[20]%
\>[20]{}\Conid{Exp}\;{}\<[25]%
\>[25]{}\Conid{Exp}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
represent the addition and multiplication of two integer expressions,
respectively.
%
In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
used to distiguish which variant of the data type we are reffering to.
%
Each data constructor is then defined as a product of zero or more types known
as \emph{fields}.
%
In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
%
Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
%
Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
very idiomatically:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
\column{27E}{@{}l@{}}%
\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
\\
\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
\\
\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
\>[27]{}\mathbin{+}{}\<[27E]%
\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
\\
\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
\>[27]{}\mathbin{*}{}\<[27E]%
\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
over each possible kind of value.
%
For the input case of a literal values, we simply return the value contained in
the \ensuremath{\Conid{Val}} constructor.
%
On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
subexpression, we recursively evaluate these subexpressions, combining them with
the apropiate operation on each case.

\subsection{\textbf{Type-Driven Generation of Random Values}}

In order to perform random testing of a Haskell codebase involving user defined
data types, most approaches require the user to provide a random data generator
for each one of them.
%
This tends to be a cumbersome and error prone task that closely follows the data
type structure.
%
For instance, consider the following definition of a \quickcheck random
generator for the type \ensuremath{\Conid{Exp}}:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
\column{7E}{@{}l@{}}%
\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathbf{where}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
\\
\>[5]{}\hsindent{2}{}\<[7]%
\>[7]{}[\mskip1.5mu {}\<[7E]%
\>[10]{}\Conid{Val}{}\<[15]%
\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
\\
\>[5]{}\hsindent{2}{}\<[7]%
\>[7]{},{}\<[7E]%
\>[10]{}\Conid{Add}{}\<[15]%
\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
\\
\>[5]{}\hsindent{2}{}\<[7]%
\>[7]{},{}\<[7E]%
\>[10]{}\Conid{Mul}{}\<[15]%
\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
 %$

This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
integer number known as the \emph{generation size}.
%
This parameter is used to limit the maximum amount of recursive call that this
random generator can perform.
%
When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
%
In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
\ensuremath{\Conid{Bool}}, etc.).
%
On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
data constructor, it also generates independently two random subexpressions
corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
invocation.

This procedure keeps calling itself recursively until the generation size
reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
particular case.
%
Strictly decreasing the the generation size by one on each recursive call
results in a useful property over the generated data: every generated value has
at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
%
This property enables us to model the generation process using the theory of
branching processes introduced by \tocite, and extended in this work as
described in Section \ref{sec:synthesis}.

Having discussed the previous random generator definition, it easy to understand
how to extend this generation procedure to any data type defined in an algebraic
fashion.
%
Fortunately, there exists different meta-programming tools to avoid the user
from having to mechanically write random generators over and over again for each
user-defined ADT.
%
The simplest tool for such purpose is \megadeth.
%
Given the name of the target data type, it synthesizes a random generator for it
that behaves pretty much like the previously defined one.
%
However, picking among different data constructors with uniform probability can
lead to a generation process biased to generate (in average) very small values,
regardless of the generation sized set by the user \tocite.

\dragen is meta-programming tool conceived to mitigate this problem.
%
Instead of setting a uniform generation probability of data constructors, this
tool considers the scenario where each one can be generated following a
different (although fixed over time) generation probability.
%
Then, this tool uses the theory of branching processes to analitically predict
the average distribution of data constructors generated on each random value.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation probability of each data constructor in
order to obtain a particular distribution of values that can be specified by the
user, offering a more flexible testing environment while still being mostly
automated.

Both \megadeth and \dragen synthesize random generators that are theoretically
capable to generate the whole space of values of the target data type, provided
that we set a generation size suficiently big.
%
However, the limitations arise quickly when we consider that the underlying
generation model is essentially the same: they pick a single data constructor
and recusively generate each required subexpression independently.
%
In practice, this procedure is too generic to generate complex data that is
useful enough to be used for random testing.

In this work we identify two sources of complications that appear when using the
aforementioned automatic derivation tools to obtain random data generators:

\begin{enumerate}
\item The target code behaves differently over inputs matching very specific
  patterns.
\item The target code behaves differently according to some dependency over
  subexpressions of its inputs.
\end{enumerate}

To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
to test a property comprising the following function:

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
\\
\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
\\
\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks

This function behaves mostly like an identity function, except for two very
specific patterns of input values defined using three nested levels of pattern
matching.
%
This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
against the data constructors of its subexpressions and sub-subexpressions.

If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
a number in the interval $[-n, n]$ with uniform distribution, we need to be able
to identify which one is a sensible generation size to set based on wether our
code branches on numbers or not, otherwise our random generator will not be able
to generate random values to match against such numbers, leaving fragments of
code completely untested.
%
Suppose that we pick the a generation size $50$, i.e. the minimum size that is
big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
%
Under this considerations, the probability of generating a value matching the
first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:

\begin{align*}
  P(match(foo_1))
  \!&&=\; &P(Add) \\
  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
\end{align*}

\todo[inline]{AM: COMPLETE}

Despite the fact that the previous example might look artificial, branching
against specific patterns of the input data is a common task.
%
For instance, the balancing function of a Red-Black tree need to consider a
quite specific combination of color, left- and right subtrees and sub-subtrees
in order to preserve the height invariant.
%
Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
needs to consider very specific sequences of instructions that may be regrouped
in a computationally more efficient way.

We believe that if a certain pattern of values is relevant enough to appear in
the codebase being tested, a practical random testing methodology should be able
to produce values satisfying such pattern in a substantial proportion.

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
\\
\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

% \begin{code}
% data IExp  =   VAL Int
%            pp  ADD IExp IExp
%            pp  IF BExp IExp IExp

% data BExp  =   TRUE | FALSE
%            pp  LE IExp IExp
%            pp  AND BExp BExp
% \end{code}

% \begin{code}
% foo :: IExp -> IExp
% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
% foo x                                = ...

% bar :: BExp -> BExp
% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
% bar x                                = ...
% \end{code}

% \begin{code}
% twice :: IExp -> IExp
% twice e = ADD e e

% equals :: IExp -> IExp -> BExp
% equals x y = AND (LE x y) (LE y x)

% and3 :: BExp -> BExp -> BExp -> BExp
% and3 x y z = AND x (AND y z)
% \end{code}
\section{Extracting Structure} \label{sec:hrep}

What to say here:

\begin{itemize}
\item Haskell ADTs
\item Type driven generation
\item Limitations
\item Key idea: Higher order representations
\item Canonical representation of the data type
\item Pattern matchings representation
\item Abstract interface representation
\item Composing representations
\end{itemize}

\begin{figure}[b]
  \centering
  \input{tikz/hrep.tex}
  \caption{The idea}
\end{figure}

Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
lacinia purus, et dictum nunc justo sit amet elit.

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
\section{Random Generators Synthesis} \label{sec:synthesis}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.
\section{Case Studies}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
\section{Related Work}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.
\section{Final Remarks}

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
\bibliographystyle{IEEEtran}
\bibliography{references.bib}

\end{document}

### 2019-01-15 paper/sections/finalremarks.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
-\end{enumerate}
-
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
-
-\begin{align*}
-  P(match(foo_1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-\todo[inline]{AM: COMPLETE}
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
-
-\begin{figure}[b]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{The idea}
-\end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
 \section{Final Remarks}
@@
 Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/sections/hrep.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
-\end{enumerate}
-
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
-
-\begin{align*}
-  P(match(foo_1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-\todo[inline]{AM: COMPLETE}
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}
 \section{Extracting Structure} \label{sec:hrep}
@@
 est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/sections/introduction.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
 \section{Introduction}
@@
 \end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
-\end{enumerate}
-
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
-
-\begin{align*}
-  P(match(foo_1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-\todo[inline]{AM: COMPLETE}
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
-
-\begin{figure}[b]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{The idea}
-\end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/sections/randomtesting.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
 \section{Type-Driven Random Generation} \label{sec:randomtesting}
@@
 %
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
+To exemplify this, consider the following definition of the data type |Exp|
 encoding integer expressions:
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\begin{code}
+data Exp  =   Val  Int
+          pp  Add  Exp  Exp
+          pp  Mul  Exp  Exp
+\end{code}
 
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
+In the previous definition, we declare |Exp| as the sum of three possible
+classes of values: |Val| represents literal values, whereas |Add| and |Mul|
 represent the addition and multiplication of two integer expressions,
@@
 %
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
+In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} and are
 used to distiguish which variant of the data type we are reffering to.
@@
 %
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
+In particular, |Val| contains a field of type |Int|, while |Add| and |Mul|
+contain two fields of type |Exp| representing the operands of each operation.
 %
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
+Note that |Exp| is used as a field of at least one of its data constructors
+(case |Add| and |Mul|), making it a recursively defined ADT.
 %
@@
 With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
+using |Exp| data constructors as |(Add (Val 2) (Mul (Val 5) (Val 6)))|.
 %
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
+Furthermore, an evaluation function from |Exp|s to integer values can be defined
 very idiomatically:
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\begin{code}
+eval :: Exp -> Int
+eval (Val n)    = n
+eval (Add x y)  = eval x  +  eval y
+eval (Mul x y)  = eval x  *  eval y
+\end{code}
 
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
+In the previous definition, |eval| is described using \emph{pattern matching}
 over each possible kind of value.
@@
 For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
+the |Val| constructor.
 %
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
+On the other hand, if the input value matches either an |Add| or a |Mul| data
+constructor, where |x| and |y| are \emph{pattern variables} that match any
 subexpression, we recursively evaluate these subexpressions, combining them with
@@
 For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
+generator for the type |Exp|:
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
+\begin{code}
+genExp :: Gen Exp
+genExp = sized gen
+  where
+    gen 0 = Val <$> arbitrary
+    gen n = oneof
+      [  Val  <$> arbitrary
+      ,  Add  <$> gen pred   <*> gen pred
+      ,  Mul  <$> gen pred   <*> gen pred ]
+\end{code} %$
 
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
+This random generator is defined using \quickcheck's combinator |sized :: (Int
+-> Gen a) -> Gen a| to paremetrize the generation process up to an external
 integer number known as the \emph{generation size}.
@@
 %
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
+When called with a positive generation size (case |gen n|), the generator is free
+to randomly pick with uniform probability among any data constructor of |Exp|.
 %
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
+In the case it picks to generate a |Val| data constructor, it also generates a
+random |Int| value using the standard overloaded generator |arbitrary|
+(\quickcheck provides standard random generators for most base types like |Int|,
+|Bool|, etc.).
 %
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
+On the other hand, when it randomly picks to generate either an |Add| or a |Mul|
 data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
+corresponding to the data constructor fields by calling itself recursively (|gen
+pred|), ensuring to decrease the generation size by a unit on each recursive
 invocation.
@@
 This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
+reaches zero (case |gen 0|), where the generator is constrained to pick only
+among terminal data constructors, being |Val| the only possible choice in our
 particular case.
@@
 results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
+at most |n| levels, where |n| is the generation size set by the user.
 %
@@
 
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
+To exemplify the first point, suppose we want to use randomly generated |Exp|s
 to test a property comprising the following function:
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\begin{code}
+foo :: Exp -> Exp
+foo (Add (Add x (Val 50)) (Add (Val 25) y))
+  = error "pattern #1"
+foo (Mul (Mul x (Val 50)) (Mul (Val 25) y))
+  = error "pattern #2"
+foo x = x
+\end{code}
 
@@
 %
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
+This is, |foo| does not only matches against the root data constructor, but also
 against the data constructors of its subexpressions and sub-subexpressions.
 
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
+If we consider that \quickcheck's default implementation for random |Int|s pick
 a number in the interval $[-n, n]$ with uniform distribution, we need to be able
@@
 Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
+big enough to produce an |Int| number equal to $50$.
 %
 Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
+first clause of |foo| (and hence triggering the error) are as follows:
 
@@
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\begin{code}
+fortyTwo :: Exp
+fortyTwo = Val 42
+
+twice :: Exp -> Exp
+twice x = Add x x
+
+square :: Exp -> Exp
+square x = Mul x x
+
+neg :: Exp -> Exp
+neg x = Mul x (Val (-1))
+
+minus :: Exp -> Exp -> Exp
+minus x y = Add x (neg y)
+\end{code}
 
@@
 % \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
-
-\begin{figure}[b]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{The idea}
-\end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/sections/relatedwork.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
-\end{enumerate}
-
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
-
-\begin{align*}
-  P(match(foo_1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-\todo[inline]{AM: COMPLETE}
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
-
-\begin{figure}[b]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{The idea}
-\end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
 \section{Related Work}
@@
 Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/sections/synthesis.tex
Sync

--- 2019-01-15 paper/main.tex
@@
-\documentclass[conference]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-master: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
-\end{enumerate}
-
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
-
-\begin{align*}
-  P(match(foo_1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-\todo[inline]{AM: COMPLETE}
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
-
-\begin{figure}[b]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{The idea}
-\end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
 \section{Random Generators Synthesis} \label{sec:synthesis}
@@
 Nam vestibulum accumsan nisl.
-\section{Case Studies}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-15 paper/tikz/hrep.tex
Sync

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,416.5); %set diagram left start at 0, and has height of 416.5

%Shape: Rectangle [id:dp9489660675592899] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (130,20) .. controls (130,14.48) and (134.48,10) .. (140,10) -- (400,10) .. controls (405.52,10) and (410,14.48) .. (410,20) -- (410,80) .. controls (410,85.52) and (405.52,90) .. (400,90) -- (140,90) .. controls (134.48,90) and (130,85.52) .. (130,80) -- cycle ;
%Snip Same Side Corner Rect [id:dp09931070350524518] 
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,30.5) -- (150.6,20) -- (189.5,20) -- (200,30.5) -- (200,50) -- (200,50) -- (140.1,50) -- (140.1,50) -- cycle ;
%Down Arrow [id:dp5984296172784] 
\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (140,159.5) -- (144,159.5) -- (144,90) -- (152,90) -- (152,159.5) -- (156,159.5) -- (148,170) -- cycle ;
%Straight Lines [id:da30879479079788386] 
\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]    (145,90) -- (151,90) ;

%Straight Lines [id:da5824111707250128] 
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (170,170) .. controls (168.33,168.33) and (168.33,166.67) .. (170,165) .. controls (171.67,163.33) and (171.67,161.67) .. (170,160) .. controls (168.33,158.33) and (168.33,156.67) .. (170,155) .. controls (171.67,153.33) and (171.67,151.67) .. (170,150) .. controls (168.33,148.33) and (168.33,146.67) .. (170,145) .. controls (171.67,143.33) and (171.67,141.67) .. (170,140) .. controls (168.33,138.33) and (168.33,136.67) .. (170,135) .. controls (171.67,133.33) and (171.67,131.67) .. (170,130) .. controls (168.33,128.33) and (168.33,126.67) .. (170,125) .. controls (171.67,123.33) and (171.67,121.67) .. (170,120) .. controls (168.33,118.33) and (168.33,116.67) .. (170,115) .. controls (171.67,113.33) and (171.67,111.67) .. (170,110) .. controls (168.33,108.33) and (168.33,106.67) .. (170,105) .. controls (171.67,103.33) and (171.67,101.67) .. (170,100) .. controls (168.33,98.33) and (168.33,96.67) .. (170,95) -- (170,91) -- (170,83) ;
\draw [shift={(170,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) .. controls (9.04,-1.82) and (4.3,-0.39) .. (0,0) .. controls (4.3,0.39) and (9.04,1.82) .. (14.21,4.28)   ;

%Straight Lines [id:da7051587981224026] 
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (270,110) .. controls (268.33,108.33) and (268.33,106.67) .. (270,105) .. controls (271.67,103.33) and (271.67,101.67) .. (270,100) .. controls (268.33,98.33) and (268.33,96.67) .. (270,95) -- (270,91) -- (270,83) ;
\draw [shift={(270,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) .. controls (9.04,-1.82) and (4.3,-0.39) .. (0,0) .. controls (4.3,0.39) and (9.04,1.82) .. (14.21,4.28)   ;

%Straight Lines [id:da6087008244886778] 
\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (370,110) .. controls (368.33,108.33) and (368.33,106.67) .. (370,105) .. controls (371.67,103.33) and (371.67,101.67) .. (370,100) .. controls (368.33,98.33) and (368.33,96.67) .. (370,95) -- (370,91) -- (370,83) ;
\draw [shift={(370,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) .. controls (9.04,-1.82) and (4.3,-0.39) .. (0,0) .. controls (4.3,0.39) and (9.04,1.82) .. (14.21,4.28)   ;

%Shape: Rectangle [id:dp701556783783656] 
\draw  [line width=1.5]  (230,115) .. controls (230,112.24) and (232.24,110) .. (235,110) -- (305,110) .. controls (307.76,110) and (310,112.24) .. (310,115) -- (310,135) .. controls (310,137.76) and (307.76,140) .. (305,140) -- (235,140) .. controls (232.24,140) and (230,137.76) .. (230,135) -- cycle ;

%Shape: Rectangle [id:dp033430127809365384] 
\draw  [line width=1.5]  (330,115) .. controls (330,112.24) and (332.24,110) .. (335,110) -- (405,110) .. controls (407.76,110) and (410,112.24) .. (410,115) -- (410,135) .. controls (410,137.76) and (407.76,140) .. (405,140) -- (335,140) .. controls (332.24,140) and (330,137.76) .. (330,135) -- cycle ;
%Shape: Rectangle [id:dp6050538306488522] 
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,50) -- (200,50) -- (200,80) -- (140.1,80) -- cycle ;
%Shape: Rectangle [id:dp9303311048687424] 
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (340,50) -- (400,50) -- (400,80) -- (340,80) -- cycle ;

%Up Arrow [id:dp3837429966878143] 
\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (360,160) -- (370,150) -- (380,160) -- (380,160) -- (380,160) -- (360,160) -- (360,160) -- cycle ;
%Up Arrow [id:dp0013771056350821986] 
\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,160) -- (270,150) -- (280,160) -- (280,160) -- (280,160) -- (260,160) -- (260,160) -- cycle ;
%Shape: Rectangle [id:dp7366907361635939] 
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.43,50) -- (300,50) -- (300,80) -- (240.43,80) -- cycle ;
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (210,65) .. controls (210,59.48) and (214.48,55) .. (220,55) .. controls (225.52,55) and (230,59.48) .. (230,65) .. controls (230,70.52) and (225.52,75) .. (220,75) .. controls (214.48,75) and (210,70.52) .. (210,65) -- cycle ; \draw  [line width=1.5]  (210,65) -- (230,65) ; \draw  [line width=1.5]  (220,55) -- (220,75) ;
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (310,65) .. controls (310,59.48) and (314.48,55) .. (320,55) .. controls (325.52,55) and (330,59.48) .. (330,65) .. controls (330,70.52) and (325.52,75) .. (320,75) .. controls (314.48,75) and (310,70.52) .. (310,65) -- cycle ; \draw  [line width=1.5]  (310,65) -- (330,65) ; \draw  [line width=1.5]  (320,55) -- (320,75) ;
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (152.93,28.79) .. controls (156.83,24.88) and (163.17,24.88) .. (167.07,28.79) .. controls (170.98,32.69) and (170.98,39.02) .. (167.07,42.93) .. controls (163.17,46.83) and (156.83,46.83) .. (152.93,42.93) .. controls (149.02,39.02) and (149.02,32.69) .. (152.93,28.79) -- cycle ; \draw  [line width=1.5]  (152.93,28.79) -- (167.07,42.93) ; \draw  [line width=1.5]  (167.07,28.79) -- (152.93,42.93) ;
%Snip Same Side Corner Rect [id:dp4221187902899477] 
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.1,30.5) -- (250.6,20) -- (289.5,20) -- (300,30.5) -- (300,50) -- (300,50) -- (240.1,50) -- (240.1,50) -- cycle ;
\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (252.93,28.79) .. controls (256.83,24.88) and (263.17,24.88) .. (267.07,28.79) .. controls (270.98,32.69) and (270.98,39.02) .. (267.07,42.93) .. controls (263.17,46.83) and (256.83,46.83) .. (252.93,42.93) .. controls (249.02,39.02) and (249.02,32.69) .. (252.93,28.79) -- cycle ; \draw  [line width=1.5]  (252.93,28.79) -- (267.07,42.93) ; \draw  [line width=1.5]  (267.07,28.79) -- (252.93,42.93) ;
%Shape: Rectangle [id:dp1692119124811613] 
\draw  [line width=1.5]  (130,170) -- (410,170) -- (410,200) -- (130,200) -- cycle ;

% Text Node
\draw (270,185) node [scale=1.2]  {$T$};
% Text Node
\draw (270,125) node [scale=1]  {$\text{Patterns}^{T}_{foo}$};
% Text Node
\draw (370,125) node [scale=1]  {$\text{Interface}^{T}_{M}$};
% Text Node
\draw (170.05,65) node [scale=0.9]  {$\text{HRep}^{T}_{T}$};
% Text Node
\draw (370,65) node [scale=0.9]  {$\text{HRep}^{T}_{M}$};
% Text Node
\draw (114.5,125) node [scale=1.2]  {$S_{T}\Downarrow T$};
% Text Node
\draw (392,25) node [scale=1.2]  {$S_{T}$};
% Text Node
\draw (270.22,65) node [scale=0.9]  {$\text{HRep}^{T}_{foo}$};
% Text Node
\draw (183.05,35) node [scale=1.44]  {$\alpha $};
% Text Node
\draw (284.05,35) node [scale=1.44]  {$\beta $};

\end{tikzpicture}

### 2019-01-15 paper/code/Exp.hs

@@ -1,0 +1,46 @@
+module Exp where
+
+data Exp  =  Val  Int
+          |  Add  Exp  Exp
+          |  Mul  Exp  Exp
+          |  If   Exp  Exp  Exp
+          deriving Show
+
+eval :: Exp -> Int
+eval (Val n) = n
+eval (Add x y) = eval x + eval y
+eval (Mul x y) = eval x * eval y
+eval (If b x y) = if (eval b /= 0) then eval x else eval y
+
+-- genExp :: Gen Exp
+-- genExp = sized gen
+--   where
+--     gen 0 = Val <$> genInt
+--     gen n = oneof
+--       [  Val  <$> arbitrary
+--       ,  Add  <$> gen pred   <*> gen pred
+--       ,  Mul  <$> gen pred   <*> gen pred
+--       ,  If   <$> arbitrary  <*> gen pred <*> gen pred ]
+
+
+-- foo :: Exp -> Exp
+-- foo (Mul (Add x (Val 42)) (Mul (Val 0) y))  = ...
+-- foo (Add (Add x y) (Val z))                 = ...
+-- foo (Add (Val 42)  (Val 24))                = ...
+-- foo x                                       = ...
+
+(.+) = Add
+(.*) = Mul
+val = Val
+
+square :: Exp -> Exp
+square x = Mul x x
+
+neg :: Exp -> Exp
+neg x = Mul x (Val (-1))
+
+minus :: Exp -> Exp -> Exp
+minus x y = Add x (neg y)
+
+pow :: Exp -> Exp -> Exp
+pow b e = If e (Mul b (pow b (minus e (Val 1)))) (Val 1)

### 2019-01-15 paper/watch.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+
+#https://facebook.github.io/watchman/docs/watchman-make.html
+watchman-make --make 'make' -p '**/[^.]*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all

### 2019-01-15 src/Test/QuickCheck/Branching.hs

@@ -1,5 +1,6 @@
 {-# OPTIONS_GHC -Wno-orphans #-}
 {-# LANGUAGE AllowAmbiguousTypes #-}
+{-# LANGUAGE BangPatterns #-}
 {-# LANGUAGE ConstraintKinds #-}
 {-# LANGUAGE DataKinds #-}
 {-# LANGUAGE FlexibleContexts #-}
@@ -13,8 +14,13 @@
 {-# LANGUAGE TypeFamilies #-}
 {-# LANGUAGE TypeOperators #-}
 {-# LANGUAGE UndecidableInstances #-}
+{-# LANGUAGE QuantifiedConstraints #-}
 
 module Test.QuickCheck.Branching where
+
+import Data.Maybe
+import Data.List
+import Data.Ord
 
 import Data.Kind
 import Data.Proxy
@@ -34,6 +40,9 @@ import qualified Data.Vector as Vector
 
 import Data.Map.Strict (Map)
 import qualified Data.Map.Strict as Map
+
+import System.IO.Unsafe
+import System.IO
 
 import Debug.Trace
 
@@ -95,6 +104,46 @@ instance (BranchingType f, BranchingType g, f #> ff, f @> ff', g #> gg, g @> gg'
     where ff' = numVal @ff'; gg' = numVal @gg'; tot' = ff' + gg'
   typeBeta = typeBeta @f <> typeBeta @g
   typeEta  = typeEta  @f <> typeEta  @g
+
+instance (forall a. BranchingType (f a)) => BranchingType (Some1 f) where
+  typeNames  = typeNames  @(f ())
+  typeAtomic = typeAtomic @(f ())
+  typeProbs  = typeProbs  @(f ())
+  typeProbs' = typeProbs' @(f ())
+  typeBeta   = typeBeta   @(f ())
+  typeEta    = typeEta    @(f ())
+
+instance (forall a b. BranchingType (f a b)) => BranchingType (Some2 f) where
+  typeNames  = typeNames  @(f () ())
+  typeAtomic = typeAtomic @(f () ())
+  typeProbs  = typeProbs  @(f () ())
+  typeProbs' = typeProbs' @(f () ())
+  typeBeta   = typeBeta   @(f () ())
+  typeEta    = typeEta    @(f () ())
+
+instance (forall a b c. BranchingType (f a b c)) => BranchingType (Some3 f) where
+  typeNames  = typeNames  @(f () () ())
[…]

### 2019-01-15 src/Test/QuickCheck/HRep.hs

@@ -74,7 +74,7 @@ deriving instance (Generic (f a), Generic (g a)) => Generic (Sum f g a)
 class Functor f => Algebra f a | f -> a where
   alg :: f a -> a
 
--- | Catamorphisms, a.k.a generic foldings over fix points
+-- | Catamorphisms
 cata :: Functor f => (f a -> a) -> Fix f -> a
 cata f = f . fmap (cata f) . unFix
 
@@ -182,7 +182,7 @@ type family HRep (t :: Symbol) :: Type -> Type
 -- | A single constructor
 type family Con (c :: k) :: Type -> Type
 
--- | A single pattern matchings
+-- | A single pattern matching
 type family Pat (p :: Symbol) (n :: Nat) :: Type -> Type
 
 -- | A single function in a module
@@ -208,7 +208,7 @@ data Some5 (f :: Type -> Type -> Type -> Type -> Type -> Type -> Type) (r :: Typ
   = forall (e :: Type) (d :: Type) (c :: Type) (b :: Type) (a :: Type)
   . Some5 (f a b c d e r)
 
-type family Apply (a :: Type) (t :: Type -> Type) where
+type family Apply (a :: Type) (t :: Type -> Type) :: Type -> Type where
   Apply a (Freq t n) = Freq (Apply a t) n
   Apply a (Term t) = Term (Apply a t)
   Apply a (Sum l r) = Sum (Apply a l) (Apply a r)

### 2019-01-15 src/Test/QuickCheck/HRep/Infix.hs

@@ -18,16 +18,16 @@ infixl 6 :*
 type (:+) = Sum
 infixl 5 :+
 
-type (spec :# k) = Lookup spec k
-infix 4 :#
+type (spec ~> k) = Lookup spec k
+infix 4 ~>
 
-type (t :@ a) = Apply a t
-infixl 3 :@
+type (t $ a) = Apply a t
+infixl 3 $
 
--- type (spec :@@ a) = ApplySpec a spec
--- infixl 3 :@@
+type (spec :@@ a) = ApplySpec a spec
+infixl 3 :@@
 
 type ((k :: kk) := (v :: vv)) = '(k, v)
 infix 2 :=
 
-type (f :> g) = Optimized f g
+-- type (f :> g) = Optimized f g

### 2019-01-15 src/Test/QuickCheck/HRep/TH.hs

@@ -1,9 +1,16 @@
+{-# LANGUAGE BangPatterns #-}
 {-# LANGUAGE RankNTypes #-}
 {-# LANGUAGE TemplateHaskell #-}
 {-# LANGUAGE FlexibleInstances #-}
 {-# LANGUAGE StandaloneDeriving #-}
+{-# LANGUAGE TypeApplications #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE AllowAmbiguousTypes #-}
 
 module Test.QuickCheck.HRep.TH where
+
+import Data.Proxy
 
 import Control.Monad.Extra
 
@@ -13,6 +20,7 @@ import Language.Haskell.TH.Desugar
 import Test.QuickCheck.HRep.TH.TypeRep
 import Test.QuickCheck.HRep.TH.FunPatsRep
 import Test.QuickCheck.HRep.TH.ModIntRep
+import Test.QuickCheck.Branching
 
 import Test.QuickCheck.HRep.TH.Common
 
@@ -47,11 +55,11 @@ module_ tyName = ModuleInterface tyName ("<" ++ nameBase tyName ++ ">") []
 
 derive :: Target -> Q [Dec]
 derive (TypeDefinition tyName tyFam)
-  = deriveTypeDefinitionRep tyName tyFam
+  = deriveTypeRep tyName tyFam
 derive (FunctionPatterns funName funArgNr tyFam)
-  = deriveFunctionPatternsRep funName funArgNr tyFam
-derive (ModuleInterface tyName modAlias _)
-  = deriveModuleInterfaceRep tyName modAlias
+  = deriveFunPatsRep funName funArgNr tyFam
+derive (ModuleInterface tyName modAlias tyFam)
+  = deriveModIntRep tyName modAlias tyFam
 
 deriveHRep :: [Name] -> [Target] -> Q [Dec]
 deriveHRep fam' targets = concatMapM derive
@@ -93,3 +101,29 @@ deriveAll tyFam targets specName = do
   arbIns   <- concatMapM (\tn -> deriveArbitrary tn specName) tyFam
   tgtReps  <- concatMapM (\target -> derive (target {fam = tyFam})) targets
   return (concat [typeReps, arbIns, tgtReps])
+
+----------------------------------------
+-- | Optimize the frequencies of a specification
+
+optimize :: forall spec root. HasSpec spec root
+         => Name -> String -> DistFun -> QCSize -> Q [Dec]
[…]

### 2019-01-15 src/Test/QuickCheck/HRep/TH/FunPatsRep.hs

@@ -20,8 +20,8 @@ import Test.QuickCheck.HRep.TH.Common
 ----------------------------------------
 -- | Derive the complete representation for a every pattern of a function,
 -- plus the function representation as a sum of each pattern.
-deriveFunctionPatternsRep :: Name -> Int -> [Name] -> Q [Dec]
-deriveFunctionPatternsRep funName argNr tyFam = do
+deriveFunPatsRep :: Name -> Int -> [Name] -> Q [Dec]
+deriveFunPatsRep funName argNr tyFam = do
 
   -- | Retrieve the LHS of the target function
   FunLHS _ funSig funPats <- reifyFunLHS funName

### 2019-01-15 src/Test/QuickCheck/HRep/TH/ModIntRep.hs

@@ -12,7 +12,7 @@ import qualified Language.Haskell.Exts as Ext
 
 import qualified Test.QuickCheck as QC
 import qualified Test.QuickCheck.HRep as HRep
--- import qualified Test.QuickCheck.Branching as Branching
+import qualified Test.QuickCheck.Branching as Branching
 
 import Test.QuickCheck.HRep.TH.Common
 
@@ -20,8 +20,8 @@ import Test.QuickCheck.HRep.TH.Common
 -- | Derive the complete representation of every combinator of a module that
 -- returns the desired type, plus the module interface representation as a sum
 -- of each combinator.
-deriveModuleInterfaceRep :: Name -> String -> Q [Dec]
-deriveModuleInterfaceRep tyName modAlias = do
+deriveModIntRep :: Name -> String -> [Name] -> Q [Dec]
+deriveModIntRep tyName modAlias tyFam = do
 
   -- | Convert the module alias to a safe identifier
   let modAliasId = toIdentStr modAlias
@@ -30,7 +30,7 @@ deriveModuleInterfaceRep tyName modAlias = do
   combNames <- filterM (returnsType tyName) =<< reifyModNames
 
   -- | Create the representation for each combinator
-  combsRep <- concatMapM (deriveCombRep tyName modAliasId) combNames
+  combsRep <- concatMapM (deriveCombRep tyName modAliasId tyFam) combNames
 
   -- | Create the default HRep type instance
   let repTyIns = DTySynInstD ''HRep.HRep repTyInsEqn
@@ -54,8 +54,8 @@ returnsType tyName funName
 ----------------------------------------
 -- | Derive the complete representation of a single value of the abstract
 -- interface of a module.
-deriveCombRep :: Name -> String -> Name -> Q [DDec]
-deriveCombRep tyName modAlias funName = do
+deriveCombRep :: Name -> String -> [Name] -> Name -> Q [DDec]
+deriveCombRep tyName modAlias tyFam funName = do
 
   -- | Some "fresh" names
   repName    <- newName ("HRep_Comb_" ++ toIdentStr modAlias ++ "_" ++ nameBase funName)
@@ -107,6 +107,7 @@ deriveCombRep tyName modAlias funName = do
 
       mkCxt v = DAppPr (DConPr ''QC.Arbitrary) (DVarT v)
 
+  -- | Representation HRep type instance
   let repFunTyIns = DTySynInstD ''HRep.Fun repFunInsEqn
[…]

### 2019-01-15 src/Test/QuickCheck/HRep/TH/TypeRep.hs

@@ -18,8 +18,8 @@ import Test.QuickCheck.HRep.TH.Common
 ----------------------------------------
 -- | Derive the complete representation for every constructor of a type, plus
 -- the type representation as a sized sum of each constructor.
-deriveTypeDefinitionRep :: Name -> [Name] -> Q [Dec]
-deriveTypeDefinitionRep tyName tyFam = do
+deriveTypeRep :: Name -> [Name] -> Q [Dec]
+deriveTypeRep tyName tyFam = do
 
   -- | Reify the original data declaration name and desugar it
   (vs, cons) <- getDataD mempty tyName

### 2019-01-15 test/PredMutRec.hs

@@ -28,233 +28,141 @@ import Test.QuickCheck.Branching
 import Test.QuickCheck.HRep
 import Test.QuickCheck.HRep.Infix
 import Test.QuickCheck.HRep.TH
-import Test.QuickCheck.HRep.TH.Common
 
 ----------------------------------------
-
-type Var = Char
-
-data IExp a
-  = Val a
-  | Add (IExp a) (IExp a)
-  | If  (BExp a) (IExp a) (IExp a)
-  deriving Show
-
-data BExp a
-  = Bool Bool
-  | LEq  (IExp a) (IExp a)
-  deriving Show
-
-foo :: IExp a -> IExp a
-foo (Add (Add (Val a)   _i1) _i2) = undefined
-foo (Add (If _b (Val a) _i1) _i2) = undefined
-foo _                             = undefined
-
-bar :: BExp a -> BExp a
-bar (LEq (Val a) _i2) = Bool True
-bar b                 = b
-
-
--- deriveHRep
---   [''IExp, ''BExp]
---   [ patterns_ 'foo
---   , patterns_ 'bar
---   , module_   ''IExp ]
--- deriveArbitrary ''ExpS [''IExp, ''BExp]
-
--- type ExpS = '[
---   "IExp" := HRep "IExp"
---          :| HRep "foo",
---   "BExp" := HRep "BExp"
---          :| HRep "bar"
---   ]
-
-type ExpS =
-  '[ "IExp"
-       := Term (Con 'Val) :* 1
-       :+ Con 'Add        :* 2
-       :+ Con 'If         :* 3
-   , "BExp"
-       := Term (Con 'Bool)
-       :+ Con 'LEq
-   ]
-
-deriveAll
-  [ ''IExp, ''BExp ]
-  [ patterns_ 'foo
-  , patterns_ 'bar
-  , module_   ''IExp ]
-  ''ExpS
-
-type instance Optimized "ExpS" "SomeDist"
-  = SetSpecFreqs (MkSpec ExpS)
-      '[ '[ 4, 5, 6]
-       , '[ 1, 1]
-       ]
-
--- dumpTHInfo ''SpecExp
-
--- type SpecExp' =
---   '[ IExp := Term (Con 'Val) :* 2
---           :| Con 'Add
---           :| Con 'If :* 3
---           :| Pat "foo" 1
---           :| Pat "foo" 2
---    , BExp := Term (Con 'Bool)
---           :| Con 'LEq
[…]

### 2019-01-15 test/PredMutRec_opt.hs

@@ -1,0 +1,28 @@
+{-# LANGUAGE AllowAmbiguousTypes #-}
+{-# LANGUAGE DataKinds #-}
+{-# LANGUAGE DeriveFunctor #-}
+{-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE ExistentialQuantification #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE FlexibleContexts #-}
+{-# LANGUAGE FlexibleInstances #-}
+{-# LANGUAGE GADTs #-}
+{-# LANGUAGE MultiParamTypeClasses #-}
+{-# LANGUAGE PolyKinds #-}
+{-# LANGUAGE Rank2Types #-}
+{-# LANGUAGE ScopedTypeVariables #-}
+{-# LANGUAGE StandaloneDeriving #-}
+{-# LANGUAGE TemplateHaskell #-}
+{-# LANGUAGE TypeApplications #-}
+{-# LANGUAGE TypeFamilies #-}
+{-# LANGUAGE TypeOperators #-}
+{-# LANGUAGE UndecidableInstances #-}
+
+module PredMutRec_opt where
+
+import PredMutRec
+import Test.QuickCheck.HRep.TH
+import Test.QuickCheck.HRep.Infix
+import Test.QuickCheck.Branching
+
+optimize @ExpS @"IExp" ''ExpS "test" uniform 10

### 2019-01-15 trunk/Exp_test_fine.hs

@@ -57,3 +57,18 @@ type instance Rep "foo" = Rep "foo#1" .* 2
 --   alg (AddEq e) = Add e e
 
 -- type Exp_pat = ExpR + Pats + AddEq
+
+-- twice :: IExp -> IExp
+-- twice e = Add e e
+
+-- thrice :: IExp -> IExp
+-- thrice e = Add e (twice e)
+
+-- weird :: IExp -> IExp -> IExp -> IExp
+-- weird x y z = If (And (LEq x y) (LEq y z)) (Add x y) (Add y z)
+
+-- equals :: IExp -> IExp -> BExp
+-- equals x y = And (LEq x y) (LEq y x)
+
+-- and3 :: BExp -> BExp -> BExp -> BExp
+-- and3 x y z = And x (And y z)

### 2019-01-15 test/TestExp.hs
Sync

@@ -19,7 +19,7 @@
 {-# LANGUAGE TypeOperators #-}
 {-# LANGUAGE UndecidableInstances #-}
 
-module PredMutRec where
+module TestExp where
 
 import GHC.Generics
 

### 2019-01-16 paper/main.tex
Sync

@@
 
-In this work we identify two sources of complications that appear when using the
-aforementioned automatic derivation tools to obtain random data generators:
+In this work we identify two sources of additional structure information which
+are not considered by the aforementioned automatic derivation tools to obtain
+random data generators:
 
@@
   patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
+\item The target code yields some of the responsability of preserving its
+  invariants to the functions of its abstract interface.
 \end{enumerate}
 
-To exemplify the first point, suppose we want to use randomly generated \ensuremath{\Conid{Exp}}s
-to test a property comprising the following function:
+This information can be reified in compile time and used to synthesize richer
+random generators automatically. We proceed to exemplify the previous points in
+detail.
+
+\subsection{\textbf{Presence of Complex Pattern Matchings}}
+
+To exemplify the firts problematic scenarion that may arise, suppose we want to
+use randomly generated \ensuremath{\Conid{Exp}}s to test a property comprising the following
+function:
 
@@
 \\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
+\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\Varid{y})){}\<[E]%
 \\
@@
 
-This function behaves mostly like an identity function, except for two very
-specific patterns of input values defined using three nested levels of pattern
-matching.
+This function behaves very much like an identity function, with the exception
+that it fails for two very specific patterns of input values defined using
+nested pattern matching.
 %
@@
 
-If we consider that \quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
-%
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this considerations, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the error) are as follows:
+\quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick a number in the
+interval $[-n, n]$ with uniform distribution.
+%
+Hence, in principle we need to recognize what is a suitable generation size,
+which should be big enough for our generator be able to generate the numbers we
+pattern match against to, otherwise we will not be able to generate any value to
+match against \ensuremath{\Varid{foo}} clauses, leaving fragments of code completely untested.
+%
+Suppose then that we pick the a generation size $50$, i.e. the minimum size that
+is big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
+%
+Under this consideration, the probability of generating a value matching the
+first clause of \ensuremath{\Varid{foo}} (and hence triggering the first error) results as follows:
 
 \begin{align*}
-  P(match(foo_1))
+  P(match(foo\#1))
   \!&&=\; &P(Add) \\
@@
 
-\todo[inline]{AM: COMPLETE}
+If we use \megadeth to obtain a random generator for \ensuremath{\Conid{Exp}}, we obtain a uniform
+generation probability distribution over constructors, i.e., $P(\ensuremath{\Conid{Val}}) =
+P(\ensuremath{\Conid{Add}}) = P(\ensuremath{\Conid{Mul}}) = 1/3$.
+%
+In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
+average, we will need to generate over than two million test cases in order to
+be able to test the first clause of \ensuremath{\Varid{foo}} only once.
+%
+This situation can be somewhat improved if we decide to use \dragen to obtain a
+random generator.
+%
+By optimizing the generation probabilities in order to benefit the generation of
+\ensuremath{\Conid{Add}} constructor in a ratio of $20:1$ with respect to the rest of them, we
+obtained a distribution of values where $P(match(foo\#1)) \sim 1/300000$.
+%
+Although this certainly improves the probability of generating a matching value,
+this probability is not substantial enough to become practical.
+%
+Additionally, by favoring the generation probabilities towards the \ensuremath{\Conid{Add}}
+constructor, we found that the probability of generating a value matching the
+second clause of \ensuremath{\Varid{foo}} (which does not matches against it) also gets diminished
+into an impractival value.
 
@@
 For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left- and right subtrees and sub-subtrees
+quite specific combination of color, left and right subtrees and sub-subtrees
 in order to preserve the height invariant.
@@
 needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
-
-We believe that if a certain pattern of values is relevant enough to appear in
+in a computationally more efficient way, to cite a few.
+
+\subsection{\textbf{Data Invariants Encoded on Abstract Interfaces}}
+
+A common choice when implementing a data structure is to transfer the
+responsability of preserving its invariants to the functions that manipulates
+its values.
+%
+Consider for instance the following possible implementation of a basic \ensuremath{\Conid{HTML}}
+manipulation library:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\Conid{Html}\mathrel{=}\Conid{Html}\;\Conid{String}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{head}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\Varid{head}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <head>\char34}\plus \Varid{inner}\plus \text{\ttfamily \char34 </head>\char34}){}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{body}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\Varid{body}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <body>\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 </body>\char34}){}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{br}\mathbin{::}\Conid{Html}{}\<[E]%
+\\
+\>[B]{}\Varid{br}\mathrel{=}\Conid{Html}\;\text{\ttfamily \char34 <br~/>\char34}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{concat}\mathbin{::}\Conid{Html}\to \Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\Varid{concat}\;(\Conid{Html}\;\Varid{x})\;(\Conid{Html}\;\Varid{y})\mathrel{=}\Conid{Html}\;(\Varid{x}\plus \Varid{y}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+In the previous definition, the \ensuremath{\Conid{Html}} data type is defined using a single data
+constructor that contains the textual representation of the Html code it
+represents in plain text.
+%
+Later, the combinator functions defined over \ensuremath{\Conid{HTML}} are the ones in charge of
+transforming this plain text representation with the invariant that, given valid
+\ensuremath{\Conid{Html}} parameters, they always return a valid \ensuremath{\Conid{Html}} value.
+%
+This is a very commmon programming pattern that can be found in a variety of
+Haskell libraries \tocite{}.
+
+If we use a standard type-driven approach to derive a random generator for
+\ensuremath{\Conid{Html}}, the only ``structural'' information that we can use in the derivation
+process is that \ensuremath{\Conid{Html}} values are composed of strings.
+%
+As a consequence, we essentially end up generating random strings, which rarely
+represents a valid (or at least well structured) Html value.
+%
+Additionally, if our test suite contains properties constrained by certain
+preconditions, this lack of domain knowleadge may lead in an impractically high
+discard ratio of randomly generated test.
+%
+For instance, suposse we write a property to verify that \ensuremath{\Varid{head}} preserves the
+invariant previously mentioned:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{prop\char95 head\char95 inv}\mathbin{::}\Conid{Html}\to \Conid{Bool}{}\<[E]%
+\\
+\>[B]{}\Varid{prop\char95 head\char95 inv}\;\Varid{x}\mathrel{=}\Varid{valid}\;\Varid{x}\Longrightarrow\Varid{valid}\;(\Varid{head}\;\Varid{x}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+In the previous definition we state that \ensuremath{\Varid{head}} always returns a valid \ensuremath{\Conid{Html}}
+value (\ensuremath{\Varid{valid}\;(\Varid{head}\;\Varid{x})}), provided that we are given a valid \ensuremath{\Conid{Html}} as input
+(\ensuremath{\Varid{valid}\;\Varid{x}}).
+%
+Then, while testing this property we rarely satisfy its precondition, in which
+case \quickcheck discards the whole test, retrying with a diferent random input,
+degradating this way the testing performance.
+
+We believe that, if certain patterns of values are relevant enough to appear in
 the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{fortyTwo}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{fortyTwo}\mathrel{=}\Conid{Val}\;\mathrm{42}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{twice}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{twice}\;\Varid{x}\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{neg}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{neg}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1})){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Varid{neg}\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+to produce values satisfying such patterns in a substantial proportion.
+%
+The next section introduces a reprentation model that let us encode the
+structure information presented here into our automatically derived random
+generators in a modular and flexible way.
+
+% Consider the following Haskell module |M| which defines combinators for the type
+% |Exp|:
 
 % \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
-
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
-
-% \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
-
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
-% \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
+% module M where
+
+% ten :: Exp
+% ten = Val 10
+
+% square :: Exp -> Exp
+% square x = Mul x x
+
+% minus :: Exp -> Exp -> Exp
+% minus x y = Add x (Mul y (Val (-1)))
 % \end{code}
@@
 
-What to say here:
-
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
+In this work, we ...
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\text{HRep}_{Val}\;{}\<[16]%
+\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Val}\;\Conid{Int}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{Add}\;{}\<[16]%
+\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Add}\;\Varid{a}\;\Varid{a}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{Mul}\;{}\<[16]%
+\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Mul}\;\Varid{a}\;\Varid{a}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+For this purpose, we employ Haskell's type classes mechanism \tocite{}.
+%
+We define an evaluation type class $(\Downarrow)$ which specifies how to perform
+a single transformation \ensuremath{\Varid{step}} from the higher-order representation back into a
+concrete value of the target:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{class}\;(\Varid{rep}\;\Downarrow\;\Varid{target})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\mathbin{::}\Varid{rep}\;\Varid{target}\to \Varid{target}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Then, we can define the following overloaded instances of the \ensuremath{\Varid{step}} operation
+for the canonical representations of data constructors simply by mapping each
+lifted constructor back into its corresponding one.
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{Val}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{Val}\;\Varid{n})\mathrel{=}\Conid{Val}\;\Varid{n}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{Add}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{Add}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{y}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{Mul}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{Mul}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{y}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+With these individual representations, we can define a type combinator
+$(\oplus)$ to compose two representations into a single one:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(\Varid{f}\;\oplus\;\Varid{g})\;\Varid{a}\mathrel{=}\Conid{L}\;(\Varid{f}\;\Varid{a})\mid \Conid{R}\;(\Varid{g}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Then, a composite representation can be transformed back into the concrete
+target type by pattern matching on the data type variant and applying the \ensuremath{\Varid{step}}
+tranformation to the inner representation:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{instance}\;(\Varid{f}\;\oplus\;\Varid{g}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\Conid{L}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\Conid{R}\;\Varid{g})\mathrel{=}\Varid{step}\;\Varid{g}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Furthermore, we can define a type combinator ($\otimes$) to tag every
+representation with an explicit generation frequency:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+This combinator is evaluated simply by piping the result from the inner
+representation.
+%
+It does not change the evaluation semantics, as it is only considered at
+generation time:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{instance}\;(\Varid{f}\;\otimes\;\Varid{n}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\Conid{Freq}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{HRep}_{Exp}{}\<[16]%
+\>[16]{}\mathrel{=}\text{HRep}_{Val}\;\!\oplus\!\;\text{HRep}_{Add}\;\!\oplus\!\;\text{HRep}_{Mul}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{TagFreq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\Varid{f}\;\otimes\;\Varid{n}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\Conid{TagFreq}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#1}\;{}\<[18]%
+\>[18]{}\Varid{a}\mathrel{=}\text{Mk}_{foo\#1}\;\Varid{a}\;\Varid{a}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#2}\;{}\<[18]%
+\>[18]{}\Varid{a}\mathrel{=}\text{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{a}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{foo\#1}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{foo\#1}\;\Varid{x}\;\Varid{y}){}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathrel{=}\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y}){}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{foo\#2}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{foo\#2}\;\Varid{x}\;\Varid{y}){}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathrel{=}\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\Varid{y}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\text{HRep}_{ten}\;{}\<[21]%
+\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{square}\;{}\<[21]%
+\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{square}\;{}\<[37]%
+\>[37]{}\Varid{a}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{minus}\;{}\<[21]%
+\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{minus}\;{}\<[37]%
+\>[37]{}\Varid{a}\;\Varid{a}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{ten}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;\text{Mk}_{ten}\mathrel{=}\Varid{ten}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{square}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{square}\;\Varid{x})\mathrel{=}\Varid{square}\;\Varid{x}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\mathbf{instance}\;(\text{HRep}_{minus}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{step}\;(\text{Mk}_{minus}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{minus}\;\Varid{x}\;\Varid{y}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
+\>[16]{}\mathrel{=}{}\<[24]%
+\>[24]{}\text{HRep}_{foo\#1}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{foo\#2}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{HRep}_{M}{}\<[14]%
+\>[14]{}\mathrel{=}{}\<[22]%
+\>[22]{}\text{HRep}_{ten}{}\<[E]%
+\\
+\>[14]{}\,\oplus\;{}\<[22]%
+\>[22]{}\text{HRep}_{square}{}\<[E]%
+\\
+\>[14]{}\,\oplus\;{}\<[22]%
+\>[22]{}\text{HRep}_{minus}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{Spec}_{Exp}{}\<[16]%
+\>[16]{}\mathrel{=}{}\<[24]%
+\>[24]{}\text{HRep}_{Exp}\;{}\<[34]%
+\>[34]{}\otimes\;\mathrm{4}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{foo}\;{}\<[34]%
+\>[34]{}\otimes\;\mathrm{2}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{M}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+
+% \begin{code}
+% data (Term f) a = TagTerm (f a)
+
+% instance (Term f down Exp) where
+%   step (TagTerm f) = step f
+% \end{code}
 \section{Random Generators Synthesis} \label{sec:synthesis}
@@
 Nam vestibulum accumsan nisl.
-\section{Case Studies}
+% %include sections/casestudies.tex
+\section{Related Work}
 
@@
 Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+\section{Final Remarks}
 
@@
 est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
 

### 2019-01-16 paper/sections/hrep.tex
Sync

@@
 
-What to say here:
+In this work, we ...
 
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
+\begin{code}
+data HRep_Val  a = Mk_Val Int
+data HRep_Add  a = Mk_Add a a
+data HRep_Mul  a = Mk_Mul a a
+\end{code}
+
+For this purpose, we employ Haskell's type classes mechanism \tocite{}.
+%
+We define an evaluation type class $(\Downarrow)$ which specifies how to perform
+a single transformation |step| from the higher-order representation back into a
+concrete value of the target:
+
+\begin{code}
+class (rep down target) where
+  step :: rep target -> target
+\end{code}
+
+Then, we can define the following overloaded instances of the |step| operation
+for the canonical representations of data constructors simply by mapping each
+lifted constructor back into its corresponding one.
+
+\begin{code}
+instance (HRep_Val down Exp) where
+  step (Mk_Val n) = Val n
+
+instance (HRep_Add down Exp) where
+  step (Mk_Add x y) = Add x y
+
+instance (HRep_Mul down Exp) where
+  step (Mk_Mul x y) = Mul x y
+\end{code}
+
+With these individual representations, we can define a type combinator
+$(\oplus)$ to compose two representations into a single one:
+
+\begin{code}
+data (f oplus g) a = L (f a) | R (g a)
+\end{code}
+
+Then, a composite representation can be transformed back into the concrete
+target type by pattern matching on the data type variant and applying the |step|
+tranformation to the inner representation:
+
+\begin{code}
+instance (f oplus g down Exp) where
+  step (L f) = step f
+  step (R g) = step g
+\end{code}
+
+Furthermore, we can define a type combinator ($\otimes$) to tag every
+representation with an explicit generation frequency:
+
+\begin{code}
+data (f otimes n) a = Freq (f a)
+\end{code}
+
+This combinator is evaluated simply by piping the result from the inner
+representation.
+%
+It does not change the evaluation semantics, as it is only considered at
+generation time:
+
+\begin{code}
+instance (f otimes n down Exp) where
+  step (Freq f) = step f
+\end{code}
+
+\begin{code}
+type HRep_Exp  = HRep_Val oplus'' HRep_Add oplus'' HRep_Mul
+\end{code}
+
+\begin{code}
+data (f otimes n) a = TagFreq (f a)
+
+instance (f otimes n down Exp) where
+  step (TagFreq f) = step f
+\end{code}
+
+\begin{code}
+data HRep_foo_1  a = Mk_foo_1 a a
+data HRep_foo_2  a = Mk_foo_2 Int a
+\end{code}
+
+\begin{code}
+instance (HRep_foo_1 down Exp) where
+  step (Mk_foo_1 x y)
+    = Add (Add x (Val 50)) (Add (Val 25) y)
+
+instance (HRep_foo_2 down Exp) where
+  step (Mk_foo_2 x y)
+    = Mul (Val 50) (Mul (Val x) y)
+\end{code}
+
+\begin{code}
+data HRep_ten       a = Mk_ten
+data HRep_square    a = Mk_square   a
+data HRep_minus     a = Mk_minus    a a
+\end{code}
+
+\begin{code}
+instance (HRep_ten down Exp) where
+  step Mk_ten = ten
+
+instance (HRep_square down Exp) where
+  step (Mk_square x) = square x
+
+instance (HRep_minus down Exp) where
+  step (Mk_minus x y) = minus x y
+\end{code}
+
+\begin{code}
+type HRep_foo  =       HRep_foo_1
+               oplus'  HRep_foo_2
+\end{code}
+
+\begin{code}
+type HRep_M  =       HRep_ten
+             oplus'  HRep_square
+             oplus'  HRep_minus
+\end{code}
+
+\begin{code}
+type Spec_Exp  =       HRep_Exp  otimes 4
+               oplus'  HRep_foo  otimes 2
+               oplus'  HRep_M
+\end{code}
 
@@
 est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+
+% \begin{code}
+% data (Term f) a = TagTerm (f a)
+
+% instance (Term f down Exp) where
+%   step (TagTerm f) = step f
+% \end{code}

### 2019-01-16 paper/sections/randomtesting.tex
Sync

@@
   patterns.
-\item The target code behaves differently according to some dependency over
-  subexpressions of its inputs.
+\item The target code yields some of the responsability of preserving its
+  invariants to the functions of its abstract interface.
 \end{enumerate}
 
-To exemplify the first point, suppose we want to use randomly generated |Exp|s
-to test a property comprising the following function:
+This information can be reified in compile time and used to synthesize richer
+random generators automatically. We proceed to exemplify the previous points in
+detail.
+
+\subsection{\textbf{Presence of Complex Pattern Matchings}}
+
+To exemplify the firts problematic scenarion that may arise, suppose we want to
+use randomly generated |Exp|s to test a property comprising the following
+function:
 
@@
   = error "pattern #1"
-foo (Mul (Mul x (Val 50)) (Mul (Val 25) y))
+foo (Mul (Val 50) (Mul (Val x) y))
   = error "pattern #2"
@@
 
-If we consider that \quickcheck's default implementation for random |Int|s pick
-a number in the interval $[-n, n]$ with uniform distribution, we need to be able
-to identify which one is a sensible generation size to set based on wether our
-code branches on numbers or not, otherwise our random generator will not be able
-to generate random values to match against such numbers, leaving fragments of
-code completely untested.
+\quickcheck's default implementation for random |Int|s pick a number in the
+interval $[-n, n]$ with uniform distribution.
 %
-Suppose that we pick the a generation size $50$, i.e. the minimum size that is
-big enough to produce an |Int| number equal to $50$.
+Hence, in principle we need to recognize what is a suitable generation size,
+which should be big enough for our generator be able to generate the numbers we
+pattern match against to, otherwise we will not be able to generate any value to
+match against |foo| clauses, leaving fragments of code completely untested.
 %
-Under this considerations, the probability of generating a value matching the
-first clause of |foo| (and hence triggering the error) are as follows:
+Suppose then that we pick the a generation size $50$, i.e. the minimum size that
+is big enough to produce an |Int| number equal to $50$.
+%
+Under this consideration, the probability of generating a value matching the
+first clause of |foo| (and hence triggering the first error) results as follows:
 
 \begin{align*}
-  P(match(foo_1))
+  P(match(foo\#1))
   \!&&=\; &P(Add) \\
@@
 
-\todo[inline]{AM: COMPLETE}
+If we use \megadeth to obtain a random generator for |Exp|, we obtain a uniform
+generation probability distribution over constructors, i.e., $P(|Val|) =
+P(|Add|) = P(|Mul|) = 1/3$.
+%
+In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
+average, we will need to generate over than two million test cases in order to
+be able to test the first clause of |foo| only once.
+%
+This situation can be somewhat improved if we decide to use \dragen to obtain a
+random generator.
+%
+By optimizing the generation probabilities in order to benefit the generation of
+|Add| constructor in a ratio of $20:1$ with respect to the rest of them, we
+obtained a distribution of values where $P(match(foo\#1)) \sim 1/300000$.
+%
+Although this certainly improves the probability of generating a matching value,
+this probability is not substantial enough to become practical.
+%
+Additionally, by favoring the generation probabilities towards the |Add|
+constructor, we found that the probability of generating a value matching the
+second clause of |foo| (which does not matches against it) also gets diminished
+into an impractival value.
 
@@
 needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way.
+in a computationally more efficient way, to cite a few.
 
-We believe that if a certain pattern of values is relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such pattern in a substantial proportion.
+\subsection{\textbf{Data Invariants Encoded on Abstract Interfaces}}
+
+A common choice when implementing a data structure is to transfer the
+responsability of preserving its invariants to the functions that manipulates
+its values.
+%
+Consider for instance the following possible implementation of a basic |HTML|
+manipulation library:
 
 \begin{code}
-fortyTwo :: Exp
-fortyTwo = Val 42
+data Html = Html String
 
-twice :: Exp -> Exp
-twice x = Add x x
+head :: Html -> Html
+head (Html inner)
+  = Html ("<head>" ++ inner ++ "</head>")
 
-square :: Exp -> Exp
-square x = Mul x x
+body :: Html -> Html
+body (Html inner)
+  = Html ("<body>" ++ t ++ "</body>")
 
-neg :: Exp -> Exp
-neg x = Mul x (Val (-1))
+br :: Html
+br = Html "<br />"
 
-minus :: Exp -> Exp -> Exp
-minus x y = Add x (neg y)
+concat :: Html -> Html -> Html
+concat (Html x) (Html y) = Html (x ++ y)
 \end{code}
 
-% \begin{code}
-% data IExp  =   VAL Int
-%            pp  ADD IExp IExp
-%            pp  IF BExp IExp IExp
+In the previous definition, the |Html| data type is defined using a single data
+constructor that contains the textual representation of the Html code it
+represents in plain text.
+%
+Later, the combinator functions defined over |HTML| are the ones in charge of
+transforming this plain text representation with the invariant that, given valid
+|Html| parameters, they always return a valid |Html| value.
+%
+This is a very commmon programming pattern that can be found in a variety of
+Haskell libraries \tocite{}.
 
-% data BExp  =   TRUE | FALSE
-%            pp  LE IExp IExp
-%            pp  AND BExp BExp
-% \end{code}
+If we use a standard type-driven approach to derive a random generator for
+|Html|, the only ``structural'' information that we can use in the derivation
+process is that |Html| values are composed of strings.
+%
+As a consequence, we essentially end up generating random strings, which rarely
+represents a valid (or at least well structured) Html value.
+%
+Additionally, if our test suite contains properties constrained by certain
+preconditions, this lack of domain knowleadge may lead in an impractically high
+discard ratio of randomly generated test.
+%
+For instance, suposse we write a property to verify that |head| preserves the
+invariant previously mentioned:
+
+\begin{code}
+prop_head_inv :: Html -> Bool
+prop_head_inv x = valid x ===> valid (head x)
+\end{code}
+
+In the previous definition we state that |head| always returns a valid |Html|
+value (|valid (head x)|), provided that we are given a valid |Html| as input
+(|valid x|).
+%
+Then, while testing this property we rarely satisfy its precondition, in which
+case \quickcheck discards the whole test, retrying with a diferent random input,
+degradating this way the testing performance.
+
+We believe that, if certain patterns of values are relevant enough to appear in
+the codebase being tested, a practical random testing methodology should be able
+to produce values satisfying such patterns in a substantial proportion.
+%
+The next section introduces a reprentation model that let us encode the
+structure information presented here into our automatically derived random
+generators in a modular and flexible way.
+
+% Consider the following Haskell module |M| which defines combinators for the type
+% |Exp|:
 
 % \begin{code}
-% foo :: IExp -> IExp
-% foo (ADD (IF TRUE (VAL 42) x) y)     = ...
-% foo (IF (LE (VAL 24) x) (VAL y)) z)  = ...
-% foo x                                = ...
+% module M where
 
-% bar :: BExp -> BExp
-% bar (LE (ADD (VAL 12) x) (VAL 34))   = ...
-% bar (LE (VAL 78) (IF FALSE x y)) z)  = ...
-% bar x                                = ...
+% ten :: Exp
+% ten = Val 10
+
+% square :: Exp -> Exp
+% square x = Mul x x
+
+% minus :: Exp -> Exp -> Exp
+% minus x y = Add x (Mul y (Val (-1)))
 % \end{code}
-
-% \begin{code}
-% twice :: IExp -> IExp
-% twice e = ADD e e
-
-% equals :: IExp -> IExp -> BExp
-% equals x y = AND (LE x y) (LE y x)
-
-% and3 :: BExp -> BExp -> BExp -> BExp
-% and3 x y z = AND x (AND y z)
-% \end{code}

### 2019-01-16 paper/tikz/hrep.tex
Sync

@@
 
-%Up Arrow [id:dp3837429966878143] 
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (360,160) -- (370,150) -- (380,160) -- (380,160) -- (380,160) -- (360,160) -- (360,160) -- cycle ;
 %Up Arrow [id:dp0013771056350821986]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,160) -- (270,150) -- (280,160) -- (280,160) -- (280,160) -- (260,160) -- (260,160) -- cycle ;
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,149.5) -- (270,140) -- (280,149.5) -- (273.5,149.5) -- (273.5,170) -- (266.5,170) -- (266.5,149.5) -- cycle ;
 %Shape: Rectangle [id:dp7366907361635939]
@@
 \draw  [line width=1.5]  (130,170) -- (410,170) -- (410,200) -- (130,200) -- cycle ;
+%Up Arrow [id:dp4999372801455]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (360,149.5) -- (370,140) -- (380,149.5) -- (373.5,149.5) -- (373.5,170) -- (366.5,170) -- (366.5,149.5) -- cycle ;
 
 % Text Node
-\draw (270,185) node [scale=1.2]  {$T$};
+\draw (270,185) node [scale=1.2]  {$Exp$};
 % Text Node
-\draw (270,125) node [scale=1]  {$\text{Patterns}^{T}_{foo}$};
+\draw (270,125) node [scale=1]  {$\text{Patterns}_{foo}$};
 % Text Node
-\draw (370,125) node [scale=1]  {$\text{Interface}^{T}_{M}$};
+\draw (370,125) node [scale=1]  {$\text{Interface}_{M}$};
 % Text Node
-\draw (170.05,65) node [scale=0.9]  {$\text{HRep}^{T}_{T}$};
+\draw (170.05,65) node [scale=1]  {$\text{HRep}_{Exp}$};
 % Text Node
-\draw (370,65) node [scale=0.9]  {$\text{HRep}^{T}_{M}$};
+\draw (370,65) node [scale=1]  {$\text{HRep}_{M}$};
+% % Text Node
+% \draw (77.5,125) node [scale=1.2]  {$Spec_{Exp}\Downarrow \ Exp$};
 % Text Node
-\draw (114.5,125) node [scale=1.2]  {$S_{T}\Downarrow T$};
+\draw (377.5,25) node [scale=1.2]  {$Spec_{Exp}$};
 % Text Node
-\draw (392,25) node [scale=1.2]  {$S_{T}$};
+\draw (270.22,65) node [scale=1]  {$\text{HRep}_{foo}$};
 % Text Node
-\draw (270.22,65) node [scale=0.9]  {$\text{HRep}^{T}_{foo}$};
+\draw (181.5,34) node [scale=1.44]  {$4$};
 % Text Node
-\draw (183.05,35) node [scale=1.44]  {$\alpha $};
-% Text Node
-\draw (284.05,35) node [scale=1.44]  {$\beta $};
+\draw (284.05,34) node [scale=1.44]  {$2$};
 

### 2019-01-16 paper/main.tex
WIP

@@
   patterns.
-\item The target code yields some of the responsability of preserving its
+\item The target code yields part of the responsability of preserving its
   invariants to the functions of its abstract interface.
@@
 
-If we use \megadeth to obtain a random generator for \ensuremath{\Conid{Exp}}, we obtain a uniform
-generation probability distribution over constructors, i.e., $P(\ensuremath{\Conid{Val}}) =
-P(\ensuremath{\Conid{Add}}) = P(\ensuremath{\Conid{Mul}}) = 1/3$.
+If we use \megadeth to automatically derive a random generator for \ensuremath{\Conid{Exp}}, we
+obtain a uniform generation probability distribution over constructors, i.e.,
+$P(\ensuremath{\Conid{Val}}) = P(\ensuremath{\Conid{Add}}) = P(\ensuremath{\Conid{Mul}}) = 1/3$.
 %
@@
 \ensuremath{\Conid{Add}} constructor in a ratio of $20:1$ with respect to the rest of them, we
-obtained a distribution of values where $P(match(foo\#1)) \sim 1/300000$.
+obtained a distribution of values where $P(match(foo\#1)) \approx 1/300000$.
+
 %
@@
 
-This combinator is evaluated simply by piping the result from the inner
-representation.
+This combinator is evaluated back to our target data type simply by piping the
+result from the inner representation.
 %
@@
 
+With the introduced combinators, we can easily create a type synonym \ensuremath{\text{HRep}_{Exp}}
+to refer to the canonical representation of our original data type \ensuremath{\Conid{Exp}},
+tagging for instance the representation of the constuctor \ensuremath{\Conid{Add}} to be generated
+in double the proportion of rest of the representations:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
 \>[B]{}\mathbf{type}\;\text{HRep}_{Exp}{}\<[16]%
-\>[16]{}\mathrel{=}\text{HRep}_{Val}\;\!\oplus\!\;\text{HRep}_{Add}\;\!\oplus\!\;\text{HRep}_{Mul}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{TagFreq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\Varid{f}\;\otimes\;\Varid{n}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\Conid{TagFreq}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
+\>[16]{}\mathrel{=}{}\<[24]%
+\>[24]{}\text{HRep}_{Val}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{Add}\;{}\<[34]%
+\>[34]{}\otimes\;\mathrm{2}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{Mul}{}\<[E]%
 \ColumnHook
@@
 \column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
+\>[16]{}\mathrel{=}{}\<[24]%
+\>[24]{}\text{HRep}_{foo\#1}{}\<[E]%
+\\
+\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{foo\#2}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{21}{@{}>{\hspre}l<{\hspost}@{}}%
@@
 \>[3]{}\Varid{step}\;(\text{Mk}_{minus}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{minus}\;\Varid{x}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{foo\#1}{}\<[E]%
-\\
-\>[16]{}\,\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{foo\#2}{}\<[E]%
 \ColumnHook
@@
 
-\begin{figure}[b]
+This previous definition can be interpreted graphically as it is shown in the
+Figure \ref{fig:hrep}.
+%
+Curly arrows represent the structural information extracted using
+meta-programming.
+
+\begin{figure}[t]
   \centering
   \input{tikz/hrep.tex}
-  \caption{The idea}
+  \caption{Higher order representation of the data type \ensuremath{\Conid{Exp}}, using structural
+    information from the function \ensuremath{\Varid{foo}} and the abstract interface of the module
+    \ensuremath{\Conid{M}}.}
+  \label{fig:hrep}
 \end{figure}
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+% \begin{code}
+% data (Term f) a = TagTerm (f a)
+
+% instance (Term f down Exp) where
+%   step (TagTerm f) = step f
+% \end{code}
+\section{Random Generators Synthesis} \label{sec:synthesis}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
 
@@
 
-% \begin{code}
-% data (Term f) a = TagTerm (f a)
-
-% instance (Term f down Exp) where
-%   step (TagTerm f) = step f
-% \end{code}
-\section{Random Generators Synthesis} \label{sec:synthesis}
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
+tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
+eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
+parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
+Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
+Nam vestibulum accumsan nisl.
+% %include sections/casestudies.tex
+\section{Related Work}
 
@@
 Nam euismod tellus id erat.
+\section{Final Remarks}
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-% %include sections/casestudies.tex
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-01-16 paper/sections/hrep.tex
WIP

@@
 
-\begin{code}
-type HRep_Exp  = HRep_Val oplus'' HRep_Add oplus'' HRep_Mul
-\end{code}
+With the introduced combinators, we can easily create a type synonym |HRep_Exp|
+to refer to the canonical representation of our original data type |Exp|,
+tagging for instance the representation of the constuctor |Add| to be generated
+in double the proportion of rest of the representations:
 
 \begin{code}
-data (f otimes n) a = TagFreq (f a)
-
-instance (f otimes n down Exp) where
-  step (TagFreq f) = step f
+type HRep_Exp  =       HRep_Val
+               oplus'  HRep_Add  otimes 2
+               oplus'  HRep_Mul
 \end{code}
@@
     = Mul (Val 50) (Mul (Val x) y)
+\end{code}
+
+\begin{code}
+type HRep_foo  =       HRep_foo_1
+               oplus'  HRep_foo_2
 \end{code}
@@
 \begin{code}
-type HRep_foo  =       HRep_foo_1
-               oplus'  HRep_foo_2
-\end{code}
-
-\begin{code}
 type HRep_M  =       HRep_ten
@@
 
-\begin{figure}[b]
+This previous definition can be interpreted graphically as it is shown in the
+Figure \ref{fig:hrep}.
+%
+Curly arrows represent the structural information extracted using
+meta-programming.
+
+\begin{figure}[t]
   \centering
   \input{tikz/hrep.tex}
-  \caption{The idea}
+  \caption{Higher order representation of the data type |Exp|, using structural
+    information from the function |foo| and the abstract interface of the module
+    |M|.}
+  \label{fig:hrep}
 \end{figure}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
 

### 2019-01-16 paper/sections/randomtesting.tex
WIP

@@
 
-If we use \megadeth to obtain a random generator for |Exp|, we obtain a uniform
-generation probability distribution over constructors, i.e., $P(|Val|) =
-P(|Add|) = P(|Mul|) = 1/3$.
+If we use \megadeth to automatically derive a random generator for |Exp|, we
+obtain a uniform generation probability distribution over constructors, i.e.,
+$P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
 %
@@
 |Add| constructor in a ratio of $20:1$ with respect to the rest of them, we
-obtained a distribution of values where $P(match(foo\#1)) \sim 1/300000$.
+obtained a distribution of values where $P(match(foo\#1)) \approx 1/300000$.
+
 %

### 2019-01-16 paper/main.tex
Unfolded figure

@@
 %
-By optimizing the generation probabilities in order to benefit the generation of
-\ensuremath{\Conid{Add}} constructor in a ratio of $20:1$ with respect to the rest of them, we
-obtained a distribution of values where $P(match(foo\#1)) \approx 1/300000$.
-
+Using this tool, we can optimize the generation probabilities in order to
+benefit the generation of some data constructors over the rest.
+%
+Considering that the first pattern matching of \ensuremath{\Varid{foo}} involves the data
+constructor \ensuremath{\Conid{Add}} as the only recursive one, we can set an target generation
+proportion of \ensuremath{\Conid{Add}} data constructors of, for instance, $20:1$ with respect to
+the rest of the generated data constructors.
+%
+By doing so, the obtained distribution of values results such that
+$P(match(foo\#1)) \approx 1/300000$.
 %
@@
 this probability is not substantial enough to become practical.
-%
+
 Additionally, by favoring the generation probabilities towards the \ensuremath{\Conid{Add}}
@@
 
+\subsection{\textbf{Representing Pattern Matchings}}
+
 \begin{hscode}\SaveRestoreHook
@@
 \end{hscode}\resethooks
+
+\subsection{\textbf{Representing Abstract Interfaces}}
 

### 2019-01-16 paper/sections/randomtesting.tex
Unfolded figure

@@
 %
-By optimizing the generation probabilities in order to benefit the generation of
-|Add| constructor in a ratio of $20:1$ with respect to the rest of them, we
-obtained a distribution of values where $P(match(foo\#1)) \approx 1/300000$.
-
+Using this tool, we can optimize the generation probabilities in order to
+benefit the generation of some data constructors over the rest.
+%
+Considering that the first pattern matching of |foo| involves the data
+constructor |Add| as the only recursive one, we can set an target generation
+proportion of |Add| data constructors of, for instance, $20:1$ with respect to
+the rest of the generated data constructors.
+%
+By doing so, the obtained distribution of values results such that
+$P(match(foo\#1)) \approx 1/300000$.
 %
@@
 this probability is not substantial enough to become practical.
-%
+
 Additionally, by favoring the generation probabilities towards the |Add|

### 2019-01-16 paper/tikz/hrep.tex
Unfolded figure

@@
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,416.5); %set diagram left start at 0, and has height of 416.5
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.9,xscale=0.9]
+%uncomment if require: \path (0,367.99999618530273); %set diagram left start at 0, and has height of 367.99999618530273
 
 %Shape: Rectangle [id:dp9489660675592899]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (130,20) ..
-controls (130,14.48) and (134.48,10) ..
-(140,10) -- (400,10) ..
-controls (405.52,10) and (410,14.48) ..
-(410,20) -- (410,80) ..
-controls (410,85.52) and (405.52,90) ..
-(400,90) -- (140,90) ..
-controls (134.48,90) and (130,85.52) ..
-(130,80) -- cycle ;
-%Snip Same Side Corner Rect [id:dp09931070350524518]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,30.5) -- (150.6,20) -- (189.5,20) -- (200,30.5) -- (200,50) -- (200,50) -- (140.1,50) -- (140.1,50) -- cycle ;
-%Down Arrow [id:dp5984296172784]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (140,159.5) -- (144,159.5) -- (144,90) -- (152,90) -- (152,159.5) -- (156,159.5) -- (148,170) -- cycle ;
-%Straight Lines [id:da30879479079788386]
-\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]    (145,90) -- (151,90) ;
-
-%Straight Lines [id:da5824111707250128]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (170,170) ..
-controls (168.33,168.33) and (168.33,166.67) ..
-(170,165) ..
-controls (171.67,163.33) and (171.67,161.67) ..
-(170,160) ..
-controls (168.33,158.33) and (168.33,156.67) ..
-(170,155) ..
-controls (171.67,153.33) and (171.67,151.67) ..
-(170,150) ..
-controls (168.33,148.33) and (168.33,146.67) ..
-(170,145) ..
-controls (171.67,143.33) and (171.67,141.67) ..
-(170,140) ..
-controls (168.33,138.33) and (168.33,136.67) ..
-(170,135) ..
-controls (171.67,133.33) and (171.67,131.67) ..
-(170,130) ..
-controls (168.33,128.33) and (168.33,126.67) ..
-(170,125) ..
-controls (171.67,123.33) and (171.67,121.67) ..
-(170,120) ..
-controls (168.33,118.33) and (168.33,116.67) ..
-(170,115) ..
-controls (171.67,113.33) and (171.67,111.67) ..
-(170,110) ..
-controls (168.33,108.33) and (168.33,106.67) ..
-(170,105) ..
-controls (171.67,103.33) and (171.67,101.67) ..
-(170,100) ..
-controls (168.33,98.33) and (168.33,96.67) ..
-(170,95) -- (170,91) -- (170,83) ;
-\draw [shift={(170,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
-(0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
-
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (110,20) ..
+controls (110,14.48) and (114.48,10) ..
+(120,10) -- (420,10) ..
+controls (425.52,10) and (430,14.48) ..
+(430,20) -- (430,220) ..
+controls (430,225.52) and (425.52,230) ..
+(420,230) -- (120,230) ..
+controls (114.48,230) and (110,225.52) ..
+(110,220) -- cycle ;
 %Straight Lines [id:da7051587981224026]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (270,110) ..
-controls (268.33,108.33) and (268.33,106.67) ..
-(270,105) ..
-controls (271.67,103.33) and (271.67,101.67) ..
-(270,100) ..
-controls (268.33,98.33) and (268.33,96.67) ..
-(270,95) -- (270,91) -- (270,83) ;
-\draw [shift={(270,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (270,250) ..
+controls (268.33,248.33) and (268.33,246.67) ..
+(270,245) ..
+controls (271.67,243.33) and (271.67,241.67) ..
+(270,240) ..
+controls (268.33,238.33) and (268.33,236.67) ..
+(270,235) ..
+controls (271.67,233.33) and (271.67,231.67) ..
+(270,230) ..
+controls (268.33,228.33) and (268.33,226.67) ..
+(270,225) ..
+controls (271.67,223.33) and (271.67,221.67) ..
+(270,220) ..
+controls (268.33,218.33) and (268.33,216.67) ..
+(270,215) ..
+controls (271.67,213.33) and (271.67,211.67) ..
+(270,210) ..
+controls (268.33,208.33) and (268.33,206.67) ..
+(270,205) -- (270,201) -- (270,193) ;
+\draw [shift={(270,190)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
 controls (9.04,-1.82) and (4.3,-0.39) ..
@@
 %Straight Lines [id:da6087008244886778]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (370,110) ..
-controls (368.33,108.33) and (368.33,106.67) ..
-(370,105) ..
-controls (371.67,103.33) and (371.67,101.67) ..
-(370,100) ..
-controls (368.33,98.33) and (368.33,96.67) ..
-(370,95) -- (370,91) -- (370,83) ;
-\draw [shift={(370,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (380,250) ..
+controls (378.33,248.33) and (378.33,246.67) ..
+(380,245) ..
+controls (381.67,243.33) and (381.67,241.67) ..
+(380,240) ..
+controls (378.33,238.33) and (378.33,236.67) ..
+(380,235) ..
+controls (381.67,233.33) and (381.67,231.67) ..
+(380,230) ..
+controls (378.33,228.33) and (378.33,226.67) ..
+(380,225) ..
+controls (381.67,223.33) and (381.67,221.67) ..
+(380,220) ..
+controls (378.33,218.33) and (378.33,216.67) ..
+(380,215) -- (380,211) -- (380,203) ;
+\draw [shift={(380,200)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
 controls (9.04,-1.82) and (4.3,-0.39) ..
@@
 %Shape: Rectangle [id:dp701556783783656]
-\draw  [line width=1.5]  (230,115) ..
-controls (230,112.24) and (232.24,110) ..
-(235,110) -- (305,110) ..
-controls (307.76,110) and (310,112.24) ..
-(310,115) -- (310,135) ..
-controls (310,137.76) and (307.76,140) ..
-(305,140) -- (235,140) ..
-controls (232.24,140) and (230,137.76) ..
-(230,135) -- cycle ;
+\draw  [line width=1.5]  (230,255) ..
+controls (230,252.24) and (232.24,250) ..
+(235,250) -- (305,250) ..
+controls (307.76,250) and (310,252.24) ..
+(310,255) -- (310,275) ..
+controls (310,277.76) and (307.76,280) ..
+(305,280) -- (235,280) ..
+controls (232.24,280) and (230,277.76) ..
+(230,275) -- cycle ;
 
 %Shape: Rectangle [id:dp033430127809365384]
-\draw  [line width=1.5]  (330,115) ..
-controls (330,112.24) and (332.24,110) ..
-(335,110) -- (405,110) ..
-controls (407.76,110) and (410,112.24) ..
-(410,115) -- (410,135) ..
-controls (410,137.76) and (407.76,140) ..
-(405,140) -- (335,140) ..
-controls (332.24,140) and (330,137.76) ..
-(330,135) -- cycle ;
+\draw  [line width=1.5]  (340,255) ..
+controls (340,252.24) and (342.24,250) ..
+(345,250) -- (415,250) ..
+controls (417.76,250) and (420,252.24) ..
+(420,255) -- (420,275) ..
+controls (420,277.76) and (417.76,280) ..
+(415,280) -- (345,280) ..
+controls (342.24,280) and (340,277.76) ..
+(340,275) -- cycle ;
+%Up Arrow [id:dp0013771056350821986]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,289.5) -- (270,280) -- (280,289.5) -- (273.5,289.5) -- (273.5,310) -- (266.5,310) -- (266.5,289.5) -- cycle ;
+%Shape: Rectangle [id:dp1692119124811613]
+\draw  [line width=1.5]  (110,310) -- (430,310) -- (430,340) -- (110,340) -- cycle ;
+%Up Arrow [id:dp4999372801455]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (370,289.5) -- (380,280) -- (390,289.5) -- (383.5,289.5) -- (383.5,310) -- (376.5,310) -- (376.5,289.5) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (223,132.93) ..
+controls (223,129.02) and (219.83,125.86) ..
+(215.93,125.86) ..
+controls (212.02,125.86) and (208.86,129.02) ..
+(208.86,132.93) ..
+controls (208.86,136.83) and (212.02,140) ..
+(215.93,140) ..
+controls (219.83,140) and (223,136.83) ..
+(223,132.93) -- cycle ; \draw  [line width=1.5]  (223,132.93) -- (208.86,132.93) ; \draw  [line width=1.5]  (215.93,125.86) -- (215.93,140) ;
+%Snip Same Side Corner Rect [id:dp11914646356873848]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (230,67.5) -- (237,60.5) -- (303,60.5) -- (310,67.5) -- (310,80.5) -- (310,80.5) -- (230,80.5) -- (230,80.5) -- cycle ;
+%Shape: Rectangle [id:dp6865365415345861]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (230,80.5) -- (310,80.5) -- (310,190.5) -- (230,190.5) -- cycle ;
+%Shape: Rectangle [id:dp39543698275133865]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (236.29,107.66) -- (303.21,107.66) -- (303.21,130.34) -- (236.29,130.34) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (247,75.5) ..
+controls (249.76,72.74) and (249.76,68.26) ..
+(247,65.5) ..
+controls (244.24,62.74) and (239.76,62.74) ..
+(237,65.5) ..
+controls (234.24,68.26) and (234.24,72.74) ..
+(237,75.5) ..
+controls (239.76,78.26) and (244.24,78.26) ..
+(247,75.5) -- cycle ; \draw  [line width=1.5]  (247,75.5) -- (237,65.5) ; \draw  [line width=1.5]  (247,65.5) -- (237,75.5) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (277.14,143.43) ..
+controls (277.14,139.52) and (273.98,136.36) ..
+(270.07,136.36) ..
+controls (266.17,136.36) and (263,139.52) ..
+(263,143.43) ..
+controls (263,147.33) and (266.17,150.5) ..
+(270.07,150.5) ..
+controls (273.98,150.5) and (277.14,147.33) ..
+(277.14,143.43) -- cycle ; \draw  [line width=1.5]  (277.14,143.43) -- (263,143.43) ; \draw  [line width=1.5]  (270.07,136.36) -- (270.07,150.5) ;
+%Shape: Rectangle [id:dp047421885346712545]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (236.08,157.66) -- (303,157.66) -- (303,180.34) -- (236.08,180.34) -- cycle ;
+%Shape: Rectangle [id:dp0425475572348164]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (340,40) -- (420,40) -- (420,200) -- (340,200) -- cycle ;
+%Shape: Rectangle [id:dp9247208762532781]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (345,66.37) -- (415,66.37) -- (415,89) -- (345,89) -- cycle ;
+%Shape: Rectangle [id:dp8712965895636344]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (345.52,117.37) -- (414.48,117.37) -- (414.48,140) -- (345.52,140) -- cycle ;
+%Shape: Rectangle [id:dp933766425327347]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (346.66,168.24) -- (413.34,168.24) -- (413.34,190.87) -- (346.66,190.87) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (387.14,103.95) ..
+controls (387.14,100.05) and (383.98,96.89) ..
+(380.07,96.89) ..
+controls (376.17,96.89) and (373,100.05) ..
+(373,103.95) ..
+controls (373,107.84) and (376.17,111) ..
+(380.07,111) ..
+controls (383.98,111) and (387.14,107.84) ..
+(387.14,103.95) -- cycle ; \draw  [line width=1.5]  (387.14,103.95) -- (373,103.95) ; \draw  [line width=1.5]  (380.07,96.89) -- (380.07,111) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (387.14,152.95) ..
+controls (387.14,149.05) and (383.98,145.89) ..
+(380.07,145.89) ..
+controls (376.17,145.89) and (373,149.05) ..
+(373,152.95) ..
+controls (373,156.84) and (376.17,160) ..
+(380.07,160) ..
+controls (383.98,160) and (387.14,156.84) ..
+(387.14,152.95) -- cycle ; \draw  [line width=1.5]  (387.14,152.95) -- (373,152.95) ; \draw  [line width=1.5]  (380.07,145.89) -- (380.07,160) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (333,132.93) ..
+controls (333,129.02) and (329.83,125.86) ..
+(325.93,125.86) ..
+controls (322.02,125.86) and (318.86,129.02) ..
+(318.86,132.93) ..
+controls (318.86,136.83) and (322.02,140) ..
+(325.93,140) ..
+controls (329.83,140) and (333,136.83) ..
+(333,132.93) -- cycle ; \draw  [line width=1.5]  (333,132.93) -- (318.86,132.93) ; \draw  [line width=1.5]  (325.93,125.86) -- (325.93,140) ;
+%Snip Same Side Corner Rect [id:dp09931070350524518]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (120,27.48) -- (126.98,20.5) -- (193.02,20.5) -- (200,27.48) -- (200,40.45) -- (200,40.45) -- (120,40.45) -- (120,40.45) -- cycle ;
 %Shape: Rectangle [id:dp6050538306488522]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,50) -- (200,50) -- (200,80) -- (140.1,80) -- cycle ;
-%Shape: Rectangle [id:dp9303311048687424]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (340,50) -- (400,50) -- (400,80) -- (340,80) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (120,40.45) -- (200,40.45) -- (200,220) -- (120,220) -- cycle ;
+%Shape: Rectangle [id:dp20494666512059623]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,67.54) -- (190,67.54) -- (190,90.17) -- (130,90.17) -- cycle ;
+%Shape: Rectangle [id:dp8741156804316592]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,137.36) -- (190,137.36) -- (190,159.99) -- (130,159.99) -- cycle ;
+%Shape: Rectangle [id:dp7972859165129229]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,187.24) -- (190,187.24) -- (190,209.87) -- (130,209.87) -- cycle ;
+%Snip Same Side Corner Rect [id:dp44067415974317603]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,124.4) -- (136.98,117.41) -- (183.02,117.41) -- (190,124.4) -- (190,137.36) -- (190,137.36) -- (130,137.36) -- (130,137.36) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (147,132.22) ..
+controls (149.76,129.46) and (149.76,125) ..
+(147,122.24) ..
+controls (144.24,119.49) and (139.76,119.49) ..
+(137,122.24) ..
+controls (134.24,125) and (134.24,129.46) ..
+(137,132.22) ..
+controls (139.76,134.97) and (144.24,134.97) ..
+(147,132.22) -- cycle ; \draw  [line width=1.5]  (147,132.22) -- (137,122.24) ; \draw  [line width=1.5]  (147,122.24) -- (137,132.22) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (137,35.46) ..
+controls (139.76,32.71) and (139.76,28.24) ..
+(137,25.49) ..
+controls (134.24,22.73) and (129.76,22.73) ..
+(127,25.49) ..
+controls (124.24,28.24) and (124.24,32.71) ..
+(127,35.46) ..
+controls (129.76,38.22) and (134.24,38.22) ..
+(137,35.46) -- cycle ; \draw  [line width=1.5]  (137,35.46) -- (127,25.49) ; \draw  [line width=1.5]  (137,25.49) -- (127,35.46) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (167.14,103.22) ..
+controls (167.14,99.33) and (163.98,96.17) ..
+(160.07,96.17) ..
+controls (156.17,96.17) and (153,99.33) ..
+(153,103.22) ..
+controls (153,107.12) and (156.17,110.27) ..
+(160.07,110.27) ..
+controls (163.98,110.27) and (167.14,107.12) ..
+(167.14,103.22) -- cycle ; \draw  [line width=1.5]  (167.14,103.22) -- (153,103.22) ; \draw  [line width=1.5]  (160.07,96.17) -- (160.07,110.27) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (167.14,173.05) ..
+controls (167.14,169.15) and (163.98,165.99) ..
+(160.07,165.99) ..
+controls (156.17,165.99) and (153,169.15) ..
+(153,173.05) ..
+controls (153,176.94) and (156.17,180.1) ..
+(160.07,180.1) ..
+controls (163.98,180.1) and (167.14,176.94) ..
+(167.14,173.05) -- cycle ; \draw  [line width=1.5]  (167.14,173.05) -- (153,173.05) ; \draw  [line width=1.5]  (160.07,165.99) -- (160.07,180.1) ;
+%Down Arrow [id:dp5984296172784]
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (160,299.5) -- (165,299.5) -- (165,230) -- (175,230) -- (175,299.5) -- (180,299.5) -- (170,310) -- cycle ;
+%Straight Lines [id:da30879479079788386]
+\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]    (166.25,230) -- (173.75,230) ;
 
-%Up Arrow [id:dp0013771056350821986]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,149.5) -- (270,140) -- (280,149.5) -- (273.5,149.5) -- (273.5,170) -- (266.5,170) -- (266.5,149.5) -- cycle ;
-%Shape: Rectangle [id:dp7366907361635939]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.43,50) -- (300,50) -- (300,80) -- (240.43,80) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (210,65) ..
-controls (210,59.48) and (214.48,55) ..
-(220,55) ..
-controls (225.52,55) and (230,59.48) ..
-(230,65) ..
-controls (230,70.52) and (225.52,75) ..
-(220,75) ..
-controls (214.48,75) and (210,70.52) ..
-(210,65) -- cycle ; \draw  [line width=1.5]  (210,65) -- (230,65) ; \draw  [line width=1.5]  (220,55) -- (220,75) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (310,65) ..
-controls (310,59.48) and (314.48,55) ..
-(320,55) ..
-controls (325.52,55) and (330,59.48) ..
-(330,65) ..
-controls (330,70.52) and (325.52,75) ..
-(320,75) ..
-controls (314.48,75) and (310,70.52) ..
-(310,65) -- cycle ; \draw  [line width=1.5]  (310,65) -- (330,65) ; \draw  [line width=1.5]  (320,55) -- (320,75) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (152.93,28.79) ..
-controls (156.83,24.88) and (163.17,24.88) ..
-(167.07,28.79) ..
-controls (170.98,32.69) and (170.98,39.02) ..
-(167.07,42.93) ..
-controls (163.17,46.83) and (156.83,46.83) ..
-(152.93,42.93) ..
-controls (149.02,39.02) and (149.02,32.69) ..
-(152.93,28.79) -- cycle ; \draw  [line width=1.5]  (152.93,28.79) -- (167.07,42.93) ; \draw  [line width=1.5]  (167.07,28.79) -- (152.93,42.93) ;
-%Snip Same Side Corner Rect [id:dp4221187902899477]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.1,30.5) -- (250.6,20) -- (289.5,20) -- (300,30.5) -- (300,50) -- (300,50) -- (240.1,50) -- (240.1,50) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (252.93,28.79) ..
-controls (256.83,24.88) and (263.17,24.88) ..
-(267.07,28.79) ..
-controls (270.98,32.69) and (270.98,39.02) ..
-(267.07,42.93) ..
-controls (263.17,46.83) and (256.83,46.83) ..
-(252.93,42.93) ..
-controls (249.02,39.02) and (249.02,32.69) ..
-(252.93,28.79) -- cycle ; \draw  [line width=1.5]  (252.93,28.79) -- (267.07,42.93) ; \draw  [line width=1.5]  (267.07,28.79) -- (252.93,42.93) ;
-%Shape: Rectangle [id:dp1692119124811613]
-\draw  [line width=1.5]  (130,170) -- (410,170) -- (410,200) -- (130,200) -- cycle ;
-%Up Arrow [id:dp4999372801455]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (360,149.5) -- (370,140) -- (380,149.5) -- (373.5,149.5) -- (373.5,170) -- (366.5,170) -- (366.5,149.5) -- cycle ;
+%Straight Lines [id:da8667104419281686]
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (190,310) ..
+controls (188.33,308.33) and (188.33,306.67) ..
+(190,305) ..
+controls (191.67,303.33) and (191.67,301.67) ..
+(190,300) ..
+controls (188.33,298.33) and (188.33,296.67) ..
+(190,295) ..
+controls (191.67,293.33) and (191.67,291.67) ..
+(190,290) ..
+controls (188.33,288.33) and (188.33,286.67) ..
+(190,285) ..
+controls (191.67,283.33) and (191.67,281.67) ..
+(190,280) ..
+controls (188.33,278.33) and (188.33,276.67) ..
+(190,275) ..
+controls (191.67,273.33) and (191.67,271.67) ..
+(190,270) ..
+controls (188.33,268.33) and (188.33,266.67) ..
+(190,265) ..
+controls (191.67,263.33) and (191.67,261.67) ..
+(190,260) ..
+controls (188.33,258.33) and (188.33,256.67) ..
+(190,255) ..
+controls (191.67,253.33) and (191.67,251.67) ..
+(190,250) ..
+controls (188.33,248.33) and (188.33,246.67) ..
+(190,245) ..
+controls (191.67,243.33) and (191.67,241.67) ..
+(190,240) ..
+controls (188.33,238.33) and (188.33,236.67) ..
+(190,235) -- (190,231) -- (190,223) ;
+\draw [shift={(190,220)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+controls (9.04,-1.82) and (4.3,-0.39) ..
+(0,0) ..
+controls (4.3,0.39) and (9.04,1.82) ..
+(14.21,4.28)   ;
 
 % Text Node
-\draw (270,185) node [scale=1.2]  {$Exp$};
+\draw (270,323) node [scale=1.2]  {$Exp$};
 % Text Node
-\draw (270,125) node [scale=1]  {$\text{Patterns}_{foo}$};
+\draw (270,265) node [scale=1]  {$\text{Patterns}_{foo}$};
 % Text Node
-\draw (370,125) node [scale=1]  {$\text{Interface}_{M}$};
+\draw (380,265) node [scale=1]  {$\text{Interface}_{M}$};
 % Text Node
-\draw (170.05,65) node [scale=1]  {$\text{HRep}_{Exp}$};
+\draw (148,51.5) node [scale=0.9]  {$\text{HRep}_{Exp}$};
 % Text Node
-\draw (370,65) node [scale=1]  {$\text{HRep}_{M}$};
-% % Text Node
-% \draw (77.5,125) node [scale=1.2]  {$Spec_{Exp}\Downarrow \ Exp$};
+\draw (273,27.5) node [scale=1]  {$Spec_{Exp}$};
 % Text Node
-\draw (377.5,25) node [scale=1.2]  {$Spec_{Exp}$};
+\draw (160,78.85) node [scale=0.9]  {$\text{HRep}_{Val}$};
 % Text Node
-\draw (270.22,65) node [scale=1]  {$\text{HRep}_{foo}$};
+\draw (160,148.68) node [scale=0.9]  {$\text{HRep}_{Add}$};
 % Text Node
-\draw (181.5,34) node [scale=1.44]  {$4$};
+\draw (160,198.55) node [scale=0.9]  {$\text{HRep}_{Mul}$};
 % Text Node
-\draw (284.05,34) node [scale=1.44]  {$2$};
+\draw (160,127.39) node [scale=1]  {$2$};
+% Text Node
+\draw (160,30.47) node [scale=1]  {$4$};
+% Text Node
+\draw (257,91.5) node [scale=0.9]  {$\text{HRep}_{foo}$};
+% Text Node
+\draw (270,70.5) node [scale=1]  {$2$};
+% Text Node
+\draw (269.75,119) node [scale=0.9]  {$\text{HRep}_{foo\#1}$};
+% Text Node
+\draw (269.54,169) node [scale=0.9]  {$\text{HRep}_{foo\#2}$};
+% Text Node
+\draw (115,269.5) node [scale=0.9]  {$Spec_{Exp}\! \Downarrow \! Exp$};
+% Text Node
+\draw (365.5,51.5) node [scale=0.9]  {$\text{HRep}_{M}$};
+% Text Node
+\draw (380,179.55) node [scale=0.9]  {$\text{HRep}_{minus}$};
+% Text Node
+\draw (380,128.69) node [scale=0.9]  {$\text{HRep}_{square}$};
+% Text Node
+\draw (380,77.69) node [scale=0.9]  {$\text{HRep}_{ten}$};
 

### 2019-01-16 paper/watch.sh

@@ -1,4 +1,4 @@
 #!/bin/sh
 
 #https://facebook.github.io/watchman/docs/watchman-make.html
-watchman-make --make 'make' -p '**/[^.]*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all
+watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' 'sections/[^.]*.tex' 'tikz/[^.]*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all

### 2019-01-17 paper/main.tex
WIP

@@
-\documentclass[conference]{IEEEtran}
+\documentclass[conference, fleqn]{IEEEtran}
 
@@
 \usepackage{xspace}
+\usepackage{relsize}
 \usetikzlibrary{fadings}
@@
 %% Macros
-\newcommand{\tocite}{\textbf{CITATION NEEDED}}
+\newcommand{\tocite}{\textbf{CITE}}
 \newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
@@
 \newcommand{\dragen}{\textbf{DRAGEN}\xspace}
+
+\newcommand{\evalrep}[2]{\ensuremath{\llbracket #1 \rrbracket_{#2}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Some parameters
+\setlength{\mathindent}{1.5\parindent}
 
@@
 %
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
+In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} (or
+constructors for short) and are used to distiguish which variant of the data
+type we are reffering to.
 %
@@
 \column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{7E}{@{}l@{}}%
-\column{10}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{genExp}\mathbin{::}\Conid{Gen}\;\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{genExp}\mathrel{=}\Varid{sized}\;\Varid{gen}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathbf{where}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\mathrm{0}\mathrel{=}\Conid{Val}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\Varid{gen}\;\Varid{n}\mathrel{=}\Varid{oneof}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu {}\<[7E]%
-\>[10]{}\Conid{Val}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{arbitrary}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Add}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},{}\<[7E]%
-\>[10]{}\Conid{Mul}{}\<[15]%
-\>[15]{}\mathop{\langle \$ \rangle}\Varid{gen}\;(n{-}1){}\<[30]%
-\>[30]{}\mathop{\langle \ast \rangle}\Varid{gen}\;(n{-}1)\mskip1.5mu]{}\<[E]%
+\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
+\column{6E}{@{}l@{}}%
+\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}gen_{Exp}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
+\\
+\>[B]{}\hsindent{4}{}\<[4]%
+\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
+\\
+\>[B]{}\hsindent{4}{}\<[4]%
+\>[4]{}\mathbf{then}\;\Conid{Val}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
+\\
+\>[B]{}\hsindent{4}{}\<[4]%
+\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{}[\mskip1.5mu {}\<[6E]%
+\>[9]{}(\mathrm{1},\Conid{Val}{}\<[18]%
+\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{},{}\<[6E]%
+\>[9]{}(\mathrm{2},\Conid{Add}{}\<[18]%
+\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp}){}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{},{}\<[6E]%
+\>[9]{}(\mathrm{1},\Conid{Mul}{}\<[18]%
+\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp})\mskip1.5mu]){}\<[E]%
 \ColumnHook
@@
 
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}\mathbin{::}(\Conid{Int}\to \Conid{Gen}\;\Varid{a})\to \Conid{Gen}\;\Varid{a}} to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
+This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
+paremetrize the generation process up to an external integer number known as the
+\emph{generation size}.
 %
@@
 %
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator is free
-to randomly pick with uniform probability among any data constructor of \ensuremath{\Conid{Exp}}.
+When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator will
+pick among any data constructor of \ensuremath{\Conid{Exp}} with a explicitly given frequency---two
+times more \ensuremath{\Conid{Add}}s than \ensuremath{\Conid{Val}} or \ensuremath{\Conid{Mul}}.
 %
@@
 data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (\ensuremath{\Varid{gen}\;(n{-}1)}), ensuring to decrease the generation size by a unit on each recursive
+corresponding to the data constructor fields by calling itself recursively
+(\ensuremath{\Varid{resize}\;(n{-}1)\;gen_{Exp}}), decreasing the generation size on each recursive
 invocation.
@@
 first clause of \ensuremath{\Varid{foo}} (and hence triggering the first error) results as follows:
-
+%
 \begin{align*}
-  P(match(foo\#1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
+  &P(match(foo\#1))\\
+  &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
+  &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
 \end{align*}
@@
 %
-This situation can be somewhat improved if we decide to use \dragen to obtain a
-random generator.
+This situation can be somewhat improved if we use \dragen to obtain a random
+generator.
 %
@@
 \\[\blanklineskip]%
-\>[B]{}\Varid{concat}\mathbin{::}\Conid{Html}\to \Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{concat}\;(\Conid{Html}\;\Varid{x})\;(\Conid{Html}\;\Varid{y})\mathrel{=}\Conid{Html}\;(\Varid{x}\plus \Varid{y}){}\<[E]%
+\>[B]{}(\mathop{\langle\! +\! \rangle})\mathbin{::}\Conid{Html}\to \Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}(\Conid{Html}\;\Varid{x})\mathop{\langle\! +\! \rangle}(\Conid{Html}\;\Varid{y})\mathrel{=}\Conid{Html}\;(\Varid{x}\plus \Varid{y}){}\<[E]%
 \ColumnHook
@@
 generators in a modular and flexible way.
-
-% Consider the following Haskell module |M| which defines combinators for the type
-% |Exp|:
+\section{Extracting Structure} \label{sec:hrep}
+
+In this section we present a compositional representation to express the random
+generation of values following the internal structure of their data types along
+with the structure present on patterns matchings and abstract interfaces.
+
+The key idea of this work is to represent different sources of structured value
+information of a data type in a homogeneous way that we call a ``higher-level
+representation'' (\ensuremath{\text{HRep}} from now on).%
+\footnote{The notion of higher-level comes from that the generation process of
+  given target data type is entirely determined by the type of the chosen
+  representation, instead of by a concrete generator defined at the term level.}
+%
+For this purpose, we use a series of automatically derived data types, each one
+representing an atomic unit of information that can be randomly generated and
+then reflected back to the corresponding value of the original data type.
+
+We wil reuse the previously defined data type \ensuremath{\Conid{Exp}} and the function \ensuremath{\Varid{foo}} to
+explain the different concepts involved all across this section.
+
+\subsection{\textbf{Representing Data Constructors}}
+
+We begin by introducing the simplest representation that we can extract from our
+codebase: the representation of single data constructor.
+%
+Each original constructor can be represented by a automatically derived data
+type consisting of a single constructor with the same fields, except for the
+recursive ones that are abstracted away.
+%
+In this light, we represent each constructor of the data type \ensuremath{\Conid{Exp}} as follows:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\text{HRep}_{Val}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Val}\;\Conid{Int}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{Add}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Add}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\text{HRep}_{Mul}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Mul}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Note that the previous definition are type parametric over the type parameter
+\ensuremath{\Varid{r}}.
+%
+This way we can replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
+possible values on each case.
+%
+For instance, the value (\ensuremath{\text{Mk}_{Add}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\text{HRep}_{Add}\;\Conid{Int}}, while the
+value (\ensuremath{\text{Mk}_{Mul}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\text{HRep}_{Mul}\;\Conid{Bool}}.
+%
+In practice, this parametricity let us instatiate \ensuremath{\Varid{a}} with the type of the
+chosen representation (which might change according to the user demands),
+without having to modify anything in the underlying machinery.
+
+For this purpose, we employ Haskell's type classes mechanism \tocite{}.
+%
+We define an evaluation type class $(\Downarrow)$ which specifies how to perform
+a single transformation \ensuremath{\Varid{step}} from an \ensuremath{\text{HRep}} back into a concrete value of the
+target:
+%
+\begin{align*}
+  \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
+\end{align*}
 
 % \begin{code}
-% module M where
-
-% ten :: Exp
-% ten = Val 10
-
-% square :: Exp -> Exp
-% square x = Mul x x
-
-% minus :: Exp -> Exp -> Exp
-% minus x y = Add x (Mul y (Val (-1)))
+% class (rep down target) where
+%   step :: rep target -> target
 % \end{code}
-\section{Extracting Structure} \label{sec:hrep}
-
-In this work, we ...
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{Val}\;{}\<[16]%
-\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Val}\;\Conid{Int}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Add}\;{}\<[16]%
-\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Add}\;\Varid{a}\;\Varid{a}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Mul}\;{}\<[16]%
-\>[16]{}\Varid{a}\mathrel{=}\text{Mk}_{Mul}\;\Varid{a}\;\Varid{a}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-For this purpose, we employ Haskell's type classes mechanism \tocite{}.
-%
-We define an evaluation type class $(\Downarrow)$ which specifies how to perform
-a single transformation \ensuremath{\Varid{step}} from the higher-order representation back into a
-concrete value of the target:
+
+Then, we can define the following overloaded instances of the \ensuremath{\Varid{step}} operation
+for the canonical \ensuremath{\text{HRep}} of data constructors simply by mapping each lifted
+constructor back into its corresponding one.
+%
+\begin{alignat*}{4}
+  &\evalrep{\ensuremath{\text{Mk}_{Val}}\ n\ \ \ &&}{\ensuremath{\Conid{Exp}}}
+    &&= \ensuremath{\Conid{Val}}\ n \\
+  &\evalrep{\ensuremath{\text{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
+    &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{\ensuremath{\Conid{Exp}}}\ \evalrep{y}{\ensuremath{\Conid{Exp}}} \\
+  &\evalrep{\ensuremath{\text{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
+    &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{\ensuremath{\Conid{Exp}}}\ \evalrep{y}{\ensuremath{\Conid{Exp}}}
+\end{alignat*}
+
+% \begin{code}
+% instance (HRep_Val down Exp) where
+%   step (Mk_Val n) = Val n
+
+% instance (HRep_Add down Exp) where
+%   step (Mk_Add x y) = Add x y
+
+% instance (HRep_Mul down Exp) where
+%   step (Mk_Mul x y) = Mul x y
+% \end{code}
 
@@
 \column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{class}\;(\Varid{rep}\;\Downarrow\;\Varid{target})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\mathbin{::}\Varid{rep}\;\Varid{target}\to \Varid{target}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Then, we can define the following overloaded instances of the \ensuremath{\Varid{step}} operation
-for the canonical representations of data constructors simply by mapping each
-lifted constructor back into its corresponding one.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{Val}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{Val}\;\Varid{n})\mathrel{=}\Conid{Val}\;\Varid{n}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{Add}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{Add}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Add}\;\Varid{x}\;\Varid{y}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{Mul}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{Mul}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{y}{}\<[E]%
+\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}gen_{Val}\;{}\<[12]%
+\>[12]{}gen_{r}{}\<[19]%
+\>[19]{}\mathrel{=}\text{Mk}_{Val}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
+\\
+\>[3]{}gen_{Add}\;{}\<[12]%
+\>[12]{}gen_{r}{}\<[19]%
+\>[19]{}\mathrel{=}\text{Mk}_{Add}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\\
+\>[3]{}gen_{Mul}\;{}\<[12]%
+\>[12]{}gen_{r}{}\<[19]%
+\>[19]{}\mathrel{=}\text{Mk}_{Mul}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
+
+\subsection{\textbf{Type Level Combinators}}
+
+In order to compose the \ensuremath{\text{HRep}}s for the single data constructors already
+presented, we will need to define a series of type level combinators.
+
+In first place, we define a type combinator ($\otimes$) to tag a \ensuremath{\text{HRep}} with an
+explicit generation frequency:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+This combinator is evaluated back to our target data type simply by piping the
+result from the inner representation.
+%
+It does not change the evaluation semantics, as it is only considered at
+generation time:
+%
+\begin{align*}
+  \evalrep{Freq\ x : f \otimes n}{t} = \evalrep{x : f}{t}
+\end{align*}
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}gen_{f \otimes n}\;gen_{r}\mathrel{=}\Conid{Term}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+% \begin{code}
+% instance (f otimes n down Exp) where
+%   step (Freq f) = step f
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f otimes n) where
+%   liftArbitrary gen_HRep = Tag <$> gen_HRep
+% \end{code} %$
+
+When generating random values, we will only allow certain terminal
+representations to be generated at the moment that the generation size gets
+exhausted.
+%
+For this, we follow the same approach as before to define a combinator (\ensuremath{\_^\bullet})
+to tag terminal representations.
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(f^\bullet)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{align*}
+  \evalrep{Term\ x : \ensuremath{f^\bullet}}{t} = \evalrep{x : f}{t}
+\end{align*}
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}gen_{f^\bullet}\;gen_{r}\mathrel{=}\Conid{Term}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
 \ColumnHook
@@
 With these individual representations, we can define a type combinator
-$(\oplus)$ to compose two representations into a single one:
+$(\oplus)$ to compose two \ensuremath{\text{HRep}}s into a single one:
 
@@
 target type by pattern matching on the data type variant and applying the \ensuremath{\Varid{step}}
-tranformation to the inner representation:
+tranformation to the inner \ensuremath{\text{HRep}}:
+%
+\begin{alignat*}{3}
+  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
+  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
+\end{alignat*}
 
@@
 \column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{instance}\;(\Varid{f}\;\oplus\;\Varid{g}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\Conid{L}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\Conid{R}\;\Varid{g})\mathrel{=}\Varid{step}\;\Varid{g}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Furthermore, we can define a type combinator ($\otimes$) to tag every
-representation with an explicit generation frequency:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This combinator is evaluated back to our target data type simply by piping the
-result from the inner representation.
-%
-It does not change the evaluation semantics, as it is only considered at
-generation time:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{instance}\;(\Varid{f}\;\otimes\;\Varid{n}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\Conid{Freq}\;\Varid{f})\mathrel{=}\Varid{step}\;\Varid{f}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{45}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}gen_{f \oplus g}\;gen_{r}{}\<[23]%
+\>[23]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}[\mskip1.5mu (\Varid{freq0}\;{}\<[17]%
+\>[17]{}\textit{@}\!f,{}\<[24]%
+\>[24]{}\Conid{L}{}\<[27]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;{}\<[38]%
+\>[38]{}gen_{r}{}\<[45]%
+\>[45]{}){}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{},(\Varid{freq0}\;{}\<[17]%
+\>[17]{}\textit{@}\!g,{}\<[24]%
+\>[24]{}\Conid{R}{}\<[27]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
+\>[38]{}gen_{r}{}\<[45]%
+\>[45]{})\mskip1.5mu]{}\<[E]%
+\\
+\>[3]{}\hsindent{2}{}\<[5]%
+\>[5]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{}[\mskip1.5mu (\Varid{freq}\;{}\<[17]%
+\>[17]{}\textit{@}\!f,{}\<[24]%
+\>[24]{}\Conid{L}{}\<[27]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;{}\<[38]%
+\>[38]{}gen_{r}{}\<[45]%
+\>[45]{}){}\<[E]%
+\\
+\>[5]{}\hsindent{2}{}\<[7]%
+\>[7]{},(\Varid{freq}\;{}\<[17]%
+\>[17]{}\textit{@}\!g,{}\<[24]%
+\>[24]{}\Conid{R}{}\<[27]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
+\>[38]{}gen_{r}{}\<[45]%
+\>[45]{})\mskip1.5mu]){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
+
+% \begin{code}
+% instance (f oplus g down Exp) where
+%   step (L f) = step f
+%   step (R g) = step g
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f oplus g) where
+%   liftArbitrary gen_HRep
+%     = frequency
+%       [ (freq at_f, L <$> gen_HRep)
+%       , (freq at_g, R <$> gen_HRep) ]
+% \end{code} %$
 
 With the introduced combinators, we can easily create a type synonym \ensuremath{\text{HRep}_{Exp}}
-to refer to the canonical representation of our original data type \ensuremath{\Conid{Exp}},
-tagging for instance the representation of the constuctor \ensuremath{\Conid{Add}} to be generated
-in double the proportion of rest of the representations:
+to refer to the canonical \ensuremath{\text{HRep}} of our original data type \ensuremath{\Conid{Exp}}, tagging for
+instance the representation of the constuctor \ensuremath{\Conid{Add}} to be generated in with
+double the frequency of rest:
 
@@
 \>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{Val}{}\<[E]%
-\\
-\>[16]{}\,\oplus\;{}\<[24]%
+\>[24]{}\text{HRep}_{Val}^\bullet{}\<[E]%
+\\
+\>[16]{}\!\;\oplus\;{}\<[24]%
 \>[24]{}\text{HRep}_{Add}\;{}\<[34]%
@@
 \\
-\>[16]{}\,\oplus\;{}\<[24]%
+\>[16]{}\!\;\oplus\;{}\<[24]%
 \>[24]{}\text{HRep}_{Mul}{}\<[E]%
@@
 
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\mathbf{data}\;\Conid{Fix}\;\Varid{f}\mathrel{=}\Conid{Fix}\;(\Varid{f}\;(\Conid{Fix}\;\Varid{f})){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{align*}
+  eval : Fix\ rep \rightarrow target
+\end{align*}
+
 \subsection{\textbf{Representing Pattern Matchings}}
@@
 \>[B]{}\mathbf{data}\;\text{HRep}_{foo\#1}\;{}\<[18]%
-\>[18]{}\Varid{a}\mathrel{=}\text{Mk}_{foo\#1}\;\Varid{a}\;\Varid{a}{}\<[E]%
+\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#1}\;\Varid{r}\;\Varid{r}{}\<[E]%
 \\
 \>[B]{}\mathbf{data}\;\text{HRep}_{foo\#2}\;{}\<[18]%
-\>[18]{}\Varid{a}\mathrel{=}\text{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{a}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{foo\#1}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{foo\#1}\;\Varid{x}\;\Varid{y}){}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathrel{=}\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y}){}\<[E]%
+\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{align*}
+  &\evalrep{\ensuremath{\text{Mk}_{foo\#1}}\ x\ y}{Exp} = \\
+  &\ \ \ensuremath{\Conid{Add}\;(\Conid{Add}\;\evalrep{x}{Exp}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\evalrep{y}{Exp})}\\
+  &\evalrep{\ensuremath{\text{Mk}_{foo\#2}}\ x\ y}{Exp} = \\
+  &\ \ \ensuremath{\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\evalrep{y}{Exp})}
+\end{align*}
+
+% \begin{code}
+% instance (HRep_foo_1 down Exp) where
+%   step (Mk_foo_1 x y)
+%     = Add (Add x (Val 50)) (Add (Val 25) y)
+
+% instance (HRep_foo_2 down Exp) where
+%   step (Mk_foo_2 x y)
+%     = Mul (Val 50) (Mul (Val x) y)
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_foo_1 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
+
+% instance Arbitrary1 HRep_foo_2 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
+% \end{code} %$
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
+\>[16]{}\mathrel{=}\text{HRep}_{foo\#1}\;\!\!\oplus\!\!\;{}\<[38]%
+\>[38]{}\text{HRep}_{foo\#2}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\subsection{\textbf{Representing Abstract Interfaces}}
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
 \\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{foo\#2}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{foo\#2}\;\Varid{x}\;\Varid{y}){}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathrel{=}\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{foo\#1}{}\<[E]%
-\\
-\>[16]{}\,\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{foo\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection{\textbf{Representing Abstract Interfaces}}
+\>[B]{}\Varid{ten}\mathbin{::}\Conid{Exp}{}\<[E]%
+\\
+\>[B]{}\Varid{ten}\mathrel{=}\Conid{Val}\;\mathrm{10}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
+\\
+\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
+\\
+\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Conid{Mul}\;\Varid{y}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1}))){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 \>[B]{}\mathbf{data}\;\text{HRep}_{ten}\;{}\<[21]%
-\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
+\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
 \\
 \>[B]{}\mathbf{data}\;\text{HRep}_{square}\;{}\<[21]%
-\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{square}\;{}\<[37]%
-\>[37]{}\Varid{a}{}\<[E]%
+\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{square}\;{}\<[37]%
+\>[37]{}\Varid{r}{}\<[E]%
 \\
 \>[B]{}\mathbf{data}\;\text{HRep}_{minus}\;{}\<[21]%
-\>[21]{}\Varid{a}\mathrel{=}\text{Mk}_{minus}\;{}\<[37]%
-\>[37]{}\Varid{a}\;\Varid{a}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{ten}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;\text{Mk}_{ten}\mathrel{=}\Varid{ten}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{square}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{square}\;\Varid{x})\mathrel{=}\Varid{square}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\mathbf{instance}\;(\text{HRep}_{minus}\;\Downarrow\;\Conid{Exp})\;\mathbf{where}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{step}\;(\text{Mk}_{minus}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{minus}\;\Varid{x}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{minus}\;{}\<[37]%
+\>[37]{}\Varid{r}\;\Varid{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{alignat*}{3}
+  &\evalrep{\ensuremath{\text{Mk}_{ten}}         &&}{Exp} = \ensuremath{\Varid{ten}} \\
+  &\evalrep{\ensuremath{\text{Mk}_{square}}\ x   &&}{Exp} = \ensuremath{\Varid{square}\;\evalrep{x}{Exp}} \\
+  &\evalrep{\ensuremath{\text{Mk}_{minus}}\ x\ y &&}{Exp} = \ensuremath{\Varid{minus}\;\evalrep{x}{Exp}\;\evalrep{y}{Exp}}
+\end{alignat*}
+
+% \begin{code}
+% instance (HRep_ten down Exp) where
+%   step Mk_ten = ten
+
+% instance (HRep_square down Exp) where
+%   step (Mk_square x) = square x
+
+% instance (HRep_minus down Exp) where
+%   step (Mk_minus x y) = minus x y
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_ten where
+%   liftArbitrary gen_HRep = pure Mk_ten
+
+% instance Arbitrary1 HRep_square where
+%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
+
+% instance Arbitrary1 HRep_minus where
+%   liftArbitrary gen_HRep
+%     = Mk_minus <$> gen_HRep <*> gen_HRep
+% \end{code} %$
 
@@
 \column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{55}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
 \>[B]{}\mathbf{type}\;\text{HRep}_{M}{}\<[14]%
-\>[14]{}\mathrel{=}{}\<[22]%
-\>[22]{}\text{HRep}_{ten}{}\<[E]%
-\\
-\>[14]{}\,\oplus\;{}\<[22]%
-\>[22]{}\text{HRep}_{square}{}\<[E]%
-\\
-\>[14]{}\,\oplus\;{}\<[22]%
-\>[22]{}\text{HRep}_{minus}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{Spec}_{Exp}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{Exp}\;{}\<[34]%
-\>[34]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[16]{}\,\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{foo}\;{}\<[34]%
-\>[34]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[16]{}\,\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{M}{}\<[E]%
+\>[14]{}\mathrel{=}\text{HRep}_{ten}\;\!\!\oplus\!\!\;{}\<[34]%
+\>[34]{}\text{HRep}_{square}\;\!\!\oplus\!\!\;{}\<[55]%
+\>[55]{}\text{HRep}_{minus}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\Conid{Spec}{}\<[12]%
+\>[12]{}\mathrel{=}{}\<[20]%
+\>[20]{}\text{HRep}_{Exp}\;{}\<[30]%
+\>[30]{}\otimes\;\mathrm{4}{}\<[E]%
+\\
+\>[12]{}\!\;\oplus\;{}\<[20]%
+\>[20]{}\text{HRep}_{foo}\;{}\<[30]%
+\>[30]{}\otimes\;\mathrm{2}{}\<[E]%
+\\
+\>[12]{}\!\;\oplus\;{}\<[20]%
+\>[20]{}\text{HRep}_{M}{}\<[E]%
 \ColumnHook
@@
   \input{tikz/hrep.tex}
-  \caption{Higher order representation of the data type \ensuremath{\Conid{Exp}}, using structural
-    information from the function \ensuremath{\Varid{foo}} and the abstract interface of the module
-    \ensuremath{\Conid{M}}.}
+  \caption{Higher level representation of the data type \ensuremath{\Conid{Exp}}, defined using
+    structural information from the function \ensuremath{\Varid{foo}} and the abstract interface of
+    the module \ensuremath{\Conid{M}}.}
   \label{fig:hrep}

### 2019-01-17 paper/sections/hrep.tex
Unfolded figure
WIP

--- 2019-01-15 paper/sections/hrep.tex
@@
 
-What to say here:
+In this section we present a compositional representation to express the random
+generation of values following the internal structure of their data types along
+with the structure present on patterns matchings and abstract interfaces.
 
-\begin{itemize}
-\item Haskell ADTs
-\item Type driven generation
-\item Limitations
-\item Key idea: Higher order representations
-\item Canonical representation of the data type
-\item Pattern matchings representation
-\item Abstract interface representation
-\item Composing representations
-\end{itemize}
+The key idea of this work is to represent different sources of structured value
+information of a data type in a homogeneous way that we call a ``higher-level
+representation'' (|HRep| from now on).%
+\footnote{The notion of higher-level comes from that the generation process of
+  given target data type is entirely determined by the type of the chosen
+  representation, instead of by a concrete generator defined at the term level.}
+%
+For this purpose, we use a series of automatically derived data types, each one
+representing an atomic unit of information that can be randomly generated and
+then reflected back to the corresponding value of the original data type.
 
-\begin{figure}[b]
+We wil reuse the previously defined data type |Exp| and the function |foo| to
+explain the different concepts involved all across this section.
+
+\subsection{\textbf{Representing Data Constructors}}
+
+We begin by introducing the simplest representation that we can extract from our
+codebase: the representation of single data constructor.
+%
+Each original constructor can be represented by a automatically derived data
+type consisting of a single constructor with the same fields, except for the
+recursive ones that are abstracted away.
+%
+In this light, we represent each constructor of the data type |Exp| as follows:
+
+\begin{code}
+data HRep_Val  r = Mk_Val Int
+data HRep_Add  r = Mk_Add r r
+data HRep_Mul  r = Mk_Mul r r
+\end{code}
+
+Note that the previous definition are type parametric over the type parameter
+|r|.
+%
+This way we can replace |r| with any concrete data type, obtaining different
+possible values on each case.
+%
+For instance, the value (|Mk_Add 10 20|) has type |HRep_Add Int|, while the
+value (|Mk_Mul True False|) has type |HRep_Mul Bool|.
+%
+In practice, this parametricity let us instatiate |a| with the type of the
+chosen representation (which might change according to the user demands),
+without having to modify anything in the underlying machinery.
+
+For this purpose, we employ Haskell's type classes mechanism \tocite{}.
+%
+We define an evaluation type class $(\Downarrow)$ which specifies how to perform
+a single transformation |step| from an |HRep| back into a concrete value of the
+target:
+%
+\begin{align*}
+  \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
+\end{align*}
+
+% \begin{code}
+% class (rep down target) where
+%   step :: rep target -> target
+% \end{code}
+
+Then, we can define the following overloaded instances of the |step| operation
+for the canonical |HRep| of data constructors simply by mapping each lifted
+constructor back into its corresponding one.
+%
+\begin{alignat*}{4}
+  &\evalrep{|Mk_Val|\ n\ \ \ &&}{|Exp|}
+    &&= |Val|\ n \\
+  &\evalrep{|Mk_Add|\ x\ y   &&}{|Exp|}
+    &&= |Add|\ \evalrep{x}{|Exp|}\ \evalrep{y}{|Exp|} \\
+  &\evalrep{|Mk_Mul|\ x\ y   &&}{|Exp|}
+    &&= |Mul|\ \evalrep{x}{|Exp|}\ \evalrep{y}{|Exp|}
+\end{alignat*}
+
+% \begin{code}
+% instance (HRep_Val down Exp) where
+%   step (Mk_Val n) = Val n
+
+% instance (HRep_Add down Exp) where
+%   step (Mk_Add x y) = Add x y
+
+% instance (HRep_Mul down Exp) where
+%   step (Mk_Mul x y) = Mul x y
+% \end{code}
+
+\begin{code}
+  gen_Val  gen_r  = Mk_Val  <$> arbitrary
+  gen_Add  gen_r  = Mk_Add  <$> smaller gen_r <*> smaller gen_r
+  gen_Mul  gen_r  = Mk_Mul  <$> smaller gen_r <*> smaller gen_r
+\end{code} %$
+
+\subsection{\textbf{Type Level Combinators}}
+
+In order to compose the |HRep|s for the single data constructors already
+presented, we will need to define a series of type level combinators.
+
+In first place, we define a type combinator ($\otimes$) to tag a |HRep| with an
+explicit generation frequency:
+
+\begin{code}
+data (f otimes n) a = Freq (f a)
+\end{code}
+
+This combinator is evaluated back to our target data type simply by piping the
+result from the inner representation.
+%
+It does not change the evaluation semantics, as it is only considered at
+generation time:
+%
+\begin{align*}
+  \evalrep{Freq\ x : f \otimes n}{t} = \evalrep{x : f}{t}
+\end{align*}
+
+\begin{code}
+gen_f_times_n gen_r = Term <$> gen_f gen_r
+\end{code}
+
+% \begin{code}
+% instance (f otimes n down Exp) where
+%   step (Freq f) = step f
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f otimes n) where
+%   liftArbitrary gen_HRep = Tag <$> gen_HRep
+% \end{code} %$
+
+When generating random values, we will only allow certain terminal
+representations to be generated at the moment that the generation size gets
+exhausted.
+%
+For this, we follow the same approach as before to define a combinator (|term|)
+to tag terminal representations.
+
+\begin{code}
+data (f_term) a = Term (f a)
+\end{code}
+
+\begin{align*}
+  \evalrep{Term\ x : |f_term|}{t} = \evalrep{x : f}{t}
+\end{align*}
+
+\begin{code}
+gen_f_term gen_r = Term <$> gen_f gen_r
+\end{code}
+
+With these individual representations, we can define a type combinator
+$(\oplus)$ to compose two |HRep|s into a single one:
+
+\begin{code}
+data (f oplus g) a = L (f a) | R (g a)
+\end{code}
+
+Then, a composite representation can be transformed back into the concrete
+target type by pattern matching on the data type variant and applying the |step|
+tranformation to the inner |HRep|:
+%
+\begin{alignat*}{3}
+  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
+  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
+\end{alignat*}
+
+\begin{code}
+  gen_f_plus_g gen_r  = sized (\size ->
+    if size == 0
+    then frequency
+      [ (freq0  at_f,  L  <$> gen_f  gen_r  )
+      , (freq0  at_g,  R  <$> gen_g  gen_r  ) ]
+    else frequency
+      [ (freq   at_f,  L  <$> gen_f  gen_r  )
+      , (freq   at_g,  R  <$> gen_g  gen_r  ) ])
+\end{code} %$
+
+% \begin{code}
+% instance (f oplus g down Exp) where
+%   step (L f) = step f
+%   step (R g) = step g
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f oplus g) where
+%   liftArbitrary gen_HRep
+%     = frequency
+%       [ (freq at_f, L <$> gen_HRep)
+%       , (freq at_g, R <$> gen_HRep) ]
+% \end{code} %$
+
+With the introduced combinators, we can easily create a type synonym |HRep_Exp|
+to refer to the canonical |HRep| of our original data type |Exp|, tagging for
+instance the representation of the constuctor |Add| to be generated in with
+double the frequency of rest:
+
+\begin{code}
+type HRep_Exp  =       HRep_Val_term
+               oplus'  HRep_Add  otimes 2
+               oplus'  HRep_Mul
+\end{code}
+
+\begin{code}
+  data Fix f = Fix (f (Fix f))
+\end{code}
+
+\begin{align*}
+  eval : Fix\ rep \rightarrow target
+\end{align*}
+
+\subsection{\textbf{Representing Pattern Matchings}}
+
+\begin{code}
+data HRep_foo_1  r = Mk_foo_1 r r
+data HRep_foo_2  r = Mk_foo_2 Int r
+\end{code}
+
+\begin{align*}
+  &\evalrep{|Mk_foo_1|\ x\ y}{Exp} = \\
+  &\ \ |Add (Add eval_x_Exp (Val 50)) (Add (Val 25) eval_y_Exp)|\\
+  &\evalrep{|Mk_foo_2|\ x\ y}{Exp} = \\
+  &\ \ |Mul (Val 50) (Mul (Val x) eval_y_Exp)|
+\end{align*}
+
+% \begin{code}
+% instance (HRep_foo_1 down Exp) where
+%   step (Mk_foo_1 x y)
+%     = Add (Add x (Val 50)) (Add (Val 25) y)
+
+% instance (HRep_foo_2 down Exp) where
+%   step (Mk_foo_2 x y)
+%     = Mul (Val 50) (Mul (Val x) y)
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_foo_1 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
+
+% instance Arbitrary1 HRep_foo_2 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
+% \end{code} %$
+
+\begin{code}
+type HRep_foo  = HRep_foo_1 oplus''  HRep_foo_2
+\end{code}
+
+\subsection{\textbf{Representing Abstract Interfaces}}
+
+\begin{code}
+module M where
+
+ten :: Exp
+ten = Val 10
+
+square :: Exp -> Exp
+square x = Mul x x
+
+minus :: Exp -> Exp -> Exp
+minus x y = Add x (Mul y (Val (-1)))
+\end{code}
+
+\begin{code}
+data HRep_ten       r = Mk_ten
+data HRep_square    r = Mk_square   r
+data HRep_minus     r = Mk_minus    r r
+\end{code}
+
+\begin{alignat*}{3}
+  &\evalrep{|Mk_ten|         &&}{Exp} = |ten| \\
+  &\evalrep{|Mk_square|\ x   &&}{Exp} = |square eval_x_Exp| \\
+  &\evalrep{|Mk_minus|\ x\ y &&}{Exp} = |minus  eval_x_Exp eval_y_Exp|
+\end{alignat*}
+
+% \begin{code}
+% instance (HRep_ten down Exp) where
+%   step Mk_ten = ten
+
+% instance (HRep_square down Exp) where
+%   step (Mk_square x) = square x
+
+% instance (HRep_minus down Exp) where
+%   step (Mk_minus x y) = minus x y
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_ten where
+%   liftArbitrary gen_HRep = pure Mk_ten
+
+% instance Arbitrary1 HRep_square where
+%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
+
+% instance Arbitrary1 HRep_minus where
+%   liftArbitrary gen_HRep
+%     = Mk_minus <$> gen_HRep <*> gen_HRep
+% \end{code} %$
+
+\begin{code}
+type HRep_M  = HRep_ten oplus''  HRep_square oplus''  HRep_minus
+\end{code}
+
+\begin{code}
+type Spec  =       HRep_Exp  otimes 4
+           oplus'  HRep_foo  otimes 2
+           oplus'  HRep_M
+\end{code}
+
+This previous definition can be interpreted graphically as it is shown in the
+Figure \ref{fig:hrep}.
+%
+Curly arrows represent the structural information extracted using
+meta-programming.
+
+\begin{figure}[t]
   \centering
   \input{tikz/hrep.tex}
-  \caption{The idea}
+  \caption{Higher level representation of the data type |Exp|, defined using
+    structural information from the function |foo| and the abstract interface of
+    the module |M|.}
+  \label{fig:hrep}
 \end{figure}
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+% \begin{code}
+% data (Term f) a = TagTerm (f a)
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+% instance (Term f down Exp) where
+%   step (TagTerm f) = step f
+% \end{code}

### 2019-01-17 paper/sections/randomtesting.tex
WIP

@@
 %
-In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
+In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} (or
+constructors for short) and are used to distiguish which variant of the data
+type we are reffering to.
 %
@@
 \begin{code}
-genExp :: Gen Exp
-genExp = sized gen
-  where
-    gen 0 = Val <$> arbitrary
-    gen n = oneof
-      [  Val  <$> arbitrary
-      ,  Add  <$> gen pred   <*> gen pred
-      ,  Mul  <$> gen pred   <*> gen pred ]
+gen_Exp = sized (\size ->
+   if size == 0
+   then Val <$> arbitrary
+   else frequency
+     [  (1, Val  <$> arbitrary)
+     ,  (2, Add  <$> smaller gen_Exp <*> smaller gen_Exp)
+     ,  (1, Mul  <$> smaller gen_Exp <*> smaller gen_Exp) ])
 \end{code} %$
 
-This random generator is defined using \quickcheck's combinator |sized :: (Int
--> Gen a) -> Gen a| to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
+This random generator is defined using \quickcheck's combinator |sized| to
+paremetrize the generation process up to an external integer number known as the
+\emph{generation size}.
 %
@@
 %
-When called with a positive generation size (case |gen n|), the generator is free
-to randomly pick with uniform probability among any data constructor of |Exp|.
+When called with a positive generation size (case |gen n|), the generator will
+pick among any data constructor of |Exp| with a explicitly given frequency---two
+times more |Add|s than |Val| or |Mul|.
 %
@@
 data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (|gen
-pred|), ensuring to decrease the generation size by a unit on each recursive
+corresponding to the data constructor fields by calling itself recursively
+(|resize pred gen_Exp|), decreasing the generation size on each recursive
 invocation.
@@
 first clause of |foo| (and hence triggering the first error) results as follows:
-
+%
 \begin{align*}
-  P(match(foo\#1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
+  &P(match(foo\#1))\\
+  &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
+  &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
 \end{align*}
@@
 
-concat :: Html -> Html -> Html
-concat (Html x) (Html y) = Html (x ++ y)
+(<+>) :: Html -> Html -> Html
+(Html x) <+> (Html y) = Html (x ++ y)
 \end{code}
@@
 generators in a modular and flexible way.
-
-% Consider the following Haskell module |M| which defines combinators for the type
-% |Exp|:
-
-% \begin{code}
-% module M where
-
-% ten :: Exp
-% ten = Val 10
-
-% square :: Exp -> Exp
-% square x = Mul x x
-
-% minus :: Exp -> Exp -> Exp
-% minus x y = Add x (Mul y (Val (-1)))
-% \end{code}

### 2019-01-17 paper/main.tex
WIP

@@
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
 Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
@@
 \item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
+  the target data type by reifying the static information present on the program
   codebase (Section \ref{sec:hrep}).
@@
 \item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
+  to perform automatic derivation of random generators capable to extract useful
+  structure information from the user source code.
   %
@@
 
-\subsection{\textbf{Type-Driven Generation of Random Values}}
+\subsection*{\textbf{Type-Driven Generation of Random Values}}
 
@@
 
-\subsection{\textbf{Presence of Complex Pattern Matchings}}
+\subsection*{\textbf{Presence of Complex Pattern Matchings}}
 
@@
 
-\subsection{\textbf{Data Invariants Encoded on Abstract Interfaces}}
+\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
 
@@
 then reflected back to the corresponding value of the original data type.
+%
+Later, the user can compose these atomic representations using the provided type
+level combinators in different ways into a ``generation specification'' that
+completely determines the generation process.
 
@@
 
-\subsection{\textbf{Representing Data Constructors}}
-
-We begin by introducing the simplest representation that we can extract from our
-codebase: the representation of single data constructor.
-%
-Each original constructor can be represented by a automatically derived data
-type consisting of a single constructor with the same fields, except for the
-recursive ones that are abstracted away.
+\subsection*{\textbf{Representing Data Constructors}}
+
+We begin by introducing the simplest data type representation that we can
+extract from our codebase: the representation of single data constructor.
+%
+Each data constructor can be represented by an automatically derived data type
+consisting of a single constructor with the same fields as the original, except
+for the recursive ones that are abstracted away.
 %
@@
 
-Note that the previous definition are type parametric over the type parameter
+Note that the previous definitions are type parametric over the type parameter
 \ensuremath{\Varid{r}}.
 %
-This way we can replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
+This allow us to replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
 possible values on each case.
@@
 %
-In practice, this parametricity let us instatiate \ensuremath{\Varid{a}} with the type of the
-chosen representation (which might change according to the user demands),
+In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
+chosen generation specification (which might be composed of several \ensuremath{\text{HRep}}s),
 without having to modify anything in the underlying machinery.
 
-For this purpose, we employ Haskell's type classes mechanism \tocite{}.
-%
-We define an evaluation type class $(\Downarrow)$ which specifies how to perform
-a single transformation \ensuremath{\Varid{step}} from an \ensuremath{\text{HRep}} back into a concrete value of the
-target:
-%
-\begin{align*}
-  \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-\end{align*}
-
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-
-Then, we can define the following overloaded instances of the \ensuremath{\Varid{step}} operation
-for the canonical \ensuremath{\text{HRep}} of data constructors simply by mapping each lifted
-constructor back into its corresponding one.
+Having the \ensuremath{\text{HRep}} of each data constructor, we can define an evaluation relation
+($\evalrep{\_}{t} : \ensuremath{\text{HRep}}_f \rightarrow t$) that maps a value from each
+representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
+%
+Then, we simply need to translate each constructor representation back into its
+corresponding one, translating the abstracted fields recursively:
 %
@@
   &\evalrep{\ensuremath{\text{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{\ensuremath{\Conid{Exp}}}\ \evalrep{y}{\ensuremath{\Conid{Exp}}} \\
+    &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
   &\evalrep{\ensuremath{\text{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{\ensuremath{\Conid{Exp}}}\ \evalrep{y}{\ensuremath{\Conid{Exp}}}
+    &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
 \end{alignat*}
 
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
+The last missing piece is to automatically derive a random generators for each
+constructor \ensuremath{\text{HRep}}.
+%
+For this purpose, it is important to consider that each constructor \ensuremath{\text{HRep}} has
+its recursive fields abstracted away with a type parameter that will be later
+instantiated with the generation specification type.
+%
+Given that this specification is unknown at the derivation time, we parametrize
+each \ensuremath{\text{HRep}} generator with a random generator \ensuremath{gen_{r}} that is used to generate
+each recursive fields:
 
@@
 
-\subsection{\textbf{Type Level Combinators}}
-
-In order to compose the \ensuremath{\text{HRep}}s for the single data constructors already
-presented, we will need to define a series of type level combinators.
-
-In first place, we define a type combinator ($\otimes$) to tag a \ensuremath{\text{HRep}} with an
-explicit generation frequency:
+\subsection*{\textbf{Type Level Combinators}}
+
+In this work we define type level combinators which let the user combine the
+automatically derived \ensuremath{\text{HRep}}s in several ways.
+%
+In first place, we define a type combinator ($\_^\bullet$) to tag a \ensuremath{\text{HRep}} to be
+terminal, i.e., a representation that is allowed be generated when the
+generation size gets exausted:
+%
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;(f^\bullet)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\text{HRep}} with an
+explicit generation frequency $n$:
 
@@
 
-This combinator is evaluated back to our target data type simply by piping the
-result from the inner representation.
-%
-It does not change the evaluation semantics, as it is only considered at
-generation time:
-%
-\begin{align*}
-  \evalrep{Freq\ x : f \otimes n}{t} = \evalrep{x : f}{t}
-\end{align*}
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{f \otimes n}\;gen_{r}\mathrel{=}\Conid{Term}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-When generating random values, we will only allow certain terminal
-representations to be generated at the moment that the generation size gets
-exhausted.
-%
-For this, we follow the same approach as before to define a combinator (\ensuremath{\_^\bullet})
-to tag terminal representations.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(f^\bullet)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{align*}
-  \evalrep{Term\ x : \ensuremath{f^\bullet}}{t} = \evalrep{x : f}{t}
-\end{align*}
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{f^\bullet}\;gen_{r}\mathrel{=}\Conid{Term}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-With these individual representations, we can define a type combinator
-$(\oplus)$ to compose two \ensuremath{\text{HRep}}s into a single one:
+The previous combinators only include information relevant to the generation
+process, in a sense that neither one adds new structure to the final
+representation.
+%
+In this light, they do not alter the evaluation semantics, and we translate them
+back to our target data type by evaluating the inner representation:
+%
+\begin{alignat*}{4}
+  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
+  &\evalrep{Term\ x &&: \ensuremath{f^\bullet}    &&}{t} &&= \evalrep{x : f}{t}
+\end{alignat*}
+
+In the previous equations we explicitly annotate (using a colon) the type of the
+evaluated term for clarity.
+%
+Later, to generate these combinators is enough to wrap a generated value from
+the inner representation with the apropriate tag:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}gen_{f^\bullet}\;{}\<[16]%
+\>[16]{}gen_{r}{}\<[23]%
+\>[23]{}\mathrel{=}\Conid{Term}{}\<[31]%
+\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
+\\
+\>[B]{}gen_{f \otimes n}\;{}\<[16]%
+\>[16]{}gen_{r}{}\<[23]%
+\>[23]{}\mathrel{=}\Conid{Freq}{}\<[31]%
+\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
+\ensuremath{\text{HRep}}s into a single one using a sum type and representing a random choice
+between them:
 
@@
 
-Then, a composite representation can be transformed back into the concrete
-target type by pattern matching on the data type variant and applying the \ensuremath{\Varid{step}}
-tranformation to the inner \ensuremath{\text{HRep}}:
+A composite representation built using $(\oplus)$ is transformed back into the
+target data type by pattern matching on the data type variant and evaluating the
+inner \ensuremath{\text{HRep}}:
 %
@@
 \end{alignat*}
+
+Generating a composite \ensuremath{\text{HRep}} is slightly more complicated than before, as we
+need to perform a random choice based on the generation size and the given
+frequencies for each sub-represention:
 
@@
 
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
+The previous definition we reflect the type level frequencies of \ensuremath{\Varid{f}} and \ensuremath{\Varid{g}}
+(\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}).
+%
+This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
+used.
+%
+Then we use these frequencies to generate each inner \ensuremath{\text{HRep}} in the apropriate
+proportion.
+%
+When the generation size gets exhausted, we reflect the terminal generation
+frequency of each inner \ensuremath{\text{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
+\ensuremath{\Varid{freq0}\;\textit{@}\!g}).
+%
+This time, however, we default the frequency reflection to $0$ for any inner
+that not tagged as terminal, avoiding to generate non-terminal constructions in
+the last step.
 
 With the introduced combinators, we can easily create a type synonym \ensuremath{\text{HRep}_{Exp}}
-to refer to the canonical \ensuremath{\text{HRep}} of our original data type \ensuremath{\Conid{Exp}}, tagging for
-instance the representation of the constuctor \ensuremath{\Conid{Add}} to be generated in with
-double the frequency of rest:
+to specify a generation schema equivalent to the one seen in the concrete random
+generator of type \ensuremath{\Conid{Exp}} presented in Section \ref{sec:randomtesting}:
 
@@
 
+However, a value of type \ensuremath{\text{HRep}_{Exp}} still has its recursive calls abstracted
+away---the type parameter \ensuremath{\Varid{r}} is implicit in the definition of \ensuremath{\text{HRep}_{Exp}}.
+%
+If we want to generate recursive values using this representation we need to
+introduce a last type level combinator to ``tie the knot'' recursively:
+
 \begin{hscode}\SaveRestoreHook
@@
 
+This datatype represents the \emph{fixed point} of a parametric data type \ensuremath{\Varid{f}},
+i.e., a data type where each recursive call gets instantiated with iself.
+%
+To translate fixed points back to our target data type we simply need translate
+the inner representation:
+%
 \begin{align*}
-  eval : Fix\ rep \rightarrow target
+  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
 \end{align*}
-
-\subsection{\textbf{Representing Pattern Matchings}}
+%
+Categorically, this generic transformation of fixed points is called a
+\emph{catamorphism}, where our evaluation relation is known as its
+\emph{algebra}.
+
+Unlike the other combinators, to randomly generate a fixed point of a certain
+representation \ensuremath{\Varid{f}}, we do not parametrize the generation of the recursive fields
+of the inner represention over an external generator \ensuremath{gen_{r}}.
+%
+Instead, we can also tie the recursive knot in our random generator by piping
+the generation of the inner \ensuremath{\text{HRep}} parametrized recursively by the itself:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;genFix_{f}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
+
+This way we obtain a concrete recursive generator for each representation \ensuremath{\Varid{f}}
+that we define.
+%
+With this generator, we can define a random generator for our target data type
+simply by evaluating a random value of the representation:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}gen_{Exp}\mathrel{=}\mathbf{do}\;{}\<[17]%
+\>[17]{}\Varid{x}\leftarrow genFix_{(\text{HRep}\,{Exp})}{}\<[E]%
+\\
+\>[17]{}\Varid{return}\;\evalrep{x}{Exp}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
+
+\subsection*{\textbf{Representing Pattern Matchings}}
 
@@
 
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{33}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{52}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}gen_{foo\#1}\;{}\<[14]%
+\>[14]{}gen_{r}{}\<[21]%
+\>[21]{}\mathrel{=}\text{Mk}_{foo\#1}{}\<[33]%
+\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[52]%
+\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\\
+\>[3]{}gen_{foo\#2}\;{}\<[14]%
+\>[14]{}gen_{r}{}\<[21]%
+\>[21]{}\mathrel{=}\text{Mk}_{foo\#2}{}\<[33]%
+\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[52]%
+\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
 
@@
 
-\subsection{\textbf{Representing Abstract Interfaces}}
+\subsection*{\textbf{Representing Abstract Interfaces}}
 
@@
 
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{35}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{54}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}gen_{ten}\;{}\<[15]%
+\>[15]{}gen_{r}{}\<[22]%
+\>[22]{}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
+\\
+\>[3]{}gen_{square}\;{}\<[15]%
+\>[15]{}gen_{r}{}\<[22]%
+\>[22]{}\mathrel{=}\text{Mk}_{square}{}\<[35]%
+\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\\
+\>[3]{}gen_{minus}\;{}\<[14]%
+\>[14]{}gen_{r}{}\<[22]%
+\>[22]{}\mathrel{=}\text{Mk}_{minus}{}\<[35]%
+\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[54]%
+\>[54]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
 
@@
 
+% \begin{align*}
+%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
+% \end{align*}
+%
+% \begin{code}
+% class (rep down target) where
+%   step :: rep target -> target
+% \end{code}
+%
 % \begin{code}
 % data (Term f) a = TagTerm (f a)
+
+% \begin{code}
+% instance (HRep_Val down Exp) where
+%   step (Mk_Val n) = Val n
+
+% instance (HRep_Add down Exp) where
+%   step (Mk_Add x y) = Add x y
+
+% instance (HRep_Mul down Exp) where
+%   step (Mk_Mul x y) = Mul x y
+% \end{code}
+
+% \begin{code}
+% instance (f otimes n down Exp) where
+%   step (Freq f) = step f
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f otimes n) where
+%   liftArbitrary gen_HRep = Tag <$> gen_HRep
+% \end{code} %$
 
@@
 % \end{code}
+
+% \begin{code}
+% instance (HRep_ten down Exp) where
+%   step Mk_ten = ten
+
+% instance (HRep_square down Exp) where
+%   step (Mk_square x) = square x
+
+% instance (HRep_minus down Exp) where
+%   step (Mk_minus x y) = minus x y
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_ten where
+%   liftArbitrary gen_HRep = pure Mk_ten
+
+% instance Arbitrary1 HRep_square where
+%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
+
+% instance Arbitrary1 HRep_minus where
+%   liftArbitrary gen_HRep
+%     = Mk_minus <$> gen_HRep <*> gen_HRep
+% \end{code} %$
+
+% \begin{code}
+% instance (HRep_foo_1 down Exp) where
+%   step (Mk_foo_1 x y)
+%     = Add (Add x (Val 50)) (Add (Val 25) y)
+
+% instance (HRep_foo_2 down Exp) where
+%   step (Mk_foo_2 x y)
+%     = Mul (Val 50) (Mul (Val x) y)
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_foo_1 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
+
+% instance Arbitrary1 HRep_foo_2 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
+% \end{code} %$
+
+% \begin{code}
+% instance (f oplus g down Exp) where
+%   step (L f) = step f
+%   step (R g) = step g
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f oplus g) where
+%   liftArbitrary gen_HRep
+%     = frequency
+%       [ (freq at_f, L <$> gen_HRep)
+%       , (freq at_g, R <$> gen_HRep) ]
+% \end{code} %$
 \section{Random Generators Synthesis} \label{sec:synthesis}
@@
 Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
 % %include sections/casestudies.tex
@@
 ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+vulputate vestibulum lorem.
 

### 2019-01-17 paper/sections/hrep.tex
WIP

@@
 
-\subsection{\textbf{Representing Data Constructors}}
+\subsection*{\textbf{Representing Data Constructors}}
 
-We begin by introducing the simplest representation that we can extract from our
-codebase: the representation of single data constructor.
+We begin by introducing the simplest data type representation that we can
+extract from our codebase: the representation of single data constructor.
 %
-Each original constructor can be represented by a automatically derived data
-type consisting of a single constructor with the same fields, except for the
-recursive ones that are abstracted away.
+Each data constructor can be represented by an automatically derived data type
+consisting of a single constructor with the same fields as the original, except
+for the recursive ones that are abstracted away.
 %
@@
 
-Note that the previous definition are type parametric over the type parameter
+Note that the previous definitions are type parametric over the type parameter
 |r|.
 %
-This way we can replace |r| with any concrete data type, obtaining different
+This allow us to replace |r| with any concrete data type, obtaining different
 possible values on each case.
@@
 %
-In practice, this parametricity let us instatiate |a| with the type of the
-chosen representation (which might change according to the user demands),
+In practice, this parametricity let us instatiate |r| with the type of the
+chosen generation specification (which might be composed of several |HRep|s),
 without having to modify anything in the underlying machinery.
 
-For this purpose, we employ Haskell's type classes mechanism \tocite{}.
+Having the |HRep| of each data constructor, we can define an evaluation relation
+($\evalrep{\_}{t} : |HRep|_f \rightarrow t$) that maps a value from each
+representation |f| back to the target data type |t|.
 %
-We define an evaluation type class $(\Downarrow)$ which specifies how to perform
-a single transformation |step| from an |HRep| back into a concrete value of the
-target:
-%
-\begin{align*}
-  \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-\end{align*}
-
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-
-Then, we can define the following overloaded instances of the |step| operation
-for the canonical |HRep| of data constructors simply by mapping each lifted
-constructor back into its corresponding one.
+Then, we simply need to translate each constructor representation back into its
+corresponding one, translating the abstracted fields recursively:
 %
@@
   &\evalrep{|Mk_Add|\ x\ y   &&}{|Exp|}
-    &&= |Add|\ \evalrep{x}{|Exp|}\ \evalrep{y}{|Exp|} \\
+    &&= |Add|\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
   &\evalrep{|Mk_Mul|\ x\ y   &&}{|Exp|}
-    &&= |Mul|\ \evalrep{x}{|Exp|}\ \evalrep{y}{|Exp|}
+    &&= |Mul|\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
 \end{alignat*}
 
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
+The last missing piece is to automatically derive a random generators for each
+constructor |HRep|.
+%
+For this purpose, it is important to consider that each constructor |HRep| has
+its recursive fields abstracted away with a type parameter that will be later
+instantiated with the generation specification type.
+%
+Given that this specification is unknown at the derivation time, we parametrize
+each |HRep| generator with a random generator |gen_r| that is used to generate
+each recursive fields:
 
@@
 
-\subsection{\textbf{Type Level Combinators}}
+\subsection*{\textbf{Type Level Combinators}}
 
-In order to compose the |HRep|s for the single data constructors already
-presented, we will need to define a series of type level combinators.
+In this work we define type level combinators which let the user combine the
+automatically derived |HRep|s in several ways.
+%
+In first place, we define a type combinator ($\_^\bullet$) to tag a |HRep| to be
+terminal, i.e., a representation that is allowed be generated when the
+generation size gets exausted:
+%
+\begin{code}
+data (f_term) a = Term (f a)
+\end{code}
 
-In first place, we define a type combinator ($\otimes$) to tag a |HRep| with an
-explicit generation frequency:
+Additionally, we define a combinator ($\otimes$) to tag a |HRep| with an
+explicit generation frequency $n$:
 
@@
 
-This combinator is evaluated back to our target data type simply by piping the
-result from the inner representation.
+The previous combinators only include information relevant to the generation
+process, in a sense that neither one adds new structure to the final
+representation.
 %
-It does not change the evaluation semantics, as it is only considered at
-generation time:
+In this light, they do not alter the evaluation semantics, and we translate them
+back to our target data type by evaluating the inner representation:
 %
-\begin{align*}
-  \evalrep{Freq\ x : f \otimes n}{t} = \evalrep{x : f}{t}
-\end{align*}
+\begin{alignat*}{4}
+  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
+  &\evalrep{Term\ x &&: |f_term|    &&}{t} &&= \evalrep{x : f}{t}
+\end{alignat*}
 
-\begin{code}
-gen_f_times_n gen_r = Term <$> gen_f gen_r
-\end{code}
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-When generating random values, we will only allow certain terminal
-representations to be generated at the moment that the generation size gets
-exhausted.
+In the previous equations we explicitly annotate (using a colon) the type of the
+evaluated term for clarity.
 %
-For this, we follow the same approach as before to define a combinator (|term|)
-to tag terminal representations.
-
-\begin{code}
-data (f_term) a = Term (f a)
-\end{code}
-
-\begin{align*}
-  \evalrep{Term\ x : |f_term|}{t} = \evalrep{x : f}{t}
-\end{align*}
+Later, to generate these combinators is enough to wrap a generated value from
+the inner representation with the apropriate tag:
 
@@
 gen_f_term     gen_r  = Term  <$> gen_f gen_r
+gen_f_times_n  gen_r  = Freq  <$> gen_f gen_r
 \end{code}
 
-With these individual representations, we can define a type combinator
-$(\oplus)$ to compose two |HRep|s into a single one:
+Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
+|HRep|s into a single one using a sum type and representing a random choice
+between them:
 
@@
 
-Then, a composite representation can be transformed back into the concrete
-target type by pattern matching on the data type variant and applying the |step|
-tranformation to the inner |HRep|:
+A composite representation built using $(\oplus)$ is transformed back into the
+target data type by pattern matching on the data type variant and evaluating the
+inner |HRep|:
 %
@@
 \end{alignat*}
+
+Generating a composite |HRep| is slightly more complicated than before, as we
+need to perform a random choice based on the generation size and the given
+frequencies for each sub-represention:
 
@@
 
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
+The previous definition we reflect the type level frequencies of |f| and |g|
+(|freq at_f| and |freq at_g|).
+%
+This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
+used.
+%
+Then we use these frequencies to generate each inner |HRep| in the apropriate
+proportion.
+%
+When the generation size gets exhausted, we reflect the terminal generation
+frequency of each inner |HRep| in the same way as before (|freq0 at_f| and
+|freq0 at_g|).
+%
+This time, however, we default the frequency reflection to $0$ for any inner
+that not tagged as terminal, avoiding to generate non-terminal constructions in
+the last step.
 
 With the introduced combinators, we can easily create a type synonym |HRep_Exp|
-to refer to the canonical |HRep| of our original data type |Exp|, tagging for
-instance the representation of the constuctor |Add| to be generated in with
-double the frequency of rest:
+to specify a generation schema equivalent to the one seen in the concrete random
+generator of type |Exp| presented in Section \ref{sec:randomtesting}:
 
@@
 
+However, a value of type |HRep_Exp| still has its recursive calls abstracted
+away---the type parameter |r| is implicit in the definition of |HRep_Exp|.
+%
+If we want to generate recursive values using this representation we need to
+introduce a last type level combinator to ``tie the knot'' recursively:
+
 \begin{code}
@@
 
+This datatype represents the \emph{fixed point} of a parametric data type |f|,
+i.e., a data type where each recursive call gets instantiated with iself.
+%
+To translate fixed points back to our target data type we simply need translate
+the inner representation:
+%
 \begin{align*}
-  eval : Fix\ rep \rightarrow target
+  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
 \end{align*}
+%
+Categorically, this generic transformation of fixed points is called a
+\emph{catamorphism}, where our evaluation relation is known as its
+\emph{algebra}.
 
-\subsection{\textbf{Representing Pattern Matchings}}
+Unlike the other combinators, to randomly generate a fixed point of a certain
+representation |f|, we do not parametrize the generation of the recursive fields
+of the inner represention over an external generator |gen_r|.
+%
+Instead, we can also tie the recursive knot in our random generator by piping
+the generation of the inner |HRep| parametrized recursively by the itself:
+
+\begin{code}
+genFix_f = Fix <$> gen_f genFix_f
+\end{code} %$
+
+This way we obtain a concrete recursive generator for each representation |f|
+that we define.
+%
+With this generator, we can define a random generator for our target data type
+simply by evaluating a random value of the representation:
+
+\begin{code}
+  gen_Exp = do  x <- genFix_HRep_Exp
+                return eval_x_Exp
+\end{code} %$
+
+\subsection*{\textbf{Representing Pattern Matchings}}
 
@@
 
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
+\begin{code}
+  gen_foo_1  gen_r  = Mk_foo_1  <$> smaller gen_r  <*> smaller gen_r
+  gen_foo_2  gen_r  = Mk_foo_2  <$> arbitrary      <*> smaller gen_r
+\end{code} %$
 
@@
 
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
+\begin{code}
+  gen_ten     gen_r  = Mk_ten
+  gen_square  gen_r  = Mk_square  <$> smaller gen_r
+  gen_minus  gen_r   = Mk_minus   <$> smaller gen_r  <*> smaller gen_r
+\end{code} %$
 
@@
 % \end{code}
+
+% \begin{code}
+% instance (HRep_ten down Exp) where
+%   step Mk_ten = ten
+
+% instance (HRep_square down Exp) where
+%   step (Mk_square x) = square x
+
+% instance (HRep_minus down Exp) where
+%   step (Mk_minus x y) = minus x y
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_ten where
+%   liftArbitrary gen_HRep = pure Mk_ten
+
+% instance Arbitrary1 HRep_square where
+%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
+
+% instance Arbitrary1 HRep_minus where
+%   liftArbitrary gen_HRep
+%     = Mk_minus <$> gen_HRep <*> gen_HRep
+% \end{code} %$
+
+% \begin{code}
+% instance (HRep_foo_1 down Exp) where
+%   step (Mk_foo_1 x y)
+%     = Add (Add x (Val 50)) (Add (Val 25) y)
+
+% instance (HRep_foo_2 down Exp) where
+%   step (Mk_foo_2 x y)
+%     = Mul (Val 50) (Mul (Val x) y)
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 HRep_foo_1 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
+
+% instance Arbitrary1 HRep_foo_2 where
+%   liftArbitrary gen_HRep
+%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
+% \end{code} %$
+
+% \begin{code}
+% instance (f oplus g down Exp) where
+%   step (L f) = step f
+%   step (R g) = step g
+% \end{code}
+
+% \begin{code}
+% instance Arbitrary1 (f oplus g) where
+%   liftArbitrary gen_HRep
+%     = frequency
+%       [ (freq at_f, L <$> gen_HRep)
+%       , (freq at_g, R <$> gen_HRep) ]
+% \end{code} %$

### 2019-01-17 paper/main.tex
WIP

@@
 % Local Variables:
-% TeX-master: "main.lhs.tex"
+% TeX-mbigstarer: "main.lhs.tex"
 % TeX-command-default: "Make"
@@
 %% Title
-\title{Synthesizing Random Generators via Higher-Order Representations}
+\title{Deriving Random Generators via\\ Higher-Level Representations}
 
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{Val}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Val}\;\Conid{Int}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Add}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Add}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Mul}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Mul}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\>[B]{}\mathbf{data}\;{H\!Rep}_{Val}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Val}\;\Conid{Int}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;{H\!Rep}_{Add}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Add}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;{H\!Rep}_{Mul}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Mul}\;\Varid{r}\;\Varid{r}{}\<[E]%
 \ColumnHook
@@
 %
-For instance, the value (\ensuremath{\text{Mk}_{Add}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\text{HRep}_{Add}\;\Conid{Int}}, while the
-value (\ensuremath{\text{Mk}_{Mul}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\text{HRep}_{Mul}\;\Conid{Bool}}.
+For instance, the value (\ensuremath{{Mk}_{Add}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{{H\!Rep}_{Add}\;\Conid{Int}}, while the
+value (\ensuremath{{Mk}_{Mul}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{{H\!Rep}_{Mul}\;\Conid{Bool}}.
 %
@@
 \begin{alignat*}{4}
-  &\evalrep{\ensuremath{\text{Mk}_{Val}}\ n\ \ \ &&}{\ensuremath{\Conid{Exp}}}
+  &\evalrep{\ensuremath{{Mk}_{Val}}\ n\ \ \ &&}{\ensuremath{\Conid{Exp}}}
     &&= \ensuremath{\Conid{Val}}\ n \\
-  &\evalrep{\ensuremath{\text{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
+  &\evalrep{\ensuremath{{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
     &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
-  &\evalrep{\ensuremath{\text{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
+  &\evalrep{\ensuremath{{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
     &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
@@
 \>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Val}{}\<[29]%
+\>[19]{}\mathrel{=}{Mk}_{Val}{}\<[29]%
 \>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
@@
 \>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Add}{}\<[29]%
+\>[19]{}\mathrel{=}{Mk}_{Add}{}\<[29]%
 \>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
@@
 \>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Mul}{}\<[29]%
+\>[19]{}\mathrel{=}{Mk}_{Mul}{}\<[29]%
 \>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
@@
 %
-In first place, we define a type combinator ($\_^\bullet$) to tag a \ensuremath{\text{HRep}} to be
+In first place, we define a type combinator (\ensuremath{\_^\bigstar}) to tag a \ensuremath{\text{HRep}} to be
 terminal, i.e., a representation that is allowed be generated when the
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(f^\bullet)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\>[B]{}\mathbf{data}\;(f^\bigstar)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
 \ColumnHook
@@
   &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Term\ x &&: \ensuremath{f^\bullet}    &&}{t} &&= \evalrep{x : f}{t}
+  &\evalrep{Term\ x &&: \ensuremath{f^\bigstar}    &&}{t} &&= \evalrep{x : f}{t}
 \end{alignat*}
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{f^\bullet}\;{}\<[16]%
+\>[B]{}gen_{f^\bigstar}\;{}\<[16]%
 \>[16]{}gen_{r}{}\<[23]%
@@
 
-With the introduced combinators, we can easily create a type synonym \ensuremath{\text{HRep}_{Exp}}
+With the introduced combinators, we can easily create a type synonym \ensuremath{{H\!Rep}_{Exp}}
 to specify a generation schema equivalent to the one seen in the concrete random
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{Exp}{}\<[16]%
+\>[B]{}\mathbf{type}\;{H\!Rep}_{Exp}{}\<[16]%
 \>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{Val}^\bullet{}\<[E]%
+\>[24]{}{H\!Rep}_{Val}^\bigstar{}\<[E]%
 \\
 \>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{Add}\;{}\<[34]%
+\>[24]{}{H\!Rep}_{Add}\;{}\<[34]%
 \>[34]{}\otimes\;\mathrm{2}{}\<[E]%
@@
 \>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{Mul}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{\text{HRep}_{Exp}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit in the definition of \ensuremath{\text{HRep}_{Exp}}.
+\>[24]{}{H\!Rep}_{Mul}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+However, a value of type \ensuremath{{H\!Rep}_{Exp}} still has its recursive calls abstracted
+away---the type parameter \ensuremath{\Varid{r}} is implicit in the definition of \ensuremath{{H\!Rep}_{Exp}}.
 %
@@
 %
-Instead, we can also tie the recursive knot in our random generator by piping
-the generation of the inner \ensuremath{\text{HRep}} parametrized recursively by the itself:
+Instead, we can tie the recursive knot in our random generator by piping the
+generation of the inner \ensuremath{\text{HRep}}, and calling itself to generate recursive values:
 
@@
 \>[3]{}gen_{Exp}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{(\text{HRep}\,{Exp})}{}\<[E]%
+\>[17]{}\Varid{x}\leftarrow genFix_{\,\text{HRep}_{Exp}}{}\<[E]%
 \\
@@
 
+So far we have introduced the representation machinery required to represent the
+random generation of a target data type considering only the structure encoded
+on its definition.
+%
+This approach has the advantage of being extendable with different sources of
+structured information.
+%
+We proceed to introduce two extensions that help to address the problematic
+testing scenarios presented in the previous section.
+
 \subsection*{\textbf{Representing Pattern Matchings}}
 
+We can follow a similar reasoning as before to represent the pattern matching
+structure from a given function.
+%
+Consider the function \ensuremath{\Varid{foo}} defined in the previous section.
+%
+We can derive a new data type for each one of its pattern matchings, whose
+fields represent each pattern variable present on the given pattern, abstracting
+away every pattern variable of type \ensuremath{\Conid{Exp}}.
+%
+Concretely, we define the following data types:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#1}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
+\>[B]{}\mathbf{data}\;{H\!Rep}_{foo\#1}\;{}\<[18]%
+\>[18]{}\Varid{r}\mathrel{=}{Mk}_{foo\#1}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;{H\!Rep}_{foo\#2}\;{}\<[18]%
+\>[18]{}\Varid{r}\mathrel{=}{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+The first pattern of \ensuremath{\Varid{foo}} conteins two pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type
+\ensuremath{\Conid{Exp}} that are abstracted in the fields of \ensuremath{{Mk}_{foo\#1}}.
+%
+Similarly, the second pattern of \ensuremath{\Varid{foo}} contains the pattern variable \ensuremath{\Varid{x}} of type
+\ensuremath{\Conid{Int}} represented by the first field of \ensuremath{{Mk}_{foo\#2}}, along with the pattern
+variable \ensuremath{\Varid{y}} of type \ensuremath{\Conid{Exp}} that also is abstracted in the second field of
+\ensuremath{{Mk}_{foo\#2}}.
+
+To evaluate these representations back to our target data type, we simply expand
+them it into the concrete value represented by the original pattern:
+%
 \begin{align*}
-  &\evalrep{\ensuremath{\text{Mk}_{foo\#1}}\ x\ y}{Exp} = \\
+  &\evalrep{\ensuremath{{Mk}_{foo\#1}}\ x\ y}{Exp} = \\
   &\ \ \ensuremath{\Conid{Add}\;(\Conid{Add}\;\evalrep{x}{Exp}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\evalrep{y}{Exp})}\\
-  &\evalrep{\ensuremath{\text{Mk}_{foo\#2}}\ x\ y}{Exp} = \\
+  &\evalrep{\ensuremath{{Mk}_{foo\#2}}\ x\ y}{Exp} = \\
   &\ \ \ensuremath{\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\evalrep{y}{Exp})}
 \end{align*}
+
+And the random generation of pattern \ensuremath{\text{HRep}}s is defined in the same way as we
+did before with data constructors \ensuremath{\text{HRep}}s:
 
@@
 \>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}\text{Mk}_{foo\#1}{}\<[33]%
+\>[21]{}\mathrel{=}{Mk}_{foo\#1}{}\<[33]%
 \>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[52]%
@@
 \>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}\text{Mk}_{foo\#2}{}\<[33]%
+\>[21]{}\mathrel{=}{Mk}_{foo\#2}{}\<[33]%
 \>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[52]%
@@
 
+Finally, we can join the pattern matching representation of each clause of \ensuremath{\Varid{foo}}
+into a single one:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
-\>[16]{}\mathrel{=}\text{HRep}_{foo\#1}\;\!\!\oplus\!\!\;{}\<[38]%
-\>[38]{}\text{HRep}_{foo\#2}{}\<[E]%
+\>[B]{}\mathbf{type}\;{H\!Rep}_{foo}{}\<[16]%
+\>[16]{}\mathrel{=}{H\!Rep}_{foo\#1}\;\!\!\oplus\!\!\;{}\<[38]%
+\>[38]{}{H\!Rep}_{foo\#2}{}\<[E]%
 \ColumnHook
@@
 \subsection*{\textbf{Representing Abstract Interfaces}}
+
+To introduce the higher level representation of module abstract interfaces,
+suppose we have a module \ensuremath{\Conid{M}} defining \ensuremath{\Conid{Exp}} combinators as follows:
 
@@
 
+We will represent each function of \ensuremath{\Conid{M}} returning a value of type \ensuremath{\Conid{Exp}} following
+the same idea as before, deriving a data type with a single data constructor for
+each one of them:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{ten}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{square}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{square}\;{}\<[37]%
+\>[B]{}\mathbf{data}\;{H\!Rep}_{ten}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}{Mk}_{ten}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;{H\!Rep}_{square}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}{Mk}_{square}\;{}\<[37]%
 \>[37]{}\Varid{r}{}\<[E]%
 \\
-\>[B]{}\mathbf{data}\;\text{HRep}_{minus}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{minus}\;{}\<[37]%
+\>[B]{}\mathbf{data}\;{H\!Rep}_{minus}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}{Mk}_{minus}\;{}\<[37]%
 \>[37]{}\Varid{r}\;\Varid{r}{}\<[E]%
@@
 
+In this case, each single constructor will have as fields the types of the
+inputs of the function that they represent.
+%
+As before, we abstract away any field representing and input of type \ensuremath{\Conid{Exp}} with
+a type parameter \ensuremath{\Varid{r}}.
+
+To evaluate these representations, we simply pipe the values on its fields as
+input parameter of each original function, returning its result.
+
 \begin{alignat*}{3}
-  &\evalrep{\ensuremath{\text{Mk}_{ten}}         &&}{Exp} = \ensuremath{\Varid{ten}} \\
-  &\evalrep{\ensuremath{\text{Mk}_{square}}\ x   &&}{Exp} = \ensuremath{\Varid{square}\;\evalrep{x}{Exp}} \\
-  &\evalrep{\ensuremath{\text{Mk}_{minus}}\ x\ y &&}{Exp} = \ensuremath{\Varid{minus}\;\evalrep{x}{Exp}\;\evalrep{y}{Exp}}
+  &\evalrep{\ensuremath{{Mk}_{ten}}         &&}{Exp} = \ensuremath{\Varid{ten}} \\
+  &\evalrep{\ensuremath{{Mk}_{square}}\ x   &&}{Exp} = \ensuremath{\Varid{square}\;\evalrep{x}{Exp}} \\
+  &\evalrep{\ensuremath{{Mk}_{minus}}\ x\ y &&}{Exp} = \ensuremath{\Varid{minus}\;\evalrep{x}{Exp}\;\evalrep{y}{Exp}}
 \end{alignat*}
+
+Moreover, the generation procedure for abstract interface representations
+follows the same pattern as before:
 
@@
 \>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
+\>[22]{}\mathrel{=}{Mk}_{ten}{}\<[E]%
 \\
@@
 \>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{square}{}\<[35]%
+\>[22]{}\mathrel{=}{Mk}_{square}{}\<[35]%
 \>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
@@
 \>[14]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{minus}{}\<[35]%
+\>[22]{}\mathrel{=}{Mk}_{minus}{}\<[35]%
 \>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[54]%
@@
  %$
+
+We will also define a type synonym to join all the representations of the module
+\ensuremath{\Conid{M}} into a single one.
 
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{M}{}\<[14]%
-\>[14]{}\mathrel{=}\text{HRep}_{ten}\;\!\!\oplus\!\!\;{}\<[34]%
-\>[34]{}\text{HRep}_{square}\;\!\!\oplus\!\!\;{}\<[55]%
-\>[55]{}\text{HRep}_{minus}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
+\>[B]{}\mathbf{type}\;{H\!Rep}_{M}{}\<[14]%
+\>[14]{}\mathrel{=}{H\!Rep}_{ten}\;\!\!\oplus\!\!\;{}\<[34]%
+\>[34]{}{H\!Rep}_{square}\;\!\!\oplus\!\!\;{}\<[55]%
+\>[55]{}{H\!Rep}_{minus}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Finally, we can put all the derived machinery together into a generation
+specification \ensuremath{\Conid{Spec}} for the type \ensuremath{\Conid{Exp}}, assigning (possibly) different
+generation frequencies to each individual \ensuremath{\text{HRep}} that constitutes it:
 
@@
 \>[12]{}\mathrel{=}{}\<[20]%
-\>[20]{}\text{HRep}_{Exp}\;{}\<[30]%
+\>[20]{}{H\!Rep}_{Exp}\;{}\<[30]%
 \>[30]{}\otimes\;\mathrm{4}{}\<[E]%
@@
 \>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}\text{HRep}_{foo}\;{}\<[30]%
+\>[20]{}{H\!Rep}_{foo}\;{}\<[30]%
 \>[30]{}\otimes\;\mathrm{2}{}\<[E]%
@@
 \>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}\text{HRep}_{M}{}\<[E]%
+\>[20]{}{H\!Rep}_{M}{}\<[E]%
 \ColumnHook
@@
 This previous definition can be interpreted graphically as it is shown in the
-Figure \ref{fig:hrep}.
-%
-Curly arrows represent the structural information extracted using
-meta-programming.
+Figure \ref{fig:hrep}, where curly arrows represent the structural information
+extracted using meta-programming.
 
@@
 \end{figure}
+
+This representation can be extended to mutually recursive and parametric types
+...
 

### 2019-01-17 paper/sections/hrep.tex
WIP

@@
 %
-In first place, we define a type combinator ($\_^\bullet$) to tag a |HRep| to be
+In first place, we define a type combinator (|term|) to tag a |HRep| to be
 terminal, i.e., a representation that is allowed be generated when the
@@
 %
-Instead, we can also tie the recursive knot in our random generator by piping
-the generation of the inner |HRep| parametrized recursively by the itself:
+Instead, we can tie the recursive knot in our random generator by piping the
+generation of the inner |HRep|, and calling itself to generate recursive values:
 
@@
 
+So far we have introduced the representation machinery required to represent the
+random generation of a target data type considering only the structure encoded
+on its definition.
+%
+This approach has the advantage of being extendable with different sources of
+structured information.
+%
+We proceed to introduce two extensions that help to address the problematic
+testing scenarios presented in the previous section.
+
 \subsection*{\textbf{Representing Pattern Matchings}}
+
+We can follow a similar reasoning as before to represent the pattern matching
+structure from a given function.
+%
+Consider the function |foo| defined in the previous section.
+%
+We can derive a new data type for each one of its pattern matchings, whose
+fields represent each pattern variable present on the given pattern, abstracting
+away every pattern variable of type |Exp|.
+%
+Concretely, we define the following data types:
 
@@
 
+The first pattern of |foo| conteins two pattern variables (|x| and |y|) of type
+|Exp| that are abstracted in the fields of |Mk_foo_1|.
+%
+Similarly, the second pattern of |foo| contains the pattern variable |x| of type
+|Int| represented by the first field of |Mk_foo_2|, along with the pattern
+variable |y| of type |Exp| that also is abstracted in the second field of
+|Mk_foo_2|.
+
+To evaluate these representations back to our target data type, we simply expand
+them it into the concrete value represented by the original pattern:
+%
 \begin{align*}
@@
 
+And the random generation of pattern |HRep|s is defined in the same way as we
+did before with data constructors |HRep|s:
+
 \begin{code}
@@
 \end{code} %$
+
+Finally, we can join the pattern matching representation of each clause of |foo|
+into a single one:
 
@@
 \subsection*{\textbf{Representing Abstract Interfaces}}
+
+To introduce the higher level representation of module abstract interfaces,
+suppose we have a module |M| defining |Exp| combinators as follows:
 
@@
 
+We will represent each function of |M| returning a value of type |Exp| following
+the same idea as before, deriving a data type with a single data constructor for
+each one of them:
+
 \begin{code}
@@
 \end{code}
+
+In this case, each single constructor will have as fields the types of the
+inputs of the function that they represent.
+%
+As before, we abstract away any field representing and input of type |Exp| with
+a type parameter |r|.
+
+To evaluate these representations, we simply pipe the values on its fields as
+input parameter of each original function, returning its result.
 
@@
 
+Moreover, the generation procedure for abstract interface representations
+follows the same pattern as before:
+
 \begin{code}
@@
 
+We will also define a type synonym to join all the representations of the module
+|M| into a single one.
+
 \begin{code}
@@
 \end{code}
+
+Finally, we can put all the derived machinery together into a generation
+specification |Spec| for the type |Exp|, assigning (possibly) different
+generation frequencies to each individual |HRep| that constitutes it:
 

### 2019-01-17 paper/tikz/hrep.tex
WIP

@@
 
+\hspace*{-4mm}
 \begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.9,xscale=0.9]
@@
 %Shape: Rectangle [id:dp9489660675592899]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (110,20) ..
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]  (110,20) ..
 controls (110,14.48) and (114.48,10) ..
@@
 %Straight Lines [id:da7051587981224026]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (270,250) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (270,250) ..
 controls (268.33,248.33) and (268.33,246.67) ..
@@
 (270,205) -- (270,201) -- (270,193) ;
-\draw [shift={(270,190)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+\draw [shift={(270,190)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
 controls (9.04,-1.82) and (4.3,-0.39) ..
@@
 %Straight Lines [id:da6087008244886778]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (380,250) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (380,250) ..
 controls (378.33,248.33) and (378.33,246.67) ..
@@
 (380,215) -- (380,211) -- (380,203) ;
-\draw [shift={(380,200)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+\draw [shift={(380,200)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
 controls (9.04,-1.82) and (4.3,-0.39) ..
@@
 %Shape: Rectangle [id:dp701556783783656]
-\draw  [line width=1.5]  (230,255) ..
+\draw  [line width=1]  (230,255) ..
 controls (230,252.24) and (232.24,250) ..
@@
 %Shape: Rectangle [id:dp033430127809365384]
-\draw  [line width=1.5]  (340,255) ..
+\draw  [line width=1]  (340,255) ..
 controls (340,252.24) and (342.24,250) ..
@@
 %Up Arrow [id:dp0013771056350821986]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,289.5) -- (270,280) -- (280,289.5) -- (273.5,289.5) -- (273.5,310) -- (266.5,310) -- (266.5,289.5) -- cycle ;
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1]  (260,289.5) -- (270,280) -- (280,289.5) -- (273.5,289.5) -- (273.5,310) -- (266.5,310) -- (266.5,289.5) -- cycle ;
 %Shape: Rectangle [id:dp1692119124811613]
-\draw  [line width=1.5]  (110,310) -- (430,310) -- (430,340) -- (110,340) -- cycle ;
+\draw  [line width=1]  (110,310) -- (430,310) -- (430,340) -- (110,340) -- cycle ;
 %Up Arrow [id:dp4999372801455]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (370,289.5) -- (380,280) -- (390,289.5) -- (383.5,289.5) -- (383.5,310) -- (376.5,310) -- (376.5,289.5) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (223,132.93) ..
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1]  (370,289.5) -- (380,280) -- (390,289.5) -- (383.5,289.5) -- (383.5,310) -- (376.5,310) -- (376.5,289.5) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (223,132.93) ..
 controls (223,129.02) and (219.83,125.86) ..
@@
 controls (219.83,140) and (223,136.83) ..
-(223,132.93) -- cycle ; \draw  [line width=1.5]  (223,132.93) -- (208.86,132.93) ; \draw  [line width=1.5]  (215.93,125.86) -- (215.93,140) ;
+(223,132.93) -- cycle ; \draw  [line width=1]  (223,132.93) -- (208.86,132.93) ; \draw  [line width=1]  (215.93,125.86) -- (215.93,140) ;
 %Snip Same Side Corner Rect [id:dp11914646356873848]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (230,67.5) -- (237,60.5) -- (303,60.5) -- (310,67.5) -- (310,80.5) -- (310,80.5) -- (230,80.5) -- (230,80.5) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (230,67.5) -- (237,60.5) -- (303,60.5) -- (310,67.5) -- (310,80.5) -- (310,80.5) -- (230,80.5) -- (230,80.5) -- cycle ;
 %Shape: Rectangle [id:dp6865365415345861]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (230,80.5) -- (310,80.5) -- (310,190.5) -- (230,190.5) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (230,80.5) -- (310,80.5) -- (310,190.5) -- (230,190.5) -- cycle ;
 %Shape: Rectangle [id:dp39543698275133865]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (236.29,107.66) -- (303.21,107.66) -- (303.21,130.34) -- (236.29,130.34) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (247,75.5) ..
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (236.29,107.66) -- (303.21,107.66) -- (303.21,130.34) -- (236.29,130.34) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (247,75.5) ..
 controls (249.76,72.74) and (249.76,68.26) ..
@@
 controls (239.76,78.26) and (244.24,78.26) ..
-(247,75.5) -- cycle ; \draw  [line width=1.5]  (247,75.5) -- (237,65.5) ; \draw  [line width=1.5]  (247,65.5) -- (237,75.5) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (277.14,143.43) ..
+(247,75.5) -- cycle ; \draw  [line width=1]  (247,75.5) -- (237,65.5) ; \draw  [line width=1]  (247,65.5) -- (237,75.5) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (277.14,143.43) ..
 controls (277.14,139.52) and (273.98,136.36) ..
@@
 controls (273.98,150.5) and (277.14,147.33) ..
-(277.14,143.43) -- cycle ; \draw  [line width=1.5]  (277.14,143.43) -- (263,143.43) ; \draw  [line width=1.5]  (270.07,136.36) -- (270.07,150.5) ;
+(277.14,143.43) -- cycle ; \draw  [line width=1]  (277.14,143.43) -- (263,143.43) ; \draw  [line width=1]  (270.07,136.36) -- (270.07,150.5) ;
 %Shape: Rectangle [id:dp047421885346712545]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (236.08,157.66) -- (303,157.66) -- (303,180.34) -- (236.08,180.34) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (236.08,157.66) -- (303,157.66) -- (303,180.34) -- (236.08,180.34) -- cycle ;
 %Shape: Rectangle [id:dp0425475572348164]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (340,40) -- (420,40) -- (420,200) -- (340,200) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (340,40) -- (420,40) -- (420,200) -- (340,200) -- cycle ;
 %Shape: Rectangle [id:dp9247208762532781]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (345,66.37) -- (415,66.37) -- (415,89) -- (345,89) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (345,66.37) -- (415,66.37) -- (415,89) -- (345,89) -- cycle ;
 %Shape: Rectangle [id:dp8712965895636344]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (345.52,117.37) -- (414.48,117.37) -- (414.48,140) -- (345.52,140) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (345.52,117.37) -- (414.48,117.37) -- (414.48,140) -- (345.52,140) -- cycle ;
 %Shape: Rectangle [id:dp933766425327347]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (346.66,168.24) -- (413.34,168.24) -- (413.34,190.87) -- (346.66,190.87) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (387.14,103.95) ..
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (346.66,168.24) -- (413.34,168.24) -- (413.34,190.87) -- (346.66,190.87) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (387.14,103.95) ..
 controls (387.14,100.05) and (383.98,96.89) ..
@@
 controls (383.98,111) and (387.14,107.84) ..
-(387.14,103.95) -- cycle ; \draw  [line width=1.5]  (387.14,103.95) -- (373,103.95) ; \draw  [line width=1.5]  (380.07,96.89) -- (380.07,111) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (387.14,152.95) ..
+(387.14,103.95) -- cycle ; \draw  [line width=1]  (387.14,103.95) -- (373,103.95) ; \draw  [line width=1]  (380.07,96.89) -- (380.07,111) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (387.14,152.95) ..
 controls (387.14,149.05) and (383.98,145.89) ..
@@
 controls (383.98,160) and (387.14,156.84) ..
-(387.14,152.95) -- cycle ; \draw  [line width=1.5]  (387.14,152.95) -- (373,152.95) ; \draw  [line width=1.5]  (380.07,145.89) -- (380.07,160) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (333,132.93) ..
+(387.14,152.95) -- cycle ; \draw  [line width=1]  (387.14,152.95) -- (373,152.95) ; \draw  [line width=1]  (380.07,145.89) -- (380.07,160) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (333,132.93) ..
 controls (333,129.02) and (329.83,125.86) ..
@@
 controls (329.83,140) and (333,136.83) ..
-(333,132.93) -- cycle ; \draw  [line width=1.5]  (333,132.93) -- (318.86,132.93) ; \draw  [line width=1.5]  (325.93,125.86) -- (325.93,140) ;
+(333,132.93) -- cycle ; \draw  [line width=1]  (333,132.93) -- (318.86,132.93) ; \draw  [line width=1]  (325.93,125.86) -- (325.93,140) ;
 %Snip Same Side Corner Rect [id:dp09931070350524518]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (120,27.48) -- (126.98,20.5) -- (193.02,20.5) -- (200,27.48) -- (200,40.45) -- (200,40.45) -- (120,40.45) -- (120,40.45) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (120,27.48) -- (126.98,20.5) -- (193.02,20.5) -- (200,27.48) -- (200,40.45) -- (200,40.45) -- (120,40.45) -- (120,40.45) -- cycle ;
 %Shape: Rectangle [id:dp6050538306488522]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (120,40.45) -- (200,40.45) -- (200,220) -- (120,220) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (120,40.45) -- (200,40.45) -- (200,220) -- (120,220) -- cycle ;
 %Shape: Rectangle [id:dp20494666512059623]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,67.54) -- (190,67.54) -- (190,90.17) -- (130,90.17) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,67.54) -- (190,67.54) -- (190,90.17) -- (130,90.17) -- cycle ;
 %Shape: Rectangle [id:dp8741156804316592]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,137.36) -- (190,137.36) -- (190,159.99) -- (130,159.99) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,137.36) -- (190,137.36) -- (190,159.99) -- (130,159.99) -- cycle ;
 %Shape: Rectangle [id:dp7972859165129229]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,187.24) -- (190,187.24) -- (190,209.87) -- (130,209.87) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,187.24) -- (190,187.24) -- (190,209.87) -- (130,209.87) -- cycle ;
 %Snip Same Side Corner Rect [id:dp44067415974317603]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (130,124.4) -- (136.98,117.41) -- (183.02,117.41) -- (190,124.4) -- (190,137.36) -- (190,137.36) -- (130,137.36) -- (130,137.36) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (147,132.22) ..
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,124.4) -- (136.98,117.41) -- (183.02,117.41) -- (190,124.4) -- (190,137.36) -- (190,137.36) -- (130,137.36) -- (130,137.36) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (147,132.22) ..
 controls (149.76,129.46) and (149.76,125) ..
@@
 controls (139.76,134.97) and (144.24,134.97) ..
-(147,132.22) -- cycle ; \draw  [line width=1.5]  (147,132.22) -- (137,122.24) ; \draw  [line width=1.5]  (147,122.24) -- (137,132.22) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (137,35.46) ..
+(147,132.22) -- cycle ; \draw  [line width=1]  (147,132.22) -- (137,122.24) ; \draw  [line width=1]  (147,122.24) -- (137,132.22) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (137,35.46) ..
 controls (139.76,32.71) and (139.76,28.24) ..
@@
 controls (129.76,38.22) and (134.24,38.22) ..
-(137,35.46) -- cycle ; \draw  [line width=1.5]  (137,35.46) -- (127,25.49) ; \draw  [line width=1.5]  (137,25.49) -- (127,35.46) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (167.14,103.22) ..
+(137,35.46) -- cycle ; \draw  [line width=1]  (137,35.46) -- (127,25.49) ; \draw  [line width=1]  (137,25.49) -- (127,35.46) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (167.14,103.22) ..
 controls (167.14,99.33) and (163.98,96.17) ..
@@
 controls (163.98,110.27) and (167.14,107.12) ..
-(167.14,103.22) -- cycle ; \draw  [line width=1.5]  (167.14,103.22) -- (153,103.22) ; \draw  [line width=1.5]  (160.07,96.17) -- (160.07,110.27) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (167.14,173.05) ..
+(167.14,103.22) -- cycle ; \draw  [line width=1]  (167.14,103.22) -- (153,103.22) ; \draw  [line width=1]  (160.07,96.17) -- (160.07,110.27) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (167.14,173.05) ..
 controls (167.14,169.15) and (163.98,165.99) ..
@@
 controls (163.98,180.1) and (167.14,176.94) ..
-(167.14,173.05) -- cycle ; \draw  [line width=1.5]  (167.14,173.05) -- (153,173.05) ; \draw  [line width=1.5]  (160.07,165.99) -- (160.07,180.1) ;
+(167.14,173.05) -- cycle ; \draw  [line width=1]  (167.14,173.05) -- (153,173.05) ; \draw  [line width=1]  (160.07,165.99) -- (160.07,180.1) ;
 %Down Arrow [id:dp5984296172784]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (160,299.5) -- (165,299.5) -- (165,230) -- (175,230) -- (175,299.5) -- (180,299.5) -- (170,310) -- cycle ;
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]  (160,299.5) -- (165,299.5) -- (165,230) -- (175,230) -- (175,299.5) -- (180,299.5) -- (170,310) -- cycle ;
 %Straight Lines [id:da30879479079788386]
-\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]    (166.25,230) -- (173.75,230) ;
+\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]    (166.25,230) -- (173.75,230) ;
 
 %Straight Lines [id:da8667104419281686]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (190,310) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (190,310) ..
 controls (188.33,308.33) and (188.33,306.67) ..
@@
 (190,235) -- (190,231) -- (190,223) ;
-\draw [shift={(190,220)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
+\draw [shift={(190,220)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
 controls (9.04,-1.82) and (4.3,-0.39) ..
@@
 % Text Node
-\draw (148,51.5) node [scale=0.9]  {$\text{HRep}_{Exp}$};
+\draw (148,51.5) node [scale=0.9]  {${H\!Rep}_{Exp}$};
 % Text Node
-\draw (273,27.5) node [scale=1]  {$Spec_{Exp}$};
+\draw (270,27.5) node [scale=1]  {$Spec$};
 % Text Node
-\draw (160,78.85) node [scale=0.9]  {$\text{HRep}_{Val}$};
+\draw (160,78.85) node [scale=0.9]  {${H\!Rep}_{Val}^\bigstar$};
 % Text Node
-\draw (160,148.68) node [scale=0.9]  {$\text{HRep}_{Add}$};
+\draw (160,148.68) node [scale=0.9]  {${H\!Rep}_{Add}$};
 % Text Node
-\draw (160,198.55) node [scale=0.9]  {$\text{HRep}_{Mul}$};
+\draw (160,198.55) node [scale=0.9]  {${H\!Rep}_{Mul}$};
 % Text Node
@@
 % Text Node
-\draw (257,91.5) node [scale=0.9]  {$\text{HRep}_{foo}$};
+\draw (257,91.5) node [scale=0.9]  {${H\!Rep}_{foo}$};
 % Text Node
@@
 % Text Node
-\draw (269.75,119) node [scale=0.9]  {$\text{HRep}_{foo\#1}$};
+\draw (269.75,119) node [scale=0.9]  {${H\!Rep}_{foo\#1}$};
 % Text Node
-\draw (269.54,169) node [scale=0.9]  {$\text{HRep}_{foo\#2}$};
+\draw (269.54,169) node [scale=0.9]  {${H\!Rep}_{foo\#2}$};
 % Text Node
-\draw (115,269.5) node [scale=0.9]  {$Spec_{Exp}\! \Downarrow \! Exp$};
+\draw (120,269.5) node [scale=0.9] {$\llbracket Fix\ Spec \rrbracket_{Exp}$};
 % Text Node
-\draw (365.5,51.5) node [scale=0.9]  {$\text{HRep}_{M}$};
+\draw (365.5,51.5) node [scale=0.9]  {${H\!Rep}_{M}$};
 % Text Node
-\draw (380,179.55) node [scale=0.9]  {$\text{HRep}_{minus}$};
+\draw (380,179.55) node [scale=0.9]  {${H\!Rep}_{minus}$};
 % Text Node
-\draw (380,128.69) node [scale=0.9]  {$\text{HRep}_{square}$};
+\draw (380,128.69) node [scale=0.9]  {${H\!Rep}_{square}$};
 % Text Node
-\draw (380,77.69) node [scale=0.9]  {$\text{HRep}_{ten}$};
+\draw (380,77.69) node [scale=0.9]  {${H\!Rep}_{ten}$};
 
 \end{tikzpicture}
+\hspace*{4mm}

### 2019-01-20 paper/main.tex
WIP

--- 2019-01-17 paper/main.tex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% HRep
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Constructor names
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% gen
+
 % Local Variables:
-% TeX-master: "main.lhs.tex"
+% TeX-mbigstarer: "main.lhs.tex"
 % TeX-command-default: "Make"
@@
 
-\newcommand{\evalrep}[2]{\ensuremath{\llbracket #1 \rrbracket_{#2}}}
+\newcommand{\hrep}[1]{\ensuremath{H\!Rep_{#1}}}
+\newcommand{\gen}[1]{\ensuremath{gen_{#1}}}
+\newcommand{\term}[1]{\ensuremath{#1^\bigstar}}
+\newcommand{\evalrep}[2]{ \ensuremath{\llbracket #1 \rrbracket_{#2}}}
 
@@
 %% Some parameters
-\setlength{\mathindent}{1.5\parindent}
+\setlength{\mathindent}{\parindent}
 
@@
 
+We remark that, altough this work focuses on Haskell data types, this technique
+is general enough to be applied to most programming languages with some level of
+support for composite types.
+
 The main contribuitions of this paper are:
@@
 \end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
+\section{Random Data Generation in Haskell} \label{sec:randomtesting}
+
+In this section we briefly introduce the common approach for automatically
+deriving random data generators in Haskell using a type-driven approach, along
+with its main drawbacks.
 
@@
 It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
+structure of their systems using data types that can be checked at compilation
 time.
@@
 %
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} (or
-constructors for short) and are used to distiguish which variant of the data
-type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
+Essentially, an ADTs is a composite type defined by combining other types in
+terms of \textbf{sums} (also known as \emph{variant types}) and
+\textbf{products} (or tuples) of other data types.
+%
+To exemplify this, consider the following type definition to encoding \ensuremath{\Conid{Html}}
+expressions:
+%
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\Conid{Html}{}\<[12]%
+\>[12]{}\mathrel{=}{}\<[16]%
+\>[16]{}\Conid{Text}\;{}\<[22]%
+\>[22]{}\Conid{String}{}\<[E]%
+\\
+\>[12]{}\,\ \vert\;{}\<[16]%
+\>[16]{}\Conid{Sing}\;{}\<[22]%
+\>[22]{}\Conid{String}{}\<[E]%
+\\
+\>[12]{}\,\ \vert\;{}\<[16]%
+\>[16]{}\Conid{Join}\;{}\<[22]%
+\>[22]{}\Conid{Html}\;{}\<[30]%
+\>[30]{}\Conid{Html}{}\<[E]%
+\\
+\>[12]{}\,\ \vert\;{}\<[16]%
+\>[16]{}\Conid{Tag}\;{}\<[22]%
+\>[22]{}\Conid{String}\;{}\<[30]%
+\>[30]{}\Conid{Html}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+In the previous definition, we declare \ensuremath{\Conid{Html}} as the sum of four possible
+constructions: \ensuremath{\Conid{Text}} represents plain text values. \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Tag}} represent
+singular and paired Html tags, and \ensuremath{\Conid{Join}} concats two expression one after
+another.
+%
+We only encode a very small subset of the actual Html specification for
+illustrative reasons.
+%
+In Haskell, \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}} are known as \emph{data
+  constructors} (or constructors for short) and are used to distiguish which
+variant of the data type we are constructing.
+%
+Each data constructor is defined as a product of zero or more types known as
+\emph{fields}.
+%
+For instance, \ensuremath{\Conid{Text}} has a field of type \ensuremath{\Conid{String}}, whereas \ensuremath{\Conid{Join}} two recusive
+fields of type \ensuremath{\Conid{Html}}.
 %
@@
 %
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
+With this reprentation, the expression ``\ensuremath{<\!\!html\!\!>\!\!hello\!\!<\!\!hr\!\!>\!\!bye\!\!<\!\!/html\!\!>}'' can be encoded as:
+%
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Conid{Tag}\;\text{\ttfamily \char34 html\char34}\;(\Conid{Join}\;(\Conid{Join}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}(\Conid{Text}\;\text{\ttfamily \char34 hello\char34})\;(\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}))\;(\Conid{Text}\;\text{\ttfamily \char34 bye\char34})){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+%
+Additionally, we can define a function \ensuremath{\Varid{render}} to serialize \ensuremath{\Conid{Html}} values as
+follows:
+%
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{render}\mathbin{::}\Conid{Html}\to \Conid{String}{}\<[E]%
+\\
+\>[B]{}\Varid{render}\;(\Conid{Text}\;\Varid{t})\mathrel{=}\Varid{t}{}\<[E]%
+\\
+\>[B]{}\Varid{render}\;(\Conid{Sing}\;\Varid{t})\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
+\\
+\>[B]{}\Varid{render}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{render}\;\Varid{x}\plus \Varid{render}\;\Varid{y}{}\<[E]%
+\\
+\>[B]{}\Varid{render}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x}){}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}\plus \Varid{render}\;\Varid{x}\plus \text{\ttfamily \char34 </\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+In the previous definition, \ensuremath{\Varid{render}} is described using \emph{pattern matching}
 over each possible kind of value.
 %
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
+Using pattern matching we can define functions idiomatically by defining
+different function clauses for each input pattern we are interested on.
+%
+Patterns can be defined to match specific constructors, literal values or
+variable subexpressions (like \ensuremath{\Varid{t}}, \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} in the definition of \ensuremath{\Varid{render}}).
+%
+They can also be nested in order to match very specific patterns of values.
 
@@
 
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
+In order to generate random value of types involving user defined ADTs, most
+approaches require the user to provide a random data generator for each one of
+them.
+%
+This is a cumbersome and error prone task that closely follows the data type
+structure.
 %
 For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
+generator for the type \ensuremath{\Conid{Html}}:
+%
 \begin{hscode}\SaveRestoreHook
@@
 \column{9}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{Exp}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
+\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}gen_{Html}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
 \\
@@
 \>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{then}\;\Conid{Val}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
+\>[4]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{}[\mskip1.5mu {}\<[6E]%
+\>[9]{}(\mathrm{2},{}\<[14]%
+\>[14]{}\Conid{Text}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{},{}\<[6E]%
+\>[9]{}(\mathrm{1},{}\<[14]%
+\>[14]{}\Conid{Sing}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary})\mskip1.5mu]{}\<[E]%
 \\
@@
 \>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{1},\Conid{Val}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
+\>[9]{}(\mathrm{2},{}\<[14]%
+\>[14]{}\Conid{Text}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
 \\
@@
 \>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{2},\Conid{Add}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp}){}\<[E]%
+\>[9]{}(\mathrm{1},{}\<[14]%
+\>[14]{}\Conid{Sing}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
 \\
@@
 \>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},\Conid{Mul}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp})\mskip1.5mu]){}\<[E]%
+\>[9]{}(\mathrm{3},{}\<[14]%
+\>[14]{}\Conid{Join}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Html}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html}){}\<[E]%
+\\
+\>[4]{}\hsindent{2}{}\<[6]%
+\>[6]{},{}\<[6E]%
+\>[9]{}(\mathrm{4},{}\<[14]%
+\>[14]{}\Conid{Tag}{}\<[20]%
+\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html})\mskip1.5mu]){}\<[E]%
 \ColumnHook
@@
  %$
-
+%
 This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
-paremetrize the generation process up to an external integer number known as the
+paremetrize the generation process up to an external natural number known as the
 \emph{generation size}.
 %
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator will
-pick among any data constructor of \ensuremath{\Conid{Exp}} with a explicitly given frequency---two
-times more \ensuremath{\Conid{Add}}s than \ensuremath{\Conid{Val}} or \ensuremath{\Conid{Mul}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively
-(\ensuremath{\Varid{resize}\;(n{-}1)\;gen_{Exp}}), decreasing the generation size on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
+This parameter is chosen by the user, and we use to limit the maximum amount of
+recursive calls that this random generator can perform.
+%
+When called with a positive generation size, this generator can pick to generate
+among any \ensuremath{\Conid{Html}} data constructor of with a explicitly given frequency that can
+be chosen by the user (2, 1, 3 and 4 for \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}}, \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}}
+constructors, respectively).
+%
+When it picks to generate a \ensuremath{\Conid{Val}} or a \ensuremath{\Conid{Sing}} data constructor, it also
+generates a random \ensuremath{\Conid{String}} value using the standard overloaded generator
+\ensuremath{\Varid{arbitrary}} (\quickcheck provides standard random generators for most base types
+like \ensuremath{\Conid{String}}, \ensuremath{\Conid{Int}}, \ensuremath{\Conid{Bool}}, etc.). \footnote{%
+  The operators \ensuremath{\mathop{\mathsmaller{\langle \$ \rangle}}} and \ensuremath{\mathop{\mathsmaller{\langle \ast \rangle}}} are used in Haskell to combine values obtained
+  from effectful computations (like calling to a random generator) and they are
+  not particularly relevant for the point being made in this work.}
+%$
+On the other hand, when it picks to generate either an \ensuremath{\Conid{Join}} constructor, it
+also generates two independent random subexpression recursively, decreasing the
+generation size by a unit on each recursive invocation (\ensuremath{\Varid{smaller}\;gen_{Html}}).
+%
+The case of random generation of \ensuremath{\Conid{Tag}} constructors follows analogously.
+
+This random process keeps calling itself recursively until the generation size
+reaches zero, where the generator is constrained to pick only among terminal
+data constructors, being \ensuremath{\Conid{Text}} and \ensuremath{\Conid{Sing}} the only possible choices in our
 particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
+
+% Strictly decreasing the the generation size by one on each recursive call
+% results in a useful property over the generated data: every generated value has
+% at most |n| levels, where |n| is the generation size set by the user.
+%
+% This property enables us to model the generation process using the theory of
+% branching processes introduced by \tocite, and extended in this work as
+% described in Section \ref{sec:synthesis}.
+
+The previous definition is rather mechanical, except perhaps for the chosen
+generation frequencies.
+%
+In this light, it easy to extend this procedure to any data type defined in an
+algebraic fashion.
 %
@@
 %
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
+The simplest tool for such purpose is \megadeth \tocite.
+%
+Given a target data type, it synthesizes a random generator for it that behaves
+similarly to the one presented above, where the generation frequencies are
+defined to be uniform across constructors.
+%
+However, picking among different data constructors with uniform frequency can
+lead to a generation process biased towards generating (in average) very small
+values, regardless of the generation sized set by the user \tocite.
 
@@
 Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
+tool this tool uses the theory of branching processes to modelize and predict
+analitically the average distribution of data constructors generated on each
+random value.
 %
 This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
+process that adjusts the generation frequency of each data constructor in order
+to obtain a particular distribution of values that can be specified by the user,
+providing this way a more flexible testing environment while still being mostly
 automated.
 
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
+Althogh both \megadeth and \dragen synthesize random generators that are
+theoretically capable to generate the whole space of values of the target data
+type. the limitations arise quickly when we consider that the underlying
 generation model is essentially the same: they pick a single data constructor
@@
 %
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of additional structure information which
-are not considered by the aforementioned automatic derivation tools to obtain
-random data generators:
+In practice, this procedure is often too generic to generate random data with
+enough structural complexity to be used for testing purposes.
+
+% In this work we identify two sources of additional structure information which
+% are not considered by the aforementioned automatic derivation tools to obtain
+% better random data generators:
+
+In this work we identify two patological situations which are not properly
+handled by the aforementioned derivation tools:
 
 \begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code yields part of the responsability of preserving its
-  invariants to the functions of its abstract interface.
+\item The target code behaves differently over inputs matching specific patterns
+  of nested values.
+\item The target code encodes a significant portion of its structure on its
+  abstract interface.
 \end{enumerate}
 
-This information can be reified in compile time and used to synthesize richer
-random generators automatically. We proceed to exemplify the previous points in
-detail.
+In Section \ref{sec:hrep} we show how this structural information can be used to
+synthesize richer random generators automatically.
+%
+We proceed to exemplify the previous points in detail.
 
@@
 
-To exemplify the firts problematic scenarion that may arise, suppose we want to
-use randomly generated \ensuremath{\Conid{Exp}}s to test a property comprising the following
-function:
-
+To exemplify the first problematic scenario, suppose we want to use randomly
+generated \ensuremath{\Conid{Htmls}}s to test a property comprising the following function:
+%
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
+\>[B]{}\Varid{simpl}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[E]%
 \\
 \>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\Varid{y})){}\<[E]%
+\>[3]{}\mathrel{=}\Conid{Text}\;(\Varid{t1}\plus \Varid{t2}){}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[E]%
 \\
 \>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves very much like an identity function, with the exception
-that it fails for two very specific patterns of input values defined using
-nested pattern matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-\quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick a number in the
-interval $[-n, n]$ with uniform distribution.
-%
-Hence, in principle we need to recognize what is a suitable generation size,
-which should be big enough for our generator be able to generate the numbers we
-pattern match against to, otherwise we will not be able to generate any value to
-match against \ensuremath{\Varid{foo}} clauses, leaving fragments of code completely untested.
-%
-Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-is big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this consideration, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the first error) results as follows:
-%
-\begin{align*}
-  &P(match(foo\#1))\\
-  &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
-  &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-If we use \megadeth to automatically derive a random generator for \ensuremath{\Conid{Exp}}, we
-obtain a uniform generation probability distribution over constructors, i.e.,
-$P(\ensuremath{\Conid{Val}}) = P(\ensuremath{\Conid{Add}}) = P(\ensuremath{\Conid{Mul}}) = 1/3$.
-%
-In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-average, we will need to generate over than two million test cases in order to
-be able to test the first clause of \ensuremath{\Varid{foo}} only once.
-%
-This situation can be somewhat improved if we use \dragen to obtain a random
-generator.
-%
-Using this tool, we can optimize the generation probabilities in order to
-benefit the generation of some data constructors over the rest.
-%
-Considering that the first pattern matching of \ensuremath{\Varid{foo}} involves the data
-constructor \ensuremath{\Conid{Add}} as the only recursive one, we can set an target generation
-proportion of \ensuremath{\Conid{Add}} data constructors of, for instance, $20:1$ with respect to
-the rest of the generated data constructors.
-%
-By doing so, the obtained distribution of values results such that
-$P(match(foo\#1)) \approx 1/300000$.
-%
-Although this certainly improves the probability of generating a matching value,
-this probability is not substantial enough to become practical.
-
-Additionally, by favoring the generation probabilities towards the \ensuremath{\Conid{Add}}
-constructor, we found that the probability of generating a value matching the
-second clause of \ensuremath{\Varid{foo}} (which does not matches against it) also gets diminished
-into an impractival value.
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left and right subtrees and sub-subtrees
-in order to preserve the height invariant.
+\>[3]{}\mathrel{=}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y}))){}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Join}\;(\Varid{simpl}\;\Varid{x})\;(\Varid{simpl}\;\Varid{y}){}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x})\mathrel{=}\Conid{Tag}\;\Varid{t}\;(\Varid{simpl}\;\Varid{x}){}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+This function simplifies sequences of \ensuremath{\Conid{Text}} constructors into a single big
+\ensuremath{\Conid{Text}} constructor.
+%
+To do so, it has to pattern match against sequences of \ensuremath{\Conid{Text}} constructors
+combined by a \ensuremath{\Conid{Join}} constructor using nested patterns (see \ensuremath{\Varid{simpl}} fist and
+second clauses).
+%
+The remaining clauses are only meant to propagate this simplification within
+nested expressions.
+
+Ideally, we would like to test each clause of the function \ensuremath{\Varid{simpl}} approximately
+the same amount of time each.
+%
+However, each data constructor is generated independently when using either
+\megadeth or \dragen, thus the probability of generating a value satisfying a
+nested pattern decreases multiplicatively with the number of constructors we
+pattern against to simultaneously.
+%
+In our tests, we found that the first two clauses of \ensuremath{\Varid{simpl}} get exercised only
+approximately $1.5\%$ of the time when using \megadeth to derive a random
+generator for \ensuremath{\Conid{HTML}}.
+%
+On the other hand, the best result we could achieve with \dragen was only able
+to exercise the first and second clauses of \ensuremath{\Varid{simpl}} approximately $3\%$ and
+$6\%$ of the time, respectively.
+%
+With both derivation tools, the most of the time was spent testing the trivial
+clauses of our function, in view of they pattern match against simpler patterns
+of input values.
+
+% In this light, we found that using a random generator obtained with \megadeth to
+% test this function will only exercise the first two clauses of |simpl| about
+% only $1.5\%$ of the time each, relegating the rest of the generated test cases
+% to the remaining and somewhat less interesting clauses, where, remarkably, half
+% of the time is spent exercising the last trivial clause.
+% %
+% Furthermore, using \dragen to ...
+
+% This function behaves very much like an identity function, with the exception
+% that it fails for two very specific patterns of input values defined using
+% nested pattern matching.
+% %
+% This is, |foo| does not only matches against the root data constructor, but also
+% against the data constructors of its subexpressions and sub-subexpressions.
+
+% \quickcheck's default implementation for random |Int|s pick a number in the
+% interval $[-n, n]$ with uniform distribution.
+% %
+% Hence, in principle we need to recognize what is a suitable generation size,
+% which should be big enough for our generator be able to generate the numbers we
+% pattern match against to, otherwise we will not be able to generate any value to
+% match against |foo| clauses, leaving fragments of code completely untested.
+% %
+% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
+% is big enough to produce an |Int| number equal to $50$.
+% %
+% Under this consideration, the probability of generating a value matching the
+% first clause of |foo| (and hence triggering the first error) results as follows:
+% %
+% \begin{align*}
+%   &P(match(foo\#1))\\
+%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
+%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
+% \end{align*}
+
+% If we use \megadeth to automatically derive a random generator for |Exp|, we
+% obtain a uniform generation probability distribution over constructors, i.e.,
+% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
+% %
+% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
+% average, we will need to generate over than two million test cases in order to
+% be able to test the first clause of |foo| only once.
+% %
+% This situation can be somewhat improved if we use \dragen to obtain a random
+% generator.
+% %
+% Using this tool, we can optimize the generation probabilities in order to
+% benefit the generation of some data constructors over the rest.
+% %
+% Considering that the first pattern matching of |foo| involves the data
+% constructor |Add| as the only recursive one, we can set an target generation
+% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
+% the rest of the generated data constructors.
+% %
+% By doing so, the obtained distribution of values results such that
+% $P(match(foo\#1)) \approx 1/300000$.
+% %
+% Although this certainly improves the probability of generating a matching value,
+% this probability is not substantial enough to become practical.
+
+% Additionally, by favoring the generation probabilities towards the |Add|
+% constructor, we found that the probability of generating a value matching the
+% second clause of |foo| (which does not matches against it) also gets diminished
+% into an impractival value.
+
+Altough the previous example might seem rather simple, branching against
+specific patterns of the input data is a common task.
+%
+For instance, balancing a Red-Black tree requires to consider specific
+combinations of color, left and right subtrees and sub-subtrees in order to
+preserve the height invariant \tocite.
 %
@@
 needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few.
+in a computationally more efficient way, to cite a few \tocite.
 
@@
 A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions that manipulates
+responsability of preserving its invariants to the functions which manipulates
 its values.
 %
-Consider for instance the following possible implementation of a basic \ensuremath{\Conid{HTML}}
-manipulation library:
-
+For this purspose, suppose we extend our \ensuremath{\Conid{Html}} data type with the following
+basic combinators:
+%
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Html}\mathrel{=}\Conid{Html}\;\Conid{String}{}\<[E]%
+\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
 \\[\blanklineskip]%
-\>[B]{}\Varid{head}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{head}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
-\\
 \>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <head>\char34}\plus \Varid{inner}\plus \text{\ttfamily \char34 </head>\char34}){}\<[E]%
+\>[3]{}\Varid{div}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{div}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 div\char34}\;\Varid{inner}{}\<[E]%
 \\[\blanklineskip]%
-\>[B]{}\Varid{body}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{body}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
-\\
 \>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <body>\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 </body>\char34}){}\<[E]%
+\>[3]{}\Varid{bold}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{bold}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 b\char34}\;\Varid{inner}{}\<[E]%
 \\[\blanklineskip]%
-\>[B]{}\Varid{br}\mathbin{::}\Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{br}\mathrel{=}\Conid{Html}\;\text{\ttfamily \char34 <br~/>\char34}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}(\mathop{\langle\! +\! \rangle})\mathbin{::}\Conid{Html}\to \Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}(\Conid{Html}\;\Varid{x})\mathop{\langle\! +\! \rangle}(\Conid{Html}\;\Varid{y})\mathrel{=}\Conid{Html}\;(\Varid{x}\plus \Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, the \ensuremath{\Conid{Html}} data type is defined using a single data
-constructor that contains the textual representation of the Html code it
-represents in plain text.
-%
-Later, the combinator functions defined over \ensuremath{\Conid{HTML}} are the ones in charge of
-transforming this plain text representation with the invariant that, given valid
-\ensuremath{\Conid{Html}} parameters, they always return a valid \ensuremath{\Conid{Html}} value.
-%
-This is a very commmon programming pattern that can be found in a variety of
-Haskell libraries \tocite{}.
-
-If we use a standard type-driven approach to derive a random generator for
-\ensuremath{\Conid{Html}}, the only ``structural'' information that we can use in the derivation
-process is that \ensuremath{\Conid{Html}} values are composed of strings.
-%
-As a consequence, we essentially end up generating random strings, which rarely
-represents a valid (or at least well structured) Html value.
-%
-Additionally, if our test suite contains properties constrained by certain
-preconditions, this lack of domain knowleadge may lead in an impractically high
-discard ratio of randomly generated test.
-%
-For instance, suposse we write a property to verify that \ensuremath{\Varid{head}} preserves the
-invariant previously mentioned:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{prop\char95 head\char95 inv}\mathbin{::}\Conid{Html}\to \Conid{Bool}{}\<[E]%
-\\
-\>[B]{}\Varid{prop\char95 head\char95 inv}\;\Varid{x}\mathrel{=}\Varid{valid}\;\Varid{x}\Longrightarrow\Varid{valid}\;(\Varid{head}\;\Varid{x}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition we state that \ensuremath{\Varid{head}} always returns a valid \ensuremath{\Conid{Html}}
-value (\ensuremath{\Varid{valid}\;(\Varid{head}\;\Varid{x})}), provided that we are given a valid \ensuremath{\Conid{Html}} as input
-(\ensuremath{\Varid{valid}\;\Varid{x}}).
-%
-Then, while testing this property we rarely satisfy its precondition, in which
-case \quickcheck discards the whole test, retrying with a diferent random input,
-degradating this way the testing performance.
-
-We believe that, if certain patterns of values are relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such patterns in a substantial proportion.
-%
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{hr}\mathbin{::}\Conid{Html}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\Varid{hr}\mathrel{=}\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+%
+% In the previous definition, the |Html| data type is defined using a single data
+% constructor that contains the textual representation of the Html code it
+% represents in plain text.
+%
+These functions encode additional information about the structure of our \ensuremath{\Conid{Html}}
+data type in the form of specific Html tags.
+
+Instead of including a new data constructor for each possible Html tag in out
+type definition, we defined a mininal representation and then extended it with
+an set of high level combinators.
+%
+% These combinators functions defined over |HTML| are the ones in charge of
+% transforming this plain text representation with the invariant that, given valid
+% |Html| parameters, they always return a valid |Html| value.
+%
+This programming pattern is often called a ``shallow embedding'', and can be
+found in a variety of Haskell libraries, being \emph{html} \tocite,
+\emph{svg-builder} \tocite some examples of this.
+
+As a consequence of this practice, type-driven derivation techniques often fail
+to synthesize useful random generators due to that most of the data type
+structure has been encoded into its abstract interface of combinators.
+%
+In our particular case, the chances of generating a \ensuremath{\Conid{Tag}} value representing a
+commonly used Html tag such as \ensuremath{\Varid{div}} are extremely low.
+
+% Additionally, if our test suite contains properties constrained by certain
+% preconditions, the lack of domain knowleadge may lead in an impractically high
+% discard ratio of randomly generated test cases.
+% %%
+% For instance, suposse we write a property to verify that |render| always outputs
+% a valid |Html| value:
+
+% \begin{code}
+% prop_render :: Html -> Bool
+% prop_render x = valid x ===> valid (parse (render x))
+% \end{code}
+
+% In the previous definition we state that |render| always returns a valid |Html|
+% value when it is parsed again (|parse (render x)|), provided that we are given a
+% valid |Html| as input (|valid x|).
+% %
+% In practice, not every string of characters constitutes a valid Html given that
+% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).
+
+% %
+% While testing this property we rarely satisfy its precondition, in which case
+% \quickcheck discards the whole test, retrying with a diferent random input,
+% degradating this way the testing performance.
+
+% \begin{code}
+% text :: String -> Bool
+% text str = Text (addEscapes str)
+% \end{code}
+
+% We believe that, if certain patterns of values are relevant enough to appear in
+% the codebase being tested, a practical random testing methodology should be able
+% to produce values satisfying such patterns in a substantial proportion.
+
+So far we have introduced two testing scenarios where type-driven derivation
+approaches are unable to capture all the available structure information from
+the user codebase.
+%
+Fortunately, this information can be automatically exploited and used to
+generate interesting and more structured random values.
+
 The next section introduces a reprentation model that let us encode the
@@
 
-In this section we present a compositional representation to express the random
-generation of values following the internal structure of their data types along
-with the structure present on patterns matchings and abstract interfaces.
-
-The key idea of this work is to represent different sources of structured value
-information of a data type in a homogeneous way that we call a ``higher-level
-representation'' (\ensuremath{\text{HRep}} from now on).%
-\footnote{The notion of higher-level comes from that the generation process of
-  given target data type is entirely determined by the type of the chosen
-  representation, instead of by a concrete generator defined at the term level.}
+In this section we present a compositional representation of values to express
+the random generation of values following the internal structure of their data
+types along with the structure present on patterns matchings and abstract
+interfaces.
+
+The key idea of this work is to represent different structured constructions of
+data in a homogeneous way that we call a ``higher-level representation'' (\ensuremath{\Conid{HRep}}
+from now on).
+%
+Instead of generating each data constructor independently, a random generator
+derived from this representation might generate composite structured values on
+each random choice it performs.
+%
+% \footnote{The notion of higher-level comes from that the generation process is
+%   entirely determined by the type of the chosen type level representation,
+%   instead of by a concrete generator defined at the term level.}
 %
@@
 level combinators in different ways into a ``generation specification'' that
-completely determines the generation process.
-
-We wil reuse the previously defined data type \ensuremath{\Conid{Exp}} and the function \ensuremath{\Varid{foo}} to
-explain the different concepts involved all across this section.
+completely determines the generation process behavior.
+
+We wil reuse the previously defined data type \ensuremath{\Conid{Html}} and the functions defined
+in the previous section to explain the different concepts involved all across
+this section.
 
@@
 %
-In this light, we represent each constructor of the data type \ensuremath{\Conid{Exp}} as follows:
+In this light, we represent each constructor of the data type \ensuremath{\Conid{Html}} as follows:
 
@@
 \column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{Val}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Val}\;\Conid{Int}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Add}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Add}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{Mul}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}\text{Mk}_{Mul}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\hrep{Text}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}Con_{Text}\;{}\<[29]%
+\>[29]{}\Conid{String}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{Sing}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}Con_{Sing}\;{}\<[29]%
+\>[29]{}\Conid{String}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{Join}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}Con_{Join}\;{}\<[29]%
+\>[29]{}\Varid{r}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{Tag}\;{}\<[16]%
+\>[16]{}\Varid{r}\mathrel{=}Con_{Tag}\;{}\<[29]%
+\>[29]{}\Conid{String}\;\Varid{r}{}\<[E]%
 \ColumnHook
@@
 %
-For instance, the value (\ensuremath{\text{Mk}_{Add}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\text{HRep}_{Add}\;\Conid{Int}}, while the
-value (\ensuremath{\text{Mk}_{Mul}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\text{HRep}_{Mul}\;\Conid{Bool}}.
-%
+For instance, the value (\ensuremath{Con_{Join}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\hrep{Join}\;\Conid{Int}}, while the
+value (\ensuremath{Con_{Join}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\hrep{Join}\;\Conid{Bool}}.
+
 In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
-chosen generation specification (which might be composed of several \ensuremath{\text{HRep}}s),
+chosen generation specification (which might be composed of several \ensuremath{\Conid{HRep}}s),
 without having to modify anything in the underlying machinery.
 
-Having the \ensuremath{\text{HRep}} of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : \ensuremath{\text{HRep}}_f \rightarrow t$) that maps a value from each
+Having the \ensuremath{\Conid{HRep}} of each data constructor, we can define an evaluation relation
+($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
 representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
@@
 \begin{alignat*}{4}
-  &\evalrep{\ensuremath{\text{Mk}_{Val}}\ n\ \ \ &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Val}}\ n \\
-  &\evalrep{\ensuremath{\text{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
-  &\evalrep{\ensuremath{\text{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
+  &\evalrep{\ensuremath{Con_{Text}}\ s &&}{Html}
+    &&= \ensuremath{\Conid{Text}}\ s \\
+  &\evalrep{\ensuremath{Con_{Sing}}\ s &&}{Html}
+    &&= \ensuremath{\Conid{Sing}}\ s \\
+  &\evalrep{\ensuremath{Con_{Join}}\ x\ y &&}{Html}
+    &&= \ensuremath{\Conid{Join}}\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
+  &\evalrep{\ensuremath{Con_{Tag}}\ s\ x &&}{Html}
+    &&= \ensuremath{\Conid{Tag}}\ s\ \evalrep{x}{Html}
 \end{alignat*}
 
-The last missing piece is to automatically derive a random generators for each
-constructor \ensuremath{\text{HRep}}.
-%
-For this purpose, it is important to consider that each constructor \ensuremath{\text{HRep}} has
+The missing piece is to automatically synthesize a random generators for each
+constructor representation.
+%
+For this purpose, it is important to consider that each constructor \ensuremath{\Conid{HRep}} has
 its recursive fields abstracted away with a type parameter that will be later
@@
 Given that this specification is unknown at the derivation time, we parametrize
-each \ensuremath{\text{HRep}} generator with a random generator \ensuremath{gen_{r}} that is used to generate
-each recursive fields:
+each \ensuremath{\Conid{HRep}} generator with a random generator \ensuremath{\gen{r}} that is used to generate
+random values for each recursive field:
 
@@
 \column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{29}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Val}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Val}{}\<[29]%
+\>[3]{}\gen{Text}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Con_{Text}{}\<[29]%
 \>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
 \\
-\>[3]{}gen_{Add}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Add}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{Mul}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}\text{Mk}_{Mul}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
+\>[3]{}\gen{Sing}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Con_{Sing}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
+\\
+\>[3]{}\gen{Join}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Con_{Join}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
+\\
+\>[3]{}\gen{Tag}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Con_{Tag}{}\<[29]%
+\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
 
@@
 
-In this work we define type level combinators which let the user combine the
-automatically derived \ensuremath{\text{HRep}}s in several ways.
-%
-In first place, we define a type combinator ($\_^\bullet$) to tag a \ensuremath{\text{HRep}} to be
+The next step is to define a series of type level combinators to enable us
+combining the automatically derived \ensuremath{\Conid{HRep}}s in several ways.
+
+In first place, we define a type combinator (\ensuremath{\term{\_}}) to tag a \ensuremath{\Conid{HRep}} to be
 terminal, i.e., a representation that is allowed be generated when the
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(f^\bullet)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\text{HRep}} with an
+\>[B]{}\mathbf{data}\;(\term{f})\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\Conid{HRep}} with an
 explicit generation frequency $n$:
@@
 \begin{alignat*}{4}
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Term\ x &&: \ensuremath{f^\bullet}    &&}{t} &&= \evalrep{x : f}{t}
+  &\evalrep{Term\ x &&: \ensuremath{\term{f}}    &&}{t} &&= \evalrep{x : f}{t} \\
+  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
 \end{alignat*}
@@
 \column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{f^\bullet}\;{}\<[16]%
-\>[16]{}gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Conid{Term}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
-\\
-\>[B]{}gen_{f \otimes n}\;{}\<[16]%
-\>[16]{}gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Conid{Freq}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
+\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\gen{f}^\bigstar\;{}\<[16]%
+\>[16]{}\gen{r}{}\<[22]%
+\>[22]{}\mathrel{=}\Conid{Term}{}\<[30]%
+\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
+\\[\blanklineskip]%
+\>[B]{}\gen{f \otimes n}\;{}\<[16]%
+\>[16]{}\gen{r}{}\<[22]%
+\>[22]{}\mathrel{=}\Conid{Freq}{}\<[30]%
+\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
 \ColumnHook
@@
 Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-\ensuremath{\text{HRep}}s into a single one using a sum type and representing a random choice
-between them:
+\ensuremath{\Conid{HRep}}s into a single one using a sum type to represent a random choice between
+them:
 
@@
 A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data type variant and evaluating the
-inner \ensuremath{\text{HRep}}:
+target data type by pattern matching on the data constructor variant and
+evaluating the inner \ensuremath{\Conid{HRep}} accodingly:
 %
@@
 
-Generating a composite \ensuremath{\text{HRep}} is slightly more complicated than before, as we
+Generating a composite \ensuremath{\Conid{HRep}} is slightly more complicated than before, as we
 need to perform a random choice based on the generation size and the given
@@
 \column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{24}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{27}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{45}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{f \oplus g}\;gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
+\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{43}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\gen{f \oplus g}\;\gen{r}{}\<[22]%
+\>[22]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
 \\
@@
 \>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{}){}\<[E]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
+\>[37]{}\gen{r}{}\<[43]%
+\>[43]{}){}\<[E]%
 \\
@@
 \>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{})\mskip1.5mu]{}\<[E]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
+\>[37]{}\gen{r}{}\<[43]%
+\>[43]{})\mskip1.5mu]{}\<[E]%
 \\
@@
 \>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{})\mskip1.5mu]){}\<[E]%
+\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
+\>[37]{}\gen{r}{}\<[43]%
+\>[43]{})\mskip1.5mu]){}\<[E]%
 \ColumnHook
@@
 
-The previous definition we reflect the type level frequencies of \ensuremath{\Varid{f}} and \ensuremath{\Varid{g}}
-(\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}).
-%
-This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
-used.
-%
-Then we use these frequencies to generate each inner \ensuremath{\text{HRep}} in the apropriate
+In the previous definition, we reflect the type level frequencies of the types
+\ensuremath{\Varid{f}} and \ensuremath{\Varid{g}} (\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}) into term-level values.
+%
+This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
+%
+Then we use these frequencies to generate each inner \ensuremath{\Conid{HRep}} in the apropriate
 proportion.
@@
 When the generation size gets exhausted, we reflect the terminal generation
-frequency of each inner \ensuremath{\text{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
+frequency of each inner \ensuremath{\Conid{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
 \ensuremath{\Varid{freq0}\;\textit{@}\!g}).
@@
 
-With the introduced combinators, we can easily create a type synonym \ensuremath{\text{HRep}_{Exp}}
-to specify a generation schema equivalent to the one seen in the concrete random
-generator of type \ensuremath{\Conid{Exp}} presented in Section \ref{sec:randomtesting}:
+Later, we can use these combinators to create a type synonym \ensuremath{\hrep{Html}} that
+specifies a generation schema equivalent to the one seen in the concrete random
+generator of type \ensuremath{\Conid{Html}} presented in Section \ref{sec:randomtesting}:
 
@@
 \column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{Exp}{}\<[16]%
+\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\hrep{Html}{}\<[16]%
 \>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\text{HRep}_{Val}^\bullet{}\<[E]%
+\>[24]{}\term{\hrep{Text}}\;{}\<[39]%
+\>[39]{}\otimes\;\mathrm{2}{}\<[E]%
 \\
 \>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{Add}\;{}\<[34]%
-\>[34]{}\otimes\;\mathrm{2}{}\<[E]%
+\>[24]{}\term{\hrep{Sing}}{}\<[E]%
 \\
 \>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\text{HRep}_{Mul}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{\text{HRep}_{Exp}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit in the definition of \ensuremath{\text{HRep}_{Exp}}.
-%
-If we want to generate recursive values using this representation we need to
-introduce a last type level combinator to ``tie the knot'' recursively:
+\>[24]{}\hrep{Join}\;{}\<[39]%
+\>[39]{}\otimes\;\mathrm{3}{}\<[E]%
+\\
+\>[16]{}\!\;\oplus\;{}\<[24]%
+\>[24]{}\hrep{Tag}\;{}\<[39]%
+\>[39]{}\otimes\;\mathrm{4}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+However, a value of type \ensuremath{\hrep{Html}} still has its recursive calls abstracted
+away---the type parameter \ensuremath{\Varid{r}} is implicit at the definition of \ensuremath{\hrep{Html}}.
+%
+We can think of it as a ``single layer'' of representation.
+%
+To make it able to represent recursive values we need to define a last type
+level combinator to ``tie the knot'':
 
@@
 %
-To translate fixed points back to our target data type we simply need translate
-the inner representation:
+Then, to translate fixed points back to our target data type we simply need
+translate the inner representation:
 %
@@
 \end{align*}
-%
-Categorically, this generic transformation of fixed points is called a
-\emph{catamorphism}, where our evaluation relation is known as its
-\emph{algebra}.
-
-Unlike the other combinators, to randomly generate a fixed point of a certain
-representation \ensuremath{\Varid{f}}, we do not parametrize the generation of the recursive fields
-of the inner represention over an external generator \ensuremath{gen_{r}}.
-%
-Instead, we can also tie the recursive knot in our random generator by piping
-the generation of the inner \ensuremath{\text{HRep}} parametrized recursively by the itself:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;genFix_{f}{}\<[E]%
+
+% Categorically, this generic transformation of fixed points is called a
+% \emph{catamorphism}, where our evaluation relation is known as its
+% \emph{algebra} \tocite.
+
+Unlike the other combinators, to define a random generator for the fixed point
+of a certain representation \ensuremath{\Varid{f}}, we do not to parametrize the generation of the
+recursive fields of \ensuremath{\Varid{f}} over an external generator \ensuremath{\gen{r}}.
+%
+% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
+Instead, we replace it with our fixed point generator, calling itself
+recursively on any recursive field of \ensuremath{\Varid{f}} that might appear inside:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;genFix_{f}{}\<[E]%
 \ColumnHook
@@
 %
-With this generator, we can define a random generator for our target data type
-simply by evaluating a random value of the representation:
+Then, we can define a random generator for our target data type simply by
+generating a random value of our chosen representation, and transforming it back
+to our target by the means of the evaluation relation:
 
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Exp}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{(\text{HRep}\,{Exp})}{}\<[E]%
-\\
-\>[17]{}\Varid{return}\;\evalrep{x}{Exp}{}\<[E]%
+\>[3]{}gen_{Html}\mathrel{=}\mathbf{do}\;{}\<[17]%
+\>[17]{}\Varid{x}\leftarrow genFix_{\hrep{Html}}{}\<[E]%
+\\
+\>[17]{}\Varid{return}\;\evalrep{x}{Html}{}\<[E]%
 \ColumnHook
@@
 
+So far we have introduced the machinery required to represent the random
+generation of a target data type considering only the structure encoded on its
+definition.
+%
+However, this approach can now be easily extended to encode different sources of
+structured information.
+%
+We proceed to introduce two extensions that help to address the problematic
+testing scenarios presented in the previous section.
+
 \subsection*{\textbf{Representing Pattern Matchings}}
 
+We can follow a similar reasoning as before to represent the pattern matching
+structure from a given function.
+%
+Consider the nested pattern matchings of function \ensuremath{\Varid{simpl}} defined in the
+previous section:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{36}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[36]%
+\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
+\\
+\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[36]%
+\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+To represent these patterns, we derive a new data for each one of them, whose
+fields represent the pattern variables that occur inside, and abstracting away
+every pattern variable of type \ensuremath{\Conid{Html}}.
+%
+Concretely, we define the following data types:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#1}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{foo\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}\text{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
+\>[B]{}\mathbf{data}\;\hrep{simpl\#1}\;{}\<[18]%
+\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#1}\;\Conid{String}\;\Conid{String}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{simpl\#2}\;{}\<[18]%
+\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#2}\;\Conid{String}\;\Varid{r}\;\Varid{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+The first pattern of \ensuremath{\Varid{simpl}} contains two pattern variables (\ensuremath{\Varid{t1}} and \ensuremath{\Varid{t2}}) of
+type \ensuremath{\Conid{String}} that are included as fields of \ensuremath{Pat_{simpl\#1}}.
+%
+Similarly, the second pattern of \ensuremath{\Varid{simpl}} contains a pattern variable \ensuremath{\Varid{t1}} of
+type \ensuremath{\Conid{String}} represented by the first field of \ensuremath{Pat_{simpl\#2}}, along with two
+pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type \ensuremath{\Conid{Html}} that are abstracted in the second
+and third fields of \ensuremath{Pat_{simpl\#2}}.
+
+To transform these representations back to our target data type, we simply need
+to expand them it into the concrete value represented by the original pattern,
+evaluating its fields back to \ensuremath{\Conid{Html}} as well:
+%
 \begin{align*}
-  &\evalrep{\ensuremath{\text{Mk}_{foo\#1}}\ x\ y}{Exp} = \\
-  &\ \ \ensuremath{\Conid{Add}\;(\Conid{Add}\;\evalrep{x}{Exp}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\evalrep{y}{Exp})}\\
-  &\evalrep{\ensuremath{\text{Mk}_{foo\#2}}\ x\ y}{Exp} = \\
-  &\ \ \ensuremath{\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\evalrep{y}{Exp})}
+  &\evalrep{\ensuremath{Pat_{simpl\#1}}\ t1\ t2}{Html} \\
+  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html})\;(\Conid{Text}\;\evalrep{t2}{Html})}\\
+  &\evalrep{\ensuremath{Pat_{simpl\#2}}\ t1\ x\ y}{Html} = \\
+  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html}))\;\evalrep{x}{Html})\;\evalrep{y}{Html})}
 \end{align*}
 
+This way, any value of type \ensuremath{\hrep{simpl\#1}} is guaranteed to satisfy the first
+pattern matching of \ensuremath{\Varid{simpl}}---the same property follows for \ensuremath{\hrep{simpl\#2}}.
+
+The random generation of pattern \ensuremath{\Conid{HRep}}s is defined in the same way as we did
+before for representations of data constructors:
+
 \begin{hscode}\SaveRestoreHook
@@
 \column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\gen{simpl\#1}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathrel{=}Pat_{simpl\#1}{}\<[16]%
+\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[31]%
+\>[31]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{arbitrary}{}\<[E]%
+\\
+\>[B]{}\gen{simpl\#2}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[E]%
+\\
+\>[B]{}\hsindent{3}{}\<[3]%
+\>[3]{}\mathrel{=}Pat_{simpl\#2}{}\<[16]%
+\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+ %$
+
+Finally, we can join the pattern matching representation of each clause of
+\ensuremath{\Varid{simpl}} into a single one:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\hrep{simpl}{}\<[17]%
+\>[17]{}\mathrel{=}\hrep{simpl\#1}\;\!\!\oplus\!\!\;{}\<[39]%
+\>[39]{}\hrep{simpl\#2}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+\subsection*{\textbf{Representing Abstract Interfaces}}
+
+To introduce the higher level representation of module abstract interfaces,
+consider module \ensuremath{\Conid{M}} defininf \ensuremath{\Conid{Html}} combinators introduced in the previous
+section.
+
+% \begin{code}
+% module M where
+
+% ten :: Exp
+% ten = Val 10
+
+% square :: Exp -> Exp
+% square x = Mul x x
+
+% minus :: Exp -> Exp -> Exp
+% minus x y = Add x (Mul y (Val (-1)))
+% \end{code}
+
+We can represent each function of \ensuremath{\Conid{M}} that returns a value of type \ensuremath{\Conid{Html}}
+following the same idea as before, deriving a data type with a single data
+constructor for each one of them:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{33}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{52}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{foo\#1}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}\text{Mk}_{foo\#1}{}\<[33]%
-\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[52]%
-\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{foo\#2}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}\text{Mk}_{foo\#2}{}\<[33]%
-\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[52]%
-\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{data}\;\hrep{div}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}Fun_{div}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{bold}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}Fun_{bold}\;\Varid{r}{}\<[E]%
+\\
+\>[B]{}\mathbf{data}\;\hrep{hr}\;{}\<[21]%
+\>[21]{}\Varid{r}\mathrel{=}Fun_{hr}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+In this case, each single constructor will have as fields the types of the
+inputs of the function that they represent.
+%
+As before, we abstract away any field representing and input of type \ensuremath{\Conid{Html}} with
+a type parameter \ensuremath{\Varid{r}}.
+
+To evaluate these representations, we simply use the values on its fields as
+input parameter of each original function, returning its result.
+%
+\begin{alignat*}{3}
+  &\evalrep{\ensuremath{Fun_{div}}\  &&x  &&}{Html} = \ensuremath{\Varid{div}\;\Varid{eval\char95 x\char95 html}} \\
+  &\evalrep{\ensuremath{Fun_{bold}}\ &&x &&}{Html} = \ensuremath{\Varid{bold}\;\Varid{eval\char95 x\char95 html}} \\
+  &\evalrep{\ensuremath{Fun_{hr}}    &&  &&}{Html} = \ensuremath{\Varid{hr}}
+\end{alignat*}
+
+Note that, by doing this, the generation process inherits any patology that the
+functions we use to generate values might have.
+%
+For instance, if the function \ensuremath{\Varid{div}} would happen to be non-terminating for some
+inputs, our generation process could suffer from this as well.
+
+Furthermore, the generation procedure for abstract interface representations
+follows the same pattern as before:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[3]{}\gen{div}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Fun_{div}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
+\\
+\>[3]{}\gen{bold}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}Fun_{bold}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
+\\
+\>[3]{}\gen{hr}\;{}\<[12]%
+\>[12]{}\gen{r}{}\<[18]%
+\>[18]{}\mathrel{=}\Varid{return}\;Fun_{hr}{}\<[E]%
 \ColumnHook
@@
 
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{foo}{}\<[16]%
-\>[16]{}\mathrel{=}\text{HRep}_{foo\#1}\;\!\!\oplus\!\!\;{}\<[38]%
-\>[38]{}\text{HRep}_{foo\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Representing Abstract Interfaces}}
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{ten}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{ten}\mathrel{=}\Conid{Val}\;\mathrm{10}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Conid{Mul}\;\Varid{y}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1}))){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+And we will also define a type synonym to join all the representations of the
+module \ensuremath{\Conid{M}} into a single one.
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{50}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;\hrep{M}{}\<[13]%
+\>[13]{}\mathrel{=}\hrep{div}\;\!\!\oplus\!\!\;{}\<[32]%
+\>[32]{}\hrep{bold}\;\!\!\oplus\!\!\;{}\<[50]%
+\>[50]{}\hrep{hr}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+Finally, we can put all the derived machinery together into a generation
+specification \ensuremath{Html_S}, assigning (possibly) different generation frequencies to
+each individual \ensuremath{\Conid{HRep}} we combine:
+
+\begin{hscode}\SaveRestoreHook
+\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
 \column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\text{HRep}_{ten}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{square}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{square}\;{}\<[37]%
-\>[37]{}\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\text{HRep}_{minus}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}\text{Mk}_{minus}\;{}\<[37]%
-\>[37]{}\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{alignat*}{3}
-  &\evalrep{\ensuremath{\text{Mk}_{ten}}         &&}{Exp} = \ensuremath{\Varid{ten}} \\
-  &\evalrep{\ensuremath{\text{Mk}_{square}}\ x   &&}{Exp} = \ensuremath{\Varid{square}\;\evalrep{x}{Exp}} \\
-  &\evalrep{\ensuremath{\text{Mk}_{minus}}\ x\ y &&}{Exp} = \ensuremath{\Varid{minus}\;\evalrep{x}{Exp}\;\evalrep{y}{Exp}}
-\end{alignat*}
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{35}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{54}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{ten}\;{}\<[15]%
-\>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{ten}{}\<[E]%
-\\
-\>[3]{}gen_{square}\;{}\<[15]%
-\>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{square}{}\<[35]%
-\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{minus}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}\text{Mk}_{minus}{}\<[35]%
-\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[54]%
-\>[54]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{55}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\text{HRep}_{M}{}\<[14]%
-\>[14]{}\mathrel{=}\text{HRep}_{ten}\;\!\!\oplus\!\!\;{}\<[34]%
-\>[34]{}\text{HRep}_{square}\;\!\!\oplus\!\!\;{}\<[55]%
-\>[55]{}\text{HRep}_{minus}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\Conid{Spec}{}\<[12]%
-\>[12]{}\mathrel{=}{}\<[20]%
-\>[20]{}\text{HRep}_{Exp}\;{}\<[30]%
-\>[30]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}\text{HRep}_{foo}\;{}\<[30]%
-\>[30]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}\text{HRep}_{M}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This previous definition can be interpreted graphically as it is shown in the
-Figure \ref{fig:hrep}.
-%
-Curly arrows represent the structural information extracted using
-meta-programming.
-
-\begin{figure}[t]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{Higher level representation of the data type \ensuremath{\Conid{Exp}}, defined using
-    structural information from the function \ensuremath{\Varid{foo}} and the abstract interface of
-    the module \ensuremath{\Conid{M}}.}
-  \label{fig:hrep}
-\end{figure}
+\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
+\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
+\>[B]{}\mathbf{type}\;Html_S{}\<[13]%
+\>[13]{}\mathrel{=}{}\<[21]%
+\>[21]{}\hrep{Html}\;{}\<[32]%
+\>[32]{}\otimes\;\mathrm{4}{}\<[E]%
+\\
+\>[13]{}\!\;\oplus\;{}\<[21]%
+\>[21]{}\hrep{simpl}\;{}\<[32]%
+\>[32]{}\otimes\;\mathrm{2}{}\<[E]%
+\\
+\>[13]{}\!\;\oplus\;{}\<[21]%
+\>[21]{}\hrep{M}{}\<[E]%
+\ColumnHook
+\end{hscode}\resethooks
+
+% This previous definition can be interpreted graphically as it is shown in the
+% Figure \ref{fig:hrep}, where curly arrows represent the structural information
+% extracted using meta-programming.
+
+% \begin{figure}[t]
+%   \centering
+%   \input{tikz/hrep.tex}
+%   \caption{Higher level representation of the data type |Exp|, defined using
+%     structural information from the function |foo| and the abstract interface of
+%     the module |M|.}
+%   \label{fig:hrep}
+% \end{figure}
+
+We want to remark that, for space reasons, we were only able to introduce the
+representation of a rather simple target data type.
+%
+In practice, this reasoning can be extended to mutually recursive and parametric
+types as well.
+
+Overall, this approach offers significant advantages over the usual type-driven
+derivation of random generators:
+%
+\begin{itemize}
+\item \textbf{Composability:} we can combine different atomic representations
+  using different structure information sources depending on what property or
+  sub-system we need to verify using randomly generated values.
+  %
+\item \textbf{Extensibility:} the developer can derive representations for new
+  sources of structure information and combine them with the existing ones
+  simply by adding them to the existing generation specification of the target
+  data type.
+  %
+\item \textbf{Predictability:} using branching processes theory, it is possible
+  to predict the average distribution of generated values in terms of number of
+  constructors.
+  %
+  This prediction is completely modular, and can be obtained for any composite
+  representation obtained using the automatically derived \ensuremath{\Conid{HRep}}s and the
+  provided combinators.
+\end{itemize}
+
+\todo[inline]{Maybe we can show an example random value from the representation
+  and its corresponding target value }
 
@@
 % \end{code} %$
+
+% \begin{figure}[b]
+%   \centering
+%   \input{tikz/val.tex}
+%   \caption{Higher level representation of the data type |Exp|, defined using
+%     structural information from the function |foo| and the abstract interface of
+%     the module |M|.}
+%   \label{fig:hrep}
+% \end{figure}
 \section{Random Generators Synthesis} \label{sec:synthesis}

### 2019-01-20 paper/sections.old/finalremarks.tex
WIP

--- 2019-01-15 paper/sections/finalremarks.tex
@@
 ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+vulputate vestibulum lorem.

### 2019-01-20 paper/sections.old/hrep.tex
WIP

--- 2019-01-17 paper/sections/hrep.tex
@@
 \footnote{The notion of higher-level comes from that the generation process of
-  given target data type is entirely determined by the type of the chosen
-  representation, instead of by a concrete generator defined at the term level.}
+  given target data type is entirely determined by the type of the chosen type
+  level representation, instead of by a concrete generator defined at the term
+  level.}
 %
@@
 Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-|HRep|s into a single one using a sum type and representing a random choice
-between them:
+|HRep|s into a single one using a sum type to represent a random choice between
+them:
 
@@
 target data type by pattern matching on the data type variant and evaluating the
-inner |HRep|:
+inner |HRep| accodingly:
 %
@@
 
-The previous definition we reflect the type level frequencies of |f| and |g|
+In the previous definition, we reflect the type level frequencies of |f| and |g|
 (|freq at_f| and |freq at_g|).
 %
-This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
-used.
+This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
 %
@@
 
-With the introduced combinators, we can easily create a type synonym |HRep_Exp|
-to specify a generation schema equivalent to the one seen in the concrete random
+Later, we can use these combinators to create a type synonym |HRep_Exp| that
+specifies a generation schema equivalent to the one seen in the concrete random
 generator of type |Exp| presented in Section \ref{sec:randomtesting}:
@@
 However, a value of type |HRep_Exp| still has its recursive calls abstracted
-away---the type parameter |r| is implicit in the definition of |HRep_Exp|.
+away---the type parameter |r| is implicit at the definition of |HRep_Exp|.
 %
-If we want to generate recursive values using this representation we need to
-introduce a last type level combinator to ``tie the knot'' recursively:
+We can think of it as a ``single layer'' of representation.
+%
+To make it able to represent recursive values we need to define a last type
+level combinator to ``tie the knot'':
 
@@
 Unlike the other combinators, to randomly generate a fixed point of a certain
-representation |f|, we do not parametrize the generation of the recursive fields
-of the inner represention over an external generator |gen_r|.
+representation |f|, we do not needt to parametrize the generation of the
+recursive fields of the inner represention over an external generator |gen_r|.
 %
-Instead, we can tie the recursive knot in our random generator by piping the
-generation of the inner |HRep|, and calling itself to generate recursive values:
+Instead, we simply pipe the generation to the inner |HRep|, calling itself on
+any recursive generation call that might appear inside:
 
@@
 
-So far we have introduced the representation machinery required to represent the
-random generation of a target data type considering only the structure encoded
-on its definition.
+So far we have introduced the machinery required to represent the random
+generation of a target data type considering only the structure encoded on its
+definition.
 %
-This approach has the advantage of being extendable with different sources of
+However, this approach can now be easily extended to encode different sources of
 structured information.
@@
 
-The first pattern of |foo| conteins two pattern variables (|x| and |y|) of type
-|Exp| that are abstracted in the fields of |Mk_foo_1|.
+The first pattern of |foo| contains two pattern variables (|x| and |y|) of type
+|Exp| that are abstracted in the parametric fields of |Mk_foo_1|.
 %
@@
 
-To evaluate these representations back to our target data type, we simply expand
-them it into the concrete value represented by the original pattern:
+To transform these representations back to our target data type, we simply
+expand them it into the concrete value represented by the original pattern,
+evaluating its fields back to |Exp| as well:
 %
@@
 And the random generation of pattern |HRep|s is defined in the same way as we
-did before with data constructors |HRep|s:
+did before for data constructors |HRep|s:
 
@@
 
-Moreover, the generation procedure for abstract interface representations
+Note that, by doing this, the generation process inherits any patology that the
+functions we use to generate values might have.
+%
+For instance, if the function |square| is non-terminating for some inputs, our
+generation process could suffer from this as well.
+%
+Furthermore, the generation procedure for abstract interface representations
 follows the same pattern as before:
@@
 Finally, we can put all the derived machinery together into a generation
-specification |Spec| for the type |Exp|, assigning (possibly) different
-generation frequencies to each individual |HRep| that constitutes it:
+specification |Exp_S|, assigning (possibly) different generation frequencies to
+each individual |HRep| we combine:
 
 \begin{code}
-type Spec  =       HRep_Exp  otimes 4
+type Exp_S  =       HRep_Exp  otimes 4
             oplus'  HRep_foo  otimes 2
@@
 
-This representation can be extended to mutually recursive and parametric types
-...
+We want to remark that, for space reasons, we were only able to introduce the
+representation of a rather simple target data type.
+%
+In practice, this reasoning can be extended to mutually recursive and parametric
+types as well.
+
+Overall, we believe that this approach has some advantages over the usual
+type-driven derivation of random generators:
+%
+\begin{itemize}
+\item \textbf{Composability:} we can combine different atomic representations
+  using different structure information sources depending on what property or
+  sub-system we need to verify using random values.
+  %
+\item \textbf{Extensibility:} the developer can derive representations for new
+  sources of structure information and combine them with the existing ones
+  simply by adding them to the exiting representation of the target data type.
+  %
+\item \textbf{Predictability:} using branching processes theory, it is possible
+  to predict the average distribution of generated values in terms of number of
+  constructors.
+  %
+  This prediction is completely modular, and can be obtained for any composite
+  representation obtained using the automatically derived |HRep|s and the
+  provided combinators.
+\end{itemize}
+
+\todo[inline]{Maybe we can show an example random value from the representation
+  and its corresponding target value }
 
@@
 % \end{code} %$
+
+% \begin{figure}[b]
+%   \centering
+%   \input{tikz/val.tex}
+%   \caption{Higher level representation of the data type |Exp|, defined using
+%     structural information from the function |foo| and the abstract interface of
+%     the module |M|.}
+%   \label{fig:hrep}
+% \end{figure}

### 2019-01-20 paper/sections/randomtesting.tex
WIP

--- 2019-01-16 paper/sections/randomtesting.tex
@@
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
+\section{Random Data Generation in Haskell} \label{sec:randomtesting}
 
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
+In this section we briefly introduce the common approach for automatically
+deriving random data generators in Haskell using a type-driven approach, along
+with its main drawbacks.
 
@@
 %
-Essentially, an ADTs is a composite type defined by combining other types.
+Essentially, an ADTs is a composite type defined by combining other types in
+terms of \textbf{sums} (also known as \emph{variant types}) and
+\textbf{products} (or tuples) of other data types.
 %
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
+To exemplify this, consider the following type definition to encoding |Html|
+expressions:
 %
-To exemplify this, consider the following definition of the data type |Exp|
-encoding integer expressions:
-
 \begin{code}
-data Exp  =   Val  Int
-          pp  Add  Exp  Exp
-          pp  Mul  Exp  Exp
+data Html  =   Text  String
+           pp  Sing  String
+           pp  Join  Html    Html
+           pp  Tag   String  Html
 \end{code}
 
-In the previous definition, we declare |Exp| as the sum of three possible
-classes of values: |Val| represents literal values, whereas |Add| and |Mul|
-represent the addition and multiplication of two integer expressions,
-respectively.
+In the previous definition, we declare |Html| as the sum of four possible
+constructions: |Text| represents plain text values. |Sing| and |Tag| represent
+singular and paired Html tags, and |Join| concats two expression one after
+another.
 %
-In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} and are
-used to distiguish which variant of the data type we are reffering to.
+We only encode a very small subset of the actual Html specification for
+illustrative reasons.
 %
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
+In Haskell, |Text|, |Sing| and |Join| and |Tag| are known as \emph{data
+  constructors} (or constructors for short) and are used to distiguish which
+variant of the data type we are constructing.
 %
-In particular, |Val| contains a field of type |Int|, while |Add| and |Mul|
-contain two fields of type |Exp| representing the operands of each operation.
+Each data constructor is defined as a product of zero or more types known as
+\emph{fields}.
 %
-Note that |Exp| is used as a field of at least one of its data constructors
-(case |Add| and |Mul|), making it a recursively defined ADT.
+For instance, |Text| has a field of type |String|, whereas |Join| two recusive
+fields of type |Html|.
 %
@@
 %
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using |Exp| data constructors as |(Add (Val 2) (Mul (Val 5) (Val 6)))|.
+With this reprentation, the expression ``|html_example|'' can be encoded as:
 %
-Furthermore, an evaluation function from |Exp|s to integer values can be defined
-very idiomatically:
-
 \begin{code}
-eval :: Exp -> Int
-eval (Val n)    = n
-eval (Add x y)  = eval x  +  eval y
-eval (Mul x y)  = eval x  *  eval y
+Tag "html" (Join (Join
+  (Text "hello") (Sing "hr")) (Text "bye"))
+\end{code}
+%
+Additionally, we can define a function |render| to serialize |Html| values as
+follows:
+%
+\begin{code}
+render :: Html -> String
+render (Text t) = t
+render (Sing t) = "<" ++ t ++ ">"
+render (Join x y) = render x ++ render y
+render (Tag t x)
+  = "<" ++ t ++ ">" ++ render x ++ "</" ++ t ++ ">"
 \end{code}
 
-In the previous definition, |eval| is described using \emph{pattern matching}
+In the previous definition, |render| is described using \emph{pattern matching}
 over each possible kind of value.
 %
-For the input case of a literal values, we simply return the value contained in
-the |Val| constructor.
+Using pattern matching we can define functions idiomatically by defining
+different function clauses for each input pattern we are interested on.
 %
-On the other hand, if the input value matches either an |Add| or a |Mul| data
-constructor, where |x| and |y| are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
+Patterns can be defined to match specific constructors, literal values or
+variable subexpressions (like |t|, |x| and |y| in the definition of |render|).
+%
+They can also be nested in order to match very specific patterns of values.
 
-\subsection{\textbf{Type-Driven Generation of Random Values}}
+\subsection*{\textbf{Type-Driven Generation of Random Values}}
 
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
+In order to generate random value of types involving user defined ADTs, most
+approaches require the user to provide a random data generator for each one of
+them.
 %
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
+This is a cumbersome and error prone task that closely follows the data type
+structure.
 %
 For instance, consider the following definition of a \quickcheck random
-generator for the type |Exp|:
+generator for the type |Html|:
+%
+\begin{code}
+genHtml = sized (\size ->
+   if size == 0
+   then frequency
+     [  (2,  Text  <$> arbitrary)
+     ,  (1,  Sing  <$> arbitrary) ]
+   else frequency
+     [  (2,  Text  <$> arbitrary)
+     ,  (1,  Sing  <$> arbitrary)
+     ,  (3,  Join  <$> smaller genHtml <*> smaller genHtml)
+     ,  (4,  Tag   <$> arbitrary <*> smaller genHtml) ])
+\end{code} %$
+%
+This random generator is defined using \quickcheck's combinator |sized| to
+paremetrize the generation process up to an external natural number known as the
+\emph{generation size}.
+%
+This parameter is chosen by the user, and we use to limit the maximum amount of
+recursive calls that this random generator can perform.
+%
+When called with a positive generation size, this generator can pick to generate
+among any |Html| data constructor of with a explicitly given frequency that can
+be chosen by the user (2, 1, 3 and 4 for |Text|, |Sing|, |Join| and |Tag|
+constructors, respectively).
+%
+When it picks to generate a |Val| or a |Sing| data constructor, it also
+generates a random |String| value using the standard overloaded generator
+|arbitrary| (\quickcheck provides standard random generators for most base types
+like |String|, |Int|, |Bool|, etc.). \footnote{%
+  The operators |<$>| and |<*>| are used in Haskell to combine values obtained
+  from effectful computations (like calling to a random generator) and they are
+  not particularly relevant for the point being made in this work.}
+%$
+On the other hand, when it picks to generate either an |Join| constructor, it
+also generates two independent random subexpression recursively, decreasing the
+generation size by a unit on each recursive invocation (|smaller genHtml|).
+%
+The case of random generation of |Tag| constructors follows analogously.
 
-\begin{code}
-genExp :: Gen Exp
-genExp = sized gen
-  where
-    gen 0 = Val <$> arbitrary
-    gen n = oneof
-      [  Val  <$> arbitrary
-      ,  Add  <$> gen pred   <*> gen pred
-      ,  Mul  <$> gen pred   <*> gen pred ]
-\end{code} %$
+This random process keeps calling itself recursively until the generation size
+reaches zero, where the generator is constrained to pick only among terminal
+data constructors, being |Text| and |Sing| the only possible choices in our
+particular case.
 
-This random generator is defined using \quickcheck's combinator |sized :: (Int
--> Gen a) -> Gen a| to paremetrize the generation process up to an external
-integer number known as the \emph{generation size}.
+% Strictly decreasing the the generation size by one on each recursive call
+% results in a useful property over the generated data: every generated value has
+% at most |n| levels, where |n| is the generation size set by the user.
 %
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
+% This property enables us to model the generation process using the theory of
+% branching processes introduced by \tocite, and extended in this work as
+% described in Section \ref{sec:synthesis}.
+
+The previous definition is rather mechanical, except perhaps for the chosen
+generation frequencies.
 %
-When called with a positive generation size (case |gen n|), the generator is free
-to randomly pick with uniform probability among any data constructor of |Exp|.
-%
-In the case it picks to generate a |Val| data constructor, it also generates a
-random |Int| value using the standard overloaded generator |arbitrary|
-(\quickcheck provides standard random generators for most base types like |Int|,
-|Bool|, etc.).
-%
-On the other hand, when it randomly picks to generate either an |Add| or a |Mul|
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively (|gen
-pred|), ensuring to decrease the generation size by a unit on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case |gen 0|), where the generator is constrained to pick only
-among terminal data constructors, being |Val| the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most |n| levels, where |n| is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
+In this light, it easy to extend this procedure to any data type defined in an
+algebraic fashion.
 %
@@
 %
-The simplest tool for such purpose is \megadeth.
+The simplest tool for such purpose is \megadeth \tocite.
 %
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
+Given a target data type, it synthesizes a random generator for it that behaves
+similarly to the one presented above, where the generation frequencies are
+defined to be uniform across constructors.
 %
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
+However, picking among different data constructors with uniform frequency can
+lead to a generation process biased towards generating (in average) very small
+values, regardless of the generation sized set by the user \tocite.
 
@@
 %
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
+In practice, this procedure is often too generic to generate random data with
+enough structural complexity to be used for testing purposes.
 
-In this work we identify two sources of additional structure information which
-are not considered by the aforementioned automatic derivation tools to obtain
-random data generators:
+% In this work we identify two sources of additional structure information which
+% are not considered by the aforementioned automatic derivation tools to obtain
+% better random data generators:
+
+In this work we identify two patological situations which are not properly
+handled by the aforementioned derivation tools:
 
 \begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code yields part of the responsability of preserving its
-  invariants to the functions of its abstract interface.
+\item The target code behaves differently over inputs matching specific patterns
+  of nested values.
+\item The target code encodes a significant portion of its structure on its
+  abstract interface.
 \end{enumerate}
 
-This information can be reified in compile time and used to synthesize richer
-random generators automatically. We proceed to exemplify the previous points in
-detail.
+In Section \ref{sec:hrep} we show how this structural information can be used to
+synthesize richer random generators automatically.
+%
+We proceed to exemplify the previous points in detail.
 
-\subsection{\textbf{Presence of Complex Pattern Matchings}}
+\subsection*{\textbf{Presence of Complex Pattern Matchings}}
 
-To exemplify the firts problematic scenarion that may arise, suppose we want to
-use randomly generated |Exp|s to test a property comprising the following
-function:
-
+To exemplify the first problematic scenario, suppose we want to use randomly
+generated |Htmls|s to test a property comprising the following function:
+%
 \begin{code}
-foo :: Exp -> Exp
-foo (Add (Add x (Val 50)) (Add (Val 25) y))
-  = error "pattern #1"
-foo (Mul (Val 50) (Mul (Val x) y))
-  = error "pattern #2"
-foo x = x
+simpl :: Html -> Html
+simpl (Join (Text t1) (Text t2))
+  = Text (t1 ++ t2)
+simpl (Join (Join (Text t1) x) y)
+  = simpl (Join (Text t1) (simpl (Join x y)))
+simpl (Join x y) = Join (simpl x) (simpl y)
+simpl (Tag t x) = Tag t (simpl x)
+simpl x = x
 \end{code}
 
-This function behaves very much like an identity function, with the exception
-that it fails for two very specific patterns of input values defined using
-nested pattern matching.
+This function simplifies sequences of |Text| constructors into a single big
+|Text| constructor.
 %
-This is, |foo| does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
+To do so, it has to pattern match against sequences of |Text| constructors
+combined by a |Join| constructor using nested patterns (see |simpl| fist and
+second clauses).
+%
+The remaining clauses are only meant to propagate this simplification within
+nested expressions.
 
-\quickcheck's default implementation for random |Int|s pick a number in the
-interval $[-n, n]$ with uniform distribution.
+Ideally, we would like to test each clause of the function |simpl| approximately
+the same amount of time each.
 %
-Hence, in principle we need to recognize what is a suitable generation size,
-which should be big enough for our generator be able to generate the numbers we
-pattern match against to, otherwise we will not be able to generate any value to
-match against |foo| clauses, leaving fragments of code completely untested.
+However, each data constructor is generated independently when using either
+\megadeth or \dragen, thus the probability of generating a value satisfying a
+nested pattern decreases multiplicatively with the number of constructors we
+pattern against to simultaneously.
 %
-Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-is big enough to produce an |Int| number equal to $50$.
+In our tests, we found that the first two clauses of |simpl| get exercised only
+approximately $1.5\%$ of the time when using \megadeth to derive a random
+generator for |HTML|.
 %
-Under this consideration, the probability of generating a value matching the
-first clause of |foo| (and hence triggering the first error) results as follows:
+On the other hand, the best result we could achieve with \dragen was only able
+to exercise the first and second clauses of |simpl| approximately $3\%$ and
+$6\%$ of the time, respectively.
+%
+With both derivation tools, the most of the time was spent testing the trivial
+clauses of our function, in view of they pattern match against simpler patterns
+of input values.
 
-\begin{align*}
-  P(match(foo\#1))
-  \!&&=\; &P(Add) \\
-  \!&&*\; &P(Add) * 1 * (P(Val) * (1/100)) \\
-  \!&&*\; &P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
+% In this light, we found that using a random generator obtained with \megadeth to
+% test this function will only exercise the first two clauses of |simpl| about
+% only $1.5\%$ of the time each, relegating the rest of the generated test cases
+% to the remaining and somewhat less interesting clauses, where, remarkably, half
+% of the time is spent exercising the last trivial clause.
+% %
+% Furthermore, using \dragen to ...
 
-If we use \megadeth to automatically derive a random generator for |Exp|, we
-obtain a uniform generation probability distribution over constructors, i.e.,
-$P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
+% This function behaves very much like an identity function, with the exception
+% that it fails for two very specific patterns of input values defined using
+% nested pattern matching.
+% %
+% This is, |foo| does not only matches against the root data constructor, but also
+% against the data constructors of its subexpressions and sub-subexpressions.
+
+% \quickcheck's default implementation for random |Int|s pick a number in the
+% interval $[-n, n]$ with uniform distribution.
+% %
+% Hence, in principle we need to recognize what is a suitable generation size,
+% which should be big enough for our generator be able to generate the numbers we
+% pattern match against to, otherwise we will not be able to generate any value to
+% match against |foo| clauses, leaving fragments of code completely untested.
+% %
+% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
+% is big enough to produce an |Int| number equal to $50$.
+% %
+% Under this consideration, the probability of generating a value matching the
+% first clause of |foo| (and hence triggering the first error) results as follows:
+% %
+% \begin{align*}
+%   &P(match(foo\#1))\\
+%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
+%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
+% \end{align*}
+
+% If we use \megadeth to automatically derive a random generator for |Exp|, we
+% obtain a uniform generation probability distribution over constructors, i.e.,
+% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
+% %
+% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
+% average, we will need to generate over than two million test cases in order to
+% be able to test the first clause of |foo| only once.
+% %
+% This situation can be somewhat improved if we use \dragen to obtain a random
+% generator.
+% %
+% Using this tool, we can optimize the generation probabilities in order to
+% benefit the generation of some data constructors over the rest.
+% %
+% Considering that the first pattern matching of |foo| involves the data
+% constructor |Add| as the only recursive one, we can set an target generation
+% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
+% the rest of the generated data constructors.
+% %
+% By doing so, the obtained distribution of values results such that
+% $P(match(foo\#1)) \approx 1/300000$.
+% %
+% Although this certainly improves the probability of generating a matching value,
+% this probability is not substantial enough to become practical.
+
+% Additionally, by favoring the generation probabilities towards the |Add|
+% constructor, we found that the probability of generating a value matching the
+% second clause of |foo| (which does not matches against it) also gets diminished
+% into an impractival value.
+
+Altough the previous example might seem rather simple, branching against
+specific patterns of the input data is a common task.
 %
-In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-average, we will need to generate over than two million test cases in order to
-be able to test the first clause of |foo| only once.
-%
-This situation can be somewhat improved if we decide to use \dragen to obtain a
-random generator.
-%
-By optimizing the generation probabilities in order to benefit the generation of
-|Add| constructor in a ratio of $20:1$ with respect to the rest of them, we
-obtained a distribution of values where $P(match(foo\#1)) \approx 1/300000$.
-
-%
-Although this certainly improves the probability of generating a matching value,
-this probability is not substantial enough to become practical.
-%
-Additionally, by favoring the generation probabilities towards the |Add|
-constructor, we found that the probability of generating a value matching the
-second clause of |foo| (which does not matches against it) also gets diminished
-into an impractival value.
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left and right subtrees and sub-subtrees
-in order to preserve the height invariant.
+For instance, balancing a Red-Black tree requires to consider specific
+combinations of color, left and right subtrees and sub-subtrees in order to
+preserve the height invariant \tocite.
 %
@@
 needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few.
+in a computationally more efficient way, to cite a few \tocite.
 
-\subsection{\textbf{Data Invariants Encoded on Abstract Interfaces}}
+\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
 
 A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions that manipulates
+responsability of preserving its invariants to the functions which manipulates
 its values.
 %
-Consider for instance the following possible implementation of a basic |HTML|
-manipulation library:
+For this purspose, suppose we extend our |Html| data type with the following
+basic combinators:
+%
+\begin{code}
+module M where
 
-\begin{code}
-data Html = Html String
+  div :: Html -> Html
+  div inner = Tag "div" inner
 
-head :: Html -> Html
-head (Html inner)
-  = Html ("<head>" ++ inner ++ "</head>")
+  bold :: Html -> Html
+  bold inner = Tag "b" inner
 
-body :: Html -> Html
-body (Html inner)
-  = Html ("<body>" ++ t ++ "</body>")
+  hr :: Html
+  hr = Sing "hr"
+\end{code}
+%
+% In the previous definition, the |Html| data type is defined using a single data
+% constructor that contains the textual representation of the Html code it
+% represents in plain text.
+%
+These functions encode additional information about the structure of our |Html|
+data type in the form of specific Html tags.
 
-br :: Html
-br = Html "<br />"
+Instead of including a new data constructor for each possible Html tag in out
+type definition, we defined a mininal representation and then extended it with
+an set of high level combinators.
+%
+% These combinators functions defined over |HTML| are the ones in charge of
+% transforming this plain text representation with the invariant that, given valid
+% |Html| parameters, they always return a valid |Html| value.
+%
+This programming pattern is often called a ``shallow embedding'', and can be
+found in a variety of Haskell libraries, being \emph{html} \tocite,
+\emph{svg-builder} \tocite some examples of this.
 
-concat :: Html -> Html -> Html
-concat (Html x) (Html y) = Html (x ++ y)
-\end{code}
+As a consequence of this practice, type-driven derivation techniques often fail
+to synthesize useful random generators due to that most of the data type
+structure has been encoded into its abstract interface of combinators.
+%
+In our particular case, the chances of generating a |Tag| value representing a
+commonly used Html tag such as |div| are extremely low.
 
-In the previous definition, the |Html| data type is defined using a single data
-constructor that contains the textual representation of the Html code it
-represents in plain text.
+% Additionally, if our test suite contains properties constrained by certain
+% preconditions, the lack of domain knowleadge may lead in an impractically high
+% discard ratio of randomly generated test cases.
+% %%
+% For instance, suposse we write a property to verify that |render| always outputs
+% a valid |Html| value:
+
+% \begin{code}
+% prop_render :: Html -> Bool
+% prop_render x = valid x ===> valid (parse (render x))
+% \end{code}
+
+% In the previous definition we state that |render| always returns a valid |Html|
+% value when it is parsed again (|parse (render x)|), provided that we are given a
+% valid |Html| as input (|valid x|).
+% %
+% In practice, not every string of characters constitutes a valid Html given that
+% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).
+
+% %
+% While testing this property we rarely satisfy its precondition, in which case
+% \quickcheck discards the whole test, retrying with a diferent random input,
+% degradating this way the testing performance.
+
+% \begin{code}
+% text :: String -> Bool
+% text str = Text (addEscapes str)
+% \end{code}
+
+% We believe that, if certain patterns of values are relevant enough to appear in
+% the codebase being tested, a practical random testing methodology should be able
+% to produce values satisfying such patterns in a substantial proportion.
+
+So far we have introduced two testing scenarios where type-driven derivation
+approaches are unable to capture all the available structure information from
+the user codebase.
 %
-Later, the combinator functions defined over |HTML| are the ones in charge of
-transforming this plain text representation with the invariant that, given valid
-|Html| parameters, they always return a valid |Html| value.
-%
-This is a very commmon programming pattern that can be found in a variety of
-Haskell libraries \tocite{}.
+Fortunately, this information can be automatically exploited and used to
+generate interesting and more structured random values.
 
-If we use a standard type-driven approach to derive a random generator for
-|Html|, the only ``structural'' information that we can use in the derivation
-process is that |Html| values are composed of strings.
-%
-As a consequence, we essentially end up generating random strings, which rarely
-represents a valid (or at least well structured) Html value.
-%
-Additionally, if our test suite contains properties constrained by certain
-preconditions, this lack of domain knowleadge may lead in an impractically high
-discard ratio of randomly generated test.
-%
-For instance, suposse we write a property to verify that |head| preserves the
-invariant previously mentioned:
-
-\begin{code}
-prop_head_inv :: Html -> Bool
-prop_head_inv x = valid x ===> valid (head x)
-\end{code}
-
-In the previous definition we state that |head| always returns a valid |Html|
-value (|valid (head x)|), provided that we are given a valid |Html| as input
-(|valid x|).
-%
-Then, while testing this property we rarely satisfy its precondition, in which
-case \quickcheck discards the whole test, retrying with a diferent random input,
-degradating this way the testing performance.
-
-We believe that, if certain patterns of values are relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such patterns in a substantial proportion.
-%
 The next section introduces a reprentation model that let us encode the

### 2019-01-20 paper/tikz/hrep.tex
WIP

--- 2019-01-16 paper/tikz/hrep.tex
@@
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,416.5); %set diagram left start at 0, and has height of 416.5
+\hspace*{-4mm}
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.9,xscale=0.9]
+%uncomment if require: \path (0,405.8571319580078); %set diagram left start at 0, and has height of 405.8571319580078
 
+%Shape: Rectangle [id:dp09506349711233386]
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=0.75]  (110,50) ..
+controls (110,44.48) and (114.48,40) ..
+(120,40) -- (470,40) ..
+controls (475.52,40) and (480,44.48) ..
+(480,50) -- (480,270) ..
+controls (480,275.52) and (475.52,280) ..
+(470,280) -- (120,280) ..
+controls (114.48,280) and (110,275.52) ..
+(110,270) -- cycle ;
 %Shape: Rectangle [id:dp9489660675592899]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (130,20) ..
-controls (130,14.48) and (134.48,10) ..
-(140,10) -- (400,10) ..
-controls (405.52,10) and (410,14.48) ..
-(410,20) -- (410,80) ..
-controls (410,85.52) and (405.52,90) ..
-(400,90) -- (140,90) ..
-controls (134.48,90) and (130,85.52) ..
-(130,80) -- cycle ;
-%Snip Same Side Corner Rect [id:dp09931070350524518]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,30.5) -- (150.6,20) -- (189.5,20) -- (200,30.5) -- (200,50) -- (200,50) -- (140.1,50) -- (140.1,50) -- cycle ;
-%Down Arrow [id:dp5984296172784]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]  (140,159.5) -- (144,159.5) -- (144,90) -- (152,90) -- (152,159.5) -- (156,159.5) -- (148,170) -- cycle ;
-%Straight Lines [id:da30879479079788386]
-\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1.5]    (145,90) -- (151,90) ;
-
-%Straight Lines [id:da5824111707250128]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (170,170) ..
-controls (168.33,168.33) and (168.33,166.67) ..
-(170,165) ..
-controls (171.67,163.33) and (171.67,161.67) ..
-(170,160) ..
-controls (168.33,158.33) and (168.33,156.67) ..
-(170,155) ..
-controls (171.67,153.33) and (171.67,151.67) ..
-(170,150) ..
-controls (168.33,148.33) and (168.33,146.67) ..
-(170,145) ..
-controls (171.67,143.33) and (171.67,141.67) ..
-(170,140) ..
-controls (168.33,138.33) and (168.33,136.67) ..
-(170,135) ..
-controls (171.67,133.33) and (171.67,131.67) ..
-(170,130) ..
-controls (168.33,128.33) and (168.33,126.67) ..
-(170,125) ..
-controls (171.67,123.33) and (171.67,121.67) ..
-(170,120) ..
-controls (168.33,118.33) and (168.33,116.67) ..
-(170,115) ..
-controls (171.67,113.33) and (171.67,111.67) ..
-(170,110) ..
-controls (168.33,108.33) and (168.33,106.67) ..
-(170,105) ..
-controls (171.67,103.33) and (171.67,101.67) ..
-(170,100) ..
-controls (168.33,98.33) and (168.33,96.67) ..
-(170,95) -- (170,91) -- (170,83) ;
-\draw [shift={(170,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (150,60) ..
+controls (150,54.48) and (154.48,50) ..
+(160,50) -- (460,50) ..
+controls (465.52,50) and (470,54.48) ..
+(470,60) -- (470,260) ..
+controls (470,265.52) and (465.52,270) ..
+(460,270) -- (160,270) ..
+controls (154.48,270) and (150,265.52) ..
+(150,260) -- cycle ;
+%Straight Lines [id:da7051587981224026]
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (310,290) ..
+controls (308.33,288.33) and (308.33,286.67) ..
+(310,285) ..
+controls (311.67,283.33) and (311.67,281.67) ..
+(310,280) ..
+controls (308.33,278.33) and (308.33,276.67) ..
+(310,275) ..
+controls (311.67,273.33) and (311.67,271.67) ..
+(310,270) ..
+controls (308.33,268.33) and (308.33,266.67) ..
+(310,265) ..
+controls (311.67,263.33) and (311.67,261.67) ..
+(310,260) ..
+controls (308.33,258.33) and (308.33,256.67) ..
+(310,255) ..
+controls (311.67,253.33) and (311.67,251.67) ..
+(310,250) ..
+controls (308.33,248.33) and (308.33,246.67) ..
+(310,245) ..
+controls (311.67,243.33) and (311.67,241.67) ..
+(310,240) -- (310,232) ;
+\draw [shift={(310,230)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
 (0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
-
-%Straight Lines [id:da7051587981224026]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (270,110) ..
-controls (268.33,108.33) and (268.33,106.67) ..
-(270,105) ..
-controls (271.67,103.33) and (271.67,101.67) ..
-(270,100) ..
-controls (268.33,98.33) and (268.33,96.67) ..
-(270,95) -- (270,91) -- (270,83) ;
-\draw [shift={(270,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
-(0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
 
 %Straight Lines [id:da6087008244886778]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (370,110) ..
-controls (368.33,108.33) and (368.33,106.67) ..
-(370,105) ..
-controls (371.67,103.33) and (371.67,101.67) ..
-(370,100) ..
-controls (368.33,98.33) and (368.33,96.67) ..
-(370,95) -- (370,91) -- (370,83) ;
-\draw [shift={(370,80)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1.5]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (420,290) ..
+controls (418.33,288.33) and (418.33,286.67) ..
+(420,285) ..
+controls (421.67,283.33) and (421.67,281.67) ..
+(420,280) ..
+controls (418.33,278.33) and (418.33,276.67) ..
+(420,275) ..
+controls (421.67,273.33) and (421.67,271.67) ..
+(420,270) ..
+controls (418.33,268.33) and (418.33,266.67) ..
+(420,265) ..
+controls (421.67,263.33) and (421.67,261.67) ..
+(420,260) ..
+controls (418.33,258.33) and (418.33,256.67) ..
+(420,255) ..
+controls (421.67,253.33) and (421.67,251.67) ..
+(420,250) -- (420,242) ;
+\draw [shift={(420,240)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
 (0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
 
 %Shape: Rectangle [id:dp701556783783656]
-\draw  [line width=1.5]  (230,115) ..
-controls (230,112.24) and (232.24,110) ..
-(235,110) -- (305,110) ..
-controls (307.76,110) and (310,112.24) ..
-(310,115) -- (310,135) ..
-controls (310,137.76) and (307.76,140) ..
-(305,140) -- (235,140) ..
-controls (232.24,140) and (230,137.76) ..
-(230,135) -- cycle ;
+\draw  [line width=0.75]  (270,295) ..
+controls (270,292.24) and (272.24,290) ..
+(275,290) -- (345,290) ..
+controls (347.76,290) and (350,292.24) ..
+(350,295) -- (350,315) ..
+controls (350,317.76) and (347.76,320) ..
+(345,320) -- (275,320) ..
+controls (272.24,320) and (270,317.76) ..
+(270,315) -- cycle ;
 
 %Shape: Rectangle [id:dp033430127809365384]
-\draw  [line width=1.5]  (330,115) ..
-controls (330,112.24) and (332.24,110) ..
-(335,110) -- (405,110) ..
-controls (407.76,110) and (410,112.24) ..
-(410,115) -- (410,135) ..
-controls (410,137.76) and (407.76,140) ..
-(405,140) -- (335,140) ..
-controls (332.24,140) and (330,137.76) ..
-(330,135) -- cycle ;
-%Shape: Rectangle [id:dp6050538306488522]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (140.1,50) -- (200,50) -- (200,80) -- (140.1,80) -- cycle ;
-%Shape: Rectangle [id:dp9303311048687424]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (340,50) -- (400,50) -- (400,80) -- (340,80) -- cycle ;
+\draw  [line width=0.75]  (380,295) ..
+controls (380,292.24) and (382.24,290) ..
+(385,290) -- (455,290) ..
+controls (457.76,290) and (460,292.24) ..
+(460,295) -- (460,315) ..
+controls (460,317.76) and (457.76,320) ..
+(455,320) -- (385,320) ..
+controls (382.24,320) and (380,317.76) ..
+(380,315) -- cycle ;
 
 %Up Arrow [id:dp0013771056350821986]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (260,149.5) -- (270,140) -- (280,149.5) -- (273.5,149.5) -- (273.5,170) -- (266.5,170) -- (266.5,149.5) -- cycle ;
-%Shape: Rectangle [id:dp7366907361635939]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.43,50) -- (300,50) -- (300,80) -- (240.43,80) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (210,65) ..
-controls (210,59.48) and (214.48,55) ..
-(220,55) ..
-controls (225.52,55) and (230,59.48) ..
-(230,65) ..
-controls (230,70.52) and (225.52,75) ..
-(220,75) ..
-controls (214.48,75) and (210,70.52) ..
-(210,65) -- cycle ; \draw  [line width=1.5]  (210,65) -- (230,65) ; \draw  [line width=1.5]  (220,55) -- (220,75) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (310,65) ..
-controls (310,59.48) and (314.48,55) ..
-(320,55) ..
-controls (325.52,55) and (330,59.48) ..
-(330,65) ..
-controls (330,70.52) and (325.52,75) ..
-(320,75) ..
-controls (314.48,75) and (310,70.52) ..
-(310,65) -- cycle ; \draw  [line width=1.5]  (310,65) -- (330,65) ; \draw  [line width=1.5]  (320,55) -- (320,75) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (152.93,28.79) ..
-controls (156.83,24.88) and (163.17,24.88) ..
-(167.07,28.79) ..
-controls (170.98,32.69) and (170.98,39.02) ..
-(167.07,42.93) ..
-controls (163.17,46.83) and (156.83,46.83) ..
-(152.93,42.93) ..
-controls (149.02,39.02) and (149.02,32.69) ..
-(152.93,28.79) -- cycle ; \draw  [line width=1.5]  (152.93,28.79) -- (167.07,42.93) ; \draw  [line width=1.5]  (167.07,28.79) -- (152.93,42.93) ;
-%Snip Same Side Corner Rect [id:dp4221187902899477]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (240.1,30.5) -- (250.6,20) -- (289.5,20) -- (300,30.5) -- (300,50) -- (300,50) -- (240.1,50) -- (240.1,50) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1.5]  (252.93,28.79) ..
-controls (256.83,24.88) and (263.17,24.88) ..
-(267.07,28.79) ..
-controls (270.98,32.69) and (270.98,39.02) ..
-(267.07,42.93) ..
-controls (263.17,46.83) and (256.83,46.83) ..
-(252.93,42.93) ..
-controls (249.02,39.02) and (249.02,32.69) ..
-(252.93,28.79) -- cycle ; \draw  [line width=1.5]  (252.93,28.79) -- (267.07,42.93) ; \draw  [line width=1.5]  (267.07,28.79) -- (252.93,42.93) ;
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=0.75]  (300,331.43) -- (310,320) -- (320,331.43) -- (313.5,331.43) -- (313.5,350) -- (306.5,350) -- (306.5,331.43) -- cycle ;
 %Shape: Rectangle [id:dp1692119124811613]
-\draw  [line width=1.5]  (130,170) -- (410,170) -- (410,200) -- (130,200) -- cycle ;
-%Up Arrow [id:dp4999372801455]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1.5]  (360,149.5) -- (370,140) -- (380,149.5) -- (373.5,149.5) -- (373.5,170) -- (366.5,170) -- (366.5,149.5) -- cycle ;
+\draw  [line width=0.75]  (110,350) -- (480,350) -- (480,380) -- (110,380) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (263,172.93) ..
+controls (263,169.02) and (259.83,165.86) ..
+(255.93,165.86) ..
+controls (252.02,165.86) and (248.86,169.02) ..
+(248.86,172.93) ..
+controls (248.86,176.83) and (252.02,180) ..
+(255.93,180) ..
+controls (259.83,180) and (263,176.83) ..
+(263,172.93) -- cycle ; \draw  [line width=0.75]  (263,172.93) -- (248.86,172.93) ; \draw  [line width=0.75]  (255.93,165.86) -- (255.93,180) ;
+%Shape: Rectangle [id:dp6865365415345861]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (270,120.5) -- (350,120.5) -- (350,230.5) -- (270,230.5) -- cycle ;
+%Shape: Rectangle [id:dp39543698275133865]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (276.29,147.66) -- (343.21,147.66) -- (343.21,170.34) -- (276.29,170.34) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (286,116.5) ..
+controls (288.76,113.74) and (288.76,109.26) ..
+(286,106.5) ..
+controls (283.24,103.74) and (278.76,103.74) ..
+(276,106.5) ..
+controls (273.24,109.26) and (273.24,113.74) ..
+(276,116.5) ..
+controls (278.76,119.26) and (283.24,119.26) ..
+(286,116.5) -- cycle ; \draw  [line width=0.75]  (286,116.5) -- (276,106.5) ; \draw  [line width=0.75]  (286,106.5) -- (276,116.5) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (317.14,183.43) ..
+controls (317.14,179.52) and (313.98,176.36) ..
+(310.07,176.36) ..
+controls (306.17,176.36) and (303,179.52) ..
+(303,183.43) ..
+controls (303,187.33) and (306.17,190.5) ..
+(310.07,190.5) ..
+controls (313.98,190.5) and (317.14,187.33) ..
+(317.14,183.43) -- cycle ; \draw  [line width=0.75]  (317.14,183.43) -- (303,183.43) ; \draw  [line width=0.75]  (310.07,176.36) -- (310.07,190.5) ;
+%Shape: Rectangle [id:dp047421885346712545]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (276.08,197.66) -- (343,197.66) -- (343,220.34) -- (276.08,220.34) -- cycle ;
+%Shape: Rectangle [id:dp0425475572348164]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (380,80) -- (460,80) -- (460,240) -- (380,240) -- cycle ;
+%Shape: Rectangle [id:dp9247208762532781]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (385,106.37) -- (455,106.37) -- (455,129) -- (385,129) -- cycle ;
+%Shape: Rectangle [id:dp8712965895636344]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (385.52,157.37) -- (454.48,157.37) -- (454.48,180) -- (385.52,180) -- cycle ;
+%Shape: Rectangle [id:dp933766425327347]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (386.66,208.24) -- (453.34,208.24) -- (453.34,230.87) -- (386.66,230.87) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (427.14,143.95) ..
+controls (427.14,140.05) and (423.98,136.89) ..
+(420.07,136.89) ..
+controls (416.17,136.89) and (413,140.05) ..
+(413,143.95) ..
+controls (413,147.84) and (416.17,151) ..
+(420.07,151) ..
+controls (423.98,151) and (427.14,147.84) ..
+(427.14,143.95) -- cycle ; \draw  [line width=0.75]  (427.14,143.95) -- (413,143.95) ; \draw  [line width=0.75]  (420.07,136.89) -- (420.07,151) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (427.14,192.95) ..
+controls (427.14,189.05) and (423.98,185.89) ..
+(420.07,185.89) ..
+controls (416.17,185.89) and (413,189.05) ..
+(413,192.95) ..
+controls (413,196.84) and (416.17,200) ..
+(420.07,200) ..
+controls (423.98,200) and (427.14,196.84) ..
+(427.14,192.95) -- cycle ; \draw  [line width=0.75]  (427.14,192.95) -- (413,192.95) ; \draw  [line width=0.75]  (420.07,185.89) -- (420.07,200) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (373,172.93) ..
+controls (373,169.02) and (369.83,165.86) ..
+(365.93,165.86) ..
+controls (362.02,165.86) and (358.86,169.02) ..
+(358.86,172.93) ..
+controls (358.86,176.83) and (362.02,180) ..
+(365.93,180) ..
+controls (369.83,180) and (373,176.83) ..
+(373,172.93) -- cycle ; \draw  [line width=0.75]  (373,172.93) -- (358.86,172.93) ; \draw  [line width=0.75]  (365.93,165.86) -- (365.93,180) ;
+%Shape: Rectangle [id:dp6050538306488522]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (160,77.59) -- (240,77.59) -- (240,260) -- (160,260) -- cycle ;
+%Shape: Rectangle [id:dp20494666512059623]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (170,107.54) -- (230,107.54) -- (230,130.17) -- (170,130.17) -- cycle ;
+%Shape: Rectangle [id:dp8741156804316592]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (170,177.36) -- (230,177.36) -- (230,199.99) -- (170,199.99) -- cycle ;
+%Shape: Rectangle [id:dp7972859165129229]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (170,227.24) -- (230,227.24) -- (230,249.87) -- (170,249.87) -- cycle ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (186,173.22) ..
+controls (188.76,170.46) and (188.76,166) ..
+(186,163.24) ..
+controls (183.24,160.49) and (178.76,160.49) ..
+(176,163.24) ..
+controls (173.24,166) and (173.24,170.46) ..
+(176,173.22) ..
+controls (178.76,175.97) and (183.24,175.97) ..
+(186,173.22) -- cycle ; \draw  [line width=0.75]  (186,173.22) -- (176,163.24) ; \draw  [line width=0.75]  (186,163.24) -- (176,173.22) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (176.01,73) ..
+controls (178.77,70.25) and (178.77,65.78) ..
+(176.01,63.02) ..
+controls (173.25,60.27) and (168.77,60.27) ..
+(166.01,63.02) ..
+controls (163.25,65.78) and (163.25,70.25) ..
+(166.01,73) ..
+controls (168.77,75.75) and (173.25,75.75) ..
+(176.01,73) -- cycle ; \draw  [line width=0.75]  (176.01,73) -- (166.01,63.02) ; \draw  [line width=0.75]  (176.01,63.02) -- (166.01,73) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (207.14,143.22) ..
+controls (207.14,139.33) and (203.98,136.17) ..
+(200.07,136.17) ..
+controls (196.17,136.17) and (193,139.33) ..
+(193,143.22) ..
+controls (193,147.12) and (196.17,150.27) ..
+(200.07,150.27) ..
+controls (203.98,150.27) and (207.14,147.12) ..
+(207.14,143.22) -- cycle ; \draw  [line width=0.75]  (207.14,143.22) -- (193,143.22) ; \draw  [line width=0.75]  (200.07,136.17) -- (200.07,150.27) ;
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=0.75]  (207.14,213.05) ..
+controls (207.14,209.15) and (203.98,205.99) ..
+(200.07,205.99) ..
+controls (196.17,205.99) and (193,209.15) ..
+(193,213.05) ..
+controls (193,216.94) and (196.17,220.1) ..
+(200.07,220.1) ..
+controls (203.98,220.1) and (207.14,216.94) ..
+(207.14,213.05) -- cycle ; \draw  [line width=0.75]  (207.14,213.05) -- (193,213.05) ; \draw  [line width=0.75]  (200.07,205.99) -- (200.07,220.1) ;
+%Down Arrow [id:dp5984296172784]
+\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=0.75]  (190,340.81) -- (195,340.81) -- (195,280) -- (205,280) -- (205,340.81) -- (210,340.81) -- (200,350) -- cycle ;
+%Straight Lines [id:da30879479079788386]
+\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=0.75]    (195,280) -- (205,280) ;
+
+%Straight Lines [id:da8667104419281686]
+\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (220,350) ..
+controls (218.33,348.33) and (218.33,346.67) ..
+(220,345) ..
+controls (221.67,343.33) and (221.67,341.67) ..
+(220,340) ..
+controls (218.33,338.33) and (218.33,336.67) ..
+(220,335) ..
+controls (221.67,333.33) and (221.67,331.67) ..
+(220,330) ..
+controls (218.33,328.33) and (218.33,326.67) ..
+(220,325) ..
+controls (221.67,323.33) and (221.67,321.67) ..
+(220,320) ..
+controls (218.33,318.33) and (218.33,316.67) ..
+(220,315) ..
+controls (221.67,313.33) and (221.67,311.67) ..
+(220,310) ..
+controls (218.33,308.33) and (218.33,306.67) ..
+(220,305) ..
+controls (221.67,303.33) and (221.67,301.67) ..
+(220,300) ..
+controls (218.33,298.33) and (218.33,296.67) ..
+(220,295) ..
+controls (221.67,293.33) and (221.67,291.67) ..
+(220,290) ..
+controls (218.33,288.33) and (218.33,286.67) ..
+(220,285) ..
+controls (221.67,283.33) and (221.67,281.67) ..
+(220,280) ..
+controls (218.33,278.33) and (218.33,276.67) ..
+(220,275) ..
+controls (221.67,273.33) and (221.67,271.67) ..
+(220,270) -- (220,262) ;
+\draw [shift={(220,260)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=0.75]    (10.93,-3.29) ..
+controls (6.95,-1.4) and (3.31,-0.3) ..
+(0,0) ..
+controls (3.31,0.3) and (6.95,1.4) ..
+(10.93,3.29)   ;
+
+%Up Arrow [id:dp7537485318007602]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=0.75]  (410,331.43) -- (420,320) -- (430,331.43) -- (423.5,331.43) -- (423.5,350) -- (416.5,350) -- (416.5,331.43) -- cycle ;
+%Rounded Same Side Corner Rect [id:dp8796153186426274]
+\draw  [line width=0.75]  (170,167.36) ..
+controls (170,161.84) and (174.48,157.36) ..
+(180,157.36) -- (220,157.36) ..
+controls (225.52,157.36) and (230,161.84) ..
+(230,167.36) -- (230,177.36) ..
+controls (230,177.36) and (230,177.36) ..
+(230,177.36) -- (170,177.36) ..
+controls (170,177.36) and (170,177.36) ..
+(170,177.36) -- cycle ;
+%Rounded Same Side Corner Rect [id:dp7576724127904904]
+\draw  [line width=0.75]  (270,110.25) ..
+controls (270,104.59) and (274.59,100) ..
+(280.25,100) -- (339.75,100) ..
+controls (345.41,100) and (350,104.59) ..
+(350,110.25) -- (350,120.5) ..
+controls (350,120.5) and (350,120.5) ..
+(350,120.5) -- (270,120.5) ..
+controls (270,120.5) and (270,120.5) ..
+(270,120.5) -- cycle ;
+%Rounded Same Side Corner Rect [id:dp9790986176556473]
+\draw  [line width=0.75]  (160,67.34) ..
+controls (160,61.68) and (164.59,57.09) ..
+(170.25,57.09) -- (229.75,57.09) ..
+controls (235.41,57.09) and (240,61.68) ..
+(240,67.34) -- (240,77.59) ..
+controls (240,77.59) and (240,77.59) ..
+(240,77.59) -- (160,77.59) ..
+controls (160,77.59) and (160,77.59) ..
+(160,77.59) -- cycle ;
 
 % Text Node
-\draw (270,185) node [scale=1.2]  {$Exp$};
+\draw (295,365) node [scale=1.2]  {$Exp$};
 % Text Node
-\draw (270,125) node [scale=1]  {$\text{Patterns}_{foo}$};
+\draw (310,305) node [scale=1]  {$\text{Patterns}_{foo}$};
 % Text Node
-\draw (370,125) node [scale=1]  {$\text{Interface}_{M}$};
+\draw (420,305) node [scale=1]  {$\text{Interface}_{M}$};
 % Text Node
-\draw (170.05,65) node [scale=1]  {$\text{HRep}_{Exp}$};
+\draw (188,91.5) node [scale=0.9]  {${H\!Rep}_{Exp}$};
 % Text Node
-\draw (370,65) node [scale=1]  {$\text{HRep}_{M}$};
-% % Text Node
-% \draw (77.5,125) node [scale=1.2]  {$Spec_{Exp}\Downarrow \ Exp$};
+\draw (309.5,60.5) node [scale=1]  {$Exp_S$};
 % Text Node
-\draw (377.5,25) node [scale=1.2]  {$Spec_{Exp}$};
+\draw (200,118.85) node [scale=0.9]  {${H\!Rep}^{\bigstar }_{Val}$};
 % Text Node
-\draw (270.22,65) node [scale=1]  {$\text{HRep}_{foo}$};
+\draw (200,188.68) node [scale=0.9]  {${H\!Rep}_{Add}$};
 % Text Node
-\draw (181.5,34) node [scale=1.44]  {$4$};
+\draw (200,238.55) node [scale=0.9]  {${H\!Rep}_{Mul}$};
 % Text Node
-\draw (284.05,34) node [scale=1.44]  {$2$};
+\draw (200,167.36) node [scale=1]  {$2$};
+% Text Node
+\draw (200,67) node [scale=1]  {$4$};
+% Text Node
+\draw (297,131.5) node [scale=0.9]  {${H\!Rep}_{foo}$};
+% Text Node
+\draw (310,110.5) node [scale=1]  {$2$};
+% Text Node
+\draw (309.75,159) node [scale=0.9]  {${H\!Rep}_{foo\#1}$};
+% Text Node
+\draw (309.54,209) node [scale=0.9]  {${H\!Rep}_{foo\#2}$};
+% Text Node
+\draw (145,310.5) node [scale=0.9]  {$\llbracket Fix\ Exp_S\rrbracket _{Exp}$};
+% Text Node
+\draw (405.5,91.5) node [scale=0.9]  {${H\!Rep}_{M}$};
+% Text Node
+\draw (420,219.55) node [scale=0.9]  {${H\!Rep}_{minus}$};
+% Text Node
+\draw (420,168.69) node [scale=0.9]  {${H\!Rep}_{square}$};
+% Text Node
+\draw (420,117.69) node [scale=0.9]  {${H\!Rep}_{ten}$};
+% Text Node
+\draw (127,160.5) node [scale=1]  {$Fix$};
 
 \end{tikzpicture}
+\hspace*{4mm}

### 2019-01-20 paper/tikz/val.tex
WIP

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.9,xscale=0.9]
%uncomment if require: \path (0,440.8571319580078); %set diagram left start at 0, and has height of 440.8571319580078

%Rounded Rect [id:dp48618213158536694]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (190,196) .. controls (190,192.69) and (192.69,190) .. (196,190) -- (254,190) .. controls (257.31,190) and (260,192.69) .. (260,196) -- (260,214) .. controls (260,217.31) and (257.31,220) .. (254,220) -- (196,220) .. controls (192.69,220) and (190,217.31) .. (190,214) -- cycle ;

%Rounded Rect [id:dp009025103452140248]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (95,236) .. controls (95,232.69) and (97.69,230) .. (101,230) -- (159,230) .. controls (162.31,230) and (165,232.69) .. (165,236) -- (165,254) .. controls (165,257.31) and (162.31,260) .. (159,260) -- (101,260) .. controls (97.69,260) and (95,257.31) .. (95,254) -- cycle ;

%Rounded Rect [id:dp9232511876637792]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (60,116) .. controls (60,112.69) and (62.69,110) .. (66,110) -- (144,110) .. controls (147.31,110) and (150,112.69) .. (150,116) -- (150,134) .. controls (150,137.31) and (147.31,140) .. (144,140) -- (66,140) .. controls (62.69,140) and (60,137.31) .. (60,134) -- cycle ;

%Rounded Rect [id:dp10823094494105145]
\draw  [color={rgb, 255:red, 255; green, 0; blue, 0 }  ,draw opacity=1 ] (30,156) .. controls (30,152.69) and (32.69,150) .. (36,150) -- (114,150) .. controls (117.31,150) and (120,152.69) .. (120,156) -- (120,174) .. controls (120,177.31) and (117.31,180) .. (114,180) -- (36,180) .. controls (32.69,180) and (30,177.31) .. (30,174) -- cycle ;

%Rounded Rect [id:dp509725264431323]
\draw  [color={rgb, 255:red, 255; green, 0; blue, 0 }  ,draw opacity=1 ] (20,196) .. controls (20,192.69) and (22.69,190) .. (26,190) -- (104,190) .. controls (107.31,190) and (110,192.69) .. (110,196) -- (110,214) .. controls (110,217.31) and (107.31,220) .. (104,220) -- (26,220) .. controls (22.69,220) and (20,217.31) .. (20,214) -- cycle ;

%Rounded Rect [id:dp3455746100780601]
\draw   (65,234) .. controls (65,231.79) and (66.79,230) .. (69,230) -- (81,230) .. controls (83.21,230) and (85,231.79) .. (85,234) -- (85,256) .. controls (85,258.21) and (83.21,260) .. (81,260) -- (69,260) .. controls (66.79,260) and (65,258.21) .. (65,256) -- cycle ;

%Rounded Rect [id:dp5622973663481885]
\draw  [color={rgb, 255:red, 65; green, 117; blue, 5 }  ,draw opacity=1 ] (140,156) .. controls (140,152.69) and (142.69,150) .. (146,150) -- (219,150) .. controls (222.31,150) and (225,152.69) .. (225,156) -- (225,174) .. controls (225,177.31) and (222.31,180) .. (219,180) -- (146,180) .. controls (142.69,180) and (140,177.31) .. (140,174) -- cycle ;

%Rounded Rect [id:dp33414100410678915]
\draw  [color={rgb, 255:red, 65; green, 117; blue, 5 }  ,draw opacity=1 ] (120,196) .. controls (120,192.69) and (122.69,190) .. (126,190) -- (174,190) .. controls (177.31,190) and (180,192.69) .. (180,196) -- (180,214) .. controls (180,217.31) and (177.31,220) .. (174,220) -- (126,220) .. controls (122.69,220) and (120,217.31) .. (120,214) -- cycle ;

%Straight Lines [id:da7735002123972574]
\draw    (140,125) -- (155,150) ;

\draw [shift={(140,125)}, rotate = 59.04] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da19830589256441322]
\draw    (125,125) -- (115,150) ;

\draw [shift={(125,125)}, rotate = 111.8] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da7234273703208924]
\draw    (95,165) -- (90,190) ;

\draw [shift={(95,165)}, rotate = 101.31] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da39520289035501]
\draw    (110,165) -- (135,190) ;

\draw [shift={(110,165)}, rotate = 45] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da4431461097782077]
\draw    (215,165) -- (215,190) ;

\draw [shift={(215,165)}, rotate = 90] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da8121818084178258]
\draw    (85,205) -- (75,230) ;

\draw [shift={(85,205)}, rotate = 111.8] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;
%Straight Lines [id:da8679389234287274]
\draw    (100,205) -- (110,230) ;

\draw [shift={(100,205)}, rotate = 68.2] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius= 2, y radius= 2]   ;

% Text Node
\draw (225,205) node   {$Con_{Val} \ 5$};
% Text Node
\draw (130,245) node   {$Con_{Val} \ 7$};
% Text Node
\draw (87,125) node   {$Con_{Add}$};
% Text Node
\draw (60,165) node   {$Pat_{foo\#1}$};
% Text Node
\draw (50,205) node   {$Pat_{foo\#2}$};
% Text Node
\draw (75,245) node   {$3$};
% Text Node
\draw (173.05,165) node   {$Fun_{square}$};
% Text Node
\draw (150,205) node   {$Fun_{ten}$};

\end{tikzpicture}

### 2019-01-20 test/Test.hs

@@ -1,0 +1,30 @@
+module Test where
+
+type Var = String
+
+data RE
+  = Empty
+  | Lit   Char
+  | Or    RE   RE
+  | Seq   RE   RE
+  | Star  RE
+  deriving Show
+
+simpl :: RE -> RE
+simpl (Star (Star re)) = Star (opt re)
+simpl (Seq re Empty)   = opt re   
+simpl (Seq Empty re)   = opt re   
+simpl x                = x
+
+try :: RE -> RE -> RE
+try f g = (f `Or` Empty) g
+
+plus :: RE -> RE
+plus re = re `Seq` (Star re)
+
+digit :: RE
+digit = '0' `Or` '1' `Or` ... `Or` 9
+
+lower :: RE
+lower = 'a' `Or` 'b' `Or` ... `Or` 'z'
+

### 2019-01-21 paper/main.tex
WIP

@@
 \usepackage{relsize}
+\usepackage{bm}
 \usetikzlibrary{fadings}
@@
 
-% This previous definition can be interpreted graphically as it is shown in the
-% Figure \ref{fig:hrep}, where curly arrows represent the structural information
-% extracted using meta-programming.
+This previous definition can be interpreted graphically as it is shown in the
+Figure \ref{fig:hrep}, where curly arrows represent the structural information
+extracted using meta-programming.
+
+\begin{figure*}[t]
+  \centering
+  \input{tikz/hrep.tex}
+  \caption{Higher-level representation of the data type \ensuremath{\Conid{Html}}, defined using
+    structural information from the function \ensuremath{\Varid{simpl}} and the abstract interface
+    of the module \ensuremath{\Conid{M}}.}
+  \label{fig:hrep}
+\end{figure*}
 
@@
 %   \centering
-%   \input{tikz/hrep.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
+%   \input{tikz/val.tex}
+%   \caption{Some caption.}
+%   \label{fig:val}
 % \end{figure}
@@
 
-\todo[inline]{Maybe we can show an example random value from the representation
+\todo[author=AM, inline]{Maybe we can show an example random value from the representation
   and its corresponding target value }

### 2019-01-21 paper/sections/hrep.tex
WIP

--- 2019-01-17 paper/sections/hrep.tex
@@
 
-In this section we present a compositional representation to express the random
-generation of values following the internal structure of their data types along
-with the structure present on patterns matchings and abstract interfaces.
+In this section we present a compositional representation of values to express
+the random generation of values following the internal structure of their data
+types along with the structure present on patterns matchings and abstract
+interfaces.
 
-The key idea of this work is to represent different sources of structured value
-information of a data type in a homogeneous way that we call a ``higher-level
-representation'' (|HRep| from now on).%
-\footnote{The notion of higher-level comes from that the generation process of
-  given target data type is entirely determined by the type of the chosen
-  representation, instead of by a concrete generator defined at the term level.}
+The key idea of this work is to represent different structured constructions of
+data in a homogeneous way that we call a ``higher-level representation'' (|HRep|
+from now on).
+%
+Instead of generating each data constructor independently, a random generator
+derived from this representation might generate composite structured values on
+each random choice it performs.
+%
+% \footnote{The notion of higher-level comes from that the generation process is
+%   entirely determined by the type of the chosen type level representation,
+%   instead of by a concrete generator defined at the term level.}
 %
@@
 level combinators in different ways into a ``generation specification'' that
-completely determines the generation process.
+completely determines the generation process behavior.
 
-We wil reuse the previously defined data type |Exp| and the function |foo| to
-explain the different concepts involved all across this section.
+We wil reuse the previously defined data type |Html| and the functions defined
+in the previous section to explain the different concepts involved all across
+this section.
 
@@
 %
-In this light, we represent each constructor of the data type |Exp| as follows:
+In this light, we represent each constructor of the data type |Html| as follows:
 
 \begin{code}
-data HRep_Val  r = Mk_Val Int
-data HRep_Add  r = Mk_Add r r
-data HRep_Mul  r = Mk_Mul r r
+data HRepText  r = ConText  String
+data HRepSing  r = ConSing  String
+data HRepJoin  r = ConJoin  r r
+data HRepTag   r = ConTag   String r
 \end{code}
@@
 %
-For instance, the value (|Mk_Add 10 20|) has type |HRep_Add Int|, while the
-value (|Mk_Mul True False|) has type |HRep_Mul Bool|.
-%
+For instance, the value (|ConJoin 10 20|) has type |HRepJoin Int|, while the
+value (|ConJoin True False|) has type |HRepJoin Bool|.
+
 In practice, this parametricity let us instatiate |r| with the type of the
@@
 Having the |HRep| of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : |HRep|_f \rightarrow t$) that maps a value from each
+($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
 representation |f| back to the target data type |t|.
@@
 \begin{alignat*}{4}
-  &\evalrep{|Mk_Val|\ n\ \ \ &&}{|Exp|}
-    &&= |Val|\ n \\
-  &\evalrep{|Mk_Add|\ x\ y   &&}{|Exp|}
-    &&= |Add|\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
-  &\evalrep{|Mk_Mul|\ x\ y   &&}{|Exp|}
-    &&= |Mul|\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
+  &\evalrep{|ConText|\ s &&}{Html}
+    &&= |Text|\ s \\
+  &\evalrep{|ConSing|\ s &&}{Html}
+    &&= |Sing|\ s \\
+  &\evalrep{|ConJoin|\ x\ y &&}{Html}
+    &&= |Join|\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
+  &\evalrep{|ConTag|\ s\ x &&}{Html}
+    &&= |Tag|\ s\ \evalrep{x}{Html}
 \end{alignat*}
 
-The last missing piece is to automatically derive a random generators for each
-constructor |HRep|.
+The missing piece is to automatically synthesize a random generators for each
+constructor representation.
 %
@@
 Given that this specification is unknown at the derivation time, we parametrize
-each |HRep| generator with a random generator |gen_r| that is used to generate
-each recursive fields:
+each |HRep| generator with a random generator |genr| that is used to generate
+random values for each recursive field:
 
 \begin{code}
-  gen_Val  gen_r  = Mk_Val  <$> arbitrary
-  gen_Add  gen_r  = Mk_Add  <$> smaller gen_r <*> smaller gen_r
-  gen_Mul  gen_r  = Mk_Mul  <$> smaller gen_r <*> smaller gen_r
-\end{code} %$
+  genText  genr  = ConText  <$> arbitrary
+  genSing  genr  = ConSing  <$> arbitrary
+  genJoin  genr  = ConJoin  <$> smaller genr <*> smaller genr
+  genTag   genr  = ConTag   <$> arbitrary <*> smaller genr
+\end{code}
 
@@
 
-In this work we define type level combinators which let the user combine the
-automatically derived |HRep|s in several ways.
-%
-In first place, we define a type combinator ($\_^\bullet$) to tag a |HRep| to be
+The next step is to define a series of type level combinators to enable us
+combining the automatically derived |HRep|s in several ways.
+
+In first place, we define a type combinator (|term|) to tag a |HRep| to be
 terminal, i.e., a representation that is allowed be generated when the
@@
 \begin{alignat*}{4}
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Term\ x &&: |f_term|    &&}{t} &&= \evalrep{x : f}{t}
+  &\evalrep{Term\ x &&: |f_term|    &&}{t} &&= \evalrep{x : f}{t} \\
+  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
 \end{alignat*}
@@
 \begin{code}
-gen_f_term     gen_r  = Term  <$> gen_f gen_r
-gen_f_times_n  gen_r  = Freq  <$> gen_f gen_r
+gen_f_term     genr  = Term  <$> genf genr
+
+gen_f_times_n  genr  = Freq  <$> genf genr
 \end{code}
@@
 A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data type variant and evaluating the
-inner |HRep|:
+target data type by pattern matching on the data constructor variant and
+evaluating the inner |HRep| accodingly:
 %
@@
 \begin{code}
-  gen_f_plus_g gen_r  = sized (\size ->
+  gen_f_plus_g genr  = sized (\size ->
     if size == 0
     then frequency
-      [ (freq0  at_f,  L  <$> gen_f  gen_r  )
-      , (freq0  at_g,  R  <$> gen_g  gen_r  ) ]
+      [ (freq0  at_f,  L  <$> genf  genr  )
+      , (freq0  at_g,  R  <$> geng  genr  ) ]
     else frequency
-      [ (freq   at_f,  L  <$> gen_f  gen_r  )
-      , (freq   at_g,  R  <$> gen_g  gen_r  ) ])
+      [ (freq   at_f,  L  <$> genf  genr  )
+      , (freq   at_g,  R  <$> geng  genr  ) ])
 \end{code} %$
 
-The previous definition we reflect the type level frequencies of |f| and |g|
-(|freq at_f| and |freq at_g|).
+In the previous definition, we reflect the type level frequencies of the types
+|f| and |g| (|freq at_f| and |freq at_g|) into term-level values.
 %
-This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
-used.
+This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
 %
@@
 
-With the introduced combinators, we can easily create a type synonym |HRep_Exp|
-to specify a generation schema equivalent to the one seen in the concrete random
-generator of type |Exp| presented in Section \ref{sec:randomtesting}:
+Later, we can use these combinators to create a type synonym |HRepHtml| that
+specifies a generation schema equivalent to the one seen in the concrete random
+generator of type |Html| presented in Section \ref{sec:randomtesting}:
 
 \begin{code}
-type HRep_Exp  =       HRep_Val_term
-               oplus'  HRep_Add  otimes 2
-               oplus'  HRep_Mul
+type HRepHtml  =       HRepText_term  otimes 2
+               oplus'  HRepSing_term
+               oplus'  HRepJoin       otimes 3
+               oplus'  HRepTag        otimes 4
 \end{code}
 
-However, a value of type |HRep_Exp| still has its recursive calls abstracted
-away---the type parameter |r| is implicit in the definition of |HRep_Exp|.
+However, a value of type |HRepHtml| still has its recursive calls abstracted
+away---the type parameter |r| is implicit at the definition of |HRepHtml|.
 %
-If we want to generate recursive values using this representation we need to
-introduce a last type level combinator to ``tie the knot'' recursively:
+We can think of it as a ``single layer'' of representation.
+%
+To make it able to represent recursive values we need to define a last type
+level combinator to ``tie the knot'':
 
@@
 \end{align*}
+
+% Categorically, this generic transformation of fixed points is called a
+% \emph{catamorphism}, where our evaluation relation is known as its
+% \emph{algebra} \tocite.
+
+Unlike the other combinators, to define a random generator for the fixed point
+of a certain representation |f|, we do not to parametrize the generation of the
+recursive fields of |f| over an external generator |genr|.
 %
-Categorically, this generic transformation of fixed points is called a
-\emph{catamorphism}, where our evaluation relation is known as its
-\emph{algebra}.
-
-Unlike the other combinators, to randomly generate a fixed point of a certain
-representation |f|, we do not parametrize the generation of the recursive fields
-of the inner represention over an external generator |gen_r|.
-%
-Instead, we can also tie the recursive knot in our random generator by piping
-the generation of the inner |HRep| parametrized recursively by the itself:
+% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
+Instead, we replace it with our fixed point generator, calling itself
+recursively on any recursive field of |f| that might appear inside:
 
 \begin{code}
-genFix_f = Fix <$> gen_f genFix_f
+genFix_f = Fix <$> genf genFix_f
 \end{code} %$
@@
 %
-With this generator, we can define a random generator for our target data type
-simply by evaluating a random value of the representation:
+Then, we can define a random generator for our target data type simply by
+generating a random value of our chosen representation, and transforming it back
+to our target by the means of the evaluation relation:
 
 \begin{code}
-  gen_Exp = do  x <- genFix_HRep_Exp
-                return eval_x_Exp
+  genHtml = do  x <- genFix_HRep_Html
+                return eval_x_Html
 \end{code} %$
+
+So far we have introduced the machinery required to represent the random
+generation of a target data type considering only the structure encoded on its
+definition.
+%
+However, this approach can now be easily extended to encode different sources of
+structured information.
+%
+We proceed to introduce two extensions that help to address the problematic
+testing scenarios presented in the previous section.
 
@@
 
+We can follow a similar reasoning as before to represent the pattern matching
+structure from a given function.
+%
+Consider the nested pattern matchings of function |simpl| defined in the
+previous section:
+
 \begin{code}
-data HRep_foo_1  r = Mk_foo_1 r r
-data HRep_foo_2  r = Mk_foo_2 Int r
+simpl (Join (Text t1) (Text t2))   = ...
+simpl (Join (Join (Text t1) x) y)  = ...
 \end{code}
 
+To represent these patterns, we derive a new data for each one of them, whose
+fields represent the pattern variables that occur inside, and abstracting away
+every pattern variable of type |Html|.
+%
+Concretely, we define the following data types:
+
+\begin{code}
+data HRepSimpl1  r = PatSimpl1 String String
+data HRepSimpl2  r = PatSimpl2 String r r
+\end{code}
+
+The first pattern of |simpl| contains two pattern variables (|t1| and |t2|) of
+type |String| that are included as fields of |PatSimpl1|.
+%
+Similarly, the second pattern of |simpl| contains a pattern variable |t1| of
+type |String| represented by the first field of |PatSimpl2|, along with two
+pattern variables (|x| and |y|) of type |Html| that are abstracted in the second
+and third fields of |PatSimpl2|.
+
+To transform these representations back to our target data type, we simply need
+to expand them it into the concrete value represented by the original pattern,
+evaluating its fields back to |Html| as well:
+%
 \begin{align*}
-  &\evalrep{|Mk_foo_1|\ x\ y}{Exp} = \\
-  &\ \ |Add (Add eval_x_Exp (Val 50)) (Add (Val 25) eval_y_Exp)|\\
-  &\evalrep{|Mk_foo_2|\ x\ y}{Exp} = \\
-  &\ \ |Mul (Val 50) (Mul (Val x) eval_y_Exp)|
+  &\evalrep{|PatSimpl1|\ t1\ t2}{Html} \\
+  &\quad = |Join (Text eval_t1_Html) (Text eval_t2_Html)|\\
+  &\evalrep{|PatSimpl2|\ t1\ x\ y}{Html} = \\
+  &\quad = |Join (Join (Text eval_t1_Html)) eval_x_Html) eval_y_Html)|
 \end{align*}
 
+This way, any value of type |HRepSimpl1| is guaranteed to satisfy the first
+pattern matching of |simpl|---the same property follows for |HRepSimpl2|.
+
+The random generation of pattern |HRep|s is defined in the same way as we did
+before for representations of data constructors:
+
 \begin{code}
-  gen_foo_1  gen_r  = Mk_foo_1  <$> smaller gen_r  <*> smaller gen_r
-  gen_foo_2  gen_r  = Mk_foo_2  <$> arbitrary      <*> smaller gen_r
+genSimpl1  genr
+  = PatSimpl1  <$> arbitrary  <*> arbitrary
+genSimpl2  genr
+  = PatSimpl2  <$> arbitrary <*> smaller genr <*> smaller genr
 \end{code} %$
 
+Finally, we can join the pattern matching representation of each clause of
+|simpl| into a single one:
+
 \begin{code}
-type HRep_foo  = HRep_foo_1 oplus''  HRep_foo_2
+type HRepSimpl  = HRepSimpl1 oplus''  HRepSimpl2
 \end{code}
@@
 
+To introduce the higher level representation of module abstract interfaces,
+consider module |M| defininf |Html| combinators introduced in the previous
+section.
+
+% \begin{code}
+% module M where
+
+% ten :: Exp
+% ten = Val 10
+
+% square :: Exp -> Exp
+% square x = Mul x x
+
+% minus :: Exp -> Exp -> Exp
+% minus x y = Add x (Mul y (Val (-1)))
+% \end{code}
+
+We can represent each function of |M| that returns a value of type |Html|
+following the same idea as before, deriving a data type with a single data
+constructor for each one of them:
+
 \begin{code}
-module M where
-
-ten :: Exp
-ten = Val 10
-
-square :: Exp -> Exp
-square x = Mul x x
-
-minus :: Exp -> Exp -> Exp
-minus x y = Add x (Mul y (Val (-1)))
+data HRepdiv        r = Fundiv r
+data HRepbold       r = Funbold r
+data HRephr         r = Funhr
 \end{code}
 
+In this case, each single constructor will have as fields the types of the
+inputs of the function that they represent.
+%
+As before, we abstract away any field representing and input of type |Html| with
+a type parameter |r|.
+
+To evaluate these representations, we simply use the values on its fields as
+input parameter of each original function, returning its result.
+%
+\begin{alignat*}{3}
+  &\evalrep{|Fundiv|\  &&x  &&}{Html} = |div eval_x_html| \\
+  &\evalrep{|Funbold|\ &&x &&}{Html} = |bold eval_x_html| \\
+  &\evalrep{|Funhr|    &&  &&}{Html} = |hr|
+\end{alignat*}
+
+Note that, by doing this, the generation process inherits any patology that the
+functions we use to generate values might have.
+%
+For instance, if the function |div| would happen to be non-terminating for some
+inputs, our generation process could suffer from this as well.
+
+Furthermore, the generation procedure for abstract interface representations
+follows the same pattern as before:
+
 \begin{code}
-data HRep_ten       r = Mk_ten
-data HRep_square    r = Mk_square   r
-data HRep_minus     r = Mk_minus    r r
+  gendiv   genr  = Fundiv <$> smaller genr
+  genbold  genr  = Funbold <$> smaller genr
+  genhr    genr  = return Funhr
+\end{code} %$
+
+And we will also define a type synonym to join all the representations of the
+module |M| into a single one.
+
+\begin{code}
+type HRepM  = HRepdiv oplus''  HRepbold oplus''  HRephr
 \end{code}
 
-\begin{alignat*}{3}
-  &\evalrep{|Mk_ten|         &&}{Exp} = |ten| \\
-  &\evalrep{|Mk_square|\ x   &&}{Exp} = |square eval_x_Exp| \\
-  &\evalrep{|Mk_minus|\ x\ y &&}{Exp} = |minus  eval_x_Exp eval_y_Exp|
-\end{alignat*}
+Finally, we can put all the derived machinery together into a generation
+specification |HtmlS|, assigning (possibly) different generation frequencies to
+each individual |HRep| we combine:
 
 \begin{code}
-  gen_ten     gen_r  = Mk_ten
-  gen_square  gen_r  = Mk_square  <$> smaller gen_r
-  gen_minus  gen_r   = Mk_minus   <$> smaller gen_r  <*> smaller gen_r
-\end{code} %$
-
-\begin{code}
-type HRep_M  = HRep_ten oplus''  HRep_square oplus''  HRep_minus
-\end{code}
-
-\begin{code}
-type Spec  =       HRep_Exp  otimes 4
-           oplus'  HRep_foo  otimes 2
-           oplus'  HRep_M
+type HtmlS  =       HRepHtml   otimes 4
+            oplus'  HRepSimpl  otimes 2
+            oplus'  HRepM
 \end{code}
@@
 This previous definition can be interpreted graphically as it is shown in the
-Figure \ref{fig:hrep}.
-%
-Curly arrows represent the structural information extracted using
-meta-programming.
+Figure \ref{fig:hrep}, where curly arrows represent the structural information
+extracted using meta-programming.
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
   \input{tikz/hrep.tex}
-  \caption{Higher level representation of the data type |Exp|, defined using
-    structural information from the function |foo| and the abstract interface of
-    the module |M|.}
+  \caption{Higher-level representation of the data type |Html|, defined using
+    structural information from the function |simpl| and the abstract interface
+    of the module |M|.}
   \label{fig:hrep}
-\end{figure}
+\end{figure*}
+
+% \begin{figure}[t]
+%   \centering
+%   \input{tikz/val.tex}
+%   \caption{Some caption.}
+%   \label{fig:val}
+% \end{figure}
+
+We want to remark that, for space reasons, we were only able to introduce the
+representation of a rather simple target data type.
+%
+In practice, this reasoning can be extended to mutually recursive and parametric
+types as well.
+
+Overall, this approach offers significant advantages over the usual type-driven
+derivation of random generators:
+%
+\begin{itemize}
+\item \textbf{Composability:} we can combine different atomic representations
+  using different structure information sources depending on what property or
+  sub-system we need to verify using randomly generated values.
+  %
+\item \textbf{Extensibility:} the developer can derive representations for new
+  sources of structure information and combine them with the existing ones
+  simply by adding them to the existing generation specification of the target
+  data type.
+  %
+\item \textbf{Predictability:} using branching processes theory, it is possible
+  to predict the average distribution of generated values in terms of number of
+  constructors.
+  %
+  This prediction is completely modular, and can be obtained for any composite
+  representation obtained using the automatically derived |HRep|s and the
+  provided combinators.
+\end{itemize}
+
+\todo[author=AM, inline]{Maybe we can show an example random value from the representation
+  and its corresponding target value }
 

### 2019-01-21 paper/tikz/hrep.tex
WIP

--- 2019-01-17 paper/tikz/hrep.tex
@@
 
-\hspace*{-4mm}
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.9,xscale=0.9]
-%uncomment if require: \path (0,367.99999618530273); %set diagram left start at 0, and has height of 367.99999618530273
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.1,xscale=1.1]
+%uncomment if require: \path (0,439.4285583496094); %set diagram left start at 0, and has height of 439.4285583496094
 
-%Shape: Rectangle [id:dp9489660675592899]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]  (110,20) ..
-controls (110,14.48) and (114.48,10) ..
-(120,10) -- (420,10) ..
-controls (425.52,10) and (430,14.48) ..
-(430,20) -- (430,220) ..
-controls (430,225.52) and (425.52,230) ..
-(420,230) -- (120,230) ..
-controls (114.48,230) and (110,225.52) ..
-(110,220) -- cycle ;
-%Straight Lines [id:da7051587981224026]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (270,250) ..
-controls (268.33,248.33) and (268.33,246.67) ..
-(270,245) ..
-controls (271.67,243.33) and (271.67,241.67) ..
-(270,240) ..
-controls (268.33,238.33) and (268.33,236.67) ..
-(270,235) ..
-controls (271.67,233.33) and (271.67,231.67) ..
-(270,230) ..
-controls (268.33,228.33) and (268.33,226.67) ..
-(270,225) ..
-controls (271.67,223.33) and (271.67,221.67) ..
-(270,220) ..
-controls (268.33,218.33) and (268.33,216.67) ..
-(270,215) ..
-controls (271.67,213.33) and (271.67,211.67) ..
-(270,210) ..
-controls (268.33,208.33) and (268.33,206.67) ..
-(270,205) -- (270,201) -- (270,193) ;
-\draw [shift={(270,190)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
-(0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
+%Right Arrow [id:dp7732858674364509]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (340,255) -- (340,195) -- (335,195) -- (345,195) -- (355,195) -- (350,195) -- (350,255) -- cycle ;
+%Right Arrow [id:dp2590973168233759]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (311,190) -- (360.31,190) -- (360.31,185) -- (376,195) -- (360.31,205) -- (360.31,200) -- (311,200) -- cycle ;
+%Shape: Rectangle [id:dp2141233912633571]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (231,125) -- (321,125) -- (321,255) -- (231,255) -- cycle ;
+%Right Arrow [id:dp06215174722463401]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (171,161.75) -- (228.69,161.75) -- (228.69,158) -- (241,165.5) -- (228.69,173) -- (228.69,169.25) -- (171,169.25) -- cycle ;
+%Right Arrow [id:dp6766910698477273]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (171,196.75) -- (228.69,196.75) -- (228.69,193) -- (241,200.5) -- (228.69,208) -- (228.69,204.25) -- (171,204.25) -- cycle ;
+%Right Arrow [id:dp3196923724356615]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (171,230.75) -- (228.69,230.75) -- (228.69,227) -- (241,234.5) -- (228.69,242) -- (228.69,238.25) -- (171,238.25) -- cycle ;
+%Right Arrow [id:dp0360549353694295]
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (431,190) -- (480.31,190) -- (480.31,185) -- (496,195) -- (480.31,205) -- (480.31,200) -- (431,200) -- cycle ;
+%Shape: Rectangle [id:dp2934677500856828]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (496,125) -- (596,125) -- (596,255) -- (496,255) -- cycle ;
+%Straight Lines [id:da7152551743774183]
+\draw    (558.5,207.5) -- (568.5,217) ;
 
-%Straight Lines [id:da6087008244886778]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (380,250) ..
-controls (378.33,248.33) and (378.33,246.67) ..
-(380,245) ..
-controls (381.67,243.33) and (381.67,241.67) ..
-(380,240) ..
-controls (378.33,238.33) and (378.33,236.67) ..
-(380,235) ..
-controls (381.67,233.33) and (381.67,231.67) ..
-(380,230) ..
-controls (378.33,228.33) and (378.33,226.67) ..
-(380,225) ..
-controls (381.67,223.33) and (381.67,221.67) ..
-(380,220) ..
-controls (378.33,218.33) and (378.33,216.67) ..
-(380,215) -- (380,211) -- (380,203) ;
-\draw [shift={(380,200)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
-(0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
+%Straight Lines [id:da3837250992572381]
+\draw    (558.5,207.5) -- (548.5,222.5) ;
 
-%Shape: Rectangle [id:dp701556783783656]
-\draw  [line width=1]  (230,255) ..
-controls (230,252.24) and (232.24,250) ..
-(235,250) -- (305,250) ..
-controls (307.76,250) and (310,252.24) ..
-(310,255) -- (310,275) ..
-controls (310,277.76) and (307.76,280) ..
-(305,280) -- (235,280) ..
-controls (232.24,280) and (230,277.76) ..
-(230,275) -- cycle ;
+%Straight Lines [id:da8680015881494354]
+\draw    (568.5,217) -- (558.08,227.11) ;
 
-%Shape: Rectangle [id:dp033430127809365384]
-\draw  [line width=1]  (340,255) ..
-controls (340,252.24) and (342.24,250) ..
-(345,250) -- (415,250) ..
-controls (417.76,250) and (420,252.24) ..
-(420,255) -- (420,275) ..
-controls (420,277.76) and (417.76,280) ..
-(415,280) -- (345,280) ..
-controls (342.24,280) and (340,277.76) ..
-(340,275) -- cycle ;
-%Up Arrow [id:dp0013771056350821986]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1]  (260,289.5) -- (270,280) -- (280,289.5) -- (273.5,289.5) -- (273.5,310) -- (266.5,310) -- (266.5,289.5) -- cycle ;
-%Shape: Rectangle [id:dp1692119124811613]
-\draw  [line width=1]  (110,310) -- (430,310) -- (430,340) -- (110,340) -- cycle ;
-%Up Arrow [id:dp4999372801455]
-\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 248; green, 231; blue, 28 }  ,fill opacity=1 ][line width=1]  (370,289.5) -- (380,280) -- (390,289.5) -- (383.5,289.5) -- (383.5,310) -- (376.5,310) -- (376.5,289.5) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (223,132.93) ..
-controls (223,129.02) and (219.83,125.86) ..
-(215.93,125.86) ..
-controls (212.02,125.86) and (208.86,129.02) ..
-(208.86,132.93) ..
-controls (208.86,136.83) and (212.02,140) ..
-(215.93,140) ..
-controls (219.83,140) and (223,136.83) ..
-(223,132.93) -- cycle ; \draw  [line width=1]  (223,132.93) -- (208.86,132.93) ; \draw  [line width=1]  (215.93,125.86) -- (215.93,140) ;
-%Snip Same Side Corner Rect [id:dp11914646356873848]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (230,67.5) -- (237,60.5) -- (303,60.5) -- (310,67.5) -- (310,80.5) -- (310,80.5) -- (230,80.5) -- (230,80.5) -- cycle ;
-%Shape: Rectangle [id:dp6865365415345861]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (230,80.5) -- (310,80.5) -- (310,190.5) -- (230,190.5) -- cycle ;
-%Shape: Rectangle [id:dp39543698275133865]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (236.29,107.66) -- (303.21,107.66) -- (303.21,130.34) -- (236.29,130.34) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (247,75.5) ..
-controls (249.76,72.74) and (249.76,68.26) ..
-(247,65.5) ..
-controls (244.24,62.74) and (239.76,62.74) ..
-(237,65.5) ..
-controls (234.24,68.26) and (234.24,72.74) ..
-(237,75.5) ..
-controls (239.76,78.26) and (244.24,78.26) ..
-(247,75.5) -- cycle ; \draw  [line width=1]  (247,75.5) -- (237,65.5) ; \draw  [line width=1]  (247,65.5) -- (237,75.5) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (277.14,143.43) ..
-controls (277.14,139.52) and (273.98,136.36) ..
-(270.07,136.36) ..
-controls (266.17,136.36) and (263,139.52) ..
-(263,143.43) ..
-controls (263,147.33) and (266.17,150.5) ..
-(270.07,150.5) ..
-controls (273.98,150.5) and (277.14,147.33) ..
-(277.14,143.43) -- cycle ; \draw  [line width=1]  (277.14,143.43) -- (263,143.43) ; \draw  [line width=1]  (270.07,136.36) -- (270.07,150.5) ;
-%Shape: Rectangle [id:dp047421885346712545]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (236.08,157.66) -- (303,157.66) -- (303,180.34) -- (236.08,180.34) -- cycle ;
-%Shape: Rectangle [id:dp0425475572348164]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (340,40) -- (420,40) -- (420,200) -- (340,200) -- cycle ;
-%Shape: Rectangle [id:dp9247208762532781]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (345,66.37) -- (415,66.37) -- (415,89) -- (345,89) -- cycle ;
-%Shape: Rectangle [id:dp8712965895636344]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (345.52,117.37) -- (414.48,117.37) -- (414.48,140) -- (345.52,140) -- cycle ;
-%Shape: Rectangle [id:dp933766425327347]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (346.66,168.24) -- (413.34,168.24) -- (413.34,190.87) -- (346.66,190.87) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (387.14,103.95) ..
-controls (387.14,100.05) and (383.98,96.89) ..
-(380.07,96.89) ..
-controls (376.17,96.89) and (373,100.05) ..
-(373,103.95) ..
-controls (373,107.84) and (376.17,111) ..
-(380.07,111) ..
-controls (383.98,111) and (387.14,107.84) ..
-(387.14,103.95) -- cycle ; \draw  [line width=1]  (387.14,103.95) -- (373,103.95) ; \draw  [line width=1]  (380.07,96.89) -- (380.07,111) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (387.14,152.95) ..
-controls (387.14,149.05) and (383.98,145.89) ..
-(380.07,145.89) ..
-controls (376.17,145.89) and (373,149.05) ..
-(373,152.95) ..
-controls (373,156.84) and (376.17,160) ..
-(380.07,160) ..
-controls (383.98,160) and (387.14,156.84) ..
-(387.14,152.95) -- cycle ; \draw  [line width=1]  (387.14,152.95) -- (373,152.95) ; \draw  [line width=1]  (380.07,145.89) -- (380.07,160) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (333,132.93) ..
-controls (333,129.02) and (329.83,125.86) ..
-(325.93,125.86) ..
-controls (322.02,125.86) and (318.86,129.02) ..
-(318.86,132.93) ..
-controls (318.86,136.83) and (322.02,140) ..
-(325.93,140) ..
-controls (329.83,140) and (333,136.83) ..
-(333,132.93) -- cycle ; \draw  [line width=1]  (333,132.93) -- (318.86,132.93) ; \draw  [line width=1]  (325.93,125.86) -- (325.93,140) ;
-%Snip Same Side Corner Rect [id:dp09931070350524518]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (120,27.48) -- (126.98,20.5) -- (193.02,20.5) -- (200,27.48) -- (200,40.45) -- (200,40.45) -- (120,40.45) -- (120,40.45) -- cycle ;
-%Shape: Rectangle [id:dp6050538306488522]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (120,40.45) -- (200,40.45) -- (200,220) -- (120,220) -- cycle ;
-%Shape: Rectangle [id:dp20494666512059623]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,67.54) -- (190,67.54) -- (190,90.17) -- (130,90.17) -- cycle ;
-%Shape: Rectangle [id:dp8741156804316592]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,137.36) -- (190,137.36) -- (190,159.99) -- (130,159.99) -- cycle ;
-%Shape: Rectangle [id:dp7972859165129229]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,187.24) -- (190,187.24) -- (190,209.87) -- (130,209.87) -- cycle ;
-%Snip Same Side Corner Rect [id:dp44067415974317603]
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (130,124.4) -- (136.98,117.41) -- (183.02,117.41) -- (190,124.4) -- (190,137.36) -- (190,137.36) -- (130,137.36) -- (130,137.36) -- cycle ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (147,132.22) ..
-controls (149.76,129.46) and (149.76,125) ..
-(147,122.24) ..
-controls (144.24,119.49) and (139.76,119.49) ..
-(137,122.24) ..
-controls (134.24,125) and (134.24,129.46) ..
-(137,132.22) ..
-controls (139.76,134.97) and (144.24,134.97) ..
-(147,132.22) -- cycle ; \draw  [line width=1]  (147,132.22) -- (137,122.24) ; \draw  [line width=1]  (147,122.24) -- (137,132.22) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (137,35.46) ..
-controls (139.76,32.71) and (139.76,28.24) ..
-(137,25.49) ..
-controls (134.24,22.73) and (129.76,22.73) ..
-(127,25.49) ..
-controls (124.24,28.24) and (124.24,32.71) ..
-(127,35.46) ..
-controls (129.76,38.22) and (134.24,38.22) ..
-(137,35.46) -- cycle ; \draw  [line width=1]  (137,35.46) -- (127,25.49) ; \draw  [line width=1]  (137,25.49) -- (127,35.46) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (167.14,103.22) ..
-controls (167.14,99.33) and (163.98,96.17) ..
-(160.07,96.17) ..
-controls (156.17,96.17) and (153,99.33) ..
-(153,103.22) ..
-controls (153,107.12) and (156.17,110.27) ..
-(160.07,110.27) ..
-controls (163.98,110.27) and (167.14,107.12) ..
-(167.14,103.22) -- cycle ; \draw  [line width=1]  (167.14,103.22) -- (153,103.22) ; \draw  [line width=1]  (160.07,96.17) -- (160.07,110.27) ;
-\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ][line width=1]  (167.14,173.05) ..
-controls (167.14,169.15) and (163.98,165.99) ..
-(160.07,165.99) ..
-controls (156.17,165.99) and (153,169.15) ..
-(153,173.05) ..
-controls (153,176.94) and (156.17,180.1) ..
-(160.07,180.1) ..
-controls (163.98,180.1) and (167.14,176.94) ..
-(167.14,173.05) -- cycle ; \draw  [line width=1]  (167.14,173.05) -- (153,173.05) ; \draw  [line width=1]  (160.07,165.99) -- (160.07,180.1) ;
-%Down Arrow [id:dp5984296172784]
-\draw  [color={rgb, 255:red, 155; green, 155; blue, 155 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]  (160,299.5) -- (165,299.5) -- (165,230) -- (175,230) -- (175,299.5) -- (180,299.5) -- (170,310) -- cycle ;
-%Straight Lines [id:da30879479079788386]
-\draw [color={rgb, 255:red, 249; green, 245; blue, 245 }  ,draw opacity=1 ][fill={rgb, 255:red, 249; green, 249; blue, 249 }  ,fill opacity=1 ][line width=1]    (166.25,230) -- (173.75,230) ;
+%Straight Lines [id:da7327748351239962]
+\draw    (548.5,222.5) -- (553.11,242.11) ;
 
-%Straight Lines [id:da8667104419281686]
-\draw [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (190,310) ..
-controls (188.33,308.33) and (188.33,306.67) ..
-(190,305) ..
-controls (191.67,303.33) and (191.67,301.67) ..
-(190,300) ..
-controls (188.33,298.33) and (188.33,296.67) ..
-(190,295) ..
-controls (191.67,293.33) and (191.67,291.67) ..
-(190,290) ..
-controls (188.33,288.33) and (188.33,286.67) ..
-(190,285) ..
-controls (191.67,283.33) and (191.67,281.67) ..
-(190,280) ..
-controls (188.33,278.33) and (188.33,276.67) ..
-(190,275) ..
-controls (191.67,273.33) and (191.67,271.67) ..
-(190,270) ..
-controls (188.33,268.33) and (188.33,266.67) ..
-(190,265) ..
-controls (191.67,263.33) and (191.67,261.67) ..
-(190,260) ..
-controls (188.33,258.33) and (188.33,256.67) ..
-(190,255) ..
-controls (191.67,253.33) and (191.67,251.67) ..
-(190,250) ..
-controls (188.33,248.33) and (188.33,246.67) ..
-(190,245) ..
-controls (191.67,243.33) and (191.67,241.67) ..
-(190,240) ..
-controls (188.33,238.33) and (188.33,236.67) ..
-(190,235) -- (190,231) -- (190,223) ;
-\draw [shift={(190,220)}, rotate = 450] [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][line width=1]    (14.21,-4.28) ..
-controls (9.04,-1.82) and (4.3,-0.39) ..
-(0,0) ..
-controls (4.3,0.39) and (9.04,1.82) ..
-(14.21,4.28)   ;
+%Straight Lines [id:da007641890318750955]
+\draw    (568.5,217) -- (566.5,235) ;
+
+%Straight Lines [id:da995333968066006]
+\draw    (568.11,146.11) -- (578.11,157.89) ;
+
+%Straight Lines [id:da015868138489393502]
+\draw    (568.11,146.11) -- (562.5,157) ;
+
+%Straight Lines [id:da7302160362706243]
+\draw    (562.5,157) -- (574.5,168.5) ;
+
+%Straight Lines [id:da46726020734917606]
+\draw    (562.5,157) -- (559.5,170.5) ;
+
+%Straight Lines [id:da4868124795808251]
+\draw    (574.5,168.5) -- (571.5,181) ;
+
+%Straight Lines [id:da6000768704792225]
+\draw    (521.89,188.11) -- (529.5,196.5) ;
+
+%Straight Lines [id:da9056461286434507]
+\draw    (521.89,188.11) -- (518.89,202.11) ;
+
+%Straight Lines [id:da5328270203547492]
+\draw    (529,178.13) -- (533.5,188) ;
+
+%Straight Lines [id:da38335809761139505]
+\draw    (521.5,165.63) -- (528.5,177.5) ;
+
+%Straight Lines [id:da39066239946769477]
+\draw    (513.11,177.11) -- (521.89,188.11) ;
+
+%Straight Lines [id:da6974822807110195]
+\draw    (513.11,177.11) -- (511.5,190.5) ;
+
+%Straight Lines [id:da06361852802164258]
+\draw    (521.5,165.63) -- (513.11,177.11) ;
+
+%Snip Single Corner Rect [id:dp4647891928859724]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (55,125) -- (161.34,125) -- (175,138.66) -- (175,255) -- (55,255) -- cycle ;
+%Shape: Rectangle [id:dp3811163355625795]
+\draw  [fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (376.5,125) -- (436.5,125) -- (436.5,255) -- (376.5,255) -- cycle ;
+%Shape: Ellipse [id:dp7838604577819246]
+\draw   (385,238.16) ..
+controls (385,231.62) and (393.95,226.32) ..
+(405,226.32) ..
+controls (416.05,226.32) and (425,231.62) ..
+(425,238.16) ..
+controls (425,244.7) and (416.05,250) ..
+(405,250) ..
+controls (393.95,250) and (385,244.7) ..
+(385,238.16) -- cycle ;
+%Shape: Circle [id:dp36073479073960346]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (408.16,230.53) ..
+controls (408.16,229.36) and (409.1,228.42) ..
+(410.26,228.42) ..
+controls (411.43,228.42) and (412.37,229.36) ..
+(412.37,230.53) ..
+controls (412.37,231.69) and (411.43,232.63) ..
+(410.26,232.63) ..
+controls (409.1,232.63) and (408.16,231.69) ..
+(408.16,230.53) -- cycle ;
+%Shape: Circle [id:dp8080098028480491]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (397.53,245.16) ..
+controls (397.53,244) and (398.47,243.05) ..
+(399.63,243.05) ..
+controls (400.79,243.05) and (401.74,244) ..
+(401.74,245.16) ..
+controls (401.74,246.32) and (400.79,247.26) ..
+(399.63,247.26) ..
+controls (398.47,247.26) and (397.53,246.32) ..
+(397.53,245.16) -- cycle ;
+%Shape: Circle [id:dp12076525384098513]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (415.37,240.95) ..
+controls (415.37,239.78) and (416.31,238.84) ..
+(417.47,238.84) ..
+controls (418.64,238.84) and (419.58,239.78) ..
+(419.58,240.95) ..
+controls (419.58,242.11) and (418.64,243.05) ..
+(417.47,243.05) ..
+controls (416.31,243.05) and (415.37,242.11) ..
+(415.37,240.95) -- cycle ;
+%Shape: Ellipse [id:dp5212893057388461]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (401.53,232.32) ..
+controls (401.53,231.15) and (402.47,230.21) ..
+(403.63,230.21) ..
+controls (404.79,230.21) and (405.74,231.15) ..
+(405.74,232.32) ..
+controls (405.74,233.48) and (404.79,234.42) ..
+(403.63,234.42) ..
+controls (402.47,234.42) and (401.53,233.48) ..
+(401.53,232.32) -- cycle ;
+%Shape: Circle [id:dp23843560123656338]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (406.79,241) ..
+controls (406.79,239.84) and (407.73,238.89) ..
+(408.89,238.89) ..
+controls (410.06,238.89) and (411,239.84) ..
+(411,241) ..
+controls (411,242.16) and (410.06,243.11) ..
+(408.89,243.11) ..
+controls (407.73,243.11) and (406.79,242.16) ..
+(406.79,241) -- cycle ;
+%Shape: Ellipse [id:dp4809698571086918]
+\draw   (385,197.5) ..
+controls (385,190.6) and (393.95,185) ..
+(405,185) ..
+controls (416.05,185) and (425,190.6) ..
+(425,197.5) ..
+controls (425,204.4) and (416.05,210) ..
+(405,210) ..
+controls (393.95,210) and (385,204.4) ..
+(385,197.5) -- cycle ;
+%Shape: Triangle [id:dp24226788884090933]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (392.5,195) -- (395,200) -- (390,200) -- cycle ;
+%Shape: Triangle [id:dp5291497962409548]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (402.5,190) -- (405,195) -- (400,195) -- cycle ;
+%Shape: Triangle [id:dp5917974674439683]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (412.5,190) -- (415,195) -- (410,195) -- cycle ;
+%Shape: Triangle [id:dp29519879870465426]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (405,200) -- (407.5,205) -- (402.5,205) -- cycle ;
+%Shape: Triangle [id:dp6574937894365762]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (415,199) -- (417.5,204) -- (412.5,204) -- cycle ;
+%Shape: Ellipse [id:dp8770682879482239]
+\draw   (385,157.81) ..
+controls (385,150.74) and (394.14,145) ..
+(405.42,145) ..
+controls (416.69,145) and (425.83,150.74) ..
+(425.83,157.81) ..
+controls (425.83,164.89) and (416.69,170.63) ..
+(405.42,170.63) ..
+controls (394.14,170.63) and (385,164.89) ..
+(385,157.81) -- cycle ;
+%Shape: Circle [id:dp5633637507570923]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (392.79,236.89) ..
+controls (392.79,235.73) and (393.73,234.79) ..
+(394.89,234.79) ..
+controls (396.06,234.79) and (397,235.73) ..
+(397,236.89) ..
+controls (397,238.06) and (396.06,239) ..
+(394.89,239) ..
+controls (393.73,239) and (392.79,238.06) ..
+(392.79,236.89) -- cycle ;
+%Shape: Rectangle [id:dp853277524439841]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (397,160.63) -- (402,160.63) -- (402,164.63) -- (397,164.63) -- cycle ;
+%Shape: Rectangle [id:dp26443574867732256]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (393,150.63) -- (398,150.63) -- (398,154.63) -- (393,154.63) -- cycle ;
+%Shape: Rectangle [id:dp3935753971313367]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (410,148.63) -- (415,148.63) -- (415,152.63) -- (410,152.63) -- cycle ;
+%Shape: Rectangle [id:dp5931653318550618]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (410,162.63) -- (415,162.63) -- (415,166.63) -- (410,166.63) -- cycle ;
+%Shape: Rectangle [id:dp7233342352369083]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (403,154.63) -- (408,154.63) -- (408,158.63) -- (403,158.63) -- cycle ;
+%Shape: Ellipse [id:dp9722655375449147]
+\draw   (540.17,225) ..
+controls (540.17,211.19) and (548.19,200) ..
+(558.08,200) ..
+controls (567.98,200) and (576,211.19) ..
+(576,225) ..
+controls (576,238.81) and (567.98,250) ..
+(558.08,250) ..
+controls (548.19,250) and (540.17,238.81) ..
+(540.17,225) -- cycle ;
+%Shape: Ellipse [id:dp9371514401907413]
+\draw  [color={rgb, 255:red, 74; green, 74; blue, 74 }  ,draw opacity=1 ] (504,184) ..
+controls (504,170.19) and (512.02,159) ..
+(521.92,159) ..
+controls (531.81,159) and (539.83,170.19) ..
+(539.83,184) ..
+controls (539.83,197.81) and (531.81,209) ..
+(521.92,209) ..
+controls (512.02,209) and (504,197.81) ..
+(504,184) -- cycle ;
+%Shape: Ellipse [id:dp26883422845191607]
+\draw   (551,165) ..
+controls (551,151.19) and (559.02,140) ..
+(568.92,140) ..
+controls (578.81,140) and (586.83,151.19) ..
+(586.83,165) ..
+controls (586.83,178.81) and (578.81,190) ..
+(568.92,190) ..
+controls (559.02,190) and (551,178.81) ..
+(551,165) -- cycle ;
+%Shape: Rectangle [id:dp5872203962807194]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (519,163.63) -- (524,163.63) -- (524,167.63) -- (519,167.63) -- cycle ;
+%Shape: Circle [id:dp7605266721377453]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (511,177.11) ..
+controls (511,175.94) and (511.94,175) ..
+(513.11,175) ..
+controls (514.27,175) and (515.21,175.94) ..
+(515.21,177.11) ..
+controls (515.21,178.27) and (514.27,179.21) ..
+(513.11,179.21) ..
+controls (511.94,179.21) and (511,178.27) ..
+(511,177.11) -- cycle ;
+%Shape: Triangle [id:dp4125869850828281]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (511.5,188) -- (514,193) -- (509,193) -- cycle ;
+%Shape: Triangle [id:dp09394805352485558]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (528.5,175) -- (531,180) -- (526,180) -- cycle ;
+%Shape: Circle [id:dp7834832784546726]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (516.79,202.11) ..
+controls (516.79,200.94) and (517.73,200) ..
+(518.89,200) ..
+controls (520.06,200) and (521,200.94) ..
+(521,202.11) ..
+controls (521,203.27) and (520.06,204.21) ..
+(518.89,204.21) ..
+controls (517.73,204.21) and (516.79,203.27) ..
+(516.79,202.11) -- cycle ;
+%Shape: Rectangle [id:dp18557164533567394]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (531,186) -- (536,186) -- (536,190) -- (531,190) -- cycle ;
+%Shape: Circle [id:dp8025931154410937]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (519.79,188.11) ..
+controls (519.79,186.94) and (520.73,186) ..
+(521.89,186) ..
+controls (523.06,186) and (524,186.94) ..
+(524,188.11) ..
+controls (524,189.27) and (523.06,190.21) ..
+(521.89,190.21) ..
+controls (520.73,190.21) and (519.79,189.27) ..
+(519.79,188.11) -- cycle ;
+%Shape: Triangle [id:dp1239939418024707]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (529.5,194) -- (532,199) -- (527,199) -- cycle ;
+%Shape: Circle [id:dp8812473863475445]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (566,146.11) ..
+controls (566,144.94) and (566.94,144) ..
+(568.11,144) ..
+controls (569.27,144) and (570.21,144.94) ..
+(570.21,146.11) ..
+controls (570.21,147.27) and (569.27,148.21) ..
+(568.11,148.21) ..
+controls (566.94,148.21) and (566,147.27) ..
+(566,146.11) -- cycle ;
+%Shape: Circle [id:dp7598993443376705]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (576,157.89) ..
+controls (576,156.73) and (576.94,155.79) ..
+(578.11,155.79) ..
+controls (579.27,155.79) and (580.21,156.73) ..
+(580.21,157.89) ..
+controls (580.21,159.06) and (579.27,160) ..
+(578.11,160) ..
+controls (576.94,160) and (576,159.06) ..
+(576,157.89) -- cycle ;
+%Shape: Rectangle [id:dp3435681036136593]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (560,155) -- (565,155) -- (565,159) -- (560,159) -- cycle ;
+%Shape: Triangle [id:dp4971718332882904]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (559.5,168) -- (562,173) -- (557,173) -- cycle ;
+%Shape: Triangle [id:dp8772068767151546]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (574.5,166) -- (577,171) -- (572,171) -- cycle ;
+%Shape: Rectangle [id:dp5389313915531764]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (569,179) -- (574,179) -- (574,183) -- (569,183) -- cycle ;
+%Shape: Triangle [id:dp9780900800848462]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (558.5,205) -- (561,210) -- (556,210) -- cycle ;
+%Shape: Triangle [id:dp6412646282331236]
+\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (548.5,220) -- (551,225) -- (546,225) -- cycle ;
+%Shape: Rectangle [id:dp6641116065775097]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (566,215) -- (571,215) -- (571,219) -- (566,219) -- cycle ;
+%Shape: Circle [id:dp09340888622817034]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (555.98,227.11) ..
+controls (555.98,225.94) and (556.92,225) ..
+(558.08,225) ..
+controls (559.25,225) and (560.19,225.94) ..
+(560.19,227.11) ..
+controls (560.19,228.27) and (559.25,229.21) ..
+(558.08,229.21) ..
+controls (556.92,229.21) and (555.98,228.27) ..
+(555.98,227.11) -- cycle ;
+%Shape: Circle [id:dp14489754689178413]
+\draw  [color={rgb, 255:red, 0; green, 0; blue, 0 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 0; blue, 0 }  ,fill opacity=1 ] (551,242.11) ..
+controls (551,240.94) and (551.94,240) ..
+(553.11,240) ..
+controls (554.27,240) and (555.21,240.94) ..
+(555.21,242.11) ..
+controls (555.21,243.27) and (554.27,244.21) ..
+(553.11,244.21) ..
+controls (551.94,244.21) and (551,243.27) ..
+(551,242.11) -- cycle ;
+%Shape: Rectangle [id:dp1568842876109131]
+\draw  [fill={rgb, 255:red, 65; green, 117; blue, 5 }  ,fill opacity=1 ] (564,233) -- (569,233) -- (569,237) -- (564,237) -- cycle ;
+%Straight Lines [id:da9505573863907721]
+\draw  [dash pattern={on 0.84pt off 2.51pt}]  (55,210) -- (175,210) ;
+
+%Straight Lines [id:da9881747537161369]
+\draw  [dash pattern={on 0.84pt off 2.51pt}]  (55,174) -- (175,174) ;
+
+%Shape: Rectangle [id:dp7912203503474478]
+\draw  [color={rgb, 255:red, 65; green, 117; blue, 5 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (241,155) -- (311,155) -- (311,175) -- (241,175) -- cycle ;
+%Shape: Rectangle [id:dp4628817838994279]
+\draw  [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (241,225) -- (311,225) -- (311,245) -- (241,245) -- cycle ;
+
+%Shape: Rectangle [id:dp14183247266409804]
+\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ][fill={rgb, 255:red, 255; green, 255; blue, 255 }  ,fill opacity=1 ] (241,190) -- (311,190) -- (311,210) -- (241,210) -- cycle ;
+
+%Straight Lines [id:da20016875561182368]
+\draw [color={rgb, 255:red, 226; green, 226; blue, 226 }  ,draw opacity=1 ][line width=1.5]    (340.5,200) -- (349.5,200) ;
 
 % Text Node
-\draw (270,323) node [scale=1.2]  {$Exp$};
+\draw (85,133.5) node [scale=0.8]  {$\text{\textbf{module}}\ M$};
 % Text Node
-\draw (270,265) node [scale=1]  {$\text{Patterns}_{foo}$};
+\draw (100,159.5) node [scale=0.8]  {$\text{\textbf{data}}\ Html\ =\ ... $};
 % Text Node
-\draw (380,265) node [scale=1]  {$\text{Interface}_{M}$};
+\draw (107,219.5) node [scale=0.8]  {$div::Html\rightarrow Html$};
 % Text Node
-\draw (148,51.5) node [scale=0.9]  {${H\!Rep}_{Exp}$};
+\draw (109,233.5) node [scale=0.8]  {$bold::Html\rightarrow Html$};
 % Text Node
-\draw (270,27.5) node [scale=1]  {$Spec$};
+\draw (86,246.5) node [scale=0.8]  {$hr::Html$};
 % Text Node
-\draw (160,78.85) node [scale=0.9]  {${H\!Rep}_{Val}^\bigstar$};
+\draw (275.5,135.5) node [scale=0.8]  {$Html_{Spec}$};
 % Text Node
-\draw (160,148.68) node [scale=0.9]  {${H\!Rep}_{Add}$};
+\draw (276,165) node [scale=0.8]  {$Html_{ADT}$};
 % Text Node
-\draw (160,198.55) node [scale=0.9]  {${H\!Rep}_{Mul}$};
+\draw (276,200) node [scale=0.8]  {$Html_{Patterns}$};
 % Text Node
-\draw (160,127.39) node [scale=1]  {$2$};
+\draw (276,235) node [scale=0.8]  {$Html_{Interface}$};
 % Text Node
-\draw (160,30.47) node [scale=1]  {$4$};
+\draw (276,182.5) node   {$\boldsymbol{\oplus}$};
 % Text Node
-\draw (257,91.5) node [scale=0.9]  {${H\!Rep}_{foo}$};
+\draw (276,217.5) node   {$\boldsymbol{\oplus}$};
 % Text Node
-\draw (270,70.5) node [scale=1]  {$2$};
+\draw (405.5,133.5) node [scale=0.8]  {$Generator$};
 % Text Node
-\draw (269.75,119) node [scale=0.9]  {${H\!Rep}_{foo\#1}$};
+\draw (406,178) node   {$\boldsymbol{\oplus}$};
 % Text Node
-\draw (269.54,169) node [scale=0.9]  {${H\!Rep}_{foo\#2}$};
+\draw (406,217.5) node   {$\boldsymbol{\oplus}$};
 % Text Node
-\draw (120,269.5) node [scale=0.9] {$\llbracket Fix\ Spec \rrbracket_{Exp}$};
+\draw (514.5,133.5) node [scale=0.8]  {$Html$};
 % Text Node
-\draw (365.5,51.5) node [scale=0.9]  {${H\!Rep}_{M}$};
+\draw (100,184.5) node [scale=0.8]  {$simpl\ ( ...) \ =\ ... \ $};
 % Text Node
-\draw (380,179.55) node [scale=0.9]  {${H\!Rep}_{minus}$};
+\draw (100,198.5) node [scale=0.8]  {$simpl\ ( ...) \ =\ ... \ $};
 % Text Node
-\draw (380,128.69) node [scale=0.9]  {${H\!Rep}_{square}$};
+\draw (202.5,142.5) node [scale=0.7] [align=left] {\textbf{Structure}\\ \textbf{reification}};
 % Text Node
-\draw (380,77.69) node [scale=0.9]  {${H\!Rep}_{ten}$};
+\draw (348.5,169.5) node [scale=0.7] [align=left] {\textbf{Generator}\\ \textbf{derivation}};
+% Text Node
+\draw (466,168.5) node [scale=0.7] [align=left] {\textbf{Random} \\ \textbf{generation}};
+% Text Node
+\draw (346.5,272.5) node [scale=0.7] [align=left] {\textbf{User desired} \\ \textbf{distribution}};
 
 \end{tikzpicture}
-\hspace*{4mm}

### 2019-01-22 paper/introduction.tex
No directory for section, it breaks all nice features of emacs and bibligraphy

--- 2019-01-17 paper/main.tex
@@
-\documentclass[conference, fleqn]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-% Local Variables:
-% TeX-mbigstarer: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usepackage{relsize}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITE}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\newcommand{\evalrep}[2]{\ensuremath{\llbracket #1 \rrbracket_{#2}}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some parameters
-\setlength{\mathindent}{1.5\parindent}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Deriving Random Generators via\\ Higher-Level Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
 \section{Introduction}
@@
 Nam euismod tellus id erat\cite{grieco2017}.
+
+We remark that, altough this work focuses on Haskell data types, this technique
+is general enough to be applied to most programming languages with some level of
+support for composite types.
 
@@
 \end{itemize}
-\section{Type-Driven Random Generation} \label{sec:randomtesting}
-
-In this section we briefly describe the common technique used for automatically
-deriving random data generators in Haskell using a type-driven approach.
-%
-Then, we introduce its main drawbacks by considering two scenarios where this
-technique gives poor results in practice.
-%
-We remark that, altough this work makes focus on Haskell data types, this
-technique is general enough to be applied to most programming languages with
-some level of support for composite types.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-properties of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types.
-%
-In the most basic conception, types can be combined by \textbf{sums} (also known
-as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
-(or tuples) of other data types.
-%
-To exemplify this, consider the following definition of the data type \ensuremath{\Conid{Exp}}
-encoding integer expressions:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{25}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Exp}{}\<[11]%
-\>[11]{}\mathrel{=}{}\<[15]%
-\>[15]{}\Conid{Val}\;{}\<[20]%
-\>[20]{}\Conid{Int}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Add}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\\
-\>[11]{}\;\;\vert\;{}\<[15]%
-\>[15]{}\Conid{Mul}\;{}\<[20]%
-\>[20]{}\Conid{Exp}\;{}\<[25]%
-\>[25]{}\Conid{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Exp}} as the sum of three possible
-classes of values: \ensuremath{\Conid{Val}} represents literal values, whereas \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-represent the addition and multiplication of two integer expressions,
-respectively.
-%
-In Haskell, \ensuremath{\Conid{Val}}, \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}} are called \emph{data constructors} (or
-constructors for short) and are used to distiguish which variant of the data
-type we are reffering to.
-%
-Each data constructor is then defined as a product of zero or more types known
-as \emph{fields}.
-%
-In particular, \ensuremath{\Conid{Val}} contains a field of type \ensuremath{\Conid{Int}}, while \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}
-contain two fields of type \ensuremath{\Conid{Exp}} representing the operands of each operation.
-%
-Note that \ensuremath{\Conid{Exp}} is used as a field of at least one of its data constructors
-(case \ensuremath{\Conid{Add}} and \ensuremath{\Conid{Mul}}), making it a recursively defined ADT.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
-using \ensuremath{\Conid{Exp}} data constructors as \ensuremath{(\Conid{Add}\;(\Conid{Val}\;\mathrm{2})\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{5})\;(\Conid{Val}\;\mathrm{6})))}.
-%
-Furthermore, an evaluation function from \ensuremath{\Conid{Exp}}s to integer values can be defined
-very idiomatically:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{27E}{@{}l@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{eval}\mathbin{::}\Conid{Exp}\to \Conid{Int}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Val}\;\Varid{n}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{n}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Add}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{+}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{eval}\;(\Conid{Mul}\;\Varid{x}\;\Varid{y}){}\<[17]%
-\>[17]{}\mathrel{=}\Varid{eval}\;\Varid{x}{}\<[27]%
-\>[27]{}\mathbin{*}{}\<[27E]%
-\>[30]{}\Varid{eval}\;\Varid{y}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{eval}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-For the input case of a literal values, we simply return the value contained in
-the \ensuremath{\Conid{Val}} constructor.
-%
-On the other hand, if the input value matches either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}} data
-constructor, where \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} are \emph{pattern variables} that match any
-subexpression, we recursively evaluate these subexpressions, combining them with
-the apropiate operation on each case.
-
-\subsection*{\textbf{Type-Driven Generation of Random Values}}
-
-In order to perform random testing of a Haskell codebase involving user defined
-data types, most approaches require the user to provide a random data generator
-for each one of them.
-%
-This tends to be a cumbersome and error prone task that closely follows the data
-type structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Exp}}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{6E}{@{}l@{}}%
-\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{Exp}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{then}\;\Conid{Val}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{1},\Conid{Val}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{2},\Conid{Add}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},\Conid{Mul}{}\<[18]%
-\>[18]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Exp}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Exp})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
-paremetrize the generation process up to an external integer number known as the
-\emph{generation size}.
-%
-This parameter is used to limit the maximum amount of recursive call that this
-random generator can perform.
-%
-When called with a positive generation size (case \ensuremath{\Varid{gen}\;\Varid{n}}), the generator will
-pick among any data constructor of \ensuremath{\Conid{Exp}} with a explicitly given frequency---two
-times more \ensuremath{\Conid{Add}}s than \ensuremath{\Conid{Val}} or \ensuremath{\Conid{Mul}}.
-%
-In the case it picks to generate a \ensuremath{\Conid{Val}} data constructor, it also generates a
-random \ensuremath{\Conid{Int}} value using the standard overloaded generator \ensuremath{\Varid{arbitrary}}
-(\quickcheck provides standard random generators for most base types like \ensuremath{\Conid{Int}},
-\ensuremath{\Conid{Bool}}, etc.).
-%
-On the other hand, when it randomly picks to generate either an \ensuremath{\Conid{Add}} or a \ensuremath{\Conid{Mul}}
-data constructor, it also generates independently two random subexpressions
-corresponding to the data constructor fields by calling itself recursively
-(\ensuremath{\Varid{resize}\;(n{-}1)\;gen_{Exp}}), decreasing the generation size on each recursive
-invocation.
-
-This procedure keeps calling itself recursively until the generation size
-reaches zero (case \ensuremath{\Varid{gen}\;\mathrm{0}}), where the generator is constrained to pick only
-among terminal data constructors, being \ensuremath{\Conid{Val}} the only possible choice in our
-particular case.
-%
-Strictly decreasing the the generation size by one on each recursive call
-results in a useful property over the generated data: every generated value has
-at most \ensuremath{\Varid{n}} levels, where \ensuremath{\Varid{n}} is the generation size set by the user.
-%
-This property enables us to model the generation process using the theory of
-branching processes introduced by \tocite, and extended in this work as
-described in Section \ref{sec:synthesis}.
-
-Having discussed the previous random generator definition, it easy to understand
-how to extend this generation procedure to any data type defined in an algebraic
-fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth.
-%
-Given the name of the target data type, it synthesizes a random generator for it
-that behaves pretty much like the previously defined one.
-%
-However, picking among different data constructors with uniform probability can
-lead to a generation process biased to generate (in average) very small values,
-regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool considers the scenario where each one can be generated following a
-different (although fixed over time) generation probability.
-%
-Then, this tool uses the theory of branching processes to analitically predict
-the average distribution of data constructors generated on each random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation probability of each data constructor in
-order to obtain a particular distribution of values that can be specified by the
-user, offering a more flexible testing environment while still being mostly
-automated.
-
-Both \megadeth and \dragen synthesize random generators that are theoretically
-capable to generate the whole space of values of the target data type, provided
-that we set a generation size suficiently big.
-%
-However, the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is too generic to generate complex data that is
-useful enough to be used for random testing.
-
-In this work we identify two sources of additional structure information which
-are not considered by the aforementioned automatic derivation tools to obtain
-random data generators:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching very specific
-  patterns.
-\item The target code yields part of the responsability of preserving its
-  invariants to the functions of its abstract interface.
-\end{enumerate}
-
-This information can be reified in compile time and used to synthesize richer
-random generators automatically. We proceed to exemplify the previous points in
-detail.
-
-\subsection*{\textbf{Presence of Complex Pattern Matchings}}
-
-To exemplify the firts problematic scenarion that may arise, suppose we want to
-use randomly generated \ensuremath{\Conid{Exp}}s to test a property comprising the following
-function:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{foo}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Add}\;(\Conid{Add}\;\Varid{x}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#1\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;(\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\Varid{y})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{error}\;\text{\ttfamily \char34 pattern~\#2\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{foo}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function behaves very much like an identity function, with the exception
-that it fails for two very specific patterns of input values defined using
-nested pattern matching.
-%
-This is, \ensuremath{\Varid{foo}} does not only matches against the root data constructor, but also
-against the data constructors of its subexpressions and sub-subexpressions.
-
-\quickcheck's default implementation for random \ensuremath{\Conid{Int}}s pick a number in the
-interval $[-n, n]$ with uniform distribution.
-%
-Hence, in principle we need to recognize what is a suitable generation size,
-which should be big enough for our generator be able to generate the numbers we
-pattern match against to, otherwise we will not be able to generate any value to
-match against \ensuremath{\Varid{foo}} clauses, leaving fragments of code completely untested.
-%
-Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-is big enough to produce an \ensuremath{\Conid{Int}} number equal to $50$.
-%
-Under this consideration, the probability of generating a value matching the
-first clause of \ensuremath{\Varid{foo}} (and hence triggering the first error) results as follows:
-%
-\begin{align*}
-  &P(match(foo\#1))\\
-  &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
-  &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
-\end{align*}
-
-If we use \megadeth to automatically derive a random generator for \ensuremath{\Conid{Exp}}, we
-obtain a uniform generation probability distribution over constructors, i.e.,
-$P(\ensuremath{\Conid{Val}}) = P(\ensuremath{\Conid{Add}}) = P(\ensuremath{\Conid{Mul}}) = 1/3$.
-%
-In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-average, we will need to generate over than two million test cases in order to
-be able to test the first clause of \ensuremath{\Varid{foo}} only once.
-%
-This situation can be somewhat improved if we use \dragen to obtain a random
-generator.
-%
-Using this tool, we can optimize the generation probabilities in order to
-benefit the generation of some data constructors over the rest.
-%
-Considering that the first pattern matching of \ensuremath{\Varid{foo}} involves the data
-constructor \ensuremath{\Conid{Add}} as the only recursive one, we can set an target generation
-proportion of \ensuremath{\Conid{Add}} data constructors of, for instance, $20:1$ with respect to
-the rest of the generated data constructors.
-%
-By doing so, the obtained distribution of values results such that
-$P(match(foo\#1)) \approx 1/300000$.
-%
-Although this certainly improves the probability of generating a matching value,
-this probability is not substantial enough to become practical.
-
-Additionally, by favoring the generation probabilities towards the \ensuremath{\Conid{Add}}
-constructor, we found that the probability of generating a value matching the
-second clause of \ensuremath{\Varid{foo}} (which does not matches against it) also gets diminished
-into an impractival value.
-
-Despite the fact that the previous example might look artificial, branching
-against specific patterns of the input data is a common task.
-%
-For instance, the balancing function of a Red-Black tree need to consider a
-quite specific combination of color, left and right subtrees and sub-subtrees
-in order to preserve the height invariant.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few.
-
-\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
-
-A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions that manipulates
-its values.
-%
-Consider for instance the following possible implementation of a basic \ensuremath{\Conid{HTML}}
-manipulation library:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Html}\mathrel{=}\Conid{Html}\;\Conid{String}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{head}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{head}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <head>\char34}\plus \Varid{inner}\plus \text{\ttfamily \char34 </head>\char34}){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{body}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{body}\;(\Conid{Html}\;\Varid{inner}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Html}\;(\text{\ttfamily \char34 <body>\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 </body>\char34}){}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{br}\mathbin{::}\Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{br}\mathrel{=}\Conid{Html}\;\text{\ttfamily \char34 <br~/>\char34}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}(\mathop{\langle\! +\! \rangle})\mathbin{::}\Conid{Html}\to \Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}(\Conid{Html}\;\Varid{x})\mathop{\langle\! +\! \rangle}(\Conid{Html}\;\Varid{y})\mathrel{=}\Conid{Html}\;(\Varid{x}\plus \Varid{y}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, the \ensuremath{\Conid{Html}} data type is defined using a single data
-constructor that contains the textual representation of the Html code it
-represents in plain text.
-%
-Later, the combinator functions defined over \ensuremath{\Conid{HTML}} are the ones in charge of
-transforming this plain text representation with the invariant that, given valid
-\ensuremath{\Conid{Html}} parameters, they always return a valid \ensuremath{\Conid{Html}} value.
-%
-This is a very commmon programming pattern that can be found in a variety of
-Haskell libraries \tocite{}.
-
-If we use a standard type-driven approach to derive a random generator for
-\ensuremath{\Conid{Html}}, the only ``structural'' information that we can use in the derivation
-process is that \ensuremath{\Conid{Html}} values are composed of strings.
-%
-As a consequence, we essentially end up generating random strings, which rarely
-represents a valid (or at least well structured) Html value.
-%
-Additionally, if our test suite contains properties constrained by certain
-preconditions, this lack of domain knowleadge may lead in an impractically high
-discard ratio of randomly generated test.
-%
-For instance, suposse we write a property to verify that \ensuremath{\Varid{head}} preserves the
-invariant previously mentioned:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{prop\char95 head\char95 inv}\mathbin{::}\Conid{Html}\to \Conid{Bool}{}\<[E]%
-\\
-\>[B]{}\Varid{prop\char95 head\char95 inv}\;\Varid{x}\mathrel{=}\Varid{valid}\;\Varid{x}\Longrightarrow\Varid{valid}\;(\Varid{head}\;\Varid{x}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition we state that \ensuremath{\Varid{head}} always returns a valid \ensuremath{\Conid{Html}}
-value (\ensuremath{\Varid{valid}\;(\Varid{head}\;\Varid{x})}), provided that we are given a valid \ensuremath{\Conid{Html}} as input
-(\ensuremath{\Varid{valid}\;\Varid{x}}).
-%
-Then, while testing this property we rarely satisfy its precondition, in which
-case \quickcheck discards the whole test, retrying with a diferent random input,
-degradating this way the testing performance.
-
-We believe that, if certain patterns of values are relevant enough to appear in
-the codebase being tested, a practical random testing methodology should be able
-to produce values satisfying such patterns in a substantial proportion.
-%
-The next section introduces a reprentation model that let us encode the
-structure information presented here into our automatically derived random
-generators in a modular and flexible way.
-\section{Extracting Structure} \label{sec:hrep}
-
-In this section we present a compositional representation to express the random
-generation of values following the internal structure of their data types along
-with the structure present on patterns matchings and abstract interfaces.
-
-The key idea of this work is to represent different sources of structured value
-information of a data type in a homogeneous way that we call a ``higher-level
-representation'' (\ensuremath{\text{HRep}} from now on).%
-\footnote{The notion of higher-level comes from that the generation process of
-  given target data type is entirely determined by the type of the chosen
-  representation, instead of by a concrete generator defined at the term level.}
-%
-For this purpose, we use a series of automatically derived data types, each one
-representing an atomic unit of information that can be randomly generated and
-then reflected back to the corresponding value of the original data type.
-%
-Later, the user can compose these atomic representations using the provided type
-level combinators in different ways into a ``generation specification'' that
-completely determines the generation process.
-
-We wil reuse the previously defined data type \ensuremath{\Conid{Exp}} and the function \ensuremath{\Varid{foo}} to
-explain the different concepts involved all across this section.
-
-\subsection*{\textbf{Representing Data Constructors}}
-
-We begin by introducing the simplest data type representation that we can
-extract from our codebase: the representation of single data constructor.
-%
-Each data constructor can be represented by an automatically derived data type
-consisting of a single constructor with the same fields as the original, except
-for the recursive ones that are abstracted away.
-%
-In this light, we represent each constructor of the data type \ensuremath{\Conid{Exp}} as follows:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;{H\!Rep}_{Val}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Val}\;\Conid{Int}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;{H\!Rep}_{Add}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Add}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;{H\!Rep}_{Mul}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}{Mk}_{Mul}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Note that the previous definitions are type parametric over the type parameter
-\ensuremath{\Varid{r}}.
-%
-This allow us to replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
-possible values on each case.
-%
-For instance, the value (\ensuremath{{Mk}_{Add}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{{H\!Rep}_{Add}\;\Conid{Int}}, while the
-value (\ensuremath{{Mk}_{Mul}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{{H\!Rep}_{Mul}\;\Conid{Bool}}.
-%
-In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
-chosen generation specification (which might be composed of several \ensuremath{\text{HRep}}s),
-without having to modify anything in the underlying machinery.
-
-Having the \ensuremath{\text{HRep}} of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : \ensuremath{\text{HRep}}_f \rightarrow t$) that maps a value from each
-representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
-%
-Then, we simply need to translate each constructor representation back into its
-corresponding one, translating the abstracted fields recursively:
-%
-\begin{alignat*}{4}
-  &\evalrep{\ensuremath{{Mk}_{Val}}\ n\ \ \ &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Val}}\ n \\
-  &\evalrep{\ensuremath{{Mk}_{Add}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Add}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp} \\
-  &\evalrep{\ensuremath{{Mk}_{Mul}}\ x\ y   &&}{\ensuremath{\Conid{Exp}}}
-    &&= \ensuremath{\Conid{Mul}}\ \evalrep{x}{Exp}\ \evalrep{y}{Exp}
-\end{alignat*}
-
-The last missing piece is to automatically derive a random generators for each
-constructor \ensuremath{\text{HRep}}.
-%
-For this purpose, it is important to consider that each constructor \ensuremath{\text{HRep}} has
-its recursive fields abstracted away with a type parameter that will be later
-instantiated with the generation specification type.
-%
-Given that this specification is unknown at the derivation time, we parametrize
-each \ensuremath{\text{HRep}} generator with a random generator \ensuremath{gen_{r}} that is used to generate
-each recursive fields:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{19}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Val}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}{Mk}_{Val}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}gen_{Add}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}{Mk}_{Add}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{Mul}\;{}\<[12]%
-\>[12]{}gen_{r}{}\<[19]%
-\>[19]{}\mathrel{=}{Mk}_{Mul}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-\subsection*{\textbf{Type Level Combinators}}
-
-In this work we define type level combinators which let the user combine the
-automatically derived \ensuremath{\text{HRep}}s in several ways.
-%
-In first place, we define a type combinator (\ensuremath{\_^\bigstar}) to tag a \ensuremath{\text{HRep}} to be
-terminal, i.e., a representation that is allowed be generated when the
-generation size gets exausted:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(f^\bigstar)\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\text{HRep}} with an
-explicit generation frequency $n$:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The previous combinators only include information relevant to the generation
-process, in a sense that neither one adds new structure to the final
-representation.
-%
-In this light, they do not alter the evaluation semantics, and we translate them
-back to our target data type by evaluating the inner representation:
-%
-\begin{alignat*}{4}
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Term\ x &&: \ensuremath{f^\bigstar}    &&}{t} &&= \evalrep{x : f}{t}
-\end{alignat*}
-
-In the previous equations we explicitly annotate (using a colon) the type of the
-evaluated term for clarity.
-%
-Later, to generate these combinators is enough to wrap a generated value from
-the inner representation with the apropriate tag:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{f^\bigstar}\;{}\<[16]%
-\>[16]{}gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Conid{Term}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
-\\
-\>[B]{}gen_{f \otimes n}\;{}\<[16]%
-\>[16]{}gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Conid{Freq}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-\ensuremath{\text{HRep}}s into a single one using a sum type and representing a random choice
-between them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\oplus\;\Varid{g})\;\Varid{a}\mathrel{=}\Conid{L}\;(\Varid{f}\;\Varid{a})\mid \Conid{R}\;(\Varid{g}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data type variant and evaluating the
-inner \ensuremath{\text{HRep}}:
-%
-\begin{alignat*}{3}
-  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
-  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
-\end{alignat*}
-
-Generating a composite \ensuremath{\text{HRep}} is slightly more complicated than before, as we
-need to perform a random choice based on the generation size and the given
-frequencies for each sub-represention:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{45}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{f \oplus g}\;gen_{r}{}\<[23]%
-\>[23]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{})\mskip1.5mu]{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{g}\;{}\<[38]%
-\>[38]{}gen_{r}{}\<[45]%
-\>[45]{})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-The previous definition we reflect the type level frequencies of \ensuremath{\Varid{f}} and \ensuremath{\Varid{g}}
-(\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}).
-%
-This frequency reflection defaults to $1$ if no frequency tag ($\otimes$) is
-used.
-%
-Then we use these frequencies to generate each inner \ensuremath{\text{HRep}} in the apropriate
-proportion.
-%
-When the generation size gets exhausted, we reflect the terminal generation
-frequency of each inner \ensuremath{\text{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
-\ensuremath{\Varid{freq0}\;\textit{@}\!g}).
-%
-This time, however, we default the frequency reflection to $0$ for any inner
-that not tagged as terminal, avoiding to generate non-terminal constructions in
-the last step.
-
-With the introduced combinators, we can easily create a type synonym \ensuremath{{H\!Rep}_{Exp}}
-to specify a generation schema equivalent to the one seen in the concrete random
-generator of type \ensuremath{\Conid{Exp}} presented in Section \ref{sec:randomtesting}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;{H\!Rep}_{Exp}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}{H\!Rep}_{Val}^\bigstar{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}{H\!Rep}_{Add}\;{}\<[34]%
-\>[34]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}{H\!Rep}_{Mul}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{{H\!Rep}_{Exp}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit in the definition of \ensuremath{{H\!Rep}_{Exp}}.
-%
-If we want to generate recursive values using this representation we need to
-introduce a last type level combinator to ``tie the knot'' recursively:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Fix}\;\Varid{f}\mathrel{=}\Conid{Fix}\;(\Varid{f}\;(\Conid{Fix}\;\Varid{f})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This datatype represents the \emph{fixed point} of a parametric data type \ensuremath{\Varid{f}},
-i.e., a data type where each recursive call gets instantiated with iself.
-%
-To translate fixed points back to our target data type we simply need translate
-the inner representation:
-%
-\begin{align*}
-  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
-\end{align*}
-%
-Categorically, this generic transformation of fixed points is called a
-\emph{catamorphism}, where our evaluation relation is known as its
-\emph{algebra}.
-
-Unlike the other combinators, to randomly generate a fixed point of a certain
-representation \ensuremath{\Varid{f}}, we do not parametrize the generation of the recursive fields
-of the inner represention over an external generator \ensuremath{gen_{r}}.
-%
-Instead, we can tie the recursive knot in our random generator by piping the
-generation of the inner \ensuremath{\text{HRep}}, and calling itself to generate recursive values:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}gen_{f}\;genFix_{f}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This way we obtain a concrete recursive generator for each representation \ensuremath{\Varid{f}}
-that we define.
-%
-With this generator, we can define a random generator for our target data type
-simply by evaluating a random value of the representation:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Exp}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{\,\text{HRep}_{Exp}}{}\<[E]%
-\\
-\>[17]{}\Varid{return}\;\evalrep{x}{Exp}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-So far we have introduced the representation machinery required to represent the
-random generation of a target data type considering only the structure encoded
-on its definition.
-%
-This approach has the advantage of being extendable with different sources of
-structured information.
-%
-We proceed to introduce two extensions that help to address the problematic
-testing scenarios presented in the previous section.
-
-\subsection*{\textbf{Representing Pattern Matchings}}
-
-We can follow a similar reasoning as before to represent the pattern matching
-structure from a given function.
-%
-Consider the function \ensuremath{\Varid{foo}} defined in the previous section.
-%
-We can derive a new data type for each one of its pattern matchings, whose
-fields represent each pattern variable present on the given pattern, abstracting
-away every pattern variable of type \ensuremath{\Conid{Exp}}.
-%
-Concretely, we define the following data types:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;{H\!Rep}_{foo\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}{Mk}_{foo\#1}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;{H\!Rep}_{foo\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}{Mk}_{foo\#2}\;\Conid{Int}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The first pattern of \ensuremath{\Varid{foo}} conteins two pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type
-\ensuremath{\Conid{Exp}} that are abstracted in the fields of \ensuremath{{Mk}_{foo\#1}}.
-%
-Similarly, the second pattern of \ensuremath{\Varid{foo}} contains the pattern variable \ensuremath{\Varid{x}} of type
-\ensuremath{\Conid{Int}} represented by the first field of \ensuremath{{Mk}_{foo\#2}}, along with the pattern
-variable \ensuremath{\Varid{y}} of type \ensuremath{\Conid{Exp}} that also is abstracted in the second field of
-\ensuremath{{Mk}_{foo\#2}}.
-
-To evaluate these representations back to our target data type, we simply expand
-them it into the concrete value represented by the original pattern:
-%
-\begin{align*}
-  &\evalrep{\ensuremath{{Mk}_{foo\#1}}\ x\ y}{Exp} = \\
-  &\ \ \ensuremath{\Conid{Add}\;(\Conid{Add}\;\evalrep{x}{Exp}\;(\Conid{Val}\;\mathrm{50}))\;(\Conid{Add}\;(\Conid{Val}\;\mathrm{25})\;\evalrep{y}{Exp})}\\
-  &\evalrep{\ensuremath{{Mk}_{foo\#2}}\ x\ y}{Exp} = \\
-  &\ \ \ensuremath{\Conid{Mul}\;(\Conid{Val}\;\mathrm{50})\;(\Conid{Mul}\;(\Conid{Val}\;\Varid{x})\;\evalrep{y}{Exp})}
-\end{align*}
-
-And the random generation of pattern \ensuremath{\text{HRep}}s is defined in the same way as we
-did before with data constructors \ensuremath{\text{HRep}}s:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{33}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{52}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{foo\#1}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}{Mk}_{foo\#1}{}\<[33]%
-\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[52]%
-\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{foo\#2}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[21]%
-\>[21]{}\mathrel{=}{Mk}_{foo\#2}{}\<[33]%
-\>[33]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[52]%
-\>[52]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-Finally, we can join the pattern matching representation of each clause of \ensuremath{\Varid{foo}}
-into a single one:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{38}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;{H\!Rep}_{foo}{}\<[16]%
-\>[16]{}\mathrel{=}{H\!Rep}_{foo\#1}\;\!\!\oplus\!\!\;{}\<[38]%
-\>[38]{}{H\!Rep}_{foo\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Representing Abstract Interfaces}}
-
-To introduce the higher level representation of module abstract interfaces,
-suppose we have a module \ensuremath{\Conid{M}} defining \ensuremath{\Conid{Exp}} combinators as follows:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{ten}\mathbin{::}\Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{ten}\mathrel{=}\Conid{Val}\;\mathrm{10}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{square}\mathbin{::}\Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{square}\;\Varid{x}\mathrel{=}\Conid{Mul}\;\Varid{x}\;\Varid{x}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\Varid{minus}\mathbin{::}\Conid{Exp}\to \Conid{Exp}\to \Conid{Exp}{}\<[E]%
-\\
-\>[B]{}\Varid{minus}\;\Varid{x}\;\Varid{y}\mathrel{=}\Conid{Add}\;\Varid{x}\;(\Conid{Mul}\;\Varid{y}\;(\Conid{Val}\;(\mathbin{-}\mathrm{1}))){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-We will represent each function of \ensuremath{\Conid{M}} returning a value of type \ensuremath{\Conid{Exp}} following
-the same idea as before, deriving a data type with a single data constructor for
-each one of them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;{H\!Rep}_{ten}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}{Mk}_{ten}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;{H\!Rep}_{square}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}{Mk}_{square}\;{}\<[37]%
-\>[37]{}\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;{H\!Rep}_{minus}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}{Mk}_{minus}\;{}\<[37]%
-\>[37]{}\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In this case, each single constructor will have as fields the types of the
-inputs of the function that they represent.
-%
-As before, we abstract away any field representing and input of type \ensuremath{\Conid{Exp}} with
-a type parameter \ensuremath{\Varid{r}}.
-
-To evaluate these representations, we simply pipe the values on its fields as
-input parameter of each original function, returning its result.
-
-\begin{alignat*}{3}
-  &\evalrep{\ensuremath{{Mk}_{ten}}         &&}{Exp} = \ensuremath{\Varid{ten}} \\
-  &\evalrep{\ensuremath{{Mk}_{square}}\ x   &&}{Exp} = \ensuremath{\Varid{square}\;\evalrep{x}{Exp}} \\
-  &\evalrep{\ensuremath{{Mk}_{minus}}\ x\ y &&}{Exp} = \ensuremath{\Varid{minus}\;\evalrep{x}{Exp}\;\evalrep{y}{Exp}}
-\end{alignat*}
-
-Moreover, the generation procedure for abstract interface representations
-follows the same pattern as before:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{35}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{54}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{ten}\;{}\<[15]%
-\>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}{Mk}_{ten}{}\<[E]%
-\\
-\>[3]{}gen_{square}\;{}\<[15]%
-\>[15]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}{Mk}_{square}{}\<[35]%
-\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\\
-\>[3]{}gen_{minus}\;{}\<[14]%
-\>[14]{}gen_{r}{}\<[22]%
-\>[22]{}\mathrel{=}{Mk}_{minus}{}\<[35]%
-\>[35]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{r}{}\<[54]%
-\>[54]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-We will also define a type synonym to join all the representations of the module
-\ensuremath{\Conid{M}} into a single one.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{55}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;{H\!Rep}_{M}{}\<[14]%
-\>[14]{}\mathrel{=}{H\!Rep}_{ten}\;\!\!\oplus\!\!\;{}\<[34]%
-\>[34]{}{H\!Rep}_{square}\;\!\!\oplus\!\!\;{}\<[55]%
-\>[55]{}{H\!Rep}_{minus}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Finally, we can put all the derived machinery together into a generation
-specification \ensuremath{\Conid{Spec}} for the type \ensuremath{\Conid{Exp}}, assigning (possibly) different
-generation frequencies to each individual \ensuremath{\text{HRep}} that constitutes it:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\Conid{Spec}{}\<[12]%
-\>[12]{}\mathrel{=}{}\<[20]%
-\>[20]{}{H\!Rep}_{Exp}\;{}\<[30]%
-\>[30]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}{H\!Rep}_{foo}\;{}\<[30]%
-\>[30]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[12]{}\!\;\oplus\;{}\<[20]%
-\>[20]{}{H\!Rep}_{M}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This previous definition can be interpreted graphically as it is shown in the
-Figure \ref{fig:hrep}, where curly arrows represent the structural information
-extracted using meta-programming.
-
-\begin{figure}[t]
-  \centering
-  \input{tikz/hrep.tex}
-  \caption{Higher level representation of the data type \ensuremath{\Conid{Exp}}, defined using
-    structural information from the function \ensuremath{\Varid{foo}} and the abstract interface of
-    the module \ensuremath{\Conid{M}}.}
-  \label{fig:hrep}
-\end{figure}
-
-This representation can be extended to mutually recursive and parametric types
-...
-
-% \begin{align*}
-%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-% \end{align*}
-%
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-%
-% \begin{code}
-% data (Term f) a = TagTerm (f a)
-
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-% instance (Term f down Exp) where
-%   step (TagTerm f) = step f
-% \end{code}
-
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-% %include sections/casestudies.tex
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-01-22 paper/watch.sh
Sections no more

@@ -1,4 +1,4 @@
 #!/bin/sh
 
 #https://facebook.github.io/watchman/docs/watchman-make.html
-watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' 'sections/[^.]*.tex' 'tikz/[^.]*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all
+watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' '*.tex' 'tikz/[^.]*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all

### 2019-01-22 paper/introduction.tex
WIP - Intro

@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
+% Scenario of random testing
+Random testing is a promising approach to finding bugs {\tocite}.
+%
+QuickCheck \cite{ClaessenH00} is a dominant tool of this sort used by the
+Haskell community.
+%
+It requires developers to specify testing properties which describe programs
+behavior and random data generator based on the \emph{type} of the expected
+inputs. %, e.g., an integer, an string, or an abstract data type (ADT).
+%
+QuickCheck then generates a large number of random \emph{test cases} and reports
+those violating testing properties.
+%
+
+% Situation with Quick Check, ADTs, tools
+While QuickCheck comes equipped with random generators for built-in types, there
+exists additional tools to automatically derive generators for user-defined
+abstract data types (ADT) (e.g.,
+\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}).
+%
+The main difference about these tools lies on the guarantees provided when it
+comes to ensuring a terminating generation process or distribution of the random
+values.
+%
+When generating ADT values, these tools guarantee that such values are
+\emph{well-typed}.
+%
+In that manner, generated values follow the structure described by the
+definition of the ADT, which proves useful when testing software with highly
+structured inputs like compilers \cite{Palka11,MidtgaardJKNN17}.
+%
+Generating values from ADTs also proves to be a fruitful approach when looking
+for vulnerabilities with fuzzers \cite{GriecoCB16,grieco2017}.
+%
+
+%% Problem
+
+%% Our proposal
 
@@
 \end{itemize}
+
+% Local Variables:
+% TeX-master: "main.lhs.tex"
+% TeX-command-default: "Make"
+% End:

### 2019-01-23 paper/introduction.tex
Adding citations
Intro, WIP

--- 2019-01-22 paper/introduction.tex
@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
+% Scenario of random testing
+Random testing is a promising approach to finding bugs
+\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
+%
+\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
+Haskell community.
+%
+It requires developers to specify (i) \emph{testing properties} which describe
+programs behavior and (ii) \emph{random data generator} based on the
+\underline{\emph{types}} of the expected inputs (e.g., an integer, an string,
+etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
+%
+QuickCheck then generates random test cases and reports those violating testing
+properties.
+%
 
-We remark that, altough this work focuses on Haskell data types, this technique
-is general enough to be applied to most programming languages with some level of
-support for composite types.
+% Situation with Quick Check, ADTs, tools
+QuickCheck comes equipped with random generators for built-in types, while it
+requires to manually write generators for user-defined abstract data types
+(ADTs).
+%
+Recently, there have been a proliferation of tools to automatically derive
+QuickCheck generators for ADTs
+\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
+%
+The main difference about these tools lies on the guarantees provided to ensure
+\emph{the termination of the generation process} and the \emph{distribution of
+  random values}.
+%
+Despite their differences, these tools guarantee that generated values are
+\emph{well-typed}.
+%
+In other words, generated values follow the structure described by the
+definition of the ADT.
+%
 
-The main contribuitions of this paper are:
+%% How to use random generated ADTs
+Well-typed ADT values are specially useful when testing programs which expect
+highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
 %
-\begin{itemize}
+Generating ADT values also proves fruitful when looking for vulnerabilities with
+fuzzers \cite{GriecoCB16,grieco2017}.
 %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
+%% Problem
 %
-\item We present a generation technique able to encode stronger properties of
-  the target data type by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
+Despite these success stories, ADT type-definitions do not often capture all the
+invariants expected from the data that they are intended to model.
 %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
+As a result, even if random values are well-typed, they might not present enough
+structure to penetrate into deep layers of software.
+
+%% Our proposal
+In this work, we propose a novel improvement in the generation process of ADT
+values by exploiting some static information found in the codebase.
 %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
+More specifically, to refine the structure of generated values, we propose a
+generation process that is capable to consider how programs pattern-matched on
+ADTs values as well as how they get manipulated via interfaces.
 %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract useful
-  structure information from the user source code.
+Furthermore, we show how to predict (at compile time) the distribution of the
+\emph{expected} numbers of ADT constructors, values fitting a certain pattern,
+and calls to interfaces.
 %
-\end{itemize}
+\todo[inline,author=Ale]{Mention branching processes}
+
+%
+We implement our ideas in a tool, called {\dragenp}, that is capable to
+automatically synthesize QuickCheck generators for ADT values, where the
+distributions of random values can be adjusted at compile-time to what
+developers might want based on our predictions.
+%
+\todo[inline,author=Ale]{Add here later about test cases when we know what they
+  are}
+
+We remark that, although this work focuses on Haskell algebraic data types, this
+technique is general enough to be applied to most programming languages.
+%with some level of support for composite types.
+
+% The main contribuitions of this paper are:
+% %
+% \begin{itemize}
+%   %
+% \item We identify two patological scenarios for which standard type-driven
+%   automatic derivation tools fail to synthesize practical random generators, due
+%   to a lack of either type structure or domain knowleadge (Section
+%   \ref{sec:randomtesting}).
+%   %
+% \item We present a generation technique able to encode stronger properties of
+%   the target data type by reifying the static information present on the program
+%   codebase (Section \ref{sec:hrep}).
+%   %
+% \item We apply and extend the theory of branching processes to analitically
+%   predict the average distribution of generated values.
+%   %
+%   Furthermore, we use the predictions to perform simulation-based optimization
+%   of the random generation parameters (Section \ref{sec:synthesis}).
+%   %
+% \item We provide an implementation of our ideas in the form of a Haskell library
+%   to perform automatic derivation of random generators capable to extract useful
+%   structure information from the user source code.
+%   %
+% \end{itemize}
+
+% Local Variables:
+% TeX-master: "main.lhs.tex"
+% TeX-command-default: "Make"
+% End:

### 2019-01-23 paper/abstract.tex
Abstract

@@
 
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
+  % Status
+  Automatic generation of random values described by abstract data types (ADTs)
+  is often a hard task.
+  % data types (ADTs), % ---specially when considering ADTs which
+  % represent HTML pages, images, programs or other structural rich objects.
+  % Situation
+  State-of-the-art random testing tools often can automatically synthesize
+  random data generators based on ADTs definitions.
+  %
+  In that manner, generated values comply with the structure described by
+  ADTs---something that proves useful when testing software which expects
+  complex input like web pages, images, or even programs.
+  %
+  % Problem
+  However, it sometimes becomes necessary to generate structural richer ADTs
+  values in order to test deeper software layers.
+  %
+  % Solution
+  In this light, we propose to leverage static information found in the
+  codebase as a manner to improve the generation process.
+  %
+  Namely, our generators are capable to consider how programs branch on input
+  data as well as how ADTs values are built via interfaces.
+  %
+  % Concretely, we propose to consider not only ADTs definitions, but also how
+  % programs branch on input data as well as how ADTs values get manipulated via
+  % interfaces.
+  %
+  We implement a tool, called {\dragenp}, responsible to synthezise generators
+  for ADTs values while providing compile-time guarantees about their
+  distributions.
+  %
+  Our compile-time predictions allow {\dragenp} to provide an heuristic that
+  tries to adjust the distribution of generators to what developers might want.
+  \todo[inline,author=Ale]{Case study}
 

### 2019-01-23 paper/randomtesting.tex

\section{Background} \label{sec:randomtesting}

In this section, we briefly introduce the common approach for automatically
deriving random data generators for ADTs in QuickCheck by using the information
found at the type-level.
%type-driven approach, along
%with its main drawbacks.
%
% Haskell is a strongly typed programming language with a powerful type system.
% %
% It lets programmers encode a considerable amount of information about the
% structure of their systems using data types that can be checked at compilation
% time.
% %
% One of its key aspects is the support for Algebraic Data Types (ADTs).
% %
% Essentially, an ADTs is a composite type defined by combining other types in
% terms of \textbf{sums} (also known as \emph{variant types}) and
% \textbf{products} (or tuples) of other data types.
% %
To exemplify this, and for illustrative purposes, let us consider the following
ADT definition to encoding simple |Html| pages:
%
\begin{code}
data Html  =   Text    String
           pp  Single  String
           pp  Tag     String  Html
           pp  Join    Html    Html

\end{code}
%
The type |Html| allows to build pages via four possible constructions:
|Text|---which represents plain text values---, |Single| and |Tag|---which
represent singular and paired Html tags, respectively---, and |Join|---which
concats two Htmls pages one after another.
%
% In the previous definition, we declare |Html| as the sum of four possible
% constructions: |Text| represents plain text values. |Sing| and |Tag| represent
% singular and paired Html tags, and |Join| concats two expression one after
% another.
% %
% We only encode a very small subset of the actual Html specification for
% illustrative reasons.
% %
In Haskell, |Text|, |Single|, |Tag|, and |Join| are known as \emph{data
  constructors} (or constructors for short) and are used to distiguish which
variant of the ADT we are constructing.
%
Each data constructor is defined as a product of zero or more types known as
\emph{fields}.
%
For instance, |Text| has a field of type |String|, whereas |Join| two recusive
fields of type |Html|.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this representation, the page ``|html_example|'' can be encoded as:
%
\begin{code}
Tag "html" (Join (Join
  (Text "hello") (Sing "hr")) (Text "bye"))
\end{code}
%
% Additionally, we can define a function |render| to serialize |Html| values as
% follows:
% %
% \begin{code}
% render :: Html -> String
% render (Text t) = t
% render (Sing t) = "<" ++ t ++ ">"
% render (Join x y) = render x ++ render y
% render (Tag t x)
%   = "<" ++ t ++ ">" ++ render x ++ "</" ++ t ++ ">"
% \end{code}

% In the previous definition, |render| is described using \emph{pattern matching}
% over each possible kind of value.
% %
% Using pattern matching we can define functions idiomatically by defining
% different function clauses for each input pattern we are interested on.
% %
% Patterns can be defined to match specific constructors, literal values or
% variable subexpressions (like |t|, |x| and |y| in the definition of |render|).
% %
% They can also be nested in order to match very specific patterns of values.

\subsection*{\textbf{Type-Driven Generation of Random Values}}

In order to generate random value of types involving ADTs, most approaches
require users to provide a random data generator for each one of them.
%
This is a cumbersome and error prone task and usually \emph{follows closely the ADT
structure}.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Html|:
%
\begin{code}
genHtml = sized (\size ->
   if size == 0  then  frequency
                       [ (2,  Text    <$> genString)
                       , (1,  Single  <$> genString) ]
   else frequency
     [  (2,  Text    <$> genString)
     ,  (1,  Single  <$> genString)
     ,  (4,  Tag     <$> genString <*> smaller genHtml)
     ,  (3,  Join    <$> smaller genHtml <*> smaller genHtml) ])
\end{code} %$
%
We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
elements, respecitvely, e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.
%
The random generator |genHtml| is defined using \quickcheck's function |sized|
to paremetrize the generation process up to an external natural number known as
the \emph{generation size}.
%
This parameter is chosen by the user, and it is used to limit the maximum amount
of recursive calls that this random generator can perform and thus ensuring
the termination of the generation process.
%
When called with a positive generation size, this generator can pick to generate
among any |Html| data constructor of with an explicitly given frequency that can
be chosen by the user---in this example, 2, 1, 4 and 3 for |Text|, |Single|,
|Tag|, and |Join|, respectively.
%
When it picks to generate a |Text| or a |Single| data constructor, it also
generates a random |String| value using the standard QuickCheck generator
|genString|.% (\quickcheck provides standard random generators for most base types
% like |String|, |Int|, |Bool|, etc.)
\footnote{%
  The operators |<$>| and |<*>| are used in Haskell to combine values obtained
  from calling to a random generator and they are not particularly relevant for
  the point being made in this work.}
%$
On the other hand, when it picks to generate a |Join| constructor, it also
generates two independent random subexpression recursively, thus decreasing the
generation size by a unit on each recursive invocation (|smaller genHtml|).
%
The case of random generation of |Tag| constructors follows analogously.
%
This random process keeps calling itself recursively until the generation size
reaches zero, where the generator is constrained to pick only among terminal
data constructors, being |Text| and |Single| the only possible choices in our
particular case.

% Strictly decreasing the the generation size by one on each recursive call
% results in a useful property over the generated data: every generated value has
% at most |n| levels, where |n| is the generation size set by the user.
%
% This property enables us to model the generation process using the theory of
% branching processes introduced by \tocite, and extended in this work as
% described in Section \ref{sec:synthesis}.

The previous definition is rather mechanical, except perhaps for the chosen
generation frequencies.
%
% In this light, it easy to extend this procedure to any data type defined in an
% algebraic fashion.
% %
% Fortunately, there exists different meta-programming tools to avoid the user
% from having to mechanically write random generators over and over again for each
% user-defined ADT.
% %
% The simplest tool for such purpose is \megadeth \tocite.
% %
% Given a target data type, it synthesizes a random generator for it that behaves
% similarly to the one presented above, where the generation frequencies are
% defined to be uniform across constructors.
% %
% However, picking among different data constructors with uniform frequency can
% lead to a generation process biased towards generating (in average) very small
% values, regardless of the generation sized set by the user \tocite.
%
Fortunately, there exists a tool called \dragen
\cite{DBLP:conf/haskell/MistaRH18} conceived to mitigate this problem.
%
Instead of setting a uniform frequency (or probability) of data constructors,
{\dragen} uses the theory of branching processes \cite{gw1875} to modelize and
predict analytically the expected number of generated data constructors.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation frequency of each data constructor in order
to obtain a particular distribution of values that can be specified by the
user---thus providing a flexible testing environment while still being mostly
automated.

As many other tools \tocite, {\dragen} synthesizes random generators similar to
the one shown before, where the generation process is limited to pick \emph{a
  single data constructor at the time and then recursively generate each
  required subexpression independently}.
%
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same:
%
% Althogh both \megadeth and \dragen synthesize random generators that are
% theoretically capable to generate the whole space of values of the target data
% type. the limitations arise quickly when we consider that the underlying
% generation model is essentially the same: they pick a single data constructor
% and recusively generate each required subexpression independently.
%
In practice, this procedure is often too generic to generate random data with
enough structural complexity required for testing certain applications.
% to be used for testing purposes.

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End:

### 2019-01-23 paper/main.tex

@@
 \usepackage{bm}
+\usepackage{microtype}
+\usepackage{pgfplots}
 \usetikzlibrary{fadings}
@@
 \newcommand{\dragen}{\textbf{DRAGEN}\xspace}
+\newcommand{\dragenp}{\textbf{DRAGEN\!+}\xspace}
 
@@
 %% Title
-\title{Deriving Random Generators via\\ Higher-Level Representations}
+\title{Random Generation of Rich \\ Abstract Data Type Values }
 
@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-We remark that, altough this work focuses on Haskell data types, this technique
-is general enough to be applied to most programming languages with some level of
-support for composite types.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data type by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract useful
-  structure information from the user source code.
-  %
-\end{itemize}
+% Scenario of random testing
+Random testing is a promising approach to finding bugs
+\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
+%
+\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
+Haskell community.
+%
+It requires developers to specify (i) \emph{testing properties} which describe
+programs behavior and (ii) \emph{random data generator} based on the
+\underline{\emph{types}} of the expected inputs (e.g., an integer, an string,
+etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
+%
+QuickCheck then generates random test cases and reports those violating testing
+properties.
+%
+
+% Situation with Quick Check, ADTs, tools
+QuickCheck comes equipped with random generators for built-in types, while it
+requires to manually write generators for user-defined abstract data types
+(ADTs).
+%
+Recently, there have been a proliferation of tools to automatically derive
+QuickCheck generators for ADTs
+\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
+%
+The main difference about these tools lies on the guarantees provided to ensure
+\emph{the termination of the generation process} and the \emph{distribution of
+  random values}.
+%
+Despite their differences, these tools guarantee that generated values are
+\emph{well-typed}.
+%
+In other words, generated values follow the structure described by the
+definition of the ADT.
+%
+
+%% How to use random generated ADTs
+Well-typed ADT values are specially useful when testing programs which expect
+highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
+%
+Generating ADT values also proves fruitful when looking for vulnerabilities with
+fuzzers \cite{GriecoCB16,grieco2017}.
+%
+%% Problem
+%
+Despite these success stories, ADT type-definitions do not often capture all the
+invariants expected from the data that they are intended to model.
+%
+As a result, even if random values are well-typed, they might not present enough
+structure to penetrate into deep layers of software.
+
+%% Our proposal
+In this work, we propose a novel improvement in the generation process of ADT
+values by exploiting some static information found in the codebase.
+%
+More specifically, to refine the structure of generated values, we propose a
+generation process that is capable to consider how programs pattern-matched on
+ADTs values as well as how they get manipulated via interfaces.
+%
+Furthermore, we show how to predict (at compile time) the distribution of the
+\emph{expected} numbers of ADT constructors, values fitting a certain pattern,
+and calls to interfaces.
+%
+For that, we extend some recent results on applying \emph{branching
+  processes}---a simple stochastic model conceived to study population
+growth---to predict the distribution of QuickCheck
+generators\cite{DBLP:conf/haskell/MistaRH18}.
+%
+We implement our ideas in a tool, called {\dragenp}, that is capable to
+automatically synthesize QuickCheck generators for ADT values, where the
+distributions of random values can be adjusted at compile-time to what
+developers might want based on our predictions.
+%
+\todo[inline,author=Ale]{Add here later about test cases when we know what they
+  are}
+
+We remark that, although this work focuses on Haskell algebraic data types, this
+technique is general enough to be applied to most programming languages.
+%with some level of support for composite types.
+
+% The main contribuitions of this paper are:
+% %
+% \begin{itemize}
+%   %
+% \item We identify two patological scenarios for which standard type-driven
+%   automatic derivation tools fail to synthesize practical random generators, due
+%   to a lack of either type structure or domain knowleadge (Section
+%   \ref{sec:randomtesting}).
+%   %
+% \item We present a generation technique able to encode stronger properties of
+%   the target data type by reifying the static information present on the program
+%   codebase (Section \ref{sec:hrep}).
+%   %
+% \item We apply and extend the theory of branching processes to analitically
+%   predict the average distribution of generated values.
+%   %
+%   Furthermore, we use the predictions to perform simulation-based optimization
+%   of the random generation parameters (Section \ref{sec:synthesis}).
+%   %
+% \item We provide an implementation of our ideas in the form of a Haskell library
+%   to perform automatic derivation of random generators capable to extract useful
+%   structure information from the user source code.
+%   %
+% \end{itemize}
+
+% Local Variables:
+% TeX-master: "main.lhs.tex"
+% TeX-command-default: "Make"
+% End:
+%\newpage
+
 \section{Random Data Generation in Haskell} \label{sec:randomtesting}
@@
 generators in a modular and flexible way.
+
+\newpage
+
 \section{Extracting Structure} \label{sec:hrep}
@@
 % \end{figure}
+
+\newpage
+
 \section{Random Generators Synthesis} \label{sec:synthesis}
@@
 Nam euismod tellus id erat.
-% %include sections/casestudies.tex
-\section{Related Work}
+
+\newpage
+
+\section{Case Studies}
 
@@
 Nam euismod tellus id erat.
+
+\begin{figure}[t]
+  \centering
+  \input{tikz/html.tex}
+  \caption{Something something.}
+  \label{fig:html}
+\end{figure}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
+tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
+eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
+parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
+Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
+Nam vestibulum accumsan nisl.
+
+Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
+facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
+vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
+Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
+varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
+ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
+vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
+orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
+est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+
+\newpage
+
+\section{Related Work}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+\newpage
+
 \section{Final Remarks}

### 2019-01-24 paper/randomtesting.tex
Done for today
Till Section III

@@
 %
-With this representation, the page ``|html_example|'' can be encoded as:
+With this representation, the page \texttt{<html>hello<hr>bye</html>} can be
+encoded as:
 %
@@
 Tag "html" (Join (Join
-  (Text "hello") (Sing "hr")) (Text "bye"))
+  (Text "hello") (Single "hr")) (Text "bye"))
 \end{code}
@@
 
-\subsection*{\textbf{Type-Driven Generation of Random Values}}
+\subsection{Type-driven generation of random values}
 

### 2019-01-25 paper/introduction.tex
Intro, first draft
Pass over intro

@@
 programs behavior and (ii) \emph{random data generator} based on the
-\underline{\emph{types}} of the expected inputs (e.g., an integer, an string,
+{\emph{types}} of the expected inputs (e.g., an integer, an string,
 etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
@@
 QuickCheck comes equipped with random generators for built-in types, while it
-requires to manually write generators for user-defined abstract data types
-(ADTs).
+requires to manually write generators for user-defined ADTs.
 %
@@
 %
-As a result, even if random values are well-typed, they might not present enough
-structure to penetrate into deep layers of software.
+As a result, even if random values are well-typed, they might not be built with
+enough structure to penetrate into deep layers of software.
 
@@
 More specifically, to refine the structure of generated values, we propose a
-generation process that is capable to consider how programs pattern-matched on
-ADTs values as well as how they get manipulated via interfaces.
+generation process that is capable to consider how programs branch based on ADTs
+values' patterns as well as how they get manipulated by abstract interfaces.
 %
-Furthermore, we show how to predict (at compile time) the distribution of the
-\emph{expected} numbers of ADT constructors, values fitting a certain pattern,
-and calls to interfaces.
+Furthermore, we show how to predict the distribution of the \emph{expected}
+numbers of ADT constructors, values fitting certain branching patterns, and
+calls to interfaces.
 %
-\todo[inline,author=Ale]{Mention branching processes}
-
+For that, we extend some recent results on applying \emph{branching
+  processes}\cite{gw1875}---a simple stochastic model conceived to study
+population growth.
+%---to predict the distribution of QuickCheck
+%generators\cite{DBLP:conf/haskell/MistaRH18}.
 %
 We implement our ideas in a tool, called {\dragenp}, that is capable to
-automatically synthesize QuickCheck generators for ADT values, where the
-distributions of random values can be adjusted at compile-time to what
+automatically synthesize QuickCheck generators for \emph{rich ADT values}, where
+the distributions of random values can be adjusted at compile-time to what
 developers might want based on our predictions.

### 2019-01-25 paper/randomtesting.tex
Pass over Section II

@@
 To exemplify this, and for illustrative purposes, let us consider the following
-ADT definition to encoding simple |Html| pages:
+ADT definition to encode simple |Html| pages:
 %
@@
 to paremetrize the generation process up to an external natural number known as
-the \emph{generation size}.
+the \emph{generation size}---capture in the code with variable |size|.
 %
@@
 When called with a positive generation size, this generator can pick to generate
-among any |Html| data constructor of with an explicitly given frequency that can
-be chosen by the user---in this example, 2, 1, 4 and 3 for |Text|, |Single|,
-|Tag|, and |Join|, respectively.
+among any |Html| data constructor of with an explicitly given \emph{generation
+  frequency} that can be chosen by the user---in this example, 2, 1, 4 and 3 for
+|Text|, |Single|, |Tag|, and |Join|, respectively.
 %
@@
   The operators |<$>| and |<*>| are used in Haskell to combine values obtained
-  from calling to a random generator and they are not particularly relevant for
-  the point being made in this work.}
+  from calling random generators and they are not particularly relevant for the
+  point being made in this work.}
 %$
@@
 %
-Fortunately, there exists a tool called \dragen
-\cite{DBLP:conf/haskell/MistaRH18} conceived to mitigate this problem.
+\dragen \cite{DBLP:conf/haskell/MistaRH18} is a tool conceived to mitigate the
+problem of finding the appropriated generation frequencies.
 %
-Instead of setting a uniform frequency (or probability) of data constructors,
-{\dragen} uses the theory of branching processes \cite{gw1875} to modelize and
-predict analytically the expected number of generated data constructors.
+It uses the theory of branching processes \cite{gw1875} to modelize and predict
+analytically the expected number of generated data constructors.
 %
@@
 
-As many other tools \tocite, {\dragen} synthesizes random generators similar to
-the one shown before, where the generation process is limited to pick \emph{a
-  single data constructor at the time and then recursively generate each
-  required subexpression independently}.
+As many other tools for automatic derivation of generators
+(e.g.,\cite{GriecoCB16,RuncimanNL08,mitchell2007,DuregardJW12}), {\dragen}
+synthesizes random generators similar to the one shown before, where the
+generation process is limited to pick \emph{a single data constructor at the
+  time and then recursively generate each required subexpression independently}.
 %

### 2019-01-26 paper/tikz/hrep.tex
Done for today
Added evaluation

@@
 %Right Arrow [id:dp7732858674364509]
-\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (340,255) -- (340,195) -- (335,195) -- (345,195) -- (355,195) -- (350,195) -- (350,255) -- cycle ;
+\draw  [fill={rgb, 255:red, 226; green, 226; blue, 226 }  ,fill opacity=1 ] (340,230) -- (340,195) -- (335,195) -- (345,195) -- (355,195) -- (350,195) -- (350,230) -- cycle ;
 %Right Arrow [id:dp2590973168233759]
@@
 % Text Node
-\draw (202.5,142.5) node [scale=0.7] [align=left] {\textbf{Structure}\\ \textbf{reification}};
+\draw (202.5,142.5) node [scale=0.7] [align=left] {\textbf{Structure}\\ \textbf{specification}};
 % Text Node
@@
 % Text Node
-\draw (346.5,272.5) node [scale=0.7] [align=left] {\textbf{User desired} \\ \textbf{distribution}};
+\draw (348,245) node [scale=0.7] [align=left] {\textbf{User desired} \\ \textbf{distribution}};
 

### 2019-01-28 paper/casestudies.tex
No directory for section, it breaks all nice features of emacs and bibligraphy
Figures

--- 2019-01-15 paper/sections/synthesis.tex
@@
-\section{Random Generators Synthesis} \label{sec:synthesis}
+\section{Case Studies}
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+Intuitively, adding structure information to a random generator should provide
+better results when verifying a system using random testing.
+%
+% For instance, by generating values that are no only syntactically but also
+% semantically meaningful we would expect to reach deeper stages of our program
+% pipeline other than merely input parsing/validation.
+%
+This section describes two case studies showing that considering structure
+information when deriving generators consistenly produces better testing
+results.
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+Instead of restricting our scope to Haskell, in this work we followed a broader
+evaluation approach taken previously to compare the performance of other
+state-of-the-art derivation techniques
+\cite{grieco2017,DBLP:conf/haskell/MistaRH18}.
+%
+Concretely, we evaluate how including additional structure information when
+generating a set of random test cases (often referred as a \emph{corpus})
+affects the code coverage obtained when using them to test a given target
+program.
+%
+For this purpose, we targeted two external programs which expect highly
+structured inputs, namely \emph{GNU CLISP}---the GNU Common Lisp compiler, and
+\emph{HTML Tidy}---a well known HTML refactoring and correction utility.
+%
+We remark that these applications are not written in Haskell.
+%
+However, there exist Haskell libraries defining ADTs encoding their input
+structure, i.e., Lisp and HTML values respectively. These libraries are:
+\emph{hs-zuramaru}, implementing an embedded Lisp intepreter for a small subset
+of this programming language, and \emph{html}, defining a combinator library for
+constructing HTML values.
+%
+These libraries also come with serialization functions to map Haskell values
+into corresponding test case files.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+We started by compiling instrumented versions of the target programs in a way
+that they also return the execution path followed in the source code every time
+we run them with a given input test case.
+%
+This let us distinguish the amount of different execution paths that a randomly
+generated corpus can trigger.
+%
+On the other side, we used the ADTs defined on the chosen libraries to derive
+random generators using \dragen and \dragenp, including structure information
+extracted from the library's codebase in the case of the latter.
+%
+Then, we evaluated the code coverage triggered by independent, randomly
+generated corpora of different sizes varying from 100 to 1000 test cases each.
+%
+In order to remove any external bias, we derived random generators optimized to
+follow a uniform distribution of constructors (and pattern matchings or function
+calls in the case \dragenp), and carefully adjusted their generation sizes to
+match the average test case size in bytes.
+%
+This way, any noticeable difference in the code coverage can be attributed to
+the present (or lack thereof) structure information when generating the test
+cases.
+%
+Additionally, to achieve statistical significance we repeated each experiment 30
+times with independently generated sets of random test cases.
+
+\begin{figure*}[t]
+  \centering
+  \input{tikz/lisp.tex}
+  \hspace{5pt}%
+  \input{tikz/html.tex}
+  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
+    \dragenp (\ref{exp:dragenp}). }
+  \label{fig:coverage}
+\end{figure*}
+
+Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
+triggered for each permutation of corpus size and derivation tool, including
+error bars indicating the standard error of the mean on each case.
+%
+We proceed to describe each case study in detail.
+
+\paragraph{Generating Lisp code using pattern matching information}
+
+In this first case study we wanted to evaluate the observed code coverage
+differences when considering structure information present on functions pattern
+matchings.
+
+Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
+represented as plain strings; and literal values like booleans or integers.
+%
+In order to interpret Lisp programs, this unified representation of data and
+code requires this library to pattern match against common patterns like
+let-bindings, if-then-else expressions and arithmetic operators among others.
+%
+In particular, each one of these patterns match a against special symbol of the
+Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
+sub-expressions.
+
+In this case study we extracted the structure information from the pattern
+matchings present on this library and included it into the generation
+specification of our random Lisp values, which were generated by randomly
+picking from a total of 6 data constructors and 8 different pattern matchings.
+%
+By doing this we obtained a consistent code coverage improvement of
+approximately $4\%$ using \dragenp with respect to the one obtained with
+\dragen (see Figure \ref{fig:coverage} (a)).
+%
+\todo[inline, author=AM]{Shall we say anything justifing this little
+  improvement?}
+
+\paragraph{Generating HTML pages using abstract interface information}
+
+For our second case study, we wanted to evaluate how including structural
+information coming from abstract interfaces when generating random HTML values
+might improve the testing performance.
+
+The library we used for this purpose represents HTML values very much in the
+same way as we exemplify in Section \ref{sec:randomtesting}, i.e., defining a
+small set of general constructions representing plain text and tags---altough
+this library also supports HTML tag attributes as well.
+%
+Then, this representation is extended with a large abstract interface consisting
+of combinators representing common HTML tags and tag attributes---equivalent to
+the combinators |div|, |bold| and |hr| illustrated before.
+
+In this case study we included the structure information present on the abstract
+interface of this library into the generation specification of random HTML
+values, resulting in a generation process that randomly picked among 4 data
+constructors and 163 abstract combinators.
+%
+With this large amount of additional structure information, we observed an
+increase of up to $83\%$ in the code coverage obtained with \dragenp with
+respect to the one observed with \dragen (see Figure \ref{fig:coverage} (b)).
+
+A manual inspection of the corpora generated with each tool revealed us that, in
+general terms, the test cases generated with \dragen rarely represent
+syntactically correct HTML values, consisting to a large extent of random
+strings within and between HTML tag delimeters (|"<"|, |">"| and |"/>"|).
+%
+On the other hand, test cases generated with \dragenp encode much more
+interesting structural information, being mostly syntactically correct.
+%
+We found that, in many cases, the test cases generated with \dragenp were
+parsed, analysed and reported as valid HTML values by the target application.
+
+With these results we are confident that including the structural information
+present on the user codebase improves the testing performance.
+%
+We consider that our approach is particularly useful when the data types
+encoding the shape of our data are vague or not sufficiently structured to be
+used to derive powerful random generators with the common type-driven
+derivation techniques.

### 2019-01-28 paper/relatedwork.tex

--- 2019-01-20 paper/main.tex
@@
-\documentclass[conference, fleqn]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% HRep
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Constructor names
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% gen
-
-% Local Variables:
-% TeX-mbigstarer: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usepackage{relsize}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITE}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\newcommand{\hrep}[1]{\ensuremath{H\!Rep_{#1}}}
-\newcommand{\gen}[1]{\ensuremath{gen_{#1}}}
-\newcommand{\term}[1]{\ensuremath{#1^\bigstar}}
-\newcommand{\evalrep}[2]{ \ensuremath{\llbracket #1 \rrbracket_{#2}}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some parameters
-\setlength{\mathindent}{\parindent}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Deriving Random Generators via\\ Higher-Level Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-We remark that, altough this work focuses on Haskell data types, this technique
-is general enough to be applied to most programming languages with some level of
-support for composite types.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data type by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract useful
-  structure information from the user source code.
-  %
-\end{itemize}
-\section{Random Data Generation in Haskell} \label{sec:randomtesting}
-
-In this section we briefly introduce the common approach for automatically
-deriving random data generators in Haskell using a type-driven approach, along
-with its main drawbacks.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-structure of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types in
-terms of \textbf{sums} (also known as \emph{variant types}) and
-\textbf{products} (or tuples) of other data types.
-%
-To exemplify this, consider the following type definition to encoding \ensuremath{\Conid{Html}}
-expressions:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Html}{}\<[12]%
-\>[12]{}\mathrel{=}{}\<[16]%
-\>[16]{}\Conid{Text}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Sing}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Join}\;{}\<[22]%
-\>[22]{}\Conid{Html}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Tag}\;{}\<[22]%
-\>[22]{}\Conid{String}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Html}} as the sum of four possible
-constructions: \ensuremath{\Conid{Text}} represents plain text values. \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Tag}} represent
-singular and paired Html tags, and \ensuremath{\Conid{Join}} concats two expression one after
-another.
-%
-We only encode a very small subset of the actual Html specification for
-illustrative reasons.
-%
-In Haskell, \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}} are known as \emph{data
-  constructors} (or constructors for short) and are used to distiguish which
-variant of the data type we are constructing.
-%
-Each data constructor is defined as a product of zero or more types known as
-\emph{fields}.
-%
-For instance, \ensuremath{\Conid{Text}} has a field of type \ensuremath{\Conid{String}}, whereas \ensuremath{\Conid{Join}} two recusive
-fields of type \ensuremath{\Conid{Html}}.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``\ensuremath{<\!\!html\!\!>\!\!hello\!\!<\!\!hr\!\!>\!\!bye\!\!<\!\!/html\!\!>}'' can be encoded as:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Conid{Tag}\;\text{\ttfamily \char34 html\char34}\;(\Conid{Join}\;(\Conid{Join}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}(\Conid{Text}\;\text{\ttfamily \char34 hello\char34})\;(\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}))\;(\Conid{Text}\;\text{\ttfamily \char34 bye\char34})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-Additionally, we can define a function \ensuremath{\Varid{render}} to serialize \ensuremath{\Conid{Html}} values as
-follows:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{render}\mathbin{::}\Conid{Html}\to \Conid{String}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Text}\;\Varid{t})\mathrel{=}\Varid{t}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Sing}\;\Varid{t})\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{render}\;\Varid{x}\plus \Varid{render}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}\plus \Varid{render}\;\Varid{x}\plus \text{\ttfamily \char34 </\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{render}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-Using pattern matching we can define functions idiomatically by defining
-different function clauses for each input pattern we are interested on.
-%
-Patterns can be defined to match specific constructors, literal values or
-variable subexpressions (like \ensuremath{\Varid{t}}, \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} in the definition of \ensuremath{\Varid{render}}).
-%
-They can also be nested in order to match very specific patterns of values.
-
-\subsection*{\textbf{Type-Driven Generation of Random Values}}
-
-In order to generate random value of types involving user defined ADTs, most
-approaches require the user to provide a random data generator for each one of
-them.
-%
-This is a cumbersome and error prone task that closely follows the data type
-structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Html}}:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{6E}{@{}l@{}}%
-\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{Html}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary})\mskip1.5mu]{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{3},{}\<[14]%
-\>[14]{}\Conid{Join}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Html}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{4},{}\<[14]%
-\>[14]{}\Conid{Tag}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-%
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
-paremetrize the generation process up to an external natural number known as the
-\emph{generation size}.
-%
-This parameter is chosen by the user, and we use to limit the maximum amount of
-recursive calls that this random generator can perform.
-%
-When called with a positive generation size, this generator can pick to generate
-among any \ensuremath{\Conid{Html}} data constructor of with a explicitly given frequency that can
-be chosen by the user (2, 1, 3 and 4 for \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}}, \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}}
-constructors, respectively).
-%
-When it picks to generate a \ensuremath{\Conid{Val}} or a \ensuremath{\Conid{Sing}} data constructor, it also
-generates a random \ensuremath{\Conid{String}} value using the standard overloaded generator
-\ensuremath{\Varid{arbitrary}} (\quickcheck provides standard random generators for most base types
-like \ensuremath{\Conid{String}}, \ensuremath{\Conid{Int}}, \ensuremath{\Conid{Bool}}, etc.). \footnote{%
-  The operators \ensuremath{\mathop{\mathsmaller{\langle \$ \rangle}}} and \ensuremath{\mathop{\mathsmaller{\langle \ast \rangle}}} are used in Haskell to combine values obtained
-  from effectful computations (like calling to a random generator) and they are
-  not particularly relevant for the point being made in this work.}
-%$
-On the other hand, when it picks to generate either an \ensuremath{\Conid{Join}} constructor, it
-also generates two independent random subexpression recursively, decreasing the
-generation size by a unit on each recursive invocation (\ensuremath{\Varid{smaller}\;gen_{Html}}).
-%
-The case of random generation of \ensuremath{\Conid{Tag}} constructors follows analogously.
-
-This random process keeps calling itself recursively until the generation size
-reaches zero, where the generator is constrained to pick only among terminal
-data constructors, being \ensuremath{\Conid{Text}} and \ensuremath{\Conid{Sing}} the only possible choices in our
-particular case.
-
-% Strictly decreasing the the generation size by one on each recursive call
-% results in a useful property over the generated data: every generated value has
-% at most |n| levels, where |n| is the generation size set by the user.
-%
-% This property enables us to model the generation process using the theory of
-% branching processes introduced by \tocite, and extended in this work as
-% described in Section \ref{sec:synthesis}.
-
-The previous definition is rather mechanical, except perhaps for the chosen
-generation frequencies.
-%
-In this light, it easy to extend this procedure to any data type defined in an
-algebraic fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth \tocite.
-%
-Given a target data type, it synthesizes a random generator for it that behaves
-similarly to the one presented above, where the generation frequencies are
-defined to be uniform across constructors.
-%
-However, picking among different data constructors with uniform frequency can
-lead to a generation process biased towards generating (in average) very small
-values, regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool this tool uses the theory of branching processes to modelize and predict
-analitically the average distribution of data constructors generated on each
-random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation frequency of each data constructor in order
-to obtain a particular distribution of values that can be specified by the user,
-providing this way a more flexible testing environment while still being mostly
-automated.
-
-Althogh both \megadeth and \dragen synthesize random generators that are
-theoretically capable to generate the whole space of values of the target data
-type. the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is often too generic to generate random data with
-enough structural complexity to be used for testing purposes.
-
-% In this work we identify two sources of additional structure information which
-% are not considered by the aforementioned automatic derivation tools to obtain
-% better random data generators:
-
-In this work we identify two patological situations which are not properly
-handled by the aforementioned derivation tools:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching specific patterns
-  of nested values.
-\item The target code encodes a significant portion of its structure on its
-  abstract interface.
-\end{enumerate}
-
-In Section \ref{sec:hrep} we show how this structural information can be used to
-synthesize richer random generators automatically.
-%
-We proceed to exemplify the previous points in detail.
-
-\subsection*{\textbf{Presence of Complex Pattern Matchings}}
-
-To exemplify the first problematic scenario, suppose we want to use randomly
-generated \ensuremath{\Conid{Htmls}}s to test a property comprising the following function:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Text}\;(\Varid{t1}\plus \Varid{t2}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y}))){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Join}\;(\Varid{simpl}\;\Varid{x})\;(\Varid{simpl}\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x})\mathrel{=}\Conid{Tag}\;\Varid{t}\;(\Varid{simpl}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function simplifies sequences of \ensuremath{\Conid{Text}} constructors into a single big
-\ensuremath{\Conid{Text}} constructor.
-%
-To do so, it has to pattern match against sequences of \ensuremath{\Conid{Text}} constructors
-combined by a \ensuremath{\Conid{Join}} constructor using nested patterns (see \ensuremath{\Varid{simpl}} fist and
-second clauses).
-%
-The remaining clauses are only meant to propagate this simplification within
-nested expressions.
-
-Ideally, we would like to test each clause of the function \ensuremath{\Varid{simpl}} approximately
-the same amount of time each.
-%
-However, each data constructor is generated independently when using either
-\megadeth or \dragen, thus the probability of generating a value satisfying a
-nested pattern decreases multiplicatively with the number of constructors we
-pattern against to simultaneously.
-%
-In our tests, we found that the first two clauses of \ensuremath{\Varid{simpl}} get exercised only
-approximately $1.5\%$ of the time when using \megadeth to derive a random
-generator for \ensuremath{\Conid{HTML}}.
-%
-On the other hand, the best result we could achieve with \dragen was only able
-to exercise the first and second clauses of \ensuremath{\Varid{simpl}} approximately $3\%$ and
-$6\%$ of the time, respectively.
-%
-With both derivation tools, the most of the time was spent testing the trivial
-clauses of our function, in view of they pattern match against simpler patterns
-of input values.
-
-% In this light, we found that using a random generator obtained with \megadeth to
-% test this function will only exercise the first two clauses of |simpl| about
-% only $1.5\%$ of the time each, relegating the rest of the generated test cases
-% to the remaining and somewhat less interesting clauses, where, remarkably, half
-% of the time is spent exercising the last trivial clause.
-% %
-% Furthermore, using \dragen to ...
-
-% This function behaves very much like an identity function, with the exception
-% that it fails for two very specific patterns of input values defined using
-% nested pattern matching.
-% %
-% This is, |foo| does not only matches against the root data constructor, but also
-% against the data constructors of its subexpressions and sub-subexpressions.
-
-% \quickcheck's default implementation for random |Int|s pick a number in the
-% interval $[-n, n]$ with uniform distribution.
-% %
-% Hence, in principle we need to recognize what is a suitable generation size,
-% which should be big enough for our generator be able to generate the numbers we
-% pattern match against to, otherwise we will not be able to generate any value to
-% match against |foo| clauses, leaving fragments of code completely untested.
-% %
-% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-% is big enough to produce an |Int| number equal to $50$.
-% %
-% Under this consideration, the probability of generating a value matching the
-% first clause of |foo| (and hence triggering the first error) results as follows:
-% %
-% \begin{align*}
-%   &P(match(foo\#1))\\
-%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
-%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
-% \end{align*}
-
-% If we use \megadeth to automatically derive a random generator for |Exp|, we
-% obtain a uniform generation probability distribution over constructors, i.e.,
-% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
-% %
-% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-% average, we will need to generate over than two million test cases in order to
-% be able to test the first clause of |foo| only once.
-% %
-% This situation can be somewhat improved if we use \dragen to obtain a random
-% generator.
-% %
-% Using this tool, we can optimize the generation probabilities in order to
-% benefit the generation of some data constructors over the rest.
-% %
-% Considering that the first pattern matching of |foo| involves the data
-% constructor |Add| as the only recursive one, we can set an target generation
-% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
-% the rest of the generated data constructors.
-% %
-% By doing so, the obtained distribution of values results such that
-% $P(match(foo\#1)) \approx 1/300000$.
-% %
-% Although this certainly improves the probability of generating a matching value,
-% this probability is not substantial enough to become practical.
-
-% Additionally, by favoring the generation probabilities towards the |Add|
-% constructor, we found that the probability of generating a value matching the
-% second clause of |foo| (which does not matches against it) also gets diminished
-% into an impractival value.
-
-Altough the previous example might seem rather simple, branching against
-specific patterns of the input data is a common task.
-%
-For instance, balancing a Red-Black tree requires to consider specific
-combinations of color, left and right subtrees and sub-subtrees in order to
-preserve the height invariant \tocite.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few \tocite.
-
-\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
-
-A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions which manipulates
-its values.
-%
-For this purspose, suppose we extend our \ensuremath{\Conid{Html}} data type with the following
-basic combinators:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 div\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 b\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathbin{::}\Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathrel{=}\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-% In the previous definition, the |Html| data type is defined using a single data
-% constructor that contains the textual representation of the Html code it
-% represents in plain text.
-%
-These functions encode additional information about the structure of our \ensuremath{\Conid{Html}}
-data type in the form of specific Html tags.
-
-Instead of including a new data constructor for each possible Html tag in out
-type definition, we defined a mininal representation and then extended it with
-an set of high level combinators.
-%
-% These combinators functions defined over |HTML| are the ones in charge of
-% transforming this plain text representation with the invariant that, given valid
-% |Html| parameters, they always return a valid |Html| value.
-%
-This programming pattern is often called a ``shallow embedding'', and can be
-found in a variety of Haskell libraries, being \emph{html} \tocite,
-\emph{svg-builder} \tocite some examples of this.
-
-As a consequence of this practice, type-driven derivation techniques often fail
-to synthesize useful random generators due to that most of the data type
-structure has been encoded into its abstract interface of combinators.
-%
-In our particular case, the chances of generating a \ensuremath{\Conid{Tag}} value representing a
-commonly used Html tag such as \ensuremath{\Varid{div}} are extremely low.
-
-% Additionally, if our test suite contains properties constrained by certain
-% preconditions, the lack of domain knowleadge may lead in an impractically high
-% discard ratio of randomly generated test cases.
-% %%
-% For instance, suposse we write a property to verify that |render| always outputs
-% a valid |Html| value:
-
-% \begin{code}
-% prop_render :: Html -> Bool
-% prop_render x = valid x ===> valid (parse (render x))
-% \end{code}
-
-% In the previous definition we state that |render| always returns a valid |Html|
-% value when it is parsed again (|parse (render x)|), provided that we are given a
-% valid |Html| as input (|valid x|).
-% %
-% In practice, not every string of characters constitutes a valid Html given that
-% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).
-
-% %
-% While testing this property we rarely satisfy its precondition, in which case
-% \quickcheck discards the whole test, retrying with a diferent random input,
-% degradating this way the testing performance.
-
-% \begin{code}
-% text :: String -> Bool
-% text str = Text (addEscapes str)
-% \end{code}
-
-% We believe that, if certain patterns of values are relevant enough to appear in
-% the codebase being tested, a practical random testing methodology should be able
-% to produce values satisfying such patterns in a substantial proportion.
-
-So far we have introduced two testing scenarios where type-driven derivation
-approaches are unable to capture all the available structure information from
-the user codebase.
-%
-Fortunately, this information can be automatically exploited and used to
-generate interesting and more structured random values.
-
-The next section introduces a reprentation model that let us encode the
-structure information presented here into our automatically derived random
-generators in a modular and flexible way.
-\section{Extracting Structure} \label{sec:hrep}
-
-In this section we present a compositional representation of values to express
-the random generation of values following the internal structure of their data
-types along with the structure present on patterns matchings and abstract
-interfaces.
-
-The key idea of this work is to represent different structured constructions of
-data in a homogeneous way that we call a ``higher-level representation'' (\ensuremath{\Conid{HRep}}
-from now on).
-%
-Instead of generating each data constructor independently, a random generator
-derived from this representation might generate composite structured values on
-each random choice it performs.
-%
-% \footnote{The notion of higher-level comes from that the generation process is
-%   entirely determined by the type of the chosen type level representation,
-%   instead of by a concrete generator defined at the term level.}
-%
-For this purpose, we use a series of automatically derived data types, each one
-representing an atomic unit of information that can be randomly generated and
-then reflected back to the corresponding value of the original data type.
-%
-Later, the user can compose these atomic representations using the provided type
-level combinators in different ways into a ``generation specification'' that
-completely determines the generation process behavior.
-
-We wil reuse the previously defined data type \ensuremath{\Conid{Html}} and the functions defined
-in the previous section to explain the different concepts involved all across
-this section.
-
-\subsection*{\textbf{Representing Data Constructors}}
-
-We begin by introducing the simplest data type representation that we can
-extract from our codebase: the representation of single data constructor.
-%
-Each data constructor can be represented by an automatically derived data type
-consisting of a single constructor with the same fields as the original, except
-for the recursive ones that are abstracted away.
-%
-In this light, we represent each constructor of the data type \ensuremath{\Conid{Html}} as follows:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{Text}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Text}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Sing}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Sing}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Join}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Join}\;{}\<[29]%
-\>[29]{}\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Tag}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Tag}\;{}\<[29]%
-\>[29]{}\Conid{String}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Note that the previous definitions are type parametric over the type parameter
-\ensuremath{\Varid{r}}.
-%
-This allow us to replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
-possible values on each case.
-%
-For instance, the value (\ensuremath{Con_{Join}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\hrep{Join}\;\Conid{Int}}, while the
-value (\ensuremath{Con_{Join}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\hrep{Join}\;\Conid{Bool}}.
-
-In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
-chosen generation specification (which might be composed of several \ensuremath{\Conid{HRep}}s),
-without having to modify anything in the underlying machinery.
-
-Having the \ensuremath{\Conid{HRep}} of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
-representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
-%
-Then, we simply need to translate each constructor representation back into its
-corresponding one, translating the abstracted fields recursively:
-%
-\begin{alignat*}{4}
-  &\evalrep{\ensuremath{Con_{Text}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Text}}\ s \\
-  &\evalrep{\ensuremath{Con_{Sing}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Sing}}\ s \\
-  &\evalrep{\ensuremath{Con_{Join}}\ x\ y &&}{Html}
-    &&= \ensuremath{\Conid{Join}}\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
-  &\evalrep{\ensuremath{Con_{Tag}}\ s\ x &&}{Html}
-    &&= \ensuremath{\Conid{Tag}}\ s\ \evalrep{x}{Html}
-\end{alignat*}
-
-The missing piece is to automatically synthesize a random generators for each
-constructor representation.
-%
-For this purpose, it is important to consider that each constructor \ensuremath{\Conid{HRep}} has
-its recursive fields abstracted away with a type parameter that will be later
-instantiated with the generation specification type.
-%
-Given that this specification is unknown at the derivation time, we parametrize
-each \ensuremath{\Conid{HRep}} generator with a random generator \ensuremath{\gen{r}} that is used to generate
-random values for each recursive field:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{Text}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Text}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Sing}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Sing}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Join}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Join}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{Tag}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Tag}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Type Level Combinators}}
-
-The next step is to define a series of type level combinators to enable us
-combining the automatically derived \ensuremath{\Conid{HRep}}s in several ways.
-
-In first place, we define a type combinator (\ensuremath{\term{\_}}) to tag a \ensuremath{\Conid{HRep}} to be
-terminal, i.e., a representation that is allowed be generated when the
-generation size gets exausted:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\term{f})\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\Conid{HRep}} with an
-explicit generation frequency $n$:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The previous combinators only include information relevant to the generation
-process, in a sense that neither one adds new structure to the final
-representation.
-%
-In this light, they do not alter the evaluation semantics, and we translate them
-back to our target data type by evaluating the inner representation:
-%
-\begin{alignat*}{4}
-  &\evalrep{Term\ x &&: \ensuremath{\term{f}}    &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
-\end{alignat*}
-
-In the previous equations we explicitly annotate (using a colon) the type of the
-evaluated term for clarity.
-%
-Later, to generate these combinators is enough to wrap a generated value from
-the inner representation with the apropriate tag:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{f}^\bigstar\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Term}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\gen{f \otimes n}\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Freq}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-\ensuremath{\Conid{HRep}}s into a single one using a sum type to represent a random choice between
-them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\oplus\;\Varid{g})\;\Varid{a}\mathrel{=}\Conid{L}\;(\Varid{f}\;\Varid{a})\mid \Conid{R}\;(\Varid{g}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data constructor variant and
-evaluating the inner \ensuremath{\Conid{HRep}} accodingly:
-%
-\begin{alignat*}{3}
-  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
-  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
-\end{alignat*}
-
-Generating a composite \ensuremath{\Conid{HRep}} is slightly more complicated than before, as we
-need to perform a random choice based on the generation size and the given
-frequencies for each sub-represention:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{43}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{f \oplus g}\;\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-In the previous definition, we reflect the type level frequencies of the types
-\ensuremath{\Varid{f}} and \ensuremath{\Varid{g}} (\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}) into term-level values.
-%
-This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
-%
-Then we use these frequencies to generate each inner \ensuremath{\Conid{HRep}} in the apropriate
-proportion.
-%
-When the generation size gets exhausted, we reflect the terminal generation
-frequency of each inner \ensuremath{\Conid{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
-\ensuremath{\Varid{freq0}\;\textit{@}\!g}).
-%
-This time, however, we default the frequency reflection to $0$ for any inner
-that not tagged as terminal, avoiding to generate non-terminal constructions in
-the last step.
-
-Later, we can use these combinators to create a type synonym \ensuremath{\hrep{Html}} that
-specifies a generation schema equivalent to the one seen in the concrete random
-generator of type \ensuremath{\Conid{Html}} presented in Section \ref{sec:randomtesting}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{Html}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\term{\hrep{Text}}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\term{\hrep{Sing}}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Join}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{3}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Tag}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{4}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{\hrep{Html}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit at the definition of \ensuremath{\hrep{Html}}.
-%
-We can think of it as a ``single layer'' of representation.
-%
-To make it able to represent recursive values we need to define a last type
-level combinator to ``tie the knot'':
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Fix}\;\Varid{f}\mathrel{=}\Conid{Fix}\;(\Varid{f}\;(\Conid{Fix}\;\Varid{f})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This datatype represents the \emph{fixed point} of a parametric data type \ensuremath{\Varid{f}},
-i.e., a data type where each recursive call gets instantiated with iself.
-%
-Then, to translate fixed points back to our target data type we simply need
-translate the inner representation:
-%
-\begin{align*}
-  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
-\end{align*}
-
-% Categorically, this generic transformation of fixed points is called a
-% \emph{catamorphism}, where our evaluation relation is known as its
-% \emph{algebra} \tocite.
-
-Unlike the other combinators, to define a random generator for the fixed point
-of a certain representation \ensuremath{\Varid{f}}, we do not to parametrize the generation of the
-recursive fields of \ensuremath{\Varid{f}} over an external generator \ensuremath{\gen{r}}.
-%
-% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
-Instead, we replace it with our fixed point generator, calling itself
-recursively on any recursive field of \ensuremath{\Varid{f}} that might appear inside:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;genFix_{f}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This way we obtain a concrete recursive generator for each representation \ensuremath{\Varid{f}}
-that we define.
-%
-Then, we can define a random generator for our target data type simply by
-generating a random value of our chosen representation, and transforming it back
-to our target by the means of the evaluation relation:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Html}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{\hrep{Html}}{}\<[E]%
-\\
-\>[17]{}\Varid{return}\;\evalrep{x}{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-So far we have introduced the machinery required to represent the random
-generation of a target data type considering only the structure encoded on its
-definition.
-%
-However, this approach can now be easily extended to encode different sources of
-structured information.
-%
-We proceed to introduce two extensions that help to address the problematic
-testing scenarios presented in the previous section.
-
-\subsection*{\textbf{Representing Pattern Matchings}}
-
-We can follow a similar reasoning as before to represent the pattern matching
-structure from a given function.
-%
-Consider the nested pattern matchings of function \ensuremath{\Varid{simpl}} defined in the
-previous section:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{36}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-To represent these patterns, we derive a new data for each one of them, whose
-fields represent the pattern variables that occur inside, and abstracting away
-every pattern variable of type \ensuremath{\Conid{Html}}.
-%
-Concretely, we define the following data types:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{simpl\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#1}\;\Conid{String}\;\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{simpl\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#2}\;\Conid{String}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The first pattern of \ensuremath{\Varid{simpl}} contains two pattern variables (\ensuremath{\Varid{t1}} and \ensuremath{\Varid{t2}}) of
-type \ensuremath{\Conid{String}} that are included as fields of \ensuremath{Pat_{simpl\#1}}.
-%
-Similarly, the second pattern of \ensuremath{\Varid{simpl}} contains a pattern variable \ensuremath{\Varid{t1}} of
-type \ensuremath{\Conid{String}} represented by the first field of \ensuremath{Pat_{simpl\#2}}, along with two
-pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type \ensuremath{\Conid{Html}} that are abstracted in the second
-and third fields of \ensuremath{Pat_{simpl\#2}}.
-
-To transform these representations back to our target data type, we simply need
-to expand them it into the concrete value represented by the original pattern,
-evaluating its fields back to \ensuremath{\Conid{Html}} as well:
-%
-\begin{align*}
-  &\evalrep{\ensuremath{Pat_{simpl\#1}}\ t1\ t2}{Html} \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html})\;(\Conid{Text}\;\evalrep{t2}{Html})}\\
-  &\evalrep{\ensuremath{Pat_{simpl\#2}}\ t1\ x\ y}{Html} = \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html}))\;\evalrep{x}{Html})\;\evalrep{y}{Html})}
-\end{align*}
-
-This way, any value of type \ensuremath{\hrep{simpl\#1}} is guaranteed to satisfy the first
-pattern matching of \ensuremath{\Varid{simpl}}---the same property follows for \ensuremath{\hrep{simpl\#2}}.
-
-The random generation of pattern \ensuremath{\Conid{HRep}}s is defined in the same way as we did
-before for representations of data constructors:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{simpl\#1}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#1}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[B]{}\gen{simpl\#2}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#2}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-Finally, we can join the pattern matching representation of each clause of
-\ensuremath{\Varid{simpl}} into a single one:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{simpl}{}\<[17]%
-\>[17]{}\mathrel{=}\hrep{simpl\#1}\;\!\!\oplus\!\!\;{}\<[39]%
-\>[39]{}\hrep{simpl\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Representing Abstract Interfaces}}
-
-To introduce the higher level representation of module abstract interfaces,
-consider module \ensuremath{\Conid{M}} defininf \ensuremath{\Conid{Html}} combinators introduced in the previous
-section.
-
-% \begin{code}
-% module M where
-
-% ten :: Exp
-% ten = Val 10
-
-% square :: Exp -> Exp
-% square x = Mul x x
-
-% minus :: Exp -> Exp -> Exp
-% minus x y = Add x (Mul y (Val (-1)))
-% \end{code}
-
-We can represent each function of \ensuremath{\Conid{M}} that returns a value of type \ensuremath{\Conid{Html}}
-following the same idea as before, deriving a data type with a single data
-constructor for each one of them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{div}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{div}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{bold}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{bold}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{hr}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In this case, each single constructor will have as fields the types of the
-inputs of the function that they represent.
-%
-As before, we abstract away any field representing and input of type \ensuremath{\Conid{Html}} with
-a type parameter \ensuremath{\Varid{r}}.
-
-To evaluate these representations, we simply use the values on its fields as
-input parameter of each original function, returning its result.
-%
-\begin{alignat*}{3}
-  &\evalrep{\ensuremath{Fun_{div}}\  &&x  &&}{Html} = \ensuremath{\Varid{div}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{bold}}\ &&x &&}{Html} = \ensuremath{\Varid{bold}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{hr}}    &&  &&}{Html} = \ensuremath{\Varid{hr}}
-\end{alignat*}
-
-Note that, by doing this, the generation process inherits any patology that the
-functions we use to generate values might have.
-%
-For instance, if the function \ensuremath{\Varid{div}} would happen to be non-terminating for some
-inputs, our generation process could suffer from this as well.
-
-Furthermore, the generation procedure for abstract interface representations
-follows the same pattern as before:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{div}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{div}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{bold}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{bold}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{hr}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}\Varid{return}\;Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-And we will also define a type synonym to join all the representations of the
-module \ensuremath{\Conid{M}} into a single one.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{50}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{M}{}\<[13]%
-\>[13]{}\mathrel{=}\hrep{div}\;\!\!\oplus\!\!\;{}\<[32]%
-\>[32]{}\hrep{bold}\;\!\!\oplus\!\!\;{}\<[50]%
-\>[50]{}\hrep{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Finally, we can put all the derived machinery together into a generation
-specification \ensuremath{Html_S}, assigning (possibly) different generation frequencies to
-each individual \ensuremath{\Conid{HRep}} we combine:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;Html_S{}\<[13]%
-\>[13]{}\mathrel{=}{}\<[21]%
-\>[21]{}\hrep{Html}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{simpl}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{M}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% This previous definition can be interpreted graphically as it is shown in the
-% Figure \ref{fig:hrep}, where curly arrows represent the structural information
-% extracted using meta-programming.
-
-% \begin{figure}[t]
-%   \centering
-%   \input{tikz/hrep.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-
-We want to remark that, for space reasons, we were only able to introduce the
-representation of a rather simple target data type.
-%
-In practice, this reasoning can be extended to mutually recursive and parametric
-types as well.
-
-Overall, this approach offers significant advantages over the usual type-driven
-derivation of random generators:
-%
-\begin{itemize}
-\item \textbf{Composability:} we can combine different atomic representations
-  using different structure information sources depending on what property or
-  sub-system we need to verify using randomly generated values.
-  %
-\item \textbf{Extensibility:} the developer can derive representations for new
-  sources of structure information and combine them with the existing ones
-  simply by adding them to the existing generation specification of the target
-  data type.
-  %
-\item \textbf{Predictability:} using branching processes theory, it is possible
-  to predict the average distribution of generated values in terms of number of
-  constructors.
-  %
-  This prediction is completely modular, and can be obtained for any composite
-  representation obtained using the automatically derived \ensuremath{\Conid{HRep}}s and the
-  provided combinators.
-\end{itemize}
-
-\todo[inline]{Maybe we can show an example random value from the representation
-  and its corresponding target value }
-
-% \begin{align*}
-%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-% \end{align*}
-%
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-%
-% \begin{code}
-% data (Term f) a = TagTerm (f a)
-
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-% instance (Term f down Exp) where
-%   step (TagTerm f) = step f
-% \end{code}
-
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
-
-% \begin{figure}[b]
-%   \centering
-%   \input{tikz/val.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-% %include sections/casestudies.tex
 \section{Related Work}
@@
 Nam euismod tellus id erat.
-\section{Final Remarks}
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
+tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
+eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
+parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
+Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
+Nam vestibulum accumsan nisl.
 
@@
 ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}
+vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
+orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
+est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

### 2019-01-28 paper/synthesis.tex

\section{Random Generators Prediction and Synthesis} \label{sec:synthesis}

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{align*}
  % M_{|Html|} =
  \begin{tikzpicture}[baseline=-0.65ex]
    \matrix
    [ matrix of math nodes
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      M_{ADT\!\rightarrow\!ADT} \&
      M_{ADT\!\rightarrow\!Patterns} \&
      M_{ADT\!\rightarrow\!Interface} \\
      M_{Patterns\!\rightarrow\!ADT} \&
      M_{Patterns\!\rightarrow\!Patterns} \&
      M_{Patterns\!\rightarrow\!Interface} \\
      M_{Interface\!\rightarrow\!ADT} \&
      M_{Interface\!\rightarrow\!Patterns} \&
      M_{Interface\!\rightarrow\!Interface} \\
    };
  \end{tikzpicture}
\end{align*}

% \begin{align*}
%   M_C
% %   % =
% %   % \begin{bmatrix}
% %   %   m_{ij}  \\
% %   % \end{bmatrix}_{i, j \in \{|Leaf|, |NodeA|, |NodeB|\}}
% %   = \qquad\quad

% %   % \begin{tikzpicture}[baseline=-0.65ex]
% %   %   \matrix[
% %   %   matrix of math nodes,
% %   %   column sep=0ex,
% %   %   left delimiter={[},
% %   %   right delimiter={]}
% %   %   ] (m)
% %   %   {
% %   %     0               & 0                & 0\\
% %   %     2 \cdot p_{|Leaf|} & 2 \cdot p_{|NodeA|} &2 \cdot p_{|NodeB|}\\
% %   %      p_{|Leaf|}  &  p_{|NodeA|} &  p_{|NodeB|}\\
% %   %   };
% %   %   % \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle |Leaf|$};
% %   %   % \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle |NodeA|$};
% %   %   % \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle |NodeB|$};
% %   %   % \node[left,overlay] at ([xshift=-4.5ex]m-1-1.west) {$\scriptstyle |Leaf|$};
% %   %   % \node[left,overlay] at ([xshift=-1.5ex]m-2-1.west) {$\scriptstyle |NodeA|$};
% %   %   % \node[left,overlay] at ([xshift=-3.0ex]m-3-1.west) {$\scriptstyle |NodeB|$};
% %   % \end{tikzpicture}
% %   \label{eq:treepmatrix}
% \end{align*}

### 2019-01-29 paper/casestudies.tex

@@
 
-Intuitively, adding structure information to a random generator should provide
-better results when verifying a system using random testing.
+% Intuitively, adding structure information to a random generator should provide
+% better results when verifying a system using random testing.
 %
@@
 %
-This section describes two case studies showing that considering structure
-information when deriving generators consistenly produces better testing
-results.
-
-Instead of restricting our scope to Haskell, in this work we followed a broader
-evaluation approach taken previously to compare the performance of other
-state-of-the-art derivation techniques
+This section describes two case studies showing that considering additional
+structural information when deriving generators can consistently produce better
+testing results in terms of code coverage.
+%
+Instead of restricting our scope to Haskell, in this work we follow a broader
+evaluation approach taken previously to compare state-of-the-art techniques to
+derive random data generators based on ADT definitions
 \cite{grieco2017,DBLP:conf/haskell/MistaRH18}.
 %
-Concretely, we evaluate how including additional structure information when
-generating a set of random test cases (often referred as a \emph{corpus})
-affects the code coverage obtained when using them to test a given target
-program.
+
+We evaluate how including additional structural information when generating a
+set of random test cases (often referred as a \emph{corpus}) affects the code
+coverage obtained when testing a given target program.
 %
-For this purpose, we targeted two external programs which expect highly
-structured inputs, namely \emph{GNU CLISP}---the GNU Common Lisp compiler, and
-\emph{HTML Tidy}---a well known HTML refactoring and correction utility.
+For that, we considered two external programs which expect highly structured
+inputs, namely \emph{GNU CLISP}---the GNU Common Lisp
+compiler\footnote{\href{https://www.gnu.org/software/gcl/}{https://www.gnu.org/software/gcl/}},
+and \emph{HTML Tidy}
+\footnote{\href{http://www.html-tidy.org}{http://www.html-tidy.org}}---a well
+known HTML refactoring and correction utility.
 %
@@
 structure, i.e., Lisp and HTML values respectively. These libraries are:
-\emph{hs-zuramaru}, implementing an embedded Lisp intepreter for a small subset
-of this programming language, and \emph{html}, defining a combinator library for
-constructing HTML values.
+\emph{hs-zuramaru}\footnote{\href{http://hackage.haskell.org/package/zuramaru}{http://hackage.haskell.org/package/zuramaru}},
+implementing an embedded Lisp intepreter for a small subset of this programming
+language, and
+\emph{html}\footnote{\href{http://hackage.haskell.org/package/html}{http://hackage.haskell.org/package/html}},
+defining a combinator library for constructing HTML values.
 %
@@
 
-We started by compiling instrumented versions of the target programs in a way
-that they also return the execution path followed in the source code every time
-we run them with a given input test case.
+We firstly compiled instrumented versions of the target programs in a way that
+they also return the execution path followed in the source code every time we
+run them with a given input test case.
 %
@@
 %
-On the other side, we used the ADTs defined on the chosen libraries to derive
-random generators using \dragen and \dragenp, including structure information
+We then used the ADTs defined on the chosen libraries to derive random
+generators using \dragen and \dragenp, including structural information
 extracted from the library's codebase in the case of the latter.
 %
-Then, we evaluated the code coverage triggered by independent, randomly
-generated corpora of different sizes varying from 100 to 1000 test cases each.
+Then, we proceeded to evaluate the code coverage triggered by independent,
+randomly generated corpora of different sizes varying from 100 to 1000 test
+cases each.
 %
-In order to remove any external bias, we derived random generators optimized to
-follow a uniform distribution of constructors (and pattern matchings or function
+In order to remove any external bias, we derived generators optimized to follow
+\emph{a uniform distribution of constructors (and pattern matchings or function
   calls in the case \dragenp), and carefully adjusted their generation sizes to
-match the average test case size in bytes.
+  match the average test case size in bytes}.
 %
 This way, any noticeable difference in the code coverage can be attributed to
-the present (or lack thereof) structure information when generating the test
+the present (or lack thereof) structural information when generating the test
 cases.
@@
 \end{figure*}
-
+%
 Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
-triggered for each permutation of corpus size and derivation tool, including
+triggered for different sizes of the corpus size and derivation tool, including
 error bars indicating the standard error of the mean on each case.
 %
-We proceed to describe each case study in detail.
+We proceed to describe each case study and our findings in detail as follows.
 
-\paragraph{Generating Lisp code using pattern matching information}
+\subsection{Branching on input data}
 
 In this first case study we wanted to evaluate the observed code coverage
-differences when considering structure information present on functions pattern
+differences when considering structural information present on functions pattern
 matchings.
@@
 sub-expressions.
-
-In this case study we extracted the structure information from the pattern
-matchings present on this library and included it into the generation
-specification of our random Lisp values, which were generated by randomly
+%
+We extracted this structural information and included it into the generation
+specification of our random Lisp values---which were generated by randomly
 picking from a total of 6 data constructors and 8 different pattern matchings.
 %
-By doing this we obtained a consistent code coverage improvement of
-approximately $4\%$ using \dragenp with respect to the one obtained with
-\dragen (see Figure \ref{fig:coverage} (a)).
+By doing this, we obtained a code coverage improvement of approximately $4\%$
+using \dragenp with respect to the one obtained with \dragen (see Figure
+\ref{fig:coverage} (a)).
 %
-\todo[inline, author=AM]{Shall we say anything justifing this little
-  improvement?}
+While it seems a little improvement, it is worth noticing that we are generating
+highly structured input, i.e., Lisp programs.
+%
+When testing interpreters, we argue that an improvement of $4\%$ is not
+negligible considering the little effort that took us to specify the pattern
+matchings.
+%\todo[inline, author=AM]{Shall we say anything justifing this little
+%  improvement?}
 
-\paragraph{Generating HTML pages using abstract interface information}
+\subsection{Abstract interfaces}
 
@@
 same way as we exemplify in Section \ref{sec:randomtesting}, i.e., defining a
-small set of general constructions representing plain text and tags---altough
+small set of general constructions representing plain text and tags---although
 this library also supports HTML tag attributes as well.
@@
 of combinators representing common HTML tags and tag attributes---equivalent to
-the combinators |div|, |bold| and |hr| illustrated before.
+the combinators |div|, |bold| and |hr| illustrated in Section \ref{sec:sources}.
 
@@
 %
-With this large amount of additional structure information, we observed an
+With this large amount of additional structural information, we observed an
 increase of up to $83\%$ in the code coverage obtained with \dragenp with
 respect to the one observed with \dragen (see Figure \ref{fig:coverage} (b)).
-
+%
 A manual inspection of the corpora generated with each tool revealed us that, in
@@
 encoding the shape of our data are vague or not sufficiently structured to be
-used to derive powerful random generators with the common type-driven
-derivation techniques.
+used to derive powerful random generators with the common derivation techniques
+based only on ADT definitions.

### 2019-01-29 paper/synthesis.tex

@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+Characterizing the distribution of values of an arbitrary random generator is a
+hard task.
+%
+It requires to model every random choice that it could possibly make to generate
+a value.
+%
+In a recent work \tocite, we have shown that it is possible to
+\emph{analitically} predict the average distribution of data constructors
+produced by a random generators automatically derived considering only ADT
+definitions---like the one presented on Section \ref{sec:randomtesting}.
+%
+For this purpose, we found that random generation of ADT values can be
+characterized using the theory of \emph{branching processes} \tocite.
+%
+This probabilistic theory was originally conceived to predict the growth and
+extinction of royal family trees the Victorian Era, being later applied to a
+wide variety of research areas.
+%
+In this work we extend this model to predict the average distribution of values
+of random generators derived considering structural information coming from
+pattern matchings and abstract interfaces.
+
+Essentially, a branching process is a special kind of Markov process that models
+the evolution of a population of \emph{individuals} of different kinds across
+discrete time steps known as \emph{generations}.
+%
+Each kind of individual is expected to produce an average number of offspring of
+(possibly) different kinds from one generation to the next one.
+%
+In our original setting, we shown that each different data constructor can be
+considered as an individual on its own.
+%
+On the other hand, any ADT value can be seen as a tree where each node
+represents a root data constructor and has its sub-expressions as
+sub-trees---hence note the similarity with family trees.
+%
+In this light, each tree level of a generated value can be seen as a generation
+of our generation model.
+
+We note $G_n$ to the vector of real numbers that represents the average
+distribution of constructors that a random generator produces in the $n$-th
+generation, where each element of this vector represents each different data
+constructor involved in the generation process.
+%
+Then, it is possible to predict expected number of constructors a the generation
+$n$ by the means of the next formula:
+%
+\begin{align*}
+  E[G_n]^T = E[G_0]^T \cdot M^n
+\end{align*}
 
 \begin{align*}
-  % M_{|Html|} =
-  \begin{tikzpicture}[baseline=-0.65ex]
+  E \left[ P_n \right]
+  =
+  \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
+\end{align*}
+
+\begin{align*}
+  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n+1}
+\end{align*}
+
+\newcommand{\ph}{\phantom{\square}}
+\newcommand{\el}{\square}
+\begin{equation*}
+  M_{T} = \ \
+  \begin{tikzpicture}
+    [ baseline=-0.65ex
+    , every left delimiter/.style={xshift=.5em}
+    , every right delimiter/.style={xshift=-.5em}
+    ]
     \matrix
     [ matrix of math nodes
+    , column sep=-1pt
+    , row sep=-1pt
     , ampersand replacement=\&
@@
     {
-      M_{ADT\!\rightarrow\!ADT} \&
-      M_{ADT\!\rightarrow\!Patterns} \&
-      M_{ADT\!\rightarrow\!Interface} \\
-      M_{Patterns\!\rightarrow\!ADT} \&
-      M_{Patterns\!\rightarrow\!Patterns} \&
-      M_{Patterns\!\rightarrow\!Interface} \\
-      M_{Interface\!\rightarrow\!ADT} \&
-      M_{Interface\!\rightarrow\!Patterns} \&
-      M_{Interface\!\rightarrow\!Interface} \\
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
+      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
     };
+    % Color squares
+    \fill[red, opacity=0.075]
+    (m-1-1.north west)
+    -|| ([xshift=-1pt]m-1-3.north east)
+    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
+    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
+    -|| cycle;
+    \fill[blue, opacity=0.075]
+    ([yshift=-1pt]m-4-1.north west)
+    -|| ([xshift=1pt]m-1-3.north east)
+    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
+    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
+    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
+    -|| cycle;
+    % Upper labels
+    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
+    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
+    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
+    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
+    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
+    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
+    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
+    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
+    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
+    % Left labels
+    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
+    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
+    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
+    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
+    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
+    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
+    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
+    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
+    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
+    % Dashed lines
+    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
+    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
+    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
+    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
+    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
+    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
+    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
+    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
   \end{tikzpicture}
+\end{equation*}
+
+\begin{equation*}
+  \beta_{Html} = \qquad \qquad
+  \begin{tikzpicture}
+    [ baseline=-0.65ex
+    , every left delimiter/.style={xshift=.5em}
+    , every right delimiter/.style={xshift=-.5em}
+    ]
+    \matrix
+    [ matrix of math nodes
+    , column sep=0ex
+    , row sep=-1pt
+    , ampersand replacement=\&
+    , left delimiter={[}
+    , right delimiter={]}
+    ] (v)
+    {
+      0 \\ % Text
+      0 \\ % Sing
+      1 \\ % Tag
+      2 \\ % Join
+      0 \\ % simplify#1
+      2 \\ % simplify#1
+      0 \\ % hr
+      1 \\ % div
+      1 \\ % bold
+    };
+    \draw[densely dashed] ([xshift=-3pt]v-4-1.south west) -- ([xshift=3pt]v-4-1.south east);
+    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
+    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
+    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
+    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
+    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
+    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
+    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
+    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
+    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
+    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
+  \end{tikzpicture}
+  \hspace{20pt}
+  \mathcal{P}_{Html} =
+  \begin{tikzpicture}
+    [ baseline=-0.65ex
+    , every left delimiter/.style={xshift=.5em}
+    , every right delimiter/.style={xshift=-.5em}
+    ]
+    \matrix
+    [ matrix of math nodes
+    , column sep=0ex
+    , row sep=-1pt
+    , ampersand replacement=\&
+    , left delimiter={[}
+    , right delimiter={]}
+    ] (v)
+    {
+      p_{Text} \\
+      p_{Single} \\
+      p_{Tag}  \\
+      p_{Join} \\
+      p_{simplify\#1} \\
+      p_{simplify\#2} \\
+      p_{hr} \\
+      p_{div} \\
+      p_{bold} \\
+    };
+    \draw[densely dashed] ([xshift=-15pt]v-4-1.south west) -- ([xshift=15pt]v-4-1.south east);
+    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
+  \end{tikzpicture}
+\end{equation*}
+
+\paragraph{End-to-End Prediction}
+%
+% So far we have shown that our prection model is able to predict the distribution
+% of data constructors, composite pattern values and interface function calls that
+% our automatically derived random generators will produce on average.
+
+It is possible to provide further ``end-to-end'' guarantees on the generated
+values if we restrict the generation process to only bare data constructors and
+pattern matching values.
+%
+In particular, we can stop considering pattern matching values as atomic
+constructions and start seeing them as compositions of data constructors.
+%
+This way is possible to obtain the average ``total'' number of data constructors
+that our generators will produce, regardless of if they are generated on their
+own, or as part of a pattern matching value.
+%
+We note this number as $E^\downarrow\![\_]$.
+
+The prediction only needs to consider the prediction of each data constructor
+and add the average number of constructors that we will additionally generate
+within pattern matching values.
+%
+For instance, we can calculate the total number of constructors |Text| and
+|Join| that we will generate by expanding the average number of pattern values
+$simplify\#1$ and $simplify\#2$ that we generate on average as follows:
+%
+{\small
+  \begin{align*}
+    E^\downarrow\!\left[ Text \right]
+    = E\left[ Text \right]
+    + 2\! \cdot\! E\left[ simplify\#1 \right]
+    + E\left[ simplify\#2 \right] \\
+    E^\downarrow\!\left[ Join \right]
+    = E\left[ Join \right]
+    + E\left[ simplify\#1 \right]
+    + 2\! \cdot\! E\left[ simplify\#2 \right]
   \end{align*}
+}
+%
+Where $simplify\#1$ carries two |Text| and one |Join| data constructors---and
+viceversa for $simplify\#2$.
 
-% \begin{align*}
-%   M_C
-% %   % =
-% %   % \begin{bmatrix}
-% %   %   m_{ij}  \\
-% %   % \end{bmatrix}_{i, j \in \{|Leaf|, |NodeA|, |NodeB|\}}
-% %   = \qquad\quad
-
-% %   % \begin{tikzpicture}[baseline=-0.65ex]
-% %   %   \matrix[
-% %   %   matrix of math nodes,
-% %   %   column sep=0ex,
-% %   %   left delimiter={[},
-% %   %   right delimiter={]}
-% %   %   ] (m)
-% %   %   {
-% %   %     0               & 0                & 0\\
-% %   %     2 \cdot p_{|Leaf|} & 2 \cdot p_{|NodeA|} &2 \cdot p_{|NodeB|}\\
-% %   %      p_{|Leaf|}  &  p_{|NodeA|} &  p_{|NodeB|}\\
-% %   %   };
-% %   %   % \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle |Leaf|$};
-% %   %   % \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle |NodeA|$};
-% %   %   % \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle |NodeB|$};
-% %   %   % \node[left,overlay] at ([xshift=-4.5ex]m-1-1.west) {$\scriptstyle |Leaf|$};
-% %   %   % \node[left,overlay] at ([xshift=-1.5ex]m-2-1.west) {$\scriptstyle |NodeA|$};
-% %   %   % \node[left,overlay] at ([xshift=-3.0ex]m-3-1.west) {$\scriptstyle |NodeB|$};
-% %   % \end{tikzpicture}
-% %   \label{eq:treepmatrix}
-% \end{align*}
+Note that this expansion cannot be applied if we also generate random values
+containing abstract interface function calls, as we cannot predict the output of
+an arbitrary function without solving the halting problem in the process.

### 2019-01-29 paper/synthesis.tex
WIP

@@
-\section{Random Generators Prediction and Synthesis} \label{sec:synthesis}
+\section{Prediction of Distributions} \label{sec:synthesis}
 
@@
 of random generators derived considering structural information coming from
-pattern matchings and abstract interfaces.
+functions pattern matchings and abstract interfaces.
 
 Essentially, a branching process is a special kind of Markov process that models
-the evolution of a population of \emph{individuals} of different kinds across
+the evolution of a population of \emph{individuals of different kinds} across
 discrete time steps known as \emph{generations}.
@@
 %
-In this light, each tree level of a generated value can be seen as a generation
-of our generation model.
+In this light, each tree level of a random value can be seen as a generation of
+individuals in this model.
 
-We note $G_n$ to the vector of real numbers that represents the average
-distribution of constructors that a random generator produces in the $n$-th
-generation, where each element of this vector represents each different data
-constructor involved in the generation process.
+Then, we can characterize the distribution of constructors that a random
+generator produces in the $n$-th generaton as a vector $G_n$ that groups the
+number of constructors of each kind produced in that generation.
+We note $G_n$ to the vector of real numbers that represents the distribution of
 %
-Then, it is possible to predict expected number of constructors a the generation
-$n$ by the means of the next formula:
+Then, the we can predict the expected distribution of constructors at the
+generation $n$ (noted as $E[G_n]$) using the following characteristic equation
+from branching processes theory:
 %
-\begin{align*}
+\begin{align}
   E[G_n]^T = E[G_0]^T \cdot M^n
-\end{align*}
+  \label{eqn:prediction}
+\end{align}
 
-\begin{align*}
-  E \left[ P_n \right]
-  =
-  \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
-\end{align*}
+Where $E[G_0]$ represents the initial distribuition of constructors that our
+generator produces, which simply consists of the generation probability of each
+one.
+%
+The interesting aspect of the prediction mechanism is encoded in the matrix $M$,
+known as a the \emph{mean matrix} of this stochastic process.
+%
+$M$ is a squared matrix with as rows and columns as different data constructors
+involved in the generation process.
+%
+Each element $M_{i,j}$ of this matrix encodes the average number of data
+constructors of kind $j$ that gets generated in a given level, provided that we
+generated a constructor of kind $i$ at the previous level.
+%
+Is in this sense that this matrix encodes the ``branching'' behavior of our
+random generation from one generation to the next one.
+%
+In our previous work, we show how each element of this matrix can be calculated
+by exploiting type information from data type definitions.
+%
+For instance, the average number of |Text| data constructors that we will
+generate provided that we generated a |Join| data constructor on the previous
+level (corresponding to the element $M_{Join, Text}$) results $2\cdot p_{Text}$,
+where $2$ is the number of the recursive fields of type |Html| present on the
+constructor |Join| and $p_{Text}$ is the probability of generating the
+constructor |Text|.
+%
+This pattern can be used to build the rest of the mean matrix as well.
 
-\begin{align*}
-  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n+1}
-\end{align*}
+\todo[inline]{Complete this!}
+
+% \begin{align*}
+%   E \left[ P_n \right]
+%   = \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
+% \end{align*}
+
+Our realization in this work is that we can also consider that each different
+pattern matching value and abstract interface function call can be considered as
+another kind of individual belonging to the random generation process.
+%
+In this light, we can extend our mean matrix with this information, adding a row
+and a column for each different pattern matching value and function call that we
+add to the generation process, as we illustrate as follows:
 
@@
 \begin{equation*}
-  M_{T} = \ \
+  \hspace{20pt}
+  M = \ \
   \begin{tikzpicture}
@@
 
+To extend $M$ to consider the new possible constructions, we need to calculate
+$M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
+and $j$, respectively.
+%
+Fortunately, it is possible to follow the same pattern as we did on our original
+work, this time considering the amount of recursive pattern variables if it the
+case of a pattern matching value being a father individual, and the amount of
+recursive arguments if we deal with a abstract interface function call as a
+father individual.
+%
+For instance, the average amount of function calls to |div| that we will produce
+provided that we generated a value matching the second pattern of |simplify| as a
+father results $2\cdot p_{div}$.
+%
+Similarly, the average amount of values satisfying the first pattern matching of
+|simplify| provided that we generated a |Tag| constructor as a father results
+$1 \cdot p_{simplify\# 1}$.
+
+Furthermore, we found that the prediction process can be factored in terms of
+two vectors $\beta$ and $\mathcal{P}$, such that $\beta$ represents the number
+of recursive sub-terms of each construction that we generate, whereas
+$\mathcal{P}$ simply represents the probability of generating that construction
+(this probability can be easily calculated based on the generation frequencies
+used in the random generator).
+%
+Then, the equation \ref{eqn:prediction} can be rewriten as:
+%
+\begin{align*}
+  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
+\end{align*}
+
+The vector $\beta$ can be obtained by analyzing the shape of each source of
+structural information that we consider, and joining them together.
+%
+For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
+values are as follows (where dashed lines indicate the composition of the
+diffent sub-vectors that can be calculated separately):
+
 \begin{equation*}
-  \beta_{Html} = \qquad \qquad
+  \hspace{20pt}
+  \beta = \qquad \qquad \
   \begin{tikzpicture}
@@
   \hspace{20pt}
-  \mathcal{P}_{Html} =
+  \mathcal{P} =
   \begin{tikzpicture}
@@
 
+Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
+distribution of our random genrator in a way that can be always characterized.
+%
+In this light, \dragenp follows the same approach as \dragen, using our adapted
+prediction mechanism to tune the generation probabilities of each each source of
+structural information.
+%
+This is done by runing a simulation-based optimization process at compile-time.
+%
+This process is parametrized by the desired distribution of values set by the
+user.
+%
+This way she can specify, for instance, a uniform distribution of data
+constructors, pattern matching values and function calls or, alternative, a
+distribution of values with some constructions appearing in a different
+proportion as others, e.g., two times more functions calls to |div| than |Join|
+constructors.
+
 \paragraph{End-to-End Prediction}
@@
 In particular, we can stop considering pattern matching values as atomic
-constructions and start seeing them as compositions of data constructors.
+constructions and start seeing them as compositions of several data
+constructors.
 %
-This way is possible to obtain the average ``total'' number of data constructors
+Then, is possible to obtain the average ``total'' number of data constructors
 that our generators will produce, regardless of if they are generated on their
@@
 
-The prediction only needs to consider the prediction of each data constructor
-and add the average number of constructors that we will additionally generate
-within pattern matching values.
+The prediction only needs to add the expected number of constructors that we
+will additionally generate within pattern matching values to the prediction of
+bare data constructors.
 %
-For instance, we can calculate the total number of constructors |Text| and
-|Join| that we will generate by expanding the average number of pattern values
-$simplify\#1$ and $simplify\#2$ that we generate on average as follows:
+For instance, we can calculate the total average number of constructors |Text|
+and |Join| that we will generate simply by expanding the average number of
+generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
+corresponding data constructors:
 %
@@
   \begin{align*}
-    E^\downarrow\!\left[ Text \right]
+    E^\downarrow\!\!\left[ Text \right]
     = E\left[ Text \right]
-    + 2\! \cdot\! E\left[ simplify\#1 \right]
-    + E\left[ simplify\#2 \right] \\
-    E^\downarrow\!\left[ Join \right]
+    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
+    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
+    E^\downarrow\!\!\left[ Join \right]
     = E\left[ Join \right]
-    + E\left[ simplify\#1 \right]
-    + 2\! \cdot\! E\left[ simplify\#2 \right]
+    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
+    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
   \end{align*}
@@
 %
-Where $simplify\#1$ carries two |Text| and one |Join| data constructors---and
-viceversa for $simplify\#2$.
+Given that each time we generate a value satisfying the first pattern matching
+of the function |simplify| we add two |Text| and one |Join| data constructors to
+our random value.
+%
+The case of the second pattern matching of |simplify| follows analogously.
 

### 2019-01-29 paper/synthesis.tex
Added logo
Check point

@@
-\section{Prediction of Distributions} \label{sec:synthesis}
+\section{Predicting Distributions} \label{sec:synthesis}
 
@@
 %
-It requires to model every random choice that it could possibly make to generate
-a value.
+It requires to model every random choice that a generator could possibly make to
+generate a value.
 %
-In a recent work \tocite, we have shown that it is possible to
-\emph{analitically} predict the average distribution of data constructors
-produced by a random generators automatically derived considering only ADT
-definitions---like the one presented on Section \ref{sec:randomtesting}.
+In a recent work \cite{DBLP:conf/haskell/MistaRH18}, we have shown that it is
+possible to \emph{analitically} predict the average distribution of data
+constructors produced by a random generators automatically derived considering
+only ADT definitions---like the one presented on Section
+\ref{sec:randomtesting}.
 %
 For this purpose, we found that random generation of ADT values can be
-characterized using the theory of \emph{branching processes} \tocite.
+characterized using the theory of \emph{branching processes} \cite{gw1875}.
+%
+\footnote{
 %
@@
 extinction of royal family trees the Victorian Era, being later applied to a
-wide variety of research areas.
+wide variety of research areas.}
 %
@@
 an arbitrary function without solving the halting problem in the process.
+
+% Local Variables:
+% TeX-master: "main.lhs.tex"
+% TeX-command-default: "Make"
+% End:

### 2019-01-29 paper/synthesis.tex
WIP

@@
 %
-\footnote{
-%
 This probabilistic theory was originally conceived to predict the growth and
 extinction of royal family trees the Victorian Era, being later applied to a
-wide variety of research areas.}
+wide variety of research areas.
 %
-In this work we extend this model to predict the average distribution of values
-of random generators derived considering structural information coming from
-functions pattern matchings and abstract interfaces.
+In this work, we adapt this model to also predict the average distribution of
+values of random generators derived considering structural information coming
+from functions pattern matchings and abstract interfaces.
 
@@
 %
-In our original setting, we shown that each different data constructor can be
-considered as an individual on its own.
+Mista el at. \cite{DBLP:conf/haskell/MistaRH18} show that branching processes
+can be adapted to predict the generation of ADT values by simply considering
+each data constructor as an individual on its own.
 %
-On the other hand, any ADT value can be seen as a tree where each node
-represents a root data constructor and has its sub-expressions as
-sub-trees---hence note the similarity with family trees.
+In fact, any ADT value can be seen as a tree where each node represents a root
+data constructor and has its sub-expressions as sub-trees---hence note the
+similarity with family trees.
 %
@@
 
-Then, we can characterize the distribution of constructors that a random
-generator produces in the $n$-th generaton as a vector $G_n$ that groups the
-number of constructors of each kind produced in that generation.
-We note $G_n$ to the vector of real numbers that represents the distribution of
+We characterize the numbers of constructors that a random generator produces in
+the $n$-th generaton as a vector $G_n$, a vector that groups the number of
+constructors of each kind produced in that generation---in our |Html| example,
+this vector has four components, i.e., one for each constructor.
 %
-Then, the we can predict the expected distribution of constructors at the
-generation $n$ (noted as $E[G_n]$) using the following characteristic equation
-from branching processes theory:
+%We note $G_n$ to the vector of real numbers that represents the distribution of
+%
+From branching processes theory, the following equation captures the expected
+distribution of constructors at the generation $n$, noted $E[G_n]$, as follows:
 %
@@
 
-Where $E[G_0]$ represents the initial distribuition of constructors that our
+Vector $E[G_0]$ represents the initial distribuition of constructors that our
 generator produces, which simply consists of the generation probability of each
@@
 %
-Is in this sense that this matrix encodes the ``branching'' behavior of our
-random generation from one generation to the next one.
+In this sense, this matrix encodes the ``branching'' behavior of our random
+generation from one generation to the next one.
 %
-In our previous work, we show how each element of this matrix can be calculated
-by exploiting type information from data type definitions.
+Each element of the matrix can be automatically calculated by exploiting ADT
+definitions \cite{DBLP:conf/haskell/MistaRH18}---specially the recursive
+fields---as well as the individual probability of generating each constructor.
 %
@@
 generate provided that we generated a |Join| data constructor on the previous
-level (corresponding to the element $M_{Join, Text}$) results $2\cdot p_{Text}$,
-where $2$ is the number of the recursive fields of type |Html| present on the
-constructor |Join| and $p_{Text}$ is the probability of generating the
-constructor |Text|.
+level is calculated as follows:
+%
+$$ M_{Join, Text} = 2\cdot p_{Text} $$
+%
+where $2$ is the number of holes present when generating a partial ADT value
+|Join| (i.e., |Join hole hole|) and $p_{Text}$ is the probability of
+individually generating the constructor |Text|.
 %
@@
 
-\todo[inline]{Complete this!}
+%\todo[inline]{Complete this!}
 
@@
 
-Our realization in this work is that we can also consider that each different
-pattern matching value and abstract interface function call can be considered as
-another kind of individual belonging to the random generation process.
+In this work, we show how to naturally fit structural information beyond ADT
+definitions into the prediction mechanism of branching processes.
 %
-In this light, we can extend our mean matrix with this information, adding a row
-and a column for each different pattern matching value and function call that we
-add to the generation process, as we illustrate as follows:
+It is our realization that it suffices to simply consider \emph{each different
+  pattern matching and functional call as a kind of individual on its own}.
+%
+In that manner, we can extend our mean matrix $M$ by adding a row and a column
+for each different pattern matching and function call as follows:
 
@@
 \end{align*}
-
+%
 The vector $\beta$ can be obtained by analyzing the shape of each source of

### 2019-01-29 paper/synthesis.tex
Pass over branching processes

@@
 
+\subsection{Extending predictions for structural information}
+
 In this work, we show how to naturally fit structural information beyond ADT
@@
 
-To extend $M$ to consider the new possible constructions, we need to calculate
-$M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
-and $j$, respectively.
+Symbol $C_1 \cdots C_i$ denotes constructors, $P_1 \cdots P_j$ pattern
+matchings, and $F_1 \cdots F_k$ function calls.
 %
-Fortunately, it is possible to follow the same pattern as we did on our original
-work, this time considering the amount of recursive pattern variables if it the
-case of a pattern matching value being a father individual, and the amount of
-recursive arguments if we deal with a abstract interface function call as a
-father individual.
+The red-light colored matrix is what we had before where the light-blue colored
+cells are new---we recommend readers to get a colored copy of the paper.
 %
-For instance, the average amount of function calls to |div| that we will produce
-provided that we generated a value matching the second pattern of |simplify| as a
-father results $2\cdot p_{div}$.
+% To extend $M$ to consider the new possible constructions, we need to calculate
+% $M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
+% and $j$, respectively.
+% %
+
+The new cells are filled as before: we need to consider the amount of holes when
+generating partial pattern matching and function calls as well as their
+individual probabilities.
 %
-Similarly, the average amount of values satisfying the first pattern matching of
-|simplify| provided that we generated a |Tag| constructor as a father results
-$1 \cdot p_{simplify\# 1}$.
+% Fortunately, it is possible to follow the same pattern as we did on our original
+% work, this time considering the amount of recursive pattern variables if it the
+% case of a pattern matching value being a father individual, and the amount of
+% recursive arguments if we deal with a abstract interface function call as a
+% father individual.
+%
+For instance, if we consider $P_j$ as the second pattern of simplify and $F_1$
+as function |div|, then the marked cell above has the value $2\cdot p_{div}$,
+i.e., the amount of holes in a partially generated pattern |(Join (Join (Text s)
+hole) hole)|, where |s| is some random string, times the probability to generate
+a call to function |div|.
+%
+The rest follows analogously.
 
-Furthermore, we found that the prediction process can be factored in terms of
-two vectors $\beta$ and $\mathcal{P}$, such that $\beta$ represents the number
-of recursive sub-terms of each construction that we generate, whereas
-$\mathcal{P}$ simply represents the probability of generating that construction
-(this probability can be easily calculated based on the generation frequencies
-used in the random generator).
+% the average amount of function calls to |div| that we will produce
+% provided that we generated a value matching the second pattern of |simplify| as a
+% father results $2\cdot p_{div}$.
+%
+% Similarly, the average amount of values satisfying the first pattern matching of
+% |simplify| provided that we generated a |Tag| constructor as a father results
+% $1 \cdot p_{simplify\# 1}$.
+
+As another contribution, we found that the whole prediction process can be
+factored in terms of two vectors $\beta$ and $\mathcal{P}$, such that $\beta$
+represents the number of recursive sub-terms of each construction that we
+generate, whereas $\mathcal{P}$ simply represents the probability of generating
+that partial ADT value.%  (this probability can be easily calculated based on the
+% generation frequencies used in the random generator).
 %
@@
 %
-The vector $\beta$ can be obtained by analyzing the shape of each source of
-structural information that we consider, and joining them together.
+The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
+source of structural information that we consider, and joining them together.
 %
@@
 Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
-distribution of our random genrator in a way that can be always characterized.
+distribution of our random genrator in a way that can be always characterized
+and predicted.
 %
-In this light, \dragenp follows the same approach as \dragen, using our adapted
-prediction mechanism to tune the generation probabilities of each each source of
-structural information.
+\dragenp follows a similar approach as \dragen and uses an heuristic to tune the
+generation probabilities of each source of structural information.
 %
@@
 %
-This way she can specify, for instance, a uniform distribution of data
-constructors, pattern matching values and function calls or, alternative, a
+In this manner, developers can specify, for instance, a uniform distribution of
+data constructors, pattern matching values and function calls or, alternative, a
 distribution of values with some constructions appearing in a different
@@
 
-\paragraph{End-to-End Prediction}
+\subsection{Overall prediction}
+%\paragraph{End-to-End Prediction}
 %
@@
 
-It is possible to provide further ``end-to-end'' guarantees on the generated
-values if we restrict the generation process to only bare data constructors and
-pattern matching values.
+It is possible to provide an overall prediction of the expected number of
+constructors when restricting the generation process to only bare data
+constructors and pattern matching values.
 %
-In particular, we can stop considering pattern matching values as atomic
-constructions and start seeing them as compositions of several data
-constructors.
+To achieve that, we should stop considering pattern matching values as atomic
+values and start seeing them as compositions of several data constructors.
 %
-Then, is possible to obtain the average ``total'' number of data constructors
-that our generators will produce, regardless of if they are generated on their
-own, or as part of a pattern matching value.
+In that manner, it is possible to obtain the expected \emph{total} number of
+generated data constructors that our generators will produce---regardless if they
+are generated on their own, or as part of a pattern matching value.
 %
-We note this number as $E^\downarrow\![\_]$.
-
-The prediction only needs to add the expected number of constructors that we
-will additionally generate within pattern matching values to the prediction of
-bare data constructors.
+We note this number as $E^\downarrow\![\_]$ and, to calculate it, we only need
+to add the expected number of bare constructors that are included within each
+pattern matching.
 %
-For instance, we can calculate the total average number of constructors |Text|
-and |Join| that we will generate simply by expanding the average number of
+For instance, we can calculate the total expected number of constructors |Text|
+and |Join| that we will generate by simply expanding the expected number of
 generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
@@
 %
-Given that each time we generate a value satisfying the first pattern matching
-of the function |simplify| we add two |Text| and one |Join| data constructors to
-our random value.
+Observe that each time we generate a value satisfying the first pattern matching
+of the function |simplify|, we add two |Text| and one |Join| data constructors
+to our random value.
 %
 The case of the second pattern matching of |simplify| follows analogously.
-
-Note that this expansion cannot be applied if we also generate random values
-containing abstract interface function calls, as we cannot predict the output of
-an arbitrary function without solving the halting problem in the process.
+%
+Note that the overall prediction cannot be applied if we also generate random
+values containing function calls, as we cannot predict the output of an abstract
+function without solving the halting problem in the process.
 

### 2019-01-29 paper/relatedwork.tex
WIP @ relatedwork

@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+The literature on random generation of structured data is vast, both in
+practical and in theoretical terms.
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+%
+% Boltzmann models
+%
+Boltzmann models \tocite are a general approach to randomly generating
+combinatorial structures such as trees and graphs, and recently, closed
+simply-typed lambda terms \tocite.
+%
+A random generator built around such models uniformly generates values of a
+target size with a certain size tolerance.
+%
+However, it has been argued that this approach has theoretical and practical
+limitations in the context of software testing \tocite.
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+%
+% GodelTest
+%
+From the practical point of view, Feldt and Poulding propose G\"odelTest
+\tocite, a search-based framework for generating biased data.
+%
+G\"odelTest uses non-determinism to generate a wide range of data structures.
+%
+Similar to our approach, it works optimizing the parameters governing the
+desired biases on the generated data.
+%
+However, the optimization mechanism uses metaheuristic search for the best
+parameters at runtime.
+%
+\dragenp on the other hand implements a analytic and composable prediction
+mechanism that is used only used at compile time to optimize the generation
+parameters, thus avoiding to perform any kind of runtime reinforcement.
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+%
+% OO
+%
+In the context of object oriented programming, ...

### 2019-01-29 paper/relatedwork.tex
WIP

@@
 Boltzmann models \tocite are a general approach to randomly generating
-combinatorial structures such as trees and graphs, and recently, closed
-simply-typed lambda terms \tocite.
+combinatorial structures such as trees and graphs, closed simply-typed lambda
+terms, etc.
 %
@@
 limitations in the context of software testing \tocite.
+%
+In a recent work, AUTHOR provides a framework called \emph{boltzmann-brain} to
+specify and synthesize standalone Haskell random generators based on Boltzmann
+models.
+%
+This framework mixes parameter tunning and rejection of unwanted samples to
+approximate the desired distribution of values according to user demands.
+%
+In this approach, the overall discard ratio depends on how constrained is this
+desired distribution.
+%
+On the other hand, our work is focused on approximating the desired distribution
+as much as possible via parameter optimization, without discarding any generated
+value at runtime.
+%
+Altough promising, we found several difficulties to compare both approaches in
+practice due that boltzmann-brain is considered a standalone utility that
+produces self-contained samplers.
+%
+In this light, data specifications have to be manually written using a special
+syntax, and cannot include Haskell ground types like |String| or |Int|,
+difficulting the task of integrating this tool to existing Haskel codebases like
+the ones we evaluate on this work.
 
@@
 %
-% OO
+% DART
 %
-In the context of object oriented programming, ...
+Directed Automated Software Testing ...

### 2019-01-30 paper/relatedwork.tex
WIP @ relatedwork

@@
 %
-Boltzmann models \tocite are a general approach to randomly generating
+Boltzmann models \cite{Duchon2004} are a general approach to randomly generating
 combinatorial structures such as trees and graphs, closed simply-typed lambda
@@
 However, it has been argued that this approach has theoretical and practical
-limitations in the context of software testing \tocite.
+limitations in the context of software testing \cite{feldt2013}.
 %
-In a recent work, AUTHOR provides a framework called \emph{boltzmann-brain} to
-specify and synthesize standalone Haskell random generators based on Boltzmann
-models.
+In a recent work, Bendkowski provides a framework called \emph{boltzmann-brain}
+to specify and synthesize standalone Haskell random generators based on
+Boltzmann models \cite{bendkowski2018}.
 %
@@
 From the practical point of view, Feldt and Poulding propose G\"odelTest
-\tocite, a search-based framework for generating biased data.
+\cite{feldt2013}, a search-based framework for generating biased data.
 %
@@
 %
-Directed Automated Software Testing ...
+Directed Automated Software Testing (DART) is a technique that combines random
+testing with symbolic execution for C programs \cite{godefroid2005dart}.
+%
+I requires to instrument the target programs in order to introduce testing
+assertions and obtain feedback from previous testing executions that is used to
+explore new paths in the source code.
+%
+This technique has been shown to be remarkably useful, altough it forces a
+strong coupling between the testing suite and the target code.
+%
+Our tool intends to provide better random generation of values following an
+undirected fashion, without having to instrument the target code, but still
+extracting useful structural information from it.

### 2019-01-30 paper/watch.sh

@@ -1,4 +1,4 @@
 #!/bin/sh
 
 #https://facebook.github.io/watchman/docs/watchman-make.html
-watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' '*.tex' 'tikz/*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all
+watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' '[^.]*.tex' 'tikz/*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all

### 2019-01-30 paper/abstract.tex
General pass. Paper still needs compressing

@@
   % Status
-  Automatic generation of random values described by abstract data types (ADTs)
+  Automatic generation of random values described by algebraic data types (ADTs)
   is often a hard task.
@@
   % Situation
-  State-of-the-art random testing tools often can automatically synthesize
-  random data generators based on ADTs definitions.
+  State-of-the-art random testing tools can automatically synthesize random data
+  generators based on ADTs definitions.
   %
-  In that manner, generated values comply with the structure described by
-  ADTs---something that proves useful when testing software which expects
-  complex input like web pages, images, or even programs.
+  % In that manner, generated values comply with the structure described by
+  % ADTs---something that proves useful when testing software which expects
+  % complex input like web pages, images, or even programs.
+  %
+  In that manner, generated values comply with the structure described by ADTs,
+  something that proves useful when testing software which expects complex
+  inputs.
   %
@@
   % Solution
-  In this light, we propose to leverage static information found in the
-  codebase as a manner to improve the generation process.
+  % In this light, we propose to leverage static information found in the
+  In this work we propose to leverage static information found in the codebase
+  as a manner to improve the generation process.
   %
@@
   %
-  We implement a tool, called {\dragenp}, responsible to synthezise generators
-  for ADTs values while providing compile-time guarantees about their
-  distributions.
+  We implement a tool, responsible to synthezise generators for ADTs values
+  while providing compile-time guarantees about their distributions.
   %
-  Our compile-time predictions allow {\dragenp} to provide an heuristic that
-  tries to adjust the distribution of generators to what developers might want.
-  \todo[inline,author=Ale]{Case study}
-
+  Using compile-time predictions, we provide an heuristic that tries to adjust
+  the distribution of generators to what developers might want.
+  % \todo[inline,author=Ale]{Case study}
+  Preliminary experiments of our approach are very encouraging.
 \end{abstract}

### 2019-01-30 paper/casestudies.tex
General pass. Paper still needs compressing

@@
-\section{Case Studies}
+\section{Case Studies} \label{sec:casestudies}
 
@@
 Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
-triggered for different sizes of the corpus size and derivation tool, including
-error bars indicating the standard error of the mean on each case.
+triggered for different combinations of corpus size and derivation tool,
+including error bars indicating the standard error of the mean on each case.
 %
@@
 While it seems a little improvement, it is worth noticing that we are generating
-highly structured input, i.e., Lisp programs.
+highly structured inputs, i.e., Lisp programs.
 %

### 2019-01-30 paper/introduction.tex
General pass. Paper still needs compressing

@@
 %
-It requires developers to specify (i) \emph{testing properties} which describe
-programs behavior and (ii) \emph{random data generator} based on the
-{\emph{types}} of the expected inputs (e.g., an integer, an string,
+It requires developers to specify (i) \emph{testing properties} describing
+programs expected behavior and (ii) \emph{random data generators} based on the
+\emph{types} of the expected inputs (e.g., an integer, an string,
 etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
 %
-QuickCheck then generates random test cases and reports those violating testing
+\quickcheck then generates random test cases and reports those violating testing
 properties.
-%
 
 % Situation with Quick Check, ADTs, tools
-QuickCheck comes equipped with random generators for built-in types, while it
+\quickcheck comes equipped with random generators for built-in types, while it
 requires to manually write generators for user-defined ADTs.
@@
 Recently, there have been a proliferation of tools to automatically derive
-QuickCheck generators for ADTs
+\quickcheck generators for ADTs
 \cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
@@
 definition of the ADT.
-%
 
@@
 %
-Generating ADT values also proves fruitful when looking for vulnerabilities with
-fuzzers \cite{GriecoCB16,grieco2017}.
+Generating ADT values also proves fruitful when looking for vulnerabilities in
+combination with fuzzers \cite{GriecoCB16,grieco2017}.
 %
@@
 %% Our proposal
-In this work, we propose a novel improvement in the generation process of ADT
-values by exploiting some static information found in the codebase.
+In this work, we identify two different sources of structural information that
+can be statically exploited to improve the generation process of ADT values
+(Section \ref{sec:sources}).
 %
-More specifically, to refine the structure of generated values, we propose a
-generation process that is capable to consider how programs branch based on ADTs
-values' patterns as well as how they get manipulated by abstract interfaces.
+% In this work, we propose a novel improvement in the generation process of ADT
+% values by exploiting some static information found in the codebase.
+%
+Then, we provide a compositional approach to capture this information into our
+automatically derived random generators.
+%
+More specifically, we propose a generation process that is capable to consider
+how programs branch based on ADTs values' patterns as well as how they get
+manipulated by abstract interfaces (Section \ref{sec:hrepcont}).
 %
@@
   processes}\cite{gw1875}---a simple stochastic model conceived to study
-population growth.
+population growth (Section \ref{sec:synthesis}).
+%
 %---to predict the distribution of QuickCheck
@@
 %
-We implement our ideas in a tool, called {\dragenp}, that is capable to
-automatically synthesize QuickCheck generators for \emph{rich ADT values}, where
-the distributions of random values can be adjusted at compile-time to what
+We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
+  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
+automatically synthesize \quickcheck generators for \emph{rich ADT values},
+where the distributions of random values can be adjusted at compile-time to what
 developers might want based on our predictions.
 %
-\todo[inline,author=Ale]{Add here later about test cases when we know what they
-  are}
+Finally, we provide empirical evaluations showing that including static
+information from the user codebase improves the code coverage of two external
+applications when tested using random values generated following our ideas
+(Section \ref{sec:casestudies}).
+
+% \todo[inline,author=Ale]{Add here later about test cases when we know what they
+%   are}
 
@@
 technique is general enough to be applied to most programming languages.
+%
 %with some level of support for composite types.

### 2019-01-30 paper/randomtesting.tex
General pass. Paper still needs compressing

@@
 found at the type-level.
+%
 %type-driven approach, along
@@
 % %
+%
 To exemplify this, and for illustrative purposes, let us consider the following
@@
 %
-With this representation, the page \texttt{<html>hello<hr>bye</html>} can be
-encoded as:
+With this representation, the example page \texttt{<html>hello<hr>bye</html>}
+can be encoded as:
 %
@@
 %
-This is a cumbersome and error prone task and usually \emph{follows closely the ADT
-structure}.
+This is a cumbersome and error prone task that usually \emph{follows closely the
+  ADT structure}.
 %
@@
 genHtml = sized (\size ->
-   if size == 0  then  frequency
+  if size == 0
+  then frequency
   [  (2,  Text    <$> genString)
@@
 to paremetrize the generation process up to an external natural number known as
-the \emph{generation size}---capture in the code with variable |size|.
+the \emph{generation size}---captured in the code with variable |size|.
 %
@@
 On the other hand, when it picks to generate a |Join| constructor, it also
-generates two independent random subexpression recursively, thus decreasing the
+generates two independent random sub-expression recursively, decreasing the
 generation size by a unit on each recursive invocation (|smaller genHtml|).
@@
 generation process is limited to pick \emph{a single data constructor at the
-  time and then recursively generate each required subexpression independently}.
+  time and then recursively generate each required sub-expression
+  independently}.
 %

### 2019-01-30 paper/relatedwork.tex
General pass. Paper still needs compressing

@@
 Altough promising, we found several difficulties to compare both approaches in
-practice due that boltzmann-brain is considered a standalone utility that
+practice due that \emph{boltzmann-brain} is considered a standalone utility that
 produces self-contained samplers.
@@
 syntax, and cannot include Haskell ground types like |String| or |Int|,
-difficulting the task of integrating this tool to existing Haskel codebases like
-the ones we evaluate on this work.
+difficulting the integration of this tool to existing Haskell codebases like the
+ones we consider in this work.
 
@@
 %
-From the practical point of view, Feldt and Poulding propose G\"odelTest
+From the practical point of view, Feldt and Poulding propose \emph{G\"odelTest}
 \cite{feldt2013}, a search-based framework for generating biased data.
 %
-G\"odelTest uses non-determinism to generate a wide range of data structures.
+% \emph{G\"odelTest} uses non-determinism to generate a wide range of data structures.
 %
-Similar to our approach, it works optimizing the parameters governing the
-desired biases on the generated data.
+Similar to our approach, \emph{G\"odelTest} works by optimizing the parameters
+governing the desired biases on the generated data.
 %
-However, the optimization mechanism uses metaheuristic search for the best
+However, the optimization mechanism uses metaheuristic search to find the best
 parameters at runtime.
@@
 %
-I requires to instrument the target programs in order to introduce testing
-assertions and obtain feedback from previous testing executions that is used to
-explore new paths in the source code.
+It requires to instrument the target programs in order to introduce testing
+assertions and obtain feedback from previous testing executions, which is used
+to explore new paths in the source code.
 %

### 2019-01-30 paper/synthesis.tex
General pass. Paper still needs compressing

@@
 It is our realization that it suffices to simply consider \emph{each different
-  pattern matching and functional call as a kind of individual on its own}.
+  pattern matching and function call as a kind of individual on its own}.
 %
@@
 for each different pattern matching and function call as follows:
-
+%
 \newcommand{\ph}{\phantom{\square}}
@@
       \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
-      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
+      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
       \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
@@
     \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
-    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
-    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
+    % \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
+    % \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
     \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
@@
 The red-light colored matrix is what we had before where the light-blue colored
-cells are new---we recommend readers to get a colored copy of the paper.
+cells are new---we encourage readers to obtain a colored copy of this work.
 %
@@
 %
-The rest follows analogously.
+The rest of this matrix can be computed analogously.
 
@@
 represents the number of recursive sub-terms of each construction that we
-generate, whereas $\mathcal{P}$ simply represents the probability of generating
-that partial ADT value.%  (this probability can be easily calculated based on the
+generate (i.e., data constructors, pattern matching values or function calls),
+whereas $\mathcal{P}$ simply represents the probability of generating that
+partial ADT value.% (this probability can be easily calculated based on the
 % generation frequencies used in the random generator).
@@
 values are as follows (where dashed lines indicate the composition of the
-diffent sub-vectors that can be calculated separately):
+diffent sub-vectors that are calculated separately):
 
@@
 In this manner, developers can specify, for instance, a uniform distribution of
-data constructors, pattern matching values and function calls or, alternative, a
-distribution of values with some constructions appearing in a different
+data constructors, pattern matching values and function calls or, alternatively,
+a distribution of values with some constructions appearing in a different
 proportion as others, e.g., two times more functions calls to |div| than |Join|

### 2019-01-30 paper/synthesis.tex
Added syntax highlighting (config is in notation.fmt)

@@
 Each element of the matrix can be automatically calculated by exploiting ADT
-definitions \cite{DBLP:conf/haskell/MistaRH18}---specially the recursive
-fields---as well as the individual probability of generating each constructor.
+definitions \cite{DBLP:conf/haskell/MistaRH18}
+%
+% ---specially the recursive fields---
+%
+as well as the individual probability of generating each constructor.
 %
@@
 %
-It is our realization that it suffices to simply consider \emph{each different
-  pattern matching and function call as a kind of individual on its own}.
+Our realization is that it suffices to consider \emph{each different pattern
+  matching and function call as a kind of individual on its own}.
 %
-In that manner, we can extend our mean matrix $M$ by adding a row and a column
-for each different pattern matching and function call as follows:
+In that manner, we can extend our mean matrix $M$ adding a row and a column for
+each different pattern matching and function call as shown in Fig.
+\ref{fig:meanmatrix}.
 %
+\begin{wrapfigure}{r}{0.26\textwidth}
+  \vspace{-8pt}
   \newcommand{\ph}{\phantom{\square}}
   \newcommand{\el}{\square}
-\begin{equation*}
-  \hspace{20pt}
-  M = \ \
+  \centering
+  \resizebox{4.5cm}{!}{
     \begin{tikzpicture}
-    [ baseline=-0.65ex
+      [ baseline=-0.5ex
+      % , transform canvas={scale=0.9}
       , every left delimiter/.style={xshift=.5em}
@@
       % Color squares
-    \fill[red, opacity=0.075]
+      \fill[red, opacity=0.15]
       (m-1-1.north west)
-    -|| ([xshift=-1pt]m-1-3.north east)
-    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
-    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
+      -|| (m-1-3.north east)
+      -|| (m-3-3.south east)
+      -|| ([xshift=1pt]m-3-1.south west)
       -|| cycle;
-    \fill[blue, opacity=0.075]
+      \fill[blue, opacity=0.15]
       ([yshift=-1pt]m-4-1.north west)
-    -|| ([xshift=1pt]m-1-3.north east)
-    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
-    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
+      -|| (m-1-3.north east)
+      -|| ([yshift=-1pt]m-1-9.south east)
+      -|| (m-9-9.south east)
       -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
@@
       % Upper labels
-    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
-    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
-    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
-    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
-    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
-    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
-    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
-    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
-    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
+      \node[above,text depth=1pt] at (m-1-1.north) {$ C_1$};
+      \node[above,text depth=1pt] at (m-1-2.north) {$ \cdots$};
+      \node[above,text depth=1pt] at (m-1-3.north) {$ C_i$};
+      \node[above,text depth=1pt] at (m-1-4.north) {$ P_1$};
+      \node[above,text depth=1pt] at (m-1-5.north) {$ \cdots$};
+      \node[above,text depth=1pt] at (m-1-6.north) {$ P_j$};
+      \node[above,text depth=1pt] at (m-1-7.north) {$ F_1$};
+      \node[above,text depth=1pt] at (m-1-8.north) {$ \cdots$};
+      \node[above,text depth=1pt] at (m-1-9.north) {$ F_k$};
       % Left labels
-    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
-    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
-    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
-    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
-    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
-    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
-    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
-    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
-    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
+      \node[left,overlay] at (m-1-1.west) {$ C_1$};
+      \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$ \vdots$};
+      \node[left,overlay] at (m-3-1.west) {$ C_i$};
+      \node[left,overlay] at (m-4-1.west) {$ P_1$};
+      \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$ \vdots$};
+      \node[left,overlay] at (m-6-1.west) {$ P_j$};
+      \node[left,overlay] at (m-7-1.west) {$ F_1$};
+      \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$ \vdots$};
+      \node[left,overlay] at (m-9-1.west) {$ F_k$};
       % Dashed lines
-    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
-    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
-    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
-    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
+      \draw[densely dashed, semithick] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
+      \draw[densely dashed, semithick] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
+      \draw[densely dashed, semithick] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
+      \draw[densely dashed, semithick] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
       % \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
       % \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
-    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
-    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
+      \draw[dotted, thick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
+      \draw[dotted, thick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
     \end{tikzpicture}
-\end{equation*}
-
-Symbol $C_1 \cdots C_i$ denotes constructors, $P_1 \cdots P_j$ pattern
-matchings, and $F_1 \cdots F_k$ function calls.
+}
+  \caption{Mean matrix $M$ including pattern matching and function calls
+    information.}
+  \label{fig:meanmatrix}
+  \vspace{-6pt}
+\end{wrapfigure}
+%
+Symbol $C_1\! \cdots C_i$ denotes constructors, $P_1\! \cdots P_j$ pattern
+matchings, and $F_1\! \cdots F_k$ function calls.
 %
@@
 For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
-values are as follows (where dashed lines indicate the composition of the
-diffent sub-vectors that are calculated separately):
+values are as shown in Fig. \ref{fig:vectors} (where dashed lines indicate the
+composition of the diffent sub-vectors that are calculated separately).
 
+\begin{figure}[t]
+  \resizebox{!}{2cm}{
 \begin{equation*}
@@
   \end{tikzpicture}
-  \hspace{20pt}
+  \hspace{30pt}
   \mathcal{P} =
@@
 \end{equation*}
+}
+\vspace{-5pt}
+\caption{Prediction vectors of our |Html| generation specification.}
+\label{fig:vectors}
+\vspace{-10pt}
+\end{figure}
 

### 2019-01-31 paper/introduction.tex
Pass over intro

@@
 % Scenario of random testing
-Random testing is a promising approach to finding bugs
+Random testing is a promising approach to finding\todoale{find?} bugs
 \cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
@@
 %
-\quickcheck then generates random test cases and reports those violating testing
+\quickcheck then generates random test cases and reports violating testing
 properties.
@@
 %
-In other words, generated values follow the structure described by the
-definition of the ADT.
+In other words, generated values follow the structure described by ADT
+definitions.
 
@@
 Well-typed ADT values are specially useful when testing programs which expect
-highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
+highly structured inputs like compilers \cite{Palka11,MidtgaardJKNN17}.
 %
@@
 As a result, even if random values are well-typed, they might not be built with
-enough structure to penetrate into deep layers of software.
+enough structure to penetrate into deep software layers.
 
@@
 %
-Then, we provide a compositional approach to capture this information into our
-automatically derived random generators.
+Then, we show how to capture this information into our (automatically) derived
+random generators.
 %
 More specifically, we propose a generation process that is capable to consider
-how programs branch based on ADTs values' patterns as well as how they get
-manipulated by abstract interfaces (Section \ref{sec:hrepcont}).
+how programs branch on input ADTs values as well as how they get manipulated by
+abstract interfaces (Section \ref{sec:hrepcont}).
 %
@@
   at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
-automatically synthesize \quickcheck generators for \emph{rich ADT values},
+automatically synthesize \quickcheck generators capable to produce \emph{rich ADT values},
 where the distributions of random values can be adjusted at compile-time to what
-developers might want based on our predictions.
+developers might want. %based on our predictions.
 %

### 2019-01-31 paper/randomtesting.tex
Pass over section 2

@@
 In this section, we briefly introduce the common approach for automatically
-deriving random data generators for ADTs in QuickCheck by using the information
-found at the type-level.
+deriving random data generators for ADTs in QuickCheck.
+%by using the information
+%found at the type-level.
 %
@@
 %
-For instance, |Text| has a field of type |String|, whereas |Join| two recusive
-fields of type |Html|.
+For instance, |Text| has a field of type |String|, whereas |Join| has two
+recursive fields of type |Html|.
 %
@@
 
-In order to generate random value of types involving ADTs, most approaches
-require users to provide a random data generator for each one of them.
+In order to generate random ADTs values, most approaches require users to provide
+a random data generator for each ADT definition.
 %
-This is a cumbersome and error prone task that usually \emph{follows closely the
-  ADT structure}.
+This is a cumbersome and error prone task that usually \emph{follows closely
+  ADTs structures}.
 %
@@
 We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
-elements, respecitvely, e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.
+elements, respecitvely (e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.).
 %
@@
 When called with a positive generation size, this generator can pick to generate
-among any |Html| data constructor of with an explicitly given \emph{generation
-  frequency} that can be chosen by the user---in this example, 2, 1, 4 and 3 for
-|Text|, |Single|, |Tag|, and |Join|, respectively.
+among any |Html| data constructor with an explicitly \emph{generation frequency}
+that can be chosen by the user---in this example, 2, 1, 4 and 3 for |Text|,
+|Single|, |Tag|, and |Join|, respectively.
 %
@@
 This random process keeps calling itself recursively until the generation size
-reaches zero, where the generator is constrained to pick only among terminal
-data constructors, being |Text| and |Single| the only possible choices in our
+reaches zero, where the generator is constrained to pick among terminal data
+constructors, being |Text| and |Single| the only possible choices in our
 particular case.

### 2019-01-31 paper/synthesis.tex
Pass over predictions

@@
 can be adapted to predict the generation of ADT values by simply considering
-each data constructor as an individual on its own.
+each data constructor as a kind of its own.
 %
@@
 Each element $M_{i,j}$ of this matrix encodes the average number of data
-constructors of kind $j$ that gets generated in a given level, provided that we
-generated a constructor of kind $i$ at the previous level.
+constructors of kind $j$ that gets generated in a given generation, provided that we
+generated a constructor of kind $i$ at the previous one.
 %
@@
 Each element of the matrix can be automatically calculated by exploiting ADT
-definitions \cite{DBLP:conf/haskell/MistaRH18}
+definitions. % \cite{DBLP:conf/haskell/MistaRH18}
 %
@@
 %
-For instance, if we consider $P_j$ as the second pattern of simplify and $F_1$
-as function |div|, then the marked cell above has the value $2\cdot p_{div}$,
-i.e., the amount of holes in a partially generated pattern |(Join (Join (Text s)
-hole) hole)|, where |s| is some random string, times the probability to generate
-a call to function |div|.
+For instance, if we consider $P_j$ as the second pattern of function |simplify|
+and $F_1$ as function |div|, then the marked cell above has the value
+$2\cdot p_{div}$, i.e., the amount of holes in the partially generated pattern
+|(Join (Join (Text s) hole) hole)|, where |s| is some random string, times the
+probability to generate a call to function |div|.
 %
@@
 factored in terms of two vectors $\beta$ and $\mathcal{P}$, such that $\beta$
-represents the number of recursive sub-terms of each construction that we
-generate (i.e., data constructors, pattern matching values or function calls),
+represents the number of holes in each partial ADT value that we generate,
 whereas $\mathcal{P}$ simply represents the probability of generating that
@@
 %
-The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
-source of structural information that we consider, and joining them together.
+% The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
+% source of structural information that we consider, and joining them together.
 %
 For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
-values are as shown in Fig. \ref{fig:vectors} (where dashed lines indicate the
-composition of the diffent sub-vectors that are calculated separately).
+values are as shown in Fig. \ref{fig:vectors}.
+%
+We note |simplify#1| and |simplify#2| to the patterns occurring in the first and
+second clauses of |simplify|, respectively.
+%(Dashed lines indicate the composition of the diffent sub-vectors
+%that are calculated separately).
 
@@
 Note that the overall prediction cannot be applied if we also generate random
-values containing function calls, as we cannot predict the output of an abstract
-function without solving the halting problem in the process.
+values containing function calls, as we cannot predict the output of an
+arbitrary function.
+%without solving the halting problem in the process.
 

### 2019-01-31 paper/abstract.tex
Pass over abstract
Minor fixes

@@
   %
-  We implement a tool, responsible to synthezise generators for ADTs values
+  We implement a tool, responsible for synthesizing generators for ADTs values
   while providing compile-time guarantees about their distributions.
   %
-  Using compile-time predictions, we provide an heuristic that tries to adjust
+  Using compile-time predictions, we provide a heuristic that tries to adjust
   the distribution of generators to what developers might want.
   % \todo[inline,author=Ale]{Case study}
-  Preliminary experiments of our approach are very encouraging.
+  We report on preliminary experiments where our approach shows encouraging
+  results.
 \end{abstract}

### 2019-01-31 paper/introduction.tex
Minor fixes

@@
 % Scenario of random testing
-Random testing is a promising approach to finding\todoale{find?} bugs
+Random testing is a promising approach for finding bugs
 \cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
@@
 %
-Recently, there have been a proliferation of tools to automatically derive
+Recently, there has been a proliferation of tools to automatically derive
 \quickcheck generators for ADTs
@@
 %
-More specifically, we propose a generation process that is capable to consider
-how programs branch on input ADTs values as well as how they get manipulated by
-abstract interfaces (Section \ref{sec:hrepcont}).
+More specifically, we propose a generation process that is capable of
+considering how programs branch on input ADTs values as well as how they get
+manipulated by abstract interfaces (Section \ref{sec:hrepcont}).
 %
-Furthermore, we show how to predict the distribution of the \emph{expected}
-numbers of ADT constructors, values fitting certain branching patterns, and
-calls to interfaces.
+Furthermore, we show how to predict the \emph{expected} distribution of the ADT
+constructors, values fitting certain branching patterns, and calls to interfaces
+that our random generators produce.
 %
@@
 We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
-  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
-automatically synthesize \quickcheck generators capable to produce \emph{rich ADT values},
-where the distributions of random values can be adjusted at compile-time to what
-developers might want. %based on our predictions.
+  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable of
+automatically synthesizing \quickcheck generators which produce \emph{rich ADT
+  values}, where the distributions of random values can be adjusted at
+compile-time to what developers might want. %based on our predictions.
 %

### 2019-01-31 paper/synthesis.tex
Minor fixes

--- 2019-01-30 paper/synthesis.tex
@@
 %
-It requires to model every random choice that a generator could possibly make to
+It requires modeling every random choice that a generator could possibly make to
 generate a value.
@@
 In a recent work \cite{DBLP:conf/haskell/MistaRH18}, we have shown that it is
-possible to \emph{analitically} predict the average distribution of data
-constructors produced by a random generators automatically derived considering
+possible to \emph{analytically} predict the average distribution of data
+constructors produced by random generators automatically derived considering
 only ADT definitions---like the one presented on Section
@@
 This probabilistic theory was originally conceived to predict the growth and
-extinction of royal family trees the Victorian Era, being later applied to a
+extinction of royal family trees the Victorian Era, later being applied to a
 wide variety of research areas.
 %
-In this work, we adapt this model to also predict the average distribution of
-values of random generators derived considering structural information coming
-from functions pattern matchings and abstract interfaces.
+In this work, we adapt this model to predict the average distribution of values
+of random generators derived considering structural information coming from
+functions' pattern matchings and abstract interfaces.
 
@@
 %
-The red-light colored matrix is what we had before where the light-blue colored
-cells are new---we encourage readers to obtain a colored copy of this work.
+The light-red colored matrix is what we had before, whereas the light-blue
+colored cells are new---we encourage readers to obtain a colored copy of this
+work.
 %
@@
 \begin{align*}
-  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
+  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P}^T)^{n}
 \end{align*}
@@
 To achieve that, we should stop considering pattern matching values as atomic
-values and start seeing them as compositions of several data constructors.
+constructions and start seeing them as compositions of several data
+constructors.
 %
@@
 Note that the overall prediction cannot be applied if we also generate random
-values containing function calls, as we cannot predict the output of an abstract
-function without solving the halting problem in the process.
+values containing function calls, as we cannot predict the output of an
+arbitrary function.
 

### 2019-01-31 paper/casestudies.tex

@@
 \begin{figure*}[t]
+  \vspace{-5pt}%
   \centering
@@
   \label{fig:coverage}
+  \vspace{-10pt}%
 \end{figure*}
@@
 %
-While it seems a little improvement, it is worth noticing that we are generating
-highly structured inputs, i.e., Lisp programs.
+While it seems an small improvement, we argue that an improvement of $4\%$ is
+not negligible considering (a) the little effort that took us to specify the pattern
+matchings and (b) that we are testing a compiler.
+%it is worth noticing that we are generating
+%highly structured inputs, i.e., Lisp programs.
 %
-When testing interpreters, we argue that an improvement of $4\%$ is not
-negligible considering the little effort that took us to specify the pattern
-matchings.
+%When testing interpreters,
 %\todo[inline, author=AM]{Shall we say anything justifing this little
@@
 
-In this case study we included the structure information present on the abstract
+In this case study we included the structural information present on the abstract
 interface of this library into the generation specification of random HTML
 values, resulting in a generation process that randomly picked among 4 data
-constructors and 163 abstract combinators.
+constructors and 163 abstract functions.
 %
@@
 %
-We consider that our approach is particularly useful when the data types
-encoding the shape of our data are vague or not sufficiently structured to be
-used to derive powerful random generators with the common derivation techniques
-based only on ADT definitions.
+%% Space reasons
+%We consider that our approach is particularly useful when the data types
+%encoding the shape of our data are vague or not sufficiently structured to be
+%used to derive powerful random generators with the common derivation techniques
+%based only on ADT definitions.

### 2019-01-31 paper/synthesis.tex

--- 2019-01-31 paper/synthesis.tex
@@
 arbitrary function.
-%without solving the halting problem in the process.
 

### 2019-01-31 paper/relatedwork.tex
Pass on related work
Minor fixes

@@
 
-The literature on random generation of structured data is vast, both in
-practical and in theoretical terms.
+%The literature on random generation of structured data is vast, both in
+%practical and in theoretical terms.
 
@@
 %
-This framework mixes parameter tunning and rejection of unwanted samples to
+This framework mixes parameter tuning and rejection of unwanted samples to
 approximate the desired distribution of values according to user demands.
 %
-In this approach, the overall discard ratio depends on how constrained is this
-desired distribution.
+In this approach, the overall discard ratio depends on how constrained the
+desired distribution is.
 %
@@
 %
-\dragenp on the other hand implements a analytic and composable prediction
+\dragenp on the other hand implements an analytic and composable prediction
 mechanism that is used only used at compile time to optimize the generation
-parameters, thus avoiding to perform any kind of runtime reinforcement.
+parameters, thus avoiding perfomring any kind of runtime reinforcement.
 
@@
 %
-Directed Automated Software Testing (DART) is a technique that combines random
+Directed Automated Random Testing (DART) is a technique that combines random
 testing with symbolic execution for C programs \cite{godefroid2005dart}.
 %
-It requires to instrument the target programs in order to introduce testing
+It requires instrumenting the target programs in order to introduce testing
 assertions and obtain feedback from previous testing executions, which is used

### 2019-01-31 paper/finalremarks.tex
Compression (2 citations away from the goal)
Added final remarks, tweaks on highlighting

--- 2019-01-20 paper/sections.old/finalremarks.tex
@@
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem.
+We extended the standard approach for automatically deriving random generators
+in Haskell.
+%
+Our generators are capable of producing complex and interesting random values by
+exploiting static structural information found in the user codebase.
+%
+Based on the theory of branching processes, we adapt our previous prediction
+mechanism to characterize the distribution of random values representing the
+different sources of structural information that our generators might produce.
+%
+This predictions let us optimize the generation parameters in compile time,
+resulting in an improved testing performance according to our experiments.
+
+In the future we aim to extract stronger invariants from the user codebase into
+our automatically derived generators, as well as evaluating how introducing
+random mutations into structurally complex values might improve testing
+performace, specially when generating ``almost correct'' random data.

### 2019-01-31 paper/randomtesting.tex
Minor fixes
Added final remarks, tweaks on highlighting

@@
 We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
-elements, respecitvely (e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.).
+elements, respecitvely (e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.)
 %
@@
 %
-It uses the theory of branching processes \cite{gw1875} to modelize and predict
+It uses the theory of branching processes \cite{gw1875} to model and predict
 analytically the expected number of generated data constructors.

### 2019-01-31 paper/synthesis.tex
Added final remarks, tweaks on highlighting

@@
 %
-$$ M_{Join, Text} = 2\cdot p_{Text} $$
+$$ M_{\mathrm{Join, Text}} = 2\cdot p_{\mathrm{Text}} $$
 %
 where $2$ is the number of holes present when generating a partial ADT value
-|Join| (i.e., |Join hole hole|) and $p_{Text}$ is the probability of
+|Join| (i.e., |Join hole hole|) and $p_{\mathrm{Text}}$ is the probability of
 individually generating the constructor |Text|.
@@
 %
-We note |simplify#1| and |simplify#2| to the patterns occurring in the first and
+We note |simplify1| and |simplify2| to the patterns occurring in the first and
 second clauses of |simplify|, respectively.
@@
     \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
-    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
-    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
-    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
-    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
-    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
-    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
-    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
-    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
-    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
+    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle \mathrm{Text}$};
+    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle \mathrm{Single}$};
+    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle \mathrm{Tag}$};
+    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle \mathrm{Join}$};
+    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle \mathrm{simplify}\#1$};
+    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle \mathrm{simplify}\#2$};
+    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle \mathrm{hr}$};
+    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle \mathrm{div}$};
+    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle \mathrm{bold}$};
   \end{tikzpicture}
@@
     {
-      p_{Text} \\
-      p_{Single} \\
-      p_{Tag}  \\
-      p_{Join} \\
-      p_{simplify\#1} \\
-      p_{simplify\#2} \\
-      p_{hr} \\
-      p_{div} \\
-      p_{bold} \\
+      p_{\mathrm{Text}} \\
+      p_{\mathrm{Single}} \\
+      p_{\mathrm{Tag}}  \\
+      p_{\mathrm{Join}} \\
+      p_{\mathrm{simplify\#1}} \\
+      p_{\mathrm{simplify\#2}} \\
+      p_{\mathrm{hr}} \\
+      p_{\mathrm{div}} \\
+      p_{\mathrm{bold}} \\
     };
@@
 {\small
-  \begin{align*}
-    E^\downarrow\!\!\left[ Text \right]
-    = E\left[ Text \right]
-    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
-    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
-    E^\downarrow\!\!\left[ Join \right]
-    = E\left[ Join \right]
-    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
-    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
-  \end{align*}
+  \begin{alignat*}{4}
+    &E^\downarrow\!\left[ |Text| \right]
+    &= E\left[ |Text| \right]
+      &+ 2 \cdot E\!\left[ |simplify1| \right]\
+      &+\ 1 \cdot E\!\left[ |simplify2| \right] \\
+    &E^\downarrow\!\left[ |Join| \right]
+    &= E\left[ |Join| \right]
+      &+ 1 \cdot E\!\left[ |simplify1| \right]\
+      &+\ 2 \cdot E\!\left[ |simplify2| \right]
+  \end{alignat*}
 }

### 2019-02-01 paper/casestudies.tex
Spell checking

@@
 \emph{hs-zuramaru}\footnote{\href{http://hackage.haskell.org/package/zuramaru}{http://hackage.haskell.org/package/zuramaru}},
-implementing an embedded Lisp intepreter for a small subset of this programming
+implementing an embedded Lisp interpreter for a small subset of this programming
 language, and
@@
 syntactically correct HTML values, consisting to a large extent of random
-strings within and between HTML tag delimeters (|"<"|, |">"| and |"/>"|).
+strings within and between HTML tag delimiters (|"<"|, |">"| and |"/>"|).
 %
@@
 We found that, in many cases, the test cases generated with \dragenp were
-parsed, analysed and reported as valid HTML values by the target application.
+parsed, analyzed and reported as valid HTML values by the target application.
 

### 2019-02-01 paper/finalremarks.tex
Spell checking

@@
 random mutations into structurally complex values might improve testing
-performace, specially when generating ``almost correct'' random data.
+performance, specially when generating ``almost correct'' random data.

### 2019-02-01 paper/randomtesting.tex
Spell checking

@@
 |Text|---which represents plain text values---, |Single| and |Tag|---which
-represent singular and paired Html tags, respectively---, and |Join|---which
-concats two Htmls pages one after another.
+represent singular and paired HTML tags, respectively---, and |Join|---which
+concatenates two HTMLs pages one after another.
 %
@@
 In Haskell, |Text|, |Single|, |Tag|, and |Join| are known as \emph{data
-  constructors} (or constructors for short) and are used to distiguish which
+  constructors} (or constructors for short) and are used to distinguish which
 variant of the ADT we are constructing.
@@
 %
-This is a cumbersome and error prone task that usually \emph{follows closely
-  ADTs structures}.
+This is a cumbersome and error prone task that usually \emph{follows closely the
+  structure of the ADTs}.
 %
@@
 We use the Haskell syntax |[]| and |(,)| for denoting lists and pairs of
-elements, respecitvely (e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.)
+elements, respectively (e.g., |[(1,2),(3,4)]| is a list of pairs of numbers.)
 %
 The random generator |genHtml| is defined using \quickcheck's function |sized|
-to paremetrize the generation process up to an external natural number known as
+to parameterize the generation process up to an external natural number known as
 the \emph{generation size}---captured in the code with variable |size|.

### 2019-02-01 paper/relatedwork.tex
Spell checking

@@
 %
-Altough promising, we found several difficulties to compare both approaches in
+Although promising, we found several difficulties to compare both approaches in
 practice due that \emph{boltzmann-brain} is considered a standalone utility that
@@
 %
-However, the optimization mechanism uses metaheuristic search to find the best
+However, the optimization mechanism uses meta-heuristic search to find the best
 parameters at runtime.
@@
 mechanism that is used only used at compile time to optimize the generation
-parameters, thus avoiding perfomring any kind of runtime reinforcement.
+parameters, thus avoiding performing any kind of runtime reinforcement.
 
@@
 %
-This technique has been shown to be remarkably useful, altough it forces a
+This technique has been shown to be remarkably useful, Although it forces a
 strong coupling between the testing suite and the target code.

### 2019-02-01 paper/synthesis.tex
Spell checking

@@
 We characterize the numbers of constructors that a random generator produces in
-the $n$-th generaton as a vector $G_n$, a vector that groups the number of
+the $n$-th generation as a vector $G_n$, a vector that groups the number of
 constructors of each kind produced in that generation---in our |Html| example,
@@
 
-Vector $E[G_0]$ represents the initial distribuition of constructors that our
+Vector $E[G_0]$ represents the initial distribution of constructors that our
 generator produces, which simply consists of the generation probability of each
@@
 %
-Then, the equation \ref{eqn:prediction} can be rewriten as:
+Then, the equation \ref{eqn:prediction} can be rewritten as:
 %
@@
 Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
-distribution of our random genrator in a way that can be always characterized
+distribution of our random generator in a way that can be always characterized
 and predicted.
@@
 %
-This is done by runing a simulation-based optimization process at compile-time.
+This is done by running a simulation-based optimization process at compile-time.
 %
-This process is parametrized by the desired distribution of values set by the
+This process is parameterized by the desired distribution of values set by the
 user.

### 2019-02-01 paper/casestudies.tex
Minor fixes

@@
 For that, we considered two external programs which expect highly structured
-inputs, namely \emph{GNU CLISP}---the GNU Common Lisp
-compiler\footnote{\href{https://www.gnu.org/software/gcl/}{https://www.gnu.org/software/gcl/}},
-and \emph{HTML Tidy}
+inputs, namely \emph{GNU
+  CLISP} \footnote{\href{https://www.gnu.org/software/gcl/}{https://www.gnu.org/software/gcl/}}---the
+GNU Common Lisp compiler, and \emph{HTML Tidy}
 \footnote{\href{http://www.html-tidy.org}{http://www.html-tidy.org}}---a well
@@
 This way, any noticeable difference in the code coverage can be attributed to
-the present (or lack thereof) structural information when generating the test
+the presence (or lack thereof) structural information when generating the test
 cases.
@@
 %
-Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
-triggered for different combinations of corpus size and derivation tool,
+Figure \ref{fig:coverage} illustrates the mean number of different execution
+paths triggered for different combinations of corpus size and derivation tool,
 including error bars indicating the standard error of the mean on each case.
@@
 While it seems an small improvement, we argue that an improvement of $4\%$ is
-not negligible considering (a) the little effort that took us to specify the pattern
-matchings and (b) that we are testing a compiler.
+not negligible considering (a) the little effort that took us to specify the
+pattern matchings and (b) that we are testing a full-fledged compiler.
 %it is worth noticing that we are generating
@@
 With these results we are confident that including the structural information
-present on the user codebase improves the testing performance.
+present on the user codebase improves the overall testing performance.
 %

### 2019-02-01 paper/introduction.tex
Minor fixes

@@
 It requires developers to specify (i) \emph{testing properties} describing
-programs expected behavior and (ii) \emph{random data generators} based on the
+programs' expected behavior and (ii) \emph{random data generators} based on the
 \emph{types} of the expected inputs (e.g., an integer, an string,

### 2019-02-01 paper/randomtesting.tex
Minor fixes

@@
 represent singular and paired HTML tags, respectively---, and |Join|---which
-concatenates two HTMLs pages one after another.
+concatenates two HTML pages one after another.
 %

### 2019-02-01 paper/relatedwork.tex
Minor fixes

@@
 \dragenp on the other hand implements an analytic and composable prediction
-mechanism that is used only used at compile time to optimize the generation
+mechanism that is only used at compile time to optimize the generation
 parameters, thus avoiding performing any kind of runtime reinforcement.
@@
 %
-This technique has been shown to be remarkably useful, Although it forces a
+This technique has been shown to be remarkably useful, although it forces a
 strong coupling between the testing suite and the target code.

### 2019-02-01 paper/synthesis.tex
Minor fixes

@@
 Each element of the matrix can be automatically calculated by exploiting ADT
-definitions. % \cite{DBLP:conf/haskell/MistaRH18}
+definitions, % \cite{DBLP:conf/haskell/MistaRH18}
 %
@@
 %
-This pattern can be used to build the rest of the mean matrix as well.
+This reasoning can be used to build the rest of the mean matrix analogously.
 
@@
 In that manner, we can extend our mean matrix $M$ adding a row and a column for
-each different pattern matching and function call as shown in Fig.
+each different pattern matching and function call as shown in Figure
 \ref{fig:meanmatrix}.
@@
 The new cells are filled as before: we need to consider the amount of holes when
-generating partial pattern matching and function calls as well as their
+generating partial pattern matching values and function calls as well as their
 individual probabilities.
@@
 whereas $\mathcal{P}$ simply represents the probability of generating that
-partial ADT value.% (this probability can be easily calculated based on the
+partial ADT value.
+% (this probability can be easily calculated based on the
 % generation frequencies used in the random generator).
 %
-Then, the equation \ref{eqn:prediction} can be rewritten as:
+Then, the equation (\ref{eqn:prediction}) can be rewritten as:
 %
@@
 For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
-values are as shown in Fig. \ref{fig:vectors}.
+values are as shown in Figure \ref{fig:vectors}.
 %
@@
 and |Join| that we will generate by simply expanding the expected number of
-generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
+generated pattern matching values |simplify1| and |simplify2| into their
 corresponding data constructors:

### 2019-02-06 paper/relatedwork.tex
Minor fixes

@@
 %
-In a recent work, Bendkowski provides a framework called \emph{boltzmann-brain}
-to specify and synthesize standalone Haskell random generators based on
-Boltzmann models \cite{bendkowski2018}.
+In a recent work, Bendkowski et al. provides a framework called
+\emph{boltzmann-brain} to specify and synthesize standalone Haskell random
+generators based on Boltzmann models \cite{bendkowski2018}.
 %
-This framework mixes parameter tuning and rejection of unwanted samples to
-approximate the desired distribution of values according to user demands.
+This framework mixes parameter tuning and rejection of samples of unwanted sizes
+to approximate the desired distribution of values according to user demands.
 %
-In this approach, the overall discard ratio depends on how constrained the
-desired distribution is.
+The overall discard ratio then depends on how constrained the desired sizes of
+values are.
 %
@@
 %
-Although promising, we found several difficulties to compare both approaches in
-practice due that \emph{boltzmann-brain} is considered a standalone utility that
-produces self-contained samplers.
+Although promising, we found difficulties to compare both approaches in practice
+due that \emph{boltzmann-brain} is considered a conceptual standalone utility
+that produces self-contained samplers.
 %

### 2019-03-12 paper/synthesis.tex
Preparing for camera-ready version

@@
 \end{align}
-
+%
 Vector $E[G_0]$ represents the initial distribution of constructors that our
@@
 For instance, the average number of |Text| data constructors that we will
-generate provided that we generated a |Join| data constructor on the previous
-level is calculated as follows:
+generate provided that we generated a |Join| constructor on the previous
+level results:
 %
 $$ M_{\mathrm{Join, Text}} = 2\cdot p_{\mathrm{Text}} $$
-%
+
 where $2$ is the number of holes present when generating a partial ADT value

### 2019-03-12 paper/finalremarks.tex

--- 2019-01-31 paper/finalremarks.tex
@@
 our automatically derived generators, as well as evaluating how introducing
-random mutations into structurally complex values might improve testing
-performace, specially when generating ``almost correct'' random data.
+random mutations into structurally complex values might improve the testing
+performance, specially when generating ``almost correct'' random
+data.\looseness=-1

### 2019-03-12 paper/introduction.tex
Added citation to Robby Findler on randomized testing for PLT Redex

@@
 programs' expected behavior and (ii) \emph{random data generators} based on the
-\emph{types} of the expected inputs (e.g., an integer, an string,
+\emph{types} of the expected inputs (e.g., integers, strings,
 etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
@@
 Well-typed ADT values are specially useful when testing programs which expect
-highly structured inputs like compilers \cite{Palka11,MidtgaardJKNN17}.
+highly structured inputs like compilers
+\cite{Palka11,MidtgaardJKNN17,klein2009randomized}.
 %
@@
 %
-We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
-  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable of
-automatically synthesizing \quickcheck generators which produce \emph{rich ADT
-  values}, where the distributions of random values can be adjusted at
-compile-time to what developers might want. %based on our predictions.
+We implement our ideas as an extension of the already existing derivation tool
+called \dragen\cite{DBLP:conf/haskell/MistaRH18}. We call our extension as
+\dragenp\footnote{\dragenp is available at
+  \url{http://github.com/OctopiChalmers/dragen-struct}} to make it easy the
+distinction for the reader.
+%
+\dragenp is capable of automatically synthesizing \quickcheck generators which
+produce \emph{rich ADT values}, where the distributions of random values can be
+adjusted at compile-time to what developers might want.
+% based on our predictions.
 %

### 2019-03-12 paper/randomtesting.tex

@@
 When it picks to generate a |Text| or a |Single| data constructor, it also
-generates a random |String| value using the standard QuickCheck generator
+generates a random |String| value using the standard \quickcheck generator
 |genString|.% (\quickcheck provides standard random generators for most base types
@@
 On the other hand, when it picks to generate a |Join| constructor, it also
-generates two independent random sub-expression recursively, decreasing the
+generates two independent random sub-expressions recursively, decreasing the
 generation size by a unit on each recursive invocation (|smaller genHtml|).

### 2019-03-12 paper/tikz/hrep.tex
Minor fixes
Added final remarks, tweaks on highlighting

@@
 
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1.1,xscale=1.1]
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
 %uncomment if require: \path (0,439.4285583496094); %set diagram left start at 0, and has height of 439.4285583496094
@@
 % Text Node
-\draw (85,133.5) node [scale=0.8]  {$\text{\textbf{module}}\ M$};
+\draw (85,133.5) node [scale=0.8]  {$\text{\textbf{module}}\ \mathrm{M}$};
 % Text Node
-\draw (100,159.5) node [scale=0.8]  {$\text{\textbf{data}}\ Html\ =\ ... $};
+\draw (100,159.5) node [scale=0.8]  {$\text{\textbf{data}}\ \mathrm{Html}\ =\ ... $};
 % Text Node
-\draw (107,219.5) node [scale=0.8]  {$div::Html\rightarrow Html$};
+\draw (107,219.5) node [scale=0.8]  {$\mathrm{div}::\mathrm{Html}\rightarrow \mathrm{Html}$};
 % Text Node
-\draw (109,233.5) node [scale=0.8]  {$bold::Html\rightarrow Html$};
+\draw (109,233.5) node [scale=0.8]  {$\mathrm{bold}::\mathrm{Html}\rightarrow \mathrm{Html}$};
 % Text Node
-\draw (86,246.5) node [scale=0.8]  {$hr::Html$};
+\draw (84,246.5) node [scale=0.8]  {$\mathrm{hr}::\mathrm{Html}$};
 % Text Node
-\draw (275.5,135.5) node [scale=0.8]  {$Html_{Spec}$};
+\draw (275.5,140) node [scale=0.8] [align=center] {$\mathrm{Html}$ \text{generation}\\ \text{description}};
 % Text Node
-\draw (276,165) node [scale=0.8]  {$Html_{ADT}$};
+\draw (276,165) node [scale=0.8]  {$\mathrm{Html_{ADT}}$};
 % Text Node
-\draw (276,200) node [scale=0.8]  {$Html_{Patterns}$};
+\draw (276,200) node [scale=0.8]  {$\mathrm{Html_{Patterns}}$};
 % Text Node
-\draw (276,235) node [scale=0.8]  {$Html_{Interface}$};
+\draw (276,235) node [scale=0.8]  {$\mathrm{Html_{Interface}}$};
 % Text Node
@@
 % Text Node
-\draw (405.5,133.5) node [scale=0.8]  {$Generator$};
+\draw (405.5,133.5) node [scale=0.8]  {$\text{Generator}$};
 % Text Node
@@
 % Text Node
-\draw (514.5,133.5) node [scale=0.8]  {$Html$};
+\draw (523,140) node [scale=0.8] [align=left] {$\mathrm{Html}$\\\text{test cases}};
 % Text Node
-\draw (100,184.5) node [scale=0.8]  {$simpl\ ( ...) \ =\ ... \ $};
+\draw (107,184.5) node [scale=0.8]  {$\mathrm{simplify}\ ( ...) \ =\ ... \ $};
 % Text Node
-\draw (100,198.5) node [scale=0.8]  {$simpl\ ( ...) \ =\ ... \ $};
+\draw (107,198.5) node [scale=0.8] {$\mathrm{simplify}\ ( ...) \ =\ ... \ $};
 % Text Node
-\draw (202.5,142.5) node [scale=0.7] [align=left] {\textbf{Structure}\\ \textbf{specification}};
+\draw (202.5,142.5) node [scale=0.7] [align=center] {\textbf{Structure}\\\textbf{specification}};
 % Text Node
-\draw (348.5,169.5) node [scale=0.7] [align=left] {\textbf{Generator}\\ \textbf{derivation}};
+\draw (348.5,169.5) node [scale=0.7] [align=center] {\textbf{Generator}\\ \textbf{derivation}};
 % Text Node
-\draw (466,168.5) node [scale=0.7] [align=left] {\textbf{Random} \\ \textbf{generation}};
+\draw (466,168.5) node [scale=0.7] [align=center] {\textbf{Random} \\ \textbf{generation}};
 % Text Node
-\draw (348,245) node [scale=0.7] [align=left] {\textbf{User desired} \\ \textbf{distribution}};
+\draw (348,245) node [scale=0.7] [align=center] {\textbf{User desired} \\ \textbf{distribution}};
 

### 2019-03-12 src/Test/Dragen/Struct.hs
Renamed everything from Dragen.Struct to Dragen2

@@ -1,25 +1,0 @@
-module Test.Dragen.Struct
-  ( module Test.Dragen.Struct.Algebra
-  , module Test.Dragen.Struct.BoundedArbitrary
-  , module Test.Dragen.Struct.Branching
-  , module Test.Dragen.Struct.Countable
-  , module Test.Dragen.Struct.Infix
-  , module Test.Dragen.Struct.Optimization
-  , module Test.Dragen.Struct.Prediction
-  , module Test.Dragen.Struct.Rep
-  , module Test.Dragen.Struct.Spec
-  , module Test.Dragen.Struct.TH
-  , module Test.Dragen.Struct.TypeLevel
-  ) where
-
-import Test.Dragen.Struct.Algebra
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.Branching
-import Test.Dragen.Struct.Countable
-import Test.Dragen.Struct.Infix
-import Test.Dragen.Struct.Optimization
-import Test.Dragen.Struct.Prediction
-import Test.Dragen.Struct.Rep
-import Test.Dragen.Struct.Spec
-import Test.Dragen.Struct.TH
-import Test.Dragen.Struct.TypeLevel

### 2019-03-12 src/Test/Dragen/Struct/TH/Ppr.hs

@@ -1,7 +1,0 @@
-module Test.Dragen.Struct.TH.Ppr
-  ( module Test.Dragen.Struct.TH.Ppr.Ppr
-  , module Test.Dragen.Struct.TH.Ppr.PprLib
-  ) where
-
-import Test.Dragen.Struct.TH.Ppr.Ppr
-import Test.Dragen.Struct.TH.Ppr.PprLib

### 2019-03-12 src/Test/Dragen2.hs

@@ -1,0 +1,25 @@
+module Test.Dragen2
+  ( module Test.Dragen2.Algebra
+  , module Test.Dragen2.BoundedArbitrary
+  , module Test.Dragen2.Branching
+  , module Test.Dragen2.Countable
+  , module Test.Dragen2.Infix
+  , module Test.Dragen2.Optimization
+  , module Test.Dragen2.Prediction
+  , module Test.Dragen2.Rep
+  , module Test.Dragen2.Spec
+  , module Test.Dragen2.TH
+  , module Test.Dragen2.TypeLevel
+  ) where
+
+import Test.Dragen2.Algebra
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.Branching
+import Test.Dragen2.Countable
+import Test.Dragen2.Infix
+import Test.Dragen2.Optimization
+import Test.Dragen2.Prediction
+import Test.Dragen2.Rep
+import Test.Dragen2.Spec
+import Test.Dragen2.TH
+import Test.Dragen2.TypeLevel

### 2019-03-12 src/Test/Dragen2/Algebra.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Algebra where
+module Test.Dragen2.Algebra where
 
 import Data.Kind
 import GHC.Generics

### 2019-03-12 src/Test/Dragen2/BoundedArbitrary.hs

@@ -1,11 +1,11 @@
-module Test.Dragen.Struct.BoundedArbitrary where
+module Test.Dragen2.BoundedArbitrary where
 
 import GHC.TypeLits
 import Data.Kind
 import Test.QuickCheck
 
-import Test.Dragen.Struct.Algebra
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.Algebra
+import Test.Dragen2.TypeLevel
 
 ----------------------------------------
 -- | Depth-bounded arbitrary generation

### 2019-03-12 src/Test/Dragen2/Branching.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Branching where
+module Test.Dragen2.Branching where
 
 import Data.Kind
 

### 2019-03-12 src/Test/Dragen2/Countable.hs

@@ -1,7 +1,7 @@
 {-# LANGUAGE KindSignatures  #-}
 {-# LANGUAGE DefaultSignatures  #-}
 
-module Test.Dragen.Struct.Countable where
+module Test.Dragen2.Countable where
 
 import GHC.Generics
 import Data.Kind

### 2019-03-12 src/Test/Dragen2/Infix.hs

@@ -1,7 +1,7 @@
-module Test.Dragen.Struct.Infix where
+module Test.Dragen2.Infix where
 
-import Test.Dragen.Struct.Rep
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.Rep
+import Test.Dragen2.TypeLevel
 
 -- | Infix representation of operators
 

### 2019-03-12 src/Test/Dragen2/Optimization.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Optimization where
+module Test.Dragen2.Optimization where
 
 import Data.Proxy
 import Data.Ord
@@ -10,8 +10,8 @@ import System.IO.Unsafe
 
 import qualified Data.Map.Strict as Map
 
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.Spec
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.Spec
 
 ----------------------------------------
 -- | Distance functions

### 2019-03-12 src/Test/Dragen2/Prediction.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Prediction where
+module Test.Dragen2.Prediction where
 
 import Data.Proxy
 import GHC.TypeLits
@@ -8,10 +8,10 @@ import Data.Matrix (Matrix, (<->), (<|>))
 import qualified Data.Vector as Vector
 import qualified Data.Matrix as Matrix
 
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.Branching
-import Test.Dragen.Struct.Rep
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.Branching
+import Test.Dragen2.Rep
+import Test.Dragen2.TypeLevel
 
 
 

### 2019-03-12 src/Test/Dragen2/Rep.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Rep where
+module Test.Dragen2.Rep where
 
 import GHC.TypeLits
 import GHC.Generics
@@ -7,10 +7,10 @@ import Data.Kind
 
 import Test.QuickCheck hiding (Fun)
 
-import Test.Dragen.Struct.Algebra
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.Branching
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.Algebra
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.Branching
+import Test.Dragen2.TypeLevel
 
 ----------------------------------------
 -- | Type-level combinators

### 2019-03-12 src/Test/Dragen2/Spec.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.Spec where
+module Test.Dragen2.Spec where
 
 import Data.Kind
 import GHC.TypeLits
@@ -10,13 +10,13 @@ import qualified Data.Map.Strict as Map
 
 import Test.QuickCheck
 
-import Test.Dragen.Struct.Algebra
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.Branching
-import Test.Dragen.Struct.Countable
-import Test.Dragen.Struct.Prediction
-import Test.Dragen.Struct.Rep
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.Algebra
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.Branching
+import Test.Dragen2.Countable
+import Test.Dragen2.Prediction
+import Test.Dragen2.Rep
+import Test.Dragen2.TypeLevel
 
 
 ----------------------------------------

### 2019-03-12 src/Test/Dragen2/TH.hs

@@ -1,7 +1,7 @@
 {-# LANGUAGE NamedFieldPuns #-}
 {-# LANGUAGE BangPatterns #-}
 
-module Test.Dragen.Struct.TH where
+module Test.Dragen2.TH where
 
 import Data.Proxy
 
@@ -11,16 +11,16 @@ import Language.Haskell.TH
 import Language.Haskell.TH.Desugar
 
 import Test.QuickCheck
-import Test.Dragen.Struct.Rep
-import Test.Dragen.Struct.Spec
-import Test.Dragen.Struct.Optimization
-import Test.Dragen.Struct.BoundedArbitrary
-import Test.Dragen.Struct.TypeLevel
+import Test.Dragen2.Rep
+import Test.Dragen2.Spec
+import Test.Dragen2.Optimization
+import Test.Dragen2.BoundedArbitrary
+import Test.Dragen2.TypeLevel
 
-import Test.Dragen.Struct.TH.Util
-import Test.Dragen.Struct.TH.TypeDef
-import Test.Dragen.Struct.TH.TypeInt
-import Test.Dragen.Struct.TH.FunPats
+import Test.Dragen2.TH.Util
+import Test.Dragen2.TH.TypeDef
+import Test.Dragen2.TH.TypeInt
+import Test.Dragen2.TH.FunPats
 
 ----------------------------------------
 -- | Derivation targets and dispatcher

### 2019-03-12 src/Test/Dragen2/TH/Extract.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.TH.Extract where
+module Test.Dragen2.TH.Extract where
 
 import Data.List
 
@@ -9,7 +9,7 @@ import Language.Haskell.Exts
 import qualified Language.Haskell.TH as TH
 import qualified Language.Haskell.TH.Desugar as THD
 
-import Test.Dragen.Struct.TH.Util
+import Test.Dragen2.TH.Util
 
 
 ----------------------------------------

### 2019-03-12 src/Test/Dragen2/TH/FunPats.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.TH.FunPats where
+module Test.Dragen2.TH.FunPats where
 
 import Data.Foldable
 import Control.Monad.Extra
@@ -6,14 +6,14 @@ import Control.Monad.Extra
 import Language.Haskell.TH
 import Language.Haskell.TH.Desugar
 
-import Test.Dragen.Struct.TH.Util
-import Test.Dragen.Struct.TH.Extract
+import Test.Dragen2.TH.Util
+import Test.Dragen2.TH.Extract
 
 import qualified Test.QuickCheck as QC
-import qualified Test.Dragen.Struct.Rep as Rep
-import qualified Test.Dragen.Struct.Algebra as Algebra
-import qualified Test.Dragen.Struct.BoundedArbitrary as BoundedArb
-import qualified Test.Dragen.Struct.Branching as Branching
+import qualified Test.Dragen2.Rep as Rep
+import qualified Test.Dragen2.Algebra as Algebra
+import qualified Test.Dragen2.BoundedArbitrary as BoundedArb
+import qualified Test.Dragen2.Branching as Branching
 
 ----------------------------------------
 -- | Derive the complete representation for a every pattern of a function, plus

### 2019-03-12 src/Test/Dragen2/TH/Ppr.hs

@@ -1,0 +1,7 @@
+module Test.Dragen2.TH.Ppr
+  ( module Test.Dragen2.TH.Ppr.Ppr
+  , module Test.Dragen2.TH.Ppr.PprLib
+  ) where
+
+import Test.Dragen2.TH.Ppr.Ppr
+import Test.Dragen2.TH.Ppr.PprLib

### 2019-03-12 src/Test/Dragen2/TH/Ppr/Ppr.hs

@@ -3,14 +3,14 @@
 -- | contains a prettyprinter for the
 -- Template Haskell datatypes
 
-module Test.Dragen.Struct.TH.Ppr.Ppr where
+module Test.Dragen2.TH.Ppr.Ppr where
     -- All of the exports from this module should
     -- be "public" functions.  The main module TH
     -- re-exports them all.
 
 import Text.PrettyPrint (renderStyle, Style(..), style)
 -- import Language.Haskell.TH.PprLib
-import Test.Dragen.Struct.TH.Ppr.PprLib
+import Test.Dragen2.TH.Ppr.PprLib
 import Language.Haskell.TH.Syntax
 import Data.Word ( Word8 )
 import Data.Char ( toLower, chr)

### 2019-03-12 src/Test/Dragen2/TH/Ppr/PprLib.hs

@@ -2,7 +2,7 @@
 
 -- | Monadic front-end to Text.PrettyPrint
 
-module Test.Dragen.Struct.TH.Ppr.PprLib (
+module Test.Dragen2.TH.Ppr.PprLib (
 
         -- * The document type
         Doc,            -- Abstract, instance of Show

### 2019-03-12 src/Test/Dragen2/TH/TypeDef.hs

@@ -1,5 +1,5 @@
 {-# LANGUAGE OverloadedStrings #-}
-module Test.Dragen.Struct.TH.TypeDef where
+module Test.Dragen2.TH.TypeDef where
 
 import Data.Bool
 import Data.Ord
@@ -9,13 +9,13 @@ import Control.Monad.Extra
 import Language.Haskell.TH
 import Language.Haskell.TH.Desugar
 
-import Test.Dragen.Struct.TH.Util
+import Test.Dragen2.TH.Util
 
 import qualified Test.QuickCheck as QC
-import qualified Test.Dragen.Struct.Rep as Rep
-import qualified Test.Dragen.Struct.Algebra as Algebra
-import qualified Test.Dragen.Struct.BoundedArbitrary as BoundedArb
-import qualified Test.Dragen.Struct.Branching as Branching
+import qualified Test.Dragen2.Rep as Rep
+import qualified Test.Dragen2.Algebra as Algebra
+import qualified Test.Dragen2.BoundedArbitrary as BoundedArb
+import qualified Test.Dragen2.Branching as Branching
 
 
 ----------------------------------------

### 2019-03-12 src/Test/Dragen2/TH/TypeInt.hs

@@ -1,4 +1,4 @@
-module Test.Dragen.Struct.TH.TypeInt where
+module Test.Dragen2.TH.TypeInt where
 
 import Data.List
 import Data.Foldable
@@ -7,14 +7,14 @@ import Control.Monad.Extra
 import Language.Haskell.TH
 import Language.Haskell.TH.Desugar
 
-import Test.Dragen.Struct.TH.Util
-import Test.Dragen.Struct.TH.Extract
+import Test.Dragen2.TH.Util
+import Test.Dragen2.TH.Extract
 
 import qualified Test.QuickCheck as QC
-import qualified Test.Dragen.Struct.Rep as Rep
-import qualified Test.Dragen.Struct.Algebra as Algebra
-import qualified Test.Dragen.Struct.BoundedArbitrary as BoundedArb
-import qualified Test.Dragen.Struct.Branching as Branching
+import qualified Test.Dragen2.Rep as Rep
+import qualified Test.Dragen2.Algebra as Algebra
+import qualified Test.Dragen2.BoundedArbitrary as BoundedArb
+import qualified Test.Dragen2.Branching as Branching
 
 ----------------------------------------
 -- | Derive a complete representation for every function in a module returning

### 2019-03-12 src/Test/Dragen2/TH/Util.hs

@@ -2,7 +2,7 @@
 {-# LANGUAGE StandaloneDeriving #-}
 {-# LANGUAGE CPP #-}
 
-module Test.Dragen.Struct.TH.Util where
+module Test.Dragen2.TH.Util where
 
 import System.FilePath
 import System.Directory
@@ -13,7 +13,7 @@ import Control.Monad.Extra
 import Control.Monad.IO.Class
 import System.Console.ANSI
 
-import Test.Dragen.Struct.TH.Ppr
+import Test.Dragen2.TH.Ppr
 import Language.Haskell.TH hiding (Ppr, ppr, pprint)
 import Language.Haskell.TH.PprLib
 import Language.Haskell.TH.Desugar
@@ -21,8 +21,8 @@ import Language.Haskell.TH.Desugar
 import qualified Data.Vector as Vector
 import qualified GHC.Generics as Generics
 import qualified Test.QuickCheck as QC
-import qualified Test.Dragen.Struct.Rep as Rep
-import qualified Test.Dragen.Struct.BoundedArbitrary as BoundedArb
+import qualified Test.Dragen2.Rep as Rep
+import qualified Test.Dragen2.BoundedArbitrary as BoundedArb
 
 
 -- | Derive "syntactic" equivalence relations for DTypes

### 2019-03-12 src/Test/Dragen2/TypeLevel.hs

@@ -1,5 +1,5 @@
 {-# OPTIONS_GHC -fno-warn-unticked-promoted-constructors #-}
-module Test.Dragen.Struct.TypeLevel where
+module Test.Dragen2.TypeLevel where
 
 import Data.Kind
 import Data.Proxy

### 2019-03-12 test/Main.hs

@@ -2,7 +2,7 @@ module Main where
 
 import qualified Data.Vector as Vector
 
-import Test.Dragen.Struct
+import Test.Dragen2
 
 import TreeOpt
 import Tree

### 2019-03-12 test/Tree.hs

@@ -1,6 +1,6 @@
 module Tree where
 
-import Test.Dragen.Struct
+import Test.Dragen2
 
 
 data Tree a

### 2019-03-12 test/TreeOpt.hs

@@ -1,6 +1,6 @@
 module TreeOpt where
 
-import Test.Dragen.Struct
+import Test.Dragen2
 
 import Tree
 

### 2019-03-20 papers/ast19/sections.old/introduction.tex
Added template with base structure and a bunch of code

--- 2019-01-15 paper/sections/introduction.tex
@@
 \section{Introduction}
-
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
 

### 2019-03-20 papers/ast19/sections.old/randomtesting.tex
Added template with base structure and a bunch of code

\section{Type-Driven Random Generation} \label{sec:randomtesting}

In this section we briefly describe the common technique used for automatically
deriving random data generators in Haskell using a type-driven approach.
%
Then, we introduce its main drawbacks by considering two scenarios where this
technique gives poor results in practice.
%
We remark that, altough this work makes focus on Haskell data types, this
technique is general enough to be applied to most programming languages with
some level of support for composite types.

Haskell is a strongly typed programming language with a powerful type system.
%
It lets programmers encode a considerable amount of information about the
properties of their systems using data types that can be checked at compilation
time.
%
One of its key aspects is the support for Algebraic Data Types (ADTs).
%
Essentially, an ADTs is a composite type defined by combining other types.
%
In the most basic conception, types can be combined by \textbf{sums} (also known
as \emph{variant types} or \emph{tagged disjoint unions}) and \textbf{products}
(or tuples) of other data types.
%
To exemplify this, consider the following definition of the data type |Exp|
encoding integer expressions:

\begin{code}
data Exp  =   Val  Int
          pp  Add  Exp  Exp
          pp  Mul  Exp  Exp
\end{code}

In the previous definition, we declare |Exp| as the sum of three possible
classes of values: |Val| represents literal values, whereas |Add| and |Mul|
represent the addition and multiplication of two integer expressions,
respectively.
%
In Haskell, |Val|, |Add| and |Mul| are called \emph{data constructors} (or
constructors for short) and are used to distiguish which variant of the data
type we are reffering to.
%
Each data constructor is then defined as a product of zero or more types known
as \emph{fields}.
%
In particular, |Val| contains a field of type |Int|, while |Add| and |Mul|
contain two fields of type |Exp| representing the operands of each operation.
%
Note that |Exp| is used as a field of at least one of its data constructors
(case |Add| and |Mul|), making it a recursively defined ADT.
%
In general, we will say that a data constructor with no recursive fields is
\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
one field of such nature.
%
With this reprentation, the expression ``2 + (5 * 6)'' can be encoded simply by
using |Exp| data constructors as |(Add (Val 2) (Mul (Val 5) (Val 6)))|.
%
Furthermore, an evaluation function from |Exp|s to integer values can be defined
very idiomatically:

\begin{code}
eval :: Exp -> Int
eval (Val n)    = n
eval (Add x y)  = eval x  +  eval y
eval (Mul x y)  = eval x  *  eval y
\end{code}

In the previous definition, |eval| is described using \emph{pattern matching}
over each possible kind of value.
%
For the input case of a literal values, we simply return the value contained in
the |Val| constructor.
%
On the other hand, if the input value matches either an |Add| or a |Mul| data
constructor, where |x| and |y| are \emph{pattern variables} that match any
subexpression, we recursively evaluate these subexpressions, combining them with
the apropiate operation on each case.

\subsection*{\textbf{Type-Driven Generation of Random Values}}

In order to perform random testing of a Haskell codebase involving user defined
data types, most approaches require the user to provide a random data generator
for each one of them.
%
This tends to be a cumbersome and error prone task that closely follows the data
type structure.
%
For instance, consider the following definition of a \quickcheck random
generator for the type |Exp|:

\begin{code}
gen_Exp = sized (\size ->
   if size == 0
   then Val <$> arbitrary
   else frequency
     [  (1, Val  <$> arbitrary)
     ,  (2, Add  <$> smaller gen_Exp <*> smaller gen_Exp)
     ,  (1, Mul  <$> smaller gen_Exp <*> smaller gen_Exp) ])
\end{code} %$

This random generator is defined using \quickcheck's combinator |sized| to
paremetrize the generation process up to an external integer number known as the
\emph{generation size}.
%
This parameter is used to limit the maximum amount of recursive call that this
random generator can perform.
%
When called with a positive generation size (case |gen n|), the generator will
pick among any data constructor of |Exp| with a explicitly given frequency---two
times more |Add|s than |Val| or |Mul|.
%
In the case it picks to generate a |Val| data constructor, it also generates a
random |Int| value using the standard overloaded generator |arbitrary|
(\quickcheck provides standard random generators for most base types like |Int|,
|Bool|, etc.).
%
On the other hand, when it randomly picks to generate either an |Add| or a |Mul|
data constructor, it also generates independently two random subexpressions
corresponding to the data constructor fields by calling itself recursively
(|resize pred gen_Exp|), decreasing the generation size on each recursive
invocation.

This procedure keeps calling itself recursively until the generation size
reaches zero (case |gen 0|), where the generator is constrained to pick only
among terminal data constructors, being |Val| the only possible choice in our
particular case.
%
Strictly decreasing the the generation size by one on each recursive call
results in a useful property over the generated data: every generated value has
at most |n| levels, where |n| is the generation size set by the user.
%
This property enables us to model the generation process using the theory of
branching processes introduced by \tocite, and extended in this work as
described in Section \ref{sec:synthesis}.

Having discussed the previous random generator definition, it easy to understand
how to extend this generation procedure to any data type defined in an algebraic
fashion.
%
Fortunately, there exists different meta-programming tools to avoid the user
from having to mechanically write random generators over and over again for each
user-defined ADT.
%
The simplest tool for such purpose is \megadeth.
%
Given the name of the target data type, it synthesizes a random generator for it
that behaves pretty much like the previously defined one.
%
However, picking among different data constructors with uniform probability can
lead to a generation process biased to generate (in average) very small values,
regardless of the generation sized set by the user \tocite.

\dragen is meta-programming tool conceived to mitigate this problem.
%
Instead of setting a uniform generation probability of data constructors, this
tool considers the scenario where each one can be generated following a
different (although fixed over time) generation probability.
%
Then, this tool uses the theory of branching processes to analitically predict
the average distribution of data constructors generated on each random value.
%
This prediction mechanism is used to feedback a simulation-based optimization
process that adjusts the generation probability of each data constructor in
order to obtain a particular distribution of values that can be specified by the
user, offering a more flexible testing environment while still being mostly
automated.

Both \megadeth and \dragen synthesize random generators that are theoretically
capable to generate the whole space of values of the target data type, provided
that we set a generation size suficiently big.
%
However, the limitations arise quickly when we consider that the underlying
generation model is essentially the same: they pick a single data constructor
and recusively generate each required subexpression independently.
%
In practice, this procedure is too generic to generate complex data that is
useful enough to be used for random testing.

In this work we identify two sources of additional structure information which
are not considered by the aforementioned automatic derivation tools to obtain
random data generators:

\begin{enumerate}
\item The target code behaves differently over inputs matching very specific
  patterns.
\item The target code yields part of the responsability of preserving its
  invariants to the functions of its abstract interface.
\end{enumerate}

This information can be reified in compile time and used to synthesize richer
random generators automatically. We proceed to exemplify the previous points in
detail.

\subsection*{\textbf{Presence of Complex Pattern Matchings}}

To exemplify the firts problematic scenarion that may arise, suppose we want to
use randomly generated |Exp|s to test a property comprising the following
function:

\begin{code}
foo :: Exp -> Exp
foo (Add (Add x (Val 50)) (Add (Val 25) y))
  = error "pattern #1"
foo (Mul (Val 50) (Mul (Val x) y))
  = error "pattern #2"
foo x = x
\end{code}

This function behaves very much like an identity function, with the exception
that it fails for two very specific patterns of input values defined using
nested pattern matching.
%
This is, |foo| does not only matches against the root data constructor, but also
against the data constructors of its subexpressions and sub-subexpressions.

\quickcheck's default implementation for random |Int|s pick a number in the
interval $[-n, n]$ with uniform distribution.
%
Hence, in principle we need to recognize what is a suitable generation size,
which should be big enough for our generator be able to generate the numbers we
pattern match against to, otherwise we will not be able to generate any value to
match against |foo| clauses, leaving fragments of code completely untested.
%
Suppose then that we pick the a generation size $50$, i.e. the minimum size that
is big enough to produce an |Int| number equal to $50$.
%
Under this consideration, the probability of generating a value matching the
first clause of |foo| (and hence triggering the first error) results as follows:
%
\begin{align*}
  &P(match(foo\#1))\\
  &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
  &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
\end{align*}

If we use \megadeth to automatically derive a random generator for |Exp|, we
obtain a uniform generation probability distribution over constructors, i.e.,
$P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
%
In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
average, we will need to generate over than two million test cases in order to
be able to test the first clause of |foo| only once.
%
This situation can be somewhat improved if we use \dragen to obtain a random
generator.
%
Using this tool, we can optimize the generation probabilities in order to
benefit the generation of some data constructors over the rest.
%
Considering that the first pattern matching of |foo| involves the data
constructor |Add| as the only recursive one, we can set an target generation
proportion of |Add| data constructors of, for instance, $20:1$ with respect to
the rest of the generated data constructors.
%
By doing so, the obtained distribution of values results such that
$P(match(foo\#1)) \approx 1/300000$.
%
Although this certainly improves the probability of generating a matching value,
this probability is not substantial enough to become practical.

Additionally, by favoring the generation probabilities towards the |Add|
constructor, we found that the probability of generating a value matching the
second clause of |foo| (which does not matches against it) also gets diminished
into an impractival value.

Despite the fact that the previous example might look artificial, branching
against specific patterns of the input data is a common task.
%
For instance, the balancing function of a Red-Black tree need to consider a
quite specific combination of color, left and right subtrees and sub-subtrees
in order to preserve the height invariant.
%
Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
needs to consider very specific sequences of instructions that may be regrouped
in a computationally more efficient way, to cite a few.

\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}

A common choice when implementing a data structure is to transfer the
responsability of preserving its invariants to the functions that manipulates
its values.
%
Consider for instance the following possible implementation of a basic |HTML|
manipulation library:

\begin{code}
data Html = Html String

head :: Html -> Html
head (Html inner)
  = Html ("<head>" ++ inner ++ "</head>")

body :: Html -> Html
body (Html inner)
  = Html ("<body>" ++ t ++ "</body>")

br :: Html
br = Html "<br />"

(<+>) :: Html -> Html -> Html
(Html x) <+> (Html y) = Html (x ++ y)
\end{code}

In the previous definition, the |Html| data type is defined using a single data
constructor that contains the textual representation of the Html code it
represents in plain text.
%
Later, the combinator functions defined over |HTML| are the ones in charge of
transforming this plain text representation with the invariant that, given valid
|Html| parameters, they always return a valid |Html| value.
%
This is a very commmon programming pattern that can be found in a variety of
Haskell libraries \tocite{}.

If we use a standard type-driven approach to derive a random generator for
|Html|, the only ``structural'' information that we can use in the derivation
process is that |Html| values are composed of strings.
%
As a consequence, we essentially end up generating random strings, which rarely
represents a valid (or at least well structured) Html value.
%
Additionally, if our test suite contains properties constrained by certain
preconditions, this lack of domain knowleadge may lead in an impractically high
discard ratio of randomly generated test.
%
For instance, suposse we write a property to verify that |head| preserves the
invariant previously mentioned:

\begin{code}
prop_head_inv :: Html -> Bool
prop_head_inv x = valid x ===> valid (head x)
\end{code}

In the previous definition we state that |head| always returns a valid |Html|
value (|valid (head x)|), provided that we are given a valid |Html| as input
(|valid x|).
%
Then, while testing this property we rarely satisfy its precondition, in which
case \quickcheck discards the whole test, retrying with a diferent random input,
degradating this way the testing performance.

We believe that, if certain patterns of values are relevant enough to appear in
the codebase being tested, a practical random testing methodology should be able
to produce values satisfying such patterns in a substantial proportion.
%
The next section introduces a reprentation model that let us encode the
structure information presented here into our automatically derived random
generators in a modular and flexible way.

### 2019-03-20 papers/ast19/sections.old/synthesis.tex
Added template with base structure and a bunch of code

--- 2019-01-15 paper/sections/synthesis.tex
@@
 Nam euismod tellus id erat.
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.

### 2019-03-20 papers/ast19/tikz/val.tex
Added template with base structure and a bunch of code

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-0.8,xscale=0.8]
%uncomment if require: \path (0,440.8571319580078); %set diagram left start at 0, and has height of 440.8571319580078

%Rounded Rect [id:dp4443669613398804]
\draw  [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ] (100,66) .. controls (100,62.69) and (102.69,60) .. (106,60) -- (244,60) .. controls (247.31,60) and (250,62.69) .. (250,66) -- (250,84) .. controls (250,87.31) and (247.31,90) .. (244,90) -- (106,90) .. controls (102.69,90) and (100,87.31) .. (100,84) -- cycle ;

%Rounded Rect [id:dp2445219502409255]
\draw  [color={rgb, 255:red, 65; green, 117; blue, 5 }  ,draw opacity=1 ] (25,66) .. controls (25,62.69) and (27.69,60) .. (31,60) -- (84,60) .. controls (87.31,60) and (90,62.69) .. (90,66) -- (90,84) .. controls (90,87.31) and (87.31,90) .. (84,90) -- (31,90) .. controls (27.69,90) and (25,87.31) .. (25,84) -- cycle ;
%Rounded Rect [id:dp9597491817026895]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (30,26) .. controls (30,22.69) and (32.69,20) .. (36,20) -- (109,20) .. controls (112.31,20) and (115,22.69) .. (115,26) -- (115,44) .. controls (115,47.31) and (112.31,50) .. (109,50) -- (36,50) .. controls (32.69,50) and (30,47.31) .. (30,44) -- cycle ;

%Straight Lines [id:da4767955800901629]
\draw    (90,35) -- (75,60) ;

\draw [shift={(90,35)}, rotate = 120.96] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;
%Straight Lines [id:da737226333560641]
\draw    (105,35) -- (115,60) ;

\draw [shift={(105,35)}, rotate = 68.2] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;
%Rounded Rect [id:dp9147190163305616]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (40,106) .. controls (40,102.69) and (42.69,100) .. (46,100) -- (144,100) .. controls (147.31,100) and (150,102.69) .. (150,106) -- (150,124) .. controls (150,127.31) and (147.31,130) .. (144,130) -- (46,130) .. controls (42.69,130) and (40,127.31) .. (40,124) -- cycle ;
%Straight Lines [id:da2709254544231703]
\draw    (80,75) -- (85,100) ;

\draw [shift={(80,75)}, rotate = 78.69] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;

%Shape: Rectangle [id:dp9665612509271984]
\draw   (15,10) -- (260,10) -- (260,140) -- (15,140) -- cycle ;

%Rounded Rect [id:dp8911386762864086]
\draw  [color={rgb, 255:red, 208; green, 2; blue, 27 }  ,draw opacity=1 ] (121,214.86) .. controls (121,211.54) and (123.69,208.86) .. (127,208.86) -- (325,208.86) .. controls (328.31,208.86) and (331,211.54) .. (331,214.86) -- (331,232.86) .. controls (331,236.17) and (328.31,238.86) .. (325,238.86) -- (127,238.86) .. controls (123.69,238.86) and (121,236.17) .. (121,232.86) -- cycle ;

%Rounded Rect [id:dp28519778620668634]
\draw  [color={rgb, 255:red, 65; green, 117; blue, 5 }  ,draw opacity=1 ] (26,214.86) .. controls (26,211.54) and (28.69,208.86) .. (32,208.86) -- (105,208.86) .. controls (108.31,208.86) and (111,211.54) .. (111,214.86) -- (111,232.86) .. controls (111,236.17) and (108.31,238.86) .. (105,238.86) -- (32,238.86) .. controls (28.69,238.86) and (26,236.17) .. (26,232.86) -- cycle ;
%Rounded Rect [id:dp42988108204777453]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (71,174.86) .. controls (71,171.54) and (73.69,168.86) .. (77,168.86) -- (130,168.86) .. controls (133.31,168.86) and (136,171.54) .. (136,174.86) -- (136,192.86) .. controls (136,196.17) and (133.31,198.86) .. (130,198.86) -- (77,198.86) .. controls (73.69,198.86) and (71,196.17) .. (71,192.86) -- cycle ;
%Straight Lines [id:da7584739344782256]
\draw    (111,183.86) -- (96,208.86) ;

\draw [shift={(111,183.86)}, rotate = 120.96] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;
%Straight Lines [id:da9525312529597059]
\draw    (126,183.86) -- (141,208.86) ;

\draw [shift={(126,183.86)}, rotate = 59.04] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;
%Rounded Rect [id:dp10243358267189961]
\draw  [color={rgb, 255:red, 74; green, 144; blue, 226 }  ,draw opacity=1 ] (81,254.86) .. controls (81,251.54) and (83.69,248.86) .. (87,248.86) -- (160,248.86) .. controls (163.31,248.86) and (166,251.54) .. (166,254.86) -- (166,272.86) .. controls (166,276.17) and (163.31,278.86) .. (160,278.86) -- (87,278.86) .. controls (83.69,278.86) and (81,276.17) .. (81,272.86) -- cycle ;

%Straight Lines [id:da01173054188383471]
\draw    (101,223.86) -- (106,248.86) ;

\draw [shift={(101,223.86)}, rotate = 78.69] [color={rgb, 255:red, 0; green, 0; blue, 0 }  ][fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.75]      (0, 0) circle [x radius=2.5, y radius=2.5]   ;

%Shape: Rectangle [id:dp5499992274881935]
\draw   (16,158.86) -- (340,158.86) -- (340,290) -- (16,290) -- cycle ;

%Down Arrow [id:dp3191824787828028]
\draw  [fill={rgb, 225:red, 225; green, 225; blue, 225 }  ,fill opacity=1 ] (215,164) -- (225,164) -- (225,125) -- (245,125) -- (245,164) -- (255,164) -- (235,190) -- cycle ;

% Text Node
\draw (175,75) node [scale=0.8]  {$Pat_{simpl\#1}\ "abc"\ "xyz"$};
% Text Node
\draw (120.5,115.5) node  [scale=0.8] {$"hello"$};
% Text Node
\draw (51.5,75.5) node  [scale=0.8] {$Fun_{div}$};
% Text Node
\draw (57,35) node  [scale=0.8] {$Con_{Join}$};
% Text Node
\draw (68,115.5) node  [scale=0.8] {$Con_{Text}$};
% Text Node
\draw (62.5,225) node  [scale=0.8] {$Tag\ "div"\ $};
% Text Node
\draw (123,263.36) node  [scale=0.8] {$Text\ "hello"$};
% Text Node
\draw (87.5,182.36) node  [scale=0.8] {$Join$};
% Text Node
\draw (227.5,225) node  [scale=0.8] {$Join\ (Text\ "abc")\ (Text\ "xyz")$};
% Text Node
\draw (247,22.5) node [scale=1]  {$x$};
% Text Node
\draw (307.5,175) node [scale=1]  {$\llbracket x\rrbracket _{Html}$};

\end{tikzpicture}

### 2019-03-20 papers/haskell19/extensions.tex
Added template with base structure and a bunch of code

--- 2019-01-15 paper/sections/synthesis.tex
@@
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+\section{Extensions}
+\label{sec:extensions}
 
@@
 est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.

### 2019-03-20 papers/haskell19/final.tex
Added template with base structure and a bunch of code

--- 2019-01-15 paper/sections/introduction.tex
@@
-\section{Introduction}
+\section{Final Remarks}
+\label{sec:final}
 
@@
 lacinia purus, et dictum nunc justo sit amet elit.
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}

### 2019-03-20 papers/haskell19/introduction.tex
Added template with base structure and a bunch of code

--- 2019-01-15 paper/sections/introduction.tex
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+\begin{code}
+data Html
+  =  Text  String
+  |  Sing  String
+  |  Tag   String  Html
+  |  Join  Html    Html
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
+hr :: Html
+hr = Sing "hr"
 
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}
+bold :: Html -> Html
+bold x = Tag "b" x
+
+div :: Html -> Html
+div x = Tag "div" x
+\end{code}

### 2019-03-20 papers/haskell19/relatedwork.tex
Added template with base structure and a bunch of code

--- 2019-01-20 paper/main.tex
@@
-\documentclass[conference, fleqn]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% HRep
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Constructor names
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% gen
-
-% Local Variables:
-% TeX-mbigstarer: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usepackage{relsize}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITE}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\newcommand{\hrep}[1]{\ensuremath{H\!Rep_{#1}}}
-\newcommand{\gen}[1]{\ensuremath{gen_{#1}}}
-\newcommand{\term}[1]{\ensuremath{#1^\bigstar}}
-\newcommand{\evalrep}[2]{ \ensuremath{\llbracket #1 \rrbracket_{#2}}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some parameters
-\setlength{\mathindent}{\parindent}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Deriving Random Generators via\\ Higher-Level Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-We remark that, altough this work focuses on Haskell data types, this technique
-is general enough to be applied to most programming languages with some level of
-support for composite types.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data type by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract useful
-  structure information from the user source code.
-  %
-\end{itemize}
-\section{Random Data Generation in Haskell} \label{sec:randomtesting}
-
-In this section we briefly introduce the common approach for automatically
-deriving random data generators in Haskell using a type-driven approach, along
-with its main drawbacks.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-structure of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types in
-terms of \textbf{sums} (also known as \emph{variant types}) and
-\textbf{products} (or tuples) of other data types.
-%
-To exemplify this, consider the following type definition to encoding \ensuremath{\Conid{Html}}
-expressions:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Html}{}\<[12]%
-\>[12]{}\mathrel{=}{}\<[16]%
-\>[16]{}\Conid{Text}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Sing}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Join}\;{}\<[22]%
-\>[22]{}\Conid{Html}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Tag}\;{}\<[22]%
-\>[22]{}\Conid{String}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Html}} as the sum of four possible
-constructions: \ensuremath{\Conid{Text}} represents plain text values. \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Tag}} represent
-singular and paired Html tags, and \ensuremath{\Conid{Join}} concats two expression one after
-another.
-%
-We only encode a very small subset of the actual Html specification for
-illustrative reasons.
-%
-In Haskell, \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}} are known as \emph{data
-  constructors} (or constructors for short) and are used to distiguish which
-variant of the data type we are constructing.
-%
-Each data constructor is defined as a product of zero or more types known as
-\emph{fields}.
-%
-For instance, \ensuremath{\Conid{Text}} has a field of type \ensuremath{\Conid{String}}, whereas \ensuremath{\Conid{Join}} two recusive
-fields of type \ensuremath{\Conid{Html}}.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``\ensuremath{<\!\!html\!\!>\!\!hello\!\!<\!\!hr\!\!>\!\!bye\!\!<\!\!/html\!\!>}'' can be encoded as:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Conid{Tag}\;\text{\ttfamily \char34 html\char34}\;(\Conid{Join}\;(\Conid{Join}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}(\Conid{Text}\;\text{\ttfamily \char34 hello\char34})\;(\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}))\;(\Conid{Text}\;\text{\ttfamily \char34 bye\char34})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-Additionally, we can define a function \ensuremath{\Varid{render}} to serialize \ensuremath{\Conid{Html}} values as
-follows:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{render}\mathbin{::}\Conid{Html}\to \Conid{String}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Text}\;\Varid{t})\mathrel{=}\Varid{t}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Sing}\;\Varid{t})\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{render}\;\Varid{x}\plus \Varid{render}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}\plus \Varid{render}\;\Varid{x}\plus \text{\ttfamily \char34 </\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{render}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-Using pattern matching we can define functions idiomatically by defining
-different function clauses for each input pattern we are interested on.
-%
-Patterns can be defined to match specific constructors, literal values or
-variable subexpressions (like \ensuremath{\Varid{t}}, \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} in the definition of \ensuremath{\Varid{render}}).
-%
-They can also be nested in order to match very specific patterns of values.
-
-\subsection*{\textbf{Type-Driven Generation of Random Values}}
-
-In order to generate random value of types involving user defined ADTs, most
-approaches require the user to provide a random data generator for each one of
-them.
-%
-This is a cumbersome and error prone task that closely follows the data type
-structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Html}}:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{6E}{@{}l@{}}%
-\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{Html}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary})\mskip1.5mu]{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{3},{}\<[14]%
-\>[14]{}\Conid{Join}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Html}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{4},{}\<[14]%
-\>[14]{}\Conid{Tag}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-%
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
-paremetrize the generation process up to an external natural number known as the
-\emph{generation size}.
-%
-This parameter is chosen by the user, and we use to limit the maximum amount of
-recursive calls that this random generator can perform.
-%
-When called with a positive generation size, this generator can pick to generate
-among any \ensuremath{\Conid{Html}} data constructor of with a explicitly given frequency that can
-be chosen by the user (2, 1, 3 and 4 for \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}}, \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}}
-constructors, respectively).
-%
-When it picks to generate a \ensuremath{\Conid{Val}} or a \ensuremath{\Conid{Sing}} data constructor, it also
-generates a random \ensuremath{\Conid{String}} value using the standard overloaded generator
-\ensuremath{\Varid{arbitrary}} (\quickcheck provides standard random generators for most base types
-like \ensuremath{\Conid{String}}, \ensuremath{\Conid{Int}}, \ensuremath{\Conid{Bool}}, etc.). \footnote{%
-  The operators \ensuremath{\mathop{\mathsmaller{\langle \$ \rangle}}} and \ensuremath{\mathop{\mathsmaller{\langle \ast \rangle}}} are used in Haskell to combine values obtained
-  from effectful computations (like calling to a random generator) and they are
-  not particularly relevant for the point being made in this work.}
-%$
-On the other hand, when it picks to generate either an \ensuremath{\Conid{Join}} constructor, it
-also generates two independent random subexpression recursively, decreasing the
-generation size by a unit on each recursive invocation (\ensuremath{\Varid{smaller}\;gen_{Html}}).
-%
-The case of random generation of \ensuremath{\Conid{Tag}} constructors follows analogously.
-
-This random process keeps calling itself recursively until the generation size
-reaches zero, where the generator is constrained to pick only among terminal
-data constructors, being \ensuremath{\Conid{Text}} and \ensuremath{\Conid{Sing}} the only possible choices in our
-particular case.
-
-% Strictly decreasing the the generation size by one on each recursive call
-% results in a useful property over the generated data: every generated value has
-% at most |n| levels, where |n| is the generation size set by the user.
-%
-% This property enables us to model the generation process using the theory of
-% branching processes introduced by \tocite, and extended in this work as
-% described in Section \ref{sec:synthesis}.
-
-The previous definition is rather mechanical, except perhaps for the chosen
-generation frequencies.
-%
-In this light, it easy to extend this procedure to any data type defined in an
-algebraic fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth \tocite.
-%
-Given a target data type, it synthesizes a random generator for it that behaves
-similarly to the one presented above, where the generation frequencies are
-defined to be uniform across constructors.
-%
-However, picking among different data constructors with uniform frequency can
-lead to a generation process biased towards generating (in average) very small
-values, regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool this tool uses the theory of branching processes to modelize and predict
-analitically the average distribution of data constructors generated on each
-random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation frequency of each data constructor in order
-to obtain a particular distribution of values that can be specified by the user,
-providing this way a more flexible testing environment while still being mostly
-automated.
-
-Althogh both \megadeth and \dragen synthesize random generators that are
-theoretically capable to generate the whole space of values of the target data
-type. the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is often too generic to generate random data with
-enough structural complexity to be used for testing purposes.
-
-% In this work we identify two sources of additional structure information which
-% are not considered by the aforementioned automatic derivation tools to obtain
-% better random data generators:
-
-In this work we identify two patological situations which are not properly
-handled by the aforementioned derivation tools:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching specific patterns
-  of nested values.
-\item The target code encodes a significant portion of its structure on its
-  abstract interface.
-\end{enumerate}
-
-In Section \ref{sec:hrep} we show how this structural information can be used to
-synthesize richer random generators automatically.
-%
-We proceed to exemplify the previous points in detail.
-
-\subsection*{\textbf{Presence of Complex Pattern Matchings}}
-
-To exemplify the first problematic scenario, suppose we want to use randomly
-generated \ensuremath{\Conid{Htmls}}s to test a property comprising the following function:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Text}\;(\Varid{t1}\plus \Varid{t2}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y}))){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Join}\;(\Varid{simpl}\;\Varid{x})\;(\Varid{simpl}\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x})\mathrel{=}\Conid{Tag}\;\Varid{t}\;(\Varid{simpl}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function simplifies sequences of \ensuremath{\Conid{Text}} constructors into a single big
-\ensuremath{\Conid{Text}} constructor.
-%
-To do so, it has to pattern match against sequences of \ensuremath{\Conid{Text}} constructors
-combined by a \ensuremath{\Conid{Join}} constructor using nested patterns (see \ensuremath{\Varid{simpl}} fist and
-second clauses).
-%
-The remaining clauses are only meant to propagate this simplification within
-nested expressions.
-
-Ideally, we would like to test each clause of the function \ensuremath{\Varid{simpl}} approximately
-the same amount of time each.
-%
-However, each data constructor is generated independently when using either
-\megadeth or \dragen, thus the probability of generating a value satisfying a
-nested pattern decreases multiplicatively with the number of constructors we
-pattern against to simultaneously.
-%
-In our tests, we found that the first two clauses of \ensuremath{\Varid{simpl}} get exercised only
-approximately $1.5\%$ of the time when using \megadeth to derive a random
-generator for \ensuremath{\Conid{HTML}}.
-%
-On the other hand, the best result we could achieve with \dragen was only able
-to exercise the first and second clauses of \ensuremath{\Varid{simpl}} approximately $3\%$ and
-$6\%$ of the time, respectively.
-%
-With both derivation tools, the most of the time was spent testing the trivial
-clauses of our function, in view of they pattern match against simpler patterns
-of input values.
-
-% In this light, we found that using a random generator obtained with \megadeth to
-% test this function will only exercise the first two clauses of |simpl| about
-% only $1.5\%$ of the time each, relegating the rest of the generated test cases
-% to the remaining and somewhat less interesting clauses, where, remarkably, half
-% of the time is spent exercising the last trivial clause.
-% %
-% Furthermore, using \dragen to ...
-
-% This function behaves very much like an identity function, with the exception
-% that it fails for two very specific patterns of input values defined using
-% nested pattern matching.
-% %
-% This is, |foo| does not only matches against the root data constructor, but also
-% against the data constructors of its subexpressions and sub-subexpressions.
-
-% \quickcheck's default implementation for random |Int|s pick a number in the
-% interval $[-n, n]$ with uniform distribution.
-% %
-% Hence, in principle we need to recognize what is a suitable generation size,
-% which should be big enough for our generator be able to generate the numbers we
-% pattern match against to, otherwise we will not be able to generate any value to
-% match against |foo| clauses, leaving fragments of code completely untested.
-% %
-% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-% is big enough to produce an |Int| number equal to $50$.
-% %
-% Under this consideration, the probability of generating a value matching the
-% first clause of |foo| (and hence triggering the first error) results as follows:
-% %
-% \begin{align*}
-%   &P(match(foo\#1))\\
-%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
-%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
-% \end{align*}
-
-% If we use \megadeth to automatically derive a random generator for |Exp|, we
-% obtain a uniform generation probability distribution over constructors, i.e.,
-% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
-% %
-% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-% average, we will need to generate over than two million test cases in order to
-% be able to test the first clause of |foo| only once.
-% %
-% This situation can be somewhat improved if we use \dragen to obtain a random
-% generator.
-% %
-% Using this tool, we can optimize the generation probabilities in order to
-% benefit the generation of some data constructors over the rest.
-% %
-% Considering that the first pattern matching of |foo| involves the data
-% constructor |Add| as the only recursive one, we can set an target generation
-% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
-% the rest of the generated data constructors.
-% %
-% By doing so, the obtained distribution of values results such that
-% $P(match(foo\#1)) \approx 1/300000$.
-% %
-% Although this certainly improves the probability of generating a matching value,
-% this probability is not substantial enough to become practical.
-
-% Additionally, by favoring the generation probabilities towards the |Add|
-% constructor, we found that the probability of generating a value matching the
-% second clause of |foo| (which does not matches against it) also gets diminished
-% into an impractival value.
-
-Altough the previous example might seem rather simple, branching against
-specific patterns of the input data is a common task.
-%
-For instance, balancing a Red-Black tree requires to consider specific
-combinations of color, left and right subtrees and sub-subtrees in order to
-preserve the height invariant \tocite.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few \tocite.
-
-\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
-
-A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions which manipulates
-its values.
-%
-For this purspose, suppose we extend our \ensuremath{\Conid{Html}} data type with the following
-basic combinators:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 div\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 b\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathbin{::}\Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathrel{=}\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-% In the previous definition, the |Html| data type is defined using a single data
-% constructor that contains the textual representation of the Html code it
-% represents in plain text.
-%
-These functions encode additional information about the structure of our \ensuremath{\Conid{Html}}
-data type in the form of specific Html tags.
-
-Instead of including a new data constructor for each possible Html tag in out
-type definition, we defined a mininal representation and then extended it with
-an set of high level combinators.
-%
-% These combinators functions defined over |HTML| are the ones in charge of
-% transforming this plain text representation with the invariant that, given valid
-% |Html| parameters, they always return a valid |Html| value.
-%
-This programming pattern is often called a ``shallow embedding'', and can be
-found in a variety of Haskell libraries, being \emph{html} \tocite,
-\emph{svg-builder} \tocite some examples of this.
-
-As a consequence of this practice, type-driven derivation techniques often fail
-to synthesize useful random generators due to that most of the data type
-structure has been encoded into its abstract interface of combinators.
-%
-In our particular case, the chances of generating a \ensuremath{\Conid{Tag}} value representing a
-commonly used Html tag such as \ensuremath{\Varid{div}} are extremely low.
-
-% Additionally, if our test suite contains properties constrained by certain
-% preconditions, the lack of domain knowleadge may lead in an impractically high
-% discard ratio of randomly generated test cases.
-% %%
-% For instance, suposse we write a property to verify that |render| always outputs
-% a valid |Html| value:
-
-% \begin{code}
-% prop_render :: Html -> Bool
-% prop_render x = valid x ===> valid (parse (render x))
-% \end{code}
-
-% In the previous definition we state that |render| always returns a valid |Html|
-% value when it is parsed again (|parse (render x)|), provided that we are given a
-% valid |Html| as input (|valid x|).
-% %
-% In practice, not every string of characters constitutes a valid Html given that
-% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).
-
-% %
-% While testing this property we rarely satisfy its precondition, in which case
-% \quickcheck discards the whole test, retrying with a diferent random input,
-% degradating this way the testing performance.
-
-% \begin{code}
-% text :: String -> Bool
-% text str = Text (addEscapes str)
-% \end{code}
-
-% We believe that, if certain patterns of values are relevant enough to appear in
-% the codebase being tested, a practical random testing methodology should be able
-% to produce values satisfying such patterns in a substantial proportion.
-
-So far we have introduced two testing scenarios where type-driven derivation
-approaches are unable to capture all the available structure information from
-the user codebase.
-%
-Fortunately, this information can be automatically exploited and used to
-generate interesting and more structured random values.
-
-The next section introduces a reprentation model that let us encode the
-structure information presented here into our automatically derived random
-generators in a modular and flexible way.
-\section{Extracting Structure} \label{sec:hrep}
-
-In this section we present a compositional representation of values to express
-the random generation of values following the internal structure of their data
-types along with the structure present on patterns matchings and abstract
-interfaces.
-
-The key idea of this work is to represent different structured constructions of
-data in a homogeneous way that we call a ``higher-level representation'' (\ensuremath{\Conid{HRep}}
-from now on).
-%
-Instead of generating each data constructor independently, a random generator
-derived from this representation might generate composite structured values on
-each random choice it performs.
-%
-% \footnote{The notion of higher-level comes from that the generation process is
-%   entirely determined by the type of the chosen type level representation,
-%   instead of by a concrete generator defined at the term level.}
-%
-For this purpose, we use a series of automatically derived data types, each one
-representing an atomic unit of information that can be randomly generated and
-then reflected back to the corresponding value of the original data type.
-%
-Later, the user can compose these atomic representations using the provided type
-level combinators in different ways into a ``generation specification'' that
-completely determines the generation process behavior.
-
-We wil reuse the previously defined data type \ensuremath{\Conid{Html}} and the functions defined
-in the previous section to explain the different concepts involved all across
-this section.
-
-\subsection*{\textbf{Representing Data Constructors}}
-
-We begin by introducing the simplest data type representation that we can
-extract from our codebase: the representation of single data constructor.
-%
-Each data constructor can be represented by an automatically derived data type
-consisting of a single constructor with the same fields as the original, except
-for the recursive ones that are abstracted away.
-%
-In this light, we represent each constructor of the data type \ensuremath{\Conid{Html}} as follows:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{Text}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Text}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Sing}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Sing}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Join}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Join}\;{}\<[29]%
-\>[29]{}\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Tag}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Tag}\;{}\<[29]%
-\>[29]{}\Conid{String}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Note that the previous definitions are type parametric over the type parameter
-\ensuremath{\Varid{r}}.
-%
-This allow us to replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
-possible values on each case.
-%
-For instance, the value (\ensuremath{Con_{Join}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\hrep{Join}\;\Conid{Int}}, while the
-value (\ensuremath{Con_{Join}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\hrep{Join}\;\Conid{Bool}}.
-
-In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
-chosen generation specification (which might be composed of several \ensuremath{\Conid{HRep}}s),
-without having to modify anything in the underlying machinery.
-
-Having the \ensuremath{\Conid{HRep}} of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
-representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
-%
-Then, we simply need to translate each constructor representation back into its
-corresponding one, translating the abstracted fields recursively:
-%
-\begin{alignat*}{4}
-  &\evalrep{\ensuremath{Con_{Text}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Text}}\ s \\
-  &\evalrep{\ensuremath{Con_{Sing}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Sing}}\ s \\
-  &\evalrep{\ensuremath{Con_{Join}}\ x\ y &&}{Html}
-    &&= \ensuremath{\Conid{Join}}\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
-  &\evalrep{\ensuremath{Con_{Tag}}\ s\ x &&}{Html}
-    &&= \ensuremath{\Conid{Tag}}\ s\ \evalrep{x}{Html}
-\end{alignat*}
-
-The missing piece is to automatically synthesize a random generators for each
-constructor representation.
-%
-For this purpose, it is important to consider that each constructor \ensuremath{\Conid{HRep}} has
-its recursive fields abstracted away with a type parameter that will be later
-instantiated with the generation specification type.
-%
-Given that this specification is unknown at the derivation time, we parametrize
-each \ensuremath{\Conid{HRep}} generator with a random generator \ensuremath{\gen{r}} that is used to generate
-random values for each recursive field:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{Text}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Text}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Sing}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Sing}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Join}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Join}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{Tag}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Tag}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Type Level Combinators}}
-
-The next step is to define a series of type level combinators to enable us
-combining the automatically derived \ensuremath{\Conid{HRep}}s in several ways.
-
-In first place, we define a type combinator (\ensuremath{\term{\_}}) to tag a \ensuremath{\Conid{HRep}} to be
-terminal, i.e., a representation that is allowed be generated when the
-generation size gets exausted:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\term{f})\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\Conid{HRep}} with an
-explicit generation frequency $n$:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The previous combinators only include information relevant to the generation
-process, in a sense that neither one adds new structure to the final
-representation.
-%
-In this light, they do not alter the evaluation semantics, and we translate them
-back to our target data type by evaluating the inner representation:
-%
-\begin{alignat*}{4}
-  &\evalrep{Term\ x &&: \ensuremath{\term{f}}    &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
-\end{alignat*}
-
-In the previous equations we explicitly annotate (using a colon) the type of the
-evaluated term for clarity.
-%
-Later, to generate these combinators is enough to wrap a generated value from
-the inner representation with the apropriate tag:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{f}^\bigstar\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Term}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\gen{f \otimes n}\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Freq}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-\ensuremath{\Conid{HRep}}s into a single one using a sum type to represent a random choice between
-them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\oplus\;\Varid{g})\;\Varid{a}\mathrel{=}\Conid{L}\;(\Varid{f}\;\Varid{a})\mid \Conid{R}\;(\Varid{g}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data constructor variant and
-evaluating the inner \ensuremath{\Conid{HRep}} accodingly:
-%
-\begin{alignat*}{3}
-  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
-  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
-\end{alignat*}
-
-Generating a composite \ensuremath{\Conid{HRep}} is slightly more complicated than before, as we
-need to perform a random choice based on the generation size and the given
-frequencies for each sub-represention:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{43}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{f \oplus g}\;\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-In the previous definition, we reflect the type level frequencies of the types
-\ensuremath{\Varid{f}} and \ensuremath{\Varid{g}} (\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}) into term-level values.
-%
-This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
-%
-Then we use these frequencies to generate each inner \ensuremath{\Conid{HRep}} in the apropriate
-proportion.
-%
-When the generation size gets exhausted, we reflect the terminal generation
-frequency of each inner \ensuremath{\Conid{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
-\ensuremath{\Varid{freq0}\;\textit{@}\!g}).
-%
-This time, however, we default the frequency reflection to $0$ for any inner
-that not tagged as terminal, avoiding to generate non-terminal constructions in
-the last step.
-
-Later, we can use these combinators to create a type synonym \ensuremath{\hrep{Html}} that
-specifies a generation schema equivalent to the one seen in the concrete random
-generator of type \ensuremath{\Conid{Html}} presented in Section \ref{sec:randomtesting}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{Html}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\term{\hrep{Text}}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\term{\hrep{Sing}}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Join}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{3}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Tag}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{4}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{\hrep{Html}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit at the definition of \ensuremath{\hrep{Html}}.
-%
-We can think of it as a ``single layer'' of representation.
-%
-To make it able to represent recursive values we need to define a last type
-level combinator to ``tie the knot'':
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Fix}\;\Varid{f}\mathrel{=}\Conid{Fix}\;(\Varid{f}\;(\Conid{Fix}\;\Varid{f})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This datatype represents the \emph{fixed point} of a parametric data type \ensuremath{\Varid{f}},
-i.e., a data type where each recursive call gets instantiated with iself.
-%
-Then, to translate fixed points back to our target data type we simply need
-translate the inner representation:
-%
-\begin{align*}
-  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
-\end{align*}
-
-% Categorically, this generic transformation of fixed points is called a
-% \emph{catamorphism}, where our evaluation relation is known as its
-% \emph{algebra} \tocite.
-
-Unlike the other combinators, to define a random generator for the fixed point
-of a certain representation \ensuremath{\Varid{f}}, we do not to parametrize the generation of the
-recursive fields of \ensuremath{\Varid{f}} over an external generator \ensuremath{\gen{r}}.
-%
-% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
-Instead, we replace it with our fixed point generator, calling itself
-recursively on any recursive field of \ensuremath{\Varid{f}} that might appear inside:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;genFix_{f}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This way we obtain a concrete recursive generator for each representation \ensuremath{\Varid{f}}
-that we define.
-%
-Then, we can define a random generator for our target data type simply by
-generating a random value of our chosen representation, and transforming it back
-to our target by the means of the evaluation relation:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Html}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{\hrep{Html}}{}\<[E]%
-\\
-\>[17]{}\Varid{return}\;\evalrep{x}{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-So far we have introduced the machinery required to represent the random
-generation of a target data type considering only the structure encoded on its
-definition.
-%
-However, this approach can now be easily extended to encode different sources of
-structured information.
-%
-We proceed to introduce two extensions that help to address the problematic
-testing scenarios presented in the previous section.
-
-\subsection*{\textbf{Representing Pattern Matchings}}
-
-We can follow a similar reasoning as before to represent the pattern matching
-structure from a given function.
-%
-Consider the nested pattern matchings of function \ensuremath{\Varid{simpl}} defined in the
-previous section:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{36}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-To represent these patterns, we derive a new data for each one of them, whose
-fields represent the pattern variables that occur inside, and abstracting away
-every pattern variable of type \ensuremath{\Conid{Html}}.
-%
-Concretely, we define the following data types:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{simpl\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#1}\;\Conid{String}\;\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{simpl\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#2}\;\Conid{String}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The first pattern of \ensuremath{\Varid{simpl}} contains two pattern variables (\ensuremath{\Varid{t1}} and \ensuremath{\Varid{t2}}) of
-type \ensuremath{\Conid{String}} that are included as fields of \ensuremath{Pat_{simpl\#1}}.
-%
-Similarly, the second pattern of \ensuremath{\Varid{simpl}} contains a pattern variable \ensuremath{\Varid{t1}} of
-type \ensuremath{\Conid{String}} represented by the first field of \ensuremath{Pat_{simpl\#2}}, along with two
-pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type \ensuremath{\Conid{Html}} that are abstracted in the second
-and third fields of \ensuremath{Pat_{simpl\#2}}.
-
-To transform these representations back to our target data type, we simply need
-to expand them it into the concrete value represented by the original pattern,
-evaluating its fields back to \ensuremath{\Conid{Html}} as well:
-%
-\begin{align*}
-  &\evalrep{\ensuremath{Pat_{simpl\#1}}\ t1\ t2}{Html} \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html})\;(\Conid{Text}\;\evalrep{t2}{Html})}\\
-  &\evalrep{\ensuremath{Pat_{simpl\#2}}\ t1\ x\ y}{Html} = \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html}))\;\evalrep{x}{Html})\;\evalrep{y}{Html})}
-\end{align*}
-
-This way, any value of type \ensuremath{\hrep{simpl\#1}} is guaranteed to satisfy the first
-pattern matching of \ensuremath{\Varid{simpl}}---the same property follows for \ensuremath{\hrep{simpl\#2}}.
-
-The random generation of pattern \ensuremath{\Conid{HRep}}s is defined in the same way as we did
-before for representations of data constructors:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{simpl\#1}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#1}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[B]{}\gen{simpl\#2}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#2}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-Finally, we can join the pattern matching representation of each clause of
-\ensuremath{\Varid{simpl}} into a single one:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{simpl}{}\<[17]%
-\>[17]{}\mathrel{=}\hrep{simpl\#1}\;\!\!\oplus\!\!\;{}\<[39]%
-\>[39]{}\hrep{simpl\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Representing Abstract Interfaces}}
-
-To introduce the higher level representation of module abstract interfaces,
-consider module \ensuremath{\Conid{M}} defininf \ensuremath{\Conid{Html}} combinators introduced in the previous
-section.
-
-% \begin{code}
-% module M where
-
-% ten :: Exp
-% ten = Val 10
-
-% square :: Exp -> Exp
-% square x = Mul x x
-
-% minus :: Exp -> Exp -> Exp
-% minus x y = Add x (Mul y (Val (-1)))
-% \end{code}
-
-We can represent each function of \ensuremath{\Conid{M}} that returns a value of type \ensuremath{\Conid{Html}}
-following the same idea as before, deriving a data type with a single data
-constructor for each one of them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{div}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{div}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{bold}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{bold}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{hr}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In this case, each single constructor will have as fields the types of the
-inputs of the function that they represent.
-%
-As before, we abstract away any field representing and input of type \ensuremath{\Conid{Html}} with
-a type parameter \ensuremath{\Varid{r}}.
-
-To evaluate these representations, we simply use the values on its fields as
-input parameter of each original function, returning its result.
-%
-\begin{alignat*}{3}
-  &\evalrep{\ensuremath{Fun_{div}}\  &&x  &&}{Html} = \ensuremath{\Varid{div}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{bold}}\ &&x &&}{Html} = \ensuremath{\Varid{bold}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{hr}}    &&  &&}{Html} = \ensuremath{\Varid{hr}}
-\end{alignat*}
-
-Note that, by doing this, the generation process inherits any patology that the
-functions we use to generate values might have.
-%
-For instance, if the function \ensuremath{\Varid{div}} would happen to be non-terminating for some
-inputs, our generation process could suffer from this as well.
-
-Furthermore, the generation procedure for abstract interface representations
-follows the same pattern as before:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{div}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{div}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{bold}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{bold}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{hr}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}\Varid{return}\;Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-And we will also define a type synonym to join all the representations of the
-module \ensuremath{\Conid{M}} into a single one.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{50}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{M}{}\<[13]%
-\>[13]{}\mathrel{=}\hrep{div}\;\!\!\oplus\!\!\;{}\<[32]%
-\>[32]{}\hrep{bold}\;\!\!\oplus\!\!\;{}\<[50]%
-\>[50]{}\hrep{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Finally, we can put all the derived machinery together into a generation
-specification \ensuremath{Html_S}, assigning (possibly) different generation frequencies to
-each individual \ensuremath{\Conid{HRep}} we combine:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;Html_S{}\<[13]%
-\>[13]{}\mathrel{=}{}\<[21]%
-\>[21]{}\hrep{Html}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{simpl}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{M}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% This previous definition can be interpreted graphically as it is shown in the
-% Figure \ref{fig:hrep}, where curly arrows represent the structural information
-% extracted using meta-programming.
-
-% \begin{figure}[t]
-%   \centering
-%   \input{tikz/hrep.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-
-We want to remark that, for space reasons, we were only able to introduce the
-representation of a rather simple target data type.
-%
-In practice, this reasoning can be extended to mutually recursive and parametric
-types as well.
-
-Overall, this approach offers significant advantages over the usual type-driven
-derivation of random generators:
-%
-\begin{itemize}
-\item \textbf{Composability:} we can combine different atomic representations
-  using different structure information sources depending on what property or
-  sub-system we need to verify using randomly generated values.
-  %
-\item \textbf{Extensibility:} the developer can derive representations for new
-  sources of structure information and combine them with the existing ones
-  simply by adding them to the existing generation specification of the target
-  data type.
-  %
-\item \textbf{Predictability:} using branching processes theory, it is possible
-  to predict the average distribution of generated values in terms of number of
-  constructors.
-  %
-  This prediction is completely modular, and can be obtained for any composite
-  representation obtained using the automatically derived \ensuremath{\Conid{HRep}}s and the
-  provided combinators.
-\end{itemize}
-
-\todo[inline]{Maybe we can show an example random value from the representation
-  and its corresponding target value }
-
-% \begin{align*}
-%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-% \end{align*}
-%
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-%
-% \begin{code}
-% data (Term f) a = TagTerm (f a)
-
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-% instance (Term f down Exp) where
-%   step (TagTerm f) = step f
-% \end{code}
-
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
-
-% \begin{figure}[b]
-%   \centering
-%   \input{tikz/val.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-\section{Random Generators Synthesis} \label{sec:synthesis}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-% %include sections/casestudies.tex
 \section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
+\label{sec:relatedwork}
 

### 2019-03-20 papers/haskell19/representation.tex
Added template with base structure and a bunch of code

\section{Unified Representation of Constructions}
\label{sec:representation}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

\begin{code}
data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
\end{code}
\begin{code}
data ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
\end{code}
\begin{code}
data Term (f :: * -> *) a = Term (f a)
\end{code}

\begin{code}
  data Fix (f :: * -> *) = Fix { unFix :: (f (Fix f)) }
\end{code}

\begin{code}
class Functor f => Algebra f a | f -> a where
  alg :: f a -> a
\end{code}

% cata :: Functor f => (f a -> a) -> Fix f -> a
% cata f = f . fmap (cata f) . unFix

\begin{code}
eval :: Algebra f a => Fix f -> a
eval = alg . fmap eval . unFix
\end{code}

\begin{code}
instance (Algebra f a, Algebra g a) => Algebra (f oplus g) a
  where  alg (InL f) = alg f
         alg (InR g) = alg g

instance Algebra f a => Algebra (f otimes n) a
  where alg (Freq f) = alg f

instance Algebra f a => Algebra (Term f) a
  where alg (Term f) = alg f
\end{code}

\begin{code}
instance BoundedArb1 f => BoundedArb1 (f otimes n) where
  liftBounded gen lvl = Freq <$> liftBounded gen lvl

instance BoundedArb1 f => BoundedArb1 (Term f) where
  liftBounded gen lvl = Term <$> liftBounded gen lvl
\end{code}

\begin{code}
type family FreqOf (f :: * -> *) :: Nat where
  FreqOf (f oplus g)   = FreqOf f + FreqOf g
  FreqOf (f otimes n)  = n * FreqOf f
  FreqOf (Term f)      = FreqOf f
  FreqOf _             = 1
\end{code}

\begin{code}
type family FreqOf' (f :: * -> *) :: Nat where
  FreqOf' (f oplus g)   = FreqOf' f  +  FreqOf' g
  FreqOf' (f otimes n)  = n  *  FreqOf' f
  FreqOf' (Term f)      = FreqOf f
  FreqOf' _             = 0
\end{code}

\begin{code}
instance (BoundedArb1 f, BoundedArb1 g)
  => BoundedArb1 (f oplus g) where
  liftBounded gen lvl =
    if lvl > 0
    then frequency
    [ (natVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
    , (natVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
    else frequency
    [ (natVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
    , (natVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
\end{code}

% Depth bounded generators
\begin{code}
type BoundedGen a = Int -> Gen a

class BoundedArb (a :: *) where
  boundedArb :: BoundedGen a
\end{code}

\begin{code}
class BoundedArb1 (f :: * -> *) where
  liftBounded :: BoundedGen a -> BoundedGen (f a)
\end{code}

\begin{code}
instance BoundedArb1 f => BoundedArb (Fix f) where
  boundedArb lvl = Fix <$> liftBounded boundedArb lvl
\end{code} %$

\begin{code}
genFix  ::  forall f dot BoundedArb1 f =>  BoundedGen (Fix f)
genFix lvl = boundedArb at(Fix f) lvl
\end{code}

\begin{code}
genEval :: forall f a dot (BoundedArb1 f, Algebra f a) =>  BoundedGen a
genEval lvl = eval <$> genFix at f lvl
\end{code} %$

\begin{code}
data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
\end{code}

\begin{code}
type family (t :: * -> *) apply (a :: *) :: * -> *  where
  (Some t)      apply  a  = t a
  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
  (f otimes n)  apply  a  = (f apply a) otimes n
  (Term t)      apply  a  = Term (t apply a)
  t             apply  a  = t
\end{code}

\begin{code}
type family Rep (t :: Symbol) :: * -> *

type family Con (c :: Symbol) :: * -> *
type family Fun (f :: Symbol) :: * -> *
type family Pat (p :: Symbol) (n :: Nat) :: * -> *

type Con' a    = Term  (Con a)
type Fun' a    = Term  (Fun a)
type Pat' a n  = Term  (Pat a n)
\end{code}

\begin{code}
type HtmlSpec
  =      Con'  "Text"
  oplus  Con   "Join"      otimes 2
  oplus  Fun'  "hr"        otimes 4
  oplus  Fun   "div"       otimes 9
  oplus  Pat   "simpl"  1  otimes 3
  oplus  Pat   "simpl"  2  otimes 5
\end{code}

### 2019-03-20 papers/haskell19/watch.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+
+#https://facebook.github.io/watchman/docs/watchman-make.html
+watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' '[^.]*.tex' 'tikz/*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all

### 2019-03-26 papers/haskell19/introduction.tex
More code

@@
   |  Tag   String Html
-  |  Join  Html    Html
+  |  Html  :+: Html
+\end{code}
 
-hr :: Html
-hr = Sing "hr"
+\begin{code}
+br :: Html
+br = Sing "br"
 
 bold :: Html -> Html
-bold x = Tag "b" x
+bold = Tag "b"
 
-div :: Html -> Html
-div x = Tag "div" x
+(<+>) :: Html -> Html -> Html
+(<+>) x y = x :+: br :+: y
+
+list :: [Html] -> Html
+list xs = Tag "ul" (foldl1 (:+:) (Tag "li" <$> xs))
+\end{code} %$
+
+Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
+commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
+lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
+felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
+dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
+quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
+Nam euismod tellus id erat.
+
+\begin{code}
+simplify :: Html -> Html
+simplify (Text t1 :+: Text t2)
+  = Text (t1 ++ t2)
+simplify ((Text t1 :+: x) :+: y)
+  = simplify (Text t1 :+: simplify (x :+: y))
+simplify (x :+: y)
+  = simplify x :+: simplify y
+simplify (Tag t x)
+  = Tag t (simplify x)
+simplify x = x
 \end{code}
+
+Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
+tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
+leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
+wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
+mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
+commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
+vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
+varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
+tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
+lacinia purus, et dictum nunc justo sit amet elit.

### 2019-03-26 papers/haskell19/representation.tex
More code

@@
 \begin{code}
+class Arbitrary (a :: *) where
+  arbitrary :: Gen a
+\end{code}
+
+\begin{code}
+class Arbitrary1 (f :: * -> *) where
+  liftGen :: Gen a -> Gen (f a)
+\end{code}
+
+\begin{code}
+instance Arbitrary Html where
+  arbitrary = sized gen
+    where
+      gen 0 = frequency
+        [  (2,  Text    <$> arbitrary)
+        ,  (1,  Single  <$> arbitrary) ]
+      gen n = frequency
+        [  (2,  Text    <$> arbitrary)
+        ,  (1,  Single  <$> arbitrary)
+        ,  (4,  Tag     <$> arbitrary  <*> gen (n-1))
+        ,  (3,  Join    <$> gen (n-1)  <*> gen (n-1)) ]
+\end{code} %$
+
+\begin{code}
+type BoundedGen a = Int -> Gen a
+
+class BoundedArb (a :: *) where
+  boundedArb :: BoundedGen a
+\end{code}
+
+\begin{code}
 data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
 \end{code}
+
 \begin{code}
@@
 \end{code}
+
 \begin{code}
@@
   where alg (Term f) = alg f
+\end{code}
+
+\begin{code}
+data Con_Text   r = Con_Text String
+data Con_Sing   r = Con_Sing String
+data Con_Tag    r = Con_Tag  String r
+data Con_infix  r = Con_infix r r
+\end{code}
+
+\begin{code}
+data Fun_br     r = Fun_br
+data Fun_bold   r = Fun_bold r
+data Fun_infix  r = Fun_infix r r
+data Fun_list   r = Fun_list [r]
+\end{code}
+
+\begin{code}
+data Pat_simplify_1 r = Pat_simplify_1 String String
+data Pat_simplify_2 r = Pat_simplify_2 String r r
+\end{code}
+
+\begin{code}
+instance Algebra Con_Text Html where
+  alg (Con_Text x) = Text x
+
+instance Algebra Con_Sing Html where
+  alg (Con_Sing x) = Sing x
+
+instance Algebra Con_Tag Html where
+  alg (Con_Tag t x) = Tag t x
+
+instance Algebra Con_infix Html where
+  alg (Con_infix x y) = x :+: y
+\end{code}
+
+\begin{code}
+instance Algebra Fun_br Html where
+  alg Fun_br = br
+
+instance Algebra Fun_bold Html where
+  alg (Fun_bold x) = bold x
+
+instance Algebra Fun_infix Html where
+  alg (Fun_infix x y) = x <+> y
+
+instance Algebra Fun_list Html where
+  alg (Fun_list xs) = list xs
+\end{code}
+
+\begin{code}
+instance Algebra Pat_simplify_1 Html where
+  alg (Pat_simplify_1 t1 t2) = Text t1 :+: Text t2
+
+instance Algebra Pat_simplify_2 Html where
+  alg (Pat_simplify_1 t x y) = (Text t :+: x) :+: y
+\end{code}
+
+\begin{code}
+instance BoundedArb1 Con_Text where
+  liftBounded gen lvl = Con_Text <$> arbitrary
+
+instance BoundedArb1 Con_Sing where
+  liftBounded gen lvl = Con_Sing <$> arbitrary
+
+instance BoundedArb1 Con_Tag where
+  liftBounded gen lvl
+    = Con_Tag <$> arbitrary <*> gen (lvl - 1)
+
+instance BoundedArb1 Con_infix where
+  liftBounded gen lvl
+    = Con_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+\end{code} %$
+
+\begin{code}
+instance BoundedArb1 Fun_br where
+  liftBounded gen lvl = pure Fun_br
+
+instance BoundedArb1 Fun_bold where
+  liftBounded gen lvl = Fun_bold <$> gen (lvl - 1)
+
+instance BoundedArb1 Fun_infix where
+  liftBounded gen lvl
+    = Fun_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+
+instance BoundedArb1 Fun_list where
+  liftBounded gen lvl
+    = Fun_list <$> liftGen (gen (lvl - 1))
+\end{code} %$
+
+\begin{code}
+instance BoundedArb1 Pat_simplify_1 where
+  liftBounded gen lvl
+    = Pat_simplify_1 <$> arbitrary <*> arbitrary
+
+instance BoundedArb1 Pat_simplify_2 where
+  liftBounded gen lvl
+    = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
 \end{code}
@@
     , (natVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
-\end{code}
-
-% Depth bounded generators
-\begin{code}
-type BoundedGen a = Int -> Gen a
-
-class BoundedArb (a :: *) where
-  boundedArb :: BoundedGen a
 \end{code}
@@
   oplus  Con   "Join"         otimes 2
-  oplus  Fun'  "hr"        otimes 4
-  oplus  Fun   "div"       otimes 9
-  oplus  Pat   "simpl"  1  otimes 3
-  oplus  Pat   "simpl"  2  otimes 5
+  oplus  Fun'  "br"           otimes 4
+  oplus  Fun   "<+>"          otimes 9
+  oplus  Pat   "simplify"  1  otimes 3
+  oplus  Pat   "simplify"  2  otimes 5
 \end{code}

### 2019-03-27 papers/haskell19/2-sources.tex
Reorganized sections and code

--- 2019-03-26 papers/haskell19/introduction.tex
@@
-\section{Introduction}
-\label{sec:introduction}
+\section{Sources of Constructions}
+\label{sec:sources}
 
@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-
 \begin{code}
@@
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+\begin{code}
+class Arbitrary (a :: *) where
+  arbitrary :: Gen a
+\end{code}
+
+\begin{code}
+instance Arbitrary Html where
+  arbitrary = sized gen
+    where
+      gen 0 = frequency
+        [  (2,  Text    <$> arbitrary)
+        ,  (1,  Single  <$> arbitrary) ]
+      gen n = frequency
+        [  (2,  Text    <$> arbitrary)
+        ,  (1,  Single  <$> arbitrary)
+        ,  (4,  Tag     <$> arbitrary  <*> gen (n-1))
+        ,  (3,  Join    <$> gen (n-1)  <*> gen (n-1)) ]
+\end{code} %$

### 2019-03-27 papers/haskell19/3-representation.tex
Reorganized sections and code

--- 2019-03-20 papers/haskell19/representation.tex
@@
 data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
-\end{code}
-\begin{code}
-data ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
-\end{code}
-\begin{code}
-data Term (f :: * -> *) a = Term (f a)
 \end{code}
@@
          alg (InR g) = alg g
-
-instance Algebra f a => Algebra (f otimes n) a
-  where alg (Freq f) = alg f
-
-instance Algebra f a => Algebra (Term f) a
-  where alg (Term f) = alg f
 \end{code}
@@
 \begin{code}
-instance BoundedArb1 f => BoundedArb1 (f otimes n) where
-  liftBounded gen lvl = Freq <$> liftBounded gen lvl
-
-instance BoundedArb1 f => BoundedArb1 (Term f) where
-  liftBounded gen lvl = Term <$> liftBounded gen lvl
+data Con_Text   r = Con_Text String
+data Con_Sing   r = Con_Sing String
+data Con_Tag    r = Con_Tag  String r
+data Con_infix  r = Con_infix r r
 \end{code}
@@
 \begin{code}
-type family FreqOf (f :: * -> *) :: Nat where
-  FreqOf (f oplus g)   = FreqOf f + FreqOf g
-  FreqOf (f otimes n)  = n * FreqOf f
-  FreqOf (Term f)      = FreqOf f
-  FreqOf _             = 1
+data Fun_br     r = Fun_br
+data Fun_bold   r = Fun_bold r
+data Fun_infix  r = Fun_infix r r
+data Fun_list   r = Fun_list [r]
 \end{code}
@@
 \begin{code}
-type family FreqOf' (f :: * -> *) :: Nat where
-  FreqOf' (f oplus g)   = FreqOf' f  +  FreqOf' g
-  FreqOf' (f otimes n)  = n  *  FreqOf' f
-  FreqOf' (Term f)      = FreqOf f
-  FreqOf' _             = 0
+data Pat_simplify_1 r = Pat_simplify_1 String String
+data Pat_simplify_2 r = Pat_simplify_2 String r r
 \end{code}
@@
 \begin{code}
-instance (BoundedArb1 f, BoundedArb1 g)
-  => BoundedArb1 (f oplus g) where
-  liftBounded gen lvl =
-    if lvl > 0
-    then frequency
-    [ (natVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
-    , (natVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
-    else frequency
-    [ (natVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
-    , (natVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
-\end{code}
+instance Algebra Con_Text Html where
+  alg (Con_Text x) = Text x
 
-% Depth bounded generators
-\begin{code}
-type BoundedGen a = Int -> Gen a
+instance Algebra Con_Sing Html where
+  alg (Con_Sing x) = Sing x
 
-class BoundedArb (a :: *) where
-  boundedArb :: BoundedGen a
+instance Algebra Con_Tag Html where
+  alg (Con_Tag t x) = Tag t x
+
+instance Algebra Con_infix Html where
+  alg (Con_infix x y) = x :+: y
 \end{code}
@@
 \begin{code}
-class BoundedArb1 (f :: * -> *) where
-  liftBounded :: BoundedGen a -> BoundedGen (f a)
+instance Algebra Fun_br Html where
+  alg Fun_br = br
+
+instance Algebra Fun_bold Html where
+  alg (Fun_bold x) = bold x
+
+instance Algebra Fun_infix Html where
+  alg (Fun_infix x y) = x <+> y
+
+instance Algebra Fun_list Html where
+  alg (Fun_list xs) = list xs
 \end{code}
@@
 \begin{code}
-instance BoundedArb1 f => BoundedArb (Fix f) where
-  boundedArb lvl = Fix <$> liftBounded boundedArb lvl
-\end{code} %$
+instance Algebra Pat_simplify_1 Html where
+  alg (Pat_simplify_1 t1 t2) = Text t1 :+: Text t2
 
-\begin{code}
-genFix  ::  forall f dot BoundedArb1 f =>  BoundedGen (Fix f)
-genFix lvl = boundedArb at(Fix f) lvl
+instance Algebra Pat_simplify_2 Html where
+  alg (Pat_simplify_1 t x y) = (Text t :+: x) :+: y
 \end{code}
-
-\begin{code}
-genEval :: forall f a dot (BoundedArb1 f, Algebra f a) =>  BoundedGen a
-genEval lvl = eval <$> genFix at f lvl
-\end{code} %$
 
@@
 type family Pat (p :: Symbol) (n :: Nat) :: * -> *
-
-type Con' a    = Term  (Con a)
-type Fun' a    = Term  (Fun a)
-type Pat' a n  = Term  (Pat a n)
 \end{code}
-
-\begin{code}
-type HtmlSpec
-  =      Con'  "Text"
-  oplus  Con   "Join"      otimes 2
-  oplus  Fun'  "hr"        otimes 4
-  oplus  Fun   "div"       otimes 9
-  oplus  Pat   "simpl"  1  otimes 3
-  oplus  Pat   "simpl"  2  otimes 5
-\end{code}

### 2019-03-27 papers/haskell19/4-generators.tex
Reorganized sections and code

--- 2019-03-20 papers/haskell19/representation.tex
@@
-\section{Unified Representation of Constructions}
-\label{sec:representation}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+\section{Generating Composable Data via Type Specfications}
+\label{sec:generators}
 
 \begin{code}
-data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
+type BoundedGen a = Int -> Gen a
+
+class BoundedArb (a :: *) where
+  boundedArb :: BoundedGen a
 \end{code}
+
+\begin{code}
+class Arbitrary1 (f :: * -> *) where
+  liftGen :: Gen a -> Gen (f a)
+\end{code}
+
+\begin{code}
+class BoundedArb1 (f :: * -> *) where
+  liftBounded :: BoundedGen a -> BoundedGen (f a)
+\end{code}
+
 \begin{code}
@@
 \begin{code}
-  data Fix (f :: * -> *) = Fix { unFix :: (f (Fix f)) }
+instance Algebra f a => Algebra (f otimes n) a
+  where alg (Freq f) = alg f
 \end{code}
@@
 \begin{code}
-class Functor f => Algebra f a | f -> a where
-  alg :: f a -> a
-\end{code}
-
-% cata :: Functor f => (f a -> a) -> Fix f -> a
-% cata f = f . fmap (cata f) . unFix
-
-\begin{code}
-eval :: Algebra f a => Fix f -> a
-eval = alg . fmap eval . unFix
-\end{code}
-
-\begin{code}
-instance (Algebra f a, Algebra g a) => Algebra (f oplus g) a
-  where  alg (InL f) = alg f
-         alg (InR g) = alg g
-
-instance Algebra f a => Algebra (f otimes n) a
-  where alg (Freq f) = alg f
-
 instance Algebra f a => Algebra (Term f) a
@@
     then frequency
-    [ (natVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
-    , (natVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
+    [ (freqVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
+    , (freqVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
     else frequency
-    [ (natVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
-    , (natVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
-\end{code}
-
-% Depth bounded generators
-\begin{code}
-type BoundedGen a = Int -> Gen a
-
-class BoundedArb (a :: *) where
-  boundedArb :: BoundedGen a
+    [ (freqVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
+    , (freqVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
 \end{code}
@@
 \begin{code}
-class BoundedArb1 (f :: * -> *) where
-  liftBounded :: BoundedGen a -> BoundedGen (f a)
+instance BoundedArb1 Con_Text where
+  liftBounded gen lvl = Con_Text <$> arbitrary
+
+instance BoundedArb1 Con_Sing where
+  liftBounded gen lvl = Con_Sing <$> arbitrary
+
+instance BoundedArb1 Con_Tag where
+  liftBounded gen lvl
+    = Con_Tag <$> arbitrary <*> gen (lvl - 1)
+
+instance BoundedArb1 Con_infix where
+  liftBounded gen lvl
+    = Con_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+\end{code} %$
+
+\begin{code}
+instance BoundedArb1 Fun_br where
+  liftBounded gen lvl = pure Fun_br
+
+instance BoundedArb1 Fun_bold where
+  liftBounded gen lvl = Fun_bold <$> gen (lvl - 1)
+
+instance BoundedArb1 Fun_infix where
+  liftBounded gen lvl
+    = Fun_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+
+instance BoundedArb1 Fun_list where
+  liftBounded gen lvl
+    = Fun_list <$> liftGen (gen (lvl - 1))
+\end{code} %$
+
+\begin{code}
+instance BoundedArb1 Pat_simplify_1 where
+  liftBounded gen lvl
+    = Pat_simplify_1 <$> arbitrary <*> arbitrary
+
+instance BoundedArb1 Pat_simplify_2 where
+  liftBounded gen lvl
+    = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
 \end{code}
@@
 \begin{code}
-data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
-\end{code}
-
-\begin{code}
-type family (t :: * -> *) apply (a :: *) :: * -> *  where
-  (Some t)      apply  a  = t a
-  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
-  (f otimes n)  apply  a  = (f apply a) otimes n
-  (Term t)      apply  a  = Term (t apply a)
-  t             apply  a  = t
-\end{code}
-
-\begin{code}
-type family Rep (t :: Symbol) :: * -> *
-
-type family Con (c :: Symbol) :: * -> *
-type family Fun (f :: Symbol) :: * -> *
-type family Pat (p :: Symbol) (n :: Nat) :: * -> *
-
 type Con' a    = Term  (Con a)

### 2019-03-27 papers/haskell19/5-casestudies.tex
Reorganized sections and code

--- 2019-01-20 paper/main.tex
@@
-\documentclass[conference, fleqn]{IEEEtran}
-
-%% ODER: format ==         = "\mathrel{==}"
-%% ODER: format /=         = "\neq "
-%
-%
-\makeatletter
-\@ifundefined{lhs2tex.lhs2tex.sty.read}%
-  {\@namedef{lhs2tex.lhs2tex.sty.read}{}%
-   \newcommand\SkipToFmtEnd{}%
-   \newcommand\EndFmtInput{}%
-   \long\def\SkipToFmtEnd#1\EndFmtInput{}%
-  }\SkipToFmtEnd
-
-\newcommand\ReadOnlyOnce[1]{\@ifundefined{#1}{\@namedef{#1}{}}\SkipToFmtEnd}
-\usepackage{amstext}
-\usepackage{amssymb}
-\usepackage{stmaryrd}
-\DeclareFontFamily{OT1}{cmtex}{}
-\DeclareFontShape{OT1}{cmtex}{m}{n}
-  {<5><6><7><8>cmtex8
-   <9>cmtex9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmtex10}{}
-\DeclareFontShape{OT1}{cmtex}{m}{it}
-  {<-> ssub * cmtt/m/it}{}
-\newcommand{\texfamily}{\fontfamily{cmtex}\selectfont}
-\DeclareFontShape{OT1}{cmtt}{bx}{n}
-  {<5><6><7><8>cmtt8
-   <9>cmbtt9
-   <10><10.95><12><14.4><17.28><20.74><24.88>cmbtt10}{}
-\DeclareFontShape{OT1}{cmtex}{bx}{n}
-  {<-> ssub * cmtt/bx/n}{}
-\newcommand{\tex}[1]{\text{\texfamily#1}}	% NEU
-
-\newcommand{\Sp}{\hskip.33334em\relax}
-
-\newcommand{\Conid}[1]{\mathit{#1}}
-\newcommand{\Varid}[1]{\mathit{#1}}
-\newcommand{\anonymous}{\kern0.06em \vbox{\hrule\@width.5em}}
-\newcommand{\plus}{\mathbin{+\!\!\!+}}
-\newcommand{\bind}{\mathbin{>\!\!\!>\mkern-6.7mu=}}
-\newcommand{\rbind}{\mathbin{=\mkern-6.7mu<\!\!\!<}}% suggested by Neil Mitchell
-\newcommand{\sequ}{\mathbin{>\!\!\!>}}
-\renewcommand{\leq}{\leqslant}
-\renewcommand{\geq}{\geqslant}
-\usepackage{polytable}
-
-%mathindent has to be defined
-\@ifundefined{mathindent}%
-  {\newdimen\mathindent\mathindent\leftmargini}%
-  {}%
-
-\def\resethooks{%
-  \global\let\SaveRestoreHook\empty
-  \global\let\ColumnHook\empty}
-\newcommand*{\savecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\savecolumns[#1]}}
-\newcommand*{\restorecolumns}[1][default]%
-  {\g@addto@macro\SaveRestoreHook{\restorecolumns[#1]}}
-\newcommand*{\aligncolumn}[2]%
-  {\g@addto@macro\ColumnHook{\column{#1}{#2}}}
-
-\resethooks
-
-\newcommand{\onelinecommentchars}{\quad-{}- }
-\newcommand{\commentbeginchars}{\enskip\{-}
-\newcommand{\commentendchars}{-\}\enskip}
-
-\newcommand{\visiblecomments}{%
-  \let\onelinecomment=\onelinecommentchars
-  \let\commentbegin=\commentbeginchars
-  \let\commentend=\commentendchars}
-
-\newcommand{\invisiblecomments}{%
-  \let\onelinecomment=\empty
-  \let\commentbegin=\empty
-  \let\commentend=\empty}
-
-\visiblecomments
-
-\newlength{\blanklineskip}
-\setlength{\blanklineskip}{0.66084ex}
-
-\newcommand{\hsindent}[1]{\quad}% default is fixed indentation
-\let\hspre\empty
-\let\hspost\empty
-\newcommand{\NB}{\textbf{NB}}
-\newcommand{\Todo}[1]{$\langle$\textbf{To do:}~#1$\rangle$}
-
-\EndFmtInput
-\makeatother
-%
-%
-%
-%
-%
-%
-% This package provides two environments suitable to take the place
-% of hscode, called "plainhscode" and "arrayhscode". 
-%
-% The plain environment surrounds each code block by vertical space,
-% and it uses \abovedisplayskip and \belowdisplayskip to get spacing
-% similar to formulas. Note that if these dimensions are changed,
-% the spacing around displayed math formulas changes as well.
-% All code is indented using \leftskip.
-%
-% Changed 19.08.2004 to reflect changes in colorcode. Should work with
-% CodeGroup.sty.
-%
-\ReadOnlyOnce{polycode.fmt}%
-\makeatletter
-
-\newcommand{\hsnewpar}[1]%
-  {{\parskip=0pt\parindent=0pt\par\vskip #1\noindent}}
-
-% can be used, for instance, to redefine the code size, by setting the
-% command to \small or something alike
-\newcommand{\hscodestyle}{}
-
-% The command \sethscode can be used to switch the code formatting
-% behaviour by mapping the hscode environment in the subst directive
-% to a new LaTeX environment.
-
-\newcommand{\sethscode}[1]%
-  {\expandafter\let\expandafter\hscode\csname #1\endcsname
-   \expandafter\let\expandafter\endhscode\csname end#1\endcsname}
-
-% "compatibility" mode restores the non-polycode.fmt layout.
-
-\newenvironment{compathscode}%
-  {\par\noindent
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed\)%
-   \par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\compaths}{\sethscode{compathscode}}
-
-% "plain" mode is the proposed default.
-% It should now work with \centering.
-% This required some changes. The old version
-% is still available for reference as oldplainhscode.
-
-\newenvironment{plainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\hspre\(\let\hspost\)%
-   \pboxed}%
-  {\endpboxed%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newenvironment{oldplainhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-% Here, we make plainhscode the default environment.
-
-\newcommand{\plainhs}{\sethscode{plainhscode}}
-\newcommand{\oldplainhs}{\sethscode{oldplainhscode}}
-\plainhs
-
-% The arrayhscode is like plain, but makes use of polytable's
-% parray environment which disallows page breaks in code blocks.
-
-\newenvironment{arrayhscode}%
-  {\hsnewpar\abovedisplayskip
-   \advance\leftskip\mathindent
-   \hscodestyle
-   \let\\=\@normalcr
-   \(\parray}%
-  {\endparray\)%
-   \hsnewpar\belowdisplayskip
-   \ignorespacesafterend}
-
-\newcommand{\arrayhs}{\sethscode{arrayhscode}}
-
-% The mathhscode environment also makes use of polytable's parray 
-% environment. It is supposed to be used only inside math mode 
-% (I used it to typeset the type rules in my thesis).
-
-\newenvironment{mathhscode}%
-  {\parray}{\endparray}
-
-\newcommand{\mathhs}{\sethscode{mathhscode}}
-
-% texths is similar to mathhs, but works in text mode.
-
-\newenvironment{texthscode}%
-  {\(\parray}{\endparray\)}
-
-\newcommand{\texths}{\sethscode{texthscode}}
-
-% The framed environment places code in a framed box.
-
-\def\codeframewidth{\arrayrulewidth}
-\RequirePackage{calc}
-
-\newenvironment{framedhscode}%
-  {\parskip=\abovedisplayskip\par\noindent
-   \hscodestyle
-   \arrayrulewidth=\codeframewidth
-   \tabular{@{}|p{\linewidth-2\arraycolsep-2\arrayrulewidth-2pt}|@{}}%
-   \hline\framedhslinecorrect\\{-1.5ex}%
-   \let\endoflinesave=\\
-   \let\\=\@normalcr
-   \(\pboxed}%
-  {\endpboxed\)%
-   \framedhslinecorrect\endoflinesave{.5ex}\hline
-   \endtabular
-   \parskip=\belowdisplayskip\par\noindent
-   \ignorespacesafterend}
-
-\newcommand{\framedhslinecorrect}[2]%
-  {#1[#2]}
-
-\newcommand{\framedhs}{\sethscode{framedhscode}}
-
-% The inlinehscode environment is an experimental environment
-% that can be used to typeset displayed code inline.
-
-\newenvironment{inlinehscode}%
-  {\(\def\column##1##2{}%
-   \let\>\undefined\let\<\undefined\let\\\undefined
-   \newcommand\>[1][]{}\newcommand\<[1][]{}\newcommand\\[1][]{}%
-   \def\fromto##1##2##3{##3}%
-   \def\nextline{}}{\) }%
-
-\newcommand{\inlinehs}{\sethscode{inlinehscode}}
-
-% The joincode environment is a separate environment that
-% can be used to surround and thereby connect multiple code
-% blocks.
-
-\newenvironment{joincode}%
-  {\let\orighscode=\hscode
-   \let\origendhscode=\endhscode
-   \def\endhscode{\def\hscode{\endgroup\def\@currenvir{hscode}\\}\begingroup}
-   %\let\SaveRestoreHook=\empty
-   %\let\ColumnHook=\empty
-   %\let\resethooks=\empty
-   \orighscode\def\hscode{\endgroup\def\@currenvir{hscode}}}%
-  {\origendhscode
-   \global\let\hscode=\orighscode
-   \global\let\endhscode=\origendhscode}%
-
-\makeatother
-\EndFmtInput
-%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% HRep
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Constructor names
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% gen
-
-% Local Variables:
-% TeX-mbigstarer: "main.lhs.tex"
-% TeX-command-default: "Make"
-% End:
-
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-
-\usepackage{todonotes}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{mathdots}
-\usepackage{yhmath}
-\usepackage{cancel}
-\usepackage{color}
-\usepackage{siunitx}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{xspace}
-\usepackage{relsize}
-\usetikzlibrary{fadings}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Macros
-\newcommand{\tocite}{\textbf{CITE}}
-\newcommand{\quickcheck}{\emph{QuickCheck}\xspace}
-\newcommand{\megadeth}{\emph{MegaDeTH}\xspace}
-\newcommand{\dragen}{\textbf{DRAGEN}\xspace}
-
-\newcommand{\hrep}[1]{\ensuremath{H\!Rep_{#1}}}
-\newcommand{\gen}[1]{\ensuremath{gen_{#1}}}
-\newcommand{\term}[1]{\ensuremath{#1^\bigstar}}
-\newcommand{\evalrep}[2]{ \ensuremath{\llbracket #1 \rrbracket_{#2}}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some parameters
-\setlength{\mathindent}{\parindent}
-
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Title
-\title{Deriving Random Generators via\\ Higher-Level Representations}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Authors
-\author{
-  \IEEEauthorblockN{
-    % 1\textsuperscript{st}
-    Agust\'in Mista
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    mista@chalmers.se
-  }
-\and
-\IEEEauthorblockN{
-  % 2\textsuperscript{nd}
-  Alejandro Russo
-  }
-  \IEEEauthorblockA{
-    % \textit{Department of Computer Science and Engineering} \\
-    \textit{Chalmers University of Technology}\\
-    Gothenburg, Sweden \\
-    russo@chalmers.se
-  }
-}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Abstract
-
-\begin{abstract}
-
-  Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-  tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-  leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-  wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-  mollis ultricies. Mauris mollis tincidunt felis. Aliquam feugiat tellus ut
-  neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus lacinia
-  purus, et dictum nunc justo sit amet elit.
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Keywords
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Sections
-
-\section{Introduction}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-We remark that, altough this work focuses on Haskell data types, this technique
-is general enough to be applied to most programming languages with some level of
-support for composite types.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data type by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract useful
-  structure information from the user source code.
-  %
-\end{itemize}
-\section{Random Data Generation in Haskell} \label{sec:randomtesting}
-
-In this section we briefly introduce the common approach for automatically
-deriving random data generators in Haskell using a type-driven approach, along
-with its main drawbacks.
-
-Haskell is a strongly typed programming language with a powerful type system.
-%
-It lets programmers encode a considerable amount of information about the
-structure of their systems using data types that can be checked at compilation
-time.
-%
-One of its key aspects is the support for Algebraic Data Types (ADTs).
-%
-Essentially, an ADTs is a composite type defined by combining other types in
-terms of \textbf{sums} (also known as \emph{variant types}) and
-\textbf{products} (or tuples) of other data types.
-%
-To exemplify this, consider the following type definition to encoding \ensuremath{\Conid{Html}}
-expressions:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\Conid{Html}{}\<[12]%
-\>[12]{}\mathrel{=}{}\<[16]%
-\>[16]{}\Conid{Text}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Sing}\;{}\<[22]%
-\>[22]{}\Conid{String}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Join}\;{}\<[22]%
-\>[22]{}\Conid{Html}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\\
-\>[12]{}\,\ \vert\;{}\<[16]%
-\>[16]{}\Conid{Tag}\;{}\<[22]%
-\>[22]{}\Conid{String}\;{}\<[30]%
-\>[30]{}\Conid{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, we declare \ensuremath{\Conid{Html}} as the sum of four possible
-constructions: \ensuremath{\Conid{Text}} represents plain text values. \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Tag}} represent
-singular and paired Html tags, and \ensuremath{\Conid{Join}} concats two expression one after
-another.
-%
-We only encode a very small subset of the actual Html specification for
-illustrative reasons.
-%
-In Haskell, \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}} and \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}} are known as \emph{data
-  constructors} (or constructors for short) and are used to distiguish which
-variant of the data type we are constructing.
-%
-Each data constructor is defined as a product of zero or more types known as
-\emph{fields}.
-%
-For instance, \ensuremath{\Conid{Text}} has a field of type \ensuremath{\Conid{String}}, whereas \ensuremath{\Conid{Join}} two recusive
-fields of type \ensuremath{\Conid{Html}}.
-%
-In general, we will say that a data constructor with no recursive fields is
-\emph{terminal}, and \emph{non-terminal} or \emph{recursive} if it has at least
-one field of such nature.
-%
-With this reprentation, the expression ``\ensuremath{<\!\!html\!\!>\!\!hello\!\!<\!\!hr\!\!>\!\!bye\!\!<\!\!/html\!\!>}'' can be encoded as:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Conid{Tag}\;\text{\ttfamily \char34 html\char34}\;(\Conid{Join}\;(\Conid{Join}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}(\Conid{Text}\;\text{\ttfamily \char34 hello\char34})\;(\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}))\;(\Conid{Text}\;\text{\ttfamily \char34 bye\char34})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-Additionally, we can define a function \ensuremath{\Varid{render}} to serialize \ensuremath{\Conid{Html}} values as
-follows:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{render}\mathbin{::}\Conid{Html}\to \Conid{String}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Text}\;\Varid{t})\mathrel{=}\Varid{t}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Sing}\;\Varid{t})\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Varid{render}\;\Varid{x}\plus \Varid{render}\;\Varid{y}{}\<[E]%
-\\
-\>[B]{}\Varid{render}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\text{\ttfamily \char34 <\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}\plus \Varid{render}\;\Varid{x}\plus \text{\ttfamily \char34 </\char34}\plus \Varid{t}\plus \text{\ttfamily \char34 >\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In the previous definition, \ensuremath{\Varid{render}} is described using \emph{pattern matching}
-over each possible kind of value.
-%
-Using pattern matching we can define functions idiomatically by defining
-different function clauses for each input pattern we are interested on.
-%
-Patterns can be defined to match specific constructors, literal values or
-variable subexpressions (like \ensuremath{\Varid{t}}, \ensuremath{\Varid{x}} and \ensuremath{\Varid{y}} in the definition of \ensuremath{\Varid{render}}).
-%
-They can also be nested in order to match very specific patterns of values.
-
-\subsection*{\textbf{Type-Driven Generation of Random Values}}
-
-In order to generate random value of types involving user defined ADTs, most
-approaches require the user to provide a random data generator for each one of
-them.
-%
-This is a cumbersome and error prone task that closely follows the data type
-structure.
-%
-For instance, consider the following definition of a \quickcheck random
-generator for the type \ensuremath{\Conid{Html}}:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{4}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{6}{@{}>{\hspre}c<{\hspost}@{}}%
-\column{6E}{@{}l@{}}%
-\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{14}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{20}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}gen_{Html}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary})\mskip1.5mu]{}\<[E]%
-\\
-\>[B]{}\hsindent{4}{}\<[4]%
-\>[4]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{}[\mskip1.5mu {}\<[6E]%
-\>[9]{}(\mathrm{2},{}\<[14]%
-\>[14]{}\Conid{Text}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{1},{}\<[14]%
-\>[14]{}\Conid{Sing}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{3},{}\<[14]%
-\>[14]{}\Conid{Join}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;gen_{Html}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html}){}\<[E]%
-\\
-\>[4]{}\hsindent{2}{}\<[6]%
-\>[6]{},{}\<[6E]%
-\>[9]{}(\mathrm{4},{}\<[14]%
-\>[14]{}\Conid{Tag}{}\<[20]%
-\>[20]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;gen_{Html})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-%
-This random generator is defined using \quickcheck's combinator \ensuremath{\Varid{sized}} to
-paremetrize the generation process up to an external natural number known as the
-\emph{generation size}.
-%
-This parameter is chosen by the user, and we use to limit the maximum amount of
-recursive calls that this random generator can perform.
-%
-When called with a positive generation size, this generator can pick to generate
-among any \ensuremath{\Conid{Html}} data constructor of with a explicitly given frequency that can
-be chosen by the user (2, 1, 3 and 4 for \ensuremath{\Conid{Text}}, \ensuremath{\Conid{Sing}}, \ensuremath{\Conid{Join}} and \ensuremath{\Conid{Tag}}
-constructors, respectively).
-%
-When it picks to generate a \ensuremath{\Conid{Val}} or a \ensuremath{\Conid{Sing}} data constructor, it also
-generates a random \ensuremath{\Conid{String}} value using the standard overloaded generator
-\ensuremath{\Varid{arbitrary}} (\quickcheck provides standard random generators for most base types
-like \ensuremath{\Conid{String}}, \ensuremath{\Conid{Int}}, \ensuremath{\Conid{Bool}}, etc.). \footnote{%
-  The operators \ensuremath{\mathop{\mathsmaller{\langle \$ \rangle}}} and \ensuremath{\mathop{\mathsmaller{\langle \ast \rangle}}} are used in Haskell to combine values obtained
-  from effectful computations (like calling to a random generator) and they are
-  not particularly relevant for the point being made in this work.}
-%$
-On the other hand, when it picks to generate either an \ensuremath{\Conid{Join}} constructor, it
-also generates two independent random subexpression recursively, decreasing the
-generation size by a unit on each recursive invocation (\ensuremath{\Varid{smaller}\;gen_{Html}}).
-%
-The case of random generation of \ensuremath{\Conid{Tag}} constructors follows analogously.
-
-This random process keeps calling itself recursively until the generation size
-reaches zero, where the generator is constrained to pick only among terminal
-data constructors, being \ensuremath{\Conid{Text}} and \ensuremath{\Conid{Sing}} the only possible choices in our
-particular case.
-
-% Strictly decreasing the the generation size by one on each recursive call
-% results in a useful property over the generated data: every generated value has
-% at most |n| levels, where |n| is the generation size set by the user.
-%
-% This property enables us to model the generation process using the theory of
-% branching processes introduced by \tocite, and extended in this work as
-% described in Section \ref{sec:synthesis}.
-
-The previous definition is rather mechanical, except perhaps for the chosen
-generation frequencies.
-%
-In this light, it easy to extend this procedure to any data type defined in an
-algebraic fashion.
-%
-Fortunately, there exists different meta-programming tools to avoid the user
-from having to mechanically write random generators over and over again for each
-user-defined ADT.
-%
-The simplest tool for such purpose is \megadeth \tocite.
-%
-Given a target data type, it synthesizes a random generator for it that behaves
-similarly to the one presented above, where the generation frequencies are
-defined to be uniform across constructors.
-%
-However, picking among different data constructors with uniform frequency can
-lead to a generation process biased towards generating (in average) very small
-values, regardless of the generation sized set by the user \tocite.
-
-\dragen is meta-programming tool conceived to mitigate this problem.
-%
-Instead of setting a uniform generation probability of data constructors, this
-tool this tool uses the theory of branching processes to modelize and predict
-analitically the average distribution of data constructors generated on each
-random value.
-%
-This prediction mechanism is used to feedback a simulation-based optimization
-process that adjusts the generation frequency of each data constructor in order
-to obtain a particular distribution of values that can be specified by the user,
-providing this way a more flexible testing environment while still being mostly
-automated.
-
-Althogh both \megadeth and \dragen synthesize random generators that are
-theoretically capable to generate the whole space of values of the target data
-type. the limitations arise quickly when we consider that the underlying
-generation model is essentially the same: they pick a single data constructor
-and recusively generate each required subexpression independently.
-%
-In practice, this procedure is often too generic to generate random data with
-enough structural complexity to be used for testing purposes.
-
-% In this work we identify two sources of additional structure information which
-% are not considered by the aforementioned automatic derivation tools to obtain
-% better random data generators:
-
-In this work we identify two patological situations which are not properly
-handled by the aforementioned derivation tools:
-
-\begin{enumerate}
-\item The target code behaves differently over inputs matching specific patterns
-  of nested values.
-\item The target code encodes a significant portion of its structure on its
-  abstract interface.
-\end{enumerate}
-
-In Section \ref{sec:hrep} we show how this structural information can be used to
-synthesize richer random generators automatically.
-%
-We proceed to exemplify the previous points in detail.
-
-\subsection*{\textbf{Presence of Complex Pattern Matchings}}
-
-To exemplify the first problematic scenario, suppose we want to use randomly
-generated \ensuremath{\Conid{Htmls}}s to test a property comprising the following function:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Conid{Text}\;(\Varid{t1}\plus \Varid{t2}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y}))){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;\Varid{x}\;\Varid{y})\mathrel{=}\Conid{Join}\;(\Varid{simpl}\;\Varid{x})\;(\Varid{simpl}\;\Varid{y}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Tag}\;\Varid{t}\;\Varid{x})\mathrel{=}\Conid{Tag}\;\Varid{t}\;(\Varid{simpl}\;\Varid{x}){}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;\Varid{x}\mathrel{=}\Varid{x}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This function simplifies sequences of \ensuremath{\Conid{Text}} constructors into a single big
-\ensuremath{\Conid{Text}} constructor.
-%
-To do so, it has to pattern match against sequences of \ensuremath{\Conid{Text}} constructors
-combined by a \ensuremath{\Conid{Join}} constructor using nested patterns (see \ensuremath{\Varid{simpl}} fist and
-second clauses).
-%
-The remaining clauses are only meant to propagate this simplification within
-nested expressions.
-
-Ideally, we would like to test each clause of the function \ensuremath{\Varid{simpl}} approximately
-the same amount of time each.
-%
-However, each data constructor is generated independently when using either
-\megadeth or \dragen, thus the probability of generating a value satisfying a
-nested pattern decreases multiplicatively with the number of constructors we
-pattern against to simultaneously.
-%
-In our tests, we found that the first two clauses of \ensuremath{\Varid{simpl}} get exercised only
-approximately $1.5\%$ of the time when using \megadeth to derive a random
-generator for \ensuremath{\Conid{HTML}}.
-%
-On the other hand, the best result we could achieve with \dragen was only able
-to exercise the first and second clauses of \ensuremath{\Varid{simpl}} approximately $3\%$ and
-$6\%$ of the time, respectively.
-%
-With both derivation tools, the most of the time was spent testing the trivial
-clauses of our function, in view of they pattern match against simpler patterns
-of input values.
-
-% In this light, we found that using a random generator obtained with \megadeth to
-% test this function will only exercise the first two clauses of |simpl| about
-% only $1.5\%$ of the time each, relegating the rest of the generated test cases
-% to the remaining and somewhat less interesting clauses, where, remarkably, half
-% of the time is spent exercising the last trivial clause.
-% %
-% Furthermore, using \dragen to ...
-
-% This function behaves very much like an identity function, with the exception
-% that it fails for two very specific patterns of input values defined using
-% nested pattern matching.
-% %
-% This is, |foo| does not only matches against the root data constructor, but also
-% against the data constructors of its subexpressions and sub-subexpressions.
-
-% \quickcheck's default implementation for random |Int|s pick a number in the
-% interval $[-n, n]$ with uniform distribution.
-% %
-% Hence, in principle we need to recognize what is a suitable generation size,
-% which should be big enough for our generator be able to generate the numbers we
-% pattern match against to, otherwise we will not be able to generate any value to
-% match against |foo| clauses, leaving fragments of code completely untested.
-% %
-% Suppose then that we pick the a generation size $50$, i.e. the minimum size that
-% is big enough to produce an |Int| number equal to $50$.
-% %
-% Under this consideration, the probability of generating a value matching the
-% first clause of |foo| (and hence triggering the first error) results as follows:
-% %
-% \begin{align*}
-%   &P(match(foo\#1))\\
-%   &\quad = P(Add)       * P(Add) * 1 * (P(Val) * (1/100)) \\
-%   &\phantom{xxxxxxxxxx} * P(Add) * (P(Val) * (1/100)) * 1
-% \end{align*}
-
-% If we use \megadeth to automatically derive a random generator for |Exp|, we
-% obtain a uniform generation probability distribution over constructors, i.e.,
-% $P(|Val|) = P(|Add|) = P(|Mul|) = 1/3$.
-% %
-% In this setting, $P(match(foo\#1))$ results $1/2430000$, meaning that, in
-% average, we will need to generate over than two million test cases in order to
-% be able to test the first clause of |foo| only once.
-% %
-% This situation can be somewhat improved if we use \dragen to obtain a random
-% generator.
-% %
-% Using this tool, we can optimize the generation probabilities in order to
-% benefit the generation of some data constructors over the rest.
-% %
-% Considering that the first pattern matching of |foo| involves the data
-% constructor |Add| as the only recursive one, we can set an target generation
-% proportion of |Add| data constructors of, for instance, $20:1$ with respect to
-% the rest of the generated data constructors.
-% %
-% By doing so, the obtained distribution of values results such that
-% $P(match(foo\#1)) \approx 1/300000$.
-% %
-% Although this certainly improves the probability of generating a matching value,
-% this probability is not substantial enough to become practical.
-
-% Additionally, by favoring the generation probabilities towards the |Add|
-% constructor, we found that the probability of generating a value matching the
-% second clause of |foo| (which does not matches against it) also gets diminished
-% into an impractival value.
-
-Altough the previous example might seem rather simple, branching against
-specific patterns of the input data is a common task.
-%
-For instance, balancing a Red-Black tree requires to consider specific
-combinations of color, left and right subtrees and sub-subtrees in order to
-preserve the height invariant \tocite.
-%
-Moreover, Common Subexpression Elimination (CSE) is a compiler optimization that
-needs to consider very specific sequences of instructions that may be regrouped
-in a computationally more efficient way, to cite a few \tocite.
-
-\subsection*{\textbf{Data Invariants Encoded on Abstract Interfaces}}
-
-A common choice when implementing a data structure is to transfer the
-responsability of preserving its invariants to the functions which manipulates
-its values.
-%
-For this purspose, suppose we extend our \ensuremath{\Conid{Html}} data type with the following
-basic combinators:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{module}\;\Conid{M}\;\mathbf{where}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{div}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 div\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\mathbin{::}\Conid{Html}\to \Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{bold}\;\Varid{inner}\mathrel{=}\Conid{Tag}\;\text{\ttfamily \char34 b\char34}\;\Varid{inner}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathbin{::}\Conid{Html}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\Varid{hr}\mathrel{=}\Conid{Sing}\;\text{\ttfamily \char34 hr\char34}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-%
-% In the previous definition, the |Html| data type is defined using a single data
-% constructor that contains the textual representation of the Html code it
-% represents in plain text.
-%
-These functions encode additional information about the structure of our \ensuremath{\Conid{Html}}
-data type in the form of specific Html tags.
-
-Instead of including a new data constructor for each possible Html tag in out
-type definition, we defined a mininal representation and then extended it with
-an set of high level combinators.
-%
-% These combinators functions defined over |HTML| are the ones in charge of
-% transforming this plain text representation with the invariant that, given valid
-% |Html| parameters, they always return a valid |Html| value.
-%
-This programming pattern is often called a ``shallow embedding'', and can be
-found in a variety of Haskell libraries, being \emph{html} \tocite,
-\emph{svg-builder} \tocite some examples of this.
-
-As a consequence of this practice, type-driven derivation techniques often fail
-to synthesize useful random generators due to that most of the data type
-structure has been encoded into its abstract interface of combinators.
-%
-In our particular case, the chances of generating a \ensuremath{\Conid{Tag}} value representing a
-commonly used Html tag such as \ensuremath{\Varid{div}} are extremely low.
-
-% Additionally, if our test suite contains properties constrained by certain
-% preconditions, the lack of domain knowleadge may lead in an impractically high
-% discard ratio of randomly generated test cases.
-% %%
-% For instance, suposse we write a property to verify that |render| always outputs
-% a valid |Html| value:
-
-% \begin{code}
-% prop_render :: Html -> Bool
-% prop_render x = valid x ===> valid (parse (render x))
-% \end{code}
-
-% In the previous definition we state that |render| always returns a valid |Html|
-% value when it is parsed again (|parse (render x)|), provided that we are given a
-% valid |Html| as input (|valid x|).
-% %
-% In practice, not every string of characters constitutes a valid Html given that
-% some special characters need to be escaped (``<'', ``>'', ``&'', etc.).
-
-% %
-% While testing this property we rarely satisfy its precondition, in which case
-% \quickcheck discards the whole test, retrying with a diferent random input,
-% degradating this way the testing performance.
-
-% \begin{code}
-% text :: String -> Bool
-% text str = Text (addEscapes str)
-% \end{code}
-
-% We believe that, if certain patterns of values are relevant enough to appear in
-% the codebase being tested, a practical random testing methodology should be able
-% to produce values satisfying such patterns in a substantial proportion.
-
-So far we have introduced two testing scenarios where type-driven derivation
-approaches are unable to capture all the available structure information from
-the user codebase.
-%
-Fortunately, this information can be automatically exploited and used to
-generate interesting and more structured random values.
-
-The next section introduces a reprentation model that let us encode the
-structure information presented here into our automatically derived random
-generators in a modular and flexible way.
-\section{Extracting Structure} \label{sec:hrep}
-
-In this section we present a compositional representation of values to express
-the random generation of values following the internal structure of their data
-types along with the structure present on patterns matchings and abstract
-interfaces.
-
-The key idea of this work is to represent different structured constructions of
-data in a homogeneous way that we call a ``higher-level representation'' (\ensuremath{\Conid{HRep}}
-from now on).
-%
-Instead of generating each data constructor independently, a random generator
-derived from this representation might generate composite structured values on
-each random choice it performs.
-%
-% \footnote{The notion of higher-level comes from that the generation process is
-%   entirely determined by the type of the chosen type level representation,
-%   instead of by a concrete generator defined at the term level.}
-%
-For this purpose, we use a series of automatically derived data types, each one
-representing an atomic unit of information that can be randomly generated and
-then reflected back to the corresponding value of the original data type.
-%
-Later, the user can compose these atomic representations using the provided type
-level combinators in different ways into a ``generation specification'' that
-completely determines the generation process behavior.
-
-We wil reuse the previously defined data type \ensuremath{\Conid{Html}} and the functions defined
-in the previous section to explain the different concepts involved all across
-this section.
-
-\subsection*{\textbf{Representing Data Constructors}}
-
-We begin by introducing the simplest data type representation that we can
-extract from our codebase: the representation of single data constructor.
-%
-Each data constructor can be represented by an automatically derived data type
-consisting of a single constructor with the same fields as the original, except
-for the recursive ones that are abstracted away.
-%
-In this light, we represent each constructor of the data type \ensuremath{\Conid{Html}} as follows:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{Text}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Text}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Sing}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Sing}\;{}\<[29]%
-\>[29]{}\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Join}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Join}\;{}\<[29]%
-\>[29]{}\Varid{r}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{Tag}\;{}\<[16]%
-\>[16]{}\Varid{r}\mathrel{=}Con_{Tag}\;{}\<[29]%
-\>[29]{}\Conid{String}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Note that the previous definitions are type parametric over the type parameter
-\ensuremath{\Varid{r}}.
-%
-This allow us to replace \ensuremath{\Varid{r}} with any concrete data type, obtaining different
-possible values on each case.
-%
-For instance, the value (\ensuremath{Con_{Join}\;\mathrm{10}\;\mathrm{20}}) has type \ensuremath{\hrep{Join}\;\Conid{Int}}, while the
-value (\ensuremath{Con_{Join}\;\Conid{True}\;\Conid{False}}) has type \ensuremath{\hrep{Join}\;\Conid{Bool}}.
-
-In practice, this parametricity let us instatiate \ensuremath{\Varid{r}} with the type of the
-chosen generation specification (which might be composed of several \ensuremath{\Conid{HRep}}s),
-without having to modify anything in the underlying machinery.
-
-Having the \ensuremath{\Conid{HRep}} of each data constructor, we can define an evaluation relation
-($\evalrep{\_}{t} : \hrep{f} \rightarrow t$) that maps a value from each
-representation \ensuremath{\Varid{f}} back to the target data type \ensuremath{\Varid{t}}.
-%
-Then, we simply need to translate each constructor representation back into its
-corresponding one, translating the abstracted fields recursively:
-%
-\begin{alignat*}{4}
-  &\evalrep{\ensuremath{Con_{Text}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Text}}\ s \\
-  &\evalrep{\ensuremath{Con_{Sing}}\ s &&}{Html}
-    &&= \ensuremath{\Conid{Sing}}\ s \\
-  &\evalrep{\ensuremath{Con_{Join}}\ x\ y &&}{Html}
-    &&= \ensuremath{\Conid{Join}}\ \evalrep{x}{Html}\ \evalrep{y}{Html} \\
-  &\evalrep{\ensuremath{Con_{Tag}}\ s\ x &&}{Html}
-    &&= \ensuremath{\Conid{Tag}}\ s\ \evalrep{x}{Html}
-\end{alignat*}
-
-The missing piece is to automatically synthesize a random generators for each
-constructor representation.
-%
-For this purpose, it is important to consider that each constructor \ensuremath{\Conid{HRep}} has
-its recursive fields abstracted away with a type parameter that will be later
-instantiated with the generation specification type.
-%
-Given that this specification is unknown at the derivation time, we parametrize
-each \ensuremath{\Conid{HRep}} generator with a random generator \ensuremath{\gen{r}} that is used to generate
-random values for each recursive field:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{29}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{Text}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Text}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Sing}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Sing}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[3]{}\gen{Join}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Join}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{Tag}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Con_{Tag}{}\<[29]%
-\>[29]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Type Level Combinators}}
-
-The next step is to define a series of type level combinators to enable us
-combining the automatically derived \ensuremath{\Conid{HRep}}s in several ways.
-
-In first place, we define a type combinator (\ensuremath{\term{\_}}) to tag a \ensuremath{\Conid{HRep}} to be
-terminal, i.e., a representation that is allowed be generated when the
-generation size gets exausted:
-%
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\term{f})\;\Varid{a}\mathrel{=}\Conid{Term}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Additionally, we define a combinator ($\otimes$) to tag a \ensuremath{\Conid{HRep}} with an
-explicit generation frequency $n$:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\otimes\;\Varid{n})\;\Varid{a}\mathrel{=}\Conid{Freq}\;(\Varid{f}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The previous combinators only include information relevant to the generation
-process, in a sense that neither one adds new structure to the final
-representation.
-%
-In this light, they do not alter the evaluation semantics, and we translate them
-back to our target data type by evaluating the inner representation:
-%
-\begin{alignat*}{4}
-  &\evalrep{Term\ x &&: \ensuremath{\term{f}}    &&}{t} &&= \evalrep{x : f}{t} \\
-  &\evalrep{Freq\ x &&: f \otimes n &&}{t} &&= \evalrep{x : f}{t}
-\end{alignat*}
-
-In the previous equations we explicitly annotate (using a colon) the type of the
-evaluated term for clarity.
-%
-Later, to generate these combinators is enough to wrap a generated value from
-the inner representation with the apropriate tag:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{30}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{f}^\bigstar\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Term}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\\[\blanklineskip]%
-\>[B]{}\gen{f \otimes n}\;{}\<[16]%
-\>[16]{}\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Conid{Freq}{}\<[30]%
-\>[30]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Perhaps more interesting, we define a combinator $(\oplus)$ to compose two
-\ensuremath{\Conid{HRep}}s into a single one using a sum type to represent a random choice between
-them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;(\Varid{f}\;\oplus\;\Varid{g})\;\Varid{a}\mathrel{=}\Conid{L}\;(\Varid{f}\;\Varid{a})\mid \Conid{R}\;(\Varid{g}\;\Varid{a}){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-A composite representation built using $(\oplus)$ is transformed back into the
-target data type by pattern matching on the data constructor variant and
-evaluating the inner \ensuremath{\Conid{HRep}} accodingly:
-%
-\begin{alignat*}{3}
-  &\evalrep{L\ x &&: f \oplus g}{t} = \evalrep{x : f &&}{t} \\
-  &\evalrep{R\ x &&: f \oplus g}{t} = \evalrep{y : g &&}{t}
-\end{alignat*}
-
-Generating a composite \ensuremath{\Conid{HRep}} is slightly more complicated than before, as we
-need to perform a random choice based on the generation size and the given
-frequencies for each sub-represention:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{7}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{22}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{27}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{37}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{43}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{f \oplus g}\;\gen{r}{}\<[22]%
-\>[22]{}\mathrel{=}\Varid{sized}\;(\lambda \Varid{size}\to {}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{if}\;\Varid{size}\equiv \mathrm{0}{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{then}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq0}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]{}\<[E]%
-\\
-\>[3]{}\hsindent{2}{}\<[5]%
-\>[5]{}\mathbf{else}\;\Varid{frequency}{}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{}[\mskip1.5mu (\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!f,{}\<[24]%
-\>[24]{}\Conid{L}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{}){}\<[E]%
-\\
-\>[5]{}\hsindent{2}{}\<[7]%
-\>[7]{},(\Varid{freq}\;{}\<[17]%
-\>[17]{}\textit{@}\!g,{}\<[24]%
-\>[24]{}\Conid{R}{}\<[27]%
-\>[27]{}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{g}\;{}\<[37]%
-\>[37]{}\gen{r}{}\<[43]%
-\>[43]{})\mskip1.5mu]){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-In the previous definition, we reflect the type level frequencies of the types
-\ensuremath{\Varid{f}} and \ensuremath{\Varid{g}} (\ensuremath{\Varid{freq}\;\textit{@}\!f} and \ensuremath{\Varid{freq}\;\textit{@}\!g}) into term-level values.
-%
-This reflection defaults to $1$ if the frequency tag ($\otimes$) is not present.
-%
-Then we use these frequencies to generate each inner \ensuremath{\Conid{HRep}} in the apropriate
-proportion.
-%
-When the generation size gets exhausted, we reflect the terminal generation
-frequency of each inner \ensuremath{\Conid{HRep}} in the same way as before (\ensuremath{\Varid{freq0}\;\textit{@}\!f} and
-\ensuremath{\Varid{freq0}\;\textit{@}\!g}).
-%
-This time, however, we default the frequency reflection to $0$ for any inner
-that not tagged as terminal, avoiding to generate non-terminal constructions in
-the last step.
-
-Later, we can use these combinators to create a type synonym \ensuremath{\hrep{Html}} that
-specifies a generation schema equivalent to the one seen in the concrete random
-generator of type \ensuremath{\Conid{Html}} presented in Section \ref{sec:randomtesting}:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{24}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{Html}{}\<[16]%
-\>[16]{}\mathrel{=}{}\<[24]%
-\>[24]{}\term{\hrep{Text}}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\term{\hrep{Sing}}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Join}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{3}{}\<[E]%
-\\
-\>[16]{}\!\;\oplus\;{}\<[24]%
-\>[24]{}\hrep{Tag}\;{}\<[39]%
-\>[39]{}\otimes\;\mathrm{4}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-However, a value of type \ensuremath{\hrep{Html}} still has its recursive calls abstracted
-away---the type parameter \ensuremath{\Varid{r}} is implicit at the definition of \ensuremath{\hrep{Html}}.
-%
-We can think of it as a ``single layer'' of representation.
-%
-To make it able to represent recursive values we need to define a last type
-level combinator to ``tie the knot'':
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\mathbf{data}\;\Conid{Fix}\;\Varid{f}\mathrel{=}\Conid{Fix}\;(\Varid{f}\;(\Conid{Fix}\;\Varid{f})){}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-This datatype represents the \emph{fixed point} of a parametric data type \ensuremath{\Varid{f}},
-i.e., a data type where each recursive call gets instantiated with iself.
-%
-Then, to translate fixed points back to our target data type we simply need
-translate the inner representation:
-%
-\begin{align*}
-  \evalrep{Fix\ x : Fix\ f}{t} = \evalrep{x : f}{t}
-\end{align*}
-
-% Categorically, this generic transformation of fixed points is called a
-% \emph{catamorphism}, where our evaluation relation is known as its
-% \emph{algebra} \tocite.
-
-Unlike the other combinators, to define a random generator for the fixed point
-of a certain representation \ensuremath{\Varid{f}}, we do not to parametrize the generation of the
-recursive fields of \ensuremath{\Varid{f}} over an external generator \ensuremath{\gen{r}}.
-%
-% Instead, we simply pipe the generation to the inner |HRep|, calling itself on
-Instead, we replace it with our fixed point generator, calling itself
-recursively on any recursive field of \ensuremath{\Varid{f}} that might appear inside:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}genFix_{f}\mathrel{=}\Conid{Fix}\mathop{\mathsmaller{\langle \$ \rangle}}\gen{f}\;genFix_{f}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-This way we obtain a concrete recursive generator for each representation \ensuremath{\Varid{f}}
-that we define.
-%
-Then, we can define a random generator for our target data type simply by
-generating a random value of our chosen representation, and transforming it back
-to our target by the means of the evaluation relation:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}gen_{Html}\mathrel{=}\mathbf{do}\;{}\<[17]%
-\>[17]{}\Varid{x}\leftarrow genFix_{\hrep{Html}}{}\<[E]%
-\\
-\>[17]{}\Varid{return}\;\evalrep{x}{Html}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-So far we have introduced the machinery required to represent the random
-generation of a target data type considering only the structure encoded on its
-definition.
-%
-However, this approach can now be easily extended to encode different sources of
-structured information.
-%
-We proceed to introduce two extensions that help to address the problematic
-testing scenarios presented in the previous section.
-
-\subsection*{\textbf{Representing Pattern Matchings}}
-
-We can follow a similar reasoning as before to represent the pattern matching
-structure from a given function.
-%
-Consider the nested pattern matchings of function \ensuremath{\Varid{simpl}} defined in the
-previous section:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{36}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;(\Conid{Text}\;\Varid{t2})){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\\
-\>[B]{}\Varid{simpl}\;(\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\Varid{t1})\;\Varid{x})\;\Varid{y}){}\<[36]%
-\>[36]{}\mathrel{=}\mathbin{...}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-To represent these patterns, we derive a new data for each one of them, whose
-fields represent the pattern variables that occur inside, and abstracting away
-every pattern variable of type \ensuremath{\Conid{Html}}.
-%
-Concretely, we define the following data types:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{simpl\#1}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#1}\;\Conid{String}\;\Conid{String}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{simpl\#2}\;{}\<[18]%
-\>[18]{}\Varid{r}\mathrel{=}Pat_{simpl\#2}\;\Conid{String}\;\Varid{r}\;\Varid{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-The first pattern of \ensuremath{\Varid{simpl}} contains two pattern variables (\ensuremath{\Varid{t1}} and \ensuremath{\Varid{t2}}) of
-type \ensuremath{\Conid{String}} that are included as fields of \ensuremath{Pat_{simpl\#1}}.
-%
-Similarly, the second pattern of \ensuremath{\Varid{simpl}} contains a pattern variable \ensuremath{\Varid{t1}} of
-type \ensuremath{\Conid{String}} represented by the first field of \ensuremath{Pat_{simpl\#2}}, along with two
-pattern variables (\ensuremath{\Varid{x}} and \ensuremath{\Varid{y}}) of type \ensuremath{\Conid{Html}} that are abstracted in the second
-and third fields of \ensuremath{Pat_{simpl\#2}}.
-
-To transform these representations back to our target data type, we simply need
-to expand them it into the concrete value represented by the original pattern,
-evaluating its fields back to \ensuremath{\Conid{Html}} as well:
-%
-\begin{align*}
-  &\evalrep{\ensuremath{Pat_{simpl\#1}}\ t1\ t2}{Html} \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html})\;(\Conid{Text}\;\evalrep{t2}{Html})}\\
-  &\evalrep{\ensuremath{Pat_{simpl\#2}}\ t1\ x\ y}{Html} = \\
-  &\quad = \ensuremath{\Conid{Join}\;(\Conid{Join}\;(\Conid{Text}\;\evalrep{t1}{Html}))\;\evalrep{x}{Html})\;\evalrep{y}{Html})}
-\end{align*}
-
-This way, any value of type \ensuremath{\hrep{simpl\#1}} is guaranteed to satisfy the first
-pattern matching of \ensuremath{\Varid{simpl}}---the same property follows for \ensuremath{\hrep{simpl\#2}}.
-
-The random generation of pattern \ensuremath{\Conid{HRep}}s is defined in the same way as we did
-before for representations of data constructors:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{31}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\gen{simpl\#1}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#1}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}{}\<[31]%
-\>[31]{}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{arbitrary}{}\<[E]%
-\\
-\>[B]{}\gen{simpl\#2}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[E]%
-\\
-\>[B]{}\hsindent{3}{}\<[3]%
-\>[3]{}\mathrel{=}Pat_{simpl\#2}{}\<[16]%
-\>[16]{}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{arbitrary}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}\mathop{\mathsmaller{\langle \ast \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-Finally, we can join the pattern matching representation of each clause of
-\ensuremath{\Varid{simpl}} into a single one:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{17}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{39}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{simpl}{}\<[17]%
-\>[17]{}\mathrel{=}\hrep{simpl\#1}\;\!\!\oplus\!\!\;{}\<[39]%
-\>[39]{}\hrep{simpl\#2}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-\subsection*{\textbf{Representing Abstract Interfaces}}
-
-To introduce the higher level representation of module abstract interfaces,
-consider module \ensuremath{\Conid{M}} defininf \ensuremath{\Conid{Html}} combinators introduced in the previous
-section.
-
-% \begin{code}
-% module M where
-
-% ten :: Exp
-% ten = Val 10
-
-% square :: Exp -> Exp
-% square x = Mul x x
-
-% minus :: Exp -> Exp -> Exp
-% minus x y = Add x (Mul y (Val (-1)))
-% \end{code}
-
-We can represent each function of \ensuremath{\Conid{M}} that returns a value of type \ensuremath{\Conid{Html}}
-following the same idea as before, deriving a data type with a single data
-constructor for each one of them:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{data}\;\hrep{div}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{div}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{bold}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{bold}\;\Varid{r}{}\<[E]%
-\\
-\>[B]{}\mathbf{data}\;\hrep{hr}\;{}\<[21]%
-\>[21]{}\Varid{r}\mathrel{=}Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-In this case, each single constructor will have as fields the types of the
-inputs of the function that they represent.
-%
-As before, we abstract away any field representing and input of type \ensuremath{\Conid{Html}} with
-a type parameter \ensuremath{\Varid{r}}.
-
-To evaluate these representations, we simply use the values on its fields as
-input parameter of each original function, returning its result.
-%
-\begin{alignat*}{3}
-  &\evalrep{\ensuremath{Fun_{div}}\  &&x  &&}{Html} = \ensuremath{\Varid{div}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{bold}}\ &&x &&}{Html} = \ensuremath{\Varid{bold}\;\Varid{eval\char95 x\char95 html}} \\
-  &\evalrep{\ensuremath{Fun_{hr}}    &&  &&}{Html} = \ensuremath{\Varid{hr}}
-\end{alignat*}
-
-Note that, by doing this, the generation process inherits any patology that the
-functions we use to generate values might have.
-%
-For instance, if the function \ensuremath{\Varid{div}} would happen to be non-terminating for some
-inputs, our generation process could suffer from this as well.
-
-Furthermore, the generation procedure for abstract interface representations
-follows the same pattern as before:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{12}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{18}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[3]{}\gen{div}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{div}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{bold}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}Fun_{bold}\mathop{\mathsmaller{\langle \$ \rangle}}\Varid{smaller}\;\gen{r}{}\<[E]%
-\\
-\>[3]{}\gen{hr}\;{}\<[12]%
-\>[12]{}\gen{r}{}\<[18]%
-\>[18]{}\mathrel{=}\Varid{return}\;Fun_{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
- %$
-
-And we will also define a type synonym to join all the representations of the
-module \ensuremath{\Conid{M}} into a single one.
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{50}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;\hrep{M}{}\<[13]%
-\>[13]{}\mathrel{=}\hrep{div}\;\!\!\oplus\!\!\;{}\<[32]%
-\>[32]{}\hrep{bold}\;\!\!\oplus\!\!\;{}\<[50]%
-\>[50]{}\hrep{hr}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-Finally, we can put all the derived machinery together into a generation
-specification \ensuremath{Html_S}, assigning (possibly) different generation frequencies to
-each individual \ensuremath{\Conid{HRep}} we combine:
-
-\begin{hscode}\SaveRestoreHook
-\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{13}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{21}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{32}{@{}>{\hspre}l<{\hspost}@{}}%
-\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
-\>[B]{}\mathbf{type}\;Html_S{}\<[13]%
-\>[13]{}\mathrel{=}{}\<[21]%
-\>[21]{}\hrep{Html}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{4}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{simpl}\;{}\<[32]%
-\>[32]{}\otimes\;\mathrm{2}{}\<[E]%
-\\
-\>[13]{}\!\;\oplus\;{}\<[21]%
-\>[21]{}\hrep{M}{}\<[E]%
-\ColumnHook
-\end{hscode}\resethooks
-
-% This previous definition can be interpreted graphically as it is shown in the
-% Figure \ref{fig:hrep}, where curly arrows represent the structural information
-% extracted using meta-programming.
-
-% \begin{figure}[t]
-%   \centering
-%   \input{tikz/hrep.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-
-We want to remark that, for space reasons, we were only able to introduce the
-representation of a rather simple target data type.
-%
-In practice, this reasoning can be extended to mutually recursive and parametric
-types as well.
-
-Overall, this approach offers significant advantages over the usual type-driven
-derivation of random generators:
-%
-\begin{itemize}
-\item \textbf{Composability:} we can combine different atomic representations
-  using different structure information sources depending on what property or
-  sub-system we need to verify using randomly generated values.
-  %
-\item \textbf{Extensibility:} the developer can derive representations for new
-  sources of structure information and combine them with the existing ones
-  simply by adding them to the existing generation specification of the target
-  data type.
-  %
-\item \textbf{Predictability:} using branching processes theory, it is possible
-  to predict the average distribution of generated values in terms of number of
-  constructors.
-  %
-  This prediction is completely modular, and can be obtained for any composite
-  representation obtained using the automatically derived \ensuremath{\Conid{HRep}}s and the
-  provided combinators.
-\end{itemize}
-
-\todo[inline]{Maybe we can show an example random value from the representation
-  and its corresponding target value }
-
-% \begin{align*}
-%   \llbracket \_ \rrbracket_{target}\ :\ rep_{target}\ \rightarrow\ target
-% \end{align*}
-%
-% \begin{code}
-% class (rep down target) where
-%   step :: rep target -> target
-% \end{code}
-%
-% \begin{code}
-% data (Term f) a = TagTerm (f a)
-
-% \begin{code}
-% instance (HRep_Val down Exp) where
-%   step (Mk_Val n) = Val n
-
-% instance (HRep_Add down Exp) where
-%   step (Mk_Add x y) = Add x y
-
-% instance (HRep_Mul down Exp) where
-%   step (Mk_Mul x y) = Mul x y
-% \end{code}
-
-% \begin{code}
-% instance (f otimes n down Exp) where
-%   step (Freq f) = step f
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f otimes n) where
-%   liftArbitrary gen_HRep = Tag <$> gen_HRep
-% \end{code} %$
-
-% instance (Term f down Exp) where
-%   step (TagTerm f) = step f
-% \end{code}
-
-% \begin{code}
-% instance (HRep_ten down Exp) where
-%   step Mk_ten = ten
-
-% instance (HRep_square down Exp) where
-%   step (Mk_square x) = square x
-
-% instance (HRep_minus down Exp) where
-%   step (Mk_minus x y) = minus x y
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_ten where
-%   liftArbitrary gen_HRep = pure Mk_ten
-
-% instance Arbitrary1 HRep_square where
-%   liftArbitrary gen_HRep = Mk_square <$> gen_HRep
-
-% instance Arbitrary1 HRep_minus where
-%   liftArbitrary gen_HRep
-%     = Mk_minus <$> gen_HRep <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (HRep_foo_1 down Exp) where
-%   step (Mk_foo_1 x y)
-%     = Add (Add x (Val 50)) (Add (Val 25) y)
-
-% instance (HRep_foo_2 down Exp) where
-%   step (Mk_foo_2 x y)
-%     = Mul (Val 50) (Mul (Val x) y)
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 HRep_foo_1 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_1 <$> gen_HRep <*> gen_HRep
-
-% instance Arbitrary1 HRep_foo_2 where
-%   liftArbitrary gen_HRep
-%     = Mk_foo_2 <$> arbitrary <*> gen_HRep
-% \end{code} %$
-
-% \begin{code}
-% instance (f oplus g down Exp) where
-%   step (L f) = step f
-%   step (R g) = step g
-% \end{code}
-
-% \begin{code}
-% instance Arbitrary1 (f oplus g) where
-%   liftArbitrary gen_HRep
-%     = frequency
-%       [ (freq at_f, L <$> gen_HRep)
-%       , (freq at_g, R <$> gen_HRep) ]
-% \end{code} %$
-
-% \begin{figure}[b]
-%   \centering
-%   \input{tikz/val.tex}
-%   \caption{Higher level representation of the data type |Exp|, defined using
-%     structural information from the function |foo| and the abstract interface of
-%     the module |M|.}
-%   \label{fig:hrep}
-% \end{figure}
-\section{Random Generators Synthesis} \label{sec:synthesis}
+\section{Case Studies}
+\label{sec:casestudies}
 
@@
 Nam euismod tellus id erat.
-% %include sections/casestudies.tex
-\section{Related Work}
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
-\section{Final Remarks}
-
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Bibliography
-\bibliographystyle{IEEEtran}
-\bibliography{references.bib}
-
-\end{document}

### 2019-04-16 papers/haskell19/2-sources.tex
Code refactor

@@
-\section{Sources of Constructions}
+\section{Random Testing using Hints}
 \label{sec:sources}
@@
   |  Tag   String Html
-  |  Html  :+: Html
+  |  Join  Html Html
 \end{code}
 
-\begin{code}
-br :: Html
-br = Sing "br"
-
-bold :: Html -> Html
-bold = Tag "b"
-
-(<+>) :: Html -> Html -> Html
-(<+>) x y = x :+: br :+: y
-
-list :: [Html] -> Html
-list xs = Tag "ul" (foldl1 (:+:) (Tag "li" <$> xs))
-\end{code} %$
+A value of type |Html| represents either plain text strings, via the |Text| data
+constructor; single and paired HTML tags using the |Sing| and |Tag| data
+constructors, respectively; and concatenation of ...
 
 \begin{code}
-simplify :: Html -> Html
-simplify (Text t1 :+: Text t2)
-  = Text (t1 ++ t2)
-simplify ((Text t1 :+: x) :+: y)
-  = simplify (Text t1 :+: simplify (x :+: y))
-simplify (x :+: y)
-  = simplify x :+: simplify y
-simplify (Tag t x)
-  = Tag t (simplify x)
-simplify x = x
+class Arbitrary (a :: *) where
+  arbitrary :: Gen a
 \end{code}
@@
 \end{code} %$
+
+\begin{code}
+br :: Html
+br = Sing "br"
+
+bold :: Html -> Html
+bold = Tag "b"
+
+(<+>) :: Html -> Html -> Html
+(<+>) x y = Join x (Join br y)
+
+list :: [Html] -> Html
+list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))
+\end{code} %$
+
+\begin{code}
+simplify :: Html -> Html
+simplify (Join (Text t1) (Text t2))
+  = Text (t1 ++ t2)
+simplify (Join (Join (Text t1) x) y)
+  = simplify (Join (Text t1) (simplify (Join x y)))
+simplify (Join x y)
+  = Join (simplify x) (simplify y)
+simplify (Tag t x)
+  = Tag t (simplify x)
+simplify x = x
+\end{code}

### 2019-04-16 papers/haskell19/3-representation.tex
Code refactor

@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+% --------------------------------------
+% The basic idea
+%
+This section introduces a unified representation for the different constructions
+we might want to consider when generating random values.
+%
+The key idea of this work is to lift each different source of structural
+information to the type level.
+%
+In this light, we will define a set of different simple ``open'' data types,
+each one in charge of representing a single construction.
+%
+Later, these types can be
+\begin{inparaenum}[(i)]
+  \item combined in several ways depending on the desired shape of our test data,
+\item randomly generated (see Section \ref{sec:generators}), and finally,
+\item transformed back to corresponding values of the target data type automatically.
+\end{inparaenum}
+%
+This lifted representation of constructions can be automatically derived from
+our source code in compile-time, relieving programmers of the burden of writing
+random generators by hand.
+
+% --------------------------------------
+% Representing data constructors
+%
+\paragraph{Representing Data Constructors}
+%
+The simplest piece of meaningful%
+\footnote{We avoid considering $\bot$ as it does not contain any useful
+  information.}
+%
+information we can consider when generating values of an algebraic data type, is
+the one given by each one of its data constructors.
+%
+In this light, the representation of each individual data constructor of the
+target data type is simply given by a new data type containing a single
+constructor with the same fields as the corresponding one, except for the
+recursive fields which are abstracted away using a type parameter.
+
+Recalling our |Html| example, each one of it data constructors can be
+represented as follows:
 
 \begin{code}
-data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
-\end{code}
-
-\begin{code}
-  data Fix (f :: * -> *) = Fix { unFix :: (f (Fix f)) }
+data Con_Text   r = Con_Text String
+data Con_Sing   r = Con_Sing String
+data Con_Tag    r = Con_Tag  String r
+data Con_Join   r = Con_Join r r
 \end{code}
@@
   alg :: f a -> a
-\end{code}
-
-% cata :: Functor f => (f a -> a) -> Fix f -> a
-% cata f = f . fmap (cata f) . unFix
-
-\begin{code}
-eval :: Algebra f a => Fix f -> a
-eval = alg . fmap eval . unFix
-\end{code}
-
-\begin{code}
-instance (Algebra f a, Algebra g a) => Algebra (f oplus g) a
-  where  alg (InL f) = alg f
-         alg (InR g) = alg g
-\end{code}
-
-\begin{code}
-data Con_Text   r = Con_Text String
-data Con_Sing   r = Con_Sing String
-data Con_Tag    r = Con_Tag  String r
-data Con_infix  r = Con_infix r r
-\end{code}
-
-\begin{code}
-data Fun_br     r = Fun_br
-data Fun_bold   r = Fun_bold r
-data Fun_infix  r = Fun_infix r r
-data Fun_list   r = Fun_list [r]
-\end{code}
-
-\begin{code}
-data Pat_simplify_1 r = Pat_simplify_1 String String
-data Pat_simplify_2 r = Pat_simplify_2 String r r
 \end{code}
@@
 
-instance Algebra Con_infix Html where
-  alg (Con_infix x y) = x :+: y
+instance Algebra Con_Join Html where
+  alg (Con_Join x y) = Join x y
+\end{code}
+
+\begin{code}
+data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
+\end{code}
+
+\begin{code}
+instance (Algebra f a, Algebra g a) => Algebra (f oplus g) a
+  where  alg (InL f) = alg f
+         alg (InR g) = alg g
+\end{code}
+
+\begin{code}
+  data Fix (f :: * -> *) = Fix { unFix :: (f (Fix f)) }
+\end{code}
+
+\begin{code}
+eval :: Algebra f a => Fix f -> a
+eval = alg . fmap eval . unFix
+\end{code}
+
+\begin{code}
+data Fun_br     r = Fun_br
+data Fun_bold   r = Fun_bold r
+data Fun_cat    r = Fun_cat r r
+data Fun_list   r = Fun_list [r]
+\end{code}
+
+\begin{code}
+data Pat_simplify_1 r = Pat_simplify_1 String String
+data Pat_simplify_2 r = Pat_simplify_2 String r r
 \end{code}
@@
 
-instance Algebra Fun_infix Html where
-  alg (Fun_infix x y) = x <+> y
+instance Algebra Fun_cat Html where
+  alg (Fun_cat x y) = x <+> y
 
@@
 instance Algebra Pat_simplify_1 Html where
-  alg (Pat_simplify_1 t1 t2) = Text t1 :+: Text t2
+  alg (Pat_simplify_1 t1 t2) = Join (Text t1) (Text t2)
 
 instance Algebra Pat_simplify_2 Html where
-  alg (Pat_simplify_1 t x y) = (Text t :+: x) :+: y
+  alg (Pat_simplify_1 t x y) = Join (Join (Text t) x) y
 \end{code}
@@
 \begin{code}
-type family Rep (t :: Symbol) :: * -> *
-
 type family Con (c :: Symbol)             :: * -> *
@@
 \end{code}
+
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
+tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
+eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
+parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
+Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
+Nam vestibulum accumsan nisl\cite{godefroid2005dart}.

### 2019-04-16 papers/haskell19/4-generators.tex
Code refactor

@@
-\section{Generating Composable Data via Type Specfications}
+\section{Generating Composable Data via Type-level Specfications}
 \label{sec:generators}
@@
 \begin{code}
-type BoundedGen a = Int -> Gen a
+type BGen a = Int -> Gen a
 
-class BoundedArb (a :: *) where
-  boundedArb :: BoundedGen a
+class BArbitrary (a :: *) where
+  bounded :: BGen a
 \end{code}
@@
 \begin{code}
-class BoundedArb1 (f :: * -> *) where
-  liftBounded :: BoundedGen a -> BoundedGen (f a)
+class BArbitrary1 (f :: * -> *) where
+  liftBGen :: BGen a -> BGen (f a)
 \end{code}
@@
 \begin{code}
-instance Algebra f a => Algebra (f otimes n) a
-  where alg (Freq f) = alg f
+instance Algebra f a => Algebra (f otimes n) a where
+  alg (Freq f) = alg f
 \end{code}
@@
 \begin{code}
-instance Algebra f a => Algebra (Term f) a
-  where alg (Term f) = alg f
+instance Algebra f a => Algebra (Term f) a where
+  alg (Term f) = alg f
 \end{code}
@@
 \begin{code}
-instance BoundedArb1 f => BoundedArb1 (f otimes n) where
-  liftBounded gen lvl = Freq <$> liftBounded gen lvl
+instance BArbitrary1 f => BArbitrary1 (f otimes n) where
+  liftBGen gen lvl = Freq <$> liftBGen gen lvl
 
-instance BoundedArb1 f => BoundedArb1 (Term f) where
-  liftBounded gen lvl = Term <$> liftBounded gen lvl
+instance BArbitrary1 f => BArbitrary1 (Term f) where
+  liftBGen gen lvl = Term <$> liftBGen gen lvl
 \end{code}
@@
 \begin{code}
-instance (BoundedArb1 f, BoundedArb1 g)
-  => BoundedArb1 (f oplus g) where
-  liftBounded gen lvl =
+instance (BArbitrary1 f, BArbitrary1 g)
+  => BArbitrary1 (f oplus g) where
+  liftBGen gen lvl =
     if lvl > 0
     then frequency
-    [ (freqVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
-    , (freqVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
+      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen lvl)
+      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen lvl) ]
     else frequency
-    [ (freqVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
-    , (freqVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
+      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen lvl)
+      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen lvl) ]
 \end{code}
@@
 \begin{code}
-instance BoundedArb1 Con_Text where
-  liftBounded gen lvl = Con_Text <$> arbitrary
+instance BArbitrary1 Con_Text where
+  liftBGen gen lvl = Con_Text <$> arbitrary
 
-instance BoundedArb1 Con_Sing where
-  liftBounded gen lvl = Con_Sing <$> arbitrary
+instance BArbitrary1 Con_Sing where
+  liftBGen gen lvl = Con_Sing <$> arbitrary
 
-instance BoundedArb1 Con_Tag where
-  liftBounded gen lvl
-    = Con_Tag <$> arbitrary <*> gen (lvl - 1)
+instance BArbitrary1 Con_Tag where
+  liftBGen gen lvl = Con_Tag <$> arbitrary <*> gen (lvl - 1)
 
-instance BoundedArb1 Con_infix where
-  liftBounded gen lvl
-    = Con_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+instance BArbitrary1 Con_Join where
+  liftBGen gen lvl = Con_Join <$> gen (lvl - 1) <*> gen (lvl - 1)
 \end{code} %$
@@
 \begin{code}
-instance BoundedArb1 Fun_br where
-  liftBounded gen lvl = pure Fun_br
+instance BArbitrary1 Fun_br where
+  liftBGen gen lvl = pure Fun_br
 
-instance BoundedArb1 Fun_bold where
-  liftBounded gen lvl = Fun_bold <$> gen (lvl - 1)
+instance BArbitrary1 Fun_bold where
+  liftBGen gen lvl = Fun_bold <$> gen (lvl - 1)
 
-instance BoundedArb1 Fun_infix where
-  liftBounded gen lvl
-    = Fun_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
+instance BArbitrary1 Fun_cat where
+  liftBGen gen lvl = Fun_cat <$> gen (lvl - 1) <*> gen (lvl - 1)
 
-instance BoundedArb1 Fun_list where
-  liftBounded gen lvl
-    = Fun_list <$> liftGen (gen (lvl - 1))
+instance BArbitrary1 Fun_list where
+  liftBGen gen lvl = Fun_list <$> liftGen (gen (lvl - 1))
 \end{code} %$
@@
 \begin{code}
-instance BoundedArb1 Pat_simplify_1 where
-  liftBounded gen lvl
+instance BArbitrary1 Pat_simplify_1 where
+  liftBGen gen lvl
     = Pat_simplify_1 <$> arbitrary <*> arbitrary
 
-instance BoundedArb1 Pat_simplify_2 where
-  liftBounded gen lvl
+instance BArbitrary1 Pat_simplify_2 where
+  liftBGen gen lvl
     = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
@@
 \begin{code}
-instance BoundedArb1 f => BoundedArb (Fix f) where
-  boundedArb lvl = Fix <$> liftBounded boundedArb lvl
+instance BArbitrary1 f => BArbitrary (Fix f) where
+  bounded lvl = Fix <$> liftBGen bounded lvl
 \end{code} %$
@@
 \begin{code}
-genFix  ::  forall f dot BoundedArb1 f =>  BoundedGen (Fix f)
-genFix lvl = boundedArb at(Fix f) lvl
+genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)
+genFix lvl = bounded at(Fix f) lvl
 \end{code}
@@
 \begin{code}
-genEval :: forall f a dot (BoundedArb1 f, Algebra f a) =>  BoundedGen a
+genEval :: forall f a dot (BArbitrary1 f, Algebra f a) => BGen a
 genEval lvl = eval <$> genFix at f lvl

### 2019-04-16 papers/haskell19/3-representation.tex
Code refactor

@@
 \begin{code}
-instance (Algebra f a, Algebra g a) => Algebra (f oplus g) a
-  where  alg (InL f) = alg f
+instance (Algebra f a, Algebra g a)
+  => Algebra (f oplus g) a where
+  alg (InL f) = alg f
   alg (InR g) = alg g
@@
 type family Con (c :: Symbol)             :: * -> *
+type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
 type family Fun (f :: Symbol)             :: * -> *
-type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
+\end{code}
+
+\begin{code}
+type instance Con "Text"        = Con_Text
+type instance Pat "simplify" 1  = Pat_simplify_1
+type instance Fun "<+>"         = Fun_cat
 \end{code}

### 2019-04-16 papers/haskell19/3-representation.tex
Added file

--- 2019-04-16 papers/haskell19/3-representation.tex
@@
 
-\begin{code}
-data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
-\end{code}
-
-\begin{code}
-type family (t :: * -> *) apply (a :: *) :: * -> *  where
-  (Some t)      apply  a  = t a
-  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
-  (f otimes n)  apply  a  = (f apply a) otimes n
-  (Term t)      apply  a  = Term (t apply a)
-  t             apply  a  = t
-\end{code}
-
-\begin{code}
-type family Con (c :: Symbol)             :: * -> *
-type family Fun (f :: Symbol)             :: * -> *
-type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
-\end{code}
-
 Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor

### 2019-04-16 papers/haskell19/4-generators.tex
Code refactor
Added file

@@
-\section{Generating Composable Data via Type-level Specfications}
+\section{Generating Random Constructions}
 \label{sec:generators}
@@
 \end{code} %$
-
-\begin{code}
-type Con' a    = Term  (Con a)
-type Fun' a    = Term  (Fun a)
-type Pat' a n  = Term  (Pat a n)
-\end{code}
-
-\begin{code}
-type HtmlSpec
-  =      Con'  "Text"
-  oplus  Con   "Join"         otimes 2
-  oplus  Fun'  "br"           otimes 4
-  oplus  Fun   "<+>"          otimes 9
-  oplus  Pat   "simplify"  1  otimes 3
-  oplus  Pat   "simplify"  2  otimes 5
-\end{code}

### 2019-04-16 papers/haskell19/5-typelevel.tex
Added file

\section{Type-Level Generation Specifications}
\label{sec:typelevel}

\begin{code}
type family Con (c :: Symbol)             :: * -> *
type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
type family Fun (f :: Symbol)             :: * -> *
\end{code}

\begin{code}
data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
\end{code}

\begin{code}
type family (t :: * -> *) apply (a :: *) :: * -> *  where
  (Some t)      apply  a  = t a
  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
  (f otimes n)  apply  a  = (f apply a) otimes n
  (Term t)      apply  a  = Term (t apply a)
  t             apply  a  = t
\end{code}

\begin{code}
type instance Con "Text"        = Con_Text
type instance Pat "simplify" 1  = Pat_simplify_1
type instance Fun "<+>"         = Fun_cat
\end{code}

\begin{code}
type Con' a    = Term  (Con a)
type Fun' a    = Term  (Fun a)
type Pat' a n  = Term  (Pat a n)
\end{code}

\begin{code}
type Html_Spec
  =      Con'  "Text"
  oplus  Con   "Join"         otimes 2
  oplus  Fun'  "br"           otimes 4
  oplus  Fun   "<+>"          otimes 9
  oplus  Pat   "simplify"  1  otimes 3
  oplus  Pat   "simplify"  2  otimes 5
\end{code}

\begin{code}
instance Arbitrary Html where
  arbitrary = sized (genEval at Html_Spec)
\end{code}

### 2019-04-16 papers/haskell19/8-conclusions.tex
Added file

--- 2019-03-27 papers/haskell19/2-sources.tex
@@
-\section{Sources of Constructions}
-\label{sec:sources}
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 lacinia purus, et dictum nunc justo sit amet elit.
-
-\begin{code}
-data Html
-  =  Text  String
-  |  Sing  String
-  |  Tag   String Html
-  |  Html  :+: Html
-\end{code}
-
-\begin{code}
-br :: Html
-br = Sing "br"
-
-bold :: Html -> Html
-bold = Tag "b"
-
-(<+>) :: Html -> Html -> Html
-(<+>) x y = x :+: br :+: y
-
-list :: [Html] -> Html
-list xs = Tag "ul" (foldl1 (:+:) (Tag "li" <$> xs))
-\end{code} %$
-
-\begin{code}
-simplify :: Html -> Html
-simplify (Text t1 :+: Text t2)
-  = Text (t1 ++ t2)
-simplify ((Text t1 :+: x) :+: y)
-  = simplify (Text t1 :+: simplify (x :+: y))
-simplify (x :+: y)
-  = simplify x :+: simplify y
-simplify (Tag t x)
-  = Tag t (simplify x)
-simplify x = x
-\end{code}
-
-\begin{code}
-class Arbitrary (a :: *) where
-  arbitrary :: Gen a
-\end{code}
-
-\begin{code}
-instance Arbitrary Html where
-  arbitrary = sized gen
-    where
-      gen 0 = frequency
-        [  (2,  Text    <$> arbitrary)
-        ,  (1,  Single  <$> arbitrary) ]
-      gen n = frequency
-        [  (2,  Text    <$> arbitrary)
-        ,  (1,  Single  <$> arbitrary)
-        ,  (4,  Tag     <$> arbitrary  <*> gen (n-1))
-        ,  (3,  Join    <$> gen (n-1)  <*> gen (n-1)) ]
-\end{code} %$

### 2019-04-16 papers/haskell19/4-generators.tex
WIP

@@
 instance BArbitrary1 f => BArbitrary1 (f otimes n) where
-  liftBGen gen lvl = Freq <$> liftBGen gen lvl
+  liftBGen gen d = Freq <$> liftBGen gen d
 
 instance BArbitrary1 f => BArbitrary1 (Term f) where
-  liftBGen gen lvl = Term <$> liftBGen gen lvl
+  liftBGen gen d = Term <$> liftBGen gen d
 \end{code}
@@
   => BArbitrary1 (f oplus g) where
-  liftBGen gen lvl =
-    if lvl > 0
+  liftBGen gen d =
+    if d > 0
     then frequency
-      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen lvl)
-      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen lvl) ]
+      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen d)
+      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen d) ]
     else frequency
-      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen lvl)
-      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen lvl) ]
+      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen d)
+      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen d) ]
+\end{code}
+
+\begin{code}
+freqVal :: forall n dot KnownNat n => Int
 \end{code}
@@
 instance BArbitrary1 Con_Text where
-  liftBGen gen lvl = Con_Text <$> arbitrary
+  liftBGen gen d = Con_Text <$> arbitrary
 
 instance BArbitrary1 Con_Sing where
-  liftBGen gen lvl = Con_Sing <$> arbitrary
+  liftBGen gen d = Con_Sing <$> arbitrary
 
 instance BArbitrary1 Con_Tag where
-  liftBGen gen lvl = Con_Tag <$> arbitrary <*> gen (lvl - 1)
+  liftBGen gen d = Con_Tag <$> arbitrary <*> gen (d - 1)
 
 instance BArbitrary1 Con_Join where
-  liftBGen gen lvl = Con_Join <$> gen (lvl - 1) <*> gen (lvl - 1)
+  liftBGen gen d = Con_Join <$> gen (d - 1) <*> gen (d - 1)
 \end{code} %$
@@
 instance BArbitrary1 Fun_br where
-  liftBGen gen lvl = pure Fun_br
+  liftBGen gen d = pure Fun_br
 
 instance BArbitrary1 Fun_bold where
-  liftBGen gen lvl = Fun_bold <$> gen (lvl - 1)
+  liftBGen gen d = Fun_bold <$> gen (d - 1)
 
 instance BArbitrary1 Fun_cat where
-  liftBGen gen lvl = Fun_cat <$> gen (lvl - 1) <*> gen (lvl - 1)
+  liftBGen gen d = Fun_cat <$> gen (d - 1) <*> gen (d - 1)
 
 instance BArbitrary1 Fun_list where
-  liftBGen gen lvl = Fun_list <$> liftGen (gen (lvl - 1))
+  liftBGen gen d = Fun_list <$> liftGen (gen (d - 1))
 \end{code} %$
@@
 instance BArbitrary1 Pat_simplify_1 where
-  liftBGen gen lvl
+  liftBGen gen d
     = Pat_simplify_1 <$> arbitrary <*> arbitrary
@@
 instance BArbitrary1 Pat_simplify_2 where
-  liftBGen gen lvl
-    = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
+  liftBGen gen d
+    = Pat_simplify_2 <$> gen (d - 1) <*> gen (d - 1)
 \end{code}
@@
 instance BArbitrary1 f => BArbitrary (Fix f) where
-  bounded lvl = Fix <$> liftBGen bounded lvl
+  bounded d = Fix <$> liftBGen bounded d
 \end{code} %$
@@
 genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)
-genFix lvl = bounded at(Fix f) lvl
+genFix d = bounded at(Fix f) d
 \end{code}
@@
 genEval :: forall f a dot (BArbitrary1 f, Algebra f a) => BGen a
-genEval lvl = eval <$> genFix at f lvl
+genEval d = eval <$> genFix at f d
 \end{code} %$

### 2019-04-16 papers/haskell19/5-typelevel.tex
WIP

@@
 \label{sec:typelevel}
-
-\begin{code}
-type family Con (c :: Symbol)             :: * -> *
-type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
-type family Fun (f :: Symbol)             :: * -> *
-\end{code}
-
-\begin{code}
-data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
-\end{code}
-
-\begin{code}
-type family (t :: * -> *) apply (a :: *) :: * -> *  where
-  (Some t)      apply  a  = t a
-  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
-  (f otimes n)  apply  a  = (f apply a) otimes n
-  (Term t)      apply  a  = Term (t apply a)
-  t             apply  a  = t
-\end{code}
-
-\begin{code}
-type instance Con "Text"        = Con_Text
-type instance Pat "simplify" 1  = Pat_simplify_1
-type instance Fun "<+>"         = Fun_cat
-\end{code}
-
-\begin{code}
-type Con' a    = Term  (Con a)
-type Fun' a    = Term  (Fun a)
-type Pat' a n  = Term  (Pat a n)
-\end{code}
 
@@
 \begin{code}
+instance BArbitrary Html where
+  bounded = genEval at Html_Spec
+
 instance Arbitrary Html where
-  arbitrary = sized (genEval at Html_Spec)
+  arbitrary = sized bounded
 \end{code}
+
+\begin{code}
+type family Con (c :: Symbol)             :: * -> *
+type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
+type family Fun (f :: Symbol)             :: * -> *
+\end{code}
+
+\begin{code}
+type instance Con "Text"        = Con_Text
+type instance Pat "simplify" 1  = Pat_simplify_1
+type instance Fun "<+>"         = Fun_cat
+\end{code}
+
+\begin{code}
+type Con' a    = Term  (Con a)
+type Pat' a n  = Term  (Pat a n)
+type Fun' a    = Term  (Fun a)
+\end{code}
+
+\begin{code}
+data Tree a = Leaf a | Node (Tree a) (Tree a)
+\end{code}
+
+\begin{code}
+data Con_Leaf a r = Con_Leaf a
+data Con_Node a r = Con_Node r r
+\end{code}
+
+\begin{code}
+instance Algebra (Con_Leaf a) (Tree a) where ...
+instance Algebra (Con_Node a) (Tree a) where ...
+\end{code}
+
+\begin{code}
+instance Arbitrary a => BArbitrary1 (Con_Leaf a) where ...
+instance Arbitrary a => BArbitrary1 (Con_Node a) where ...
+\end{code}
+
+\begin{code}
+data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
+\end{code}
+
+\begin{code}
+type instance Con "Leaf" = Some Con_Leaf
+type instance Con "Node" = Some Con_Node
+\end{code}
+
+\begin{code}
+type Tree_Spec
+  =      Con "Leaf" otimes 2
+  oplus  Con "Node" otimes 3
+\end{code}
+
+\begin{code}
+type family (f :: * -> *) apply (a :: *) :: * -> *  where
+  (Some f)      apply  a  = f a
+  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
+  (f otimes n)  apply  a  = (f apply a) otimes n
+  (Term f)      apply  a  = Term (t apply a)
+  f             apply  a  = f
+\end{code}
+
+\begin{code}
+instance Arbitrary a => BArbitrary (Tree a) where
+  bounded = genEval at (Tree_Spec apply a)
+\end{code}

### 2019-04-17 papers/haskell19/2-sources.tex
WIP

@@
         [  (2,  Text    <$> arbitrary)
-        ,  (1,  Single  <$> arbitrary) ]
+        ,  (1,  Sing    <$> arbitrary) ]
       gen n = frequency
         [  (2,  Text    <$> arbitrary)
-        ,  (1,  Single  <$> arbitrary)
+        ,  (1,  Sing    <$> arbitrary)
         ,  (4,  Tag     <$> arbitrary  <*> gen (n-1))

### 2019-04-17 papers/haskell19/3-representation.tex
WIP

@@
 The key idea of this work is to lift each different source of structural
-information to the type level.
+information into the type level.
 %
-In this light, we will define a set of different simple ``open'' data types,
-each one in charge of representing a single construction.
+In this light, the ``shape'' of our random data is determined entirely by
+the type we use to represent it during the generation process.
 %
-Later, these types can be
+For this, we will define a set of different simple ``open'' data types,
+each one in charge of representing a single construction of our target data
+type.
+%
+These types can be later
 \begin{inparaenum}[(i)]
@@
 \item randomly generated (see Section \ref{sec:generators}), and finally,
-\item transformed back to corresponding values of the target data type automatically.
+\item transformed back into the corresponding values of the target data type
+  automatically.
 \end{inparaenum}
@@
 This lifted representation of constructions can be automatically derived from
-our source code in compile-time, relieving programmers of the burden of writing
-random generators by hand.
+our source code in compile-time, relieving programmers of the burden of manually
+implementing this required machinery.
 
@@
 %
-\paragraph{Representing Data Constructors}
+When generating values of algebraic data types, the simplest piece of meaningful
+information we ought to consider is the one given by each one of its data
+constructors.
 %
-The simplest piece of meaningful%
-\footnote{We avoid considering $\bot$ as it does not contain any useful
-  information.}
+\footnote{We explicitly avoid considering $\bot$ when generating random values.}
 %
-information we can consider when generating values of an algebraic data type, is
-the one given by each one of its data constructors.
+In this light, each individual data constructor of the target data type can be
+represented simply by a new data type with a single constructor.
 %
-In this light, the representation of each individual data constructor of the
-target data type is simply given by a new data type containing a single
-constructor with the same fields as the corresponding one, except for the
-recursive fields which are abstracted away using a type parameter.
-
-Recalling our |Html| example, each one of it data constructors can be
-represented as follows:
+Recalling our |Html| example, its data constructors can be represented as
+follows:
 
@@
 
+Where each representation has the same fields as its corresponding constructor,
+except for the recursive ones which are abstracted away using a type parameter
+|r|.
+%
+This parametricity lets us leave the type of each recursive sub-term unspecified
+until we have decided the overall shape of our random data.
+%
+Then, for instance, the value |Con_Tag "div" x :: Con_Tag y| represents the
+|Html| value |Tag "div" x|, for some sub-term |x :: y| that can be transformed
+back to |Html| as well.
+%
+It is easy to note how each one of these simple representations encode the
+minimal amount of information they need, leaving everything else unspecified.
+
+An important property of these parametric representations is that, in most
+cases, they form a functor over its type parameter, and thus we can use
+Haskell's |deriving| mechanism to obtain a suitable |Functor| instances for
+free:
+
+\begin{code}
+deriving instance Functor Con_Text
+deriving instance Functor Con_Sing
+deriving instance Functor Con_Tag
+deriving instance Functor Con_Join
+\end{code}
+
+% --------------------------------------
+% F-Algebras
+%
+The next building block of our approach consists of providing a mapping from
+each constructor representation to its actual target value, provided that each
+recursive sub-term has already been translated to its corresponding target
+value.
+%
+For this purpose, we will define a type class |Algebra| as follows:
+
 \begin{code}
@@
 \end{code}
+
+Where |f| is the functor type used to represent a construction of the target
+type |a|.
+%
+Note the functional dependency |f -> a|, ensuring that a given representation
+|f| can be translated to a unique target type |a|.
+%
+This mapping is usually referred categorically as an F-Algebra.
+%
 

### 2019-04-23 papers/haskell19/3-representation.tex
Section 3

@@
 %
+\subsection{Representing Data Constructors}
+
 When generating values of algebraic data types, the simplest piece of meaningful
@@
 \end{code}
-
-Where each representation has the same fields as its corresponding constructor,
-except for the recursive ones which are abstracted away using a type parameter
-|r|.
+%
+In the previous definitions, each representation has the same fields as its
+corresponding constructor, except for the recursive ones which are abstracted
+away using a type parameter |r|.
 %
@@
 %
-The next building block of our approach consists of providing a mapping from
-each constructor representation to its actual target value, provided that each
-recursive sub-term has already been translated to its corresponding target
+Then, the next building block of our approach consists of providing a mapping
+from each constructor representation to its actual target value, provided that
+each recursive sub-term has already been translated to its corresponding target
 value.
 %
-For this purpose, we will define a type class |Algebra| as follows:
+Categorically, this is often referred as an F-Algebra over the functor we use to
+encode our different constructions.
+%
+We will represent this mapping defining a type class |Algebra| with a single
+method |alg| as follows:
 
@@
 \end{code}
-
-Where |f| is the functor type used to represent a construction of the target
-type |a|.
 %
-Note the functional dependency |f -> a|, ensuring that a given representation
-|f| can be translated to a unique target type |a|.
+In the previous definition, |f| is the functor type used to represent a
+construction of the target type |a|.
 %
-This mapping is usually referred categorically as an F-Algebra.
+Note the functional dependency |f -> a|, helping the type checker to solve the
+free type variable |a| at the right hand side of the |=>|.
 %
+In practical terms, it ensures that each representation type |f| will be
+translated to a unique target type |a|.
+%
+Then, we need to instantiate this type class for each data constructor
+representation we are considering, providing an appropriate implementation for
+the overloaded |alg| function.
+%
+To do so we can simply transform each representation into its corresponding
+value by picking its analogous data constructor, and piping its fields
+unchanged:
 
@@
 \end{code}
+%
+Note how every definition regarding each different representation so far was
+obtained rather mechanically.
+%
+This let us automate the process quite a bit using Haskell's meta-programming
+capabilities.
+
+% --------------------------------------
+% Generic sums
+%
+\subsection{Composing Representations}
+
+So far we have seen how to represent each data constructor of our |Html| data
+type independently.
+%
+In order to represent interesting values, we need to be able to combine these
+single representations into (possibly complex) composite ones.
+%
+For this purpose, we will define a generic functor type |oplus| to encode the
+choice between two given representation types as follows:
 
 \begin{code}
-data ((f :: * -> *) oplus (g :: * -> *)) a = InL (f a) | InR (g a)
+data ((f :: * -> *) oplus'' (g :: * -> *)) r = InL (f r) | InR (g r)
+
+instance (Functor f, Functor g) => Functor (f oplus'' g)
+  where  fmap h (InL f) = InL (fmap h f)
+         fmap h (InR g) = InR (fmap h g)
+
+\end{code}
+%
+This infix type-level operator let us combine two representations |f| and |g|
+into a composite one |f oplus'' g|, representing either a value from |f| (via
+the |InL| constructor) or a value from |g| (via the |InR| constructor).
+%
+This operator works pretty much in the same way as Haskell's |Either| data type,
+except that, instead of combining two base types, it works combining two
+\emph{parametric type constructors}, hence the kind signature |* -> *| in both
+|f| and |g|.
+%
+For instance, we can use the type |Con_Text oplus'' Con_Tag| to encode values
+representing either plain text HTMLs or paired tags.
+%
+Such values can be constructed easily using the injections |InL| and |InR| on
+each each case, respectively.
+
+The next step consists of providing an |Algebra| type instance for composite
+representations, provided that each side of the sum can be translated back to
+the same target type:
+
+\begin{code}
+instance (Algebra f a, Algebra g a) => Algebra (f oplus'' g) a
+  where  alg (InL f) = alg f
+         alg (InR g) = alg g
 \end{code}
 
-\begin{code}
-instance (Algebra f a, Algebra g a)
-  => Algebra (f oplus g) a where
-  alg (InL f) = alg f
-  alg (InR g) = alg g
-\end{code}
+In the previous definition, we simply use the appropriate |Algebra| instance for
+the inner representation, based on which injection was used to construct the
+composite value.
+%
+It is worth remarking that this instance also lets us handle multiple
+compositions transparently, i.e., for every set of types |f1|, |f2|, ..., |fk|
+representing a target type |t| and having suitable |Algebra| type instances,
+then the composition |f1 oplus'' f2 oplus'' dots oplus'' fk| can be mapped back
+to |t| using this same mechanism.
+%
+Additionally, the order in which we associate each operand of |oplus| results
+irrelevant, thus we can omit parentheses when describing composite
+representations.
+%
+For simplicity, we will assume |oplus| as having rigth associativity, i.e., |f
+oplus'' g oplus'' h == f oplus'' (g oplus'' h)|.
+
+% --------------------------------------
+% Fixed points
+%
+\subsection{Tying the Knot}
+
+Even though we have already seen how to encode single and composite
+representations for our target data types, there is as piece of machinery still
+missing: our representations are not recursive, but parametric on its recursive
+fields.
+%
+We can think of them as a encoding a single ``layer'' of our target data.
+%
+In order to represent recursive values we need to close them ``tying the knot''
+recusively, i.e., once we have decided the final representation of our target
+data, we want each one of its recursive fields to be instantiated with itself.
+%
+This can be easily achieved by using a type-level fixed point operator:
 
@@
 \end{code}
+%
+Given a parametric type |f| of kind |* -> *| representing some target data type,
+the type |Fix f| instantiates each recursive field of |f| with |Fix f|, thus
+closing the definition of |f| into itself---note thus that the kind of |Fix f|
+results |*|.
+
+In general, if |f| is a representation for a given target type, we will refer to
+|Fix f| as a \emph{final representation}, since it cannot be further combined or
+extended---the |oplus| operator has to be applied \emph{within} the |Fix| type
+constructor.
+
+The effect of a fixed point combinator is much easier to interpret considering
+an example.
+%
+Let us imagine we want to represent our |Html| data type using all of its data
+constructors.
+%
+For this purpose we use the following type to represent it:
+
+\begin{code}
+type Html' = Con_Text oplus'' Con_Sing oplus'' Con_Tag oplus'' Con_Join
+\end{code}
+%
+Then, for instance, the value:
+
+\begin{code}
+  x :: Html
+  x = Join (Text "hello") (Sing "hr")
+\end{code}
+%
+can be represented with |Html'| as:
+
+\begin{code}
+  x' ::  Fix Html'
+  x' =   Fix  (InR (InR (InR (Con_Join
+              (Fix (InL (Con_Text "hello")))
+              (Fix (InR (InL (Con_Sing "hr"))))))))
+\end{code}
+%
+Where the sequences of |InL| and |InR| data constructors inject each value from
+an individual representation into the appropriate position of our composite
+representation |Html'|.
+
+Finally, given the value |x' :: Fix Html'|, we can evaluate it back to its
+corresponding |Html| value as follows:
 
@@
 \end{code}
+%
+This function exploits the |Functor| structure of our representations,
+unwrapping the fixed points and mapping their algebras to the result of
+evaluating recursively each recursive sub-term.
+%
+This recusion schema using functors and F-Algebras to fold data generically has
+been well studied and is known categorically a \emph{catamorphism}.
+
+In our particular example, this function satisfies the equation |eval x' == x|.
+%
+More specifically, one can prove that the types |Html| and |Fix Html'| are in
+fact isomorphic, with |eval| as the witness of one side of this
+isomorphism---though this is not usually the case for every representation we
+choose for our data.
+
+% --------------------------------------
+% Representing other constructions
+%
+\subsection{Representing Additional Constructions}
+
+The representation mechanism we have developed so far let us determine the shape
+of our target data based on the type we use to represent it.
+%
+However, if we only use it to represent data types based solely on their data
+constructors, then it is hardly useful for performing random testing, as it only
+adds unnecessary complexity to the random generation process.
+%
+It is not until we start considering more complex constructions that this
+approach becomes particularly appealing.
+
+%
+% Abstract interfaces
+%
+Fistly, let us consider the case of generating values obtained by abstract
+interface functions.
+%
+If we recall our |Html| example, we can observe that it is possible to use the
+functions on its abstract interface to obtain |Html| values based on different
+input arguments:
+
+\begin{code}
+br     ::                        Html
+bold   ::     Html   ->          Html
+list   ::  [  Html]  ->          Html
+(<+>)  ::     Html   -> Html ->  Html
+\end{code}
+
+Fortunately, it is easy to extend our approach to incorporate the interesting
+structure arising from these functions into our representations.
+%
+As before, we first need to define a set of simple data types to encode each
+construction:
 
@@
 data Fun_bold   r = Fun_bold r
+data Fun_list   r = Fun_list [r]
 data Fun_cat    r = Fun_cat  r r
-data Fun_list   r = Fun_list [r]
 \end{code}
+%
+Each one of these data types represents a value resulting from evaluating its
+corresponding function, using as input arguments the values encoded in its
+fields.
+%
+Once again, we replace each recursive field (representing a recursive input
+argument) with a type parameter |r| in order to leave the type of the recursive
+sub-terms unspecified until we have decided the overall shape of our data.
 
-\begin{code}
-data Pat_simplify_1 r = Pat_simplify_1 String String
-data Pat_simplify_2 r = Pat_simplify_2 String r r
-\end{code}
+Note that, by representing values obtained from function application this way,
+we are not performing any actual computation---we simply store the functions'
+input arguments.
+%
+Instead, the actual functions are evaluated when transforming our
+representations back to their target type, again by the means of an |Algebra|:
 
@@
 
+instance Algebra Fun_list Html where
+  alg (Fun_list xs) = list xs
+
 instance Algebra Fun_cat Html where
   alg (Fun_cat x y) = x <+> y
+\end{code}
+%
+In the previous |Algebra| instances, we simply return the result of evaluating
+the appropriate functions, using each representation field as an input argument.
 
-instance Algebra Fun_list Html where
-  alg (Fun_list xs) = list xs
+It is important to remark that this approach inherits any possible downside from
+the functions we use to represent our target data.
+%
+In particular, using non-terminating functions to represent our data might
+produce a non-terminating behavior when calling to the |eval| function.
+
+%
+% Functions pattern matchings
+%
+The second source of structural information that we consider in this work is the
+one present in functions' pattern matchings.
+%
+If we recall to our |simplify| from Section \ref{sec:sources}, we can observe it
+has two complex, non-trivial patterns that we might want to satisfy when
+generating random values:
+
+\begin{code}
+simplify (Join (Text t1) (Text t2))   = dots
+simplify (Join (Join (Text t1) x) y)  = dots
 \end{code}
+
+Once again, we can extend our approach to represent these patterns as well.
+%
+We start defining single data types for each one of them, this time encoding in
+the fields of each single data constructor the free pattern variables (or
+wildcards) appearing in corresponding patterns from left to right:
+
+\begin{code}
+data Pat_simplify_1 r = Pat_simplify_1 String String
+data Pat_simplify_2 r = Pat_simplify_2 String r r
+\end{code}
+%
+Where the number after the |#| simply distinguishes the different patterns from
+the function |simplify| using index of the clause they belong to.
+%
+Once again, we abstracted away every recursive field (corresponding to a
+recursive pattern variable) using a type variable |r| until we have decided the
+overall shape of our target data.
+
+Then, the task of our |Algebra| instances is to expand each pattern
+representation into the corresponding target value resembling such pattern,
+where each pattern variable gets instantiated using the values stored in its
+representation fields:
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl\cite{godefroid2005dart}.
+%
+% Closing
+%
+Using this extensions, we are now able to represent values of our target data
+using a higher abstraction level.
+%
+Instead of constructing values solely by combining different data constructors,
+this mechanism lets us also consider complex constructions like computing
+abstract interface functions and functions' patterns, treating them uniformingly
+and in a fine grained fashion.
+
+As we have introduced in the Section \ref{sec:sources}, having an extensible,
+fine-grained framework for generating random data might be of great advantage,
+as we would like to generate random test cases following different strategies
+when testing different properties.
+%
+For instance, generating values satisfying the pattern matchings of the function
+|simplify| could be relevant only when testing properties involving such
+function.
+%
+For this purpose, we could represent |Html| values as follows:
+%
+\begin{code}
+type Html_simplify  =      Con_Text oplus'' Con_Sing oplus'' Con_Tag oplus'' Con_Join
+                    oplus  Pat_simplify_1 oplus'' Pat_simplify_2
+\end{code}
+
+%
+Furthermore, if we want to test properties expecting valid |Html| values as
+inputs, we could avoid generating random |Sing| and |Tag| data constructors, as
+they might represent invalid tags.
+%
+Instead, we could specify that our random data can be constructed using only
+``safe'' constructions, using perhaps a representation type like follows:
+
+\begin{code}
+type ValidHtml = Con_Text oplus'' Con_Join oplus'' Fun_br oplus'' Fun_bold
+\end{code}
+
+The next section introduces a generic way to generate random values from our
+different representations, extending them with a set of type-level to encode
+information relevant to the generation process directly on their types.

### 2019-04-23 src/Test/Dragen2/Infix.hs

@@ -9,7 +9,7 @@ type (:*) = Freq
 infixl 6 :*
 
 type (:+) = Sum
-infixl 5 :+
+infixr 5 :+
 
 type (m :> k) = Lookup m k
 infix 4 :>

### 2019-04-24 papers/haskell19/2-sources.tex
Section 3
WIP Section 4

@@
         ,  (1,  Sing    <$> arbitrary) ]
-      gen n = frequency
+      gen d = frequency
         [  (2,  Text    <$> arbitrary)
         ,  (1,  Sing    <$> arbitrary)
-        ,  (4,  Tag     <$> arbitrary  <*> gen (n-1))
-        ,  (3,  Join    <$> gen (n-1)  <*> gen (n-1)) ]
+        ,  (4,  Tag     <$> arbitrary  <*> gen (d-1))
+        ,  (3,  Join    <$> gen (d-1)  <*> gen (d-1)) ]
 \end{code} %$
@@
 
+list :: [Html] -> Html
+list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))
+
 (<+>) :: Html -> Html -> Html
 (<+>) x y = Join x (Join br y)
-
-list :: [Html] -> Html
-list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))
 \end{code} %$

### 2019-04-24 papers/haskell19/3-representation.tex
WIP Section 4

@@
 
-Once again, we can extend our approach to represent these patterns as well.
+We can extend our approach further in order to represent these patterns as well.
 %
@@
 Where the number after the |#| simply distinguishes the different patterns from
-the function |simplify| using index of the clause they belong to.
+the function |simplify| using the index of the clause they belong to.
 %
-Once again, we abstracted away every recursive field (corresponding to a
-recursive pattern variable) using a type variable |r| until we have decided the
-overall shape of our target data.
+Once again, we abstracted away every recursive field (this time corresponding to
+a recursive pattern variable or wildcard) using a type variable |r| until we
+have decided the overall shape of our target data.
 
@@
 The next section introduces a generic way to generate random values from our
-different representations, extending them with a set of type-level to encode
-information relevant to the generation process directly on their types.
+different representations, extending them with a set of type-level combinators
+to encode information relevant to the generation process directly on their
+types.

### 2019-04-24 papers/haskell19/4-generators.tex
WIP Section 4

@@
 
+% --------------------------------------
+% Intro
+%
+So far we have seen how to encode different constructions representing
+interesting values from our target data types.
+%
+Such representations follow a modular approach, where each different
+construction is considered independent from the rest.
+%
+This modularity allows us to derive each different construction independently,
+as well to specify the shape of our target data with with a fine granularity.
+
+In this section we introduce the machinery required to randomly generate the
+values encoded using our representations.
+%
+For this purpose, this step also follows the modular, fine-grained fashion,
+resulting in a generation process entirely compositional.
+%
+In this light, our random generators are built from smaller ones (each one
+representing a single construction), and they are based solely on the types we
+use to represent the shape of our random data.
+
+Ideally, our aim is to be able to obtain random generators with a behavior
+similar to the one presented for |Html| in Section \ref{sec:sources}.
+%
+If we take a closer look at its definition, we can observe three factors
+happening simultaneously:
+%
+\begin{itemize}
+\item We use the QuickCheck's generation size to limit the depth of the
+  generated values, reducing it by one on each recursive call of the local
+  auxiliary function |gen|.
+\item We differentiate between \emph{terminal} and \emph{non-terminal
+    constructors}, limiting our generator to pick only among terminal ones when
+  we have reached the maximum depth (case |gen 0|).
+\item We generate different constructions in a (possibly) different frequency
+  that is fixed over time.
+\end{itemize}
+%
+Then, in order to achieve the desired generation behavior, we will need to model
+these previous aspects in our modular framework, and in such a way that we do
+not compromise the compositionality obtained so far.
+
+% --------------------------------------
+% Depth-bounded generators
+%
+\subsection{Depth-Bounded Modular Generators}
+
+The first obstacle that arises when trying to generate random values with a
+limited depth using our approach is related to modularity.
+%
+If we recall the random generator for |Html| from Section \ref{sec:sources}, we
+can observe that the depth parameter |d| is threaded to the different recursive
+call of our generator, but always within the scope of the |gen| local function.
+%
+Since each construction will have an specialized random generator, we cannot
+group them as we did before using an internal |gen| function.
+%
+Instead, we will define a depth-bounded generator data type, wrapping
+QuickCheck's |Gen| data type with a external parameter representing the current
+maximum recursive depth:
+
 \begin{code}
 type BGen a = Int -> Gen a
+\end{code}
 
+A |BGen| is essentially a normal |Gen| with the maximum recursive depth as an
+input parameter.
+%
+Using this previous definition, we can generalize QuickCheck's |Arbitrary| class
+to work with depth-bounded generators simply as follows:
+
+\begin{code}
 class BArbitrary (a :: *) where
@@
 
+From now on, we will use this type class as a more flexibe substitute of
+|Arbitrary|, given that now we have two parameters to tune: the maximum
+recursive depth, and the QuickCheck generation size.
+%
+Where the former is useful to tune the overall size of our random data, whereas
+the latter can be use for tuning the values of the ``leaves'', such as the
+maximum length of random |String| or the maximum/minumum random |Int|s.
+
+Here we want to actively remark that, even though we could have reused the
+QuickCheck generation size parameter to simultaneously model the maximum
+recursive depth and the maximum size of the leaves, doing so would imply
+generating random leaves with a decresing size as we move deeper within a random
+value, obtaining for instance, random trees with all zeroes on its leaves, or
+random lists showing a hight tendency to be ordered in decreasing order.
+%
+In addition, one can always obtain a trivial |Arbitrary| instance from a
+|BArbitrary| one by using QuickCheck's generation size both for the maximum
+recursive depth as well as for generating leaves:
+
+\begin{code}
+instance BArbitrary a => Arbitrary a where
+  arbitrary = sized bounded
+\end{code} %$
+
+Even though we have extended QuickCheck's generators to be depth-aware, here we
+also need to consider the parametric nature of our representations.
+%
+In the previous section, we defined each construction representation as being
+parametric on the type of the recursive sub-terms, as a manner to defer this
+choice until we have specified the overal shape of our target data.
+%
+Hence, each construction representation is of kind |* -> *|.
+%
+If we want to define our generators in a modular way, we also need to
+parametrize somehow the generation of the recursive subterms!
+%
+Fortunately, QuickCheck already defines a lifted type class |Arbitrary1| for
+types for parametric types of kind |*->*|:
+
 \begin{code}
 class Arbitrary1 (f :: * -> *) where
-  liftGen :: Gen a -> Gen (f a)
+  liftArbitrary :: Gen a -> Gen (f a)
 \end{code}
+
+This type class allows us to write modular generators, leaving unspecified the
+specific generator used to generate values of its type argument.
+%
+For instance, the default parametric generator for the data type |Maybe| looks
+like:
+
+\begin{code}
+instance Arbitrary1 Maybe where
+  liftArbitrary gen = frequency
+    [  (1,  pure Nothing),  (3,  Just <$> gen) ]
+\end{code} %$
+
+Then, we can use the same mechanism for our modular generators, extending it to
+be depth-aware as follows:
 
@@
 
+Note the similarity between the |Arbitrary1| and |BArbitrary1| type classes.
+%
+We will use this type class to implement random generators for each construction
+we are considering.
+%
+Recalling our |Html| example, we can finally define modular random generator for
+the constructions representing data constructors as follows:
+
 \begin{code}
-data ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
+instance BArbitrary1 Con_Text where
+  liftBGen gen d = Con_Text <$> arbitrary
+
+instance BArbitrary1 Con_Sing where
+  liftBGen gen d = Con_Sing <$> arbitrary
+
+instance BArbitrary1 Con_Tag where
+  liftBGen gen d = Con_Tag <$> arbitrary <*> gen (d - 1)
+
+instance BArbitrary1 Con_Join where
+  liftBGen gen d = Con_Join <$> gen (d - 1) <*> gen (d - 1)
+\end{code} %$
+
+Note in the previous definitions how each instance is defined to be independent
+of the maximum depth (paramtrized by the input variable |d|) and the random
+generator used for the recursive sub-terms (parametrized by the input variable
+|gen|).
+%
+Every other non-recursive sub-term can be generated using a normal |Arbitrary|
+instance---we use this to generate random |String|s in the previous definitions.
+
+The rest of our construction representations can be generated using the same
+idea.
+%
+For example, the |BArbitrary1| instances for |Fun_bold| and |Pat_simplify_2| are
+as follows:
+
+\begin{code}
+instance BArbitrary1 Fun_bold where
+  liftBGen gen d = Fun_bold <$> gen (d - 1)
+
+instance BArbitrary1 Pat_simplify_2 where
+  liftBGen gen d = Pat_simplify_2  <$> arbitrary
+                                   <*> gen (d - 1)
+                                   <*> gen (d - 1)
+\end{code} %$
+
+% \begin{code}
+% instance BArbitrary1 Fun_br where
+%   liftBGen gen d = pure Fun_br
+
+% instance BArbitrary1 Fun_bold where
+%   liftBGen gen d = Fun_bold <$> gen (d - 1)
+
+% instance BArbitrary1 Fun_cat where
+%   liftBGen gen d = Fun_cat <$> gen (d - 1) <*> gen (d - 1)
+
+% instance BArbitrary1 Fun_list where
+%   liftBGen gen d = Fun_list <$> liftArbitrary (gen (d - 1))
+% \end{code} %$
+
+% \begin{code}
+% instance BArbitrary1 Pat_simplify_1 where
+%   liftBGen gen d
+%     = Pat_simplify_1 <$> arbitrary <*> arbitrary
+
+% instance BArbitrary1 Pat_simplify_2 where
+%   liftBGen gen d
+%     = Pat_simplify_2 <$> gen (d - 1) <*> gen (d - 1)
+% \end{code}
+
+Having all the modular generators for each construction in place, we can finally
+obtain a concrete depth-aware generator (of kind |*|) for any final
+representation |f| as follows:
+
+\begin{code}
+instance BArbitrary1 f => BArbitrary (Fix f) where
+  bounded d = Fix <$> liftBGen bounded d
+\end{code} %$
+
+The previous instance uses the |BArbitrary1| of our final representation to
+generate sub-terms recursively using itself (|bounded d|) as the parametrized
+input generator, wrapping each recursive sub-term with a |Fix| data constructor.
+
+The machinery developed so far lets us generate single random constructions in a
+modular fashion.
+%
+However, we still need to develop our generation mechanism a bit further in
+order to generate composite representations built using the |oplus''| type-level
+operator.
+%
+This is the objective of the next sub-section.
+
+% --------------------------------------
+% Encoding Generation Instructions Using Types
+%
+\subsection{Encoding Generation Instructions Using Types}
+
+\begin{code}
+newtype ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
 \end{code}
@@
 \begin{code}
-data Term (f :: * -> *) a = Term (f a)
+newtype Term (f :: * -> *) a = Term (f a)
 \end{code}
@@
 \begin{code}
-instance BArbitrary1 Con_Text where
-  liftBGen gen d = Con_Text <$> arbitrary
-
-instance BArbitrary1 Con_Sing where
-  liftBGen gen d = Con_Sing <$> arbitrary
-
-instance BArbitrary1 Con_Tag where
-  liftBGen gen d = Con_Tag <$> arbitrary <*> gen (d - 1)
-
-instance BArbitrary1 Con_Join where
-  liftBGen gen d = Con_Join <$> gen (d - 1) <*> gen (d - 1)
-\end{code} %$
-
-\begin{code}
-instance BArbitrary1 Fun_br where
-  liftBGen gen d = pure Fun_br
-
-instance BArbitrary1 Fun_bold where
-  liftBGen gen d = Fun_bold <$> gen (d - 1)
-
-instance BArbitrary1 Fun_cat where
-  liftBGen gen d = Fun_cat <$> gen (d - 1) <*> gen (d - 1)
-
-instance BArbitrary1 Fun_list where
-  liftBGen gen d = Fun_list <$> liftGen (gen (d - 1))
-\end{code} %$
-
-\begin{code}
-instance BArbitrary1 Pat_simplify_1 where
-  liftBGen gen d
-    = Pat_simplify_1 <$> arbitrary <*> arbitrary
-
-instance BArbitrary1 Pat_simplify_2 where
-  liftBGen gen d
-    = Pat_simplify_2 <$> gen (d - 1) <*> gen (d - 1)
-\end{code}
-
-\begin{code}
-instance BArbitrary1 f => BArbitrary (Fix f) where
-  bounded d = Fix <$> liftBGen bounded d
-\end{code} %$
-
-\begin{code}
 genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)

### 2019-04-24 papers/haskell19/3-representation.tex
Section 4

@@
 \begin{code}
-data Fix (f :: * -> *) = Fix { unFix :: (f (Fix f)) }
+newtype Fix (f :: * -> *) = Fix { unFix :: f (Fix f) }
 \end{code}
@@
 results |*|.
+%
+In addition, note that we defined |Fix| as a |newtype|.
+%
+This way, the compiler will avoid producing a runtime representation for it,
+thus avoiding unnecessary overhead.
+%
+This is possible since we only use it as a wrapper to obtain a self-recursive
+type from a parametric one.
 

### 2019-04-24 papers/haskell19/4-generators.tex
Section 4

@@
 In the previous section, we defined each construction representation as being
-parametric on the type of the recursive sub-terms, as a manner to defer this
-choice until we have specified the overal shape of our target data.
+parametric on the type of its recursive sub-terms, as a way to defer this choice
+until we have specified the overal shape of our target data.
 %
@@
 %
-Fortunately, QuickCheck already defines a lifted type class |Arbitrary1| for
-types for parametric types of kind |*->*|:
+If we look at QuickCheck, this library already defines a type class |Arbitrary1|
+for parametric types of kind |* -> *|:
 
@@
 This type class allows us to write modular generators, leaving unspecified the
-specific generator used to generate values of its type argument.
+specific generator used when generating values of its type argument.
 %
@@
 
-Then, we can use the same mechanism for our modular generators, extending it to
+Where the random generator use to generate the value stored in a |Just|
+constructor is parametrized by the |gen| input variable.
+%
+We can use this same mechanism for our modular generators, first extending it to
 be depth-aware as follows:
@@
 \end{code}
-
+%
 Note the similarity between the |Arbitrary1| and |BArbitrary1| type classes.
 %
-We will use this type class to implement random generators for each construction
-we are considering.
+Then, we will use this type class to implement random generators for each
+construction we are considering.
 %
-Recalling our |Html| example, we can finally define modular random generator for
-the constructions representing data constructors as follows:
+Recalling our |Html| example, we can define modular random generator for the
+constructions representing data constructors as follows:
 
@@
 
-The rest of our construction representations can be generated using the same
-idea.
+The rest of our construction representations can be generated analogously.
 %
@@
 
-The previous instance uses the |BArbitrary1| of our final representation to
+The previous instance employs the |BArbitrary1| of our final representation to
 generate sub-terms recursively using itself (|bounded d|) as the parametrized
@@
 
+%
+% Wrapping this subsection
+%
 The machinery developed so far lets us generate single random constructions in a
@@
 However, we still need to develop our generation mechanism a bit further in
-order to generate composite representations built using the |oplus''| type-level
+order to generate composite representations built using the |oplus| type-level
 operator.
@@
 % --------------------------------------
-% Encoding Generation Instructions Using Types
+% Encoding Generation Behavior Using Types
 %
-\subsection{Encoding Generation Instructions Using Types}
+\subsection{Encoding Generation Behavior Using Types}
 
-\begin{code}
-newtype ((f :: * -> *) otimes (n :: Nat)) a = Freq (f a)
-\end{code}
+As we have shown so far, generating each construction representation is rather
+straightforward: there is only one data constructor to select, and every field
+is generating following a mechanical recipe.
+%
+In our approach, most of the generation complexity is encoded in the random
+generator used for composite representations built upon the |oplus| operator.
+%
+Before introducing it, we need to define some additional machinery to encode the
+notions of terminal construction and generation frequency.
+
+Recalling the random generator for |Html| presented in Section
+\ref{sec:sources}, we can observe that the last generation level (see |gen 0|)
+is constrained to generate only the non-recursive subset of all the available
+constructions.
+%
+In order to model this behavior in our framework, we can start by defining a
+data type |Term| to ``tag'' every terminal construction explicitly:
 
@@
 
-\begin{code}
-instance Algebra f a => Algebra (f otimes n) a where
-  alg (Freq f) = alg f
-\end{code}
+Then, if |f| is a non-recursive construction, the type |Term f oplus' g| can be
+interpreted as representing values generated using values drawn both from |f|
+and |g|, but ``closed'' using only values from |f|.
+%
+Since this data type will not add any semantic information to our represented
+values, we can define suitable |Algebra| and |BArbitrary1| instances for it
+simply by delegating the work to the inner type:
 
@@
   alg (Term f) = alg f
-\end{code}
-
-\begin{code}
-instance BArbitrary1 f => BArbitrary1 (f otimes n) where
-  liftBGen gen d = Freq <$> liftBGen gen d
 
@@
   liftBGen gen d = Term <$> liftBGen gen d
+\end{code} %$
+
+The next building block of our framework consists in a way of specifying the
+generation frequency of each construction.
+%
+For this purpose, we can follow the same reasoning as before, defining a data
+type-level operator |otimes| to explicitly tag the generation frequency of a
+given representation:
+
+\begin{code}
+newtype ((f :: * -> *) otimes'' (n :: Nat)) a = Freq (f a)
 \end{code}
+
+In the definition above, the operator |otimes| is parametrized by a type-level
+natural number |n| (of kind |Nat|) representing the desired generation
+frequency.
+%
+In this light, the type |(f otimes'' 3) oplus'' (g otimes'' 1)| represents
+values generated using values from both |f| and |g|, where |f| is randomly
+chosen three times more frequently than |g|.
+%
+In practice, we defined |otimes| such that it associates more strongly than
+|oplus|, thus avoiding the need of parenthesis in types like the previous one.
+%
+Analogously as |Term|, the operator |otimes| does not add any semantic
+information to the values it represents, so we can define its |Algebra| and
+|BAbitrary1| instance by delegating the work to the inner type as before:
+
+\begin{code}
+instance Algebra f a => Algebra (f otimes'' n) a where
+  alg (Freq f) = alg f
+
+instance BArbitrary1 f => BArbitrary1 (f otimes'' n) where
+  liftBGen gen d = Freq <$> liftBGen gen d
+\end{code} %$
+
+With these two new type level combinators |Term| and |otimes|, we are now able
+to express the behavior of our entire generation process based solely on the
+type we are generating.
+%
+In addition to these combinators, we will need to perform some type-level
+computations based on them in order to define our random generator for composite
+representations.
+%
+Consider for instance the following type defined using parenthesis for clarity:
+
+\begin{code}
+(f otimes'' 2) oplus'' ((g otimes'' 3) oplus'' (Term h))
+\end{code}
+
+Our generation process will traverse this type one combinator at a time,
+processing each occurrence of |oplus| independently.
+%
+This means that, in order to select the appropriate generation frequency of each
+operand we need to calculate the overall sum of frequencies on each side of the
+|oplus|.
+%
+For this purpose, we can implement a closed type family |FreqOf| to compute the
+overal sum of frequencies of a given representation type:
 
@@
 type family FreqOf (f :: * -> *) :: Nat where
-  FreqOf (f oplus g)   = FreqOf f + FreqOf g
-  FreqOf (f otimes n)  = n * FreqOf f
+  FreqOf (f oplus'' g)   = FreqOf f + FreqOf g
+  FreqOf (f otimes'' n)  = n * FreqOf f
   FreqOf (Term f)        = FreqOf f
@@
 
+This type-level function takes a representation type as an input and traverses
+it recursively, adding up each frequency tag found in the process, and returning
+a type-level natural number.
+%
+Note how in the second equation we multiply the frequency encoded in the
+|otimes| tag with the frequency of the type it is wrapping.
+%
+This way, the type |((f otimes'' 2) oplus'' g) otimes'' 3| is equivalent to |(f
+otimes'' 6) oplus'' (g otimes'' 3)|, following the natural intuition for the
+addition and multiplication operations over numbers.
+%
+Moreover, if a type does have an explicit frequency then its frequency is simply
+defaulted to one.
+
+On the other hand, the last step of our generation process which only generates
+terminal constructions could be seen as generating values considering all the
+available constructions, but instead treating the non-terminal ones as having
+generation frequency zero.
+%
+This way we can reuse our previous type level computation to calculate the
+\emph{terminal generation frequency} |FreqOf'| of a given representation type as
+follows:
+
 \begin{code}
 type family FreqOf' (f :: * -> *) :: Nat where
-  FreqOf' (f oplus g)   = FreqOf' f  +  FreqOf' g
-  FreqOf' (f otimes n)  = n  *  FreqOf' f
+  FreqOf' (f oplus'' g)   = FreqOf' f  +  FreqOf' g
+  FreqOf' (f otimes'' n)  = n  *  FreqOf' f
   FreqOf' (Term f)        = FreqOf f
@@
 \end{code}
+
+Similar to |FreqOf|, the type family above traverses its input type adding the
+terminal frequency of each sub-type.
+%
+However, |FreqOf'| only considers the frequency of those representation
+sub-types that are explicitly tagged as terminal, returning zero in any other
+case.
+
+Using the typle level combinators introduced at the beginning of this
+sub-section along with the previous type-level computations over frequencies, we
+are finally in position of defining our random generator for composite
+representations:
 
@@
 
+Like the generator introduced in Section \ref{sec:sources} for |Html|, this
+generator branches over the current generation depth |d|.
+%
+In the case we can still generate values from any construction (|d > 0|), we
+will use QuickCheck's |frequency| operation to randomly choose between
+generating a value of each side of the |oplus|, i.e., either a value of |f| or a
+value of |g|, following the generation frequencies specified for both of them,
+and wrapping the values with the appropriate injection |InL| or |InR| on each
+case.
+%
+Such frequencies are obtained by \emph{reflecting} the type-level natural values
+obtained from applying |FreqOf| to both |f| and |g|, using a type-dependent
+function |freqVal| that returns the number corresponding to the type-level
+natural value we apply to it:
+
 \begin{code}
@@
 
-\begin{code}
-genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)
-genFix d = bounded at(Fix f) d
-\end{code}
+Note that the type of |freqVal| is ambiguous, since it quantifies over every
+possible known type-level natural value |n|.
+%
+We need to use a \emph{visible type application} (employing the |at (...)|
+syntax) to desambiguate which type-level natural value we are refering to.
+%
+Then, for instance, the value |freqVal at (FreqOf (f otimes'' 5))| will evaluate
+to |5 :: Int|.
+
+The remaining part of our random generator works analogously, except that this
+time we only want to genrate terminal constructors, hence we use the |FreqOf'|
+type family to compute the terminal generation frequency of each operand.
+%
+If any of |FreqOf' f| or |FreqOf' g| evaluates to zero, then it means that that
+operand does not contain any terminal construction, and |frequency| will not
+consider it when generating terminal values.
+
+If it happens that both |FreqOf' f| and |FreqOf' g| compute to zero when
+genrating a composite value, then it will produce a runtime error triggered by
+the function |frequency|, as it does not have anything with a positive frequency
+to generate.
+%
+This kind of exceptions will arise if we forget to tag at least a single
+construction within a composite representation as terminal.
+%
+Fortunately, such runtime exceptions can be catched in compile time by defining
+a type constraint that ensures we are trying to generate values using a
+representation with a strictly positive terminal generation frequency---thus
+containing at least a single terminal construction:
 
 \begin{code}
-genEval :: forall f a dot (BArbitrary1 f, Algebra f a) => BGen a
-genEval d = eval <$> genFix at f d
+type family HasTerminals (f :: * -> *) :: Constraint where
+  HasTerminals f = IsPositive (FreqOf' f)
+
+type family IsPositive (n :: Nat) :: Constraint where
+  IsPositive 0 = TypeError "No terminals"
+  IsPositive _ = ()
+\end{code}
+%
+The previous type families compute the terminal generation frequency of a
+representation type |f|, returning a type error if its result is zero, or an
+empty constraint |()| that is always trivially satisfied in any other case.
+%
+Finally, we can use this constraint to define a safe generation primitive
+|genEval| to obtain a random depth-bounded generator for every target type |a|
+specified using a safe generation representation |f|:
+
+\begin{code}
+genEval  :: forall f a dot
+  (BArbitrary1 f, HasTerminals f, Algebra f a) => BGen a
+genEval d = eval <$> bounded at (Fix f) d
 \end{code} %$
+
+Note how this primitive is also ambiguous in the type used for the
+representation.
+%
+This allows us to use a visible type application to obtain values from the same
+target type but generated using a different underlying representations.
+
+So far we have seen how to represent and generate values for a target data type
+by combining a wide selection of structural constructions, as well as a series
+of type-level combinators to encode the generation behavior.
+%
+The next section refines our type-level machinery in order to export an elegant
+domain specific language for specifying composable random generators.

### 2019-04-25 papers/haskell19/4-generators.tex
Section 5

@@
 The remaining part of our random generator works analogously, except that this
-time we only want to genrate terminal constructors, hence we use the |FreqOf'|
+time we only want to generate terminal constructions, hence we use the |FreqOf'|
 type family to compute the terminal generation frequency of each operand.
@@
 If any of |FreqOf' f| or |FreqOf' g| evaluates to zero, then it means that that
-operand does not contain any terminal construction, and |frequency| will not
+operand does not contain any terminal constructions, and |frequency| will not
 consider it when generating terminal values.
@@
 %
-This kind of exceptions will arise if we forget to tag at least a single
-construction within a composite representation as terminal.
+This kind of exceptions will arise for examplke if we forget to tag at least a
+single construction within a composite representation as terminal.
 %
-Fortunately, such runtime exceptions can be catched in compile time by defining
+Fortunately, such runtime exceptions can be catched in compile-time by defining
 a type constraint that ensures we are trying to generate values using a
@@
 target type but generated using a different underlying representations.
+%
+For instance, we can obtain two different concrete generators for our |Html|
+type simply by changing its representation type as follows:
 
+\begin{code}
+genHtmlPats :: BGen Html
+genHtmlPats = genEval at Html_simplify
+
+genValidHtml :: BGen Html
+genValidHtml = genEval  at ValidHtml
+\end{code}
+
+Where |Html_simplify| and |ValidHtml| are the types defined at the end of
+Section \ref{sec:representation}.
+
+% --------------------------------------
+% Closing
+%
 So far we have seen how to represent and generate values for a target data type

### 2019-04-25 papers/haskell19/5-typelevel.tex
Section 5

@@
 \label{sec:typelevel}
+
+% --------------------------------------
+% Intro
+%
+This section introduces a set refinements to our basic language for describing
+random generators, making them more flexible and robust in order to fit
+real-world usage scenarios.
+
+% --------------------------------------
+% Naming conventions
+%
+To begin with, let us consider the name conventions we used when defining our
+representation data types along this paper.
+%
+Since our aim is to automate this process as much as possible using Haskell's
+meta-programming capabilities, the actual names used when deriving construction
+representations will have to be generated in such a way that they are compliant
+with Haskell syntax and are unique within our namespace.
+%
+This means that, in practice, type names like |Fun_cat| or |Pat_simplify_1| are
+technically not valid Haskell data type names, thus they will have to be
+synthesized as something like |Fun_lt_plus_gt_543| and |Pat_simplify_1_325|,
+where the last sequence of numbers is automatically inserted by Template Haskell
+to ensure uniqueness.
+
+This naming convention results hard to use, specially if we consider that we do
+not now the actual name types until they are synthesized in compile-time, due to
+their unique suffixes.
+%
+Fortunately, it is easy to solve this problem again using some type-level
+machinery.
+%
+Instead of imposing a naming convention in our representation synthesizer, we
+can define a set of open type families to hide each kind of construction using
+meaningful names:
+
+\begin{code}
+type family Con (c :: Symbol)             :: * -> *
+type family Fun (f :: Symbol)             :: * -> *
+type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
+\end{code}
+
+Where |Symbol| is the kind of type-level strings in Haskell.
+%
+Then, our derivation process will synthesize each representation using unique
+names, along with a type instance of the corresponding type family for the end
+user to manipulate it.
+%
+For instance, along with the representations of |Con_Text|, |Fun_cat| and
+|Pat_simplify_1|, we will automatically derive the following type instances:
+
+\begin{code}
+type instance Con "Text"        = Con_Text_123
+type instance Fun "<+>"         = Fun_lt_plus_gt_543
+type instance Pat "simplify" 1  = Pat_simplify_1_325
+\end{code}
+
+And the end user can simply refer to each particular construction by using these
+synonyms, e.g., with representation types like |Con "Text" oplus' Fun "<+>"|.
+
+Moreover, we can define a \emph{terminal} version of each kind of construction
+to avoid having to explicitly wrap constructions using |Term| combinators:
+
+\begin{code}
+type Con' a    = Term  (Con a)
+type Pat' a n  = Term  (Pat a n)
+type Fun' a    = Term  (Fun a)
+\end{code}
+
+This way, any quoted construction will considered to be terminal.
+%
+With these type-level manipulations, we are now able to write generation
+specifications in a clear and concise way.
+%
+For instance, we can specify a complex random generator for |Html| values as
+follows:
 
@@
 
+And simply obtain a concrete generator by refering to this previous type when
+using |genEval|:
+
 \begin{code}
-instance BArbitrary Html where
-  bounded = genEval at Html_Spec
-
-instance Arbitrary Html where
-  arbitrary = sized bounded
+genHtml :: BGen Html
+genHtml = genEval at Html_Spec
 \end{code}
 
-\begin{code}
-type family Con (c :: Symbol)             :: * -> *
-type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
-type family Fun (f :: Symbol)             :: * -> *
-\end{code}
+% --------------------------------------
+% Parametric target data types
+%
+\subsection{Parametric Target Data Types }
 
-\begin{code}
-type instance Con "Text"        = Con_Text
-type instance Pat "simplify" 1  = Pat_simplify_1
-type instance Fun "<+>"         = Fun_cat
-\end{code}
-
-\begin{code}
-type Con' a    = Term  (Con a)
-type Pat' a n  = Term  (Pat a n)
-type Fun' a    = Term  (Fun a)
-\end{code}
+So far we have seen how specify random generators for our simple self-contained
+|Html| data type.
+%
+In practice, however, we are often required to write random generators for
+parametric target data types as well.
+%
+Consider for example the following |Tree| data type definition encoding binary
+trees with generic information of type |a| in the leaves:
 
@@
 \end{code}
+
+In order to represent its data constructors, we can follow the same recipe
+presented in Section \ref{sec:representation}, but also parametrizing our
+representations over the type variable |a| as well:
 
@@
 
+Then, the rest of the machinery can be derived in the same way as before,
+carrying this type parameter and including the appropriate |Arbitrary|
+constraints all along the way:
+
 \begin{code}
@@
 instance Algebra (Con_Node a) (Tree a) where ...
-\end{code}
 
-\begin{code}
 instance Arbitrary a => BArbitrary1 (Con_Leaf a) where ...
@@
 
+The biggest different appears when deriving the required type instances for the
+|Con|, |Fun| and |Pat| synonyms.
+%
+Instead of carrying this type parameter in our generation specifications, we can
+avoid it by hiding them using an existential type:
+
 \begin{code}
@@
 \end{code}
+
+In the previous definition, |Some| is a wrapper for a 2-parametric type that
+hides the first type variable by using an explicit existential quantificator.
+%
+Note thus that the type parameter |a| does not appear at the left hand side of
+|Some| in its defition.
+%
+Then, when deriving any |Con|, |Fun| or |Pat| type synonym, we will wrap each
+construction representations using a |Some|:
 
@@
 
+This way we can write generation specification for our |Tree| data type without
+having to refer to its type parameter anywhere.
+%
+For instance:
+
 \begin{code}
 type Tree_Spec
-  =      Con "Leaf" otimes 2
+  =      Con'  "Leaf"  otimes 2
   oplus  Con   "Node"  otimes 3
@@
 
+Instead, we defer handling this type parameter until we actually use it to
+define a concrete generator.
+%
+For instance, we can write a concrete generator of |Tree Int| as follows:
+
 \begin{code}
-type family (f :: * -> *) apply (a :: *) :: * -> *  where
-  (Some f)      apply  a  = f a
-  (f oplus g)   apply  a  = (f apply a) oplus (g apply a)
-  (f otimes n)  apply  a  = (f apply a) otimes n
-  (Term f)      apply  a  = Term (t apply a)
-  f             apply  a  = f
+genIntTree :: BGen (Tree Int)
+genIntTree = genEval at (Tree_Spec apply' Int)
 \end{code}
 
+Where |apply| is a closed type family that simply traverses our generation
+specification, applying the |Int| type to every occurrence of |Some|:
+
 \begin{code}
-instance Arbitrary a => BArbitrary (Tree a) where
-  bounded = genEval at (Tree_Spec apply a)
+type family (f :: * -> *) apply' (a :: *) :: * -> * space where
+  (Some f)         apply'  a  = f a
+  (f oplus''  g)   apply'  a  = (f apply' a) oplus'' (g apply' a)
+  (f otimes'' n)   apply'  a  = (f apply' a) otimes'' n
+  (Term f)         apply'  a  = Term (t apply' a)
+  f                apply'  a  = f
 \end{code}
+
+As a result, in our definition of |genIntTree| above, the |apply| operator will
+reduce the type |(Tree_Spec apply'' Int)| to the following concrete type:
+
+\begin{code}
+(Term (Con_Leaf Int) otimes' 2) oplus' ((Con_Node Int) otimes' 3)
+\end{code}
+
+Along with our automated constructions derivation mechanism, the machinery
+introduced in this section allows us to specify random generators using a
+concise and flexible type-level domain specific language.
+%
+The next section evaluates our approach in terms of performance using a set of
+case studies extracted from real-world Haskell libraries.

### 2019-04-28 bench/BlockTable.hs
Work on benchmarks, bug fixes on implementation

@@ -1,0 +1,175 @@
+-----------------------------------------------------------------------------
+-- |
+-- Module      :  Text.Html.BlockTable
+-- Copyright   :  (c) Andy Gill and OGI, 1999-2001
+-- License     :  BSD-style (see the file libraries/base/LICENSE)
+-- 
+-- Maintainer  :  Andy Gill <andy@galconn.com>
+-- Stability   :  provisional
+-- Portability :  portable
+--
+-- An Html combinator library
+--
+-----------------------------------------------------------------------------
+
+module BlockTable (
+
+-- Datatypes:
+
+      BlockTable,             -- abstract
+
+-- Contruction Functions: 
+
+      single,
+      empty,
+      above,
+      beside,
+
+-- Investigation Functions: 
+
+      getMatrix,
+      showsTable,
+      showTable,
+
+      ) where
+
+infixr 4 `beside`
+infixr 3 `above`
+
+-- These combinators can be used to build formated 2D tables.
+-- The specific target useage is for HTML table generation.
+
+{-
+   Examples of use:
+
+  	> table1 :: BlockTable String
+  	> table1 = single "Hello"	+-----+
+					|Hello|
+	  This is a 1x1 cell		+-----+
+	  Note: single has type
+	 
+		single :: a -> BlockTable a
+	
+	  So the cells can contain anything.
+	
+	> table2 :: BlockTable String
+	> table2 = single "World"	+-----+
+					|World|
+					+-----+
+
+
+	> table3 :: BlockTable String
+	> table3 = table1 %-% table2	+-----%-----+
+					|Hello%World|
+	 % is used to indicate		+-----%-----+
+	 the join edge between
+	 the two Tables.  
+
+	> table4 :: BlockTable String
+	> table4 = table3 %/% table2	+-----+-----+
+					|Hello|World|
+	  Notice the padding on the	%%%%%%%%%%%%%
+	  smaller (bottom) cell to	|World      |
+	  force the table to be a	+-----------+
+	  rectangle.
+
+	> table5 :: BlockTable String
+	> table5 = table1 %-% table4	+-----%-----+-----+
+					|Hello%Hello|World|
+	  Notice the padding on the	|     %-----+-----+
+	  leftmost cell, again to	|     %World      |
+	  force the table to be a	+-----%-----------+
+	  rectangle.
+ 
+   Now the table can be rendered with processTable, for example:
[…]

### 2019-04-28 bench/Html.hs

@@ -1,0 +1,1466 @@
+{-# OPTIONS_GHC -Wno-orphans #-}
+{-# OPTIONS_GHC -Wno-unused-imports #-}
+{-# LANGUAGE TemplateHaskell #-}
+{-# LANGUAGE TypeOperators #-}
+{-# LANGUAGE TypeApplications #-}
+{-# LANGUAGE DataKinds #-}
+{-# LANGUAGE TypeFamilies #-}
+{-# LANGUAGE GADTs #-}
+{-# LANGUAGE MultiParamTypeClasses #-}
+{-# LANGUAGE UndecidableInstances #-}
+{-# LANGUAGE DeriveFunctor #-}
+{-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE FlexibleInstances #-}
+{-# LANGUAGE StandaloneDeriving #-}
+{-# LANGUAGE IncoherentInstances #-}
+
+-----------------------------------------------------------------------------
+-- |
+-- Module      :  Text.Html
+-- Copyright   :  (c) Andy Gill and OGI, 1999-2001
+-- License     :  BSD-style (see the file libraries/base/LICENSE)
+--
+-- Maintainer  :  Andy Gill <andy@galconn.com>
+-- Stability   :  provisional
+-- Portability :  portable
+--
+-- An Html combinator library
+--
+-----------------------------------------------------------------------------
+
+module Html (
+      module Html,
+      ) where
+
+
+import Control.DeepSeq
+import GHC.Generics hiding (Rep)
+
+import Test.QuickCheck hiding (Fun)
+import Test.Dragen2
+
+import qualified BlockTable as BT
+
+infixr 3 `above`  -- combining table cells
+infixr 4 `beside`  -- combining table cells
+infixr 2 `combine`  -- combining Html
+infixr 7 `nest`   -- nesting Html
+infixl 8 !    -- adding optional arguments
+
+deriving instance Generic Char
+deriving instance Generic HtmlAttr
+deriving instance Generic HtmlElement
+deriving instance Generic Html
+
+instance NFData HtmlAttr
+instance NFData HtmlElement
+instance NFData Html
+
+-- A important property of Html is that all strings inside the
+-- structure are already in Html friendly format.
+-- For example, use of &gt;,etc.
+
+data HtmlElement
+{-
+ -    ..just..plain..normal..text... but using &copy; and &amb;, etc.
+ -}
+      = HtmlString String
+{-
+ -    <thetag {..attrs..}> ..content.. </thetag>
+ -}
[…]

### 2019-04-28 bench/Main.hs

@@ -1,0 +1,38 @@
+module Main where
+
+import Criterion
+import Criterion.Main
+
+import Test.Dragen2
+import Test.QuickCheck
+
+import qualified RBT
+import qualified RE
+import qualified Html
+
+samples :: Int
+samples = 1000
+
+size :: Int
+size = 30
+
+depth :: Int
+depth = 5
+
+genSamples :: Show a => BoundedGen a -> IO [a]
+genSamples gen
+  = sequence
+  . replicate samples
+  . generate
+  . resize size
+  $ gen depth
+
+main :: IO ()
+main = defaultMain
+  [ bench "RBT/Dragen"  $ nfIO (genSamples (RBT.genTree' @Int))
+  , bench "RBT/Manual"  $ nfIO (genSamples (RBT.genTree  @Int))
+  , bench "RE/Dragen"   $ nfIO (genSamples (RE.genR'     @Char))
+  , bench "RE/Manual"   $ nfIO (genSamples (RE.genR      @Char))
+  , bench "Html/Dragen" $ nfIO (genSamples (Html.genHtml'))
+  , bench "RE/Manual"   $ nfIO (genSamples (Html.genHtml ))
+  ]

### 2019-04-28 bench/RBT.hs

@@ -1,0 +1,258 @@
+module RBT where
+
+import GHC.Generics hiding (Rep)
+import Control.DeepSeq
+
+import Test.QuickCheck hiding (Fun)
+import Test.Dragen2    hiding (empty)
+
+
+data Color = R | B deriving (Show, Generic)
+
+data Tree a = E | T Color (Tree a) a (Tree a) deriving (Show, Generic)
+
+instance NFData Color
+instance NFData a => NFData (Tree a)
+
+
+-- Invariants
+-- 1. No red node has a red parent
+-- 2. Every path from the root node to an empty node contains the same number of
+-- black nodes
+-- 3. The root and leaves of the tree are black
+
+
+-- | Simple Tree operations
+empty :: Tree a
+empty = E
+
+member :: (Ord a) => a -> Tree a -> Bool
+member _ E    = False
+member x (T _ a y b)
+  | x < y     = member x a
+  | x == y    = True
+  | otherwise = member x b
+
+-- | Insertion
+insert :: (Ord a) => a -> Tree a -> Tree a
+insert x s = makeBlack $ ins s
+  where ins E  = T R E x E
+        ins (T color a y b)
+          | x < y  = balance color (ins a) y b
+          | x == y = T color a y b
+          | x > y  = balance color a y (ins b)
+        makeBlack (T _ a y b) = T B a y b
+
+balance :: Color -> Tree a -> a -> Tree a -> Tree a
+balance B (T R (T R a x b) y c) z d = T R (T B a x b) y (T B c z d)
+balance B (T R a x (T R b y c)) z d = T R (T B a x b) y (T B c z d)
+balance B a x (T R (T R b y c) z d) = T R (T B a x b) y (T B c z d)
+balance B a x (T R b y (T R c z d)) = T R (T B a x b) y (T B c z d)
+balance color a x b = T color a x b
+
+-- This is simple helper to reduce typing
+balance' :: Tree a -> Tree a
+balance' (T color left value right) = balance color left value right
+balance' x = x
+
+-- | Deletion
+delete :: (Ord a) => a -> Tree a -> Tree a
+delete x t = makeBlack $ del x t
+  where makeBlack (T _ a y b) = T B a y b
+        makeBlack E           = E
+
+-- Delete with consecutive red nodes at the top which is rectified in delete
[…]

### 2019-04-28 bench/RE.hs

@@ -1,0 +1,199 @@
+module RE where
+
+import Control.Monad(liftM, liftM2)
+
+import GHC.Generics hiding (Rep, R, L)
+import Control.DeepSeq
+import Test.Dragen2 hiding ((.+.))
+import Test.QuickCheck hiding ((.&.), Fun)
+
+
+infix  5 ~~
+infixl 7 .+.
+infixl 8 .>.
+
+--------------------------------------------------------------------------------
+-- regular expressions with intersection
+
+data R a
+  = Nil
+  | Eps
+  | Atom a
+  | R a :+: R a
+  | R a :&: R a
+  | R a :>: R a
+  | Star (R a)
+ deriving (Eq, Ord, Show, Generic)
+
+instance NFData a => NFData (R a)
+
+(.&.) :: R a -> R a -> R a
+Nil .&.  _   = Nil
+_   .&.  Nil = Nil
+p   .&.  q   = p :&: q
+
+(.+.) :: R a -> R a -> R a
+Nil .+. q   = q
+p   .+. Nil = p
+p   .+. q   = p :+: q
+
+(.>.) :: R a -> R a -> R a
+Nil .>. _   = Nil
+_   .>. Nil = Nil
+Eps .>. q   = q
+p   .>. Eps = p
+p   .>. q   = p :>: q
+
+--------------------------------------------------------------------------------
+-- can the RE parse the empty string?
+
+eps :: R a -> Bool
+eps Eps       = True
+eps (p :+: q) = eps p || eps q
+eps (p :&: q) = eps p && eps q
+eps (p :>: q) = eps p && eps q
+eps (Star _)  = True
+eps _         = False
+
+epsR :: R a  -> R a
+epsR p | eps p     = Eps
+       | otherwise = Nil
+
+--------------------------------------------------------------------------------
+-- small step operational semantics, Brzozowsky-style
+
+step :: Eq a => R a -> a -> R a
+step (Atom a)  x
+  | a == x    = Eps
+  | otherwise = Nil
+step (p :+: q) x = step p x .+. step q x
+step (p :&: q) x = step p x .&. step q x
+step (p :>: q) x = (step p x .>. q) .+. if eps p then step q x else Nil
+step (Star p)  x = step p x .>. Star p
+step _         _ = Nil
+
+rec :: Eq a => R a -> [a] -> Bool
[…]

### 2019-04-28 src/Test/Dragen2/BoundedArbitrary.hs

@@ -26,16 +26,10 @@ newtype DepthBounded (n :: Nat) (a :: Type)
 instance (KnownNat d, BoundedArbitrary a) => Arbitrary (DepthBounded d a) where
   arbitrary = MkDepthBounded <$> boundedArbitrary (max 0 (numVal @d))
 
-
 -- | This global instance exists as a workaround of TH's stage restriction.
 -- More info coming soon.
 instance {-# OVERLAPPABLE #-} BoundedArbitrary a where
   boundedArbitrary = error "BoundedArbitrary: default dummy instance called"
-
--- | This global orphan instance exists just for convenience.
--- It uses the same size for inner calls to arbitrary and depth limit.
-instance {-# OVERLAPPABLE #-} BoundedArbitrary a => Arbitrary a where
-  arbitrary = sized boundedArbitrary
 
 ----------------------------------------
 -- | Depth-bounded arbitrary generation of parametric types

### 2019-04-28 src/Test/Dragen2/Countable.hs

@@ -1,7 +1,14 @@
 {-# LANGUAGE KindSignatures  #-}
 {-# LANGUAGE DefaultSignatures  #-}
 
-module Test.Dragen2.Countable where
+module Test.Dragen2.Countable
+  ( ConsCount
+  , ConsAvg
+  , Countable(..)
+  , GCountable
+  , defaultCount
+  , consAvg
+  ) where
 
 import GHC.Generics
 import Data.Kind

### 2019-04-28 src/Test/Dragen2/Infix.hs

@@ -11,11 +11,15 @@ infixl 6 :*
 type (:+) = Sum
 infixr 5 :+
 
-type (m :> k) = Lookup m k
-infix 4 :>
+type (m :! k) = Lookup m k
+infix 4 :!
 
-type (t :$ a) = Apply a t
-infixl 3 :$
+type (t <| a) = Apply a t
+infixl 3 <|
 
 type ((k :: kk) := (v :: vv)) = '(k, v)
 infix 2 :=
+
+(.=>) :: a -> b -> (b, a)
+(.=>) = flip (,)
+infix 2 .=>

### 2019-04-28 src/Test/Dragen2/TH.hs

@@ -3,6 +3,7 @@
 
 module Test.Dragen2.TH where
 
+import Data.Maybe
 import Data.Proxy
 
 import Control.Monad.Extra
@@ -32,29 +33,35 @@ data Target
   = TypeDef
     { typeName_  :: Name
     , blacklist_ :: Blacklist
-    , typeFam_   :: MutRecFam }
+    , typeFam_   :: MutRecFam
+    , branching_ :: Bool }
   | TypeInt
     { typeName_  :: Name
     , alias_     :: String
     , blacklist_ :: Blacklist
     , path_      :: Maybe FilePath
-    , typeFam_   :: MutRecFam }
+    , typeFam_   :: MutRecFam
+    , branching_ :: Bool }
   | FunPats
-    { funName_ :: Name
-    , argNum_  :: Int
-    , path_    :: Maybe FilePath
-    , typeFam_ :: MutRecFam }
+    { funName_   :: Name
+    , argNum_    :: Int
+    , path_      :: Maybe FilePath
+    , typeFam_   :: MutRecFam
+    , branching_ :: Bool }
 
 
 -- | Simple smart constructors
 constructors :: Name -> Target
-constructors tname = TypeDef tname [] []
+constructors tname
+  = TypeDef tname [] [] False
 
 interface :: Name -> Target
-interface tname = TypeInt tname ("(-> " ++ nameBase tname ++ ")") [] Nothing []
+interface tname
+  = TypeInt tname ("(-> " ++ nameBase tname ++ ")") [] Nothing [] False
 
 patterns :: Name -> Target
-patterns fname = FunPats fname 1 Nothing []
+patterns fname
+  = FunPats fname 1 Nothing [] False
 
 -- | Simple target modifiers
 (|>) :: Target -> (Target -> Target) -> Target
@@ -74,15 +81,17 @@ argNum n t = t { argNum_ = n }
 path :: FilePath -> Target -> Target
 path p t = t { path_ = Just p }
 
+branching :: Target -> Target
+branching t = t { branching_ = True }
 
 -- | TH target dispatcher
 deriveTarget :: Target -> Q [Dec]
-deriveTarget (TypeDef typeName _blacklist typeFam)
-  = sweeten <$> deriveTypeDef typeName _blacklist typeFam
-deriveTarget (TypeInt typeName _alias _blacklist _path typeFam)
-  = sweeten <$> deriveTypeInt typeName _alias _blacklist _path typeFam
-deriveTarget (FunPats funName _argNum _path typeFam)
[…]

### 2019-04-28 src/Test/Dragen2/TH/Extract.hs

@@ -1,6 +1,8 @@
 module Test.Dragen2.TH.Extract where
 
 import Data.List
+import Data.Maybe
+import Data.Foldable
 
 import Control.Monad.Extra
 import Control.Monad.IO.Class
@@ -49,30 +51,45 @@ defaultExts = parseExtension <$>
 extractVarNames :: Module SrcSpanInfo -> TH.Q [TH.Name]
 extractVarNames (Module _ mbModHead _ _ decs) = do
 
-  let srcNames = foldr extractBinds [] decs
+  let extractBinds (PatBind _ ps@(PVar {}) _ _) = patBindName ps
+      extractBinds (FunBind _ (ms : _)) = funBindName ms
+      extractBinds _ = return []
 
-      extractBinds (PatBind _ ps@(PVar {}) _ _) xs = patBindName ps : xs
-      extractBinds (FunBind _ ms) xs               = map funBindName ms ++ xs
-      extractBinds _ xs = xs
+      funBindName (Match _ n _ _ _) = do
+        dragenMsg "reifying name: " [rawName n]
+        maybeToList <$> TH.lookupValueName (rawName n)
+      funBindName (InfixMatch _ _ n _ _ _) = do
+        dragenMsg "reifying infix name: " [rawName n]
+        maybeToList <$> TH.lookupValueName (rawName n)
 
-      funBindName (Match _ n _ _ _)        = TH.mkName (qualify mbModName n)
-      funBindName (InfixMatch _ _ n _ _ _) = TH.mkName (qualify mbModName n)
+      patBindName (PVar _ n) = do
+        dragenMsg "reifying name: " [rawName n]
+        maybeToList <$> TH.lookupValueName (rawName n)
+      patBindName p
+        = unsupported 'extractVarNames p
 
-      patBindName (PVar _ n) = TH.mkName (qualify mbModName n)
-      patBindName p          = unsupported 'extractVarNames p
+      -- funBindName (Match _ n _ _ _)        = TH.mkName (qualify mbModName n)
+      -- funBindName (InfixMatch _ _ n _ _ _) = TH.mkName (qualify mbModName n)
+
+      -- patBindName (PVar _ n) = return [TH.mkName (qualify mbModName n)]
+      -- patBindName p          = unsupported 'extractVarNames p
+
+
+      -- qualify (Just modName) n = prettyPrint modName ++ "." ++ prettyPrint n
+      -- qualify _              n = prettyPrint n
 
[…]

### 2019-04-28 src/Test/Dragen2/TH/FunPats.hs

@@ -18,15 +18,15 @@ import qualified Test.Dragen2.Branching as Branching
 ----------------------------------------
 -- | Derive the complete representation for a every pattern of a function, plus
 -- the function representation as a sum of each pattern.
-deriveFunPats :: Name -> Int -> Maybe FilePath -> [Name] -> Q [DDec]
-deriveFunPats funName argNum path typeFam = do
+deriveFunPats :: Name -> Int -> Maybe FilePath -> [Name] -> Bool -> Q [DDec]
+deriveFunPats funName argNum path typeFam branching = do
 
   -- Parse the appropriate file to obtain the target function pattern matching
   modulePath <- maybeM currentFile pure (pure path)
   funPats <- extractDPats funName =<< parseModule modulePath
 
   -- Extract the patterns from the function for the given argument
-  (funArgsTypes, _) <- reifyFunSig funName
+  (_, _, funArgsTypes, _) <- reifyFunSig funName
   let funArgType = funArgsTypes !! (argNum - 1)
       funArgPats = dropTrivialPats ((!! (argNum - 1))  <$> funPats)
 
@@ -51,7 +51,8 @@ deriveFunPats funName argNum path typeFam = do
   patsVarsTypes <- mapM collectPatTypes funArgPats
 
   -- Create the representation for each pattern matching of the function
-  let derivePatRep' = derivePatRep funName typeFam funArgType' funArgDefTypeVars
+  let derivePatRep'
+        = derivePatRep funName typeFam funArgType' funArgDefTypeVars branching
   patReps <- concatMapM (uncurry3 derivePatRep')
                         (zip3 patsVarsTypes [1..] funArgPats)
 
@@ -62,9 +63,9 @@ deriveFunPats funName argNum path typeFam = do
 
 ----------------------------------------
 -- | Derive the complete representation for a single pattern of a function.
-derivePatRep :: Name -> [Name] -> DType -> [DTyVarBndr]
+derivePatRep :: Name -> [Name] -> DType -> [DTyVarBndr] -> Bool
              -> [DType] -> Int -> DPat -> Q [DDec]
-derivePatRep funName typeFam funArgType funArgDefTypeVars
+derivePatRep funName typeFam funArgType funArgDefTypeVars branching
              funArgPatVarsTypes patNum targetPat = do
 
   -- Create the new names of the representation and single constructor
[…]

### 2019-04-28 src/Test/Dragen2/TH/TypeDef.hs

@@ -21,8 +21,8 @@ import qualified Test.Dragen2.Branching as Branching
 ----------------------------------------
 -- | Derive a complete representation for every constructor of a type, plus
 -- the type representation as a sized sum of each constructor.
-deriveTypeDef :: Name -> [Name] -> [Name] -> Q [DDec]
-deriveTypeDef typeName blacklist typeFam = do
+deriveTypeDef :: Name -> [Name] -> [Name] -> Bool -> Q [DDec]
+deriveTypeDef typeName blacklist typeFam branching = do
 
   -- Reify the data constructors of the target data type
   allCons <- getDCons typeName
@@ -39,7 +39,7 @@ deriveTypeDef typeName blacklist typeFam = do
     dragenError "nothing to derive" [typeName]
 
   -- Derive all the stuff for each target data constructor
-  consReps <- concatMapM (deriveConRep typeFam) targetCons
+  consReps <- concatMapM (deriveConRep typeFam branching) targetCons
 
   -- Create the default `Rep` type instance
   repTypeIns <- deriveRepTypeIns typeFam typeName targetCons
@@ -50,8 +50,8 @@ deriveTypeDef typeName blacklist typeFam = do
 
 ----------------------------------------
 -- | Derive the complete representation for a single constructor of a type
-deriveConRep :: [Name] -> DCon -> Q [DDec]
-deriveConRep typeFam (DCon conTyVars conCxt conName conFields conType) = do
+deriveConRep :: [Name] -> Bool -> DCon -> Q [DDec]
+deriveConRep typeFam branching (DCon conTyVars conCxt conName conFields conType) = do
 
   -- Create the new names of the representation and single constructor
   repTypeName <- newName $ toValidIdent "Rep_Con" (nameBase conName)
@@ -101,7 +101,29 @@ deriveConRep typeFam (DCon conTyVars conCxt conName conFields conType) = do
 
       mkCxt v = DAppPr (DConPr ''QC.Arbitrary) (dTyVarBndrToDType v)
 
+  -- Create the function Rep type instance
+  let repConTyIns = DTySynInstD ''Rep.Con repInsEqn
+      repInsEqn = DTySynEqn [thNameTyLit conName] someTy
+      someTy | null conTyVars = DConT repTypeName
+             | otherwise      = thSome (length conTyVars) (DConT repTypeName)
+
   -- Create the representation BranchingType instance
[…]

### 2019-04-28 src/Test/Dragen2/TH/TypeInt.hs

@@ -19,8 +19,9 @@ import qualified Test.Dragen2.Branching as Branching
 ----------------------------------------
 -- | Derive a complete representation for every function in a module returning
 -- the target type, plus the api representation as a sized sum of each function.
-deriveTypeInt :: Name -> String -> [Name] -> Maybe FilePath -> [Name] -> Q [DDec]
-deriveTypeInt typeName alias blacklist path typeFam = do
+deriveTypeInt :: Name -> String -> [Name] -> Maybe FilePath
+              -> [Name] -> Bool -> Q [DDec]
+deriveTypeInt typeName alias blacklist path typeFam branching = do
 
   -- Parse the appropriate file to obtain interface function names
   modulePath <- maybeM currentFile pure (pure path)
@@ -35,7 +36,7 @@ deriveTypeInt typeName alias blacklist path typeFam = do
     dragenMsg "ignored constructors:" [droppedFuncs]
 
   -- Derive all the stuff for each target interface function
-  funReps <- concatMapM (deriveFunRep typeName typeFam) targetFuncs
+  funReps <- concatMapM (deriveFunRep typeFam branching) targetFuncs
 
   -- Create the default `Rep` type instance
   repTypeIns <- deriveRepTypeIns typeFam alias targetFuncs
@@ -51,16 +52,16 @@ returnsTargetType typeName funName = do
 
 
 ----------------------------------------
--- | Derive the complete representation for a single constructor of a type
-deriveFunRep :: Name -> [Name] -> Name -> Q [DDec]
-deriveFunRep typeName typeFam funName = do
+-- | Derive the complete representation for a single abstract interface function
+deriveFunRep :: [Name] -> Bool -> Name -> Q [DDec]
+deriveFunRep typeFam branching funName = do
 
   -- Create the new names of the representation and single constructor
   repTypeName <- newName $ toValidIdent "Rep_Fun" (nameBase funName)
   repConName  <- newName $ toValidIdent "Mk_Fun"  (nameBase funName)
 
   -- Reify the type signature of the function
-  (funArgTypes, funReturnType) <- reifyFunSig funName
+  (_, funCxt, funArgTypes, funReturnType) <- reifyFunSig funName
   let funRetTypeInsVars = typeArgs funReturnType
 
[…]

### 2019-04-28 src/Test/Dragen2/TH/Util.hs

@@ -6,7 +6,6 @@ module Test.Dragen2.TH.Util where
 
 import System.FilePath
 import System.Directory
-import Data.Ratio
 import Data.List
 import Control.Arrow
 import Control.Monad.Extra
@@ -15,7 +14,6 @@ import System.Console.ANSI
 
 import Test.Dragen2.TH.Ppr
 import Language.Haskell.TH hiding (Ppr, ppr, pprint)
-import Language.Haskell.TH.PprLib
 import Language.Haskell.TH.Desugar
 
 import qualified Data.Vector as Vector
@@ -58,11 +56,12 @@ reifyName name = do
     Nothing -> dragenError "could not reify name" [name]
 
 -- | Retrieve a function type signature
-reifyFunSig :: Name -> Q ([DType], DType)
+reifyFunSig :: Name -> Q ([DTyVarBndr], [DPred], [DType], DType)
 reifyFunSig funName = do
   DVarI _ funTy _ <- reifyName funName
-  let funSig = splitSignature funTy
-  return (init funSig, last funSig)
+  -- let funSig = splitSignature funTy
+  -- return (init funSig, last funSig)
+  return (unravel funTy)
 
 -- | Reify the constructors of a given type name
 getDCons :: Name -> Q [DCon]
@@ -255,7 +254,7 @@ dropTrivialPats (p : ps) = p : dropTrivialPats ps
 -- | Transform a pattern into an expression using a list of variables names for
 -- the pattern variables
 dPatToDExpWithVars :: [Name] -> DPat -> DExp
-dPatToDExpWithVars vs = fst . go vs
+dPatToDExpWithVars vars = fst . go vars
   where
     go vs (DLitPa l) = (DLitE l, vs)
     go vs (DVarPa _) = (DVarE (head vs), tail vs)
@@ -517,13 +516,13 @@ dragenWarning msg inputs = runIO $ do
     forM_ inputs $ \i -> do
       mapM_ dragenLog' (lines (dump i))
 
-dragenError :: Ppr a => String -> [a] -> Q b
+dragenError :: Show a => String -> [a] -> Q b
 dragenError msg inputs = runIO $ do
   withColor Red $ do
     dragenLog "an error happened:"
     dragenLog msg
     dragenLog "input was:"
     forM_ inputs $ \i -> do
-      mapM_ dragenLog' (lines (dump i))
+      mapM_ dragenLog' (lines (show i))
 
   error "DRAGEN derivation error"

### 2019-04-28 src/Test/Dragen2/TypeLevel.hs

@@ -20,7 +20,7 @@ type family
   Lookup ('(k, v) : m) k = v
   Lookup (_       : m) k = Lookup m k
   Lookup '[]           k
-    = TypeError (Text "Lookup: key not found" :<>: ShowType k)
+    = TypeError (Text "Lookup: key not found " :<>: ShowType k)
 
 type family
   Ix (map :: [(k, v)]) (key :: k) :: Nat where

### 2019-04-29 papers/haskell19/6-casestudies.tex
Added case studies

--- 2019-03-27 papers/haskell19/5-casestudies.tex
@@
 Nam euismod tellus id erat.
+
+\begin{tikzpicture}
+  \begin{axis}
+    [ ybar
+    , enlargelimits=0.15
+    , enlarge x limits=0.3
+    , /pgf/bar width=15pt
+    , legend style={
+        at={(0.5,-0.15)},
+        anchor=north,legend columns=3}
+    , ylabel={Relative Generation Time}
+    , symbolic x coords={SExp, RBT, Html}
+    , xtick=data
+    % , nodes near coords
+    , nodes near coords align={vertical}
+    ]
+    \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
+    \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
+    \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
+    \legend{Manual, Modular/Linear, Modular/Balanced}
+\end{axis}
+\end{tikzpicture}

### 2019-04-29 bench/.#Html.hs

@@ -1,0 +1,1 @@
+agustin@palpatine.7239:1555097683

### 2019-04-29 bench/Html/BlockTable.hs

@@ -3,7 +3,7 @@
 -- Module      :  Text.Html.BlockTable
 -- Copyright   :  (c) Andy Gill and OGI, 1999-2001
 -- License     :  BSD-style (see the file libraries/base/LICENSE)
--- 
+--
 -- Maintainer  :  Andy Gill <andy@galconn.com>
 -- Stability   :  provisional
 -- Portability :  portable
@@ -12,20 +12,20 @@
 --
 -----------------------------------------------------------------------------
 
-module BlockTable (
+module Html.BlockTable (
 
 -- Datatypes:
 
       BlockTable,             -- abstract
 
--- Contruction Functions: 
+-- Contruction Functions:
 
       single,
       empty,
       above,
       beside,
 
--- Investigation Functions: 
+-- Investigation Functions:
 
       getMatrix,
       showsTable,
@@ -47,11 +47,11 @@ infixr 3 `above`
 					|Hello|
 	  This is a 1x1 cell		+-----+
 	  Note: single has type
-	 
+
 		single :: a -> BlockTable a
-	
+
 	  So the cells can contain anything.
-	
+
 	> table2 :: BlockTable String
 	> table2 = single "World"	+-----+
 					|World|
@@ -63,7 +63,7 @@ infixr 3 `above`
 					|Hello%World|
 	 % is used to indicate		+-----%-----+
 	 the join edge between
-	 the two Tables.  
+	 the two Tables.
 
 	> table4 :: BlockTable String
 	> table4 = table3 %/% table2	+-----+-----+
@@ -80,14 +80,14 @@ infixr 3 `above`
 	  leftmost cell, again to	|     %World      |
 	  force the table to be a	+-----%-----------+
 	  rectangle.
- 
+
    Now the table can be rendered with processTable, for example:
 	Main> processTable table5
 	[[("Hello",(1,2)),
 	  ("Hello",(1,1)),
 	  ("World",(1,1))],
 	 [("World",(2,1))]] :: [[([Char],(Int,Int))]]
-	Main> 
+	Main>
 -}
 
 -- ---------------------------------------------------------------------------
@@ -143,9 +143,9 @@ t1 `beside` t2 = combine t1 t2 (\ lst1 lst2 r ->
 trans :: BlockTable a -> BlockTable a
 trans (Table f1 x1 y1) = Table (flip f1) y1 x1
 
-combine :: BlockTable a 
-      -> BlockTable b 
[…]

### 2019-04-29 bench/Html/Html.hs

@@ -28,8 +28,8 @@
 --
 -----------------------------------------------------------------------------
 
-module Html (
-      module Html,
+module Html.Html (
+      module Html.Html,
       ) where
 
 
@@ -39,7 +39,7 @@ import GHC.Generics hiding (Rep)
 import Test.QuickCheck hiding (Fun)
 import Test.Dragen2
 
-import qualified BlockTable as BT
+import qualified Html.BlockTable as BT
 
 infixr 3 `above`  -- combining table cells
 infixr 4 `beside`  -- combining table cells
@@ -1316,6 +1316,149 @@ derive
 
 type Html_S =
   '[ "Html"
+       := Con' "Html"
+       :+ Fun' "area"
+       :+ Fun' "basefont"
+       :+ Fun' "br"
+       :+ Fun' "hr"
+       :+ Fun' "image"
+       :+ Fun' "input"
+       :+ Fun' "meta"
+       :+ Fun' "param"
+       :+ Fun' "thebase"
+       :+ Fun  "address"
+       :+ Fun  "anchor"
+       :+ Fun  "applet"
+       :+ Fun  "big"
+       :+ Fun  "blockquote"
+       :+ Fun  "body"
+       :+ Fun  "bold"
+       :+ Fun  "caption"
+       :+ Fun  "center"
+       :+ Fun  "cite"
+       :+ Fun  "ddef"
+       :+ Fun  "define"
+       :+ Fun  "dlist"
+       :+ Fun  "dterm"
+       :+ Fun  "emphasize"
+       :+ Fun  "fieldset"
+       :+ Fun  "font"
+       :+ Fun  "form"
+       :+ Fun  "frame"
+       :+ Fun  "frameset"
+       :+ Fun  "h1"
+       :+ Fun  "h2"
+       :+ Fun  "h3"
+       :+ Fun  "h4"
+       :+ Fun  "h5"
+       :+ Fun  "h6"
+       :+ Fun  "header"
+       :+ Fun  "italics"
+       :+ Fun  "keyboard"
+       :+ Fun  "legend"
+       :+ Fun  "li"
+       :+ Fun  "noframes"
+       :+ Fun  "olist"
+       :+ Fun  "option"
+       :+ Fun  "paragraph"
+       :+ Fun  "pre"
+       :+ Fun  "samp"
+       :+ Fun  "select"
[…]

### 2019-04-29 bench/Lisp/Lisp.hs

@@ -1,0 +1,149 @@
+module Lisp.Lisp where
+
+import GHC.Generics hiding (Rep)
+import Control.DeepSeq
+import Test.Dragen2
+import Test.QuickCheck hiding ((.&.), Fun)
+
+newtype Symbol = Symbol String
+  deriving (Show, Eq, Generic)
+
+data SExpr
+  = Cons SExpr SExpr  -- ^ Appending list and list
+  | Null               -- ^ A representation of empty list
+  | AtomInt Int       -- ^ A pattern of the atom for `Int`
+  | AtomBool Bool     -- ^ A pattern of the atom for `Bool`
+  | AtomSymbol Symbol -- ^ A pattern of the atom for `MaruSymbol`
+  | Quote SExpr       -- ^ Delays the evaluation of a 'SExpr'
+  deriving (Show, Eq, Generic)
+
+instance NFData Symbol
+instance NFData SExpr
+
+execute :: SExpr -> ()
+execute (Cons (AtomSymbol (Symbol "let"))
+         (Cons (Cons (AtomSymbol var) (Cons val Null))
+          (Cons body Null))) = ()
+execute (Cons (AtomSymbol (Symbol "if"))
+          (Cons cond (Cons x (Cons y Null)))) = ()
+execute (Cons (Cons (AtomSymbol (Symbol "defun")) (Cons params (Cons body Null))) args) = ()
+execute (Cons (AtomSymbol (Symbol "print")) s) = ()
+execute (Cons (AtomSymbol (Symbol "list")) s) = ()
+
+constantFold :: SExpr -> ()
+constantFold (AtomSymbol (Symbol "+")) = ()
+constantFold (AtomSymbol (Symbol "-")) = ()
+constantFold (AtomSymbol (Symbol "*")) = ()
+constantFold (AtomSymbol (Symbol "/")) = ()
+
+----------------------------------------
+-- Size
+
+lispSize :: SExpr -> Int
+lispSize Null = 1
+lispSize (AtomInt _) = 1
+lispSize (AtomBool _) = 2
+lispSize (AtomSymbol _) = 2
+lispSize (Quote x) = 1 + lispSize x
+lispSize (Cons x y) = 1 + lispSize x + lispSize y
+
+----------------------------------------
+-- Manual generator
+
+genSExpr :: BoundedGen SExpr
+genSExpr d = go d
+  where
+    go 0 = frequency
+      [ (1, pure Null)
[…]

### 2019-04-29 bench/Main.hs

@@ -2,13 +2,15 @@ module Main where
 
 import Criterion
 import Criterion.Main
+import Criterion.Types
 
 import Test.Dragen2
 import Test.QuickCheck
 
-import qualified RBT
-import qualified RE
-import qualified Html
+import qualified RBT.RBT   as RBT
+import qualified RE.RE     as RE
+import qualified Html.Html as Html
+import qualified Lisp.Lisp as Lisp
 
 samples :: Int
 samples = 1000
@@ -28,11 +30,17 @@ genSamples gen
   $ gen depth
 
 main :: IO ()
-main = defaultMain
-  [ bench "RBT/Dragen"  $ nfIO (genSamples (RBT.genTree' @Int))
-  , bench "RBT/Manual"  $ nfIO (genSamples (RBT.genTree  @Int))
-  , bench "RE/Dragen"   $ nfIO (genSamples (RE.genR'     @Char))
-  , bench "RE/Manual"   $ nfIO (genSamples (RE.genR      @Char))
-  , bench "Html/Dragen" $ nfIO (genSamples (Html.genHtml'))
-  , bench "RE/Manual"   $ nfIO (genSamples (Html.genHtml ))
+main = defaultMainWith (defaultConfig { verbosity = Verbose })
+  [ 
+  -- bench "RBT/Manual"      $ nfIO (genSamples (RBT.genTree   @Int))
+  -- , bench "RBT/Dragen"      $ nfIO (genSamples (RBT.genTree'  @Int))
+   bench "RBT/Dragen/Bal"  $ nfIO (genSamples (RBT.genTree'' @Int))
+  -- , bench "Html/Manual"     $ nfIO (genSamples (Html.genHtml ))
+  -- , bench "Html/Dragen/Bal" $ nfIO (genSamples (Html.genHtml'))
+  -- , bench "Lisp/Manual"     $ nfIO (genSamples (Lisp.genSExpr))
+  -- , bench "Lisp/Dragen"     $ nfIO (genSamples (Lisp.genSExpr'))
+  -- , bench "Lisp/Dragen/Bal" $ nfIO (genSamples (Lisp.genSExpr''))
+  -- , bench "RE/Manual"       $ nfIO (genSamples (RE.genR       @Char))
+  -- , bench "RE/Dragen"       $ nfIO (genSamples (RE.genR'      @Char))
+  -- , bench "RE/Dragen/Bal"   $ nfIO (genSamples (RE.genR''     @Char))
   ]

### 2019-04-29 bench/RBT/RBT.hs

@@ -1,4 +1,4 @@
-module RBT where
+module RBT.RBT where
 
 import GHC.Generics hiding (Rep)
 import Control.DeepSeq
@@ -134,9 +134,6 @@ treeSize (T _ l _ r) = 2 + treeSize l + treeSize r
 ----------------------------------------
 -- Manual generator
 
-genColor :: Gen Color
-genColor = elements [B, R]
-
 genTree :: (Arbitrary a, Ord a) => BoundedGen (Tree a)
 genTree depth = go depth
   where
@@ -148,7 +145,7 @@ genTree depth = go depth
       [ -- Con "E"
         (1, pure E)
         -- Con "T"
-      , (1, T <$> genColor <*> go (n-1) <*> arbitrary <*> go (n-1))
+      , (1, T <$> arbitrary <*> go (n-1) <*> arbitrary <*> go (n-1))
         -- Fun' "empty"
       , (1, pure empty)
         -- Fun  "insert"
@@ -222,37 +219,49 @@ genTree depth = go depth
 ----------------------------------------
 -- Dragen2
 
-genTree' :: (Ord a, Arbitrary a) => BoundedGen (Tree a)
-genTree' = boundedArbitrary
+instance Arbitrary Color where
+  arbitrary = elements [R, B]
 
-derive [''Color, ''Tree]
-  [ constructors ''Color
-  , constructors ''Tree
-  , interface    ''Tree  |> blacklist ['balance]
+derive [''Tree]
+  [ constructors ''Tree
+  , interface    ''Tree |> blacklist ['balance]
   , patterns     'balL
   , patterns     'balR
   ]
 
-type RBT_Spec =
-  '[ "Color"
-       := Rep "Color"
-   , "Tree"
-       := (((Con' "E"
-       :+ Con  "T")
-       :+ (Fun' "empty"
-       :+ Fun  "insert"))
-       :+ ((Fun  "balance'"
-       :+ Fun  "delete")
-       :+ (Fun  "fuse"
-       :+ Pat  "balL" 1)))
-       :+ (((Pat  "balL" 2
-       :+ Pat  "balL" 3)
-       :+ (Pat  "balR" 1
[…]

### 2019-04-29 bench/RE/RE.hs

@@ -1,4 +1,4 @@
-module RE where
+module RE.RE where
 
 import Control.Monad(liftM, liftM2)
 
@@ -195,5 +195,21 @@ type R_Spec
   :+ Fun  ".>."
   :+ Fun  "epsR"
 
+type R_Spec_Bal
+  =  (((Con' "Nil"
+  :+ Con' "Eps")
+  :+ (Con' "Atom"
+  :+ Con  ":+:"))
+  :+ ((Con  ":&:"
+  :+ Con  ":>:")
+  :+ (Con  "Star"
+  :+ Fun  ".&.")))
+  :+ (((Fun  ".+."
+  :+ Fun  ".>.")
+  :+ Fun  "epsR"))
+
 genR' :: forall a. (Ord a, Arbitrary a) => BoundedGen (R a)
 genR' = genEval @(R_Spec <| a)
+
+genR'' :: forall a. (Ord a, Arbitrary a) => BoundedGen (R a)
+genR'' = genEval @(R_Spec_Bal <| a)

### 2019-04-29 papers/haskell19/6-casestudies.tex
WIP case studies

@@
 
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat.
+The random generation framework presented throughout this work allows us to
+write extensible generators in a very concise way.
+%
+However, this expresiveness comes attached to a perceptible runtime overhead.
+%
+In this section we evaluate the implicit cost of composing generators using
+three real-world case studies.
 
+As we have shown in Section \ref{sec:generators}, the random generation
+process we propose in this paper can be seen as having two phases.
+%
+First, generate random values from the representation types used to specify the
+shape of our data; and then we use their algebras to translate them to the
+corresponding values of our target data types.
+%
+In particular, the tranformation step is expected to pattern match repeatedly
+against the |InL| and |InR| constructors of the |oplus| operators when
+traversing each construction injection.
+%
+Because of this, in general we expect a performance impact with respect to
+manually-written concrete generators.
+
+As recently analyzed by \citeauthor{KiriyamaOptimizingDTC}, this slowdown is
+expected to be linear in the depth of our representation type
+\cite{KiriyamaOptimizingDTC}.
+%
+Hence, one can drastically reduce the runtime overhead by associating each
+|oplus| operator in a balanced fashion.
+%
+So, for instance, instead of writing |(f oplus'' g oplus'' h oplus'' i)|, which
+is implicitly parsed as |(f oplus'' (g oplus'' (h oplus'' i)))|; we can
+associate constructions as |((f oplus'' g) oplus'' (h oplus'' i))|, thus
+reducing the depth of our representation from four to three levels and, in
+general, from a $\mathcal{O}(n)$ to a $\mathcal{O}(log(n))$ complexity in the
+runtime overhead, where $n$ is the amount of constructions under consideration.
+%
+To apply this optimization automatically, and relying on more sophisticated
+techniques like prefix-free encodings to obtain optimal representations, is a
+undergoing path of our future work.
+
+To evaluate our approach, we analyzed the performance of generating random
+values using three real-world case studies:
+\begin{inparaenum}[(i)]
+\item Red-Black Trees, inspired by Okasaki's formulation,
+\item Lisp S-expressions, inspired by the package \emph{hs-zuramaru}, and
+\item HTML expressions, inspired by the \emph{html} package, and following the
+  same structure as our motivating |Html| example.
+\end{inparaenum}
+%
+The expanse of each case study can be summarized as follows:
+
+\begin{table}[H]
+  \begin{tabular}{||c||c||c||c||c||}
+    \hline
+    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
+    \hline
+    RBT        & 2      & 6      & 5      & 13                  \\ \hline
+    SExp       & 6      & 9      & -      & 15                  \\ \hline
+    Html       & 4      & -      & 132    & 136                 \\ \hline
+  \end{tabular}
+\end{table}
+
+For each case study, we benchmarked the execution time required to generate 1000
+random values using both manually-written concrete generators, and modular
+generators obtained using our approach.
+%
+Additionally, our modular generators were tested using both linear and balanced
+generation specifications.
+%
+Figure \ref{fig:times} illustrates the relative execution time of each case
+study, with respect to each manually-written random generator---we encourage the
+reader to obtain a colored version of this work.
+
+\begin{figure}[b]
+  \centering
   \begin{tikzpicture}
     \begin{axis}
-    [ ybar
-    , enlargelimits=0.15
-    , enlarge x limits=0.3
-    , /pgf/bar width=15pt
+      [ height=6cm
+      , width=9cm
+      , ybar
+      , ymajorgrids=true
+      , ytick={0,1,...,5}
+      , enlargelimits=0.25
+      , enlarge x limits=0.25
+      , /pgf/bar width=17pt
       , legend style={
@@
         anchor=north,legend columns=3}
+      , legend image code/.code={%
+        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
       , ylabel={Relative Generation Time}
-    , symbolic x coords={SExp, RBT, Html}
+      , symbolic x coords={RBT, SExp, Html}
       , xtick=data
@@
   \end{tikzpicture}
+  \caption{Generation time comparison between concrete and modular generators.}
+  \label{fig:times}
+\end{figure}

### 2019-04-30 papers/haskell19/6-casestudies.tex
Case studies

@@
 runtime overhead, where $n$ is the amount of constructions under consideration.
-%
-To apply this optimization automatically, and relying on more sophisticated
-techniques like prefix-free encodings to obtain optimal representations, is a
-undergoing path of our future work.
 
@@
     SExp       & 6      & 9      & -      & 15                  \\ \hline
-    Html       & 4      & -      & 132    & 136                 \\ \hline
+    HTML       & 4      & -      & 132    & 136                 \\ \hline
   \end{tabular}
@@
 
-For each case study, we benchmarked the execution time required to generate 1000
-random values using both manually-written concrete generators, and modular
-generators obtained using our approach.
+We benchmarked the execution time required to generate and fully evaluate 1000
+random values corresponding to each case study, comparing both manually-written
+concrete generators and modular generators obtained using our approach.
+%
+For this purpose, we used the \emph{Criterion} benchmarking tool for Haskell,
+and limited the maximum depth of the generated values to five levels.
 %
@@
       , ytick={0,1,...,5}
+      , ymax=4.5
       , enlargelimits=0.25
@@
       , ylabel={Relative Generation Time}
-      , symbolic x coords={RBT, SExp, Html}
+      , symbolic x coords={RBT, SExp, HTML}
       , xtick=data
@@
       ]
-      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
-      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
-      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
+      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (HTML,1)};
+      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (HTML,4.88)};
+      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (HTML,1.70)};
       \legend{Manual, Modular/Linear, Modular/Balanced}
@@
 \end{figure}
+
+As it can be observed, our approach suffers from a noticeable runtime overhead,
+specially when considering the HTML case study with a large amount of
+construction involved at the generation process.
+%
+However, by balancing our representation types we found that the generation
+performance gets dramatically improved.
+%
+To apply this optimization automatically, and relying on more sophisticated
+techniques like prefix-free encodings to obtain optimal representations, is a
+undergoing path of our future work.
+
+On the other hand, it has been argued that the generation time is often not
+substantial with respect to the rest of the testing process, especially when
+testing complex properties over monadic code, as well as using randomly
+generated values for penetration testing.
+
+Nonetheless, we consider that these results are in turn very encouraging, given
+that the flexibility obtained from using our approach does not produce severe
+slowdowns in practice, nor explodes more than linearly as we increase the
+complexity of our random generators.

### 2019-04-30 papers/haskell19/1-intro.tex
Reorganized sections and code
Section 3
Added contributions

--- 2019-03-20 papers/haskell19/introduction.tex
@@
 \section{Introduction}
-\label{sec:introduction}
+\label{sec:intro}
 
@@
 
-\begin{code}
-data Html
-  =  Text  String
-  |  Sing  String
-  |  Tag   String  Html
-  |  Join  Html    Html
+The main contribution of this paper are:
+%
+\begin{itemize}
+\item We identify some of the most important limitations of performing random
+  testing with automatically derived generators using state-of-the-art tools
+  (Section \ref{sec:sources}).
+\item We present an extensible mechanism for representing random values built
+  upon different source constructions using the Data Types \`a la Carte approach
+  (Section \ref{sec:representation}).
+\item We develop a modular generation schema, exploiting the extensible nature
+  of our representations, which are extended to encode information relevant to
+  the generation process at the type level (Section \ref{sec:generation}).
+\item We present a very simple yet powerful domain specific language for
+  describing extensible generators solely based on the types we used to
+  represent the desired shape of our random data (Section \ref{sec:type-level}).
+\item We provide a Template Haskell tool for automatically deriving all the
+  required machinery presented throughout this paper, and evaluate its
+  generation performance using three real-world case studies (Section
+  \ref{sec:casestudies}).
+\end{itemize}
 
-hr :: Html
-hr = Sing "hr"
-
-bold :: Html -> Html
-bold x = Tag "b" x
-
-div :: Html -> Html
-div x = Tag "div" x
-\end{code}
+Overall, we present a novel technique for reusing automatically derived
+generators in a composable fashion, in contrast with the usual paradigm of
+automatically deriving ``closed'' monolithic generators.

### 2019-04-30 papers/haskell19/1-intro.tex
WIP section 2

@@
 \item We present an extensible mechanism for representing random values built
-  upon different source constructions using the Data Types \`a la Carte approach
-  (Section \ref{sec:representation}).
-\item We develop a modular generation schema, exploiting the extensible nature
+  upon different source constructions using Data Types \`a la Carte (Section
+  \ref{sec:representation}).
+\item We develop a modular generation schema, exploiting the composable nature
   of our representations, which are extended to encode information relevant to
-  the generation process at the type level (Section \ref{sec:generation}).
+  the generation process at the type level (Section \ref{sec:generators}).
 \item We present a very simple yet powerful domain specific language for
-  describing extensible generators solely based on the types we used to
-  represent the desired shape of our random data (Section \ref{sec:type-level}).
+  describing extensible generators solely based on the types used to represent
+  the desired shape of our random data (Section \ref{sec:typelevel}).
 \item We provide a Template Haskell tool for automatically deriving all the

### 2019-04-30 papers/haskell19/2-sources.tex
Section 4
Added contributions
WIP section 2

@@
-\section{Random Testing using Hints}
+\section{Random Testing with Source Hints}
 \label{sec:sources}
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+This section introduces the common workflow for generating random values using
+automatically derived QuickCheck generators, along with its common limitations.
+%
+To illustrate this, let us consider the following data type definition encoding
+basic HTML pages:
 
@@
 
-A value of type |Html| represents either plain text strings, via the |Text| data
-constructor; single and paired HTML tags using the |Sing| and |Tag| data
-constructors, respectively; and concatenation of ...
+This type allows us to build HTML pages via four possible data constructors:
+|Text| is used for plain text values; |Sing| and |Tag| represent singular and
+paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
+one after another.
+%
+Note that the constructors |Tag| and |Join| are recursive, as they have at least
+one field of their own type.
+%
+In this paper, and in the scope of random generation in general, we will refer
+to any non-recursive data constructor as \emph{terminal}, and
+\emph{non-terminal} in any other case.
+
+Then, using this representation the example page\linebreak
+%
+\texttt{<html>hello<hr><b>world!</b></html>}
+%
+can be encoded simply as:
+%
+\begin{code}
+Tag "html" $  Text "hello"
+              iJoin Single "hr"
+              iJoin Tag "b" $ Text "world!"
+\end{code} %$
+
+Then, in order to generate random values of type |Html|, QuickCheck requires us
+to provide a random data generator for it.
 
@@
 list :: [Html] -> Html
-list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))
+list []  = Text "empty list"
+list xs  = Tag "ul" $ foldl1 Join (Tag "li" <$> xs)
 
@@
 \begin{code}
+  gen d = frequency
+    [ ...
+    , (1, pure br)
+    , (5, bold   <$> gen (d-1))
+    , (2, list   <$> listOf (gen (d-1)))
+    , (3, (<+>)  <$> gen (d-1) <*> gen (d-1)) ]
+\end{code}%$
+
+\begin{code}
 simplify :: Html -> Html
-simplify (Join (Text t1) (Text t2))
+simplify (Text t1 iJoin Text t2)
   = Text (t1 ++ t2)
-simplify (Join (Join (Text t1) x) y)
-  = simplify (Join (Text t1) (simplify (Join x y)))
-simplify (Join x y)
-  = Join (simplify x) (simplify y)
+simplify ((Text t1 iJoin x) iJoin y)
+  = simplify (Text t1 iJoin simplify (x iJoin y))
+simplify (x iJoin y)
+  = simplify x iJoin simplify y
 simplify (Tag t x)
@@
 \end{code}
+
+\begin{code}
+  gen d = frequency
+    [ ...
+    , (2, do  t1  <- arbitrary
+              t2  <- arbitrary
+              return (Text t1 iJoin Text t2))
+    , (4, do  t2  <- arbitrary
+              x   <- go (d-1)
+              y   <- go (d-1)
+              return ((Text t1 iJoin x) iJoin y))]
+\end{code}%$

### 2019-05-01 papers/haskell19/2-sources.tex
Section 2

@@
 
+% --------------------------------------
+% Motivating Example
+%
 This section introduces the common workflow for generating random values using
@@
 
+% --------------------------------------
+% Random Generators
+%
 Then, in order to generate random values of type |Html|, QuickCheck requires us
 to provide a random data generator for it.
+%
+For this purpose, QuickCheck defines the |Arbitrary| type-class for default
+defining random generators as:
 
@@
 \end{code}
-
-\begin{code}
-class Arbitrary (a :: *) where
-  arbitrary :: Gen a
-\end{code}
+%
+where the overloaded symbol |arbitrary :: Gen a| denotes a monadic generator for
+values of type |a|.
+%
+Using this mechanism, we can define a sensible random generator for our |Html|
+data type as follows:
 
@@
 \end{code} %$
+
+At the top level, this definition parametrizes the generation process using
+QuickCheck's |sized| combinator, which lets us build our generator using an
+auxliary local function |gen :: Int -> Gen Html|.
+%
+The |Int| passed to |gen| is known as the \emph{generation size}, and is
+threaded seamlessly by QuickCheck on each call to arbitrary.
+%
+Then, we use this parameter to limit the maximum amount of recursive calls that
+our generator can perform, and thus the maximum depth of the generated values.
+%
+If the generation size is positive (case |gen d|), our random generator picks a
+random constructor from our |Html| data type with a given generation frequency
+(using QuickCheck's |frequency| combinator), and proceeds to fill its fields
+using randomly generated sub-terms---here using Haskell's applicative notation
+and the default |Arbitrary| instance for |String|s.
+%
+For the case of the recursive sub-terms, this generator simply calls itself
+recursively with a smaller generation size (|gen (d-1)|).
+%
+This process repeats until we reach the base case (|gen 0|) on each recursive
+sub-term.
+%
+At this point, our random generator is limited to only pick among the |Html|
+terminal constructors, hence terminating the generation process.
+
+% --------------------------------------
+% Automatic derivation tools
+%
+As one can observe, the previous definition is quite mechanical, and depends
+only on the generation frequencies we choose for each constructor.
+%
+Fortunately, there exists a variety of meta-programming tools for obtaining
+random generators automatically based on their types.
+%
+Tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random generators
+using the same procedure as the one shown above.
+%
+\emph{MegaDeTH} simply picks among data constructors uniformingly, whereas
+\emph{DRAGEN} analyzes the type definitions and tunes the generation frequencies
+based on a simulation-based optimization approach, in order to obtain random
+distributions of values based on the users' desires.
+
+% --------------------------------------
+% Limitation
+%
+These tools provide a useful mechanism for automating the boring and error-prone
+task of writing random generators by hand.
+%
+However, this simple derivation procedure which only considers the structural
+information encoded on data type definitions, and generates one single
+constructor at a time; is often too generic to synthesize useful generators in
+some practical scenarios.
+%
+For instance, because of the types we use for deriving our random generators are
+too lax or, alternatively, because our properties need to be tested against very
+specific patterns of input cases in order to trigger bugs.
+
+This problem, however, can be alleviated by considering additional sources of
+structural information from our codebase other than just data constructors.
+%
+We proceed to illustrate this idea for consideting the structural information
+found on abstract interfaces and functions' pattern matchings.
+
+% --------------------------------------
+% Abstract interfaces
+%
+\paragraph{Extracting Information From Interfaces}
+
+A common choice when implementing abstract data types is to transfer the
+resposiblity of preserving their invariants to the functions on their abstract
+interface.
+%
+Take for example our |Html| data type.
+%
+Instead of defining a different data constructor for each possible HTML tag, we
+opted for a very small generic representation that can be extended via a set of
+combinators making use of this generic interface:
 
@@
 
+These functions let us create interesting structured values representing HTML
+constructions like linebreaks (|br|), bold blocks (|bold|), unordered lists
+(|list|) and concatenation of values one below another (|<+>|).
+
+Note how difficult would be to generate random values containing, for example,
+structurally valid HTML lists if we only consider the structural information
+encoded in our |Html| data type definion.
+%
+After all, much of the valid structure of HTML has been encoded on its abstract
+interface.
+
+An automatically derived random generator could integrate this structural
+information into the generation process by simply creating random values arising
+from applying such functions using randomly generated inputs:
+
 \begin{code}
@@
 
+where the symbol (|...|) represents the rest of the possible sub-generators for
+each data constructor seen before.
+%
+From now on, we will refer to each possible choice given to |frequency| as a
+different \emph{random construction}, as we are not considering only generating
+only single data constructors anymore.
+
+% --------------------------------------
+% Functions' pattern matchings
+%
+\paragraph{Extracting Information From Patterns}
+
+A different challenge appears when we try to test functions comprising complex
+pattern matchings. Consider for instance the following function for simplifying
+|Html| values:
+
 \begin{code}
 simplify :: Html -> Html
-simplify (Text t1 iJoin Text t2)
-  = Text (t1 ++ t2)
+simplify (Text t1 iJoin Text t2) = Text (t1 ++ t2)
 simplify ((Text t1 iJoin x) iJoin y)
   = simplify (Text t1 iJoin simplify (x iJoin y))
-simplify (x iJoin y)
-  = simplify x iJoin simplify y
-simplify (Tag t x)
-  = Tag t (simplify x)
+simplify (x iJoin y) = simplify x iJoin simplify y
+simplify (Tag t x) = Tag t (simplify x)
 simplify x = x
 \end{code}
+
+This function simply traverses |Html| values, joining together every contiguous
+pair of |Text| constructors.
+%
+Ideally, we would like to put approximately the same testing effort info each
+clause of |simplify|, or perhaps even more to first two ones, since those are
+the ones performing actual simplifications.
+%
+However, these two clauses are the most difficult ones to test in practice.
+%
+The probabilty of generating a random value satisfying nested patterns decreases
+multiplicatively with the number of constructors we match against to
+simultaneously.
+%
+In our tests, we were not able to exercise any of these two patterns more than
+6\% of the overall testing time (and considering recursive sub-calls), using
+random generators derived using either \emph{MegaDeTH} and \emph{DRAGEN}.
+%
+As expected, most of the random test cases were exercising the simplest (an
+rather uninteresting) patterns of this function.
+
+To solve this issue, we could opt to consider each complex pattern as a new kind
+of random construction.
+%
+In this light, we can simply generate values satisfying pattern directly by
+returning its corresponding expression, where each variable or wildcard pattern
+is filled using a random sub-expression:
 
@@
 \end{code}%$
+
+% --------------------------------------
+% Wrapping up
+%
+In general, we think the solution we is far from ideal, as in many cases we
+would like to select which random constructions are of our interest when testing
+different properties or sub-systems.
+%
+In this light, having ``monolithic'' random generators derived automatically
+would complicate our desired workflow.
+%
+We would need to derive different concrete random generators for each use case
+instead of having a single |Arbitrary| instance that can be used everywhere.
+%
+In contrast, in this work we opted for developing an extensible fashion, where
+the generation machinery is automatically derived only once, and every variant
+of our random generators is expressed at the type-level.
+%
+The following sections of this paper cover our alternative approach in depth.

### 2019-05-01 papers/haskell19/8-conclusions.tex
WIP conclusions

@@
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+We presented a novel approach for automatically deriving flexibe random
+generators inspired by the seminal work on Data Types \`a la Carte.
+%
+Furthermore, we incorporate valuable structural information into our generation
+process by considering not only data constructors but also the information
+statically available in abstract interfaces and functions' pattern matchings.
+
+In the future, something something...

### 2019-05-01 papers/haskell19/abstract.tex
WIP conclusions

--- 2019-01-15 paper/sections/introduction.tex
@@
-\section{Introduction}
-
 Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
@@
 commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
 tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
@@
 Nam vestibulum accumsan nisl.
-
-Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
-commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
-lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
-felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
-dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
-quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
-Nam euismod tellus id erat\cite{grieco2017}.
-
-The main contribuitions of this paper are:
-%
-\begin{itemize}
-  %
-\item We identify two patological scenarios for which standard type-driven
-  automatic derivation tools fail to synthesize practical random generators, due
-  to a lack of either type structure or domain knowleadge (Section
-  \ref{sec:randomtesting}).
-  %
-\item We present a generation technique able to encode stronger properties of
-  the target data by reifying the static information present on the program
-  codebase (Section \ref{sec:hrep}).
-  %
-\item We apply and extend the theory of branching processes to analitically
-  predict the average distribution of generated values.
-  %
-  Furthermore, we use the predictions to perform simulation-based optimization
-  of the random generation parameters (Section \ref{sec:synthesis}).
-  %
-\item We provide an implementation of our ideas in the form of a Haskell library
-  to perform automatic derivation of random generators capable to extract
-  useful structure information from the user source code.
-  %
-\end{itemize}

### 2019-05-02 papers/haskell19/7-related.tex
Added file
WIP conclusions
Related work

--- 2019-03-20 papers/haskell19/relatedwork.tex
@@
 \section{Related Work}
-\label{sec:relatedwork}
+\label{sec:related}
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+\citeauthor{SwierstraDTC} \cite{SwierstraDTC} proposes the Data Types \`a la
+Carte approach for building extensible data types and unified interpreters as a
+solution for the expression problem coined by \citeauthor{wadler1998expression}
+\cite{wadler1998expression}.
+%
+In this work, we take ideas from this approach and extend them to work in the
+scope of random data generation, where other parameters come into play apart
+from just combining construction, like generation frequency and terminal
+constructions.
+
+From the practical point of view, \citeauthor{KiriyamaOptimizingDTC} propose an
+optimization mechanism for Data Types \`a Carte, where a concrete data type is
+derived for each different variant of the compositional data types we use in our
+codebase \cite{KiriyamaOptimizingDTC}.
+%
+This solution avoids much of the runtime overhead introduced when internally
+pattern matching against sequencies of |InL| and |InR| data constructors.
+%
+However, we see this solution as not entirely compositional, since we still need
+to rely on Template Haskell to derive the machinery for each specialized data
+type.
+%
+In our particular setting, we found that our solution has a fairly acceptable
+overhead when the representation types are balanced.
+%
+Hence, our aim in the future is to keep our fully extensible approach, but
+automatically computing the best arrangement of constructions in order to reduce
+the overhead as much as possible.
+
+\emph{DRAGEN} is a random testing derivation tool that synthesizes random
+generators for our target data types, tuning their generation frequencies using
+a simulation-based optimization process parametrized by the distribution of
+values desired by the user \cite{DBLP:conf/haskell/MistaRH18}.
+%
+This simulation is based on the theory of \emph{branching processes}, which
+models the growth and extinction of populations across succesive generations.
+%
+In this setting, populations consist of randomly generated data constructors,
+where generators correspond to each level of the generated values.
+%
+This tool has shown to improve the code coverage obtained over our codebase,
+with respect to other fully-automated generators derivation mechanism.
+%
+Recently, \citeauthor{Mista2019GeneratingRS} extended this approach to generate
+random values considering also the other sources of structural information
+covered here, namely abstract interfaces and function pattern machings.
+%
+Using this extension, the authors show how the usage of extra information when
+generating random values is extremely valuable, particularly in situations like
+the ones described in Section \ref{sec:sources}, where the usual derivation
+approaches fail to synthesize useful generators due to a lack of structural
+information.
+%
+Our work is inspired by the same general idea. In constrast, in this paper we
+tackle the representation problem, showing how these ideas can be effectively
+implemented in Haskell using advanced type-level features.

### 2019-05-02 papers/haskell19/3-representation.tex
AddedAle's feedback

@@
 The key idea of this work is to lift each different source of structural
-information into the type level.
+information at the type level.
 %
-In this light, the ``shape'' of our random data is determined entirely by
-the type we use to represent it during the generation process.
+In this light, the ``shape'' of our random data is determined entirely by the
+type we use to represent it during the generation process.
 %
-For this, we will define a set of different simple ``open'' data types,
-each one in charge of representing a single construction of our target data
-type.
+For this, we will define a set of different simple ``open'' types, each one in
+charge of representing a single construction of our target data type.
 %
 These types can be later
+%
 \begin{inparaenum}[(i)]
 \item combined in several ways depending on the desired shape of our test data,
+  applying the familiar technique of Data Types \`a la Carte;
 \item randomly generated (see Section \ref{sec:generators}), and finally,
@@
 our source code in compile-time, relieving programmers of the burden of manually
-implementing this required machinery.
+implementing the required machinery.
 
@@
 \begin{code}
-data Con_Text   r = Con_Text  String
-data Con_Sing   r = Con_Sing  String
-data Con_Tag    r = Con_Tag   String r
-data Con_Join   r = Con_Join  r r
+data Con_Text   r = Mk_Text  String
+data Con_Sing   r = Mk_Sing  String
+data Con_Tag    r = Mk_Tag   String r
+data Con_Join   r = Mk_Join  r r
 \end{code}
@@
 %
-Then, for instance, the value |Con_Tag "div" x :: Con_Tag y| represents the
-|Html| value |Tag "div" x|, for some sub-term |x :: y| that can be transformed
+Then, for instance, the value |Mk_Tag "div" x :: Con_Tag r| represents the
+|Html| value |Tag "div" x|, for some sub-term |x :: r| that can be transformed
 back to |Html| as well.
@@
 %
-In practical terms, it ensures that each representation type |f| will be
-translated to a unique target type |a|.
+% In practical terms, it ensures that each representation type |f| will be
+% translated to a unique target type |a|.
 %
@@
 To do so we can simply transform each representation into its corresponding
-value by picking its analogous data constructor, and piping its fields
+value by picking its corresponding data constructor, and piping its fields
 unchanged:
@@
 instance Algebra Con_Text Html where
-  alg (Con_Text x) = Text x
+  alg (Mk_Text x) = Text x
 
 instance Algebra Con_Sing Html where
-  alg (Con_Sing x) = Sing x
+  alg (Mk_Sing x) = Sing x
 
 instance Algebra Con_Tag Html where
-  alg (Con_Tag t x) = Tag t x
+  alg (Mk_Tag t x) = Tag t x
 
 instance Algebra Con_Join Html where
-  alg (Con_Join x y) = Join x y
+  alg (Mk_Join x y) = Join x y
 \end{code}
 %
-Note how every definition regarding each different representation so far was
+Note how every definition regarding each different constructor so far was
 obtained rather mechanically.
@@
 instance (Functor f, Functor g) => Functor (f oplus'' g)
-  where  fmap h (InL f) = InL (fmap h f)
-         fmap h (InR g) = InR (fmap h g)
+  where  fmap h (InL fa) = InL (fmap h fa)
+         fmap h (InR ga) = InR (fmap h ga)
 
@@
 instance (Algebra f a, Algebra g a) => Algebra (f oplus'' g) a
-  where  alg (InL f) = alg f
-         alg (InR g) = alg g
+  where  alg (InL fa) = alg fa
+         alg (InR ga) = alg ga
 \end{code}
 
-In the previous definition, we simply use the appropriate |Algebra| instance for
-the inner representation, based on which injection was used to construct the
-composite value.
+where we simply use the appropriate |Algebra| instance for the inner
+representation, based on which injection was used to construct the composite
+value.
 %
-It is worth remarking that this instance also lets us handle multiple
-compositions transparently, i.e., for every set of types |f1|, |f2|, ..., |fk|
-representing a target type |t| and having suitable |Algebra| type instances,
-then the composition |f1 oplus'' f2 oplus'' dots oplus'' fk| can be mapped back
-to |t| using this same mechanism.
+It is worth remarking that multiple compositions are handled transparently,
+i.e., for every set of types |f1|, |f2|, ..., |fk| representing a target type
+|t| and having suitable |Algebra| type instances, then the composition |f1
+oplus'' f2 oplus'' dots oplus'' fk| can be mapped back to |t| using this same
+mechanism.
 %
 Additionally, the order in which we associate each operand of |oplus| results
-irrelevant, thus we can omit parentheses when describing composite
+semantically irrelevant, thus we can omit parentheses when describing composite
 representations.
 %
-For simplicity, we will assume |oplus| as having rigth associativity, i.e., |f
-oplus'' g oplus'' h == f oplus'' (g oplus'' h)|.
+However, we found that in practice, the way in which we decide to associate our
+constructions affects the generation speed dramatically.
+%
+This phenomena is addressed in detail in Section \ref{sec:casestudies}.
+%
+For simplicity here, we will assume |oplus| as having rigth associativity, i.e.,
+|f oplus'' g oplus'' h == f oplus'' (g oplus'' h)|.
 
@@
 %
-In order to represent recursive values we need to close them ``tying the knot''
+In order to represent recursive values, we need to close them ``tying the knot''
 recusively, i.e., once we have decided the final representation of our target
@@
 \begin{code}
-newtype Fix (f :: * -> *) = Fix { unFix :: f (Fix f) }
+newtype Fix (f :: * -> *) = FixC { unFix :: f (Fix f) }
 \end{code}
@@
 This way, the compiler will avoid producing a runtime representation for it,
-thus avoiding unnecessary overhead.
+helping to reduce some of the unnecessary overhead.
 %
@@
   x' ::  Fix Html'
-  x' =   Fix  (InR (InR (InR (Con_Join
-              (Fix (InL (Con_Text "hello")))
-              (Fix (InR (InL (Con_Sing "hr"))))))))
+  x' =   FixC  (InR  (InR (InR (Mk_Join
+               (FixC (InL (Mk_Text "hello")))
+               (FixC (InR (InL (Mk_Sing "hr"))))))))
 \end{code}
@@
 The representation mechanism we have developed so far let us determine the shape
-of our target data based on the type we use to represent it.
+of our target data based on the type we use to represent its constructors.
 %
-However, if we only use it to represent data types based solely on their data
-constructors, then it is hardly useful for performing random testing, as it only
-adds unnecessary complexity to the random generation process.
+However, it is hardly useful for performing random testing, as it only adds
+unnecessary complexity to the random generation process.
 %
@@
 %
+\subsubsection{Abstract Interfaces}
+
 Fistly, let us consider the case of generating values obtained by abstract
@@
 \begin{code}
-data Fun_br     r = Fun_br
-data Fun_bold   r = Fun_bold r
-data Fun_list   r = Fun_list [r]
-data Fun_cat    r = Fun_cat  r r
+data Fun_br     r = Mk_br
+data Fun_bold   r = Mk_bold r
+data Fun_list   r = Mk_list [r]
+data Fun_cat    r = Mk_cat  r r
 \end{code}
@@
 instance Algebra Fun_br Html where
-  alg Fun_br = br
+  alg Mk_br = br
 
 instance Algebra Fun_bold Html where
-  alg (Fun_bold x) = bold x
+  alg (Mk_bold x) = bold x
 
 instance Algebra Fun_list Html where
-  alg (Fun_list xs) = list xs
+  alg (Mk_list xs) = list xs
 
 instance Algebra Fun_cat Html where
-  alg (Fun_cat x y) = x <+> y
+  alg (Mk_cat x y) = x <+> y
 \end{code}
 %
-In the previous |Algebra| instances, we simply return the result of evaluating
-the appropriate functions, using each representation field as an input argument.
+There, we simply return the result of evaluating the appropriate functions,
+using each representation field as an input argument.
 
@@
 %
+\subsubsection{Functions' pattern matchings}
+
 The second source of structural information that we consider in this work is the
@@
 \begin{code}
-data Pat_simplify_1 r = Pat_simplify_1 String String
-data Pat_simplify_2 r = Pat_simplify_2 String r r
+data Pat_simplify_1 r = Mk_simplify_1 String String
+data Pat_simplify_2 r = Mk_simplify_2 String r r
 \end{code}
@@
 where each pattern variable gets instantiated using the values stored in its
-representation fields:
+representation field:
 
@@
 instance Algebra Pat_simplify_1 Html where
-  alg (Pat_simplify_1 t1 t2) = Join (Text t1) (Text t2)
+  alg (Mk_simplify_1 t1 t2) = Join (Text t1) (Text t2)
 
 instance Algebra Pat_simplify_2 Html where
-  alg (Pat_simplify_1 t x y) = Join (Join (Text t) x) y
+  alg (Mk_simplify_1 t x y) = Join (Join (Text t) x) y
 \end{code}
@@
 %
+\subsubsection{Ensambling Altogether}
+
 Using this extensions, we are now able to represent values of our target data
@@
 Instead of constructing values solely by combining different data constructors,
-this mechanism lets us also consider complex constructions like computing
-abstract interface functions and functions' patterns, treating them uniformingly
-and in a fine grained fashion.
+this mechanism lets us also consider complex constructions like abstract
+interface functions and functions' patterns, treating them uniformingly and in a
+fine grained fashion.
 
@@
 
-%
 Furthermore, if we want to test properties expecting valid |Html| values as

### 2019-05-02 papers/haskell19/4-generators.tex
AddedAle's feedback

@@
 
-In this section we introduce the machinery required to randomly generate the
+In this section, we introduce the machinery required to randomly generate the
 values encoded using our representations.
 %
-For this purpose, this step also follows the modular, fine-grained fashion,
-resulting in a generation process entirely compositional.
+This step also follows the modular, fine-grained fashion, resulting in a
+generation process entirely compositional.
 %
 In this light, our random generators are built from smaller ones (each one
-representing a single construction), and they are based solely on the types we
-use to represent the shape of our random data.
+representing a single random construction), and they are based solely on the
+types we use to represent the shape of our random data.
 
@@
   we have reached the maximum depth (case |gen 0|).
-\item We generate different constructions in a (possibly) different frequency
-  that is fixed over time.
+\item We generate different constructions in a (possibly) different frequency.
 \end{itemize}
 %
-Then, in order to achieve the desired generation behavior, we will need to model
-these previous aspects in our modular framework, and in such a way that we do
-not compromise the compositionality obtained so far.
+For the rest of this section, we will focus on modelling these previous aspects
+in our modular framework, and in such a way that we do not compromise the
+compositionality obtained so far.
 
@@
 can observe that the depth parameter |d| is threaded to the different recursive
-call of our generator, but always within the scope of the |gen| local function.
+call of our generator, but always within the scope of the local function |gen|.
 %
@@
 
-A |BGen| is essentially a normal |Gen| with the maximum recursive depth as an
-input parameter.
+A |BGen| is essentially a normal QuickCheck |Gen| with the maximum recursive
+depth as an input parameter.
 %
@@
 class BArbitrary (a :: *) where
-  bounded :: BGen a
+  barbitrary :: BGen a
 \end{code}
@@
 \begin{code}
-instance BArbitrary a => Arbitrary a where
-  arbitrary = sized bounded
+instance BArbitrary MyType => Arbitrary MyType where
+  arbitrary = sized barbitrary
 \end{code} %$
@@
 
-This type class allows us to write modular generators, leaving unspecified the
-specific generator used when generating values of its type argument.
-%
-For instance, the default parametric generator for the data type |Maybe| looks
-like:
+% This type class allows us to write modular generators, leaving unspecified the
+% specific generator used when generating values of its type argument.
+% %
+% For instance, the default parametric generator for the data type |Maybe| looks
+% like:
 
-\begin{code}
-instance Arbitrary1 Maybe where
-  liftArbitrary gen = frequency
-    [  (1,  pure Nothing),  (3,  Just <$> gen) ]
-\end{code} %$
+% \begin{code}
+% instance Arbitrary1 Maybe where
+%   liftArbitrary gen = frequency
+%     [  (1,  pure Nothing),  (3,  Just <$> gen) ]
+% \end{code} %$
 
-Where the random generator use to generate the value stored in a |Just|
-constructor is parametrized by the |gen| input variable.
+% Where the random generator use to generate the value stored in a |Just|
+% constructor is parametrized by the |gen| input variable.
 %
@@
 instance BArbitrary1 Con_Text where
-  liftBGen gen d = Con_Text <$> arbitrary
+  liftBGen bgen d = Mk_Text <$> arbitrary
 
 instance BArbitrary1 Con_Sing where
-  liftBGen gen d = Con_Sing <$> arbitrary
+  liftBGen bgen d = Mk_Sing <$> arbitrary
 
 instance BArbitrary1 Con_Tag where
-  liftBGen gen d = Con_Tag <$> arbitrary <*> gen (d - 1)
+  liftBGen bgen d = Mk_Tag <$> arbitrary <*> bgen (d - 1)
 
 instance BArbitrary1 Con_Join where
-  liftBGen gen d = Con_Join <$> gen (d - 1) <*> gen (d - 1)
+  liftBGen bgen d = Mk_Join <$> bgen (d - 1) <*> bgen (d - 1)
 \end{code} %$
 
-Note in the previous definitions how each instance is defined to be independent
-of the maximum depth (paramtrized by the input variable |d|) and the random
-generator used for the recursive sub-terms (parametrized by the input variable
-|gen|).
+Note in the previous definitions how each instance is defined to be parametric
+of the maximum depth (using the input integer |d|) and the random generator used
+for the recursive sub-terms (using input bounded generators |bgen|).
 %
@@
 instance BArbitrary1 Fun_bold where
-  liftBGen gen d = Fun_bold <$> gen (d - 1)
+  liftBGen bgen d = Mk_bold <$> bgen (d - 1)
 
 instance BArbitrary1 Pat_simplify_2 where
-  liftBGen gen d = Pat_simplify_2  <$> arbitrary
-                                   <*> gen (d - 1)
-                                   <*> gen (d - 1)
+  liftBGen bgen d = Mk_simplify_2
+    <$> arbitrary <*> bgen (d - 1) <*> bgen (d - 1)
 \end{code} %$
@@
 instance BArbitrary1 f => BArbitrary (Fix f) where
-  bounded d = Fix <$> liftBGen bounded d
+  barbitrary d = FixC <$> liftBGen barbitrary d
 \end{code} %$
 
-The previous instance employs the |BArbitrary1| of our final representation to
-generate sub-terms recursively using itself (|bounded d|) as the parametrized
-input generator, wrapping each recursive sub-term with a |Fix| data constructor.
+Here we used the |BArbitrary1| instance of our final representation |f| to
+generate sub-terms recursively by lifting itself (|liftBGen barbitrary|) as the
+parametrized input generator, wrapping each recursive sub-term with a |FixC|
+data constructor.
 
@@
 \begin{code}
-newtype Term (f :: * -> *) a = Term (f a)
+newtype Term (f :: * -> *) a = TermC (f a)
 \end{code}
@@
 instance Algebra f a => Algebra (Term f) a where
-  alg (Term f) = alg f
+  alg (TermC f) = alg f
 
 instance BArbitrary1 f => BArbitrary1 (Term f) where
-  liftBGen gen d = Term <$> liftBGen gen d
+  liftBGen bgen d = TermC <$> liftBGen bgen d
 \end{code} %$
@@
 \begin{code}
-newtype ((f :: * -> *) otimes'' (n :: Nat)) a = Freq (f a)
+newtype ((f :: * -> *) otimes'' (n :: Nat)) a = FreqC (f a)
 \end{code}
@@
 instance Algebra f a => Algebra (f otimes'' n) a where
-  alg (Freq f) = alg f
+  alg (FreqC f) = alg f
 
 instance BArbitrary1 f => BArbitrary1 (f otimes'' n) where
-  liftBGen gen d = Freq <$> liftBGen gen d
+  liftBGen bgen d = FreqC <$> liftBGen bgen d
 \end{code} %$
@@
 %
-Consider for instance the following type defined using parenthesis for clarity:
+Consider for instance the following type---defined using parenthesis for clarity:
 
 \begin{code}
-(f otimes'' 2) oplus'' ((g otimes'' 3) oplus'' (Term h))
+(f otimes'' 2) oplus'' ((g otimes'' 3) oplus'' (Term h otimes'' 5))
 \end{code}
@@
 
-On the other hand, the last step of our generation process which only generates
-terminal constructions could be seen as generating values considering all the
-available constructions, but instead treating the non-terminal ones as having
-generation frequency zero.
+On the other hand, the last step of our generation process, which only generates
+terminal constructions, could be seen as treating the non-terminal ones as
+having generation frequency zero.
 %
-This way we can reuse our previous type level computation to calculate the
-\emph{terminal generation frequency} |FreqOf'| of a given representation type as
+This way we can introduce antoher type-level computation to calculate the
+\emph{terminal generation frequency} |FreqOf'| of a given representation as
 follows:
@@
   => BArbitrary1 (f oplus g) where
-  liftBGen gen d =
+  liftBGen bgen d =
     if d > 0
     then frequency
-      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen d)
-      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen d) ]
+      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen bgen d)
+      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen bgen d) ]
     else frequency
-      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen d)
-      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen d) ]
+      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen bgen d)
+      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen bgen d) ]
 \end{code}
@@
 We need to use a \emph{visible type application} (employing the |at (...)|
-syntax) to desambiguate which type-level natural value we are refering to.
+syntax) to desambiguate to which natural value we are actually refering to.
 %
@@
 %
-This kind of exceptions will arise for examplke if we forget to tag at least a
+This kind of exceptions will arise, for example, if we forget to tag at least a
 single construction within a composite representation as terminal.
@@
 \end{code}
-%
+
 The previous type families compute the terminal generation frequency of a
@@
 empty constraint |()| that is always trivially satisfied in any other case.
+
 %
 Finally, we can use this constraint to define a safe generation primitive
-|genEval| to obtain a random depth-bounded generator for every target type |a|
+|genEval| to obtain a concrete depth-bounded generator for every target type |a|
 specified using a safe generation representation |f|:
@@
 \begin{code}
-genEval  :: forall f a dot
-  (BArbitrary1 f, HasTerminals f, Algebra f a) => BGen a
+genEval  :: forall f a dot  ( BArbitrary1 f
+                            , HasTerminals f
+                            , Algebra f a) => BGen a
 genEval d = eval <$> bounded at (Fix f) d
@@
 
-Where |Html_simplify| and |ValidHtml| are the types defined at the end of
-Section \ref{sec:representation}.
+Where |Html_simplify| and |ValidHtml| are the representations defined at the end
+of Section \ref{sec:representation}.
 

### 2019-05-02 papers/haskell19/5-typelevel.tex
AddedAle's feedback

@@
 %
-This section introduces a set refinements to our basic language for describing
-random generators, making them more flexible and robust in order to fit
-real-world usage scenarios.
+This section introduces refinements to our basic language for describing random
+generators, making them more flexible and robust in order to fit real-world
+usage scenarios.
 
@@
 %
-To begin with, let us consider the name conventions we used when defining our
-representation data types along this paper.
+% To begin with, let us consider the name conventions we used when defining our
+% representation data types along this paper.
 %
-Since our aim is to automate this process as much as possible using Haskell's
-meta-programming capabilities, the actual names used when deriving construction
-representations will have to be generated in such a way that they are compliant
-with Haskell syntax and are unique within our namespace.
+% Since our aim is to automate this process as much as possible using Haskell's
+% meta-programming capabilities,
 %
-This means that, in practice, type names like |Fun_cat| or |Pat_simplify_1| are
-technically not valid Haskell data type names, thus they will have to be
-synthesized as something like |Fun_lt_plus_gt_543| and |Pat_simplify_1_325|,
-where the last sequence of numbers is automatically inserted by Template Haskell
-to ensure uniqueness.
+The first problem we face is that of naming conventions.
+%
+In practice, the actual names used when deriving construction representations
+need to be generated in a way that they are compliant with Haskell syntax and
+are \emph{unique within our namespace}.
+%
+This means that, type names like |Fun_cat| or |Pat_simplify_1| are technically
+not valid Haskell data type names, thus they will have to be synthesized as
+something like |Fun_lt_plus_gt_543| and |Pat_simplify_1_325|, where the last
+sequence of numbers is automatically inserted by Template Haskell to ensure
+uniqueness.
 
 This naming convention results hard to use, specially if we consider that we do
-not now the actual name types until they are synthesized in compile-time, due to
-their unique suffixes.
+not know the actual type names until they are synthesized in compile-time, due
+to their unique suffixes.
 %
-Fortunately, it is easy to solve this problem again using some type-level
-machinery.
+Fortunately, it is easy to solve this problem using some type-level machinery.
 %
-Instead of imposing a naming convention in our representation synthesizer, we
-can define a set of open type families to hide each kind of construction using
-meaningful names:
+Instead of imposing a naming convention in our tool, we can define a set of open
+type families to hide each kind of construction using meaningful names:
 
@@
 
-Where |Symbol| is the kind of type-level strings in Haskell.
+where |Symbol| is the kind of type-level strings in Haskell.
 %
@@
 
-And the end user can simply refer to each particular construction by using these
-synonyms, e.g., with representation types like |Con "Text" oplus' Fun "<+>"|.
+As a result, the end user can simply refer to each particular construction by
+using these synonyms, e.g., with representation types like |Con "Text" oplus'
+Fun "<+>"|.
 
@@
 
-This way, any quoted construction will considered to be terminal.
+This way, any construction with the supra-index |T| will considered to be
+terminal.
 %
@@
 
-And simply obtain a concrete generator by refering to this previous type when
-using |genEval|:
+And simply obtain a concrete generator by refering to |Html_Spec| when using
+|genEval|:
 
@@
 
-So far we have seen how specify random generators for our simple self-contained
-|Html| data type.
+So far we have seen how to specify random generators for our simple
+self-contained |Html| data type.
 %
@@
 %
-Consider for example the following |Tree| data type definition encoding binary
+Consider, for example, the following |Tree| data type definition encoding binary
 trees with generic information of type |a| in the leaves:
@@
 \begin{code}
-data Con_Leaf a r = Con_Leaf a
-data Con_Node a r = Con_Node r r
+data Con_Leaf a r = Mk_Leaf a
+data Con_Node a r = Mk_Node r r
 \end{code}
@@
 \begin{code}
-instance Algebra (Con_Leaf a) (Tree a) where ...
-instance Algebra (Con_Node a) (Tree a) where ...
+instance Algebra (Con_Leaf a) (Tree a) where dots
+instance Algebra (Con_Node a) (Tree a) where dots
 
-instance Arbitrary a => BArbitrary1 (Con_Leaf a) where ...
-instance Arbitrary a => BArbitrary1 (Con_Node a) where ...
+instance Arbitrary a => BArbitrary1 (Con_Leaf a) where dots
+instance Arbitrary a => BArbitrary1 (Con_Node a) where dots
 \end{code}
 
-The biggest different appears when deriving the required type instances for the
-|Con|, |Fun| and |Pat| synonyms.
+% The biggest different appears when deriving the required type instances for the
+% |Con|, |Fun| and |Pat| synonyms.
 %
@@
 \begin{code}
-data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot Some (f a r)
+data Some (f :: * -> * -> *) (r :: *) = forall (a :: *) dot SomeC (f a r)
 \end{code}
 
-In the previous definition, |Some| is a wrapper for a 2-parametric type that
-hides the first type variable by using an explicit existential quantificator.
+The type constructor |Some| is a wrapper for a 2-parametric type that hides the
+first type variable using an explicit existential quantificator.
 %
@@
 
-This way we can write generation specification for our |Tree| data type without
+We can therefore write generation specification for our |Tree| data type without
 having to refer to its type parameter anywhere.
@@
 \end{code}
-
-As a result, in our definition of |genIntTree| above, the |apply| operator will
-reduce the type |(Tree_Spec apply'' Int)| to the following concrete type:
+%
+As a result, in |genIntTree|, the |apply| operator will reduce the type
+|(Tree_Spec apply'' Int)| to the following concrete type:
 
@@
 The next section evaluates our approach in terms of performance using a set of
-case studies extracted from real-world Haskell libraries.
+case studies extracted from real-world Haskell implementations.

### 2019-05-02 papers/haskell19/6-casestudies.tex
AddedAle's feedback

@@
 %
-However, this expresiveness comes attached to a perceptible runtime overhead.
+However, this expresiveness comes attached to a perceptible runtime overhead,
+primarly inherited from the using the \`a la Carte technique.
 %
@@
 
-To evaluate our approach, we analyzed the performance of generating random
-values using three real-world case studies:
+In this work, we analyzed the performance of generating random values using
+three case studies:
+%
 \begin{inparaenum}[(i)]
@@
 \item Lisp S-expressions, inspired by the package \emph{hs-zuramaru}, and
-\item HTML expressions, inspired by the \emph{html} package, and following the
+\item HTML expressions, inspired by the \emph{html} package, which follows the
   same structure as our motivating |Html| example.
@@
     \hline
-    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
+    Case Study & \#|Con| & \#|Pat| & \#|Fun| & Total Constructions \\ \hline
     \hline

### 2019-05-02 papers/haskell19/7-related.tex
AddedAle's feedback

@@
 In this setting, populations consist of randomly generated data constructors,
-where generators correspond to each level of the generated values.
+where generations correspond to each level of the generated values.
 %
@@
 %
-Our work is inspired by the same general idea. In constrast, in this paper we
-tackle the representation problem, showing how these ideas can be effectively
+Our work is inspired by the same general idea. In constrast, this paper tackles
+the representation problem, showing how these ideas can be effectively
 implemented in Haskell using advanced type-level features.

### 2019-05-02 papers/haskell19/6-casestudies.tex
Added automatic representaition balancing

@@
 
-We benchmarked the execution time required to generate and fully evaluate 1000
+We benchmarked the execution time required to generate and fully evaluate 10000
 random values corresponding to each case study, comparing both manually-written
@@
       , ymajorgrids=true
-      , ytick={0,1,...,5}
-      , ymax=4.5
+      , ytick={0,1,...,4}
+      , ymax=4
+      , ymin=0.5
       , enlargelimits=0.25
@@
       \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (HTML,1)};
-      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (HTML,4.88)};
-      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (HTML,1.70)};
-      \legend{Manual, Modular/Linear, Modular/Balanced}
+      \addplot[fill, green!30] coordinates {(SExp,2.40) (RBT,3.57) (HTML,4.27)};
+      \addplot[fill, blue!30]  coordinates {(SExp,1.60) (RBT,2.52) (HTML,1.55)};
+      \legend{Manual, Modular (lin), Modular (bal)}
     \end{axis}
@@
 
-As it can be observed, our approach suffers from a noticeable runtime overhead,
-specially when considering the HTML case study with a large amount of
-construction involved at the generation process.
+As it can be observed, our approach suffers from a noticeable runtime overhead
+when using linearly defined representations, specially when considering the HTML
+case study with a large amount of construction involved at the generation
+process.
 %
-However, by balancing our representation types we found that the generation
-performance gets dramatically improved.
+However, we found that, by balancing our representation types the generation
+performance improves dramatically.
 %
-To apply this optimization automatically, and relying on more sophisticated
-techniques like prefix-free encodings to obtain optimal representations, is a
-undergoing path of our future work.
+At the light of these improvements in the performance, we included in our tool
+an additional type-level computation that automatically balances our
+representations in order to reduce the generation overhead.
+
+% and relying on more sophisticated techniques like prefix-free encodings to
+% obtain optimal representations, is a undergoing path of our future work.
 
@@
 
-Nonetheless, we consider that these results are in turn very encouraging, given
-that the flexibility obtained from using our approach does not produce severe
-slowdowns in practice, nor explodes more than linearly as we increase the
-complexity of our random generators.
+All in all, we consider that these results are fairly encouraging, given that
+the flexibility obtained from using our compositional approach does not produce
+severe slowdowns in whem generating random values in practice.
+
+% , nor explodes more than linearly as we increase the complexity of our random
+% generators.

### 2019-05-02 bench/Html/HtmlUnbal.hs

@@ -28,8 +28,8 @@
 --
 -----------------------------------------------------------------------------
 
-module Html.Html (
-      module Html.Html,
+module Html.HtmlUnbal (
+      module Html.HtmlUnbal,
       ) where
 
 
@@ -1456,11 +1456,12 @@ type Html_S =
        :+ Fun' "selected"
    ]
 
-deriveBoundedArbitrary
-  [ [] .=> ''Html
-  , [] .=> ''HtmlElement
-  , [] .=> ''HtmlAttr
-  ] ''Html_S
+instance BoundedArbitrary HtmlElement where
+  boundedArbitrary = genEval @(Lookup Html_S "HtmlElement")
+instance BoundedArbitrary Html where
+  boundedArbitrary = genEval @(Lookup Html_S "Html")
+instance BoundedArbitrary HtmlAttr where
+  boundedArbitrary = genEval @(Lookup Html_S "HtmlAttr")
 
 genHtml'' :: BoundedGen Html
 genHtml'' = boundedArbitrary

### 2019-05-02 bench/Lisp/Lisp.hs

@@ -125,25 +125,8 @@ type SExprS
    :+ Pat' "constantFold" 3
    :+ Pat' "constantFold" 4
 
-type SExprSBal
-   =  (((Con' "Null"
-   :+ Con' "AtomInt")
-   :+ (Con' "AtomBool"
-   :+ Con' "AtomSymbol"))
-   :+ ((Con  "Quote"
-   :+ Con  "Cons")
-   :+ (Pat  "execute" 1
-   :+ Pat  "execute" 2)))
-   :+ (((Pat  "execute" 3
-   :+ Pat  "execute" 4)
-   :+ (Pat  "execute" 5
-   :+ Pat' "constantFold" 1))
-   :+ ((Pat' "constantFold" 2
-   :+ Pat' "constantFold" 3)
-   :+ Pat' "constantFold" 4))
-
 genSExpr' :: BoundedGen SExpr
 genSExpr' = genEval @SExprS
 
 genSExpr'' :: BoundedGen SExpr
-genSExpr'' = genEval @SExprSBal
+genSExpr'' = genRep @SExprS

### 2019-05-02 bench/Main.hs

@@ -7,13 +7,14 @@ import Criterion.Types
 import Test.Dragen2
 import Test.QuickCheck
 
-import qualified RBT.RBT   as RBT
-import qualified RE.RE     as RE
-import qualified Html.Html as Html
-import qualified Lisp.Lisp as Lisp
+import qualified RBT.RBT        as RBT
+import qualified RE.RE          as RE
+-- import qualified Html.Html     as Html
+import qualified Html.HtmlUnbal as Html
+import qualified Lisp.Lisp      as Lisp
 
 samples :: Int
-samples = 1000
+samples = 10000
 
 size :: Int
 size = 30
@@ -31,15 +32,16 @@ genSamples gen
 
 main :: IO ()
 main = defaultMainWith (defaultConfig { verbosity = Verbose })
-  [ 
-  -- bench "RBT/Manual"      $ nfIO (genSamples (RBT.genTree   @Int))
-  -- , bench "RBT/Dragen"      $ nfIO (genSamples (RBT.genTree'  @Int))
-   bench "RBT/Dragen/Bal"  $ nfIO (genSamples (RBT.genTree'' @Int))
-  -- , bench "Html/Manual"     $ nfIO (genSamples (Html.genHtml ))
-  -- , bench "Html/Dragen/Bal" $ nfIO (genSamples (Html.genHtml'))
-  -- , bench "Lisp/Manual"     $ nfIO (genSamples (Lisp.genSExpr))
-  -- , bench "Lisp/Dragen"     $ nfIO (genSamples (Lisp.genSExpr'))
-  -- , bench "Lisp/Dragen/Bal" $ nfIO (genSamples (Lisp.genSExpr''))
+  [
+  --   bench "RBT/Manual"          $ nfIO (genSamples (RBT.genTree   @Int))
+  -- , bench "RBT/Dragen/Linear"   $ nfIO (genSamples (RBT.genTree'  @Int))
+  -- , bench "RBT/Dragen/Autobal"  $ nfIO (genSamples (RBT.genTree'' @Int))
+  -- , bench "Lisp/Manual"         $ nfIO (genSamples (Lisp.genSExpr))
+  -- , bench "Lisp/Dragen/Linear"  $ nfIO (genSamples (Lisp.genSExpr'))
+  -- , bench "Lisp/Dragen/Autobal" $ nfIO (genSamples (Lisp.genSExpr''))
[…]

### 2019-05-02 bench/RBT/RBT.hs

@@ -244,24 +244,9 @@ type RBT_Spec
   :+ Pat  "balR" 2
   :+ Pat  "balR" 3
 
-type RBT_Spec_Bal
-  =  (((Con' "E"
-  :+ Con  "T")
-  :+ (Fun' "empty"
-  :+ Fun  "insert"))
-  :+ ((Fun  "balance'"
-  :+ Fun  "delete")
-  :+ Fun  "fuse"))
-  :+ (((Pat  "balL" 1
-  :+ Pat  "balL" 2)
-  :+ Pat  "balL" 3)
-  :+ ((Pat  "balR" 1
-  :+ Pat  "balR" 2)
-  :+ Pat  "balR" 3))
-
 
 genTree' :: forall a. (Ord a, Arbitrary a) => BoundedGen (Tree a)
 genTree' = genEval @(RBT_Spec <| a)
 
 genTree'' :: forall a. (Ord a, Arbitrary a) => BoundedGen (Tree a)
-genTree'' = genEval @(RBT_Spec_Bal <| a)
+genTree'' = genRep @(RBT_Spec <| a)

### 2019-05-02 bench/RE/RE.hs

@@ -209,7 +209,7 @@ type R_Spec_Bal
   :+ Fun  "epsR"))
 
 genR' :: forall a. (Ord a, Arbitrary a) => BoundedGen (R a)
-genR' = genEval @(R_Spec <| a)
+genR' = genRep @(R_Spec <| a)
 
 genR'' :: forall a. (Ord a, Arbitrary a) => BoundedGen (R a)
-genR'' = genEval @(R_Spec_Bal <| a)
+genR'' = genRep @(R_Spec_Bal <| a)

### 2019-05-02 src/Test/Dragen2/Rep.hs

@@ -167,6 +167,42 @@ instance
     typeEta
       = typeEta  @f <> typeEta  @g
 
+----------------------------------------
+-- | Balancing representations
+
+type family
+  Multiply (n :: Nat) (fs :: [(Type -> Type,Nat)]) :: [(Type -> Type,Nat)] where
+  Multiply _ '[] = '[]
+  Multiply x ('(f, n) ': fs) = '(f, x * n) ': Multiply x fs
+
+type family
+  SetTerm (fs :: [(Type -> Type, Nat)]) :: [(Type -> Type, Nat)] where
+  SetTerm '[] = '[]
+  SetTerm ('(f, n) ': fs) = '(Term f, n) : SetTerm fs
+
+type family
+  ToFreqList (rep :: Type -> Type) :: [(Type -> Type, Nat)] where
+  ToFreqList (Sum f g) = ToFreqList f ++ ToFreqList g
+  ToFreqList (Freq f n) = Multiply n (ToFreqList f)
+  ToFreqList (Term f) = SetTerm (ToFreqList f)
+  ToFreqList f = '[ '(f, 1)]
+
+type family
+  BuildTree (fs :: [(Type -> Type, Nat)]) :: Type -> Type where
+  BuildTree '[ '(Term f, n)] = Term (BuildTree '[ '(f, n)])
+  BuildTree '[ '(f, 1)] = f
+  BuildTree '[ '(f, n)] = Freq f n
+  BuildTree xs = Sum (BuildTree (FirstHalf xs)) (BuildTree (SecondHalf xs))
+
+type family
+  Balance (rep :: Type -> Type) :: Type -> Type where
+  Balance f = BuildTree (ToFreqList f)
+
+genRep :: forall f a.
+  ( BoundedArbitrary1 (Balance f)
+  , Algebra (Balance f) a
+  ) => BoundedGen a
+genRep depth = genEval @(Balance f) depth
 
 ----------------------------------------
 -- | Existential types to hide type parameters

### 2019-05-02 src/Test/Dragen2/Spec.hs

@@ -75,7 +75,7 @@ type family
 type family
   Flatten (spec :: Spec) :: Spec where
   Flatten '[] = '[]
-  Flatten ('(k, t) : ts) = '(k, List2Rep (Rep2List t)) : Flatten ts
+  Flatten ('(k, t) : ts) = '(k, Balance (List2Rep (Rep2List t))) : Flatten ts
 
 type family
   Rep2List (rep :: Type -> Type) :: [Type -> Type] where

### 2019-05-02 src/Test/Dragen2/TH.hs

@@ -131,7 +131,7 @@ deriveBoundedArbitrary typeFam spec
                          `DAppT` thNameTyLit typeName
           arbLetDec = DLetDec (DFunD 'boundedArbitrary [arbClause])
           arbClause = DClause [] arbBody
-          arbBody = DAppTypeE (DVarE 'genEval) arbTyApp
+          arbBody = DAppTypeE (DVarE 'genRep) arbTyApp
 
           mkArbCxt v = DAppPr (DConPr ''Arbitrary) (dTyVarBndrToDType v)
 

### 2019-05-02 src/Test/Dragen2/TypeLevel.hs

@@ -26,7 +26,8 @@ type family
   Ix (map :: [(k, v)]) (key :: k) :: Nat where
   Ix ('(k,  _) : m) k = 0
   Ix ('(k', _) : m) k = 1 + Ix m k
-  Ix '[]            k = TypeError (Text "Ix: key not found" :<>: ShowType k)
+  Ix '[]            k
+    = TypeError (Text "Ix: key not found" :<>: ShowType k)
 
 type family
   Keys (map :: [(k, v)]) :: [k] where
@@ -47,6 +48,27 @@ type family
   (==) (f :: k) (g :: k) :: Bool where
   (==) t t = True
   (==) _ _ = False
+
+type family
+  FirstHalf (xs :: [k]) :: [k] where
+  FirstHalf xs = Take (Length xs `Div` 2) xs
+
+type family
+  SecondHalf (xs :: [k]) :: [k] where
+  SecondHalf xs = Drop (Length xs `Div` 2) xs
+
+
+type family
+  Take (n :: Nat) (xs :: [k]) where
+  Take 0  _  = '[]
+  Take _ '[] = '[]
+  Take n  (x ': xs) = x ': Take (n-1) xs
+
+type family
+  Drop (n :: Nat) (xs :: [k]) where
+  Drop 0  xs  = xs
+  Drop _ '[] = '[]
+  Drop n  (x ': xs) = Drop (n-1) xs
 
 type family
   FailWhen (n :: Bool) (f :: Type -> Type) (msg :: Symbol) :: Constraint where

### 2019-05-03 papers/euroSP/main.lhs.tex
Work on case studies/related work

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[sigplan]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

% %% For double-blind review submission, w/ CCS and ACM Reference
% \documentclass[sigplan,review,anonymous]{acmart}
% \settopmatter{printfolios=true}

% %% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
% \documentclass[sigplan,review]{acmart}
% \settopmatter{printfolios=true,printccs=false,printacmref=false}

% %% For single-blind review submission, w/ CCS and ACM Reference
% \documentclass[sigplan,review]{acmart}
% \settopmatter{printfolios=true}

% %% For final camera-ready submission, w/ required CCS and ACM Reference
% \documentclass[sigplan]{acmart}
% \settopmatter{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Lhs2Tex

%include polycode.fmt
%include notation.fmt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[Euro S\&P]{4th IEEE European Symposium on Security and Privacy}{June 17--19, 2019}{Stockhom, Sweden}
%\acmYear{2019}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some recommended packages.
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{paralist}
\usepackage{tikz}
\usepackage{pgfplots}

\usetikzlibrary{bending,patterns,matrix,fadings,positioning,shapes}
\pgfplotsset{compat=1.12}

\newcommand\quickcheck{\emph{QuickCheck}\xspace}
\newcommand\megadeth{\emph{MegaDeTH}\xspace}
\newcommand{\dragen}{DRAGEN\xspace}
\newcommand\feat{\emph{Feat}\xspace}
\newcommand\derive{\emph{derive}\xspace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document
\begin{document}

%% Title information
\title{Automated Penetration Testing with QuickCheck}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

\author{Agust\'in Mista}
\affiliation{
  \institution{Chalmers University of Technology}
}
\email{mista@@chalmers.se}

\author{Alejandro Russo}
\affiliation{
  \institution{Chalmers University of Technology}
}
\email{russo@@chalmers.se}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command

\begin{abstract}

Random property-based testing is an increasingly popular approach to finding
bugs \cite{HughesNSA16,ArtsHNS15,HughesPAN16}.
%
In the Haskell community, \quickcheck \cite{ClaessenH00} is the dominant tool of
this sort.
%
\quickcheck requires developers to specify \emph{testing properties} describing
the expected software behavior.
%
Then, it generates a large number of random \emph{test cases} and reports those
violating the testing properties.
%
\quickcheck generates random data by employing \emph{random test data
generators} or \quickcheck generators for short.
%
The generation of test cases is guided by the \emph{types} involved in the
testing properties.
%
It defines default generators for many built-in types like booleans, integers,
and lists.
%
However, when it comes to user-defined ADTs, developers are usually required to
specify the generation process.
%
The difficulty is, however, that it might become intricate to define generators
so that they result in a suitable distribution or enforce data invariants.

Rather than manually writing generators, libraries \derive \cite{mitchell2007}
and \megadeth \cite{GriecoCB16, grieco2017} allow us to automatically synthesize generators
for a given user-defined ADT.
%
The library \derive provides no guarantees that the generation process
terminates, while \megadeth pays almost no attention to the distribution of
values.
%
In contrast, \emph{Feat} \cite{DuregardJW12} provides a mechanism to uniformly
sample values from a given ADT.
%
It enumerates all the possible values of a given ADT so that sampling uniformly
from ADTs becomes sampling uniformly from the set of natural numbers.
%---we
%
% Lastly, \emph{Luck} is a domain specific language for manually writing
% \quickcheck properties in tandem with generators so that it becomes possible to
% finely control the distribution of generated values \cite{LampropoulosGHH17}.

In this work, we consider the scenario where developers are not fully aware of
the properties and invariants that input data must fulfill.
%
This constitutes a valid assumption for \emph{penetration testing}
\cite{pentest}, where testers often apply fuzzers in an attempt to make programs
crash---an anomaly which might lead to a vulnerability.
%

Our realization is that \emph{branching processes} \cite{gw1875}, a relatively
simple stochastic model conceived to study the evolution of populations, can be
applied to predict the generation distribution of ADTs' constructors in a simple
and automatable manner.
%
To the best of our knowledge, this stochastic model has not yet been applied to
this field, and we believe it may be a promising foundation to develop future
extensions.

% , we adapt results from an area of mathematics known as
% \emph{branching processes}, and show how they help to analytically predict (at
% compile-time) the expected number of generated constructors, even in the
% presence of mutually recursive or composite ADTs.
%
Using our probabilistic formulas, we design heuristics capable of automatically
adjusting probabilities in order to synthesize generators which distributions
are aligned with users' demands \cite{DBLP:conf/haskell/MistaRH18}.

Furthermore, we provide a Haskell implementation of our mechanism in a tool 
called \dragen and perform case studies with real-world applications.
%
When generating random values, our synthesized \quickcheck generators show
improvements in code coverage when compared with those automatically derived by
state-of-the-art tools.

Recently, we extended our framework to consider addiotional sources of structural information from the users' codebase \cite{Mista2019GeneratingRS}. 
% 
In this light, our random generators can produce complex patterns of values, as well as calls to functions in abstract interfaces, obtaining remarkable improvements in terms of code coverage.
%
All of this while still being able to reason about the distributions of generated values in terms of branching processes.

% Overall, our work addresses a timely problem with a neat mathematical insight
% that is backed by a complete implementation and experience on third-party
% examples.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Keywords
%% comma separated list
%% \keywords are mandatory in final camera-ready submission
%\keywords{keyword1, keyword2, keyword3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections

% Nothing here!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% Acknowledgments
% \begin{acks}                            %% acks environment is optional
%                                         %% contents suppressed with 'anonymous'
%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%   %% acknowledge financial support and will be used by metadata
%   %% extraction tools.
%   This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
\vspace{-5pt}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% That's all folks!
\end{document}

### 2019-05-03 papers/haskell19/4-generators.tex
Work on case studies/related work

@@
 \begin{code}
-genEval  :: forall f a dot  ( BArbitrary1 f
+genRep  :: forall f a dot  ( BArbitrary1 f
                            , HasTerminals f
                            , Algebra f a) => BGen a
-genEval d = eval <$> bounded at (Fix f) d
+genRep d = eval <$> bounded at (Fix f) d
 \end{code} %$
@@
 genHtmlPats :: BGen Html
-genHtmlPats = genEval at Html_simplify
+genHtmlPats = genRep at Html_simplify
 
 genValidHtml :: BGen Html
-genValidHtml = genEval  at ValidHtml
+genValidHtml = genRep  at ValidHtml
 \end{code}

### 2019-05-03 papers/haskell19/5-typelevel.tex
Work on case studies/related work

@@
 And simply obtain a concrete generator by refering to |Html_Spec| when using
-|genEval|:
+|genRep|:
 
@@
 genHtml :: BGen Html
-genHtml = genEval at Html_Spec
+genHtml = genRep at Html_Spec
 \end{code}
@@
 genIntTree :: BGen (Tree Int)
-genIntTree = genEval at (Tree_Spec apply' Int)
+genIntTree = genRep at (Tree_Spec apply' Int)
 \end{code}

### 2019-05-03 papers/haskell19/6-casestudies.tex
Work on case studies/related work

@@
 However, this expresiveness comes attached to a perceptible runtime overhead,
-primarly inherited from the using the \`a la Carte technique.
+primarly inherited from the using the Data Types \`a la Carte---a technique
+which is not often scrutinized for performance.
 %
-In this section we evaluate the implicit cost of composing generators using
+In this section, we evaluate the implicit cost of composing generators using
 three real-world case studies.
@@
 %
-Because of this, in general we expect a performance impact with respect to
+Because of this, in general, we expect a performance impact with respect to
 manually-written concrete generators.
@@
 %
-Hence, one can drastically reduce the runtime overhead by associating each
-|oplus| operator in a balanced fashion.
+In this light, one can drastically reduce the runtime overhead by associating
+each |oplus| operator in a balanced fashion.
 %
@@
 
-In this work, we analyzed the performance of generating random values using
-three case studies:
+Then, we analyzed the performance of generating random values using three case
+studies:
 %
 \begin{inparaenum}[(i)]
-\item Red-Black Trees, inspired by Okasaki's formulation,
-\item Lisp S-expressions, inspired by the package \emph{hs-zuramaru}, and
-\item HTML expressions, inspired by the \emph{html} package, which follows the
-  same structure as our motivating |Html| example.
+\item Red-Black Trees (RBT), inspired by Okasaki's formulation
+  \cite{okasaki1999red},
+\item Lisp S-expressions (SExp), inspired by the package
+  \emph{hs-zuramaru}\footnote{\url{http://hackage.haskell.org/package/zuramaru}},
+  and
+\item HTML expressions (HTML), inspired by the \emph{html}
+  package\footnote{\url{http://hackage.haskell.org/package/html}}, which follows
+  the same structure as our motivating |Html| example.
 \end{inparaenum}
@@
     \hline
-    Case Study & \#|Con| & \#|Pat| & \#|Fun| & Total Constructions \\ \hline
+    Case Study & \#|Con| & \#|Fun| & \#|Pat| & Total Constructions \\ \hline
     \hline
-    RBT        & 2      & 6      & 5      & 13                  \\ \hline
-    SExp       & 6      & 9      & -      & 15                  \\ \hline
-    HTML       & 4      & -      & 132    & 136                 \\ \hline
+    RBT        & 2       & 5       & 6       & 13                  \\ \hline
+    SExp       & 6       & -       & 9       & 15                  \\ \hline
+    HTML       & 4       & 132     & -       & 136                 \\ \hline
   \end{tabular}
@@
 
+These case studies provide a good combination of constructions, interface
+functions and patterns, and cover from smaller to larger numbers of
+constructions.
+
 We benchmarked the execution time required to generate and fully evaluate 10000
 random values corresponding to each case study, comparing both manually-written
-concrete generators and modular generators obtained using our approach.
+concrete generators and those obtained using our modular approach.
 %
-For this purpose, we used the \emph{Criterion} benchmarking tool for Haskell,
-and limited the maximum depth of the generated values to five levels.
+For this purpose, we used the \emph{Criterion} \cite{criterion} benchmarking
+tool for Haskell, and limited the maximum depth of the generated values to five
+levels.
 %
@@
 Figure \ref{fig:times} illustrates the relative execution time of each case
-study, with respect to each manually-written random generator---we encourage the
-reader to obtain a colored version of this work.
+study, normalized to their corresponding manually-written counterparts---we
+encourage the reader to obtain a colored version of this work.
 
-\begin{figure}[b]
+\begin{figure}[t]
   \centering
@@
 when using linearly defined representations, specially when considering the HTML
-case study with a large amount of construction involved at the generation
+case study, which spans a large amount of constructions in the generation
 process.
@@
 %
-At the light of these improvements in the performance, we included in our tool
-an additional type-level computation that automatically balances our
-representations in order to reduce the generation overhead.
+At the light of these improvements, our tool includes an additional type-level
+computation that automatically balances our representations in order to reduce
+the generation overhead.
 
@@
 testing complex properties over monadic code, as well as using randomly
-generated values for penetration testing.
+generated values for penetration testing
+\cite{DBLP:conf/haskell/MistaRH18,grieco2017}.
 
@@
 the flexibility obtained from using our compositional approach does not produce
-severe slowdowns in whem generating random values in practice.
+severe slowdowns in when generating random values in practice.
 

### 2019-05-03 papers/haskell19/7-related.tex
Work on case studies/related work

--- 2019-03-20 papers/haskell19/relatedwork.tex
@@
 \section{Related Work}
-\label{sec:relatedwork}
+\label{sec:related}
 
-Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
-facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
-vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
-Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
-varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
-ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
-vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
-orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
-est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
+\citeauthor{SwierstraDTC} \cite{SwierstraDTC} proposes the Data Types \`a la
+Carte approach for building extensible data types and unified interpreters as a
+solution for the expression problem coined by \citeauthor{wadler1998expression}
+\cite{wadler1998expression}.
+%
+In this work, we take ideas from this approach and extend them to work in the
+scope of random data generation, where other parameters come into play apart
+from just combining construction, e.g., generation frequency and terminal
+constructions.
+
+From the practical point of view, \citeauthor{KiriyamaOptimizingDTC} propose an
+optimization mechanism for Data Types \`a Carte, where a concrete data type has
+to be derived for each different composition of constructions defined by the
+user \cite{KiriyamaOptimizingDTC}.
+%
+This solution avoids much of the runtime overhead introduced when internally
+pattern matching against sequencies of |InL| and |InR| data constructors.
+%
+However, we see this solution as not entirely compositional, since we still need
+to rely on Template Haskell to derive the machinery for each specialized data
+type.
+%
+In our particular setting, we found that our solution has a fairly acceptable
+overhead by automatically balancing our representation types.
+
+In the past few years, there has been a bloom of derivation tools for
+synthesizing random generators.
+%
+\emph{Feat} provides a mechanism to uniformly generate values from a given data
+type \cite{DuregardJW12}.
+%
+It works by enumerating all the possible values of such type, so that sampling
+uniformly from it simply becomes sampling uniformly from the a finite preffix of
+natural numbers.
+
+\emph{MegaDeTH} is a simple derivation tool that derives generators solely based
+on their types, and paying no attention whatsoever on the generation frequency
+of each data constructor.
+%
+As a result, its automatically derived generators has been shown to be skewed to
+generating small values\cite{DBLP:conf/haskell/MistaRH18}.
+
+On the other hand, \emph{DRAGEN} is a tool that synthesizes optimized
+generators, tuning their generation frequencies using a simulation-based
+optimization process parametrized by the distribution of values desired by the
+user \cite{DBLP:conf/haskell/MistaRH18}.
+%
+This simulation is based on the theory of \emph{branching processes}, which
+models the growth and extinction of populations across succesive generations.
+%
+In this setting, populations consist of randomly generated data constructors,
+where generations correspond to each level of the generated values.
+%
+This tool has shown to improve the code coverage obtained over our codebase,
+with respect to other fully-automated generators derivation mechanism.
+%
+In a recent work, \citeauthor{Mista2019GeneratingRS} extend this approach to
+generate random values considering also the other sources of structural
+information covered here, namely abstract interfaces and function pattern
+machings.
+%
+The authors of this work focus on the generation model problem, extending the
+theory of branching processes in order to obtain sound predictions about
+distributions of random values considering these new kinds of constructions.
+%
+Using this extension, they show that using extra information when generating
+random values is extremely valuable, in particular under situations like the
+ones described in Section \ref{sec:sources}, where the usual derivation
+approaches fail to synthesize useful generators due to a lack of structural
+information.
+
+This paper tackles the representation problem, demonstrating how a compositional
+generation process can be effectively implemented and automated in Haskell using
+advanced type-level features.
+%
+Since none of the aforementioned tools are designed for composability, we
+consider that our ideas could perhaps be applied to improve these tools in the
+future.

### 2019-05-04 papers/haskell19/1-intro.tex
Section 2
MergeMoved stuff from section 2 to intro, work on abstract and conclusions

@@
 
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-vitae placerat urna tortor vitae lacus. Nullam libero mauris, consequat quis,
-varius et, dictum id, arcu. Mauris mollis tincidunt felis. Aliquam feugiat
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+% --------------------------------------
+% Intro
+%
+Random property-based testing is a quite powerful technique for finding bugs
+\cite{ArtsHNS15, HughesPAN16, HughesNSA16}.
+%
+In the Haskell comunity, QuickCheck is the predominant tool of this such.
+%
+The developers specify
+%
+\begin{inparaenum}
+\item the testing properties their systems must fullfil, and
+\item random data generators for the data types involved at their properties.
+\end{inparaenum}
+%
+Then, QuickCheck generates random seeds using these generators, and uses them to
+evaluate the system's properties in search of possible counterexamples, which
+often indicate bugs in our system.
 
+% --------------------------------------
+% Motivating Example
+%
+Although QuickCheck provides default random generators for the common base types
+like |Int| or |String|, it requires implementing random generators for any
+user-derined data type we want to randomly generate.
+%
+% This section introduces the common workflow for generating random values using
+% automatically derived QuickCheck generators, along with its common limitations.
+%
+To illustrate this, let us consider the following data type definition encoding
+basic HTML pages:
+
+\begin{code}
+data Html
+  =  Text  String
+  |  Sing  String
+  |  Tag   String Html
+  |  Join  Html Html
+\end{code}
+
+This type allows us to build HTML pages via four possible data constructors:
+|Text| is used for plain text values; |Sing| and |Tag| represent singular and
+paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
+one after another.
+%
+Note that the constructors |Tag| and |Join| are recursive, as they have at least
+one field of their own type.
+%
+In this paper, and in the scope of random generation in general, we will refer
+to any non-recursive data constructor as \emph{terminal}, and
+\emph{non-terminal} in any other case.
+
+Then, using this representation, the example page\linebreak
+%
+\texttt{<html><b>hello<br>world!</b></html>}
+%
+can be encoded simply as:
+%
+\begin{code}
+Tag "html" $ Tag "b" $
+  Text "hello" iJoin Sing "br" iJoin Text "world!"
+\end{code} %$
+
+% --------------------------------------
+% Random Generators
+%
+Then, in order to generate random values of type |Html|, we need to provide a
+random data generator for it.
+%
+For this purpose, QuickCheck defines the |Arbitrary| type-class for default
+defining random generators as:
+
+\begin{code}
+class Arbitrary (a :: *) where
+  arbitrary :: Gen a
+\end{code}
+%
+where the overloaded symbol |arbitrary :: Gen a| denotes a monadic generator for
+values of type |a|.
+%
+Using this mechanism, we can define a sensible random generator for our |Html|
+data type as follows:
+
+\begin{code}
+instance Arbitrary Html where
+  arbitrary = sized gen
+    where  gen 0 = frequency
+             [  (2,  Text    <$> arbitrary)
+             ,  (1,  Sing    <$> arbitrary) ]
+           gen d = frequency
+             [  (2,  Text    <$> arbitrary)
+             ,  (1,  Sing    <$> arbitrary)
+             ,  (4,  Tag     <$> arbitrary  <*> gen (d-1))
+             ,  (3,  Join    <$> gen (d-1)  <*> gen (d-1)) ]
+\end{code} %$
+%
+At the top level, this definition parametrizes the generation process using
+QuickCheck's |sized| combinator, which lets us build our generator using an
+auxliary local function |gen :: Int -> Gen Html|.
+%
+The |Int| passed to |gen| is known as the \emph{generation size}, and is
+threaded seamlessly by QuickCheck on each call to arbitrary.
+%
+Then, we use this parameter to limit the maximum amount of recursive calls that
+our generator can perform, and thus the maximum depth of the generated values.
+%
+If the generation size is positive (case |gen d|), our random generator picks a
+random constructor from our |Html| data type with a given generation frequency
+(using QuickCheck's |frequency| combinator), and proceeds to fill its fields
+using randomly generated sub-terms---here using Haskell's applicative notation
+\cite{Mcbride2008} and the default |Arbitrary| instance for |String|s.
+%
+For the case of the recursive sub-terms, this generator simply calls itself
+recursively with a smaller generation size (|gen (d-1)|).
+%
+This process repeats until we reach the base case (|gen 0|) on each recursive
+sub-term.
+%
+At this point, our random generator is limited to only pick among the |Html|
+terminal constructors, hence terminating the generation process.
+
+% --------------------------------------
+% Automatic derivation tools
+%
+As one can observe, the previous definition is quite mechanical, and depends
+only on the generation frequencies we choose for each constructor.
+%
+Fortunately, there exists a variety of meta-programming tools for obtaining
+random generators automatically based on their types.
+%
+Tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random generators
+using the same procedure as the one shown above.
+%
+\emph{MegaDeTH} simply picks among data constructors uniformingly, whereas
+\emph{DRAGEN} analyzes the type definitions and tunes the generation frequencies
+based on a simulation-based optimization approach, in order to obtain random
+distributions of values based on the users' desires.
+
+% --------------------------------------
+% Limitation
+%
+Such tools provide a useful mechanism for automating the boring and error-prone
+task of writing random generators by hand.
+%
+However, this simple derivation procedure, which only considers the structural
+information encoded in data type definitions, and generates one single
+constructor at a time; is often too generic to synthesize useful generators in
+some practical scenarios.
+%
+For instance, because of the types we use for deriving our random generators are
+too general, leaving a large room for ill-formed values, e.g., invalid HTML
+pages; or alternatively, because our properties need to be tested against very
+specific patterns of input cases in order to trigger bugs.
+
+This problem, however, can be alleviated by considering additional sources of
+structural information from our codebase other than just data constructors.
+%
+For instance, having a set of high-level combinators, defining an abstract
+interface of our data types, e.g., |hr :: Html| or |bold :: Html -> Html|; it is
+possible to include them into the generation process, thus obtaining more
+complex and structed values.
+%
+In the same way, having functions defined using pattern matching over the data
+types of our interest, e.g. |f (Join (Text t) (Tag s x)) = dots|, it is also
+possible to include this rich structure which exposes a relationship among
+different data constructors into our random generators.
+%
+This idea is addressed in depth in the next section, while the rest of this work
+focuses on implementing it in a compositional way.
+
+%
+% Contributions
+%
 The main contribution of this paper are:
@@
 \begin{itemize}
-\item We identify some of the most important limitations of performing random
-  testing with automatically derived generators using state-of-the-art tools
-  (Section \ref{sec:sources}).
+\item We identify some important limitations of performing random testing with
+  automatically derived generators using state-of-the-art tools (Section
+  \ref{sec:sources}).
 \item We present an extensible mechanism for representing random values built
@@
 generators in a composable fashion, in contrast with the usual paradigm of
-automatically deriving ``closed'' monolithic generators.
+automatically deriving ``rigid'', monolithic generators.

### 2019-05-04 papers/haskell19/2-sources.tex
Work on case studies/related work
MergeMoved stuff from section 2 to intro, work on abstract and conclusions

--- 2019-04-30 papers/haskell19/2-sources.tex
@@
 
-This section introduces the common workflow for generating random values using
-automatically derived QuickCheck generators, along with its common limitations.
+In this section, we briefly introduce the motivation for including extra
+information into our random generators, and discuss about how this can be
+implemented in practice.
+
+% --------------------------------------
+% Abstract interfaces
 %
-To illustrate this, let us consider the following data type definition encoding
-basic HTML pages:
+\subsection{Extracting Information From Interfaces}
 
-\begin{code}
-data Html
-  =  Text  String
-  |  Sing  String
-  |  Tag   String Html
-  |  Join  Html Html
-\end{code}
-
-This type allows us to build HTML pages via four possible data constructors:
-|Text| is used for plain text values; |Sing| and |Tag| represent singular and
-paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
-one after another.
+A common choice when implementing abstract data types is to transfer the
+resposiblity of preserving their invariants to the functions on their abstract
+interface.
 %
-Note that the constructors |Tag| and |Join| are recursive, as they have at least
-one field of their own type.
+Take for example our |Html| data type.
 %
-In this paper, and in the scope of random generation in general, we will refer
-to any non-recursive data constructor as \emph{terminal}, and
-\emph{non-terminal} in any other case.
-
-Then, using this representation the example page\linebreak
-%
-\texttt{<html>hello<hr><b>world!</b></html>}
-%
-can be encoded simply as:
-%
-\begin{code}
-Tag "html" $  Text "hello"
-              iJoin Single "hr"
-              iJoin Tag "b" $ Text "world!"
-\end{code} %$
-
-Then, in order to generate random values of type |Html|, QuickCheck requires us
-to provide a random data generator for it.
-
-\begin{code}
-class Arbitrary (a :: *) where
-  arbitrary :: Gen a
-\end{code}
-
-\begin{code}
-class Arbitrary (a :: *) where
-  arbitrary :: Gen a
-\end{code}
-
-\begin{code}
-instance Arbitrary Html where
-  arbitrary = sized gen
-    where
-      gen 0 = frequency
-        [  (2,  Text    <$> arbitrary)
-        ,  (1,  Sing    <$> arbitrary) ]
-      gen d = frequency
-        [  (2,  Text    <$> arbitrary)
-        ,  (1,  Sing    <$> arbitrary)
-        ,  (4,  Tag     <$> arbitrary  <*> gen (d-1))
-        ,  (3,  Join    <$> gen (d-1)  <*> gen (d-1)) ]
-\end{code} %$
+Instead of defining a different data constructor for each possible HTML tag, we
+opted for a very small generic representation that can be extended via a set of
+combinators making use of this generic interface:
 
@@
 \end{code} %$
+%
+These functions let us create interesting structured values representing HTML
+constructions like linebreaks (|br|), bold blocks (|bold|), unordered lists
+(|list|) and concatenation of values one below another (|<+>|).
+
+Note how difficult would be to generate random values containing, for example,
+structurally valid HTML lists if we only consider the structural information
+encoded in our |Html| data type definion.
+%
+After all, much of the valid structure of HTML has been encoded on its abstract
+interface.
+
+Then, an automatically derived generator could integrate this structural
+information into the generation process by creating random values arising from
+applying such functions to randomly generated inputs:
 
@@
 \end{code}%$
+%
+where the symbol (|...|) represents the rest of the possible sub-generators for
+each data constructor seen before.
+%
+From now on, we will refer to each possible choice given to |frequency| as a
+different \emph{random construction}, as we are not considering generating only
+ssingle data constructors anymore.
+
+% --------------------------------------
+% Functions' pattern matchings
+%
+\subsection{Extracting Information From Patterns}
+
+A different challenge appears when we try to test functions comprising complex
+pattern matchings. Consider for instance the following function for simplifying
+|Html| values:
 
@@
 simplify :: Html -> Html
-simplify (Text t1 iJoin Text t2)
-  = Text (t1 ++ t2)
+simplify (Text t1 iJoin Text t2) = Text (t1 ++ t2)
 simplify ((Text t1 iJoin x) iJoin y)
   = simplify (Text t1 iJoin simplify (x iJoin y))
-simplify (x iJoin y)
-  = simplify x iJoin simplify y
-simplify (Tag t x)
-  = Tag t (simplify x)
+simplify (x iJoin y) = simplify x iJoin simplify y
+simplify (Tag t x) = Tag t (simplify x)
 simplify x = x
 \end{code}
+%
+This function simply traverses |Html| values, joining together every contiguous
+pair of |Text| constructors.
+%
+Ideally, we would like to put approximately the same testing effort info each
+clause of |simplify|, or perhaps even more to first two ones, since those are
+the ones performing actual simplifications.
+%
+However, these two clauses are the most difficult ones to test in practice.
+%
+The probabilty of generating a random value satisfying nested patterns decreases
+multiplicatively with the number of constructors we match against to
+simultaneously.
+
+In our tests, we were not able to exercise any of these two patterns more than
+6\% of the overall testing time (and considering recursive sub-calls), using
+random generators derived using either \emph{MegaDeTH} and \emph{DRAGEN}.
+%
+As expected, most of the random test cases were exercising the simplest (an
+rather uninteresting) patterns of this function.
+
+To solve this issue, we could opt to consider each complex pattern as a new kind
+of random construction.
+%
+In this light, we can simply generate values satisfying pattern directly by
+returning its corresponding expression, where each variable or wildcard pattern
+is filled using a random sub-expression:
 
@@
 \end{code}%$
+
+% --------------------------------------
+% Wrapping up
+%
+In general, we think that including this static information in our genrators
+directly is far from ideal, as in many cases we would like to select which
+random constructions are of our interest when testing different properties or
+sub-systems.
+%
+In this light, having ``monolithic'' random generators derived automatically
+would complicate our desired workflow.
+%
+We would need to derive different concrete random generators for each use case
+instead of having a single |Arbitrary| instance that can be used everywhere.
+
+In contrast, in this work we opted for developing an extensible fashion, where
+the generation machinery is automatically derived only once, and every variant
+of our random generators is expressed at the type-level.
+%
+The following sections of this paper cover our alternative approach in depth.

### 2019-05-04 papers/haskell19/3-representation.tex
MergeMoved stuff from section 2 to intro, work on abstract and conclusions

@@
 \end{code}
-
+%
 where we simply use the appropriate |Algebra| instance for the inner
@@
 %
-However, we found that in practice, the way in which we decide to associate our
-constructions affects the generation speed dramatically.
+However, in practice, the way in which we associate our constructions affects
+the generation speed dramatically.
 %
@@
 %
-For simplicity here, we will assume |oplus| as having rigth associativity, i.e.,
+For simplicity here, we will assume |oplus| as having right associativity, i.e.,
 |f oplus'' g oplus'' h == f oplus'' (g oplus'' h)|.
@@
 %
-Where the sequences of |InL| and |InR| data constructors inject each value from
+where the sequences of |InL| and |InR| data constructors inject each value from
 an individual representation into the appropriate position of our composite
@@
 %
-Where the number after the |#| simply distinguishes the different patterns from
+where the number after the |#| simply distinguishes the different patterns from
 the function |simplify| using the index of the clause they belong to.

### 2019-05-04 papers/haskell19/8-conclusions.tex
Work on case studies/related work
MergeMoved stuff from section 2 to intro, work on abstract and conclusions

@@
 
-In the future, something something...
+In the future, we aim to minimize the runtime overhead of our technique, using
+perhaps prefix-free encodings based on generation frequencies, as a way to
+reduce the representation bureacracy as much as possible.
+%
+Additionally, we are interested on including additional sources of structural
+infrmation from users' codebases, e.g., locally defined patterns or sequencies
+of functions applications.
+%
+Interesting thoughts to motivate our future reseach.

### 2019-05-04 papers/haskell19/abstract.tex
MergeMoved stuff from section 2 to intro, work on abstract and conclusions

@@
-Pellentesque dapibus suscipit ligula. Donec posuere augue in quam. Etiam vel
-tortor sodales tellus ultricies commodo. Suspendisse potenti. Aenean in sem ac
-leo mollis blandit. Donec neque quam, dignissim in, mollis nec, sagittis eu,
-wisi. Phasellus lacus. Etiam laoreet quam sed arcu. Phasellus at dui in ligula
-mollis ultricies. Integer placerat tristique nisl. Praesent augue. Fusce
-commodo. Vestibulum convallis, lorem a tempus semper, dui dui euismod elit,
-tellus ut neque. Nulla facilisis, risus a rhoncus fermentum, tellus tellus
-lacinia purus, et dictum nunc justo sit amet elit.
+Generating ``good'' random values described by algebraic data types is often
+quite intricate.
+%
+State-of-the-art tools for synthesizing random generators serve the valuable
+purpose of helping with this task, while providing different level of invariants
+imposed over the generated values.
+%
+However, none of them are built neither for composability nor extendability, a
+useful feature when the shape of our random data makes a difference while
+testing different properties or sub-systems.
+%
+% Instead, users are forced to decide wether to manually write their random
+% generators or to use rigid automatically derived ones.
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
-tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
-eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
-Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
-Nam vestibulum accumsan nisl.
+In this work, we develop an extensible framework for deriving compositional
+generators, which can be easily combined in diferent ways in order to fit the
+developers' demands, using a simple type-level domain-specific language.
+%
+Our framework is based on familiar ideas from the \`a la Carte technique for
+writing composable interpreters in Haskell, and extended with the required
+machinery for it to work in the scope of random generation of values.
+%
+Additonally, we provide an implementation of our ideas, and evaluate its
+performance using real-world examples.

### 2019-05-06 papers/haskell19/1-intro.tex
Pass over abstract and Section 1

@@
 %
-Random property-based testing is a quite powerful technique for finding bugs
+Random property-based testing is a powerful technique for finding bugs
 \cite{ArtsHNS15, HughesPAN16, HughesNSA16}.
 %
-In the Haskell comunity, QuickCheck is the predominant tool of this such.
+In the Haskell community, QuickCheck is the predominant tool of this such.
 %
@@
 %
-\begin{inparaenum}
-\item the testing properties their systems must fullfil, and
-\item random data generators for the data types involved at their properties.
+\begin{inparaenum}[(i)]
+\item the testing properties their systems must fulfill, and
+\item random data generators (or generators for short) for the data types
+  involved at their properties.
 \end{inparaenum}
 %
-Then, QuickCheck generates random seeds using these generators, and uses them to
-evaluate the system's properties in search of possible counterexamples, which
-often indicate bugs in our system.
+Then, QuickCheck generates random seeds using these generators, which are used
+to evaluate the system's properties in search of possible counterexamples, which
+often indicate the presence of bugs.
 
@@
 %
-Although QuickCheck provides default random generators for the common base types
-like |Int| or |String|, it requires implementing random generators for any
-user-derined data type we want to randomly generate.
+Although QuickCheck provides default generators for the common base types, like
+|Int| or |String|, it requires implementing generators for any user-defined data
+type we want to generate.
 %
@@
 \begin{code}
-data Html
-  =  Text  String
+data Html  =  Text  String
            |  Sing  String
@@
 \end{code}
-
+%
 This type allows us to build HTML pages via four possible data constructors:
 |Text| is used for plain text values; |Sing| and |Tag| represent singular and
-paired HTML tags, respectively; and |Join| to simply concatenate two HTML pages
+paired HTML tags, respectively; and |Join| simply concatenates two HTML pages
 one after another.
@@
 %
-In this paper, and in the scope of random generation in general, we will refer
-to any non-recursive data constructor as \emph{terminal}, and
-\emph{non-terminal} in any other case.
+In this paper, we will refer to any non-recursive data constructor as
+\emph{terminal}, and \emph{non-terminal} in any other case.
 
-Then, using this representation, the example page\linebreak
+Using this representation, the example page \linebreak
 %
@@
 %
-can be encoded simply as:
+is encoded as:
 %
@@
 %
-Then, in order to generate random values of type |Html|, we need to provide a
-random data generator for it.
+As mentioned before, in order to generate random |Html| values, we need to
+provide a random generator for it.
 %
-For this purpose, QuickCheck defines the |Arbitrary| type-class for default
-defining random generators as:
+For this purpose, QuickCheck defines the |Arbitrary| type-class for defining
+default generators as:
 
@@
 At the top level, this definition parametrizes the generation process using
-QuickCheck's |sized| combinator, which lets us build our generator using an
-auxliary local function |gen :: Int -> Gen Html|.
+QuickCheck's |sized| combinator, which lets us build our generator via an
+auxiliary local function |gen :: Int -> Gen Html|.
 %
@@
 If the generation size is positive (case |gen d|), our random generator picks a
-random constructor from our |Html| data type with a given generation frequency
-(using QuickCheck's |frequency| combinator), and proceeds to fill its fields
-using randomly generated sub-terms---here using Haskell's applicative notation
+random |Html| constructor with a given generation frequency (using QuickCheck's
+|frequency| combinator), and proceeds to fill its fields using randomly
+generated sub-terms---here using Haskell's applicative notation
 \cite{Mcbride2008} and the default |Arbitrary| instance for |String|s.
 %
-For the case of the recursive sub-terms, this generator simply calls itself
+For the case of the recursive sub-terms, this generator simply calls |gen|
 recursively with a smaller generation size (|gen (d-1)|).
@@
 %
-At this point, our random generator is limited to only pick among the |Html|
-terminal constructors, hence terminating the generation process.
+At this point, the generator is limited to pick only among terminal |Html|
+constructors, hence ending the generation process.
 
@@
 Fortunately, there exists a variety of meta-programming tools for obtaining
-random generators automatically based on their types.
+random generators automatically, based on the information encoded on the
+definition of the types we want to generate.
 %
 Tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random generators
-using the same procedure as the one shown above.
+using roughly the same procedure as the one shown above.
 %
-\emph{MegaDeTH} simply picks among data constructors uniformingly, whereas
-\emph{DRAGEN} analyzes the type definitions and tunes the generation frequencies
-based on a simulation-based optimization approach, in order to obtain random
-distributions of values based on the users' desires.
+On one hand, \emph{MegaDeTH} derives generators which simply pick among data
+constructors with uniform probabilities, whereas \emph{DRAGEN} analyzes type
+definitions, tuning the generation frequencies based on a predictive mechanism.
+%
+This allows the users to specify the desired distribution of values of the
+derived generators.
 
@@
 %
-Such tools provide a useful mechanism for automating the boring and error-prone
-task of writing random generators by hand.
+These tools provide a useful mechanism for automating the repetitive and
+error-prone task of writing random generators by hand.
 %
-However, this simple derivation procedure, which only considers the structural
-information encoded in data type definitions, and generates one single
-constructor at a time; is often too generic to synthesize useful generators in
-some practical scenarios.
+However, this simple derivation procedure, which
 %
-For instance, because of the types we use for deriving our random generators are
-too general, leaving a large room for ill-formed values, e.g., invalid HTML
-pages; or alternatively, because our properties need to be tested against very
-specific patterns of input cases in order to trigger bugs.
+\begin{inparaenum}[(i)]
+\item only considers the structural information encoded in type definitions, and
+\item generates one single constructor at a time;
+\end{inparaenum}
+%
+is often too generic to synthesize useful generators in practical scenarios.
+%
+For instance, because of the types we use for deriving our generators are too
+general, leaving a large room for ill-formed values, e.g., invalid HTML pages;
+or alternatively, because our properties need to be tested against very specific
+input patterns in order to trigger bugs.
 
 This problem, however, can be alleviated by considering additional sources of
-structural information from our codebase other than just data constructors.
+structural information from our codebase other than just type definitions
+\cite{Mista2019GeneratingRS}.
 %
@@
 possible to include them into the generation process, thus obtaining more
-complex and structed values.
+complex and structured values.
 %
@@
 %
+% \vspace{-10pt}
+%
 \begin{itemize}
@@
   the generation process at the type level (Section \ref{sec:generators}).
-\item We present a very simple yet powerful domain specific language for
-  describing extensible generators solely based on the types used to represent
-  the desired shape of our random data (Section \ref{sec:typelevel}).
+\item We propose a simple type-level idiom for describing extensible generators,
+  based on the types used to represent the desired shape of our random data
+  (Section \ref{sec:typelevel}).\looseness=-1
 \item We provide a Template Haskell tool for automatically deriving all the

### 2019-05-06 papers/haskell19/abstract.tex
Pass over abstract and Section 1

@@
 In this work, we develop an extensible framework for deriving compositional
-generators, which can be easily combined in diferent ways in order to fit the
-developers' demands, using a simple type-level domain-specific language.
+generators, which can be easily combined in different ways in order to fit the
+developers' demands, using a simple type-level description language.
 %
-Our framework is based on familiar ideas from the \`a la Carte technique for
-writing composable interpreters in Haskell, and extended with the required
-machinery for it to work in the scope of random generation of values.
+Our framework relies on familiar ideas from the \`a la Carte technique for
+writing composable interpreters in Haskell, extended with the machinery required
+in the scope of random generation.
 %
-Additonally, we provide an implementation of our ideas, and evaluate its
-performance using real-world examples.
+We provide an implementation of our ideas, and evaluate its performance using
+real-world examples.

### 2019-05-06 papers/haskell19/2-sources.tex
Pass over section 2

@@
 A common choice when implementing abstract data types is to transfer the
-resposiblity of preserving their invariants to the functions on their abstract
+responsibility of preserving their invariants to the functions on their abstract
 interface.
@@
 opted for a very small generic representation that can be extended via a set of
-combinators making use of this generic interface:
+combinators making use of this basic interface:
 
@@
 These functions let us create interesting structured values representing HTML
-constructions like linebreaks (|br|), bold blocks (|bold|), unordered lists
+constructions like line-breaks (|br|), bold blocks (|bold|), unordered lists
 (|list|) and concatenation of values one below another (|<+>|).
 
-Note how difficult would be to generate random values containing, for example,
-structurally valid HTML lists if we only consider the structural information
-encoded in our |Html| data type definion.
+Note how difficult would it be to generate random values containing, for
+example, structurally valid HTML lists if we only consider the structural
+information encoded in our |Html| type definition.
 %
@@
 %
-where the symbol (|...|) represents the rest of the possible sub-generators for
-each data constructor seen before.
+where the symbol (|...|) represents the rest of the possible choices of data
+constructors seen before.
 %
@@
 different \emph{random construction}, as we are not considering generating only
-ssingle data constructors anymore.
+single data constructors anymore.
 
@@
 Ideally, we would like to put approximately the same testing effort info each
-clause of |simplify|, or perhaps even more to first two ones, since those are
-the ones performing actual simplifications.
+clause of |simplify|, or perhaps even more to the first two ones, since those
+are the ones performing actual simplifications.
 %
-However, these two clauses are the most difficult ones to test in practice.
+However, these two clauses are the most difficult ones to test in practice!
 %
-The probabilty of generating a random value satisfying nested patterns decreases
-multiplicatively with the number of constructors we match against to
-simultaneously.
+The probability of generating a random value satisfying nested patterns
+decreases multiplicatively with the number of constructors we simultaneously
+pattern match against to.
 
@@
 As expected, most of the random test cases were exercising the simplest (an
-rather uninteresting) patterns of this function.
+rather uninteresting) patterns of this function, corresponding to its last three
+clauses.
 
@@
 %
-In general, we think that including this static information in our genrators
+In general, we think that including this static information in our generators
 directly is far from ideal, as in many cases we would like to select which
@@
 
-In contrast, in this work we opted for developing an extensible fashion, where
-the generation machinery is automatically derived only once, and every variant
+In contrast, in this work we opted for developing an extensible mechanism, where
+the generation machinery is automatically derived just once, and every variant
 of our random generators is expressed at the type-level.

### 2019-05-06 papers/haskell19/2-sources.tex
Pass over Section 3

@@
 simplify :: Html -> Html
-simplify (Text t1 iJoin Text t2) = Text (t1 ++ t2)
-simplify ((Text t1 iJoin x) iJoin y)
+simplify (Join (Text t1) (Text t2)) = Text (t1 ++ t2)
+simplify (Join (Join (Text t) x) y)
   = simplify (Text t1 iJoin simplify (x iJoin y))

### 2019-05-06 papers/haskell19/3-representation.tex
Pass over Section 3

@@
 type we use to represent it during the generation process.
-%
+
 For this, we will define a set of different simple ``open'' types, each one in
-charge of representing a single construction of our target data type.
+charge of representing a single construction of our \emph{target} data type,
+i.e., the actual data type we want to randomly generate.
 %
@@
 \item combined in several ways depending on the desired shape of our test data,
-  applying the familiar technique of Data Types \`a la Carte;
-\item randomly generated (see Section \ref{sec:generators}), and finally,
-\item transformed back into the corresponding values of the target data type
+  applying the familiar \`a la Carte technique;
+\item randomly generated (see Section \ref{sec:generators}); and finally,
+\item transformed back to the corresponding values of the target data type
   automatically.
@@
 %
-In this light, each individual data constructor of the target data type can be
-represented simply by a new data type with a single constructor.
+In this light, each data constructor of the target data type can be represented
+using a single-constructor data type.
 %
-Recalling our |Html| example, its data constructors can be represented as
-follows:
+Recalling our |Html| example, its data constructors can be represented as:
+%
+\looseness=-1 % I hate having to use this!
 
@@
 %
-In the previous definitions, each representation has the same fields as its
-corresponding constructor, except for the recursive ones which are abstracted
-away using a type parameter |r|.
+Note that each representation has the same fields as its corresponding
+constructor, except for the recursive ones which are abstracted away using a
+type parameter |r|.
 %
@@
 %
-It is easy to note how each one of these simple representations encode the
-minimal amount of information they need, leaving everything else unspecified.
+It is easy to note how each of these simple representations encode the minimal
+amount of information they need, leaving everything else unspecified.
 
@@
 cases, they form a functor over its type parameter, and thus we can use
-Haskell's |deriving| mechanism to obtain a suitable |Functor| instances for
-free:
+Haskell's |deriving| mechanism to obtain suitable |Functor| instances for free:
 
@@
 %
-In the previous definition, |f| is the functor type used to represent a
-construction of the target type |a|.
+where |f| is the functor type used to represent a construction of the target
+type |a|.
 %
-Note the functional dependency |f -> a|, helping the type checker to solve the
-free type variable |a| at the right hand side of the |=>|.
+Note the functional dependency |f -> a|, helping the type checker to solve
+uniquely the free type variable |a| at the right hand side of the |=>|, based on
+the type |f| used to represent it.
 %
@@
 %
-To do so we can simply transform each representation into its corresponding
-value by picking its corresponding data constructor, and piping its fields
-unchanged:
+To do so, we can simply transform each constructor representation by picking its
+corresponding data constructor, and piping its fields unchanged:
 
@@
 \end{code}
-%
+
 Note how every definition regarding each different constructor so far was
@@
 %
-We can think of them as a encoding a single ``layer'' of our target data.
+We can think of them as a encoding a `'single layer'' of our target data.
 %
 In order to represent recursive values, we need to close them ``tying the knot''
-recusively, i.e., once we have decided the final representation of our target
+recursively, i.e., once we have fixed a suitable representation for our target
 data, we want each one of its recursive fields to be instantiated with itself.
@@
 %
-where the sequences of |InL| and |InR| data constructors inject each value from
-an individual representation into the appropriate position of our composite
-representation |Html'|.
+where the sequences of |InL| and |InR| data constructors \emph{inject} each
+value from an individual representation into the appropriate position of our
+composite representation |Html'|.
 
@@
 This recusion schema using functors and F-Algebras to fold data generically has
-been well studied and is known categorically a \emph{catamorphism}.
+been well studied and is known categorically a \emph{catamorphism} \tocite.
 
@@
 
-Fistly, let us consider the case of generating values obtained by abstract
+Firstly, let us consider the case of generating values obtained by abstract
 interface functions.
 %
-If we recall our |Html| example, we can observe that it is possible to use the
+If we recall our |Html| example, we note that it is possible to use the
 functions on its abstract interface to obtain |Html| values based on different
@@
 bold   ::  Html    ->          Html
-list   ::  [  Html]  ->          Html
+list   ::  [Html]  ->          Html
 (<+>)  ::  Html    -> Html ->  Html
 \end{code}
-
+%
 Fortunately, it is easy to extend our approach to incorporate the interesting
@@
 Each one of these data types represents a value resulting from evaluating its
-corresponding function, using as input arguments the values encoded in its
+corresponding function, using as input arguments the values encoded on its
 fields.
@@
 argument) with a type parameter |r| in order to leave the type of the recursive
-sub-terms unspecified until we have decided the overall shape of our data.
+sub-terms unspecified until we have decided the final shape of our data.
 
@@
 Instead, the actual functions are evaluated when transforming our
-representations back to their target type, again by the means of an |Algebra|:
+representations back to their target type, by the means of an |Algebra|:
 
@@
 %
-There, we simply return the result of evaluating the appropriate functions,
-using each representation field as an input argument.
+There, we simply return the result of evaluating each function, using its
+representation fields as an input arguments.
 
@@
 \end{code}
-
-We can extend our approach further in order to represent these patterns as well.
+%
+We can extend our approach in order to represent these patterns as well.
 %
@@
 the fields of each single data constructor the free pattern variables (or
-wildcards) appearing in corresponding patterns from left to right:
+wildcards) appearing in its corresponding pattern, from left to right:
 
@@
 %
-where the number after the |#| simply distinguishes the different patterns from
-the function |simplify| using the index of the clause they belong to.
+where the number after the |#| distinguishes the different patterns from the
+function |simplify| by the index of the clause they belong to.
 %
@@
 a recursive pattern variable or wildcard) using a type variable |r| until we
-have decided the overall shape of our target data.
+have decided the final shape of our target data.
 
-Then, the task of our |Algebra| instances is to expand each pattern
-representation into the corresponding target value resembling such pattern,
-where each pattern variable gets instantiated using the values stored in its
-representation field:
+Then, each |Algebra| instance will expand pattern representations into the
+corresponding target values resembling such patterns, where each pattern
+variable gets instantiated using the values stored in its representation field:
 
@@
 
-Using this extensions, we are now able to represent values of our target data
-using a higher abstraction level.
+Using the constructions presented so far, we are now able to represent values of
+our target data using a higher level of abstraction.
 %
@@
 as we would like to generate random test cases following different strategies
-when testing different properties.
+when testing different properties or sub-systems.
 %
@@
 inputs, we could avoid generating random |Sing| and |Tag| data constructors, as
-they might represent invalid tags.
+they might introduce invalid HTML tags.
 %
-Instead, we could specify that our random data can be constructed using only
+Instead, we could specify that our random data can be constructed only via
 ``safe'' constructions, using perhaps a representation type like follows:
@@
 \begin{code}
-type ValidHtml = Con_Text oplus'' Con_Join oplus'' Fun_br oplus'' Fun_bold
+type ValidHtml  =      Con_Text oplus'' Con_Join
+                oplus  Fun_br oplus'' Fun_bold oplus'' Fun_list oplus'' Fun_cat
 \end{code}
@@
 The next section introduces a generic way to generate random values from our
-different representations, extending them with a set of type-level combinators
-to encode information relevant to the generation process directly on their
-types.
+different representations, extending them with a set of combinators to encode
+information relevant to the generation process directly at the type level.

### 2019-05-06 papers/haskell19/4-generators.tex
Pass over Section 4

@@
 This modularity allows us to derive each different construction independently,
-as well to specify the shape of our target data with with a fine granularity.
+as well to specify the shape of our target data with a fine granularity.
 
@@
 In this light, our random generators are built from smaller ones (each one
-representing a single random construction), and they are based solely on the
-types we use to represent the shape of our random data.
+representing a single random construction), and are solely based on the types we
+use to represent the shape of our random data.
 
 Ideally, our aim is to be able to obtain random generators with a behavior
-similar to the one presented for |Html| in Section \ref{sec:sources}.
+similar to the one presented for |Html| in Section \ref{sec:intro}.
 %
@@
 \begin{itemize}
-\item We use the QuickCheck's generation size to limit the depth of the
-  generated values, reducing it by one on each recursive call of the local
-  auxiliary function |gen|.
+\item We use QuickCheck's generation size to limit the depth of the generated
+  values, reducing it by one on each recursive call of the local auxiliary
+  function |gen|.
 \item We differentiate between \emph{terminal} and \emph{non-terminal
@@
 %
-For the rest of this section, we will focus on modelling these previous aspects
-in our modular framework, and in such a way that we do not compromise the
-compositionality obtained so far.
+For the rest of this section, we will focus on modeling these aspects in our
+modular framework, in such a way that does not compromise the compositionality
+obtained so far.
 
@@
 %
-If we recall the random generator for |Html| from Section \ref{sec:sources}, we
+If we recall the random generator for |Html| from Section \ref{sec:intro}, we
 can observe that the depth parameter |d| is threaded to the different recursive
-call of our generator, but always within the scope of the local function |gen|.
+calls of our generator, always within the scope of the local function |gen|.
 %
@@
 %
-Instead, we will define a depth-bounded generator data type, wrapping
+Instead, we will define a data type for depth-bounded generators, wrapping
 QuickCheck's |Gen| data type with a external parameter representing the current
@@
 \end{code}
-
-A |BGen| is essentially a normal QuickCheck |Gen| with the maximum recursive
+%
+A |BGen| is, essentially, a normal QuickCheck |Gen| with the maximum recursive
 depth as an input parameter.
 %
-Using this previous definition, we can generalize QuickCheck's |Arbitrary| class
-to work with depth-bounded generators simply as follows:
+Using this definition, we can generalize QuickCheck's |Arbitrary| class to work
+with depth-bounded generators simply as follows:
 
@@
 
-From now on, we will use this type class as a more flexibe substitute of
+From now on, we will use this type class as a more flexible substitute of
 |Arbitrary|, given that now we have two parameters to tune: the maximum
@@
 %
-Where the former is useful to tune the overall size of our random data, whereas
-the latter can be use for tuning the values of the ``leaves'', such as the
-maximum length of random |String| or the maximum/minumum random |Int|s.
+The former is useful to tune the overall size of our random data, whereas the
+latter can be used for tuning the values of the ``leaves'', such as the maximum
+length of random |String| or the biggest/smallest random |Int|s.
 
-Here we want to actively remark that, even though we could have reused the
+Here we want to actively remark that, even though we could have used the
 QuickCheck generation size parameter to simultaneously model the maximum
 recursive depth and the maximum size of the leaves, doing so would imply
-generating random leaves with a decresing size as we move deeper within a random
+generating random leaves with a decreasing size as we move deeper within a random
 value, obtaining for instance, random trees with all zeroes on its leaves, or
-random lists showing a hight tendency to be ordered in decreasing order.
+random lists skewed to be ordered in decreasing order.
 %
@@
 parametric on the type of its recursive sub-terms, as a way to defer this choice
-until we have specified the overal shape of our target data.
+until we have specified the final shape of our target data.
 %
@@
 If we want to define our generators in a modular way, we also need to
-parametrize somehow the generation of the recursive subterms!
+parameterize somehow the generation of the recursive sub-terms!
 %
 If we look at QuickCheck, this library already defines a type class |Arbitrary1|
-for parametric types of kind |* -> *|:
+for parametric types of kind |* -> *|, which solves this issue by receiving the
+generator for the parametric sub-terms as an argument:
 
@@
 \end{code}
-
+%
 % This type class allows us to write modular generators, leaving unspecified the
@@
 % like:
-
+%
 % \begin{code}
@@
 % \end{code} %$
-
+%
 % Where the random generator use to generate the value stored in a |Just|
@@
 %
-We can use this same mechanism for our modular generators, first extending it to
-be depth-aware as follows:
+Then, we can use this same mechanism for our modular generators, extending
+|Arbitrary1| to be depth-aware as follows:
 
@@
 %
-Note the similarity between the |Arbitrary1| and |BArbitrary1| type classes.
+Note the similarities between the |Arbitrary1| and |BArbitrary1| type classes.
 %
-Then, we will use this type class to implement random generators for each
-construction we are considering.
+We will use this type class to implement random generators for each construction
+we are considering.
 %
-Recalling our |Html| example, we can define modular random generator for the
+Recalling our |Html| example, we can define modular random generators for the
 constructions representing data constructors as follows:
@@
 
-Note in the previous definitions how each instance is defined to be parametric
-of the maximum depth (using the input integer |d|) and the random generator used
-for the recursive sub-terms (using input bounded generators |bgen|).
+Note how each instance is defined to be parametric of the maximum depth (using
+the input integer |d|) and the random generator used for the recursive sub-terms
+(using the input generator |bgen|).
 %
@@
 
-The rest of our construction representations can be generated analogously.
+The rest of our representations can be generated analogously.
 %
@@
 
-Having all the modular generators for each construction in place, we can finally
+Having the modular generators for each random construction in place, we can
 obtain a concrete depth-aware generator (of kind |*|) for any final
@@
 \end{code} %$
-
+%
 Here we used the |BArbitrary1| instance of our final representation |f| to
 generate sub-terms recursively by lifting itself (|liftBGen barbitrary|) as the
-parametrized input generator, wrapping each recursive sub-term with a |FixC|
+parameterized input generator, wrapping each recursive sub-term with a |FixC|
 data constructor.
@@
 However, we still need to develop our generation mechanism a bit further in
-order to generate composite representations built using the |oplus| type-level
-operator.
+order to generate composite representations built using the |oplus| operator.
 %
@@
 
-As we have shown so far, generating each construction representation is rather
+As we have shown so far, generating each representation is rather
 straightforward: there is only one data constructor to select, and every field
-is generating following a mechanical recipe.
+is generated following a mechanical recipe.
 %
 In our approach, most of the generation complexity is encoded in the random
-generator used for composite representations built upon the |oplus| operator.
+generator used for composite representations, which are built upon the |oplus|
+operator.
 %
@@
 
-Recalling the random generator for |Html| presented in Section
-\ref{sec:sources}, we can observe that the last generation level (see |gen 0|)
-is constrained to generate only the non-recursive subset of all the available
-constructions.
+Recalling the random generator for |Html| presented in Section \ref{sec:intro},
+we can observe that the last generation level (see |gen 0|) is constrained to
+generate values only from the subset of terminal constructions.
 %
-In order to model this behavior in our framework, we can start by defining a
-data type |Term| to ``tag'' every terminal construction explicitly:
+In order to model this behavior, we will first defined a data type |Term| to
+``tag'' every terminal construction explicitly:
 
@@
 \end{code}
-
-Then, if |f| is a non-recursive construction, the type |Term f oplus' g| can be
-interpreted as representing values generated using values drawn both from |f|
-and |g|, but ``closed'' using only values from |f|.
 %
-Since this data type will not add any semantic information to our represented
+Then, if |f| is a terminal construction, the type |Term f oplus' g| can be
+interpreted as representing data generated using values drawn both from |f| and
+|g|, but ``closed'' using only values from |f|.
+%
+Since this data type will not add any semantic information to the represented
 values, we can define suitable |Algebra| and |BArbitrary1| instances for it
@@
 %
-For this purpose, we can follow the same reasoning as before, defining a data
+For this purpose, we can follow the same reasoning as before, defining a
 type-level operator |otimes| to explicitly tag the generation frequency of a
@@
 \end{code}
-
-In the definition above, the operator |otimes| is parametrized by a type-level
-natural number |n| (of kind |Nat|) representing the desired generation
-frequency.
 %
-In this light, the type |(f otimes'' 3) oplus'' (g otimes'' 1)| represents
-values generated using values from both |f| and |g|, where |f| is randomly
-chosen three times more frequently than |g|.
+This operator is parameterized by a type-level natural number |n| (of kind |Nat|)
+representing the desired generation frequency.
+%
+In this light, the type |(f otimes'' 3) oplus'' (g otimes'' 1)| represents data
+generated using values from both |f| and |g|, where |f| is randomly chosen three
+times more frequently than |g|.
 %
@@
 
-With these two new type level combinators |Term| and |otimes|, we are now able
+With these two new type level combinators, |Term| and |otimes|, we are now able
 to express the behavior of our entire generation process based solely on the
 type we are generating.
-%
+
 In addition to these combinators, we will need to perform some type-level
@@
 \end{code}
-
+%
 Our generation process will traverse this type one combinator at a time,
@@
 For this purpose, we can implement a closed type family |FreqOf| to compute the
-overal sum of frequencies of a given representation type:
+overall sum of frequencies of a given representation type:
 
@@
 \end{code}
-
+%
 This type-level function takes a representation type as an input and traverses
@@
 otimes'' 6) oplus'' (g otimes'' 3)|, following the natural intuition for the
-addition and multiplication operations over numbers.
+addition and multiplication operations over natural numbers.
 %
-Moreover, if a type does have an explicit frequency then its frequency is simply
-defaulted to one.
+Moreover, if a type does have an explicit frequency, then its frequency is
+simply defaulted to one.
 
 On the other hand, the last step of our generation process, which only generates
-terminal constructions, could be seen as treating the non-terminal ones as
+terminal constructions, could be seen as considering the non-terminal ones as
 having generation frequency zero.
 %
-This way we can introduce antoher type-level computation to calculate the
+This way, we will introduce another type-level computation to calculate the
 \emph{terminal generation frequency} |FreqOf'| of a given representation as
@@
 \end{code}
-
+%
 Similar to |FreqOf|, the type family above traverses its input type adding the
@@
 
-Using the typle level combinators introduced at the beginning of this
-sub-section along with the previous type-level computations over frequencies, we
-are finally in position of defining our random generator for composite
-representations:
+Then, using the |Term| and |otimes| combinators introduced at the beginning of
+this sub-section, along with the previous type-level computations over
+frequencies, we are finally in position of defining our random generator for
+composite representations:
 
@@
 \end{code}
-
-Like the generator introduced in Section \ref{sec:sources} for |Html|, this
+%
+Like the generator for |Html| introduced in Section \ref{sec:intro}, this
 generator branches over the current generation depth |d|.
@@
 \end{code}
-
+%
 Note that the type of |freqVal| is ambiguous, since it quantifies over every
@@
 %
-We need to use a \emph{visible type application} (employing the |at (...)|
-syntax) to desambiguate to which natural value we are actually refering to.
+We use a \emph{visible type application} (employing the |at (...)| syntax) to
+disambiguate to which natural value we are actually referring to.
 %
@@
 
-The remaining part of our random generator works analogously, except that this
-time we only want to generate terminal constructions, hence we use the |FreqOf'|
-type family to compute the terminal generation frequency of each operand.
+The rest of our random generator works analogously, except that, this time we
+only want to generate terminal constructions, hence we use the |FreqOf'| type
+family to compute the terminal generation frequency of each operand.
 %
@@
 If it happens that both |FreqOf' f| and |FreqOf' g| compute to zero when
-genrating a composite value, then it will produce a runtime error triggered by
-the function |frequency|, as it does not have anything with a positive frequency
-to generate.
+generating values from composite representation, then it will produce a runtime
+error triggered by the function |frequency|, as it does not have anything with a
+positive frequency to generate.
 %
 This kind of exceptions will arise, for example, if we forget to tag at least a
-single construction within a composite representation as terminal.
+one construction from our final representation as terminal.
 %
-Fortunately, such runtime exceptions can be catched in compile-time by defining
-a type constraint that ensures we are trying to generate values using a
-representation with a strictly positive terminal generation frequency---thus
-containing at least a single terminal construction:
+Fortunately, such runtime exceptions can be catched at compile-time.
+%
+We can define a type constraint |HasTerminals| that ensures we are trying to
+generate values using a representation with a strictly positive terminal
+generation frequency---thus containing at least a single terminal construction:
 
@@
 \end{code}
-
-The previous type families compute the terminal generation frequency of a
-representation type |f|, returning a type error if its result is zero, or an
-empty constraint |()| that is always trivially satisfied in any other case.
-
+%
+These type families compute the terminal generation frequency of a
+representation type |f|, returning either a type error, if its result is zero;
+or, alternatively, an empty constraint |()| that is always trivially satisfied.
 %
 Finally, we can use this constraint to define a safe generation primitive
-|genEval| to obtain a concrete depth-bounded generator for every target type |a|
-specified using a safe generation representation |f|:
+|genRep| to obtain a concrete depth-bounded generator for every target type |a|,
+specified using a safe representation |f|:
 
@@
 \end{code} %$
-
+%
 Note how this primitive is also ambiguous in the type used for the
@@
 This allows us to use a visible type application to obtain values from the same
-target type but generated using a different underlying representations.
+target type but generated using a different underlying representations, and
+avoiding the use of |Proxy| values everywhere.
 %
@@
 \end{code}
-
-Where |Html_simplify| and |ValidHtml| are the representations defined at the end
+%
+where |Html_simplify| and |ValidHtml| are the representations defined at the end
 of Section \ref{sec:representation}.
@@
 %
-So far we have seen how to represent and generate values for a target data type
-by combining a wide selection of structural constructions, as well as a series
-of type-level combinators to encode the generation behavior.
+So far we have seen how to represent and generate values for our target data
+type by combining a wide range of random constructions, as well as a series of
+type-level combinators to encode the desired generation behavior.
 %
-The next section refines our type-level machinery in order to export an elegant
-domain specific language for specifying composable random generators.
+The next section refines our type-level machinery in order to provide a simple
+idiom for defining composable random generators.

### 2019-05-06 papers/haskell19/5-typelevel.tex
Pass over Section 5

@@
 This section introduces refinements to our basic language for describing random
-generators, making them more flexible and robust in order to fit real-world
-usage scenarios.
+generators, making it more flexible and robust in order to fit real-world usage
+scenarios.
 
@@
 In practice, the actual names used when deriving construction representations
-need to be generated in a way that they are compliant with Haskell syntax and
-are \emph{unique within our namespace}.
+need to be generated in such a way that they are compliant with Haskell's
+syntax, and also that they are \emph{unique within our namespace}.
 %
-This means that, type names like |Fun_cat| or |Pat_simplify_1| are technically
+This means that, type names like |Fun_cat| or |Pat_simplify_1| are, technically,
 not valid Haskell data type names, thus they will have to be synthesized as
 something like |Fun_lt_plus_gt_543| and |Pat_simplify_1_325|, where the last
-sequence of numbers is automatically inserted by Template Haskell to ensure
-uniqueness.
+sequence of numbers is inserted by Template Haskell to ensure uniqueness.
 
@@
 %
-Instead of imposing a naming convention in our tool, we can define a set of open
-type families to hide each kind of construction using meaningful names:
+Instead of imposing a naming convention in our derivation tool, we define a set
+of open type families to hide each kind of construction behind meaningful names:
 
@@
 \end{code}
-
+%
 where |Symbol| is the kind of type-level strings in Haskell.
@@
 Then, our derivation process will synthesize each representation using unique
-names, along with a type instance of the corresponding type family for the end
-user to manipulate it.
+names, along with a type instance of the corresponding type family.
 %
@@
 \end{code}
-
+%
 As a result, the end user can simply refer to each particular construction by
@@
 Fun "<+>"|.
+%
+The additional |Nat| type parameter on |Pat| simply identifies each pattern
+number uniquely.
 
@@
 \end{code}
-
+%
 This way, any construction with the supra-index |T| will considered to be
 terminal.
-%
+
 With these type-level manipulations, we are now able to write generation
@@
 %
-Instead of carrying this type parameter in our generation specifications, we can
-avoid it by hiding them using an existential type:
+However, instead of carrying this type parameter in our generation
+specifications, we can avoid it by hiding it using an existential type:
 
@@
 \end{code}
-
+%
 The type constructor |Some| is a wrapper for a 2-parametric type that hides the
-first type variable using an explicit existential quantificator.
+first type variable using an explicit existential quantifier.
 %
-Note thus that the type parameter |a| does not appear at the left hand side of
-|Some| in its defition.
+Note thus that the type parameter |a| does not appears at the left hand side of
+|Some| on its defition.
 %
-Then, when deriving any |Con|, |Fun| or |Pat| type synonym, we will wrap each
-construction representations using a |Some|:
+Then, when deriving any |Con|, |Fun| or |Pat| type instance, we will wrap each
+construction representation using a |Some|:
 
@@
 
-We can therefore write generation specification for our |Tree| data type without
-having to refer to its type parameter anywhere.
+And we can therefore write generation specifications for our |Tree| data type
+without having to refer to its type parameter anywhere.
 %
@@
 \begin{code}
-type Tree_Spec
-  =      Con'  "Leaf"  otimes 2
+type Tree_Spec  =      Con'  "Leaf"  otimes 2
                 oplus  Con   "Node"  otimes 3
 \end{code}
-
+%
 Instead, we defer handling this type parameter until we actually use it to
@@
 \end{code}
-
-Where |apply| is a closed type family that simply traverses our generation
+%
+Where |apply| is a closed type family that traverses our generation
 specification, applying the |Int| type to every occurrence of |Some|:
@@
 
+It is worth mentioning that this approach for handling parametric types can be
+extended to multi-parametric data types with minor effort.
+
 Along with our automated constructions derivation mechanism, the machinery
 introduced in this section allows us to specify random generators using a
-concise and flexible type-level domain specific language.
+concise and flexible type-level combinators language.
 %

### 2019-05-06 papers/haskell19/5-typelevel.tex
Pass over Section 6

@@
 
-And simply obtain a concrete generator by refering to |Html_Spec| when using
+And simply obtain a concrete generator by referring to |Html_Spec| when using
 |genRep|:
@@
 In order to represent its data constructors, we can follow the same recipe
-presented in Section \ref{sec:representation}, but also parametrizing our
+presented in Section \ref{sec:representation}, but also parameterizing our
 representations over the type variable |a| as well:
@@
 Note thus that the type parameter |a| does not appears at the left hand side of
-|Some| on its defition.
+|Some| on its definition.
 %

### 2019-05-06 papers/haskell19/6-casestudies.tex
Pass over Section 6

@@
 
-The random generation framework presented throughout this work allows us to
+The random generation framework presented throughout this paper allows us to
 write extensible generators in a very concise way.
 %
-However, this expresiveness comes attached to a perceptible runtime overhead,
-primarly inherited from the using the Data Types \`a la Carte---a technique
-which is not often scrutinized for performance.
+However, this expressiveness comes attached to a perceptible runtime overhead,
+primarily inherited from the use of Data Types \`a la Carte---a technique which
+is not often scrutinized for performance.
 %
@@
 %
-First, generate random values from the representation types used to specify the
-shape of our data; and then we use their algebras to translate them to the
+First, we generate random values from the representation types used to specify
+the shape of our data; and then we use their algebras to translate them to the
 corresponding values of our target data types.
 %
-In particular, the tranformation step is expected to pattern match repeatedly
+In particular, the transformation step is expected to pattern match repeatedly
 against the |InL| and |InR| constructors of the |oplus| operators when
@@
 
-Then, we analyzed the performance of generating random values using three case
+Later, we analyzed the performance of generating random values using three case
 studies:
@@
 %
-The expanse of each case study can be summarized as follows:
+The magnitude of each case study can be outlined as follows:
 
@@
 
-These case studies provide a good combination of constructions, interface
+These case studies provide a good combination of data constructors, interface
 functions and patterns, and cover from smaller to larger numbers of
@@
 
-We benchmarked the execution time required to generate and fully evaluate 10000
-random values corresponding to each case study, comparing both manually-written
-concrete generators and those obtained using our modular approach.
+Then, we benchmarked the execution time required to generate and fully evaluate
+10000 random values corresponding to each case study, comparing both
+manually-written concrete generators, and those obtained using our modular
+approach.
 %
@@
 Figure \ref{fig:times} illustrates the relative execution time of each case
-study, normalized to their corresponding manually-written counterparts---we
+study, normalized to their corresponding manually-written counterpart---we
 encourage the reader to obtain a colored version of this work.
@@
 \end{figure}
-
+%
 As it can be observed, our approach suffers from a noticeable runtime overhead
 when using linearly defined representations, specially when considering the HTML
-case study, which spans a large amount of constructions in the generation
-process.
+case study, involving a large number of constructions in the generation process.
 %
-However, we found that, by balancing our representation types the generation
+However, we found that, by balancing our representation types, the generation
 performance improves dramatically.
@@
 the flexibility obtained from using our compositional approach does not produce
-severe slowdowns in when generating random values in practice.
+severe slowdowns when generating random values in practice.
 

### 2019-05-06 papers/haskell19/7-related.tex
Pass over Section 7

@@
 
-\citeauthor{SwierstraDTC} \cite{SwierstraDTC} proposes the Data Types \`a la
-Carte approach for building extensible data types and unified interpreters as a
+\citeauthor{SwierstraDTC} \cite{SwierstraDTC} proposed Data Types \`a la Carte,
+technique for building extensible data types and unified interpreters, as a
 solution for the expression problem coined by \citeauthor{wadler1998expression}
@@
 scope of random data generation, where other parameters come into play apart
-from just combining construction, e.g., generation frequency and terminal
+from just combining constructions, e.g., generation frequency and terminal
 constructions.
@@
 This solution avoids much of the runtime overhead introduced when internally
-pattern matching against sequencies of |InL| and |InR| data constructors.
+pattern matching against sequences of |InL| and |InR| data constructors.
 %
@@
 In our particular setting, we found that our solution has a fairly acceptable
-overhead by automatically balancing our representation types.
+overhead, achieved by automatically balancing our representation types.
 
@@
 %
-\emph{Feat} provides a mechanism to uniformly generate values from a given data
-type \cite{DuregardJW12}.
+\emph{Feat} provides a mechanism to uniformly generating values from a given
+data type \cite{DuregardJW12}.
 %
 It works by enumerating all the possible values of such type, so that sampling
-uniformly from it simply becomes sampling uniformly from the a finite preffix of
-natural numbers.
+uniformly from it simply becomes sampling uniformly from the a finite prefix of
+natural numbers---something easy to do.
 
 \emph{MegaDeTH} is a simple derivation tool that derives generators solely based
-on their types, and paying no attention whatsoever on the generation frequency
-of each data constructor.
+on their types, paying no attention whatsoever to the generation frequency of
+each data constructor.
 %
-As a result, its automatically derived generators has been shown to be skewed to
-generating small values\cite{DBLP:conf/haskell/MistaRH18}.
+As a result, its automatically derived generators has been shown to be skewed
+towards generating very small values\cite{DBLP:conf/haskell/MistaRH18}.
 
@@
 generators, tuning their generation frequencies using a simulation-based
-optimization process parametrized by the distribution of values desired by the
-user \cite{DBLP:conf/haskell/MistaRH18}.
+optimization process, which is parameterized by the distribution of values
+desired by the user \cite{DBLP:conf/haskell/MistaRH18}.
 %
 This simulation is based on the theory of \emph{branching processes}, which
-models the growth and extinction of populations across succesive generations.
+models the growth and extinction of populations across successive generations.
 %
@@
 %
-This tool has shown to improve the code coverage obtained over our codebase,
-with respect to other fully-automated generators derivation mechanism.
+This tool has shown to improve the code coverage over complex systems, when
+compared to other fully-automated generators derivation mechanisms.
 %
@@
 information covered here, namely abstract interfaces and function pattern
-machings.
+matchings.
 %
@@
 
-This paper tackles the representation problem, demonstrating how a compositional
-generation process can be effectively implemented and automated in Haskell using
-advanced type-level features.
+In turn, this paper tackles the representation problem, demonstrating how a
+compositional generation process can be effectively implemented and automated in
+Haskell using advanced type-level features.
 %

### 2019-05-06 papers/haskell19/8-conclusions.tex
Pass over Section 8

--- 2019-05-01 papers/haskell19/8-conclusions.tex
@@
 
-We presented a novel approach for automatically deriving flexibe random
+We presented a novel approach for automatically deriving flexible random
 generators inspired by the seminal work on Data Types \`a la Carte.
@@
 Furthermore, we incorporate valuable structural information into our generation
-process by considering not only data constructors but also the information
+process by considering not only data constructors, but also the information
 statically available in abstract interfaces and functions' pattern matchings.
 
-In the future, something something...
+In the future, we aim to minimize the runtime overhead of our technique, using
+perhaps prefix-free encodings based on generation frequencies, as a way to
+reduce the representation bureaucracy as much as possible.
+%
+Additionally, we are interested on including additional sources of structural
+information from the users' codebases, e.g., locally defined patterns or
+sequences of functions applications.
+%
+Interesting thoughts to motivate our future research.

### 2019-05-06 papers/haskell19/2-sources.tex
Nit

--- 2019-05-06 papers/haskell19/2-sources.tex
@@
 simplify :: Html -> Html
-simplify (Text t1 iJoin Text t2) = Text (t1 ++ t2)
-simplify ((Text t1 iJoin x) iJoin y)
+simplify (Join (Text t1) (Text t2)) = Text (t1 ++ t2)
+simplify (Join (Join (Text t) x) y)
   = simplify (Text t1 iJoin simplify (x iJoin y))
-simplify (x iJoin y) = simplify x iJoin simplify y
+simplify (Join x y) = simplify x iJoin simplify y
 simplify (Tag t x) = Tag t (simplify x)

### 2019-05-07 papers/haskell19/1-intro.tex
Some citations

@@
 Random property-based testing is a powerful technique for finding bugs
-\cite{ArtsHNS15, HughesPAN16, HughesNSA16}.
+\cite{ArtsHNS15, HughesPAN16, HughesNSA16, MidtgaardJKNN17}.
 %
-In the Haskell community, QuickCheck is the predominant tool of this such.
+In the Haskell community, QuickCheck is the predominant tool of this such
+\cite{ClaessenH00}.
 %
@@
 %
-Fortunately, there exists a variety of meta-programming tools for obtaining
-random generators automatically, based on the information encoded on the
-definition of the types we want to generate.
+Fortunately, there exists a variety of tools for obtaining random generators
+automatically, based on the information encoded on the definition of the types
+we want to generate.
 %
-Tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random generators
-using roughly the same procedure as the one shown above.
+Most of them implemented using Haskell's meta-programming framework named
+Template Haskell \cite{SheardJ02}.
+%
+For instance, tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random
+generators using roughly the same procedure as the one shown above.
 %
@@
   (Section \ref{sec:typelevel}).\looseness=-1
-\item We provide a Template Haskell tool for automatically deriving all the
-  required machinery presented throughout this paper, and evaluate its
-  generation performance using three real-world case studies (Section
-  \ref{sec:casestudies}).
+\item We provide a Template Haskell tool for automatically
+  deriving all the required machinery presented throughout this paper, and
+  evaluate its generation performance using three real-world case studies
+  (Section \ref{sec:casestudies}).
 \end{itemize}

### 2019-05-07 papers/haskell19/2-sources.tex
Some citations

@@
 %
-In this light, we can simply generate values satisfying pattern directly by
-returning its corresponding expression, where each variable or wildcard pattern
-is filled using a random sub-expression:
+In this light, we can simply generate values satisfying patterns directly by
+returning their corresponding expressions, where each variable or wildcard
+pattern is filled using a random sub-expression:
 

### 2019-05-07 papers/haskell19/1-intro.tex
Some compresUsing CompactItemize and arranged code to not break on different pages

@@
 %
-% \vspace{-10pt}
 %
-\begin{itemize}
+\begin{CompactItemize}
 \item We identify some important limitations of performing random testing with
@@
   (Section \ref{sec:typelevel}).\looseness=-1
-\item We provide a Template Haskell tool for automatically
-  deriving all the required machinery presented throughout this paper, and
-  evaluate its generation performance using three real-world case studies
-  (Section \ref{sec:casestudies}).
-\end{itemize}
-
+\item We provide a Template Haskell tool%
+  %
+  \footnote{Link will be made available for the final version of this paper.}
+  %
+  for automatically deriving all the required machinery presented throughout
+  this paper, and evaluate its generation performance using three real-world
+  case studies (Section \ref{sec:casestudies}).
+\end{CompactItemize}
+%
 Overall, we present a novel technique for reusing automatically derived

### 2019-05-07 papers/haskell19/2-sources.tex
Some compresUsing CompactItemize and arranged code to not break on different pages

@@
 %
-These functions let us create interesting structured values representing HTML
-constructions like line-breaks (|br|), bold blocks (|bold|), unordered lists
-(|list|) and concatenation of values one below another (|<+>|).
+These functions let us encode structured HTML constructions like line-breaks
+(|br|), bold blocks (|bold|), unordered lists (|list|) and concatenation of
+values one below another (|<+>|).
+%
+\looseness=-1
 

### 2019-05-07 papers/haskell19/5-typelevel.tex
Some compresUsing CompactItemize and arranged code to not break on different pages

@@
 \begin{code}
-type Html_Spec
-  =      Con'  "Text"
+type Html_Spec  =      Con'  "Text"
                 oplus  Con   "Join"         otimes 2
@@
                 oplus  Fun   "<+>"          otimes 9
-  oplus  Pat   "simplify"  1  otimes 3
+                oplus  Pat'  "simplify"  1  otimes 3
                 oplus  Pat   "simplify"  2  otimes 5

### 2019-05-07 papers/haskell19/1-intro.tex
Work on Ale's feedback

@@
 %
-In the Haskell community, QuickCheck is the predominant tool of this such
+In the Haskell community, QuickCheck is the predominant tool for that task
 \cite{ClaessenH00}.
@@
 %
-Then, QuickCheck generates random seeds using these generators, which are used
+Then, QuickCheck generates random values using these generators, which are used
 to evaluate the system's properties in search of possible counterexamples, which
@@
 
-Using this representation, the example page \linebreak
+Using this representation, the example page:
+
+\begin{code}
+ExampleHTML
+\end{code}
 %
-\texttt{<html><b>hello<br>world!</b></html>}
-%
-is encoded as:
+can be encoded as:
 %
@@
 %
-For the case of the recursive sub-terms, this generator simply calls |gen|
-recursively with a smaller generation size (|gen (d-1)|).
+For the case of the recursive sub-terms, this generator simply calls the
+function |gen| recursively with a smaller generation size (|gen (d-1)|).
 %
@@
 %
-For instance, tools like \emph{MegaDeTH} or \emph{DRAGEN} will synthesize random
+For instance, tools like \emph{MegaDeTH} \cite{GriecoCB16, grieco2017} or
+\emph{DRAGEN} \cite{DBLP:conf/haskell/MistaRH18} will synthesize random
 generators using roughly the same procedure as the one shown above.
@@
 derived generators.
+%
+\looseness=-1
 
@@
 For instance, because of the types we use for deriving our generators are too
-general, leaving a large room for ill-formed values, e.g., invalid HTML pages;
-or alternatively, because our properties need to be tested against very specific
-input patterns in order to trigger bugs.
+general, the generation process leaves a large room for ill-formed values, e.g.,
+invalid HTML pages; or alternatively, because it fails at generating very
+specific patterns of values which might be needed to trigger bugs.
+%
+% because our properties need to be tested against very specific input patterns in
+% order to trigger bugs.
 
@@
 %
-For instance, having a set of high-level combinators, defining an abstract
-interface of our data types, e.g., |hr :: Html| or |bold :: Html -> Html|; it is
-possible to include them into the generation process, thus obtaining more
-complex and structured values.
+% For instance, having a set of high-level combinators, defining an abstract
+% interface of our data types, e.g., |hr :: Html| or |bold :: Html -> Html|; it is
+% possible to include them into the generation process, thus obtaining more
+% complex and structured values.
 %
-In the same way, having functions defined using pattern matching over the data
-types of our interest, e.g. |f (Join (Text t) (Tag s x)) = dots|, it is also
-possible to include this rich structure which exposes a relationship among
-different data constructors into our random generators.
+For instance, it is possible to include into the generation process a set of
+high-level combinators of our data types, e.g., |hr :: Html| or |bold :: Html ->
+Html|---thus obtaining more complex and structured values.
 %
-This idea is addressed in depth in the next section, while the rest of this work
-focuses on implementing it in a compositional way.
+This methodology of generating random data using high-level combinators has
+shown to be particulary useful in the presence of monadic code
+\cite{Claessen:2002, grieco2017}.
+%
+% In the same way, having functions defined using pattern matching over the data
+% types of our interest, e.g. |f (Join (Text t) (Tag s x)) = dots|, it is also
+% possible to include this rich structure which exposes a relationship among
+% different data constructors into our random generators.
+%
+In the same way, it is possible to include into the generation process the
+pattern matchings that different functions perform over the data types of our
+interest, e.g. |f (Join (Text t) (Tag s x)) = dots|.
+%
+Such patterns often expose interesting relationships between multiple data
+constructors, a valuable asset for testing complex systems expecting highly
+structured inputs \cite{klein2009randomized}.
+%
+% This idea is addressed in depth in the next section, while the rest of this work
+% focuses on implementing it in a compositional way.
+%
+This work shows how to modularly include these extra sources of structural
+information into our automatically derived generators in a compositional manner.
 
@@
 %
-%
 \begin{CompactItemize}
@@
 \item We present an extensible mechanism for representing random values built
-  upon different source constructions using Data Types \`a la Carte (Section
-  \ref{sec:representation}).
+  upon different sources of structural information using Data Types \`a la Carte
+  (Section \ref{sec:representation}).
 \item We develop a modular generation schema, exploiting the composable nature

### 2019-05-07 papers/haskell19/2-sources.tex
Work on Ale's feedback

@@
-\section{Random Testing with Source Hints}
+\section{Additional Sources of Structural Information}
 \label{sec:sources}
@@
 %
-\subsection{Extracting Information From Interfaces}
+\subsection{Abstract Interfaces}
 
@@
 
-Note how difficult would it be to generate random values containing, for
+Note how difficult it would be to generate random values containing, for
 example, structurally valid HTML lists if we only consider the structural
@@
 %
-\subsection{Extracting Information From Patterns}
+\subsection{Pattern Matchings}
 

### 2019-05-07 papers/haskell19/3-representation.tex
Work on Ale's feedback

@@
 \begin{inparaenum}[(i)]
-\item combined in several ways depending on the desired shape of our test data,
-  applying the familiar \`a la Carte technique;
+\item combined in several ways depending on the desired shape of our test data
+  (applying the familiar \`a la Carte technique);
 \item randomly generated (see Section \ref{sec:generators}); and finally,
@@
 %
-This recusion schema using functors and F-Algebras to fold data generically has
-been well studied and is known categorically a \emph{catamorphism} \tocite.
+This recursion schema using functors and F-Algebras to fold data generically has
+been well studied and is known a \emph{catamorphism} \cite{Meijer91}.
 

### 2019-05-07 papers/haskell19/6-casestudies.tex
Work on Ale's feedback

@@
 %
-In particular, the transformation step is expected to pattern match repeatedly
-against the |InL| and |InR| constructors of the |oplus| operators when
-traversing each construction injection.
+In particular, this last step is expected to pattern match repeatedly against
+the |InL| and |InR| constructors of the |oplus| operators when traversing each
+construction injection.
 %
@@
 
+\vspace{-5pt}
 \begin{table}[H]
@@
 \end{table}
+\vspace{-5pt}
 
@@
     \begin{axis}
-      [ height=6cm
+      [ height=4.5cm
       , width=9cm
@@
       , ytick={0,1,...,4}
-      , ymax=4
-      , ymin=0.5
-      , enlargelimits=0.25
+      , ymax=3.7
+      , ymin=0.7
+      , enlargelimits=0.3
       , enlarge x limits=0.25
@@
       , legend style={
-        at={(0.5,-0.15)},
+        at={(0.5,-0.2)},
         anchor=north,legend columns=3}
@@
   \end{tikzpicture}
+  \vspace{-7pt}
   \caption{Generation time comparison between concrete and modular generators.}
   \label{fig:times}
+  \vspace{-18pt}
 \end{figure}
-%
+
 As it can be observed, our approach suffers from a noticeable runtime overhead

### 2019-05-07 papers/haskell19/7-related.tex
Work on Ale's feedback

@@
 
-\citeauthor{SwierstraDTC} \cite{SwierstraDTC} proposed Data Types \`a la Carte,
-technique for building extensible data types and unified interpreters, as a
-solution for the expression problem coined by \citeauthor{wadler1998expression}
+\citeauthor{SwierstraDTC} proposed Data Types \`a la Carte \cite{SwierstraDTC},
+a technique for building extensible data types, as a solution for the
+\emph{expression problem}, coined by \citeauthor{wadler1998expression}
 \cite{wadler1998expression}.
+%
+This technique has been successfully applied in a variety of scenarios, from
+extensible compilers, to composable machine-mechanized proofs
+\cite{Day:2013:CAL:2620678.2620680, persson2011generic, wu2014effect,
+  delaware2013meta}.
 %
@@
 As a result, its automatically derived generators has been shown to be skewed
-towards generating very small values\cite{DBLP:conf/haskell/MistaRH18}.
+towards generating very small values
+\cite{DBLP:conf/haskell/MistaRH18}.\looseness=-1
 

### 2019-05-07 papers/haskell19/8-conclusions.tex
Work on Ale's feedback

@@
 Furthermore, we incorporate valuable structural information into our generation
-process by considering not only data constructors, but also the information
-statically available in abstract interfaces and functions' pattern matchings.
+process by considering not only data constructors, but also the structural
+information statically available in abstract interfaces and functions' pattern
+matchings.
 
@@
 information from the users' codebases, e.g., locally defined patterns or
-sequences of functions applications.
-%
-Interesting thoughts to motivate our future research.
+sequences of functions applications---interesting thoughts to motivate our
+future research.

### 2019-05-07 papers/haskell19/abstract.tex
Work on Ale's feedback

@@
 Our framework relies on familiar ideas from the \`a la Carte technique for
-writing composable interpreters in Haskell, extended with the machinery required
-in the scope of random generation.
+writing composable interpreters in Haskell.
+%
+In this light, we adapt this technique with the machinery required in the scope
+of random generation.
 %
@@
 real-world examples.
+%

### 2019-05-08 papers/haskell19/1-intro.tex
Minor changes, fixed layout, added citation

@@
 %
-Then, QuickCheck generates random values using these generators, which are used
-to evaluate the system's properties in search of possible counterexamples, which
+Then, QuickCheck generates random values using these generators, and uses them
+to evaluate the testing properties in search of possible counterexamples, which
 often indicate the presence of bugs.
@@
 %
-In this paper, we will refer to any non-recursive data constructor as
-\emph{terminal}, and \emph{non-terminal} in any other case.
+In this paper, we will refer to any recursive data constructor as
+\emph{non-terminal}, and as \emph{terminal} in any other case.
 
@@
 \emph{DRAGEN} \cite{DBLP:conf/haskell/MistaRH18} will synthesize random
-generators using roughly the same procedure as the one shown above.
+generators using roughly the same generation procedure as the one shown above.
 %
@@
 %
-This allows the users to specify the desired distribution of values of the
-derived generators.
+This allows the users to specify the desired distribution of values produced by
+the derived generators.
 %
@@
 Overall, we present a novel technique for reusing automatically derived
-generators in a composable fashion, in contrast with the usual paradigm of
+generators in a composable fashion, in contrast to the usual paradigm of
 automatically deriving ``rigid'', monolithic generators.

### 2019-05-08 papers/haskell19/2-sources.tex
Minor changes, fixed layout, added citation

@@
 (<+>) :: Html -> Html -> Html
-(<+>) x y = Join x (Join br y)
+(<+>) x y = x iJoin br iJoin y
 \end{code} %$
@@
 6\% of the overall testing time (and considering recursive sub-calls), using
-random generators derived using either \emph{MegaDeTH} and \emph{DRAGEN}.
+random generators derived using both \emph{MegaDeTH} and \emph{DRAGEN}.
 %

### 2019-05-08 papers/haskell19/3-representation.tex
Minor changes, fixed layout, added citation

@@
 The key idea of this work is to lift each different source of structural
-information at the type level.
+information to the type level.
 %
 In this light, the ``shape'' of our random data is determined entirely by the
-type we use to represent it during the generation process.
+type we use to ``represent'' it during the generation process.
 
-For this, we will define a set of different simple ``open'' types, each one in
-charge of representing a single construction of our \emph{target} data type,
-i.e., the actual data type we want to randomly generate.
+For this purpose, we will define a set of different simple ``open'' types, each
+one in charge of representing a single random construction from our
+\emph{target} data type, i.e., the actual data type we want to randomly
+generate.
 %
@@
 %
-This let us automate the process quite a bit using Haskell's meta-programming
-capabilities.
+This let us automate the process quite a bit using Template Haskell.
 
@@
 For this purpose, we will define a generic functor type |oplus| to encode the
-choice between two given representation types as follows:
+choice between two given representation types:
 
@@
 %
-Such values can be constructed easily using the injections |InL| and |InR| on
-each each case, respectively.
+Such values can be constructed using the injections |InL| and |InR| on each
+case, respectively.
+%
+\looseness=-1
 
@@
 %
-This phenomena is addressed in detail in Section \ref{sec:casestudies}.
+This phenomenon is addressed in detail in Section \ref{sec:casestudies}.
 %
@@
 the type |Fix f| instantiates each recursive field of |f| with |Fix f|, thus
-closing the definition of |f| into itself---note thus that the kind of |Fix f|
-results |*|.
+closing the definition of |f| into itself---thus the kind of |Fix f| results
+|*|.
 %
@@
 %
-This is possible since we only use it as a wrapper to obtain a self-recursive
-type from a parametric one.
+% This is possible since we only use it as a wrapper to obtain a self-recursive
+% type from a parametric one.
 
@@
 %
-For this purpose we use the following type to represent it:
+For this purpose, we can use the following type:
 
@@
 %
-Instead, the actual functions are evaluated when transforming our
-representations back to their target type, by the means of an |Algebra|:
+Instead, these functions are evaluated when transforming our representations
+back to their target type, by the means of an |Algebra|:
 
@@
 %
-In particular, using non-terminating functions to represent our data might
-produce a non-terminating behavior when calling to the |eval| function.
+In particular, representing non-terminating functions might produce a
+non-terminating behavior when calling to the |eval| function.
 
@@
 %
-We start defining single data types for each one of them, this time encoding in
-the fields of each single data constructor the free pattern variables (or
-wildcards) appearing in its corresponding pattern, from left to right:
+We start by defining data types for each one of them, this time using the fields
+of each single data constructor to encode the free pattern variables (or
+wildcards) appearing on its corresponding pattern, from left to right:
 
@@
 %
-Instead of constructing values solely by combining different data constructors,
-this mechanism lets us also consider complex constructions like abstract
-interface functions and functions' patterns, treating them uniformingly and in a
-fine grained fashion.
+Instead of constructing values solely by combining data constructors, this
+mechanism lets us also consider complex constructions like abstract interface
+functions and functions' patterns, treating them uniformingly and in a fine
+grained fashion.
 
-As we have introduced in the Section \ref{sec:sources}, having an extensible,
-fine-grained framework for generating random data might be of great advantage,
-as we would like to generate random test cases following different strategies
-when testing different properties or sub-systems.
+As we have introduced in the Section \ref{sec:sources}, having an extensible
+framework for generating random data might be of great advantage, as we would
+like to generate random test cases following different strategies when testing
+different properties or sub-systems.
 %

### 2019-05-08 papers/haskell19/4-generators.tex
NIT
Some compresUsing CompactItemize and arranged code to not break on different pages
Minor changes, fixed layout, added citation

@@
 %
-So far we have seen how to encode different constructions representing
+So far we have seen how to encode different random constructions representing
 interesting values from our target data types.
 %
-Such representations follow a modular approach, where each different
-construction is considered independent from the rest.
+Such representations follow a modular approach, where each construction is
+independent from the rest.
 %
-This modularity allows us to derive each different construction independently,
-as well to specify the shape of our target data with a fine granularity.
+This modularity allows us to derive each different construction individually, as
+well to specify the shape of our target data with a fine granularity.
 
@@
 %
-This step also follows the modular, fine-grained fashion, resulting in a
-generation process entirely compositional.
+This step also follows the modular fashion, resulting in a generation process
+entirely compositional.
 %
@@
 %
-\begin{itemize}
+\vspace{2pt}
+\begin{CompactItemize}
 \item We use QuickCheck's generation size to limit the depth of the generated
@@
 \item We differentiate between \emph{terminal} and \emph{non-terminal
-    constructors}, limiting our generator to pick only among terminal ones when
-  we have reached the maximum depth (case |gen 0|).
+    constructors}, picking only among terminal ones when we have reached the
+  maximum depth (case |gen 0|).
 \item We generate different constructions in a (possibly) different frequency.
-\end{itemize}
+\end{CompactItemize}
 %
@@
 %
-Note the similarities between the |Arbitrary1| and |BArbitrary1| type classes.
+Note the similarities between |Arbitrary1| and |BArbitrary1|.
 %
@@
 
-As we have shown so far, generating each representation is rather
-straightforward: there is only one data constructor to select, and every field
-is generated following a mechanical recipe.
+As we have seen so far, generating each representation is rather
+straightforward: there is only one data constructor to pick, and every field is
+generated using a mechanical recipe.
 %
 In our approach, most of the generation complexity is encoded in the random
-generator used for composite representations, which are built upon the |oplus|
-operator.
+generator for composite representations, built upon the |oplus| operator.
 %
@@
 notions of terminal construction and generation frequency.
+%
+\looseness=-1
 
@@
 
-On the other hand, the last step of our generation process, which only generates
+Furthermore, the last step of our generation process, which only generates
 terminal constructions, could be seen as considering the non-terminal ones as
@@
 %
-This way, we will introduce another type-level computation to calculate the
+This way, we can introduce another type-level computation to calculate the
 \emph{terminal generation frequency} |FreqOf'| of a given representation as
@@
 Like the generator for |Html| introduced in Section \ref{sec:intro}, this
-generator branches over the current generation depth |d|.
+generator branches over the current depth |d|.
 %
@@
 %
-We use a \emph{visible type application} (employing the |at (...)| syntax) to
-disambiguate to which natural value we are actually referring to.
+We use a \emph{visible type application} \cite{eisenberg2016visible} (employing
+the |at (...)| syntax) to disambiguate to which natural value we are actually
+referring to.
 %
-Then, for instance, the value |freqVal at (FreqOf (f otimes'' 5))| will evaluate
-to |5 :: Int|.
+Then, for instance, the value:
 
-The rest of our random generator works analogously, except that, this time we
-only want to generate terminal constructions, hence we use the |FreqOf'| type
-family to compute the terminal generation frequency of each operand.
+\begin{code}
+  freqVal at (FreqOf (f otimes'' 5 oplus' g otimes' 4))
+\end{code}
 %
-If any of |FreqOf' f| or |FreqOf' g| evaluates to zero, then it means that that
+evaluates to the concrete value |9 :: Int|.
+
+The |else| clause of our random generator works analogously, except that, this
+time we only want to generate terminal constructions, hence we use the |FreqOf'|
+type family to compute the terminal generation frequency of each operand.
+%
+If any of |FreqOf' f| or |FreqOf' g| evaluates to zero, it means that such
 operand does not contain any terminal constructions, and |frequency| will not
@@
 
-If it happens that both |FreqOf' f| and |FreqOf' g| compute to zero when
-generating values from composite representation, then it will produce a runtime
-error triggered by the function |frequency|, as it does not have anything with a
-positive frequency to generate.
+Moreover, if it happens that both |FreqOf' f| and |FreqOf' g| compute to zero
+simultaneously, then this will produce a runtime error triggered by the function
+|frequency|, as it does not have anything with a positive frequency to generate.
 %
@@
                            , Algebra f a) => BGen a
-genRep d = eval <$> bounded at (Fix f) d
+genRep d = eval <$> barbitrary at (Fix f) d
 \end{code} %$
@@
 This allows us to use a visible type application to obtain values from the same
-target type but generated using a different underlying representations, and
+target type but generated using different underlying representations, and
 avoiding the use of |Proxy| values everywhere.

### 2019-05-08 papers/haskell19/5-typelevel.tex
Minor changes, fixed layout, added citation

@@
 %
-In practice, the actual names used when deriving construction representations
-need to be generated in such a way that they are compliant with Haskell's
-syntax, and also that they are \emph{unique within our namespace}.
+In practice, the actual name used when deriving the representation for each
+random construction needs to be generated in such a way that is complies with
+Haskell's syntax, and also that it is \emph{unique within our namespace}.
 %
@@
 %
-For instance, along with the representations of |Con_Text|, |Fun_cat| and
-|Pat_simplify_1|, we will automatically derive the following type instances:
+For instance, along with the constructions representations |Con_Text|, |Fun_cat|
+and |Pat_simplify_1|, we will automatically derive the following type instances:
 
@@
 However, instead of carrying this type parameter in our generation
-specifications, we can avoid it by hiding it using an existential type:
+specifications, we can avoid it by hiding it behind an existential type:
 
@@
 %
-Where |apply| is a closed type family that traverses our generation
-specification, applying the |Int| type to every occurrence of |Some|:
+Where |apply| is a type family that traverses our generation specification,
+applying the |Int| type to each occurrence of |Some|:
+%
+\looseness=-1
 
@@
 Along with our automated constructions derivation mechanism, the machinery
-introduced in this section allows us to specify random generators using a
-concise and flexible type-level combinators language.
+introduced in this section allows us to specify random generators using a simple
+and flexible type-level specification language.
 %

### 2019-05-08 papers/haskell19/6-casestudies.tex
Minor changes, fixed layout, added citation

@@
 
-\begin{figure}[t]
+\begin{figure}[b]
+  \vspace{-10pt}
   \centering
@@
     \begin{axis}
-      [ height=4.5cm
+      [ height=5cm
       , width=9cm
@@
       , legend style={
-        at={(0.5,-0.2)},
+        at={(0.5,-0.25)},
         anchor=north,legend columns=3}
@@
   \label{fig:times}
-  \vspace{-18pt}
+  \vspace{-5pt}
 \end{figure}

### 2019-05-08 papers/haskell19/7-related.tex
Minor changes, fixed layout, added citation

@@
 It works by enumerating all the possible values of such type, so that sampling
-uniformly from it simply becomes sampling uniformly from the a finite prefix of
+uniformly from it simply becomes sampling uniformly from a finite prefix of
 natural numbers---something easy to do.
 
-\emph{MegaDeTH} is a simple derivation tool that derives generators solely based
-on their types, paying no attention whatsoever to the generation frequency of
-each data constructor.
+\emph{MegaDeTH} is a simple derivation tool that synthesizes generators solely
+based on their types, paying no attention whatsoever to the generation frequency
+of each data constructor.
 %

### 2019-05-08 papers/haskell19/abstract.tex
Minor changes, fixed layout, added citation

@@
 generators, which can be easily combined in different ways in order to fit the
-developers' demands, using a simple type-level description language.
+developers' demands using a simple type-level description language.
 %
@@
 %
-In this light, we adapt this technique with the machinery required in the scope
+In particular, we adapt this technique with the machinery required in the scope
 of random generation.

### 2019-05-08 papers/haskell19/7-related.tex
3rd person citation to dragen2

@@
 %
-In a recent work, \citeauthor{Mista2019GeneratingRS} extend this approach to
-generate random values considering also the other sources of structural
-information covered here, namely abstract interfaces and function pattern
-matchings.
+In a recent work, the authors extended this approach to generate random values
+considering also the other sources of structural information covered here,
+namely abstract interfaces and function pattern matchings.
 %
-The authors of this work focus on the generation model problem, extending the
-theory of branching processes in order to obtain sound predictions about
-distributions of random values considering these new kinds of constructions.
+There, they focus on the generation model problem, extending the theory of
+branching processes in order to obtain sound predictions about distributions of
+random values considering these new kinds of constructions.
 %

### 2019-05-08 bench/.#Html.hs

@@ -1,1 +1,0 @@
-agustin@palpatine.7239:1555097683

### 2019-05-08 bench/Main.hs
Improved compilation performance, removed optimization pass for benchmarks

@@ -7,11 +7,11 @@ import Criterion.Types
 import Test.Dragen2
 import Test.QuickCheck
 
-import qualified RBT.RBT        as RBT
 import qualified RE.RE          as RE
--- import qualified Html.Html     as Html
-import qualified Html.HtmlUnbal as Html
+import qualified RBT.RBT        as RBT
 import qualified Lisp.Lisp      as Lisp
+import qualified Html.Html      as Html
+--import qualified Html.HtmlUnbal as Html
 
 samples :: Int
 samples = 10000
@@ -33,16 +33,12 @@ genSamples gen
 main :: IO ()
 main = defaultMainWith (defaultConfig { verbosity = Verbose })
   [
-  --   bench "RBT/Manual"          $ nfIO (genSamples (RBT.genTree   @Int))
-  -- , bench "RBT/Dragen/Linear"   $ nfIO (genSamples (RBT.genTree'  @Int))
-  -- , bench "RBT/Dragen/Autobal"  $ nfIO (genSamples (RBT.genTree'' @Int))
-  -- , bench "Lisp/Manual"         $ nfIO (genSamples (Lisp.genSExpr))
-  -- , bench "Lisp/Dragen/Linear"  $ nfIO (genSamples (Lisp.genSExpr'))
-  -- , bench "Lisp/Dragen/Autobal" $ nfIO (genSamples (Lisp.genSExpr''))
-  -- , bench "Html/Manual"         $ nfIO (genSamples (Html.genHtml ))
-  -- , bench "Html/Dragen/Linear"  $ nfIO (genSamples (HtmlU.genHtml'))
-    bench "Html/Dragen/Autobal" $ nfIO (genSamples (Html.genHtml''))
-  -- , bench "RE/Manual"       $ nfIO (genSamples (RE.genR       @Char))
-  -- , bench "RE/Dragen"       $ nfIO (genSamples (RE.genR'      @Char))
-  -- , bench "RE/Dragen/Bal"   $ nfIO (genSamples (RE.genR''     @Char))
+    bench "RBT/Manual"          $ nfIO (genSamples (RBT.genTree   @Int))
+  , bench "RBT/Dragen/Autobal"  $ nfIO (genSamples (RBT.genTree'' @Int))
+  , bench "Lisp/Manual"         $ nfIO (genSamples (Lisp.genSExpr))
[…]

### 2019-05-08 src/Test/Dragen2/TH/FunPats.hs

@@ -84,7 +84,7 @@ derivePatRep funName typeFam funArgType funArgDefTypeVars branching
 
   -- Create the representation data declaration
   let repDataDec = DDataD Data [] repTypeName extTypeVars
-                     Nothing [singleCon] derivingClause
+                     Nothing [singleCon] (derivingClauses branching)
       singleCon = DCon extTypeVars [] repConName conRepFields repType
       conRepFields = DNormalC False (mkPatRepField <$> funArgPatVarsTypes)
 

### 2019-05-08 src/Test/Dragen2/TH/TypeInt.hs

@@ -78,7 +78,7 @@ deriveFunRep typeFam branching funName = do
 
   -- Create the representation data declaration
   let repDataDec = DDataD Data [] repTypeName extTypeVars
-                     Nothing [singleCon] derivingClause
+                     Nothing [singleCon] (derivingClauses branching)
       singleCon = DCon extTypeVars [] repConName conRepFields repType
       conRepFields = DNormalC False (mkPatRepField <$> funArgTypes)
 

### 2019-05-09 papers/haskell19/1-intro.tex
3rd person citation to dragen2
Minor changes based on Nachi's feedback

@@
 Random property-based testing is a powerful technique for finding bugs
-\cite{ArtsHNS15, HughesPAN16, HughesNSA16, MidtgaardJKNN17}.
+~\cite{ArtsHNS15, HughesPAN16, HughesNSA16, MidtgaardJKNN17}.
 %
-In the Haskell community, QuickCheck is the predominant tool for that task
-\cite{ClaessenH00}.
+In Haskell, QuickCheck is the predominant tool for this task \cite{ClaessenH00}.
 %

### 2019-05-09 papers/haskell19/2-sources.tex
Minor changes based on Nachi's feedback

@@
 %
-Ideally, we would like to put approximately the same testing effort info each
+Ideally, we would like to put approximately the same testing effort into each
 clause of |simplify|, or perhaps even more to the first two ones, since those
@@
 decreases multiplicatively with the number of constructors we simultaneously
-pattern match against to.
+pattern match against.
 
@@
             t2  <- arbitrary
-              return (Text t1 iJoin Text t2))
+            return (Join (Text t1) (Text t2)))
   , (4, do  t2  <- arbitrary
@@
             y   <- go (d-1)
-              return ((Text t1 iJoin x) iJoin y))]
-\end{code}%$
+            return (Join (Join (Text t1) x) y))]
+\end{code}
 

### 2019-05-09 papers/haskell19/3-representation.tex
Minor changes based on Nachi's feedback

@@
 Note the functional dependency |f -> a|, helping the type checker to solve
-uniquely the free type variable |a| at the right hand side of the |=>|, based on
+uniquely the free type variable |a| on the right hand side of the |=>|, based on
 the type |f| used to represent it.
@@
 This recursion schema using functors and F-Algebras to fold data generically has
-been well studied and is known a \emph{catamorphism} \cite{Meijer91}.
+been well studied and is known as a \emph{catamorphism} \cite{Meijer91}.
 

### 2019-05-09 papers/haskell19/7-related.tex
Minor changes based on Nachi's feedback

@@
 From the practical point of view, \citeauthor{KiriyamaOptimizingDTC} propose an
-optimization mechanism for Data Types \`a Carte, where a concrete data type has
-to be derived for each different composition of constructions defined by the
+optimization mechanism for Data Types \`a la Carte, where a concrete data type
+has to be derived for each different composition of constructions defined by the
 user \cite{KiriyamaOptimizingDTC}.

### 2019-05-09 papers/haskell19/abstract.tex
Minor changes based on Nachi's feedback

@@
 %
-However, none of them are built neither for composability nor extendability, a
-useful feature when the shape of our random data makes a difference while
-testing different properties or sub-systems.
+However, none of them are built for composability nor extensibility, a useful
+feature when the shape of our random data makes a difference while testing
+different properties or sub-systems.
 %

### 2019-05-11 papers/haskell19/3-representation.tex
Last minute changes

@@
 This lifted representation of constructions can be automatically derived from
-our source code in compile-time, relieving programmers of the burden of manually
+our source code at compile time, relieving programmers of the burden of manually
 implementing the required machinery.
@@
 %
-\subsubsection{Ensambling Altogether}
+\subsubsection{Assembling It All Together}
 
@@
 mechanism lets us also consider complex constructions like abstract interface
-functions and functions' patterns, treating them uniformingly and in a fine
-grained fashion.
+functions and functions' patterns, treating them uniformly and in a fine grained
+fashion.
 

### 2019-05-11 papers/haskell19/4-generators.tex
Minor changes based on Nachi's feedback
NIT
Last minute changes

@@
 %
-If we take a closer look at its definition, we can observe three factors
+If we take a closer look at its definition, there we can observe three factors
 happening simultaneously:
@@
   function |gen|.
-\item We differentiate between \emph{terminal} and \emph{non-terminal
+\item We differentiated between \emph{terminal} and \emph{non-terminal
     constructors}, picking only among terminal ones when we have reached the
@@
 \end{code} %$
-
+%
 Note how each instance is defined to be parametric of the maximum depth (using
@@
 %
-In order to model this behavior, we will first defined a data type |Term| to
+In order to model this behavior, we will first define a data type |Term| to
 ``tag'' every terminal construction explicitly:
@@
 %
-Fortunately, such runtime exceptions can be catched at compile-time.
+Fortunately, such runtime exceptions can be caught at compile time.
 %

### 2019-05-11 papers/haskell19/5-typelevel.tex
Minor changes based on Nachi's feedback
NIT
Last minute changes

@@
 This naming convention results hard to use, specially if we consider that we do
-not know the actual type names until they are synthesized in compile-time, due
-to their unique suffixes.
+not know the actual type names until they are synthesized during compilation,
+due to their unique suffixes.
 %
@@
 \end{code}
-
+%
 And simply obtain a concrete generator by referring to |Html_Spec| when using
@@
 
-% The biggest different appears when deriving the required type instances for the
-% |Con|, |Fun| and |Pat| synonyms.
+% The biggest different appears when deriving the required type instances for
+% the |Con|, |Fun| and |Pat| synonyms.
 %
@@
 \end{code}
-
-And we can therefore write generation specifications for our |Tree| data type
-without having to refer to its type parameter anywhere.
+%
+And as a consequence, we can write generation specifications for our |Tree| data
+type without having to refer to its type parameter anywhere.
 %

### 2019-05-11 papers/haskell19/8-conclusions.tex
Last minute changes

@@
 matchings.
+%
 
@@
 %
-Additionally, we are interested on including additional sources of structural
-information from the users' codebases, e.g., locally defined patterns or
-sequences of functions applications---interesting thoughts to motivate our
-future research.
+In addition, using type-level representations for each random construction
+leaves the door open for extending our approach with a composable mechanism for
+predicting the distributions of random values.
+%
+This would be possible by overloading each representation type with stochastic
+information describing its branching behavior---interesting thoughts to motivate
+our future research.
+
+% Additionally, we are interested on including additional sources of structural
+% information from the users' codebases, e.g., locally defined patterns or
+% sequences of functions applications

### 2019-05-11 papers/haskell19/abstract.tex
Last minute changes

@@
 State-of-the-art tools for synthesizing random generators serve the valuable
-purpose of helping with this task, while providing different level of invariants
-imposed over the generated values.
+purpose of helping with this task, while providing different levels of
+invariants imposed over the generated values.
 %
@@
 In this work, we develop an extensible framework for deriving compositional
-generators, which can be easily combined in different ways in order to fit the
+generators, which can be easily combined in different ways in order to fit
 developers' demands using a simple type-level description language.

### 2019-05-11 papers/haskell19/8-conclusions.tex
NIT

@@
 %
-In addition, using type-level representations for each random construction
-leaves the door open for extending our approach with a composable mechanism for
-predicting the distributions of random values.
+In addition, by using type-level representations for each random construction,
+we leave the door open to extend our approach with a composable mechanism to
+predict and optimize the distributions of random values.
 %

### 2019-05-11 papers/haskell19/4-generators.tex
NIT

@@
   function |gen|.
-\item We differentiated between \emph{terminal} and \emph{non-terminal
+\item We differentiate between \emph{terminal} and \emph{non-terminal
     constructors}, picking only among terminal ones when we have reached the

### 2019-06-23 papers/haskell19/1-intro.tex
NIT

@@
 to evaluate the testing properties in search of possible counterexamples, which
-often indicate the presence of bugs.
+always indicate the presence of bugs.
 
@@
 In the same way, it is possible to include into the generation process the
-pattern matchings that different functions perform over the data types of our
+pattern matchings that different functions perform over the data types of
 interest, e.g. |f (Join (Text t) (Tag s x)) = dots|.
@@
 %
-% This idea is addressed in depth in the next section, while the rest of this work
-% focuses on implementing it in a compositional way.
+% This idea is addressed in depth in the next section, while the rest of this
+% work focuses on implementing it in a compositional way.
 %
@@
   %
-  \footnote{Link will be made available for the final version of this paper.}
+  \footnote{Link to be made available in the final version of this paper.}
   %

### 2019-06-23 papers/haskell19/2-sources.tex

@@
 In this section, we briefly introduce the motivation for including extra
-information into our random generators, and discuss about how this can be
-implemented in practice.
+information into our random generators, and discuss how this can be implemented
+in practice.
 
@@
 
-A different challenge appears when we try to test functions comprising complex
+A different challenge appears when we try to test functions involving complex
 pattern matchings. Consider for instance the following function for simplifying
@@
 %
-As expected, most of the random test cases were exercising the simplest (an
+As expected, most of the random test cases were exercising the simplest (and
 rather uninteresting) patterns of this function, corresponding to its last three
@@
 directly is far from ideal, as in many cases we would like to select which
-random constructions are of our interest when testing different properties or
+random constructions are of interest when testing different properties or
 sub-systems.

### 2019-06-24 papers/ast19/code/Exp.hs
Removed papers from this repo

@@ -1,46 +1,0 @@
-module Exp where
-
-data Exp  =  Val  Int
-          |  Add  Exp  Exp
-          |  Mul  Exp  Exp
-          |  If   Exp  Exp  Exp
-          deriving Show
-
-eval :: Exp -> Int
-eval (Val n) = n
-eval (Add x y) = eval x + eval y
-eval (Mul x y) = eval x * eval y
-eval (If b x y) = if (eval b /= 0) then eval x else eval y
-
--- genExp :: Gen Exp
--- genExp = sized gen
---   where
---     gen 0 = Val <$> genInt
---     gen n = oneof
---       [  Val  <$> arbitrary
---       ,  Add  <$> gen pred   <*> gen pred
---       ,  Mul  <$> gen pred   <*> gen pred
---       ,  If   <$> arbitrary  <*> gen pred <*> gen pred ]
-
-
--- foo :: Exp -> Exp
--- foo (Mul (Add x (Val 42)) (Mul (Val 0) y))  = ...
--- foo (Add (Add x y) (Val z))                 = ...
--- foo (Add (Val 42)  (Val 24))                = ...
--- foo x                                       = ...
-
-(.+) = Add
-(.*) = Mul
-val = Val
-
-square :: Exp -> Exp
-square x = Mul x x
-
-neg :: Exp -> Exp
-neg x = Mul x (Val (-1))
-
-minus :: Exp -> Exp -> Exp
-minus x y = Add x (neg y)
-
-pow :: Exp -> Exp -> Exp
-pow b e = If e (Mul b (pow b (minus e (Val 1)))) (Val 1)

### 2019-06-24 papers/haskell19/watch.sh

@@ -1,4 +1,0 @@
-#!/bin/sh
-
-#https://facebook.github.io/watchman/docs/watchman-make.html
-watchman-make --make 'make' -p 'main.lhs.tex' 'notation.fmt' '[^.]*.tex' 'tikz/*.tex' '**/[^.]*.bib' '**/GNUmakefile' -t all
